### 2017-01-10 draft.tex
LaTeX files are added

% Suha's Proposal
\documentclass{article}

\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{arydshln}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{color}

\newtheorem{dfn}{Definition}
\newtheorem{lem}{Lemma}

\title{Existence of Forward Simulations for Particular Data Structures}
\begin{document}
\maketitle
\input{prelim}
\input{queue}
\section{Existence of Forward Simulations for Stack Implementations that have Fixed Pop Linearization Points}
\input{commit}
\input{set}
\end{document}

### 2017-01-10 prelim.tex
LaTeX files are added

\section{Preliminaries}
Systems we consider are labeled transition systems (LTS):
\begin{dfn}
An LTS is defined over four-tuples $A=(Q,\Sigma, q_0, \delta)$ where $Q$ is the set of states, $\Sigma$ is the set of transition labels, $q_0 \in Q$ is the initial state and $\delta \subseteq Q \times \Sigma \times Q$ is the transition relation.
\end{dfn}
Executions generated by this system are alternating sequence of states and transition labels $\rho = s_0, e_0, s_1,... s_k, e_k,...$ where each $s_i \in Q$, each $e_i \in \Sigma$, $s_0 = q_0$ and each $(s_i, e_i s_{i+1}) \in \delta$. The projection of the sequence $\rho$ over the set $\Pi$ is denoted by $\rho | \Pi$, and it is the maximum subsequence of $\rho$ consisting of elements of $\Pi$. Traces of the LTS are obtained from executions by projecting them over $\Sigma$. For the rest of the paper and in all of the proofs, we consider only finite executions (denoted as $E(A)$) and/or traces (denoted as $Tr(A)$ of the LTSs in focus.

Libraries are LTSs that provide methods. Let $\mathcal{M}$ be the set of method names and $\mathcal{D}$ be the domain of values as input/output parameters for the methods. Then, this library contains transition labels of the form $inv(m,d,i)$ representing the invocation of method $m \in \mathcal{M}$ with input value $d \in \mathcal{D}$. The third field is the operation identifier for differentiating the different calls of the same method from the set $\mathcal{O}$. For simplicity, we take $\mathcal{O} = \mathbb{N}$ for the rest of the paper. We also assume that methods could have at most one input parameter. If they do not have any input arguments (like pop method of a stack), we can omit the second field from the action. They also provide actions of the form $ret(m,d,i)$ representing the return of method $m \in M$ with value $d \in D$ which has been invoked previously with action $inv(m,d',i)$. Again, we assume that the methods can return at most one parameter and we may omit the second field from the action if they have none (like enqueue method of a queue). Before starting to reason about any set of libraries, we first fix the sets $\mathcal{M}$ and $\mathcal{D}$ and libraries in our focus agree on this sets. For any transition label $e = inv(m,d,i)$ or $e=ret(m,d,i)$, we have the function $oid(e) = i$.

Since libraries are LTSs, they produce traces. A trace $e = e_1, e_2, ..., e_n$ of library $L$ is \emph{well-formed} iff (i) every return matches an earlier invocation: $e_j = ret(m,d,k)$ implies that there exists $i<j$ such that $e_i = inv(m,d',k)$ and (ii) every operation identifier is used at most one invocation/return pair: $oid(e_i) = oid(e_j) = k$ and $i<j$ implies $e_i = inv(m,d,k)$ and $e_j = ret(m,d',k)$. From now on, we assume that libraries produce well-formed traces. Let $f: \mathbb{N} \rightarrow \mathbb(N)$ be a bijection. Then, traces $e$ and $e'$ are equivalent if $e'$ is obtained from $e$ by replacing every action $inv(m,d,k)$ with $inv(m,d,f(k))$ and every action $ret(m,d,k)$ with $ret(m,d,f(k))$.

Based on these definitions on the traces of the libraries, we can define refinement between libraries:
\begin{dfn}
Let $L_1$ and $L_2$ be two libraries agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets. We define the set $A\Sigma = \{inv(m,d,i)| m \in \mathcal{M} \wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\} \cup \{ret(m,d,i)| m \in \mathcal{M}\wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\}$ as abstract transition labels. Note that $A\Sigma \subseteq \Sigma_{L_1}$ and $A\Sigma \subseteq \Sigma_{L_2}$. Then, we say $L_1$ refines $L_2$ iff for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
\end{dfn}
Linearizability is also a relation between two libraries and it is stricter than refinement. It requires $e'$ in Definition 2 to be a sequential one. A trace $e$ is sequential iff following two conditions hold for its projection to abstract transition labels $e|A\Sigma = e_1, ...e_n$: $(i)$ $e_1 = inv(m,d,k)$ for some $m \in \mathcal{M}, d \in \mathcal{D} and k \in \mathbb{N}$ and $(ii)$ for all $i \in [1,n)$, either $e_i = inv(m,d,k)$ and $e_{i+1} = ret(m,d',k)$ or $e_i = ret(m,d,k)$ and $e_{i+1} = inv(m',d',k')$ for some $m,m' \in \mathcal{M}$, $d,d' \in \mathcal{D}$ and $k,k' \in \mathbb{N}$.

We can extend the relations between libraries by introducing simulation relations. We will later show that simulation relations imply refinement.

\begin{dfn}
Let $L_1$ and $L_2$ be two libraries agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets. Then, the relation $fs \subseteq Q_{L_1} \times Q_{L_2}$ is called a forward simulation iff the following holds:
\begin{itemize}
\item[(i)] $fs[q_{0_{L_1}}] = \{q_{0_{L_2}} \}$
\item[(ii-a)] If $(s,inv(m,d,k),s') \in \delta_{L_1}$ and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u \xrightarrow{a} u'$ where $a = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
\item[(ii-b)] If $(s,ret(m,d,k),s') \in \delta_{L_1}$ and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u \xrightarrow{a} u'$ where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-c)] If $(s,t,s') \in \delta_{L_1}$ for some $t \in \Sigma_{L_1}\backslash A\Sigma$ and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u \xrightarrow{a} u'$ where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}
If $fs[s]$ is a unique state for all $s \in Q_{L_1}$ then it is called a refinement mapping/function. A dual notion of forward simulation is the backward simulation:
\begin{dfn}
Let $L_1$ and $L_2$ be two libraries agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets. Then, the relation $bs \subseteq Q_{L_1} \times Q_{L_2}$ is called a backward simulation iff the following holds:
\begin{itemize}
\item[(i)] $bs[q_{0_{L_1}}] = \{q_{0_{L_2}} \}$
\item[(ii-a)] If $(s,inv(m,d,k),s') \in \delta_{L_1}$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{a} u'$ where $a = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-b)] If $(s,ret(m,d,k),s') \in \delta_{L_1}$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{a} u'$ where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-c)] If $(s,t,s') \in \delta_{L_1}$ for some $t \in \Sigma_{L_1}\backslash A\Sigma$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{a} u'$ where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}
Simulation relations are used to prove refinement relations among libraries. Following lemmas show their soundness:
\begin{lem}
Let $L_1$ and $L_2$ be two libraries agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets. If $fs$ ($bs$) is a forward (backward) simulation relating $L_1$ to $L_2$, then $L_1$ refines $L_2$.
\end{lem}
\begin{proof}
Looks trivial and follows Lynch paper. Can be completed later.
\end{proof}

### 2017-01-11 draftCAV/app-backSim.tex
started draft for CAV

%!TEX root = draft.tex
\section{Backward simulations}\label{app:backSim}

A dual notion of forward simulation is the backward simulation:
\begin{dfn}
Let $L_1=(Q_1,\Sigma, q_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, q_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$. A relation $bs \subseteq Q_1 \times Q_2$ is called a \emph{backward simulation} from $L_1$ to $L_2$ iff the following holds:
\begin{itemize}
\item[(i)] $bs[q_0^1] = \{q_0^2 \}$
\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0<i<|@s|$.
% where $a = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-b)] If $(s,r,s') \in \delta_1$, for some $r\in R$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s| -1$.
%where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-c)] If $(s,e,s') \in \delta_1$ for some $e \in \Sigma\setminus (C\cup R)$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}

\begin{lem}
Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$. $L_1$ refines $L_2$ if{f} there exists a backward simulation from $L_1$ to $L_2$.
\end{lem}
\begin{proof}
Looks trivial and follows Lynch paper. Can be completed later.
\end{proof}

### 2017-01-11 draftCAV/commit.tex
started draft for CAV

\section{Relaxation for the Data Structures Without Fixed Remove Linearization Points}
We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$. However, we observe that, these implementations contain some points in their pop methods that logically removes the element from the pool. We call them commit points. If a method of a library has a fixed linearization  point, it is also a commit point. In this sense, commit points are weaker versions of fixed linearization points. Fixed linearization point of a pop preserves the following properties:
\begin{itemize}
\item If a $ret$ comes before a linearization point in the concrete execution, this order is preserved in the linearization of this execution.
\item If a $lin$ pop comes before another $lin$ pop in the concrete execution, this order is preserved in the linearization of this execution.
\item If a $lin$ pop comes before an $inv$ in the concrete execution, this order is preserved in the linearization of this execution. 
\end{itemize} 
A commit point is weaker than a fixed linearization point in the sense that it does not need to satisfy the first and the second conditions.

To our intuition, if a pop (remove) method has multiple finite linearization points, commit point is the latest element in this set. We have never observed an implementation in which a pop is linearized after it logically removes the element from the data structure.

From now on, we will restrict ourselves to stacks, since our example implementation that we will show linearizability of is the time-stamped (TS) stack. However, the notions and the machines we will introduce can be extended to queues easily.

We fix $\mathcal{M} = \{push, pop\}$ and $\mathcal{D} = \mathbb{N} \cup \{\texttt{EMPTY}\}$. We extend the alphabet $A\Sigma$ for stacks with commit points as $ACS\Sigma = A\Sigma \cup \{com(pop,d,k)|d \in \mathcal{D}, k \in \mathcal{M}\}$. We define cs-refinement and cs-linearizability as we defined q-refinement, q-linearizability in the previous sections. We also change definitions of backward and forward simulation relations for stacks with commit points as we do in the previous sections by replacing linearization points with commit points in this extensions. Lemma 1 of the first section still holds with new simulation relation definitions and cs-linearizability and cs-linearizability implies the original linearizability definition.

Our road map for this section is as follows: We will first introduce an intermediate stack machine $L_I$ that will be deterministic with respect to the alphabet $ACS\Sigma$. We will show that $L_I$ is equivalent to the standard abstract stack $L_A$ defined in the previous section with respect to the language $A\Sigma$. We show this by first showing that $L_I$ is a refinement of $L_A$ wrt alphabet $A\Sigma$ by finding a backward simulation relation between them and then $L_A$ is a refinement of $L_I$ with respect to alphabet $A\Sigma$ by finding a forward simulation relation between them. Since $L_I$ is deterministic wrt $ACS\Sigma$, if we have an implementation $L_C$ that is a cs-refinement of $L_I$, we can find a forward simulation between them. As an example, we will pick $L_C$ as time-stamped stack and establish a forward simulation relation between it and our $L_I$ machine.

Let us continue with defining $L_I$ first:
\begin{itemize}
\item A state of $L_I$ again consists of a partial strict order and a program counter: $Q_I \subseteq ND \times ED \times (\mathbb{N} \rightarrow Lbl_I)$ where $Lbl_I = \{N, A_0, A_1, R_0, R_1, R_2\}$ is the set of transition labels for the operations. Different from the fixed linearization point case, this time nodes are not triples but 5-tuples of the form $(k,d,st,mc,con)$. First three fields are the same as previous intermediate machines: $k \in \mathbb{N}$ is operation identifier of a push operation, $d \in \mathcal{D}$ is the data value of that push and $st \in \{\texttt{PENDING}, \texttt{CLOSED}\}$ is the current status of the push operation. The fourth field $mc \subset \mathbb{N}$ keeps the operation identifiers of pop operations such that this push was maximally closed when the pop began. A node $n$ is maximally closed in a state $s$ iff $n.st  = \texttt{CLOSED}$ and if there is an edge $n \rightarrow n' \in ed_s$, then $n'.st = \texttt{OPEN}$. The fifth field $con \subset \mathbb{N}$ is the set of operation identifiers of the pop methods that are concurrent with this push i.e. the either this node was open when the pop started or this node is created when the pop was pending.
\item The transition labels consist of invocation and return actions of both methods and commit action for only pop method. Hence $\Sigma_I = ACS\Sigma$. Number of parameters for all actions common in previous intermediate stack machine and this intermediate stack machine are the same. Commit actions contain the second data field. They are of the form: $com(pop,d,k)$ where $d \in \mathcal{D}$ and $k \in \mathbb{N}$. 
\item Initial state consists of an empty strict partial order and a function mapping ever operation to $N$: ${q_0}_I = (\emptyset, \emptyset, f_{{q_0}_I})$ where  $f_{{q_0}_I}(i) =N$ for all $i \in \mathbb{N}$.
\item We define $\delta_I$ less formally, by giving verbal explanations to the transitions, omitting the obvious updates on the $f$ part and not mentioning about the parts of the nodes that does not change:
\begin{itemize}
\item $(q, inv(push,d,k),q') \in \delta_I$ iff a new node $n=(k,d,\texttt{PENDING}, \emptyset, con_n)$ is added to $nd_{q'}$, where $con_n = \{ i \in \mathbb{N}| f_q(i) = R_0\}$; $n' \rightarrow n$ will be added to $ed_{q'}$ if $n'$ is a closed node at the state $q$.
\item $(q, ret(push,k),q') \in \delta_I$ iff either there is a \texttt{PENDING} node $n$ in state $q$ and this node becomes \texttt{CLOSED} in state $q'$ or there is no node with identifier $k$ in $q$ and nothing else than $f$ field changes in $q'$.
\item $(q, inv(pop,k),q') \in \delta_I$ iff for every open node $n \in nd_q$, $k$ is added to $n.con$ in $q'$ and for every maximally closed node $m \in nd_q$, $k$ is added to $m.mc$ in state $q'$. 
\item $(q, com(pop,k,d), q') \in \delta_I$ iff there exists a node $n$ in state $q$ such that $n.d = d$ and either $k \in n.con$ or $k \in n.mc$, this node $n$ and all the nodes adjacent to it are removed in the state $q'$, $k$ is removed from all $con$ and $mc$ fields of all nodes in state $q'$ and for all other pop operations $k'$ in $n.mc$ or in $n.con$ and for all states $n' \in nd_q$ such that $n' \rightarrow n \in ed_q$ and for all $n'' \in nd_q$, $n' \rightarrow n'' \in ed_q$ implies $k' \in n''.con$ (implies $k \notin n''.mc$ but it is stronger than this condition: if there are three closed states $p,q,r$ s.t. $p \rightarrow q$, $q \rightarrow r$, $p \rightarrow r$, $k'$ is only in $r.mc$ and we delete $r$, former condition only allows $k' \in q.mc$ whereas the latter one allows $k' \in p.mc$ in addition) we have $k' \in n'.mc$ in the state $q'$. Note that we need to assume data independence to make this action deterministic.
\item $(q, ret(pop,k,d), q') \in \delta_I$ iff $q=q'$ ignoring the $f$ fields.
\end{itemize}
\end{itemize}
$L_A$ is the same machine defined in the previous section. The common alphabet between $L_A$ and $L_I$ is $A\Sigma$. We will show that they are equivalent in terms of this alphabet.

\begin{lem}
$L_I$ is a refinement of $L_A$.
\end{lem}
\begin{proof}
We will provide a backward simulation relation $bs$ between states of $L_I$ and states of $L_A$. Our relation $bs \subseteq Q_I \times Q_A$ relates state $q=(so_q, f_q)$ to $q' =(s_{q'}, f_{q'}$ iff (i) for all operation identifiers $k \in \mathbb{N}$, if $f_q(k) \in \{N,R_1,R_2\}$, then $f_q(k) = f_{q'}(k)$; if $f_q(k) = A_0$, then $f_{q'}(k) = A_0$ and the data value $d$ associated with this add operation is not inserted to $s_{q'}$ or $f_{q'}(k) = A_1$;  if $f_q(k) = A_1$, then $f_{q'}(k) = A_2$; if $f_q(k) = R_0$, then $f_{q'}(k)=R_0$ or $f_{q'}(k)=R_1$; (ii) Let us call a pop operation pending if $f_{q}(k) = R_0$ and $PP_q$ be set of pending pops. There there exists a function $g: PP_q \rightarrow ND_q \cup \{\texttt{NONE}\}$ such that $k \in g(k).mc$ or $k \in g(k).con$ for all pop operation identifiers $k$ such that $g(k) \neq \texttt{NONE}$ and $g$ is one-to-one if we neglect \texttt{NONE}; $s_{q'}$ is obtained by extending $so_q$ to a total order in which pending nodes may not take place and nodes (open or closed) $n$ such that there exists a pop operation with identifier $k$ so that $g(k)=n$ surely do not take place, (iii) $g(k) = \texttt{NONE}$ implies $f_{q'}(k) = R_0$ and $g(k) \neq \texttt{NONE}$ implies $f_{q'}(k) =R_1$, (iv) if $g_{q'}(k) =n$ and $n$ is a pending node, then $f_{q'}(k') = A_1$ where $k'$ is the operation identifier part of $n$, (v) if there is a pending node with identifier $k$ and it takes part in the linearization, then $f_{q'}(k) = A_1$.

Now, we will show that $bs$ is a backward simulation relation:
\begin{itemize}
\item[$\langle i \rangle$] $bs[{q_0}_I] =\{{q_0}_A\}$
\item[$\langle ii-a-push \rangle$] Let $(q,inv(push,d,k),q') \in \delta_I$ and $t' \in bs[q']$. We consider two cases: Either the newly added node in $q'$ takes place in the linearization and there exists an index $i$ such that $s_{t'}(i) = d$ or this new node does not exist in the linearization. For the former case construct $s_t = \langle s_{t'}(1), s_{t'}(2),..., s_{t'}(i-1) \rangle$ and $f_t = f_{t'}$ for all the operations except the ones of which data values are linearized as $s_{t'}(j)$,  for $j\geq i$. For those nodes, $f_{t'}(k') = A_1$ whereas we assign $f_t(k') = A_0$ for $j>i$ and $f_t(k) = N$. Let operation identifiers of these nodes be $k_j$ for $j>i$. Then, $t \xrightarrow{\alpha} t'$ holds where $\alpha = inv(push,d,k), lin(push,d,k), lin(push,d_{i+1},k_{i+1}),....$. Moreover, $t \in bs[q]$ since $s_t$ is a valid linearization of $so_q$ using the same $g$ function and omitting more open nodes and $f_t$ obeys the conditions. For the latter case, we pick $s_t = s_{t'}$. We have two subcases:  There exists a pending pop with identifier $k'$ such that $g(k')$ is the new node with identifier $k$ or not. For the first subcase, we pick $f_t = f_{t'}$ except $f_t(k) = N$ and $f_t(k') = R_0$. Then $t xrightarrow{inv(push,d,k),lin(pop,d,k')} t'$ is a path in $L_A$ and $t \in bs[q]$. For the second subcase, we just pick $f_t = f_{t'}$ except $f_t(k) = N$. Then, it is easy to see that $t \xrightarrow{inv(push,d,k)} t'$ holds and $t \in bs[q]$. 
\item[$\langle ii-a-pop \rangle$] Let $(q, inv(pop,k),q') \in \delta_A$ and $t' \in bs[q']$. We will again consider two cases: When relating $t'$ to $q'$, either $g(k) = \texttt{NONE}$ or $g(k)$ is a node in $q'$. In other words, either the newly invoked pop operation $k$ did not linearize yet or it linearizes and removes an element inserted by a linearized push. The second case also splits into two cases: The element removed by pop $k$ is inserted by a push $k'$ that is still pending or the push has returned. We will look at all three cases separately. The easiest one is the first case. Construct $s_t=s_{t'}$ and $f_t = f_{t'}$ except that $f_t(k) = N$ whereas $f_{t'}(k)= R_0$. One can see that $t \xrightarrow{inv(pop,k)} t'$ is a step in $L_A$ and $t \in bs[q]$. For the first case of the second case, we construct $s_t = s_{t'} \circ \langle d \rangle$ where $d$ is the data of node identifier $k'$ and $f_t = f_{t'}$ except that $f_t(k) = N$ whereas $f_{t'}(k) = R_1$. One can see that $t \xrightarrow{inv(pop,k), lin(pop,d,k)}$ is a valid path in $L_A$. Moreover, $t \in bs[q]$ since $s_t$ is a valid linearization of $so_q$. This is true because the node with identifier $k'$ is a maximal node in $so_q$ and we can add it to the end of linearization of $so_{q'}$. For the second subcase, we obtain $s_t$   from $s_{t'}$ by the following procedure: Let $n$ be the node with identifier $k'$ and $k''$ be the node such that $k' \rightarrow k''$ is an edge in $so_{q'}$, $k''$ takes part in the linearization of $so_{q'}$ to $s_{t'}$ and it has the minimum index $i$ in the $s_{t'}$ among all such nodes. Then, $s_t = \langle s_{t'}(1), s_{t'}(2),..., s_{t'}(i-1), d \rangle$ where $d$ is the data value of node with identifier $k$. Let $k_j$ and $d_j$ be the identifiers and data values of nodes that constitute $s_{q'}(j)$ for $j>i$. Clearly, $t \xrightarrow{\alpha} t'$ is a path in $L_A$ where $\alpha = inv(pop,k), lin(pop,d,k), lin(push, d_i, k_i),...$. Moreover, $t \in bs[q]$ because $s_t$ is a valid linearization of $so_q$. It is true because $so_q = so_{q'}$, $k'$ is a maximally closed node in $so_q$ and all the nodes with identifier $k_j$ ($j>i$) are pending nodes.
\item[$\langle ii-c \rangle$] Let $(q, com(pop,d,k), q') \in \delta_I$ and $t' \in bs[q']$. We will consider two cases. The first case is commit action removes a maximally closed or an open node. For these cases, we can construct $t$ as in the second case of the previous item (invoke pop case). The same arguments apply for constructing a path between $t$ and $t'$ in $L_A$ and showing that $t \in bs[q]$.  The second case is that commit action removes a non-maximal closed node. This time, pick $t = t'$. Then, $t \xrightarrow{\epsilon} t'$ is the path in $L_A$ and one can show that $t \in bs[q]$ by choosing $g(k)$ as the node that is removed. 
\item[$\langle ii-b-push \rangle$] Let $(q,ret(push,k),q') \in \delta_I$ and $t' \in bs[q']$. We consider two cases. Either there is a node $n$ with identifier $k$ and data value $d$ in state $q$ or not. For the first case, we consider two subcases. Either this node takes part in linearization or not (if there exists a pop $k'$ such that $g(k')=n$). For the first subcase, we can pick $s_t = s_{t'}$ and $f_q = f_{q'}$ except that $f_q(k) = A_0$.  Then, $t \xrightarrow{ret(push,k)} t'$ is a path in $L_A$. Moreover, $t \in bs[q]$ since $n$ is a maximal node in $q$. For the second subcase, we pick $s_t = s_{t'} \circ \langle d \rangle$ and $f_t = f_{t'}$ except that $f_t(k) = A_1$ and $f_t(k') = R_0$. Then, $t \xrightarrow{lin(pop,d,k'), ret(push,k)} t'$ is a path $L_A$ and $t \in bs[q]$ since $n$ is a maximal open node in $q$. For the second case, we pick $s_t = s_{t'}$ and $f_t = f_{t'}$ except that $f_t(k) = A_1$. Then, $t \xrightarrow{ret(push,k)} t'$ is a path in $L_A$ and $t \in bs[q]$. 
\item[$\langle ii-b-pop \rangle$] Let $(q,ret(pop,d,k),q') \in \delta_I$ and $t' \in bs[q']$. We pick $s_t = s_{t'}$ and $f_t = f_{t'}$ except that $f_t(k) =R_1$. One can see that $t \xrightarrow{ret(pop,d,k)} t'$ is a valid action in $L_A$ and $t \in bs[q]$.
\end{itemize}
\end{proof}

\begin{lem}
$L_A$ is a refinement of $L_I$.
\end{lem}
\begin{proof}
We will construct a forward simulation relation $fs$ between $L_A$ and $L_I$. Our relation $fs \subseteq Q_A \times Q_I$ relates state $q = (s_q, f_q) \in Q_A$ to a state $q' =(so_{q'}, f_{q'}) \in Q_I$ iff (i) for all operation identifiers $k \in \mathbb{N}$, if $f_q(k) \in \{N, A_0, R_0, R_1, R_2\}$ then $f_{q'}(k) = f_q(k)$; if $f_q(k) = A_1$, then $f_{q'}(k) = A_0$; if $f_q(k) = A_2$, then $f_{q'}(k)= A_1$; (ii) We form nodes of $so_{q'}$ ($ND_{q'}$ as follows:  If $k$ is a push operation adding data value $d$ and either $f_q(k)= R_0$ or $f_q(k)=R_1$ and data added by this push exists in $s_q$, then there is a \texttt{PENDING} node in $ND_{q'}$ with identifier $k$ and data value $d$. If $f_q(k)= R_2$, then there is a \texttt{CLOSED} node in $nd_q$ with identifier $k$ and data value $d$. If there is an operation identifier $k$ such that $f_q(k) = R_0$, we call this a pending pop and this pop takes place $mc$ or $con$ fields of nodes of $q'$. If $n \in ND_{q'}$ is a \texttt{PENDING} node, then $k \in n.con$. If $n$ is maximally closed, then $k \in n.con$ or $n.mc$. If $k \in n.mc$ or $k \in n.con$ and there exists another node $n' \in ND_{q'}$ such that $n \rightarrow n' \in ED_{q'}$, then $k \in n'.con$. If $k \in n.mc$ and there exists another node $n' \in ND_{q'}$ such that $n' \rightarrow n \in ED_{q'}$, then neither $k \in n'.mc$ nor $k \in n'.con$. For any node $n \in ND_{q'}$, either $k \in n.mc$ or $k \in n.con$. (iii) We form edges of $so_{q'}$ ($ED_{q'}$) as follows: \texttt{PENDING} nodes are maximal. Edges obey the strict partial order conditions. (iv) We can find a linearization $so_{q'}$ that is equal to $s_q$. \texttt{PENDING} nodes may not participate in the linearization. Note that we do not need $g$ function for keeping track of linearized pops unlike the previous proof.

Next, we will show that $fs$ is a forward simulation relation.
\begin{itemize}
\item[$\langle i \rangle$] $fs[{q_0}_A] = \{{q_0}_I\}
$ 
\item[$\langle ii-a-push \rangle$] Let $(q, inv(push,d,k), q') \in \delta_A$ and $t \in fs[q]$. Pick $t'$ such that $(t, inv(push,d,k), t') \in \delta_I$. Since $s_q=s_{q'}$ and $so_{t'}$ contains a maximal open new node as the only difference from $so_t$, we can linearize $so_{t'}$ so that linearization is equal to $s_{q'}$. By checking the other conditions, one can observe that $t' \in fs[q']$.
\item[$\langle ii-a-pop \rangle$] Let $(q, inv(pop,k), q') \in \delta_A$ and $t \in fs[q]$. Pick $t'$ such that $(t, inv(pop,k), t') \in \delta_I$. Only difference between $t$ and $t'$ is that maximally closed and open nodes in $t'$ contain $k$ in their $mc$ or $con$ fields. Since this new addition obeys our forward simulation definition, $t' \in fs[q']$.
\item[$\langle ii-c-push \rangle$] Let $(q, lin(push,d,k) q') \in \delta_A$ and $t \in fs[q]$. Pick $t'=t$. $s_{q'}$ is still a linearization of $so_t$ since the node with identifier $k$ is a maximal node in $ND_t$ and we can linearize it at the end. So, $t' \in fs[q']$ holds.
\item[$\langle ii-c-pop \rangle$] Let $(q, lin(pop,d,k), q') \in \delta_A$ and $t \in fs[q]$. Pick $t'$ such that $(t, com(pop,d,k) t') \in \delta_I$. The action $com(pop,d,k)$ is a valid action in $L_I$ because $d$ is the last element in $s_q$. Hence, the node $n$ with identifier $k$ and data value $d$ is a maximal element in $so_q$ and either $k \in n.mc$ or $k \in n.con$ by the properties of $fs$. Hence, the node with identifier $k$ can be removed by a $com$ action. In addition, $s_{q'}$ is a linearization of $so_{t'}$ because removed node is a maximal node and $s_{q'}$ is obtained from $s_q$ by deleting the maximum node. Hence, $t' \in fs[q']$ holds.
\item[$\langle ii-b-push \rangle$] Let $(q, ret(push,k), q') \in \delta_A$ and $t \in fs[q]$. We will consider two cases: either the element inserted by the push with identifier $k$ is removed by a concurrent pop or the element is still in $s_q$. For the former case, there is no node with identifier $k$  in state $t$ and we can pick $t'$ such that $(t, ret(push,k), t') \in \delta_I$. We have $so_t = so_{t'}$ for this case. Since $s_q = s_{q'}$ also holds, $s_{t'}$ is a linearization of $so_{q'}$ and $t' \in fs[q']$. For the latter case, we again pick $t'$ such that $(t, ret(push,k), t') \in \delta_I$ holds. This time, only difference between $so_t$ and $so_{t'}$ is that the node with identifier $k$ is \texttt{CLOSED} in $so_{t'}$. Since the edges are the same, $s_{q'}$ is a valid linearization of $s_{t'}$ and $t' \in fs[q']$ holds. 
\item[$langle ii-b-pop \rangle$] Let $(q, ret(pop,d,k), q') \in \delta_A$ and $t \in fs[q]$. We pick $t'$ such that $(t, ret(pop,d,k) t') \in \delta_I$. Only difference between $t$ and $t'$ is that $f_{t'}(k) = R_2$ whereas $f_t(k) = R_1$. Since $s_q = s_{q'}$ and $so_t = so_{t'}$, $s_{q'}$ is a valid linearization of $so_{t'}$. By checking the other conditions, we see that $t' \in fs[q']$.  
\end{itemize}
\end{proof}

Now, we show that TS-Stack is linearizable by showing the concrete TS-Stack implementation $L_C$ is a cs-refinement of $L_I$. As $L_C$, we pick the simplest version that omits the \texttt{EMPTY} returns of the pop  methods, does not allow unlinking and elimination.

### 2017-01-11 draftCAV/prelim.tex
started draft for CAV

%!TEX root = draft.tex
\section{Preliminaries}

We formalize several abstraction relations between libraries using a simple
yet universal model of computation, namely labeled transition systems (LTS).
This model captures shared-memory programs with an arbitrary number of threads,
abstracting away the details of any particular programming system irrelevant to
our development.

A \emph{labeled transition system} (LTS) $A=(Q,\Sigma, q_0, \delta)$ over the 
possibly-infinite alphabet $\Sigma$ is a possibly-infinite set $Q$ of states with
initial state $q_0 \in Q$, and a transition relation $\delta \subseteq Q \times @S \times
Q$. The $i$th symbol of a sequence $@t \in @S^*$ is denoted $@t_i$, and the empty
sequence is denoted by $\epsilon$.
An \emph{execution} of $A$ is an alternating sequence of states and transition labels (called also actions)
$\rho = s_0, e_0,s_1\ldots e_{k-1},s_k$ for some $k>0$ such that $s_0 = q_0$ and $\delta(s_i, e_i, s_{i+1})$
for each $i$ such that $0\leq i<k$. We write $s_i\xrightarrow{e_i\ldots e_{j-1}} s_j$ for any $0\leq i\leq j <k$
(in particular $s_i\xrightarrow{\epsilon}s_i$).
The projection $@t| \Gamma$ of a sequence $@t$ is the maximum subsequence of $@t$ over
 alphabet $\Gamma$. This notation is extended to sets of sequences as usual.
A \emph{trace} of $A$ is the projection $\rho | \Sigma$ of an execution $\rho$ of $A$. 
The set of executions, resp., traces, of an LTS $A$ is denoted by $E(A)$, resp., $Tr(A)$.

\subsection{Libraries}

Programs interact with libraries by calling named library \emph{methods}, which
receive \emph{parameter values} and yield \emph{return values} upon completion.
We fix arbitrary sets $\<Methods>$ and $\<Vals>$ of method names and
parameter/return values.

%\begin{example}
%  \label{ex:methods}
%
%  TODO
%
%  The method and value sets for the stack implementation in
%  Figure~\ref{fig:treiber} are $\<Methods> = \set{ \<push>, \<pop> }$ and
%  $\<Vals> = \<Nats> \u \set{ {\tt EMPTY} }$.
%
%\end{example}

\noindent
We fix an arbitrary set $\<Ops>$ of operation identifiers, and for given sets
$\<Methods>$ and $\<Vals>$ of methods and values, we fix the sets
\begin{align*}
  & C = \set{ inv(m,d,k) : m \in \<Methods>, d \in \<Vals>, k \in \<Ops> }
  \text{, and } \\
  & R = \set{ ret(m,d,k) : m \in \<Methods>, d \in \<Vals>, k \in \<Ops> }  
\end{align*}
of \emph{call actions} and \emph{return actions}; each call action $inv(m,d,k)$
combines a method $m \in \<Methods>$ and value $d \in \<Vals>$ with an
\emph{operation identifier} $k \in \<Ops>$. Operation identifiers are used to
pair call and return actions. We denote the operation identifier of a
call/return action $a$ by $\<op>(a)$. Call and return actions $c \in C$ and $r
\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
We may omit the second field from a call/return action $a$ for methods that have no inputs or return values 
(e.g., the pop method of a stack).
A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
\emph{well formed} when:
\begin{itemize}

  \item Each return is preceded by a matching call: \\
  $\tau_j \in R$ implies $\tau_i \match \tau_j$ for some $i < j$.

  \item Each operation identifier is used in at most one call/return: \\
  $\<op>(\tau_i) = \<op>(\tau_j)$ and $i < j$ implies $\tau_i \match \tau_j$.

\end{itemize}
We say that the well-formed word $\tau \in @S^*$ is \emph{sequential} when
\begin{itemize}

  \item Operations do not overlap: \\
  $\tau_i, \tau_k \in C$ and $i < k$ implies $\tau_i \match \tau_j$ for some $i < j < k$.

\end{itemize}
Well-formed words represent traces of a library. We assume every set of well-formed
words is closed under isomorphic renaming of operation identifiers. For
notational convenience, we take $\<Ops>=\<Nats>$ for the rest of the paper.
When the value of a certain field in a call/return action is not important we use 
the placeholder $\_$, e.g., $inv(m,\_,k)$ instead of $inv(m,d,k)$ when the input  
$d$ can take any value.

An operation $k$ is called \emph{completed} in a well-formed trace $\tau$ when
$ret(m,d,k)$ occurs in $\tau$, for some $m$ and $d$. Otherwise, it is called \emph{pending}.
%An operation $o$ of an execution $e$ is \emph{completed}
%when both call and return actions $m(u)_o$ and $\<ret>(v)_o$ of $o$ occur in
%$e$, and is otherwise \emph{pending}.

%\begin{example}
%  \label{ex:executions}
%
%  TODO
%
%  The well-formed words
%  \scriptsize
%  \begin{align*}
%     \<push>(0)_1\ \<pop>_2\ \<pop>_3\ \<ret>_1\ \<ret>(0)_3\ \<ret>(0)_2
%    \text{\normalsize, and } 
%    \<push>(0)_1\ \<pop>_2\ \<pop>_3\ \<ret>_1\ \<ret>(0)_2
%  \end{align*}
%  \normalsize
%  represent executions in which one call to the $\<push>(0)$ method overlaps
%  with two calls to $\<pop>$. In the first execution both calls to $\<pop>$
%  have matching return actions $\<ret>(0)$, i.e., the operations $2$ and $3$ are completed,
%  while operation $3$ is pending in the second, it has no matching return.
%
%\end{example}

Libraries dictate the execution of methods between their call and return
points. Accordingly, a library cannot prevent a method from being called,
though it can decide not to return. Furthermore, any library action performed
in the interval between call and return points can also be performed should the
call have been made earlier, and/or the return made later. 
A library thus allows any sequence of
invocations to its methods made by \emph{some} program.

\begin{definition}\label{def:libraries}
A \emph{library} $L$ is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$
and each trace $\tau \in Tr(L)$ is well formed, and
  \begin{itemize}

    \item Call actions $c \in C$ cannot be disabled: \\
    $\tau \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot \tau' \in Tr(L)$
    if $\tau \cdot c \cdot \tau'$ is well formed.

    \item Call actions $c \in C$ cannot disable other actions: \\
    $\tau \cdot a \cdot c \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot a \cdot \tau' \in Tr(L)$.

    \item Return actions $r \in R$ cannot enable other actions: \\
    $\tau \cdot r \cdot a \cdot \tau' \in Tr(L)$ implies $\tau \cdot a \cdot r \cdot \tau' \in Tr(L)$.

  \end{itemize}

\end{definition}

The projection of a library trace over $C\cup R$ is called a \emph{history}. The set of histories of a library $L$ is denoted by $H(L)$.

Note that even a library that implements \emph{atomic methods}, e.g.,~by
guarding method bodies with a global-lock acquisition, admits executions in
which method calls and returns overlap. 
%A library which accesses the client's
%thread identifiers can be modeled by taking thread identifiers as method
%parameters.

Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
admit histories with weaker inter-operation ordering, in which calls may happen earlier, and/or returns later. This weakening relation
between histories can be formalized as follows: a history $h_1$ is \emph{weaker} than a history $h_2$, written $h_1 \sqsubseteq h_2$, 
if{f} there exists a well-formed history $h_1'$
obtained from $h_1$ by appending return actions, and deleting call actions,
such that:
\begin{quote}

  $h_2$ is a permutation of $h_1'$ that preserves the order between
  return and call actions, i.e.,~a given return action occurs before a given
  call action in $h_1'$ if{f} the same holds in $h_2$.

\end{quote}
$\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
history set $H$ under weakening.
A direct consequence of Definition~\ref{def:libraries} is that $H(L)=\overline{H(L)}$ for
every library $L$. A \emph{kernel} of L is any set $H$ such that $\overline{H} = L$.
A library $L$ is called atomic if it has a sequential kernel. 
Atomic libraries are often considered as specifications for concurrent objects. 
In practice, libraries can be made atomic by guarding their methods bodies with global lock acquisitions.

%\begin{example}
%  \label{ex:libraries}
%
%  TODO (replace executions with histories)
%
%  Any library which admits the execution
%  \scriptsize
%  \begin{align*}
%    \<push>(0)_1\ \<ret>_1\ \<pop>_2\ \<ret>(0)_2
%  \end{align*}
%  \normalsize
%  with sequential calls to $\<push>$ and $\<pop>$ must also admit
%  \scriptsize
%  \begin{align*}
%    \<push>(0)_1\ \<pop>_2\ \<ret>_1\ \<ret>(0)_2
%    \text{ \normalsize and }
%    \<push>(0)_1\ \<pop>_2\ \<pop>_3\ \<ret>_1\ \<ret>(0)_2
%    \text{\normalsize,}
%  \end{align*}
%  \normalsize
%  among others, yet need not admit an execution
%  \scriptsize
%  \begin{align*}
%    \<push>(0)_1\ \<pop>_2\ \<pop>_3\ \<ret>_1\ \<ret>(0)_3\ \<ret>(0)_2
%  \end{align*}
%  \normalsize
%  with two completed $\<pop>$ operations returning $0$.
%  
%\end{example}

%Systems we consider are labeled transition systems (LTS):
%\begin{dfn}
%An LTS is defined over four-tuples $A=(Q,\Sigma, q_0, \delta)$ where $Q$ is the set of states, $\Sigma$ is the set of transition labels, $q_0 \in Q$ is the initial state and $\delta \subseteq Q \times \Sigma \times Q$ is the transition relation.
%\end{dfn}
%Executions generated by this system are alternating sequence of states and transition labels $\rho = s_0, e_0, s_1,... s_k, e_k,...$ where each $s_i \in Q$, each $e_i \in \Sigma$, $s_0 = q_0$ and each $(s_i, e_i s_{i+1}) \in \delta$. The projection of the sequence $\rho$ over the set $\Pi$ is denoted by $\rho | \Pi$, and it is the maximum subsequence of $\rho$ consisting of elements of $\Pi$. Traces of the LTS are obtained from executions by projecting them over $\Sigma$. For the rest of the paper and in all of the proofs, we consider only finite executions (denoted as $E(A)$) and/or traces (denoted as $Tr(A)$ of the LTSs in focus.

%Libraries are LTSs that provide methods. Let $\mathcal{M}$ be the set of method names and $\mathcal{D}$ be the domain of values as input/output parameters for the methods. Then, this library contains transition labels of the form $inv(m,d,i)$ representing the invocation of method $m \in \mathcal{M}$ with input value $d \in \mathcal{D}$. The third field is the operation identifier for differentiating the different calls of the same method from the set $\mathcal{O}$. For simplicity, we take $\mathcal{O} = \mathbb{N}$ for the rest of the paper. We also assume that methods could have at most one input parameter. If they do not have any input arguments (like pop method of a stack), we can omit the second field from the action. They also provide actions of the form $ret(m,d,i)$ representing the return of method $m \in M$ with value $d \in D$ which has been invoked previously with action $inv(m,d',i)$. Again, we assume that the methods can return at most one parameter and we may omit the second field from the action if they have none (like enqueue method of a queue). Before starting to reason about any set of libraries, we first fix the sets $\mathcal{M}$ and $\mathcal{D}$ and libraries in our focus agree on this sets. For any transition label $e = inv(m,d,i)$ or $e=ret(m,d,i)$, we have the function $oid(e) = i$.
%
%Since libraries are LTSs, they produce traces. A trace $e = e_1, e_2, ..., e_n$ of library $L$ is \emph{well-formed} iff (i) every return matches an earlier invocation: $e_j = ret(m,d,k)$ implies that there exists $i<j$ such that $e_i = inv(m,d',k)$ and (ii) every operation identifier is used at most one invocation/return pair: $oid(e_i) = oid(e_j) = k$ and $i<j$ implies $e_i = inv(m,d,k)$ and $e_j = ret(m,d',k)$. From now on, we assume that libraries produce well-formed traces. Let $f: \mathbb{N} \rightarrow \mathbb(N)$ be a bijection. Then, traces $e$ and $e'$ are equivalent if $e'$ is obtained from $e$ by replacing every action $inv(m,d,k)$ with $inv(m,d,f(k))$ and every action $ret(m,d,k)$ with $ret(m,d,f(k))$.

\subsection{Refinement and Linearizability}

Conformance of a library $L_1$ to a specification given as an ``abstract'' library $L_2$ 
is formally captured by \emph{(observational) refinement}. Informally, given two libraries
$L_1$ and $L_2$ implementing the methods of some concurrent object, we
say $L_1$ refines $L_2$ if{f} every computation of every program
using $L_1$ would also be possible were $L_2$ used instead. As shown in~\citet{DBLP:conf/popl/BouajjaniEEH15}
this can be equivalently stated as history set inclusion between the two libraries.

\begin{dfn}
Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$~\footnote{The two libraries could be defined over different alphabets as long as they share the same set of call/return actions.}. 
%agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets. We define the set $A\Sigma = \{inv(m,d,i)| m \in \mathcal{M} \wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\} \cup \{ret(m,d,i)| m \in \mathcal{M}\wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\}$ as abstract transition labels. Note that $A\Sigma \subseteq \Sigma_{L_1}$ and $A\Sigma \subseteq \Sigma_{L_2}$. 
The library $L_1$ \emph{refines} $L_2$ if{f} $H(L_1) \subseteq H(L_2)$.
%We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
%for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
\end{dfn}
%Linearizability is also a relation between two libraries and it is stricter than refinement. It requires $e'$ in Definition 2 to be a sequential one. A trace $e$ is sequential iff following two conditions hold for its projection to abstract transition labels $e|A\Sigma = e_1, ...e_n$: $(i)$ $e_1 = inv(m,d,k)$ for some $m \in \mathcal{M}, d \in \mathcal{D} and k \in \mathbb{N}$ and $(ii)$ for all $i \in [1,n)$, either $e_i = inv(m,d,k)$ and $e_{i+1} = ret(m,d',k)$ or $e_i = ret(m,d,k)$ and $e_{i+1} = inv(m',d',k')$ for some $m,m' \in \mathcal{M}$, $d,d' \in \mathcal{D}$ and $k,k' \in \mathbb{N}$.

The linearizability criterion~\cite{journals/toplas/HerlihyW90} provides an
alternative characterization of library conformance which implies 
refinement~\cite{journals/tcs/FilipovicORY10}. 
The two criteria coincide when the reference implementation is atomic~\cite{DBLP:conf/popl/BouajjaniEEH15}.
Linearizability~\cite{journals/toplas/HerlihyW90} is defined by the weakening order $\sqsubseteq$. A 
history $h_1$ is \emph{linearizable} w.r.t.~a library $L_2$ if{f} there exists a sequential history
$h_2 \in H(L_2)$, with only
completed operations, such that $h_1 \sqsubseteq h_2$. A library $L_1$
is \emph{linearizable} w.r.t. $L_2$, written $L_1 \sqsubseteq L_2$, if{f}
each history $h_1 \in H(L_1)$ is linearizable w.r.t. $L_2$.

\begin{theorem}[\cite{DBLP:conf/popl/BouajjaniEEH15}]
  $L_1 \sqsubseteq L_2$ if{f} $L_1$ refines $L_2$,
  if $L_2$ is atomic.

\end{theorem}

The weakening order w.r.t. sequential histories can be also defined  in terms of ~\emph{linearization points}. Informally, a linearization point 
of an operation in an execution is a point in time where the operation is conceptually effectuated. It corresponds to a transition in that
execution that occurs between the call and the return action of the operation.
A \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that
\begin{itemize}
	\item every linearization point corresponds to an operation that has been called, i.e., TODO
	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
	\item the return value of an operation is fixed by the linearization point, i.e., TODO
\end{itemize}
An LP-annotation is called \emph{complete} when every completed operation has a linearization point, i.e., TODO.
An operation $k$ is called \emph{effectuated} in a complete LP-annotation $\hat{h}$ iff  $\hat{h}$ contains $lin(\_,\_,k)$.

It has been shown that $h_1 \sqsubseteq h_2$ for a sequential history $h_2$ if{f} there exists a complete LP-annotation $\hat{h_1}$ such that  
$h_2$ is a permutation of the history $h_1'$ obtained from $h_1$ by preserving only call/return actions of effectuated operations that preserves the linearization point order, i.e.,~a given return action $ret(\_,\_,k)$ occurs before a given
  call action $call(\_,\_,k')$ in $h_2$ if{f} $lin(\_,\_,k)$ occurs before $lin(\_,\_,k')$ in $\hat{h_1}$.

Refinement relations can be approximated using simulation relations, i.e., the existence of a simulation relation implies refinement. We define a notion of \emph{forward} simulation whose existence is equivalent to refinement only when the right-hand side library $L_2$ is deterministic. A dual notion of \emph{backward} simulation is defined in Appendix~\ref{app:backSim}. The existence of backward simulations is equivalent to refinement for arbitrary libraries.
%We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement. 
For a relation $R\subseteq A\times B$, $R[X]$ denotes the set of elements of $B$ related to elements of $X$ by $R$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.

%We can extend the relations between libraries by introducing simulation relations. We will later show that simulation relations imply refinement.

\begin{dfn}
Let $L_1=(Q_1,\Sigma, q_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, q_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$. A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{forward simulation} from $L_1$ to $L_2$ iff the following holds:
\begin{itemize}
\item[(i)] $\mathit{fs}[q_0^1] = \{q_0^2 \}$
\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0<i<|@s|$.
%$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s| -1$.
% where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-c)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus (C\cup R)$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}
If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function.

Simulation relations are used to prove refinement among libraries. The following lemma shows their soundness:
\begin{lem}
Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$. If there exists a forward simulation from $L_1$ to $L_2$, then $L_1$ refines $L_2$. Moreover, if $L_2$ is deterministic and $L_1$ refines $L_2$, then there exists a forward simulation from $L_1$ to $L_2$.
\end{lem}
\begin{proof}
Looks trivial and follows Lynch paper. Can be completed later.
\end{proof}

### 2017-01-11 draftCAV/queue.tex
started draft for CAV

\section{Existence of Forward Simulations for Queue Implementations that have Fixed Dequeue Linearization Points}
In this section, we will show that for any concurrent queue implementation library $L_C$ for which we know the linearization points of the dequeue operation and that is linearizable with respect to the reference implementation library $L_A$, there exists a forward simulation $fs$ relating $L_C$ to $L_I$ where $L_I$ is an intermediate library equivalent to $L_A$ i.e. $L_I$ refines $L_A$ and vice versa.

For all the queue libraries, we fix $\mathcal{M} = \{ enq, deq \}$ and $\mathcal{D} = \mathbb{N} \cup \{\texttt{EMPTY} \}$. Since we know the linearization point of dequeues, we extend the definitions of the previous sections adding this information. First, we extend the set $A\Sigma$ introduced in Definition 2 for queues in our focus as $AQ\Sigma = A\Sigma \cup \{lin(deq,d,k)| d \in \mathcal{D}, k \in \mathbb{N}\}$. For any library $L$ we consider in this section, $AQ\Sigma \subseteq \Sigma_L$. Then, we define q-refinement by replacing $A\Sigma$ with $AQ\Sigma$ in Definition 2. We also define q-linearizability, by enforcing $lin(deq,d,k)$ to appear immediately after $inv(deq,k)$ and immediately before $ret(deq,d,k)$ in a sequential execution. We also change definition of forward and backward simulation relations by replacing $A\Sigma$ with $AQ\Sigma$ in the original definitions and adding a new condition $(ii-d)$ to each of them:
\begin{itemize}
\item[Forward Simulation: (ii-d)] If $(s, lin(deq,d,k), s') \in \delta_{L_1}$ and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $(u, lin(deq,d,k), u') \in \delta_{L_2}$
\item[Backward Simulation: (ii-d)] If $(s, lin(deq,d,k), s') \in \delta_{L_1}$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $(u, lin(deq,d,k), u') \in \delta_{L_2}$
\end{itemize}
Lemma 1 of the previous section still holds if we replace refinement with q-refinement and use new definitions of backward and forward simulations.

Next, we define the abstract library $L_A$ as follows:
\begin{itemize}
\item A queue state consists of a finite sequence of natural numbers representing the queue content and a program counter for each operation. More formally: $Q_A \subseteq \mathbb{N}^*  \times (\mathbb{N} \rightarrow Lbl_A)$ where $Lbl_A = \{N, E_0, E_1, E_2, D_0, D_1, D_2 \}$ is the set of transition labels of the operations. Operations that have not started yet are mapped to $N$. $E_i$ ($D_i$) denotes particular transitions in enqueue (dequeue) operations that will be clear when we define $\delta_A$. For a state $q$, we denote the contents of the queue (first field) with $s_q$ and the function that maps operations to labels (the second field) with $f_q$.
\item Transition labels consists of invocations, returns and linearizations of enqueue and dequeue operations: $\Sigma_A = AQ\Sigma \cup \{lin(enq,d,k)| d \in \mathcal{D}, k \in \mathbb{N}\}$. Invocation of $deq$ operation does not have any input and return of $enq$ operation does not have any output value. We omit second fields from these labels.
\item Initial state is the empty queue: ${q_0}_A = (\langle \rangle, f_{{q_0}_A})$ where $f_{{q_0}_A}(i) = N$ for all $i \in \mathbb{N}$.
\item Each operation consists of invocation, linearization and return steps. These are reflected in the transition relation. For the below definitions, unchanged parts of the state are omitted from the formulae:
\begin{itemize}
\item $(q, inv(enq,d,k), q') \in \delta_A$ iff $ d \neq \texttt{EMPTY} \wedge f_q(k) = N \wedge f_{q'}(k) = E_0$,
\item $(q,lin(enq,d,k),q') \in \delta_A$ iff $f_q(k) =E_0 \wedge s_{q'} = s_q \circ \langle d \rangle \wedge f_{q'}(k) = E_1$ where $\circ$ is the operation that concatenates two finite sequences,
\item $(q,ret(enq,k),q') \in \delta_A$ iff $f_q(k) = E_1 \wedge f_{q'}(k) = E_2$,
\item $(q, inv(deq,k), q') \in \delta_A$ iff $f_q(k) = N \wedge f_{q'}(k) = D_0$
\item $(q,lin(deq,d,k),q') \in \delta_A$ iff $f_q(k) =D_0 \wedge (d \neq \texttt{EMPTY} \wedge s_q = \langle d \rangle \circ s_{q'} \vee d=\texttt{EMPTY} \wedge s_q = s_{q'} = \langle \rangle )\wedge f_{q'}(k) = D_1$, 
\item $(q,ret(deq,d,k),q') \in \delta_A$ iff $f_q(k) = D_1 \wedge f_{q'}(k) = D_2$.
\end{itemize}

\end{itemize} 
We restrict traces generated by this LTS by neglecting the invalid traces. If we project a trace to some operation identifier $k$ and obtain a sequence $\langle ..., inv(enq,d,k), lin(enq,d',k),...\rangle$ or $\langle ..., lin(deq,d,k), ret(deq,d',k),...\rangle$ where $d \neq d'$ we say that this trace is invalid.

We want to introduce $L_I$ as the next step. States of $L_I$ consists of strict orders of enqueue operations. Nodes of the strict order come from the set $ND = \mathbb{N} \times \mathcal{D} \times \{ \texttt{PENDING}, \texttt{CLOSED}\}$. Basically each node is a tuple keeping the operation ID, the value to be enqueued and if this enqueue is pending or closed. Strict order also keeps a set of directed edges $ ed \subseteq ED = ND \times ND$. Then, a strict order is a tuple $(nd, ed)$ where the set $ed$ obeys the assumption of strict order i.e. irreflexivity, asymmetry and transitivity. Then, the LTS of $L_I$ is defined as follows:
\begin{itemize}
\item A state consists of a partial order and a function keeping the program counter. More formally $Q_I \subseteq ND \times ED \times (\mathbb{N} \rightarrow Lbl_I)$ where $Lbl_I = \{N, E_0, E_1, D_0, D_1, D_2\} $ is the set of transition labels. $nd_q$ $ed_q$ and $f_q$ denotes the nodes of the strict order in state $q$, edges of the strict order in state $q$ and the function that maps operation to labels in state $q$, respectively. 
\item The transition label set consists of invocation and return action of both methods and linearization action of only the dequeue method: $\Sigma_I = AQ\Sigma$. Invocation of $deq$ operation does not have any input and return of $enq$ operation does not have any output value. We omit second fields from these labels.
\item Initial state consists of an empty strict order and a function mapping every operation to $N$: ${q_0}_{I} = (so_{{q_0}_I}, f_{{q_0}_I}$ where $so_{{q_0}_I} =(\emptyset, \emptyset)$ and $f_{{q_0}_I}(i) = N$ for all $i \in \mathbb{N}$.
\item While defining $\delta_I$, we again omit mentioning about the parts that has not changed:
\begin{itemize}
\item $(q, inv(enq,d,k), q') \in \delta_I$ iff $f_q(k) = N \wedge f_{q'}(k) = E_0 \wedge d \neq \texttt{EMPTY} \wedge (k,\_,\_) \notin nd_q \wedge nd_{q'} = nd_q \cup \{ (k,d,\texttt{PENDING}\} \wedge AddNode(ed_{q'}, ed_q,k)$ where $AddNode(ed_{q'}, ed_q,k)$ is true iff $ed_{q'}$ is obtained from $ed_q$ by adding edges from every closed node of $ed_q$ to the node with identifier $k$.
\item $(q, ret(enq,k), q') \in \delta_I$ iff $f_q(k) =E_0 \wedge f_{q'}(k) = E_1 \wedge ( (k,\_,\texttt{PENDING}) \notin nd_q \wedge so_q = so_{q'} \vee UpdateNode(so_{q'}, so_q, k) )$ where $UpdateNode(so_{q'}, so_q, k)$ is true iff $(k,d,\texttt{PENDING}) \in nd_q$ for some $d \in \mathcal{D}$, it is replaced with node $(k,d,\texttt{CLOSED}$ in the state $q'$ and all the edges adjacent to $(k,d,\texttt{PENDING})$ in state $q$ are replaced with edges adjacent to $(k,d,\texttt{CLOSED})$ in the state $q'$.
\item $(q, inv(deq,k), q') \in \delta_I$ iff $f_q(k) = N \wedge f_{q'}(k) = D_0$.
\item $(q, lin(deq,d,k), q') \in \delta_I$ iff $f_q(k) = D_0 \wedge f_{q'}(k) = D_1 \wedge (d = \texttt{EMPTY} \wedge nd_q = ed_q = nd_{q'} = ed_{q'} \emptyset \vee d \neq \texttt{EMPTY} \wedge RemoveNode(so_q, so_{q'},d))$ where $RemoveNode(so_q, so_{q'},d)$ is true iff there is a minimal node $n \in so_q$ of which data value (second field) is $d$ and $so_{q'}$ is obtained from $so_q$ by removing $n$ and all the edges adjacent to it. Note that linearization of dequeue could remove a \texttt{PENDING} node.  
\item $(q, ret(deq,d,k) q') \in \delta_I$ iff $f_q(k) = D_1 \wedge f_{q'}(k) = D_2$.
\end{itemize}
\end{itemize}
Again, we omit the inconsistent traces from $L_I$ as in $L_A$. Note that the library $L_I$ is deterministic with respect to alphabe $AQ\Sigma$.

Next, we show equivalence of $L_A$ and $L_I$ in terms of q-refinement.
\begin{lem} 
$L_I$ is a q-refinement of $L_A$.
\end{lem}
\begin{proof}
We will provide a backward simulation relation btw $L_I$ and $L_A$. Given a state $q = (so_q, f_q) \in Q_I$,  $q' \in bs[q]$ iff (i) $s_{q'}$ is a linearization of $so_q$ projected to $d$ fields (second field). This linearization may omit the pending elements (the ones of which third field is \texttt{PENDING}), but it must contain all the \texttt{CLOSED} elements. Note that pending elements of $so_q$ are maximal and they can appear after the closed elements, towards the end of the sequence. (ii) If $f_q(k) \in \{N, D_0, D_1, D_2\}$, then $f_{q'}(k) = f_q(k)$ . If $f_q(k) = E_0$, $(k,d,\texttt{PENDING}) \in nd_q$ and this node participates in the linearization $s_{q'}$. Then $f_{q'}(k) = E_1$. If it does not participate in the linearization, then $f_{q'}(k) = E_0$. If $f_q(k) = E_0$ but there is no node with operation identifier $k$ in $nd_q$, then $f_{q'} = E_1$. Lastly, if $f_q(k) = E_1$ then $f_{q'}(k) = E_2$. Next, we will show that $bs$ is a backward simulation relation.
\begin{itemize}
\item[$\langle i \rangle$] $bs[{q_0}_I] = \{ {q_0}_A \}$.
\item[$\langle ii-a-enq \rangle$] Let $(q, inv(enq,d,k), q') \in \delta_I$ and $t' \in bs[q']$. We know that $(k,d,\texttt{PENDING}$ is a maximal element in $so_{q'}$ and either $s_{t'}$ does not linearize it or $s_{t'} = \rho \circ \langle d \rangle \circ \pi$ where $\pi$ consists of only linearization of \texttt{PENDING} elements. For the first case $s_{t'} = s_t$ for some $t \in bs[q]$ and $(t,inv(enq,d,k),t') \in \delta_A$. For the latter case, $s_t = \rho$ for some $t \in bs[q]$ and $t \xrightarrow{a}_{L_A} t'$ where $a = inv(enq,d,k), lin(enq,d,k), lin(enq,d_1,k_1),...,lin(enq,d_j,k_j)$ where $\pi = d_1,...d_j$. The $f_t$ could be obtained easily by observing the sequence $a$ and it can be checked that such $t \in bs[q]$ exists.
\item[$\langle ii-a-deq \rangle$] Let $(q, inv(deq,k), q') \in \delta_I$ and $t' \in bs[q']$. We know that $f_{t'}(k) = D_0$. We know that there is a $t \in bs[q]$ such that $s_t = s_{t'}$ (since $so_q = so_{q'}$), $f_t(k) = N$ and $(t,inv(deq,k),t') \in \delta_A$.
\item[$\langle ii-d \rangle$] Let $(q,lin(deq,d,k),q') \in \delta_I$ and $t' \in bs[q']$. First consider the case $d \neq \texttt{EMPTY}$. We have two cases: Either the node with operation id $k$ in the $so_q$ was \texttt{PENDING} or \texttt{CLOSED}. For both of the cases, we can find $t \in bs[q]$ such that $s_t = \langle d \rangle \circ s_{t'}$ and $f_t(k) = D_0$. Hence $(t,lin(deq,d,k),t') \in \delta_I$. Next, consider the case $d = \texttt{EMPTY}$. Then, we know that $s_{t'} = \langle \rangle$ and there exists $t \in bs[q]$ such that $s_t = \langle \rangle$ and $f_t = D_0$. Hence $(t,lin(deq,d,k),t') \in \delta_A$.
\item[$\langle ii-b-enq \rangle$] Let $(q,ret(enq,k),q') \in \delta_I$ and $t' \in bs[q']$. There are two possible cases: There exists a node $(k,d,\texttt{PENDING}) \in nd_q$ or for all nodes $n = (k,\_,\_)$, $n \notin nd_q$. For the former case, $s_{t'} = \rho \circ \langle d \rangle \circ \pi$ where $\rho$ is linearization of closed nodes in $nd_{q'}$ and $\pi$ is linearization of some open nodes in $node_{q'}$. We also know that $f_{t'}(k) = E_2$. Then, there exists a node $t \in bs[q]$ such that $s_t = s_{t'}$ (since nodes that generate $\rho$ are closed in $q$ and that generate $\pi$ are open in $q$) and $f_t(k) = E_1$. Hence, $(t,ret(deq,k),t') \in \delta_A$. For the latter case, we know that $so_q = so_{q'}$. Therefore, there exists $t \in bs[q]$ such that $s_t = s_{t'}$. Moreover $f_t(k) = E_1$. Hence, $(t,ret(enq,k),t') \in \delta_A$.
\item[$\langle ii-b-deq \rangle$] Let $(q,ret(deq,d,k), q') \in \delta_I$ and $t' \in bs[q']$. Then, we know that $f_{t'}(k) = D_2$ and there exists a $t \in bs[q]$ such that $s_t = s_{t'}$ (since $so_q = so_{q'})$ and $f_t(k) = D_1$. Then, $(t,ret(deq,d,k) \in \delta_A$.
\end{itemize}
\end{proof}

\begin{lem}
$L_A$ is a q-refinement of $L_I$.
\end{lem}
\begin{proof}
Since $L_I$ is deterministic with respect to the alphabet $AQ\Sigma$, we should be able to find a forward simulation between $L_A$ and $L_I$ if our claim is correct. We propose a forward simulation by adding some auxiliary varibles to the state of $L_A$. In the new augmented state of $L_A$, the sequence does not only keep the elements of the queue but also the operation identifiers that enqueued them. Hence, the sequence consists of  pairs of the form $s_q(i) = (k,d)$ where $q \in Q_A$, $i \in \mathbb{N}$ is an index, $d \in \mathcal{D}\backslash \{ \texttt{EMPTY} \}$ is a value and $k \in \mathbb{N}$ is an operation identifier. We also add a set $InvEnq_q$ component to state that keeps the enqueue operations at the point $E_0$ i.e. enqueues that are invoked bu have not been linearized yet. Elements of $InvEnq_q$ are pairs of the form $(d,k)$ where  $d \in \mathcal{D}\backslash \{ \texttt{EMPTY} \}$ is a value and $k \in \mathbb{N}$ is an operation identifier. Then our forward simulation $fs \subseteq Q_A \times Q_I$ relates the state $q = (s_q, f_q) \in Q_A$ to a state $q' =(so_{q'}, f_{q'}) \in Q_I$ iff (i) $f_{q'}(k) = f_q(k)$ for all $f_q(k) \in \{N, D_0, D_1, D_2\}$, (ii) $f_q(k) = E_0$ or $f_q(k) = E_1$ implies $f_{q'}(k) = E_0$, (iii) $f_q(k) = E_2$ implies $f_{q'}(k) = E_1$, (iv) If $(d,k) \in InvEnq_q$ or there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ and $f_q(k) = E_1$, then $(k,d,\texttt{PENDING})$ is a node in $so_{q'}$, (v) if there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ and $f_q(k) = E_2$ then $(k,d,\texttt{CLOSED})$ is a node in $so_{q'}$, (vi) open nodes in $so_{q'}$ are maximal and (vii) there exists a linearization (total order) $lo_{q'}$ of $so_{q'}$ that may omit some open nodes of $so_{q'}$ such that if we project nodes of
$lo_{q'}$ to the first two fields, this linear order is equal to the sequence $s_q$.

Next, we will show that $fs$ is a forward simulation relation:
\begin{itemize}
\item[$\langle i \rangle$] $fs[(q_0)_A] = \{(q_0)_I\}$
\item[$\langle ii-a-enq \rangle$] Let $(q,inv(enq,d,k),q') \in \delta_A$ and $t \in fs[q]$. We can obtain $t'$ such that $(t,inv(enq,d,k),t') \in \delta_I$. We show that $t' \in fs[q']$ by checking seven conditions of our $fs$ relation. We omit the node $(k,d)$ while linearizing $so_{t'}$ and obtain $s_{q'}$.
\item[$\langle ii-a-deq \rangle$] Let $(q, inv(deq,k), q') \in \delta_A$ and $t \in fs[q]$. We obtain $t'$ such that $(t, inv(deq,k),t') \in \delta_I$. The only difference between $t'$ and $t$ is that $f_{t'}(k) = D_0$. We can again see that $t' \in fs[q']$. 
\item[$\langle ii-c \rangle$] Let $(q,lin(enq,d,k),q') \in \delta_A$ and $t \in fs[q]$. Pick $t' = t$. We can see that $t \in fs[q']$. While obtaining linearization of $so_{t'}$, we do not neglect the node $(k,d, \texttt{PENDING})$ this time, although we neglect it in the linearization of $so_t$.
\item[$\langle ii-d \rangle$] Let $(q, lin(deq,d,k), q') \in \delta_A$ and $t \in fs[q]$. Obtain $t'$ such that $(t, lin(deq,d,k), t') \in \delta_I$. In $s_q$, $(k,d)$ must be the minimum element. Hence, $(k,d,\_)$ is a minimal node in $so_t$ and $lin(deq,d,k)$ is an enabled action in $L_I$. This action removes the node $(k,d,\_)$ from $so_t$ and changes $f_t(k)=D_0$ to $f_{t'}(k)= D_1$. We can see that $t' \in fs[q']$ by checking the seven conditions.
\item[$\langle ii-b-enq \rangle$] Let $(q, ret(enq,k), q') \in \delta_A$ and $t \in fs[q]$. There are two cases: Either there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ or there is no such $i$. For the former case, there is no node of the form $(k,d,\_)$ in $so_t$. We obtain $t'$ such that $(t,ret(enq,k),t') \in \delta_I$. Then, there is no node of the form $(k,d,\_)$ in $so_{t'}$ neither and $so_{t'} = so_t$. since $s_q = s_{q'}$ also holds for this case, $t' \in fs[q']$ holds. For the latter case, we again pick $t'$ such that $(t, ret(enq,k), t') \in \delta_I$. Again, $so_t = so_{t'}$ and $s_q = s_{q'}$ holds and $t' \in fs[q']$ can be observed.
\item[$\langle ii-b-deq \rangle$] Let $(q, ret(deq,d,k), q') \in \delta_A$ and $t \in fs[q]$. We pick $t'$ such that $(t, ret(deq,d,k), t') \in \delta_I$. We know that $so_t = so_{t'}$ and only change is $f_{t'}(k) = D_2$. Since we know that $s_q = s_{q'}$ and $f_{q'}(k) = D_2$, $t' \in fs[q']$ holds.  
\end{itemize}
\end{proof}

### 2017-01-11 draftCAV/set.tex
started draft for CAV

\section{Existence of Forward Simulations for Set Implementations That Have Fixed Remove Linearization Points }
For all the set libraries we fix $\mathcal{M} = \{ add, rmv, cnt\}$ where $rmv$ is short for remove and $cnt$ is short for contains methods and $\mathcal{D} = \{1, \texttt{TRUE}, \texttt{FALSE} \}$. We assume that only single element can be inserted into our list. Our results for this domain extends to other domains such as when $\mathcal{D} = \mathbb{N}$. We extend the definition of $A\Sigma$ introduced in Definition 2 for the set in our focus as $AS\Sigma = A\Sigma \cup \{lin(rmv,d,k)| d \in \{1\}, k \in \mathbb{N}\}$. We define s-refinement, s-linearizability and change the definitions of forward and backward simulations as in the beginning of Section 2.

We define $L_A$ as follows:
\begin{itemize}
\item $Q_A = 2^{\{1\}} \times (\mathbb{N} \rightarrow Lbl_A)$ where $Lbl_A = \{N, A_0, A_{1T}, A_{1F} A_2, R_0, R_{1T}, R_{1F}, R_2, C_0, C_{1T}, C_{1F}, C_2 \}$. For each $q \in Q_A$, $s_q$ represents the set component (first component), and $f_q$ represents the program counter (second component).
\item Transition labels consists of invocations, returns and linearizations of methods in $\mathcal{M}$: $\Sigma_A: AS\Sigma \cup \{lin(m,d,k)| m \in \{add,cnt\}, d\in \{1\}, k \in \mathbb{N} \}$. All methods return \texttt{TRUE} or \texttt{FALSE} and all take an element in $\{1\}$ as the input value.
\item ${q_0}_A = (\emptyset, f_{{q_0}_A}$ where $f_{{q_0}_A}(k) = N$ for all $k \in \mathbb{N}$.
\item State transitions:
\begin{itemize}
\item $(q, inv(add, d,k), q') \in \delta_A$ iff $d \in \{1\} \wedge f_q(k) = N \wedge f_{q'}(k) = A_0$
\item $(q, lin(add,d,k), q') \in \delta_A$ iff $d \in \{1\} \wedge f_q(k) = A_0 \wedge (d \in s_q \wedge s_q = s_{q'} \wedge f_{q'}(k) = A_{1F} \vee d \notin s_q \wedge s_{q'} = s_q \cup \{d\} \wedge f_{q'}(k) = A_{1T})$
\item $(q, ret(add,d,k) q') \in delta_A)$ iff $(f_q(k) = A_{1T} \wedge d = \texttt{TRUE} \vee f_q(k) = A_{1F} \wedge d = \texttt{FALSE}) \wedge f_{q'}(k) = A_2$
\item $(q, inv(rmv,d,k) q') \in \delta_A$ iff $d \in \{1\} \wedge f_q(k) = N \wedge f_{q'}(k)= R_0 $
\item $(q, lin(rmv,d,k), q') \in \delta_A$ iff $d \in \{1\} \wedge f_q(k) = R_0 \wedge (d \in s_q \wedge s_q = s_{q'} \cup \{d\} \wedge f_{q'}(k) = R_{1T} \vee d \notin s_q \wedge s_q = s_{q'} \wedge f_{q'}(k) = R_{1F} )$
\item $(q, ret(rmv,d,k) q') \in delta_A)$ iff $(f_q(k) = R_{1T} \wedge d = \texttt{TRUE} \vee f_q(k) = R_{1F} \wedge d = \texttt{FALSE}) \wedge f_{q'}(k) = R_2$
\item $(q, inv(cnt,d,k) q') \in \delta_A$ iff $d \in \{1\} \wedge f_q(k) = N \wedge f_{q'}(k)= C_0 $
\item $(q, lin(cnt,d,k), q') \in \delta_A$ iff $d \in \{1\} \wedge f_q(k) = C_0 \wedge s_q = s_{q'} \wedge (d \in s_q \wedge f_{q'}(k) = C_{1T} \vee d \notin s_q \wedge f_{q'}(k) = C_{1F} )$
\item $(q, ret(cnt,d,k) q') \in delta_A)$ iff $(f_q(k) = C_{1T} \wedge d = \texttt{TRUE} \vee f_q(k) = C_{1F} \wedge d = \texttt{FALSE}) \wedge f_{q'}(k) = C_2$
\end{itemize}
\end{itemize}

We define $L_I$ as follows:
\begin{itemize}
\item A state $q \in Q_I$ is a tuple of the form $(sac_q, src_q, UBSA_q, LBSA_q, CC_q, ubi,lbi, f_q)$ where $sac_q \in \mathbb{N}$ keeps the number of adds that return true so far, $src_q \in \mathbb{N}$ keeps the number of successful linearizations of remove (ones that move program counter from $R_0$ to $R_{1T}$), $f:\mathbb{N} \rightarrow Lbl_I$ is program counter that maps every operation to a label in $Lbl_I = \{N, A_0, A_1, R_0, R_{1T}, R_{1F}, R_2, C_0, C_1\}$. Let $PA_q = \{k \in \mathbb{N}| f_q(k) = A_0\}$ be the set of pending adds at state $q$. Then, $UBSA_q, LBSA_q: \mathbb{N} \rightarrow 2^{PE_q}$ are functions such that $UBSA_q(i)$ is a set of pending adds at most $i$ of which may return true and $LBSA_q(i)$ is a set of pending adds at least $i$ of which may return true. $ubi, lbi \in \mathbb{N}$ keeps the upper bound (lower bound) set indices that a new pending enqueue will be inserted to. Let $IC_q = \{k \in \mathbb{N}| f_q(k) = C_0 \vee f_q(k) = C_1$  be the set of contains operation identifiers. Then, $CC_q: IC_q \rightarrow 2^{PA}$ is a map that keeps a set of pending adds of which at least one should return true due to a true return of a contains operation.
\item Transition labels are exactly the abstract transition labels we have defined earlier: $\Sigma_I = AS\Sigma$
\item ${q_0}_I =(sac_{{q_0}_I}, src_{{q_0}_I}, UBSA_{{q_0}_I}, LBSA_{{q_0}_I}, CC_{{q_0}_I}, ubi_{{q_0}_I}, lbi_{{q_0}_I}, f_{{q_0}_I})$ where $sac_{{q_0}_I} = src_{{q_0}_I} =  0$, $lbi_{{q_0}_I} = ubi_{{q_0}_I} = 1$ and $f_{{q_0}_I}(k) = N$ for all $k \in \mathbb{N}$. Hence $PA_{{q_0}_I} = IC_{{q_0}_I} = \emptyset$ and $UBSA_{{q_0}_I}$, $LBSA_{{q_0}_I}$ and $CC_{{q_0}_I}$ are empty mappings.
\item Instead of giving state transition relation formally, I would like to explain how $UBSA$, $LBSA$ and $CC$ works by considering the possible state transformations (consider $q$ as pre-state and $q'$ as the post state as convention for the following):
\begin{itemize}
\item[$UBSA$] Initially, $ubi_{{q_0}_I} = 1$. Hence, a newly invoked add operation $a$ will be inserted into $UBSA_{{q_0}_I}(1)$. All the newly invoked operations are inserted to $UBSA_q(1)$ until a linearization of remove comes or one of the adds return successful. If the first case happens, we increment the index of every set by 1 i.e. $UBSA_{q'}(i) = UBSA_q(i-1)$ for all $i>0$ and $UBSA_{q'}(0) = \emptyset$. We also set $ubi_{q'} = 1$, if it was set to $0$ somehow before (Consider the trace $inv(add,a,1), ret(add,true,1), inv(add,a,2), lin(rmv,a,3)$. $ubi =0$ after second event and it needs to be incremented to $1$ after the fourth event). Next, consider the case of successful add. Let the operation identifier of the add that will return successful be $k$ and $i$ be the index such that $k \in UBSA_q(i)$. Then, for all $j > i$, $UBSA_{q'} (j-1) = UBSA_q(j)$, $UBSA_{q'}(i-1) = UBSA_q(i-1) \cup UBSA_q(i) \backslash \{ k\}$ and for all $j < i-1$, $UBSA_{q'}(j) = UBSA_q(j)$. We do not allow elements in $UBSA_q(0)$ to return true. So, $i=0$ cannot be true. If $i = 1$, we change $ubi_{q'} = 0$ to denote that no new coming add can return true from now on (consider the case: $inv(add,a,1), lin(rmv,a,2):true, inv(add,a,3), inv(add,a,4), lin(rmv,a,5):true, ret(add,true,3), ret(add,true,4), inv(add,a,6)$. This last invocation should be added to set with index $0$. So, we need to change $ubi$ to $0$ after last return. Lastly, a failing return of add simply removes this add from its set, without changing its index. Let $i$ be the index of the failing add operation $k$. Then, $UBSA_{q'}(i) = UBSA_q(i) \backslash \{k\}$. Observing a failing contains operation may change the upper bound. Consider the following trace: $inv(add,a,1), inv(add,a,2), inv(cnt,a,3), ret(cnt,false,3)$. Although $UBSA_{q'}(1) = \{1,2\}$ after the last event, none of $1$ or $2$ can return true due to the restrictions we impose on contains operations that we will explain later on.
\item[$LBSA$] Initially $lbi_{{q_0}_I} = 1$. When a new event of type invoke add comes, we modify $LBSA_{q'}(lbi_{q'}) = PE_{q'}$ where $lbi_{q'} = lbi_q$. Although it looks like that we build $LBSA_{q'}$ from scratch with every new add invocation, in practice, this is merely adding new pending add to the old set of the same index unless the $lbi$ field are the same. If a new successful linearization of remove comes, first it updates $LBSA_{q'}(lbi_q) = PE_{q'}$. Then, it increments the $lbi$ value by one if it was not $0$ before. If it was $0$ before, new $lbi$ value becomes $src_{q'} - sac_{q'}+1$. Note that a successful remove does not modify $LBSA$ field if a new add is invoked after the previous linearization of a successful remove. The first new add invocation that comes after a successful remove, copies the set $LBSA_{q}(lbi_q)$ and adds the new add's operation identifier as $LBSA_{q'}(lbi_{q'})$. However, if no new add comes between two successful remove linearizations, then $LBSA$ is modified by a sucsessful remove linearization. Next, consider the return events of add operations. First, consider the successful return. If an add operation with identifier $k$ returns true, we remove this identifier from the $LBSA$ sets of which $k$ is element of and decrement the index values of these sets by $1$. We can observe that if $k$ is the element of the set with index $i$ then it is element of every set with index $j>i$ \textcolor{red}{(We need to prove this)}. Hence we decrement the index of every set bigger than some $i$ value. For this case, we may end up with a situation that two sets fall into same index (at index $i-1$). In this case, one of the sets must be strictly subset of the other one \textcolor{red}{(We need to prove this)}. In this case, we keep the subset as the set of this index and delete the larger set. If $k$ is deleted from just no sets, we change $lbi_{q'} = 0$. To rationalize this behavior, consider the trace $inv(add,a,1), lin(rmv,a,2):true, inv(add,a,3), inv(add,a,4), lin(rmv,a,5):true, ret(add,true,3), ret(add,true,4), inv(add,a,6)$. The last add invocation should be put to the set with index $0$ and after the add with identifier $4$ returns true, our procedure makes $lbi_{q'} = 0$. If the removed element is in $LBSA_q(lbi_q) \backslash LBSA_q(lbi_q -1)$, then we also set $lbi_{q'} = 0$. The rationale behind this is the trace: $inv(add,a,1), lin(rmv,a,2):true, inv(add,a,3), ret(add,true,3), inv(add,a,4)$. The last operation with ID $4$ should be able to return false. If an add operation with identifier $k$ returns false, then we remove $k$ from all $LBSA$ sets that $k$ is element of. The constraint we check on $LBSA$ sets is that a return false event of an add operation keeps $|LBSA_{q'}(i)| \geq i$ for the indices $i$ it modified. 
\item [$CC$] We need to keep a cc flag to show that now this operation may return true. It should be a map from identifiers to boolean.
\end{itemize}
\end{itemize}

### 2017-01-11 draftCAV/app-backSim.tex
in principle, finished the preliminaries

@@
 \begin{dfn}
-Let $L_1=(Q_1,\Sigma, q_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, q_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$.
+Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
 A relation $bs \subseteq Q_1 \times Q_2$ is called a \emph{backward simulation} from $L_1$ to $L_2$ iff the following holds:
 \begin{itemize}
-\item[(i)] $bs[q_0^1] = \{q_0^2 \}$
+\item[(i)] $bs[s_0^1] = \{s_0^2 \}$
 \item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0<i<|@s|$.
@@
 %where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
-\item[(ii-c)] If $(s,e,s') \in \delta_1$ for some $e \in \Sigma\setminus (C\cup R)$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
+\item[(ii-c)] If $(s,e,s') \in \delta_1$ for some $e \in \Sigma\setminus \Gamma$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
 %where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
+\item[(ii-d)] If $(s,\gamma, s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u\xrightarrow{\gamma} u' \in \delta_2$
 \end{itemize}

### 2017-01-11 draftCAV/prelim.tex
in principle, finished the preliminaries

@@
 
-A \emph{labeled transition system} (LTS) $A=(Q,\Sigma, q_0, \delta)$ over the 
+A \emph{labeled transition system} (LTS) $A=(Q,\Sigma, s_0, \delta)$ over the 
 possibly-infinite alphabet $\Sigma$ is a possibly-infinite set $Q$ of states with
-initial state $q_0 \in Q$, and a transition relation $\delta \subseteq Q \times @S \times
+initial state $s_0 \in Q$, and a transition relation $\delta \subseteq Q \times @S \times
 Q$. The $i$th symbol of a sequence $@t \in @S^*$ is denoted $@t_i$, and the empty
@@
 An \emph{execution} of $A$ is an alternating sequence of states and transition labels (called also actions)
-$\rho = s_0, e_0,s_1\ldots e_{k-1},s_k$ for some $k>0$ such that $s_0 = q_0$ and $\delta(s_i, e_i, s_{i+1})$
+$\rho = s_0, e_0,s_1\ldots e_{k-1},s_k$ for some $k>0$ such that $\delta(s_i, e_i, s_{i+1})$
 for each $i$ such that $0\leq i<k$. We write $s_i\xrightarrow{e_i\ldots e_{j-1}} s_j$ for any $0\leq i\leq j <k$
@@
 The set of executions, resp., traces, of an LTS $A$ is denoted by $E(A)$, resp., $Tr(A)$.
+An LTS is \emph{deterministic} if for any state $s$ and any sequence $@t\in \Sigma^*$, there is at most
+one state $s'$ such that $s\xrightarrow{@t}s'$.
 
@@
 say $L_1$ refines $L_2$ if{f} every computation of every program
-using $L_1$ would also be possible were $L_2$ used instead. As shown in~\citet{DBLP:conf/popl/BouajjaniEEH15}
-this can be equivalently stated as history set inclusion between the two libraries.
+using $L_1$ would also be possible were $L_2$ used instead. As shown in~\citet{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15},
+refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} when $L_2$ is an atomic library. 
+Refinement can be proved using \emph{forward/backward simulation relations} which roughly, guarantee that steps
+in the concrete library correspond to (sequences of) steps in the abstract library.
 
 \begin{dfn}
-Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$~\footnote{The two libraries could be defined over different alphabets as long as they share the same set of call/return actions.}. 
+Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$~\footnote{The two libraries could be defined over different alphabets as long as they share the set $\Gamma$.}.
 %agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets.
@@
 Note that $A\Sigma \subseteq \Sigma_{L_1}$ and $A\Sigma \subseteq \Sigma_{L_2}$.
-The library $L_1$ \emph{refines} $L_2$ if{f} $H(L_1) \subseteq H(L_2)$.
+The library $L_1$ \emph{$\Gamma$-refines} $L_2$ if{f} $Tr(L_1) | \Gamma \subseteq Tr(L_2) | \Gamma$.
 %We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
@@
 \end{dfn}
+The actions in $\Gamma$ are called \emph{observable actions}.
+Usually, $\Gamma$ consists only of call and return actions.
+As we show in the next sections, increasing the set of observable actions is useful in simplifying proofs of refinement.
 %Linearizability is also a relation between two libraries and it is stricter than refinement.
@@
 
-The linearizability criterion~\cite{journals/toplas/HerlihyW90} provides an
-alternative characterization of library conformance which implies 
-refinement~\cite{journals/tcs/FilipovicORY10}. 
-The two criteria coincide when the reference implementation is atomic~\cite{DBLP:conf/popl/BouajjaniEEH15}.
-Linearizability~\cite{journals/toplas/HerlihyW90} is defined by the weakening order $\sqsubseteq$. A 
-history $h_1$ is \emph{linearizable} w.r.t.~a library $L_2$ if{f} there exists a sequential history
+Linearizability~\cite{journals/toplas/HerlihyW90} requires that 
+the operations of a concurrent library seem to take effect instantaneously at some point in time just like atomic
+implementations even if they use fine-grained concurrency and they overlap in time. The point in time
+where an operation is conceptually effectuated is called a \emph{linearization point}. There are two ways of
+defining linearizability, using the weakening order $\sqsubseteq$ between histories or using linearization points.
+
+%For some libraries,
+%the linearization points of (some of) the operations are fixed to particular implementation actions and they
+%can be exposed
+
+A history $h_1$ is \emph{linearizable} w.r.t.~a library $L_2$ if{f} there exists a sequential history
 $h_2 \in H(L_2)$, with only
@@
 
-\begin{theorem}[\cite{DBLP:conf/popl/BouajjaniEEH15}]
-  $L_1 \sqsubseteq L_2$ if{f} $L_1$ refines $L_2$,
-  if $L_2$ is atomic.
-
-\end{theorem}
-
-The weakening order w.r.t. sequential histories can be also defined  in terms of ~\emph{linearization points}. Informally, a linearization point 
-of an operation in an execution is a point in time where the operation is conceptually effectuated. It corresponds to a transition in that
-execution that occurs between the call and the return action of the operation.
-A \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that
+The weakening order w.r.t. sequential histories can be also defined  in terms of ~\emph{linearization points}. 
+%A linearization point 
+%corresponds to a transition in that
+%execution that occurs between the call and the return action of the operation.
+A \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that:
 \begin{itemize}
@@
 An LP-annotation is called \emph{complete} when every completed operation has a linearization point, i.e., TODO.
-An operation $k$ is called \emph{effectuated} in a complete LP-annotation $\hat{h}$ iff  $\hat{h}$ contains $lin(\_,\_,k)$.
+An operation $k$ is called \emph{effectuated} in a complete LP-annotation $\hat{h}$ when  $\hat{h}$ contains $lin(\_,\_,k)$.
 
-It has been shown that $h_1 \sqsubseteq h_2$ for a sequential history $h_2$ if{f} there exists a complete LP-annotation $\hat{h_1}$ such that  
+For a history $h_1$ and a sequential history $h_2$, $h_1 \sqsubseteq h_2$ if{f} there exists a complete LP-annotation $\hat{h_1}$ such that  
 $h_2$ is a permutation of the history $h_1'$ obtained from $h_1$ by preserving only call/return actions of effectuated operations that preserves the linearization point order, i.e.,~a given return action $ret(\_,\_,k)$ occurs before a given
@@
 
-Refinement relations can be approximated using simulation relations, i.e., the existence of a simulation relation implies refinement.
-We define a notion of \emph{forward} simulation whose existence is equivalent to refinement only when the right-hand side library $L_2$ is deterministic.
+\begin{theorem}[\cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}]
+  $L_1 \sqsubseteq L_2$ if{f} $L_1$ refines $L_2$,
+  if $L_2$ is atomic.
+
+\end{theorem}
+
+We define a notion of \emph{forward} simulation whose existence implies refinement, and it is equivalent to refinement when the right-hand side library $L_2$ is deterministic.
 A dual notion of \emph{backward} simulation is defined in Appendix~\ref{app:backSim}.
@@
 \begin{dfn}
-Let $L_1=(Q_1,\Sigma, q_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, q_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$.
-A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{forward simulation} from $L_1$ to $L_2$ iff the following holds:
+Let $L_1=(Q_1,\Sigma, q_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, q_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
+A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
 \begin{itemize}
-\item[(i)] $\mathit{fs}[q_0^1] = \{q_0^2 \}$
+\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$
 \item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0<i<|@s|$.
@@
 % where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
-\item[(ii-c)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus (C\cup R)$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
+\item[(ii-c)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
 %where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
+\item[(ii-d)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u\xrightarrow{\gamma} u'  \in \delta_2$
 \end{itemize}
 \end{dfn}
-If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function.
+%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
+A $\Gamma$-forward simulation establishes that $L_2$ can mimic every step of $L_1$ using a (possibly empty) sequence of steps.
+In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action.
+The rest of the transitions in $L_1$ must be simulated by a possibly empty sequence of transitions of $L_2$ (with arbitrary labels).
 
-Simulation relations are used to prove refinement among libraries. The following lemma shows their soundness:
+Simulation relations are used to prove refinement among libraries. The following lemma shows their soundness and their completeness (when $L_2$ is deterministic).
 \begin{lem}
-Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$.
-If there exists a forward simulation from $L_1$ to $L_2$, then $L_1$ refines $L_2$.
-Moreover, if $L_2$ is deterministic and $L_1$ refines $L_2$, then there exists a forward simulation from $L_1$ to $L_2$.
+Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
+If there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$.
+Moreover, if $L_2$ is deterministic and $L_1$ $\Gamma$-refines $L_2$, then there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$.
 \end{lem}

### 2017-01-11 draftCAV/app-backSim.tex
in principle, finished the preliminaries

@@
 \begin{lem}
-Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$. $L_1$ refines $L_2$ if{f} there exists a backward simulation from $L_1$ to $L_2$.
+Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
+$L_1$ $\Gamma$-refines $L_2$ if{f} there exists a $\Gamma$-backward simulation from $L_1$ to $L_2$.
 \end{lem}

### 2017-01-11 draftCAV/app-backSim.tex
in principle, finished the preliminaries

@@
 %where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
-\item[(ii-c)] If $(s,e,s') \in \delta_1$ for some $e \in \Sigma\setminus \Gamma$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
+\item[(ii-c)] If $(s,\gamma, s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u\xrightarrow{\gamma} u' \in \delta_2$
+\item[(ii-d)] If $(s,e,s') \in \delta_1$ for some $e \in \Sigma\setminus \Gamma$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
 %where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
-\item[(ii-d)] If $(s,\gamma, s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u\xrightarrow{\gamma} u' \in \delta_2$
 \end{itemize}

### 2017-01-11 draftCAV/prelim.tex
in principle, finished the preliminaries

@@
 \begin{dfn}
-Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$~\footnote{The two libraries could be defined over different alphabets as long as they share the set $\Gamma$.}.
+Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$~\footnote{The two libraries could be defined over different alphabets as long as they share $\Gamma$.}.
 %agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets.
@@
 % where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
-\item[(ii-c)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
+\item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u\xrightarrow{\gamma} u'  \in \delta_2$.
+\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
 %where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
-\item[(ii-d)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u\xrightarrow{\gamma} u'  \in \delta_2$
 \end{itemize}

### 2017-01-12 draftCAV/prelim.tex
Comments on Preliminaries

@@
 $\rho = s_0, e_0,s_1\ldots e_{k-1},s_k$ for some $k>0$ such that $\delta(s_i, e_i, s_{i+1})$
-for each $i$ such that $0\leq i<k$. We write $s_i\xrightarrow{e_i\ldots e_{j-1}} s_j$ for any $0\leq i\leq j <k$
-(in particular $s_i\xrightarrow{\epsilon}s_i$).
+for each $i$ such that $0\leq i<k$.
+We write $s_i\xrightarrow{e_i\ldots e_{j-1}} s_j$ \textcolor{red}{as shorthand for $s_i,e_i,...,s_{j-1},e_{j-1},s_j$ subsequence of $\rho$,} for any $0\leq i\leq j <k$
+(in particular $s_i\xrightarrow{\epsilon}s_i$)\textcolor{red}{(I also propose to add subscript to the arrow mentioning the machine name like $s_i \xrightarrow{\alpha}_A s_j$ to avoid confusion between machines in the forward simulation definition.
+This will require us to change arrows in forward and backward simulation definitions.)}.
 The projection $@t| \Gamma$ of a sequence $@t$ is the maximum subsequence of $@t$ over
@@
 \in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
-We may omit the second field from a call/return action $a$ for methods that have no inputs or return values 
+We may omit the second field from a call/return action $a$ for methods that have no inputs \textcolor{red}{(e.g., the pop method of a stack)}or return values \textcolor{red}{(e.g., the push method of a stack)}
 (e.g., the pop method of a stack).
@@
 
+\textcolor{red}{For the below paragraph, I cannot see the equivalence  of informal explanation and formal definition of weakening.
+I think the formal one should be if a ret comes before a call in $h_1'$ then this order is preserved in $h_2$.
+But the definition says iff.
+Ignore this if I am wrong.}
 Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
@@
 \end{quote}
+
 $\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
@@
 A direct consequence of Definition~\ref{def:libraries} is that $H(L)=\overline{H(L)}$ for
-every library $L$. A \emph{kernel} of L is any set $H$ such that $\overline{H} = L$.
+every library $L$.
+A \emph{kernel} of L is any set $H$ such that $\overline{H} = L$ \textcolor{red}{(I think the correct expression is $\overline{H} = H(L)$ which makes the example set $H$ and the operator $H$  that extracts history from a library confusing with each other)}.
 A library $L$ is called atomic if it has a sequential kernel. 
@@
 
+\textcolor{red}{A general organizational idea for  this subsection: Linearizability is defined in terms of weakening in which observational actions are just call and return events.
+Then, we put forward the idea of enriching $\Gamma$ by adding new actions.
+In the current organization, linearizability is defined after refinement before which this $\Gamma$ extension argument is explained.
+Hence, people may confuse that notion of linearizability is based on $\Gamma$ set we picked.
+I propose to first introduce linearizability.
+Then explain refinement and give Theorem 1.}
 Conformance of a library $L_1$ to a specification given as an ``abstract'' library $L_2$ 
@@
 \begin{dfn}
+\textcolor{red}{We assign same alphabet to both libraries in the definition but we say libraries can be defined over distinct alphabets as long as they share $\Gamma$ which is the case in our week machine.
+Why don't we define refinement like that at the first place?}
 Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$~\footnote{The two libraries could be defined over different alphabets as long as they share $\Gamma$.}.
@@
 The actions in $\Gamma$ are called \emph{observable actions}.
-Usually, $\Gamma$ consists only of call and return actions.
+Usually, $\Gamma$ consists only of call and return actions \textcolor{red}{(For this case we say just refinement instead of $\Gamma-refinement$.
+We need to add this sentence since we say just refinement in Theorem 1.)}.
 As we show in the next sections, increasing the set of observable actions is useful in simplifying proofs of refinement.
@@
 
+\textcolor{red}{An organizational concern: Do we really need the second definition of linearizability?
+I think our aim here is to introduce both refinement and linearizability and relate them via Thm 1.
+I cannot see any contribution of this part until Thm 1.
+You may ignore this message if I am missing something important about the second definition.}
 The weakening order w.r.t. sequential histories can be also defined  in terms of ~\emph{linearization points}. 
@@
 
-We define a notion of \emph{forward} simulation whose existence implies refinement, and it is equivalent to refinement when the right-hand side library $L_2$ is deterministic.
+We define a notion of \emph{forward} simulation whose existence implies refinement, and it is equivalent to refinement when the right-hand side library $L_2$ is deterministic \textcolor{red}{(do we need to define deterministic?)}.
 A dual notion of \emph{backward} simulation is defined in Appendix~\ref{app:backSim}.
@@
 \begin{itemize}
-\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$
-\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0<i<|@s|$.
+\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$ \textcolor{red}{Either we need to change $s_0^1$ to $q_0^1$ here or vice versa in the first line of the definition.}
+\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$ \textcolor{red}{($@s_i\in \Sigma\setminus \Gamma$)}, for each $0<i<|@s|$.
 %$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $.
 The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
-\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s| -1$.
+\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma\setminus(C\cup R)$  \textcolor{red}{($@s_i\in \Sigma\setminus \Gamma$)}, for each $0\leq i<|@s| -1$.
 % where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
 \item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u\xrightarrow{\gamma} u'  \in \delta_2$.
-\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
+\textcolor{red}{$(u,\gamma, u') \in \delta_2$ instead of $u\xrightarrow{\gamma} u'  \in \delta_2$?)}
+\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$  \textcolor{red}{($@s_i\in \Sigma\setminus \Gamma$)}, for each $0\leq i<|@s|$.
+\textcolor{red}{(All of the above changes also apply to the backward simulation definition in appendix.)}
 %where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.

### 2017-01-12 draftCAV/app-absImplQueue.tex
added refinement proofs and abstract queue implementation

--- 2017-01-11 draftCAV/queue.tex
@@
-\section{Existence of Forward Simulations for Queue Implementations that have Fixed Dequeue Linearization Points}
+%!TEX root = draft.tex
+\section{Proof of Theorem~\ref{th:absImplQueue}}\label{app:absImplQueue}
+
 In this section, we will show that for any concurrent queue implementation library $L_C$ for which we know the linearization points of the dequeue operation and that is linearizable with respect to the reference implementation library $L_A$, there exists a forward simulation $fs$ relating $L_C$ to $L_I$ where $L_I$ is an intermediate library equivalent to $L_A$ i.e.

### 2017-01-12 draftCAV/app-backSim.tex
added refinement proofs and abstract queue implementation

@@
 \item[(i)] $bs[s_0^1] = \{s_0^2 \}$
-\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0<i<|@s|$.
+\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus\Gamma$, for each $0<i<|@s|$.
 % where $a = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
-\item[(ii-b)] If $(s,r,s') \in \delta_1$, for some $r\in R$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s| -1$.
+\item[(ii-b)] If $(s,r,s') \in \delta_1$, for some $r\in R$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma\setminus\Gamma$, for each $0\leq i<|@s| -1$.
 %where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
-\item[(ii-c)] If $(s,\gamma, s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u\xrightarrow{\gamma} u' \in \delta_2$
-\item[(ii-d)] If $(s,e,s') \in \delta_1$ for some $e \in \Sigma\setminus \Gamma$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
+\item[(ii-c)] If $(s,\gamma, s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $\delta_2(u,\gamma,u')$
+\item[(ii-d)] If $(s,e,s') \in \delta_1$ for some $e \in \Sigma\setminus \Gamma$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$ and $@s_i\in \Sigma\setminus\Gamma$, for each $0\leq i<|@s|$.
 %where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.

### 2017-01-12 draftCAV/prelim.tex
added refinement proofs and abstract queue implementation

--- 2017-01-11 draftCAV/prelim.tex
@@
 $\rho = s_0, e_0,s_1\ldots e_{k-1},s_k$ for some $k>0$ such that $\delta(s_i, e_i, s_{i+1})$
-for each $i$ such that $0\leq i<k$. We write $s_i\xrightarrow{e_i\ldots e_{j-1}} s_j$ for any $0\leq i\leq j <k$
+for each $i$ such that $0\leq i<k$. We write $s_i\xrightarrow{e_i\ldots e_{j-1}}_A s_j$ as shorthand for 
+the subsequence $s_i,e_i,...,s_{j-1},e_{j-1},s_j$ of $\rho$, for any $0\leq i\leq j <k$
 (in particular $s_i\xrightarrow{\epsilon}s_i$).
@@
 \in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
-We may omit the second field from a call/return action $a$ for methods that have no inputs or return values 
-(e.g., the pop method of a stack).
+We may omit the second field from a call/return action $a$ for methods that have no inputs (e.g., the pop method of a stack) or return values (e.g., the push method of a stack).
 A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
@@
 
+%\textcolor{red}{For the below paragraph, I cannot see the equivalence  of informal explanation and formal definition of weakening.
+I think the formal one should be if a ret comes before a call in $h_1'$ then this order is preserved in $h_2$.
+But the definition says iff.
+Ignore this if I am wrong.}
 Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
@@
   $h_2$ is a permutation of $h_1'$ that preserves the order between
-  return and call actions, i.e.,~a given return action occurs before a given
-  call action in $h_1'$ if{f} the same holds in $h_2$.
+  return and call actions, i.e.,~if a given return action occurs before a given
+  call action in $h_1'$, then the same holds in $h_2$.
 
 \end{quote}
+A call action $inv(\_,\_,k)$ may happen before a return action $ret(\_,\_,k')$ in $h_1$ (the operations $k$ and $k'$ overlapping in time) while it can be permuted after that return action in $h_2$, thus introducing a new inter-operation ordering (operation $k'$ ending before $k$).
+
 $\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
@@
 A direct consequence of Definition~\ref{def:libraries} is that $H(L)=\overline{H(L)}$ for
-every library $L$. A \emph{kernel} of L is any set $H$ such that $\overline{H} = L$.
+every library $L$. A \emph{kernel} of L is any set $K$ such that $\overline{K} = H(L)$.
 A library $L$ is called atomic if it has a sequential kernel. 
@@
 
+%\textcolor{red}{A general organizational idea for  this subsection: Linearizability is defined in terms of weakening in which observational actions are just call and return events.
+Then, we put forward the idea of enriching $\Gamma$ by adding new actions.
+In the current organization, linearizability is defined after refinement before which this $\Gamma$ extension argument is explained.
+Hence, people may confuse that notion of linearizability is based on $\Gamma$ set we picked.
+I propose to first introduce linearizability.
+Then explain refinement and give Theorem 1.}
 Conformance of a library $L_1$ to a specification given as an ``abstract'' library $L_2$ 
@@
 say $L_1$ refines $L_2$ if{f} every computation of every program
-using $L_1$ would also be possible were $L_2$ used instead. As shown in~\citet{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15},
+using $L_1$ would also be possible were $L_2$ used instead. We assume that a program can 
+interact with the library only through call and return actions, and thus refinement can be defined
+as history set inclusion. As shown in~\citet{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15},
 refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} when $L_2$ is an atomic library. 
-Refinement can be proved using \emph{forward/backward simulation relations} which roughly, guarantee that steps
-in the concrete library correspond to (sequences of) steps in the abstract library.
+We fix a set of methods $\<Methods>$ and values $\<Vals>$.
 
 \begin{dfn}
-Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$~\footnote{The two libraries could be defined over different alphabets as long as they share $\Gamma$.}.
+Let $L_1$ and $L_2$ be two libraries. 
 %agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets.
@@
 Note that $A\Sigma \subseteq \Sigma_{L_1}$ and $A\Sigma \subseteq \Sigma_{L_2}$.
-The library $L_1$ \emph{$\Gamma$-refines} $L_2$ if{f} $Tr(L_1) | \Gamma \subseteq Tr(L_2) | \Gamma$.
+The library $L_1$ \emph{refines} $L_2$ if{f} $H(L_1) \subseteq H(L_2)$.
 %We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
@@
 \end{dfn}
-The actions in $\Gamma$ are called \emph{observable actions}.
-Usually, $\Gamma$ consists only of call and return actions.
-As we show in the next sections, increasing the set of observable actions is useful in simplifying proofs of refinement.
+
 %Linearizability is also a relation between two libraries and it is stricter than refinement.
@@
 
-Linearizability~\cite{journals/toplas/HerlihyW90} requires that 
-the operations of a concurrent library seem to take effect instantaneously at some point in time just like atomic
-implementations even if they use fine-grained concurrency and they overlap in time. The point in time
-where an operation is conceptually effectuated is called a \emph{linearization point}. There are two ways of
-defining linearizability, using the weakening order $\sqsubseteq$ between histories or using linearization points.
-
-%For some libraries,
-%the linearization points of (some of) the operations are fixed to particular implementation actions and they
-%can be exposed
-
-A history $h_1$ is \emph{linearizable} w.r.t.~a library $L_2$ if{f} there exists a sequential history
+Linearizability~\cite{journals/toplas/HerlihyW90} requires that every history of a concurrent library $L_1$ can be 
+``linearized'' to a sequential history admitted by a library $L_2$ used as a specification. 
+Formally, a history $h_1$ is \emph{linearizable} w.r.t.~a library $L_2$ if{f} there exists a sequential history
 $h_2 \in H(L_2)$, with only
@@
 each history $h_1 \in H(L_1)$ is linearizable w.r.t. $L_2$.
-
-The weakening order w.r.t. sequential histories can be also defined  in terms of ~\emph{linearization points}. 
-%A linearization point 
-%corresponds to a transition in that
-%execution that occurs between the call and the return action of the operation.
-A \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that:
-\begin{itemize}
-	\item every linearization point corresponds to an operation that has been called, i.e., TODO
-	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
-	\item the return value of an operation is fixed by the linearization point, i.e., TODO
-\end{itemize}
-An LP-annotation is called \emph{complete} when every completed operation has a linearization point, i.e., TODO.
-An operation $k$ is called \emph{effectuated} in a complete LP-annotation $\hat{h}$ when  $\hat{h}$ contains $lin(\_,\_,k)$.
-
-For a history $h_1$ and a sequential history $h_2$, $h_1 \sqsubseteq h_2$ if{f} there exists a complete LP-annotation $\hat{h_1}$ such that  
-$h_2$ is a permutation of the history $h_1'$ obtained from $h_1$ by preserving only call/return actions of effectuated operations that preserves the linearization point order, i.e.,~a given return action $ret(\_,\_,k)$ occurs before a given
-  call action $call(\_,\_,k')$ in $h_2$ if{f} $lin(\_,\_,k)$ occurs before $lin(\_,\_,k')$ in $\hat{h_1}$.
 
@@
 
-We define a notion of \emph{forward} simulation whose existence implies refinement, and it is equivalent to refinement when the right-hand side library $L_2$ is deterministic.
-A dual notion of \emph{backward} simulation is defined in Appendix~\ref{app:backSim}.
-The existence of backward simulations is equivalent to refinement for arbitrary libraries.
-%We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement. 
-For a relation $R\subseteq A\times B$, $R[X]$ denotes the set of elements of $B$ related to elements of $X$ by $R$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.
-
-%We can extend the relations between libraries by introducing simulation relations. We will later show that simulation relations imply refinement.
-
-\begin{dfn}
-Let $L_1=(Q_1,\Sigma, q_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, q_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
-A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
-\begin{itemize}
-\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$
-\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0<i<|@s|$.
-%$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $.
-The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
-\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s| -1$.
-% where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
-\item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u\xrightarrow{\gamma} u'  \in \delta_2$.
-\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
-%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
-\end{itemize}
-\end{dfn}
-%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
-A $\Gamma$-forward simulation establishes that $L_2$ can mimic every step of $L_1$ using a (possibly empty) sequence of steps.
-In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action.
-The rest of the transitions in $L_1$ must be simulated by a possibly empty sequence of transitions of $L_2$ (with arbitrary labels).
-
-Simulation relations are used to prove refinement among libraries. The following lemma shows their soundness and their completeness (when $L_2$ is deterministic).
-\begin{lem}
-Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
-If there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$.
-Moreover, if $L_2$ is deterministic and $L_1$ $\Gamma$-refines $L_2$, then there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$.
-\end{lem}
-\begin{proof}
-Looks trivial and follows Lynch paper. Can be completed later.
-\end{proof}
+TODO MAKE THE CONNECTION WITH THE NEXT SECTION

### 2017-01-12 draftCAV/queue.tex
added refinement proofs and abstract queue implementation

@@
-\section{Existence of Forward Simulations for Queue Implementations that have Fixed Dequeue Linearization Points}
-In this section, we will show that for any concurrent queue implementation library $L_C$ for which we know the linearization points of the dequeue operation and that is linearizable with respect to the reference implementation library $L_A$, there exists a forward simulation $fs$ relating $L_C$ to $L_I$ where $L_I$ is an intermediate library equivalent to $L_A$ i.e.
-$L_I$ refines $L_A$ and vice versa.
+%!TEX root = draft.tex
+\section{Queues With Fixed Dequeue Linearization Points}
 
-For all the queue libraries, we fix $\mathcal{M} = \{ enq, deq \}$ and $\mathcal{D} = \mathbb{N} \cup \{\texttt{EMPTY} \}$.
-Since we know the linearization point of dequeues, we extend the definitions of the previous sections adding this information.
-First, we extend the set $A\Sigma$ introduced in Definition 2 for queues in our focus as $AQ\Sigma = A\Sigma \cup \{lin(deq,d,k)| d \in \mathcal{D}, k \in \mathbb{N}\}$.
-For any library $L$ we consider in this section, $AQ\Sigma \subseteq \Sigma_L$.
-Then, we define q-refinement by replacing $A\Sigma$ with $AQ\Sigma$ in Definition 2.
-We also define q-linearizability, by enforcing $lin(deq,d,k)$ to appear immediately after $inv(deq,k)$ and immediately before $ret(deq,d,k)$ in a sequential execution.
-We also change definition of forward and backward simulation relations by replacing $A\Sigma$ with $AQ\Sigma$ in the original definitions and adding a new condition $(ii-d)$ to each of them:
-\begin{itemize}
-\item[Forward Simulation: (ii-d)] If $(s, lin(deq,d,k), s') \in \delta_{L_1}$ and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $(u, lin(deq,d,k), u') \in \delta_{L_2}$
-\item[Backward Simulation: (ii-d)] If $(s, lin(deq,d,k), s') \in \delta_{L_1}$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $(u, lin(deq,d,k), u') \in \delta_{L_2}$
-\end{itemize}
-Lemma 1 of the previous section still holds if we replace refinement with q-refinement and use new definitions of backward and forward simulations.
+The typical abstract implementation of a concurrent queue maintains a sequence of values, the enqueue method adds a value atomically to the beginning of the sequence, and the dequeue method removes a value from the end of the sequence (if any, otherwise it returns ``empty'').
+Both methods have a fixed linearization point when the update of the shared sequence happens.
+Let $AbsQ_0$ denote this implementation (formally defined in Appendix~\ref{app:absImplQueue}).
+In this section, we describe another abstract implementation, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence.
+We show that there exists a forward simulation from any correct queue implementation where the dequeue methods have fixed linearization points to $AbsQ$.
+We are not aware of any implementation that doesn't satisfy this property.
+As a proof of concept, we describe a forward simulation from the Herlihy\&Wing queue~\cite{journals/toplas/HerlihyW90} to $AbsQ$.
 
-Next, we define the abstract library $L_A$ as follows:
-\begin{itemize}
-\item A queue state consists of a finite sequence of natural numbers representing the queue content and a program counter for each operation.
-More formally: $Q_A \subseteq \mathbb{N}^*  \times (\mathbb{N} \rightarrow Lbl_A)$ where $Lbl_A = \{N, E_0, E_1, E_2, D_0, D_1, D_2 \}$ is the set of transition labels of the operations.
-Operations that have not started yet are mapped to $N$.
-$E_i$ ($D_i$) denotes particular transitions in enqueue (dequeue) operations that will be clear when we define $\delta_A$.
-For a state $q$, we denote the contents of the queue (first field) with $s_q$ and the function that maps operations to labels (the second field) with $f_q$.
-\item Transition labels consists of invocations, returns and linearizations of enqueue and dequeue operations: $\Sigma_A = AQ\Sigma \cup \{lin(enq,d,k)| d \in \mathcal{D}, k \in \mathbb{N}\}$.
-Invocation of $deq$ operation does not have any input and return of $enq$ operation does not have any output value.
-We omit second fields from these labels.
-\item Initial state is the empty queue: ${q_0}_A = (\langle \rangle, f_{{q_0}_A})$ where $f_{{q_0}_A}(i) = N$ for all $i \in \mathbb{N}$.
-\item Each operation consists of invocation, linearization and return steps.
-These are reflected in the transition relation.
-For the below definitions, unchanged parts of the state are omitted from the formulae:
-\begin{itemize}
-\item $(q, inv(enq,d,k), q') \in \delta_A$ iff $ d \neq \texttt{EMPTY} \wedge f_q(k) = N \wedge f_{q'}(k) = E_0$,
-\item $(q,lin(enq,d,k),q') \in \delta_A$ iff $f_q(k) =E_0 \wedge s_{q'} = s_q \circ \langle d \rangle \wedge f_{q'}(k) = E_1$ where $\circ$ is the operation that concatenates two finite sequences,
-\item $(q,ret(enq,k),q') \in \delta_A$ iff $f_q(k) = E_1 \wedge f_{q'}(k) = E_2$,
-\item $(q, inv(deq,k), q') \in \delta_A$ iff $f_q(k) = N \wedge f_{q'}(k) = D_0$
-\item $(q,lin(deq,d,k),q') \in \delta_A$ iff $f_q(k) =D_0 \wedge (d \neq \texttt{EMPTY} \wedge s_q = \langle d \rangle \circ s_{q'} \vee d=\texttt{EMPTY} \wedge s_q = s_{q'} = \langle \rangle )\wedge f_{q'}(k) = D_1$,
-\item $(q,ret(deq,d,k),q') \in \delta_A$ iff $f_q(k) = D_1 \wedge f_{q'}(k) = D_2$.
-\end{itemize}
+\subsection{Abstract Queue Implementation}
 
-\end{itemize} 
-We restrict traces generated by this LTS by neglecting the invalid traces.
-If we project a trace to some operation identifier $k$ and obtain a sequence $\langle ..., inv(enq,d,k), lin(enq,d',k),...\rangle$ or $\langle ..., lin(deq,d,k), ret(deq,d',k),...\rangle$ where $d \neq d'$ we say that this trace is invalid.
+Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation.
+The linearization point of a dequeue operation with return value $d\neq{\tt EMPTY}$ is enabled only if the happens-before order stored in the current state contains a minimal enqueue operation that adds the value $d$.
+The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
+When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores no enqueue operation.
+The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.
 
-We want to introduce $L_I$ as the next step.
-States of $L_I$ consists of strict orders of enqueue operations.
-Nodes of the strict order come from the set $ND = \mathbb{N} \times \mathcal{D} \times \{ \texttt{PENDING}, \texttt{CLOSED}\}$.
-Basically each node is a tuple keeping the operation ID, the value to be enqueued and if this enqueue is pending or closed.
-Strict order also keeps a set of directed edges $ ed \subseteq ED = ND \times ND$.
-Then, a strict order is a tuple $(nd, ed)$ where the set $ed$ obeys the assumption of strict order i.e.
-irreflexivity, asymmetry and transitivity.
-Then, the LTS of $L_I$ is defined as follows:
-\begin{itemize}
-\item A state consists of a partial order and a function keeping the program counter.
-More formally $Q_I \subseteq ND \times ED \times (\mathbb{N} \rightarrow Lbl_I)$ where $Lbl_I = \{N, E_0, E_1, D_0, D_1, D_2\} $ is the set of transition labels.
-$nd_q$ $ed_q$ and $f_q$ denotes the nodes of the strict order in state $q$, edges of the strict order in state $q$ and the function that maps operation to labels in state $q$, respectively.
-\item The transition label set consists of invocation and return action of both methods and linearization action of only the dequeue method: $\Sigma_I = AQ\Sigma$.
-Invocation of $deq$ operation does not have any input and return of $enq$ operation does not have any output value.
-We omit second fields from these labels.
-\item Initial state consists of an empty strict order and a function mapping every operation to $N$: ${q_0}_{I} = (so_{{q_0}_I}, f_{{q_0}_I}$ where $so_{{q_0}_I} =(\emptyset, \emptyset)$ and $f_{{q_0}_I}(i) = N$ for all $i \in \mathbb{N}$.
-\item While defining $\delta_I$, we again omit mentioning about the parts that has not changed:
-\begin{itemize}
-\item $(q, inv(enq,d,k), q') \in \delta_I$ iff $f_q(k) = N \wedge f_{q'}(k) = E_0 \wedge d \neq \texttt{EMPTY} \wedge (k,\_,\_) \notin nd_q \wedge nd_{q'} = nd_q \cup \{ (k,d,\texttt{PENDING}\} \wedge AddNode(ed_{q'}, ed_q,k)$ where $AddNode(ed_{q'}, ed_q,k)$ is true iff $ed_{q'}$ is obtained from $ed_q$ by adding edges from every closed node of $ed_q$ to the node with identifier $k$.
-\item $(q, ret(enq,k), q') \in \delta_I$ iff $f_q(k) =E_0 \wedge f_{q'}(k) = E_1 \wedge ( (k,\_,\texttt{PENDING}) \notin nd_q \wedge so_q = so_{q'} \vee UpdateNode(so_{q'}, so_q, k) )$ where $UpdateNode(so_{q'}, so_q, k)$ is true iff $(k,d,\texttt{PENDING}) \in nd_q$ for some $d \in \mathcal{D}$, it is replaced with node $(k,d,\texttt{CLOSED}$ in the state $q'$ and all the edges adjacent to $(k,d,\texttt{PENDING})$ in state $q$ are replaced with edges adjacent to $(k,d,\texttt{CLOSED})$ in the state $q'$.
-\item $(q, inv(deq,k), q') \in \delta_I$ iff $f_q(k) = N \wedge f_{q'}(k) = D_0$.
-\item $(q, lin(deq,d,k), q') \in \delta_I$ iff $f_q(k) = D_0 \wedge f_{q'}(k) = D_1 \wedge (d = \texttt{EMPTY} \wedge nd_q = ed_q = nd_{q'} = ed_{q'} \emptyset \vee d \neq \texttt{EMPTY} \wedge RemoveNode(so_q, so_{q'},d))$ where $RemoveNode(so_q, so_{q'},d)$ is true iff there is a minimal node $n \in so_q$ of which data value (second field) is $d$ and $so_{q'}$ is obtained from $so_q$ by removing $n$ and all the edges adjacent to it.
-Note that linearization of dequeue could remove a \texttt{PENDING} node.
-\item $(q, ret(deq,d,k) q') \in \delta_I$ iff $f_q(k) = D_1 \wedge f_{q'}(k) = D_2$.
-\end{itemize}
-\end{itemize}
-Again, we omit the inconsistent traces from $L_I$ as in $L_A$. Note that the library $L_I$ is deterministic with respect to alphabe $AQ\Sigma$.
+The states of $AbsQ$ are tuples $\tup{O,<,\ell,rval,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed, $rval:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
+The initial state is $\tup{\emptyset,\emptyset,\emptyset,\emptyset,\emptyset}$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.
 
-Next, we show equivalence of $L_A$ and $L_I$ in terms of q-refinement.
-\begin{lem} 
-$L_I$ is a q-refinement of $L_A$.
-\end{lem}
-\begin{proof}
-We will provide a backward simulation relation btw $L_I$ and $L_A$.
-Given a state $q = (so_q, f_q) \in Q_I$,  $q' \in bs[q]$ iff (i) $s_{q'}$ is a linearization of $so_q$ projected to $d$ fields (second field).
-This linearization may omit the pending elements (the ones of which third field is \texttt{PENDING}), but it must contain all the \texttt{CLOSED} elements.
-Note that pending elements of $so_q$ are maximal and they can appear after the closed elements, towards the end of the sequence.
-(ii) If $f_q(k) \in \{N, D_0, D_1, D_2\}$, then $f_{q'}(k) = f_q(k)$ .
-If $f_q(k) = E_0$, $(k,d,\texttt{PENDING}) \in nd_q$ and this node participates in the linearization $s_{q'}$.
-Then $f_{q'}(k) = E_1$.
-If it does not participate in the linearization, then $f_{q'}(k) = E_0$.
-If $f_q(k) = E_0$ but there is no node with operation identifier $k$ in $nd_q$, then $f_{q'} = E_1$.
-Lastly, if $f_q(k) = E_1$ then $f_{q'}(k) = E_2$.
-Next, we will show that $bs$ is a backward simulation relation.
-\begin{itemize}
-\item[$\langle i \rangle$] $bs[{q_0}_I] = \{ {q_0}_A \}$.
-\item[$\langle ii-a-enq \rangle$] Let $(q, inv(enq,d,k), q') \in \delta_I$ and $t' \in bs[q']$.
-We know that $(k,d,\texttt{PENDING}$ is a maximal element in $so_{q'}$ and either $s_{t'}$ does not linearize it or $s_{t'} = \rho \circ \langle d \rangle \circ \pi$ where $\pi$ consists of only linearization of \texttt{PENDING} elements.
-For the first case $s_{t'} = s_t$ for some $t \in bs[q]$ and $(t,inv(enq,d,k),t') \in \delta_A$.
-For the latter case, $s_t = \rho$ for some $t \in bs[q]$ and $t \xrightarrow{a}_{L_A} t'$ where $a = inv(enq,d,k), lin(enq,d,k), lin(enq,d_1,k_1),...,lin(enq,d_j,k_j)$ where $\pi = d_1,...d_j$.
-The $f_t$ could be obtained easily by observing the sequence $a$ and it can be checked that such $t \in bs[q]$ exists.
-\item[$\langle ii-a-deq \rangle$] Let $(q, inv(deq,k), q') \in \delta_I$ and $t' \in bs[q']$.
-We know that $f_{t'}(k) = D_0$.
-We know that there is a $t \in bs[q]$ such that $s_t = s_{t'}$ (since $so_q = so_{q'}$), $f_t(k) = N$ and $(t,inv(deq,k),t') \in \delta_A$.
-\item[$\langle ii-d \rangle$] Let $(q,lin(deq,d,k),q') \in \delta_I$ and $t' \in bs[q']$.
-First consider the case $d \neq \texttt{EMPTY}$.
-We have two cases: Either the node with operation id $k$ in the $so_q$ was \texttt{PENDING} or \texttt{CLOSED}.
-For both of the cases, we can find $t \in bs[q]$ such that $s_t = \langle d \rangle \circ s_{t'}$ and $f_t(k) = D_0$.
-Hence $(t,lin(deq,d,k),t') \in \delta_I$.
-Next, consider the case $d = \texttt{EMPTY}$.
-Then, we know that $s_{t'} = \langle \rangle$ and there exists $t \in bs[q]$ such that $s_t = \langle \rangle$ and $f_t = D_0$.
-Hence $(t,lin(deq,d,k),t') \in \delta_A$.
-\item[$\langle ii-b-enq \rangle$] Let $(q,ret(enq,k),q') \in \delta_I$ and $t' \in bs[q']$.
-There are two possible cases: There exists a node $(k,d,\texttt{PENDING}) \in nd_q$ or for all nodes $n = (k,\_,\_)$, $n \notin nd_q$.
-For the former case, $s_{t'} = \rho \circ \langle d \rangle \circ \pi$ where $\rho$ is linearization of closed nodes in $nd_{q'}$ and $\pi$ is linearization of some open nodes in $node_{q'}$.
-We also know that $f_{t'}(k) = E_2$.
-Then, there exists a node $t \in bs[q]$ such that $s_t = s_{t'}$ (since nodes that generate $\rho$ are closed in $q$ and that generate $\pi$ are open in $q$) and $f_t(k) = E_1$.
-Hence, $(t,ret(deq,k),t') \in \delta_A$.
-For the latter case, we know that $so_q = so_{q'}$.
-Therefore, there exists $t \in bs[q]$ such that $s_t = s_{t'}$.
-Moreover $f_t(k) = E_1$.
-Hence, $(t,ret(enq,k),t') \in \delta_A$.
-\item[$\langle ii-b-deq \rangle$] Let $(q,ret(deq,d,k), q') \in \delta_I$ and $t' \in bs[q']$.
-Then, we know that $f_{t'}(k) = D_2$ and there exists a $t \in bs[q]$ such that $s_t = s_{t'}$ (since $so_q = so_{q'})$ and $f_t(k) = D_1$.
-Then, $(t,ret(deq,d,k) \in \delta_A$.
-\end{itemize}
-\end{proof}
+\begin{figure} [t]
+{\footnotesize
+  \centering
+  \begin{mathpar}
+    \inferrule[call-enq]{
+      k\not\in dom(cp) \\ 
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{inv(enq,d,k)}
+      O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rval,cp[k\mapsto 1]
+    }\hspace{5mm}
 
-\begin{lem}
-$L_A$ is a q-refinement of $L_I$.
-\end{lem}
-\begin{proof}
-Since $L_I$ is deterministic with respect to the alphabet $AQ\Sigma$, we should be able to find a forward simulation between $L_A$ and $L_I$ if our claim is correct.
-We propose a forward simulation by adding some auxiliary varibles to the state of $L_A$.
-In the new augmented state of $L_A$, the sequence does not only keep the elements of the queue but also the operation identifiers that enqueued them.
-Hence, the sequence consists of  pairs of the form $s_q(i) = (k,d)$ where $q \in Q_A$, $i \in \mathbb{N}$ is an index, $d \in \mathcal{D}\backslash \{ \texttt{EMPTY} \}$ is a value and $k \in \mathbb{N}$ is an operation identifier.
-We also add a set $InvEnq_q$ component to state that keeps the enqueue operations at the point $E_0$ i.e.
-enqueues that are invoked bu have not been linearized yet.
-Elements of $InvEnq_q$ are pairs of the form $(d,k)$ where  $d \in \mathcal{D}\backslash \{ \texttt{EMPTY} \}$ is a value and $k \in \mathbb{N}$ is an operation identifier.
-Then our forward simulation $fs \subseteq Q_A \times Q_I$ relates the state $q = (s_q, f_q) \in Q_A$ to a state $q' =(so_{q'}, f_{q'}) \in Q_I$ iff (i) $f_{q'}(k) = f_q(k)$ for all $f_q(k) \in \{N, D_0, D_1, D_2\}$, (ii) $f_q(k) = E_0$ or $f_q(k) = E_1$ implies $f_{q'}(k) = E_0$, (iii) $f_q(k) = E_2$ implies $f_{q'}(k) = E_1$, (iv) If $(d,k) \in InvEnq_q$ or there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ and $f_q(k) = E_1$, then $(k,d,\texttt{PENDING})$ is a node in $so_{q'}$, (v) if there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ and $f_q(k) = E_2$ then $(k,d,\texttt{CLOSED})$ is a node in $so_{q'}$, (vi) open nodes in $so_{q'}$ are maximal and (vii) there exists a linearization (total order) $lo_{q'}$ of $so_{q'}$ that may omit some open nodes of $so_{q'}$ such that if we project nodes of
-$lo_{q'}$ to the first two fields, this linear order is equal to the sequence $s_q$.
+    \inferrule[call-deq]{
+      k\not\in dom(cp) \\ 
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{inv(deq,k)}
+      O,<,\ell,rval,cp[k\mapsto 1]
+    }\hspace{5mm}
+    \inferrule[ret-deq]{
+       cp(k) = 2 \\
+       rval(k)=d  
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{ret(deq,d,k)}
+      O,<,\ell,rval,cp[k\mapsto 3]
+    }\hspace{5mm}
 
-Next, we will show that $fs$ is a forward simulation relation:
-\begin{itemize}
-\item[$\langle i \rangle$] $fs[(q_0)_A] = \{(q_0)_I\}$
-\item[$\langle ii-a-enq \rangle$] Let $(q,inv(enq,d,k),q') \in \delta_A$ and $t \in fs[q]$.
-We can obtain $t'$ such that $(t,inv(enq,d,k),t') \in \delta_I$.
-We show that $t' \in fs[q']$ by checking seven conditions of our $fs$ relation.
-We omit the node $(k,d)$ while linearizing $so_{t'}$ and obtain $s_{q'}$.
-\item[$\langle ii-a-deq \rangle$] Let $(q, inv(deq,k), q') \in \delta_A$ and $t \in fs[q]$.
-We obtain $t'$ such that $(t, inv(deq,k),t') \in \delta_I$.
-The only difference between $t'$ and $t$ is that $f_{t'}(k) = D_0$.
-We can again see that $t' \in fs[q']$.
-\item[$\langle ii-c \rangle$] Let $(q,lin(enq,d,k),q') \in \delta_A$ and $t \in fs[q]$.
-Pick $t' = t$.
-We can see that $t \in fs[q']$.
-While obtaining linearization of $so_{t'}$, we do not neglect the node $(k,d, \texttt{PENDING})$ this time, although we neglect it in the linearization of $so_t$.
-\item[$\langle ii-d \rangle$] Let $(q, lin(deq,d,k), q') \in \delta_A$ and $t \in fs[q]$.
-Obtain $t'$ such that $(t, lin(deq,d,k), t') \in \delta_I$.
-In $s_q$, $(k,d)$ must be the minimum element.
-Hence, $(k,d,\_)$ is a minimal node in $so_t$ and $lin(deq,d,k)$ is an enabled action in $L_I$.
-This action removes the node $(k,d,\_)$ from $so_t$ and changes $f_t(k)=D_0$ to $f_{t'}(k)= D_1$.
-We can see that $t' \in fs[q']$ by checking the seven conditions.
-\item[$\langle ii-b-enq \rangle$] Let $(q, ret(enq,k), q') \in \delta_A$ and $t \in fs[q]$.
-There are two cases: Either there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ or there is no such $i$.
-For the former case, there is no node of the form $(k,d,\_)$ in $so_t$.
-We obtain $t'$ such that $(t,ret(enq,k),t') \in \delta_I$.
-Then, there is no node of the form $(k,d,\_)$ in $so_{t'}$ neither and $so_{t'} = so_t$.
-since $s_q = s_{q'}$ also holds for this case, $t' \in fs[q']$ holds.
-For the latter case, we again pick $t'$ such that $(t, ret(enq,k), t') \in \delta_I$.
-Again, $so_t = so_{t'}$ and $s_q = s_{q'}$ holds and $t' \in fs[q']$ can be observed.
-\item[$\langle ii-b-deq \rangle$] Let $(q, ret(deq,d,k), q') \in \delta_A$ and $t \in fs[q]$.
-We pick $t'$ such that $(t, ret(deq,d,k), t') \in \delta_I$.
-We know that $so_t = so_{t'}$ and only change is $f_{t'}(k) = D_2$.
-Since we know that $s_q = s_{q'}$ and $f_{q'}(k) = D_2$, $t' \in fs[q']$ holds.
-\end{itemize}
-\end{proof}
+    \inferrule[ret-enq1]{
+      cp(k) = 1 \\
+      k \in O \\
+      \ell(k) = (d,{\tt PEND}) 
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{ret(enq,k)}
+      O,<,\ell[k\mapsto (d,{\tt COMP})],rval,cp[k\mapsto 2]
+    }\hspace{5mm}
+    \inferrule[ret-enq2]{
+      cp(k) = 1 \\
+      k \not\in O 
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{ret(enq,k)}
+      O,<,\ell,rval,cp[k\mapsto 2]
+    }\hspace{5mm}
+
+    \inferrule[lin-deq1]{
+       cp(k) = 1 \\
+       d\neq{\tt EMPTY} \\
+       \exists k'\in O.\ k'\in min(<) \land \ell_1(k')=d
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{lin(deq,d,k)}
+      O\setminus \{k\},<\uparrow k,\ell,rval[k\mapsto d],cp[k\mapsto 2]
+    }\hspace{5mm}
+
+    \inferrule[lin-deq2]{
+       cp(k) = 1 \\
+       O=\emptyset
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{lin(deq,{\tt EMPTY},k)}
+      O,<,\ell,rval[k\mapsto {\tt EMPTY}],cp[k\mapsto 2]
+    }\hspace{5mm}
+
+      \end{mathpar}
+  }
+ \vspace{-5mm}
+  \caption{The transition relation of $AbsQ$.}
+  \label{fig:transitions:AbsQ}
+\vspace{-6mm}
+\end{figure}
+
+The following result states that the library $AbsQ$ has exactly the same set of histories as the standard abstract library $AbsQ_0$ (see Appendix~\ref{app:absImplQueue} for a proof).
+
+\begin{theorem}\label{th:absImplQueue}
+$AbsQ$ is a refinement of $AbsQ_0$ and vice-versa.
+\end{theorem}
+
+A library $L$ is called a \emph{queue implementation} when its alphabet consists of call and return actions defined over $\<Methods>=\{enq,deq\}$.
+Let $LP(deq)$ be the set of all actions $lin(deq,d,k)$ for some $d\in\<Vals>$ and $k\in\<Ops>$.
+A trace of a queue implementation is called $LP(deq)$-complete when every completed dequeue has a linearization point, i.e., TODO.
+A queue implementation $L$ over alphabet $\Sigma$ is called \emph{with fixed dequeue linearization points} if{f} $C\cup R\cup LP(deq)\subseteq \Sigma$
+and every trace $@t\in Tr(L)$ is $LP(deq)$-complete.
+
+The following result shows that $C\cup R\cup LP(deq)$-forward simulations are a sound and complete proof method for showing the correctness of a queue implementation with fixed dequeue linearization points (up to the correctness of the linearization points).
+It is obtained from Theorem~\ref{th:absImplQueue} and Theorem~\ref{th:forSim} using the fact that the alphabet of $AbsQ$ is exactly $C\cup R\cup LP(deq)$ and $AbsQ$ is deterministic.
+
+\begin{corollary}
+A queue implementation $L$ with fixed dequeue linearization points is a $C\cup R\cup LP(deq)$-refinement of $AbsQ_0$ if{f} there exists a $C\cup R\cup LP(deq)$-forward simulation from $L$ to $AbsQ$.
+\end{corollary}
+
+\subsection{A Correctness Proof For Herlihy\&Wing Queue}

### 2017-01-12 draftCAV/refProofs.tex
added refinement proofs and abstract queue implementation

%!TEX root = draft.tex

\section{Refinement Proofs}

Refinement between two libraries $L_1$ and $L_2$ (and thus, linearizability) can be proved using simulation relations which roughly, require that 
the library $L_2$ can mimic every step of $L_1$ using a (possibly empty) sequence of steps. Mainly, there are two kinds of simulation
relations, forward or backward depending on whether the preservation of steps is proved starting from a similar state
forward or backward. In general, forward simulations are easier to establish, but they are incomplete, i.e., refinement may hold while
there exists no forward simulation. On the other hand, proofs based on backward simulations are more difficult, but modulo the
addition of history variables (auxiliary variables recording events in the implementation) backward simulations are complete for 
proving refinement~\footnote{Adding history variables has no impact on the incompleteness of forward simulations.}.

This section introduces notions used to enlarge the scope of proofs based on forward simulations, which is the main goal of this paper.
We first introduce a stronger notion of refinement that compares the set of traces of two libraries projected
on a larger alphabet of actions than the set of call/return actions. These actions are called \emph{observable actions},
and they may represent for instance, linearization points which informally, are points in time where an operation 
is conceptually effectuated. We then define a notion of forward simulation which is complete for proving
this notion of refinement provided that the alphabet of the ``abstract'' library $L_2$ equals the set of 
observable actions and $L_2$ is deterministic. We discuss an application of this result, namely,
proving the correctness of libraries with ``fixed linearization points'', where the so-called ``linearization points''
are fixed to particular implementation actions.

\begin{dfn}\label{def:gref}
Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. 
The library $L_1$ \emph{$\Gamma$-refines} $L_2$ if{f} $Tr(L_1) | \Gamma \subseteq Tr(L_2) | \Gamma$.
%We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
%for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
\end{dfn}

Notice that $\Gamma$-refinement implies refinement for any $\Gamma$ as in Definition~\ref{def:gref}.

We define a notion of \emph{forward} simulation that can be used to prove $\Gamma$-refinement. 
A dual notion of \emph{backward} simulation is defined in Appendix~\ref{app:backSim}. 
%We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement. 
For a relation $R\subseteq A\times B$, $R[X]$ denotes the set of elements of $B$ related to elements of $X$ by $R$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.

%We can extend the relations between libraries by introducing simulation relations. We will later show that simulation relations imply refinement.

\begin{dfn}
Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
\begin{itemize}
\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$ 
\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0<i<|@s|$.
%$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0\leq i<|@s| -1$.
% where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $\delta_2(u,\gamma, u')$. 
\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma_1\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0\leq i<|@s|$.  
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}
%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
A $\Gamma$-forward simulation shows that every step of $L_1$ corresponds to a (possibly empty) sequence of steps of $L_2$. In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action. The rest of the transitions in $L_1$ are matched to a possibly empty sequence of transitions of $L_2$ with arbitrary labels.

The following lemma shows the soundness and the completeness of $\Gamma$-forward simulations.
\begin{theorem}\label{th:forSim}
Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. If there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$. Moreover, if $L_1$ $\Gamma$-refines $L_2$, $\Gamma=\Sigma_2$, and $L_2$ is deterministic, then there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$.
\end{theorem}
\begin{proof}
TODO
\end{proof}

The linearization of a concurrent history can be also defined in terms of \emph{linearization points}. Informally, a linearization point of 
an operation in an execution is a point in time where the operation is conceptually effectuated; given the linearization points of 
each operation, the linearization of a concurrent history is the sequential history which takes operations in order of their linearization points.
For some libraries, the linearization points correspond to a fixed set of actions. For instance, in the case of atomic libraries  
where method bodies are guarded with a global-lock acquisition, the linearization point of every method invocation corresponds to the execution 
of the body. When the linearization points are fixed, we assume that the library is an LTS over an alphabet that includes actions 
$lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$. The action $lin(m,d,k)$ represents the linearization point of the operation $k$ 
which is an invocation of method $m$ returning value $d$.
Let $LP$ denote the set of such actions. We extend the notion of
well-formedness to sequences $@t$ over alphabet $\Sigma$,
such that $C\cup R\cup LP\subseteq \Sigma$, by adding the following constraints:
\begin{itemize}
	\item every linearization point corresponds to an operation that has been called, i.e., TODO
	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
	\item the return value of an operation is fixed by the linearization point, i.e., TODO
\end{itemize}
%The fact that a history $h_1$ is linearizable w.r.t. another sequential history $h_2$ can be stated as: there exists a way of choosing a linearization
%point 
%Formally, a \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by 
%inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that:
%\begin{itemize}
%	\item every linearization point corresponds to an operation that has been called, i.e., TODO
%	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
%	\item the return value of an operation is fixed by the linearization point, i.e., TODO
%\end{itemize}
We assume that libraries produce only well-formed traces. The projection of a library trace over $C\cup R\cup LP$ is called an 
\emph{extended history}. A trace or extended history is called \emph{LP-complete} when every completed operation has a linearization 
point, i.e., TODO. A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup LP\subseteq \Sigma$ 
and every trace $@t\in Tr(L)$ is $LP$-complete.

Proving the correctness of an implementation $L_1$ of a concurrent object such as a queue or a stack with fixed linearization points
reduces to proving that $L_1$ is a $(C\cup R\cup LP)$-refinement of an abstract implementation $L_2$ of the same object where method
bodies are guarded with a global-lock acquisition. Assuming that the abstract implementation is deterministic, which is typically the case,
by Theorem~\ref{th:forSim}, proving $(C\cup R\cup LP)$-refinement is equivalent to finding a $(C\cup R\cup LP)$-forward simulation 
from $L_1$ to $L_2$~\footnote{Note that the alphabet of $L_2$ is exactly $C\cup R\cup LP$.}.

The next sections extend this result to implementations of queues and stacks where the linearization point of the methods 
that \emph{add} values to the collection is \emph{not} fixed.

### 2017-01-13 draftCAV/commit.tex
started the push implementation

@@
+%!TEX root = draft.tex
 \section{Relaxation for the Data Structures Without Fixed Remove Linearization Points}
+
+\subsection{Abstract stack implementation}
+
+The states of $AbsS$ are tuples $\tup{O,<,\ell,rval,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked push operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}\times 2^{\<Ops>}$ labels every identifier with a value, a flag that records whether the operation is pending or completed, and a set of operation identifiers (of pop operations which are allowed to take the value of any other push operation happening after this operation), $rval:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
+The initial state is $\tup{\emptyset,\emptyset,\emptyset,\emptyset,\emptyset}$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.
+
+\begin{figure} [t]
+{\footnotesize
+  \centering
+  \begin{mathpar}
+    \inferrule[call-push]{
+      k\not\in dom(cp) \\ 
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{inv(push,d,k)}
+      O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND},\emptyset)],rval,cp[k\mapsto 1]
+    }\hspace{5mm}
+
+    \inferrule[call-pop]{
+      k\not\in dom(cp) \\
+      \forall k'\not\in maxCompl(<).\ \ell'(k') = \ell(k') \\
+      \forall k'\in maxCompl(<).\ \ell'(k') = (\ell_1(k'),\ell_2(k'),\ell_3(k')\cup\{k\}) \\
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{inv(pop,k)}
+      O,<,\ell',rval,cp[k\mapsto 1]
+    }\hspace{5mm}
+
+    \inferrule[ret-pop]{
+       cp(k) = 2 \\
+       rval(k)=d  
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{ret(pop,d,k)}
+      O,<,\ell,rval,cp[k\mapsto 3]
+    }\hspace{5mm}
+
+    \inferrule[ret-push1]{
+      cp(k) = 1 \\
+      k \in O \\
+      \ell(k) = (d,{\tt PEND},X) 
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{ret(push,k)}
+      O,<,\ell[k\mapsto (d,{\tt COMP},X)],rval,cp[k\mapsto 2]
+    }\hspace{5mm}
+
+    \inferrule[ret-push2]{
+      cp(k) = 1 \\
+      k \not\in O 
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{ret(push,k)}
+      O,<,\ell,rval,cp[k\mapsto 2]
+    }\hspace{5mm}
+
+    \inferrule[lin-pop1]{
+       cp(k) = 1 \\
+       d\neq{\tt EMPTY} \\
+       \exists k',k''\in O.\ k\in \ell_3(k') \land k'\leq k''\land \ell_1(k'')=d
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{lin(pop,d,k)}
+      O\setminus \{k\},<\uparrow k,\ell,rval[k\mapsto d],cp[k\mapsto 2]
+    }\hspace{5mm}
+
+    \inferrule[lin-deq2]{
+       cp(k) = 1 \\
+       O=\emptyset
+    }{
+      O,<,\ell,rval,cp
+      \xrightarrow{lin(deq,{\tt EMPTY},k)}
+      O,<,\ell,rval[k\mapsto {\tt EMPTY}],cp[k\mapsto 2]
+    }\hspace{5mm}
+
+      \end{mathpar}
+  }
+ \vspace{-5mm}
+  \caption{The transition relation of $AbsQ$.}
+  \label{fig:transitions:AbsQ}
+\vspace{-6mm}
+\end{figure}
+
+\subsection{Other}
+
 We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$.

### 2017-01-13 draftCAV/commit.tex
started the stack implementation

--- 2017-01-11 draftCAV/commit.tex
@@
-\section{Relaxation for the Data Structures Without Fixed Remove Linearization Points}
+%!TEX root = draft.tex
+\section{Stacks With Fixed Pop Commit Points}
+
+Differently from queues, there are implementations of concurrent stacks where the linearization points of the remove/pop operations are not fixed, e.g., Time-stamped Stack~\cite{}.
+We define an abstract implementation which simulates such implementations provided that the point in time where the return value of a pop operation is determined corresponds to a fixed action.
+To demonstrate the use of this abstract implementation we provide a correctness proof for a simplified version of Time-Stamped Stack that preserves its most complex behaviors.
+
+\subsection{Abstract stack implementation}
+
+Let $Com(pop)$ be the set of actions $com(pop,d,k)$ with $d\in\<Vals>$ and $k\in\<Ops>$, called \emph{commit points}, representing actions of a pop operation where its return value is determined.
+We define an abstract implementation $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the push operations whose value has not been yet removed.
+Since the commit points are not necessarily linearization points, the pop operations are treated differently.
+Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order and continuously tracks the push operations which are overlapping with it.
+The commit point is enabled only if the return value corresponds to one of the push operations in the initial snapshot, or to a push happening earlier when values from the initial snapshot are removed by other pop operations, or to one of the push operations that overlaps with the pop.
+
+The states of $AbsS$ are tuples $\tup{O,<,\ell,rval,cp,comp,pend}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked push operations), $<\subseteq O\times O$ is a strict partial order over $O$, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a flag that records whether the operation is pending or completed, $coPush:\<Ops> ~> 2^O$ records push operations that were maximal when a pop started or happening earlier provided that the values of all the push happening later have been removed, and $pePush: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
+The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.
+
+\begin{figure} [t]
+{\footnotesize
+  \centering
+  \begin{mathpar}
+    \inferrule[call-push]{
+      k\not\in dom(cp) \\ 
+    }{
+      O,<,\ell,rval,cp,comp,pend 
+      \xrightarrow{inv(push,d,k)} \\
+      O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND},\emptyset)],rval,cp[k\mapsto 1],comp,pend[\forall k'.\ k'\mapsto pend(k')\cup \{k\}]
+    }\hspace{5mm}
+
+    \inferrule[call-pop]{
+      k\not\in dom(cp) \\
+    }{
+      O,<,\ell,rval,cp,comp,pend
+      \xrightarrow{inv(pop,k)} \\
+      O,<,\ell,rval,cp[k\mapsto 1],comp[k\mapsto maxCompl(<)],pend[k\mapsto \{k':\ell_2(k')={\tt PEND}\}]
+    }\hspace{5mm}
+
+    \inferrule[ret-pop]{
+       cp(k) = 2 \\
+       rval(k)=d  
+    }{
+      O,<,\ell,rval,cp,comp,pend
+      \xrightarrow{ret(pop,d,k)}
+      O,<,\ell,rval,cp[k\mapsto 3],comp,pend
+    }\hspace{5mm}
+
+    \inferrule[ret-push1]{
+      cp(k) = 1 \\
+      k \in O \\
+      \ell(k) = (d,{\tt PEND},X) 
+    }{
+      O,<,\ell,rval,cp,comp,pend
+      \xrightarrow{ret(push,k)}
+      O,<,\ell[k\mapsto (d,{\tt COMP},X)],rval,cp[k\mapsto 2],comp,pend
+    }\hspace{5mm}
+
+    \inferrule[ret-push2]{
+      cp(k) = 1 \\
+      k \not\in O 
+    }{
+      O,<,\ell,rval,cp,comp,pend
+      \xrightarrow{ret(push,k)}
+      O,<,\ell,rval,cp[k\mapsto 2],comp,pend
+    }\hspace{5mm}
+
+    \inferrule[lin-pop1]{
+       cp(k) = 1 \\
+       d\neq{\tt EMPTY} \\
+       \exists k'\in O.\ k'\in comp(k)\cup pend(k)\land \ell_1(k')=d
+    }{
+      O,<,\ell,rval,cp,comp,pend
+      \xrightarrow{lin(pop,d,k)} \\
+      O\setminus \{k\},<\uparrow k,\ell,rval[k\mapsto d],cp[k\mapsto 2],comp[\forall k''.\ k'\in comp(k'') \Rightarrow k'' \mapsto (comp(k'')\cup pred_{<}(k'))\setminus\{k'\}],pend
+    }\hspace{5mm}
+
+    \inferrule[lin-deq2]{
+       cp(k) = 1 \\
+       comp(k)=\emptyset
+    }{
+      O,<,\ell,rval,cp,comp,pend
+      \xrightarrow{lin(deq,{\tt EMPTY},k)}
+      O,<,\ell,rval[k\mapsto {\tt EMPTY}],cp[k\mapsto 2],comp,pend
+    }\hspace{5mm}
+
+      \end{mathpar}
+  }
+ \vspace{-5mm}
+  \caption{The transition relation of $AbsQ$.}
+  \label{fig:transitions:AbsS}
+\vspace{-6mm}
+\end{figure}
+
+\subsection{Other}
+
 We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$.

### 2017-01-13 draftCAV/queue.tex
started the stack implementation

@@
 The states of $AbsQ$ are tuples $\tup{O,<,\ell,rval,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed, $rval:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
-The initial state is $\tup{\emptyset,\emptyset,\emptyset,\emptyset,\emptyset}$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.
+The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.
 
@@
 
+TODO NEEDS DATA INDEPENDENCE FOR THE LINEARIZATION POINT TRANSITIONS TO BE DETERMINISTIC
+
 The following result shows that $C\cup R\cup LP(deq)$-forward simulations are a sound and complete proof method for showing the correctness of a queue implementation with fixed dequeue linearization points (up to the correctness of the linearization points).

### 2017-01-14 draftCAV/app-absImplStack.tex
a pass on abstract implementations

--- 2017-01-11 draftCAV/commit.tex
@@
-\section{Relaxation for the Data Structures Without Fixed Remove Linearization Points}
+%!TEX root = draft.tex
+\section{Proof of Theorem~\ref{th:absImplStack}}\label{app:absImplStack}
+
 We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$.

### 2017-01-14 draftCAV/prelim.tex
a pass on abstract implementations

@@
 
+A library $L$ is called a \emph{queue implementation} when its alphabet contains call and return actions defined over $\<Methods>=\{enq,deq\}$ ($enq$ is the name of the methods that enqueue a value, and $deq$ is the name of the methods that dequeue a value) and $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$ where {\tt EMPTY} is the value returned by $deq$ when the queue is empty.
+Similarly, a library $L$ is called a \emph{stack implementation} when $\<Methods>=\{push,pop\}$ and  $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$.
+
 %\begin{example}

### 2017-01-14 draftCAV/queue.tex
a pass on abstract implementations

@@
 %!TEX root = draft.tex
-\section{Queues With Fixed Dequeue Linearization Points}
+\section{Queues With Fixed Dequeue Linearization Points}\label{sec:queues}
 
@@
 Both methods have a fixed linearization point when the update of the shared sequence happens.
-Let $AbsQ_0$ denote this implementation (formally defined in Appendix~\ref{app:absImplQueue}).
+Let $AbsQ_0$ denote this implementation~\footnote{For $\<Methods>=\{enq,deq\}$, the alphabet of $AbsQ_0$ is $C\cup R\cup Lin$.} (formally defined in Appendix~\ref{app:absImplQueue}).
 In this section, we describe another abstract implementation, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence.
-We show that there exists a forward simulation from any correct queue implementation where the dequeue methods have fixed linearization points to $AbsQ$.
+We show that there exists a forward simulation from any correct queue implementation where the \emph{dequeue} methods have fixed linearization points to $AbsQ$.
 We are not aware of any implementation that doesn't satisfy this property.
@@
 
-The states of $AbsQ$ are tuples $\tup{O,<,\ell,rval,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed, $rval:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
+TODO DESCRIBE THE HERLIHY \& WING, EXPLAIN THAT ENQUEUES HAVE NO FIXED LINEARIZATION POINTS, GIVE THE LINEARIZATION POINTS OF DEQUEUES, SHOW HOW $AbsQ$ CAN SIMULATE SOME PROBLEMATIC HISTORIES
+
+The states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed (this flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
 The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.
+The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points.
+The latter are denoted by $lin(deq,d,k)$ where $d\in \<Vals>$ and $k\in\<Ops>$.
+$Lin(deq)$ is the set of all such actions.
+
+Concerning enqueue operations, the rule {\sc call-enq} orders the invoked operation after all the completed enqueue operations stored in the current state, and the rules {\sc ret-enq1}/{\sc ret-enq2} flip the corresponding flag from {\tt PEND} to {\tt COMP} provided that the operation is still present in the current state.
+For dequeue operations, {\sc call-deq} has no effect other than incrementing the control point and {\sc ret-deq} checks whether the return value is the same as the one fixed at the linearization point.
+The linearization point rule {\sc lin-deq1} corresponds to the case of a non-empty queue, showing that $lin(deq,d,k)$ is enabled only if the value $d$ has been added by an enqueue which is minimal in the current happens-before order.
+When it is enabled, it removes the enqueue adding $d$ from the state.
+The linearization point rule {\sc lin-deq2} corresponds to the case of dequeue operations linearized when the queue is empty.
 
 \begin{figure} [t]
-{\footnotesize
+{\scriptsize
   \centering
@@
     }{
-      O,<,\ell,rval,cp
+      O,<,\ell,rv,cp
       \xrightarrow{inv(enq,d,k)}
-      O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rval,cp[k\mapsto 1]
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
+      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
     }\hspace{5mm}
@@
     }{
-      O,<,\ell,rval,cp
+      O,<,\ell,rv,cp
       \xrightarrow{inv(deq,k)}
-      O,<,\ell,rval,cp[k\mapsto 1]
+      O,<,\ell,rv,cp[k\mapsto 1]
     }\hspace{5mm}
@@
        cp(k) = 2 \\
-       rval(k)=d  
+       rv(k)=d  
     }{
-      O,<,\ell,rval,cp
+      O,<,\ell,rv,cp
       \xrightarrow{ret(deq,d,k)}
-      O,<,\ell,rval,cp[k\mapsto 3]
+      O,<,\ell,rv,cp[k\mapsto 3]
     }\hspace{5mm}
@@
     }{
-      O,<,\ell,rval,cp
+      O,<,\ell,rv,cp
       \xrightarrow{ret(enq,k)}
-      O,<,\ell[k\mapsto (d,{\tt COMP})],rval,cp[k\mapsto 2]
+      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto 2]
     }\hspace{5mm}
@@
     }{
-      O,<,\ell,rval,cp
+      O,<,\ell,rv,cp
       \xrightarrow{ret(enq,k)}
-      O,<,\ell,rval,cp[k\mapsto 2]
+      O,<,\ell,rv,cp[k\mapsto 2]
     }\hspace{5mm}
@@
        d\neq{\tt EMPTY} \\
-       \exists k'\in O.\ k'\in min(<) \land \ell_1(k')=d
+       k'\in min(O) \land \ell_1(k')=d
     }{
-      O,<,\ell,rval,cp
+      O,<,\ell,rv,cp
       \xrightarrow{lin(deq,d,k)}
-      O\setminus \{k\},<\uparrow k,\ell,rval[k\mapsto d],cp[k\mapsto 2]
+      O\setminus \{k\},<\uparrow k,\ell,rv[k\mapsto d],cp[k\mapsto 2]
     }\hspace{5mm}
@@
     }{
-      O,<,\ell,rval,cp
+      O,<,\ell,rv,cp
       \xrightarrow{lin(deq,{\tt EMPTY},k)}
-      O,<,\ell,rval[k\mapsto {\tt EMPTY}],cp[k\mapsto 2]
+      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto 2]
     }\hspace{5mm}
@@
  \vspace{-5mm}
-  \caption{The transition relation of $AbsQ$.}
+  \caption{The transition relation of $AbsQ$.
+We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.}
   \label{fig:transitions:AbsQ}
@@
 
-A library $L$ is called a \emph{queue implementation} when its alphabet consists of call and return actions defined over $\<Methods>=\{enq,deq\}$.
-Let $LP(deq)$ be the set of all actions $lin(deq,d,k)$ for some $d\in\<Vals>$ and $k\in\<Ops>$.
-A trace of a queue implementation is called $LP(deq)$-complete when every completed dequeue has a linearization point, i.e., TODO.
-A queue implementation $L$ over alphabet $\Sigma$ is called \emph{with fixed dequeue linearization points} if{f} $C\cup R\cup LP(deq)\subseteq \Sigma$
-and every trace $@t\in Tr(L)$ is $LP(deq)$-complete.
+A trace of a queue implementation is called \emph{$Lin(deq)$-complete} when every completed dequeue has a linearization point, i.e., TODO.
+A queue implementation $L$ over alphabet $\Sigma$ is called \emph{with fixed dequeue linearization points} if{f} $C\cup R\cup Lin(deq)\subseteq \Sigma$
+and every trace $@t\in Tr(L)$ is $Lin(deq)$-complete.
 
@@
 
-The following result shows that $C\cup R\cup LP(deq)$-forward simulations are a sound and complete proof method for showing the correctness of a queue implementation with fixed dequeue linearization points (up to the correctness of the linearization points).
-It is obtained from Theorem~\ref{th:absImplQueue} and Theorem~\ref{th:forSim} using the fact that the alphabet of $AbsQ$ is exactly $C\cup R\cup LP(deq)$ and $AbsQ$ is deterministic.
+The following result shows that $C\cup R\cup Lin(deq)$-forward simulations are a sound and complete proof method for showing the correctness of a queue implementation with fixed dequeue linearization points (up to the correctness of the linearization points).
+It is obtained from Theorem~\ref{th:absImplQueue} and Theorem~\ref{th:forSim} using the fact that the alphabet of $AbsQ$ is exactly $C\cup R\cup Lin(deq)$ and $AbsQ$ is deterministic.
 
 \begin{corollary}
-A queue implementation $L$ with fixed dequeue linearization points is a $C\cup R\cup LP(deq)$-refinement of $AbsQ_0$ if{f} there exists a $C\cup R\cup LP(deq)$-forward simulation from $L$ to $AbsQ$.
+A queue implementation $L$ with fixed dequeue linearization points is a $C\cup R\cup Lin(deq)$-refinement of $AbsQ_0$ if{f} there exists a $C\cup R\cup Lin(deq)$-forward simulation from $L$ to $AbsQ$.
 \end{corollary}

### 2017-01-14 draftCAV/refProofs.tex
a pass on abstract implementations

@@
 which is an invocation of method $m$ returning value $d$.
-Let $LP$ denote the set of such actions. We extend the notion of
+Let $Lin$ denote the set of such actions. We extend the notion of
 well-formedness to sequences $@t$ over alphabet $\Sigma$,
-such that $C\cup R\cup LP\subseteq \Sigma$, by adding the following constraints:
+such that $C\cup R\cup Lin\subseteq \Sigma$, by adding the following constraints:
 \begin{itemize}
@@
 %\end{itemize}
-We assume that libraries produce only well-formed traces. The projection of a library trace over $C\cup R\cup LP$ is called an 
-\emph{extended history}. A trace or extended history is called \emph{LP-complete} when every completed operation has a linearization 
-point, i.e., TODO. A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup LP\subseteq \Sigma$ 
-and every trace $@t\in Tr(L)$ is $LP$-complete.
+We assume that libraries produce only well-formed traces. The projection of a library trace over $C\cup R\cup Lin$ is called an 
+\emph{extended history}. A trace or extended history is called \emph{$Lin$-complete} when every completed operation has a linearization 
+point, i.e., TODO. A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup Lin\subseteq \Sigma$ 
+and every trace $@t\in Tr(L)$ is $Lin$-complete.
 
 Proving the correctness of an implementation $L_1$ of a concurrent object such as a queue or a stack with fixed linearization points
-reduces to proving that $L_1$ is a $(C\cup R\cup LP)$-refinement of an abstract implementation $L_2$ of the same object where method
+reduces to proving that $L_1$ is a $(C\cup R\cup Lin)$-refinement of an abstract implementation $L_2$ of the same object where method
 bodies are guarded with a global-lock acquisition. Assuming that the abstract implementation is deterministic, which is typically the case,
-by Theorem~\ref{th:forSim}, proving $(C\cup R\cup LP)$-refinement is equivalent to finding a $(C\cup R\cup LP)$-forward simulation 
-from $L_1$ to $L_2$~\footnote{Note that the alphabet of $L_2$ is exactly $C\cup R\cup LP$.}.
+by Theorem~\ref{th:forSim}, proving $(C\cup R\cup Lin)$-refinement is equivalent to finding a $(C\cup R\cup Lin)$-forward simulation 
+from $L_1$ to $L_2$~\footnote{Note that the alphabet of $L_2$ is exactly $C\cup R\cup Lin$.}.
 

### 2017-01-14 draftCAV/stack.tex
a pass on abstract implementations

%!TEX root = draft.tex
\section{Stacks With Fixed Pop Commit Points}

While the abstract queue implementation in Section~\ref{sec:queues} can be adapted to stacks~\footnote{The only modification is that the linearization point of a pop $lin(pop,d,k)$ with $d\neq{\tt EMPTY}$ is enabled only if $k$ is added by a push which is maximal in the happens-before order stored in the state.}, it can not simulate (through forward simulations) existing stack implementations like the Time-stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} where the linearization points of the pop operations are not fixed. In this section, we define an abstract implementation which simulates such implementations provided that the point in time where the return value of a pop operation is determined corresponds to a fixed action. To demonstrate the use of this abstract implementation we provide a correctness proof for a simplified version of Time-Stamped Stack that preserves its most complex behaviors.

\subsection{Abstract stack implementation}

We consider a set of actions $Com(pop)=\set{com(pop,d,k):d\in\<Vals>, k\in\<Ops>}$, called \emph{commit points}, representing actions of a pop operation where its return value is determined. Usually, a stack implementation stores the pushed values into a data structure, e.g., a singly-linked list, and a pop operation contains an action (typically, a compare-and-swap) that removes an element from this data structure (or sets a flag associated to this element to a ``deleted'' state). Such an action represents a commit point since the pop operation will return the value stored in this element.

TODO DESCRIBE THE TIME-STAMPED STACK, EXPLAIN THAT NO METHOD HAS A FIXED LIN POINT, GIVE THE COMMIT POINTS, SHOW THAT COMMIT POINTS ARE NOT LIN POINTS

COMMIT POINTS ARE LINEARIZATION POINTS WHEN THE LIBRARY IS INTERPRETED AS A MULTISET

We define an abstract stack implementation $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the push operations whose value has not been yet removed. However, since the commit points are not necessarily linearization points, the pop operations are treated differently. Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order and continuously tracks the push operations which are overlapping with it. The commit point is enabled only if the return value corresponds to one of the push operations in the initial snapshot, or to a push happening earlier when values from the initial snapshot are removed by other pop operations, or to one of the push operations that overlaps with the pop.

TODO AN EXAMPLE FOR EACH CASE (HISTORIES OF THE TIME-STAMPED STACK)

The states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked push operations), $<\subseteq O\times O$ is a strict partial order over $O$, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a flag that records whether the operation is pending or completed, $be:\<Ops> ~> 2^O$ records push operations that were maximal before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.

\begin{figure} [t]
{\scriptsize
  \centering
  \begin{mathpar}
    \inferrule[call-push]{
      k\not\in dom(cp) \\ 
    }{
      O,<,\ell,rv,cp,be,ov 
      \xrightarrow{inv(push,d,k)} \\
      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1],be,ov[\forall k'.\ k'\mapsto ov(k')\cup \{k\}]
    }\hspace{5mm}

    \inferrule[call-pop]{
      k\not\in dom(cp) \\
    }{
      O,<,\ell,rv,cp,be,ov
      \xrightarrow{inv(pop,k)} \\
      O,<,\ell,rv,cp[k\mapsto 1],be[k\mapsto maxCo(O)],ov[k\mapsto {\tt PEND}(O)]
    }\hspace{5mm}

    \inferrule[ret-pop]{
       cp(k) = 2 \\
       rv(k)=d  
    }{
      O,<,\ell,rv,cp,be,ov
      \xrightarrow{ret(pop,d,k)}
      O,<,\ell,rv,cp[k\mapsto 3],be,ov
    }\hspace{5mm}

    \inferrule[ret-push1]{
      cp(k) = 1 \\
      k \in O \\
      \ell(k) = (d,{\tt PEND}) 
    }{
      O,<,\ell,rv,cp,be,ov
      \xrightarrow{ret(push,k)}
      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto 2],be,ov
    }\hspace{5mm}

    \inferrule[ret-push2]{
      cp(k) = 1 \\
      k \not\in O 
    }{
      O,<,\ell,rv,cp,be,ov
      \xrightarrow{ret(push,k)}
      O,<,\ell,rv,cp[k\mapsto 2],be,ov
    }\hspace{5mm}

    \inferrule[lin-pop1]{
       cp(k) = 1 \\
       d\neq{\tt EMPTY} \\
       k'\in be(k)\cup ov(k)\land \ell_1(k')=d
    }{
      O,<,\ell,rv,cp,be,ov
      \xrightarrow{com(pop,d,k)} \\
      O\setminus \{k\},<\uparrow k,\ell,rv[k\mapsto d],cp[k\mapsto 2],be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}],ov
    }\hspace{5mm}

    \inferrule[lin-pop2]{
       cp(k) = 1 \\
       be(k)=\emptyset
    }{
      O,<,\ell,rv,cp,be,ov
      \xrightarrow{com(pop,{\tt EMPTY},k)}
      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto 2],be,ov
    }\hspace{5mm}

      \end{mathpar}
  }
 \vspace{-5mm}
  \caption{The transition relation of $AbsQ$. We use the following notions: $f[\forall k.guard(k)=> k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and $g(k)=expr(k)$ otherwise, $maxCo(O)$ is the set of maximally completed operations in $O$, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.}
  \label{fig:transitions:AbsS}
\vspace{-6mm}
\end{figure}

Let $AbsS_0$ be the standard abstract implementation of a concurrent stack (where elements are stored in a sequence, a push operation adds atomically an element at the beginning of the sequence, and a pop operation removes atomically an element from the beginning of the sequence)~\footnote{For $\<Methods>=\{push,pop\}$, the alphabet of $AbsS_0$ is $C\cup R\cup Lin$.}.
The following result states that the library $AbsS$ has exactly the same set of histories as $AbsS_0$ (see Appendix~\ref{app:absImplStack} for a proof).

\begin{theorem}\label{th:absImplStack}
$AbsS$ is a refinement of $AbsS_0$ and vice-versa.
\end{theorem}

A trace of a queue implementation is called \emph{$Com(pop)$-complete} when every completed pop has a commit point, i.e., TODO. A stack implementation $L$ over alphabet $\Sigma$ is called \emph{with fixed pop commit points} if{f} $C\cup R\cup Com(pop)\subseteq \Sigma$ 
and every trace $@t\in Tr(L)$ is $Com(pop)$-complete.

TODO NEEDS DATA INDEPENDENCE FOR THE COMMIT POINT TRANSITIONS TO BE DETERMINISTIC

As a consequence of Theorem~\ref{th:absImplStack} and Theorem~\ref{th:forSim}, $C\cup R\cup Com(pop)$-forward simulations are a sound and complete proof method for showing the correctness of a stack implementation with fixed pop commit points.

\begin{corollary}
A stack implementation $L$ with fixed pop commit points is a $C\cup R\cup Com(pop)$-refinement of $AbsS_0$ if{f} there exists a $C\cup R\cup Com(pop)$-forward simulation from $L$ to $AbsS$.
\end{corollary}

\subsection{A Correctness Proof For Time-Stamped Stack}

### 2017-01-14 draftCAV/stack.tex
a pass on abstract implementations

@@
 
-As a consequence of Theorem~\ref{th:absImplStack} and Theorem~\ref{th:forSim}, $C\cup R\cup Com(pop)$-forward simulations are a sound and complete proof method for showing the correctness of a stack implementation with fixed pop commit points.
+As a consequence of Theorem~\ref{th:forSim}, $C\cup R\cup Com(pop)$-forward simulations are a sound and complete proof method for showing the correctness of a stack implementation with fixed pop commit points (up to the correctness of the commit points).
+
+TODO EXPLAIN THAT THIS IS DIFFERENT W.R.T. QUEUES ($Abs_0$ doesn't have commit points).
 
 \begin{corollary}
-A stack implementation $L$ with fixed pop commit points is a $C\cup R\cup Com(pop)$-refinement of $AbsS_0$ if{f} there exists a $C\cup R\cup Com(pop)$-forward simulation from $L$ to $AbsS$.
+A stack implementation $L$ with fixed pop commit points is a $C\cup R\cup Com(pop)$-refinement of $AbsS$ if{f} there exists a $C\cup R\cup Com(pop)$-forward simulation from $L$ to $AbsS$.
 \end{corollary}

### 2017-01-14 draftCAV/draft.tex
started draft for CAV
added refinement proofs and abstract queue implementation
started the push implementation
a pass on abstract implementations
title

--- 2017-01-10 draft.tex
@@
 % Suha's Proposal
-\documentclass{article}
+\documentclass[orivec]{llncs}
 
@@
 \usepackage{amsmath}
-\usepackage{amsthm}
+%\usepackage{amsthm}
 \usepackage{color}
+
+\usepackage{amsfonts}
+\usepackage{xspace}
+\usepackage{latexsym}
+\usepackage{wasysym} % causing problems with llncs's bold vectors
+\usepackage{stmaryrd}
+\usepackage{alltt}
+\usepackage{mathpartir}
+\usepackage[ligature,reserved,inference]{semantic}
+\usepackage[lined]{algorithm2e}
+\def\ttat{\mtt{@}} % the at package clobbers this
+\usepackage{at}
+\usepackage{alltt}
+
+%\newcommand{\xRightarrow}[2][]{\ext@arrow 0359\Rightarrowfill@{#1}{#2}}
+
+\usepackage[numbers]{natbib}
+
+\input macros
 
@@
 
-\title{Existence of Forward Simulations for Particular Data Structures}
+\title{Linearizability Proofs Based On Forward Simulations}
+\author{ }
+\institute{ } 
 \begin{document}
+
 \maketitle
+
 \input{prelim}
+\input{refProofs}
+
 \input{queue}
-\section{Existence of Forward Simulations for Stack Implementations that have Fixed Pop Linearization Points}
-\input{commit}
-\input{set}
+%\section{Existence of Forward Simulations for Stack Implementations that have Fixed Pop Linearization Points}
+\input{stack}
+%\input{set}
+
+\bibliographystyle{abbrvnat}
+\bibliography{violin-short}
+
+\newpage
+\appendix
+
+\input{app-backSim}
+\input{app-absImplQueue}
+\input{app-absImplStack}
+
 \end{document}

### 2017-01-15 queue.tex
HW queue linearizability proof added

--- 2017-01-11 draftCAV/queue.tex
@@
 \end{proof}
+
+\subsection{Herlihy \& Wing Queue Linearizability Proof}
+We will begin with formal description of HW Queue library $L_C$:
+\begin{itemize}
+\item States of the queue are forms of the tuple $(O_e, O_d, bck, itms, i, x, rng, cp)$ where $O_e, O_d \subseteq \mathcal{O}$ are disjoint sets of operation identifiers of enqueue and dequeue operations, respectively, $bck \in \mathbb{N}$ and $itms: \mathbb{N} \rightarrow \mathcal{D}$ are global variables of HW queue, $i: (O_e \cup O_d) \rightarrow \mathbb{N}$ and $(O_e \cup O_d) \rightarrow \mathcal{D}$ are local variables common to both enqueue and dequeue operations, although they are used for different purposes, $rng: O_d \rightarrow \mathbb{N}$ is a local variable for dequeue operations and $cp: \mathcal{O} \rightarrow \mathbb{N}$.
+\item Transition labels alphabet consists of the common $AQ\Sigma$ actions.
+In addition $enq1$, $enq2$ actions for enqueue and $deq1$, $deq2$, $deq3$ and $deq4$ operations for dequeue operations are added to these transition labels.
+Semantics of the new actions are derived from the algorithm directly and described in Figure~\ref{fig:transitions:HWQ}.
+\item Initial state is the configuration $(\emptyset, \emptyset, 0, \emptyset, \emptyset, \emptyset, \emptyset)$ where $\emptyset$ maps every element in the domain to $0$ or \texttt{EMPTY}, depending on the range or the domain is the empty set for functions.
+\item Figure~\ref{fig:transitions:HWQ} describes the derivation rules for $L_C$ based on the HW queue algorithm.
+\end{itemize}
+
+We extract the derivation rules from the algorithm.
+
+\begin{figure} [t]
+{\footnotesize
+  \centering
+  \begin{mathpar}
+    \inferrule[call-enq]{
+      k\not\in dom(cp) \\ 
+      d \neq \texttt{EMPTY}\\
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{inv(enq,d,k)}
+      O_e\cup\{k\},O_d,bck, itms, i[k \mapsto 0], x[k \mapsto d], rng, cp[k \mapsto 1]
+    }\hspace{5mm}
+
+    \inferrule[enq-1]{
+      k \in O_e \\
+      cp(k) = 1
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{enq1(k)}
+      O_e, O_d, bck+1, itms, i[k \mapsto bck], x, rng, cp[k \mapsto 2]
+    }
+
+    \inferrule[enq-2]{
+      k \in O_e \\
+      cp(k) = 2
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{enq2(k)}
+      O_e, O_d, bck, itms[i(k) \mapsto x], i, x, rng, cp[k \mapsto 3]
+    }
+
+    \inferrule[ret-enq]{
+      k \in O_e \\
+      cp(k) = 3
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{ret(enq,k)}
+      O_e, O_d, bck, itms, i, x, rng, cp[k \mapsto 4]
+    }
+
+    \inferrule[call-deq]{
+      k\not\in dom(cp) \\ 
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{inv(deq,k)}
+      O_e, O_d \cup \{k\}, bck, itms, i[k \mapsto 0], x[k \mapsto 0], rng[k \mapsto 0], cp[k \mapsto 1]
+    }\hspace{5mm}
+
+    \inferrule[deq-1]{
+      k\in O_d \\ 
+      cp(k) = 1 \\
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{deq1(k)}
+      O_e, O_d, bck, itms, i, x, rng[k \mapsto bck-1], cp[k \mapsto 2]
+    }\hspace{5mm}
+
+    \inferrule[deq-2]{
+      k\in O_d \\ 
+      cp(k) = 2 \\
+      itms[i(k)] = \texttt{EMPTY} \\
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{deq2(k)}
+      O_e, O_d, bck, itms, i, x, rng, cp[k \mapsto 5]
+    }\hspace{5mm}
+
+    \inferrule[lin-deq]{
+      k \in O_d \\ 
+      cp(k) = 2 \\
+      itms(i(k)) \neq \texttt{EMPTY} \\
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{lin(deq,items[i(k)],k)}
+      O_e, O_d, bck, itms[i(k) \mapsto \texttt{EMPTY}], i, x[k \mapsto itms[i(k)] ], rng, cp[k \mapsto 3]
+    }\hspace{5mm}
+
+    \inferrule[deq-3]{
+      k\in O_d \\ 
+      cp(k) = 5 \\
+      i(k) = rng(k)-1\\
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{deq3(k)}
+      O_e, O_d, bck, itms, i[k \mapsto 0], x, rng, cp[k \mapsto 1]
+    }\hspace{5mm}
+
+    \inferrule[deq-4]{
+      k\in O_d \\ 
+      cp(k) = 5 \\
+      i(k) < rng(k)-1\\
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{deq4(k)}
+      O_e, O_d, bck, itms, i[k \mapsto i(k)+1], x, rng, cp[k \mapsto 2]
+    }\hspace{5mm}
+
+    \inferrule[ret-deq]{
+       k \in O_d \\
+       cp(k) = 3 \\  
+    }{
+      O_e, O_d, bck, itms, i, x, rng, cp
+      \xrightarrow{ret(deq,x,k)}
+      O_e, O_d, bck, itms, i, x, rng, cp[k \mapsto 4]
+    }\hspace{5mm}
+
+      \end{mathpar}
+  }
+ \vspace{-5mm}
+  \caption{The transition relation of $HWQ$.}
+  \label{fig:transitions:HWQ}
+\vspace{-6mm}
+\end{figure}
+
+Then, we proceed to prove linearizability of $L_C$.
+\begin{lem}
+$L_C$ is a q-refinement of $L_I$.
+\end{lem}
+\begin{proof}
+Since $L_I$ is deterministic w.r.t.
+alphabet $AQ\Sigma$, there exists a forward simulation relation form $L_C$ to $L_I$ if $L_C$ is a refinement of $L_I$.
+We will define the relation $fs$ and show that it is a forward simulation relation.
+
+The relation $fs$ relates $q = (O_e, O_d, bck, itms, i,x, rng, cp) \in Q_C$ to the state $q'=(O,<,\ell,rval,cp') \in Q_I$ iff:
+\begin{enumerate}
+\item For all $k \in O_e$, if $cp(k) \in \{1,2\}$, then $cp'(k) = 1$; if $cp(k) = 3$ then $cp'(k)=2$.
+For all $k \in O_d$, if $cp(k) \in \{1,2,5\}$, then $cp'(k) = 1$; if $cp(k) = 3$, then $cp'(k)=2$ and if $cp(k)=4$, then $cp'(k)=3$.
+\item $k \in O$ iff $k \in O_e$; and $cp(k) \in \{1,2\}$ or $cp(k) \in \{3,4\} \wedge itms(i(k)) \neq \texttt{EMPTY}$.
+\item $\ell(k)= (d, \texttt{PEND})$ iff $cp(k) \in \{1,2\}$ or $cp(k) =3 \wedge itms(i(k)) \neq \texttt{EMPTY}$; and $x(k) = d$ and $k \in O_e$.
+$\ell(k) = (d,\texttt{COMP}$ iff $k \in O_e$, $cp(k) = 4$, $itms(i(k)) \neq \texttt{EMPTY}$ and $x(k)=d$.
+\item $rval(k)=d$ if $k \in O_d$, $cp[k] = 3$ and $x(k) = d$.
+\item $<$ is a strict partial order and if second field of $\ell(k)$ is \texttt{PEND} for some $k \in O$, then $k$ is maximal in $<$.
+\item For two nodes $m,n \in O$ such that $cp(m), cp(n) > 1$, if $i(m) > i(n)$ then $m \nless n$.
+\item Let $m,n \in O$ be two nodes such that $i(m) < i(n)$ and $cp(m), cp(n) > 1$.
+If there exists a dequeue operation identifier $k \in O_d$ such that $cp(k) \in \{2,5 \}$ such that $i(m) < i(k) \leq i(n)$ and $i(n) < rng(k)$, then $m \nless n$.
+\end{enumerate}
+Next, we show that $fs$ is a forward simulation relation:
+\begin{itemize}
+\item[$\langle i \rangle$] $fs[{q_0}_C] = \{ {q_0}_I \}$
+\item[$\langle ii-a-enq \rangle$] Let $q' \in Q_C$ obtained from $q \in Q_C$ by applying \textsc{call-enq} rule of $L_C$ and $t \in fs[q]$.
+Premise of \textsc{call-enq} rule of $L_I$ holds for $t$ due to premise of \textsc{call-enq} rule of $L_C$ and item 1 in definition of $fs$.
+Let $t'$ be the state obtained by applying \textsc{call-enq} to $t$.
+One can observe that $t' \in fs[q']$ by checking the conditions of the $fs$ stated above.
+Condition $5$ holds because the newly added element $k$ which is \texttt{PEND}  does not occur on the left side of $<_{t'}$ relation and rules of strict partial order is preserved on $<_{q'}$.
+Conditions to re-check: 1, 2, 3 and 5.
+\item[$\langle ii-a-deq \rangle$] Let $q' \in Q_C$ obtained from $q \in Q_C$ by applying \textsc{call-deq} rule of $L_C$ and $t \in fs[q]$.
+Premise of \textsc{call-deq} rule of $L_I$ holds for $t$ due to premise of \textsc{call-deq} rule of $L_C$ and item 1 in definition of $fs$.
+Let $t'$ be the state obtained by applying \textsc{call-deq} to $t$.
+One can observe that $t' \in fs[q']$ by checking the conditions of the $fs$ stated above.
+Conditions to re-check: 1.
+\item[$\langle ii-c-enq1 \rangle$] Let $q' \in Q_C$ obtained from $q  \in Q_C$ by applying \textsc{enq-1} rule of $L_C$ and $t \in fs[q]$.
+Pick $t' = t$.
+We still have $t' \in fs[q']$.
+Conditions 1-5 are trivial to check.
+Conditions $6$ and $7$ still hold because $i_{q'}(k) = bck_{q'}$.
+Conditions to re-check: 1, 2, 3, 6 and 7.
+\item[$\langle ii-c-enq2 \rangle$] Let $q' \in Q_C$ obtained from $q  \in Q_C$ by applying \textsc{enq-2} rule of $L_C$ and $t \in fs[q]$.
+Pick $t' = t$.
+One can see that $t'=t \in fs[q']$ by checking conditions 1-7.
+Conditions to re-check: 1, 2, and 3.
+\item[$\langle ii-c-deq1 \rangle$]Let $q' \in Q_C$ obtained from $q  \in Q_C$ by applying \textsc{deq-1} rule of $L_C$ and $t \in fs[q]$.
+Pick $t' = t$.
+One can see that $t'=t \in fs[q']$ by checking conditions 1-7.
+Nontrivial case is condition $7$ and it still holds since one can see that $i_{q'}(k) =0$, although $rng_{q'}(k)$ is assigned to a new higher value.
+Conditions to re-check: 1 and 7.
+\item[$\langle ii-c-deq2 \rangle$]Let $q' \in Q_C$ obtained from $q \in Q_C$ by applying \textsc{deq-2} rule of $L_C$ and $t \in fs[q]$.
+Pick $t' = t$.
+One can see that $t'=t \in fs[q']$ by checking conditions 1-7.
+Conditions to re-check: 1.
+\item[$\langle ii-c-deq3 \rangle$]Let $q' \in Q_C$ obtained from $q \in Q_C$ by applying \textsc{deq-3} rule of $L_C$ and $t \in fs[q]$.
+Pick $t' = t$.
+One can see that $t'=t \in fs[q']$ by checking conditions 1-7.
+Only non-trivial case is condition $7$.
+Since $i_{q'}(k) = 0$, it cannot impose an extra restriction and it continues to hold.
+Conditions to re-check: 1 and 7.
+\item[$\langle ii-c-deq4 \rangle$]Let $q' \in Q_C$ obtained from $q \in Q_C$ by applying \textsc{deq-4} rule of $L_C$ and $t \in fs[q]$.
+Pick $t' = t$.
+We still have $t' \in fs[q']$.
+Only non-trivial case to check is condition $7$.
+Let $m,n \in O$ such that $i_{t'}(m) < i_{t'}(k) \leq i_{t'}(n)$ and $m <_{t'} n$ which violates condition 7.
+Since condition 7 is not violated while constructing $t$ and applying \textsc{deq-4} increments $i$ by one, the only possible case is $i_{t'}(m) = i_{t'}(k)-1$ and $m <_t n$ holds.
+One can see that this information implies $itms_t(i_t(m)) \neq \texttt{EMPTY}$ and $i_t(m) = i_t(k)$.
+We also know that at the previous step of dequeue with identifier $k$, $itms(i(k))$ was $\texttt{EMPTY}$ since the operation $k$ moved to control point $5$.
+Hence, the \textsc{enq-2} and \textsc{ret-enq} derivations of enqueue operation $m$ must be between the \textsc{deq-2} and current \textsc{deq-5} derivations of $k$.
+Then, we can also observe that $i_t(n) = i_{t'}(n) \geq rng_t(n) = rng_{t'}=n$ which contradicts our initial assumption.
+Conditions to re-check: 1 and 7.
+\item[$\langle ii-d \rangle$]Let $q' \in Q_C$ obtained from $q \in Q_C$ by applying \textsc{lin-deq} rule of $L_C$ and $t \in fs[q]$.
+Premise of the rule \textsc{lin-deq1} of $L_I$ holds for $t$ because of the premise of the rule \textsc{lin-deq} of $L_A$ and conditions on $L_C$.
+More specifically, one can show that there exists an operation $k' \in O_t$ such that $i_q(k') = i_q(k)$.
+Conditions $6$ and $7$ imply that there is no enqueue operation $k'' \in O_t$ such that $cp_q(k'') \in \{2,3,4\}$ and $k'' <_t k'$.
+Condition 3 and 5 imply that there is no enqueue operation $k'' \in O_t$ such that $cp_q(k') = 1$ and $k'' <_t k'$.
+Last two statements imply that $k'$ is minimal.
+Let $t'$ be the state obtained by appyling \textsc{lin-enq1} to $t$.
+One can see that $t' \in fs[q']$ by checking the conditions of $fs$.
+Conditions to re-check: 1, 2, 3, 4 and 5.
+\item[$\langle ii-b-enq \rangle$] Let $q' \in Q_C$ obtained from $q \in Q_C$ by applying \textsc{ret-enq} rule of $L_C$ and $t \in fs[q]$.
+We consider two cases while constructing $t'$.
+If $k \in O_t$, then premises of \textsc{ret-enq1} rule of $L_I$ holds due to premises of \textsc{ret-enq} and conditions of $fs$ and we obtain $t'$ by applying this rule to $t$.
+One can see that $t' \in fs[q']$ still holds by checking the conditions.
+Conditions to re-check:1, 2 and 3.
+The latter case is $k \notin O_t$.
+The premises of \textsc{ret-enq2} are still satisfied by $t$ since $q$ satisfies premises of \textsc{ret-enq} of $L_C$ and $t \in fs$.
+We obtain $t'$ by applying \textsc{ret-enq2} to $t$.
+One can observe that $t' \in fs[q']$ by checking the conditions on $fs$.
+Conditions to re-check: 1.
+\item[$\langle ii-b-deq \rangle$] Let $q' \in Q_C$ obtained from $q \in Q_C$ by applying \textsc{ret-deq} rule of $L_C$ and $t \in fs[q]$.
+State $t$ satisfies premises of the rule \textsc{ret-deq} of $L_I$ since $q$ satisfies premises of the rule \textsc{ret-deq} of $L_C$ and $t \in fs[q]$.
+We apply \textsc{ret-deq} rule of $L_I$ to $t$ to obtain $t'$.
+One can see that $t' \in fs[q']$ by checking the conditions on $fs$.
+Conditions to re-check: 1.
+\end{itemize}
+\end{proof}

### 2017-01-16 draftCAV/prelim.tex
more on queues

@@
 An LTS is \emph{deterministic} if for any state $s$ and any sequence $@t\in \Sigma^*$, there is at most
-one state $s'$ such that $s\xrightarrow{@t}s'$.
+one state $s'$ such that $s\xrightarrow{@t}s'$. More generally, for an alphabet $\Gamma\subseteq \Sigma$,
+an LTS is \emph{$\Gamma$-deterministic} if for any state s and any sequence $@t\in \Gamma^*$, there
+is at most one state $s'$ such that $s\xrightarrow{@t'}s'$ and $@t$ is a subsequence of $@t'$.
 
@@
 refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} when $L_2$ is an atomic library. 
-We fix a set of methods $\<Methods>$ and values $\<Vals>$.
+%We fix a set of methods $\<Methods>$ and values $\<Vals>$.
 
@@
 ``linearized'' to a sequential history admitted by a library $L_2$ used as a specification. 
-Formally, a history $h_1$ is \emph{linearizable} w.r.t.~a library $L_2$ if{f} there exists a sequential history
-$h_2 \in H(L_2)$, with only
-completed operations, such that $h_1 \sqsubseteq h_2$. A library $L_1$
+Formally, a sequential history $h_2$ with only complete operations is called a \emph{linearization} of a history $h_1$ when $h_1 \sqsubseteq h_2$.
+A history $h_1$ is \emph{linearizable} w.r.t.~a library $L_2$ if{f} there exists a linearization $h_2$ of $h_1$ such that 
+$h_2 \in H(L_2)$. A library $L_1$
 is \emph{linearizable} w.r.t. $L_2$, written $L_1 \sqsubseteq L_2$, if{f}
@@
 
-TODO MAKE THE CONNECTION WITH THE NEXT SECTION
+In the rest of the paper, we discuss methods for proving refinement (and thus, linearizability) focusing mainly on queue and stack implementations.

### 2017-01-16 draftCAV/queue.tex
more on queues

@@
 
-The typical abstract implementation of a concurrent queue maintains a sequence of values, the enqueue method adds a value atomically to the beginning of the sequence, and the dequeue method removes a value from the end of the sequence (if any, otherwise it returns ``empty'').
-Both methods have a fixed linearization point when the update of the shared sequence happens.
-Let $AbsQ_0$ denote this implementation~\footnote{For $\<Methods>=\{enq,deq\}$, the alphabet of $AbsQ_0$ is $C\cup R\cup Lin$.} (formally defined in Appendix~\ref{app:absImplQueue}).
+The typical abstract implementation of a concurrent queue, denoted by $AbsQ_0$, maintains a sequence of values, the enqueue method adds a value atomically to the beginning of the sequence, and the dequeue method removes a value from the end of the sequence (if any, otherwise it returns ``empty'').
 In this section, we describe another abstract implementation, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence.
-We show that there exists a forward simulation from any correct queue implementation where the \emph{dequeue} methods have fixed linearization points to $AbsQ$.
-We are not aware of any implementation that doesn't satisfy this property.
-As a proof of concept, we describe a forward simulation from the Herlihy\&Wing queue~\cite{journals/toplas/HerlihyW90} to $AbsQ$.
+We show that there exists a forward simulation from any correct queue implementation where the \emph{dequeue} methods have fixed linearization points (the enqueue methods are unconstrained) to $AbsQ$.
+This covers all the queue implementations that we are aware of.
+We describe a forward simulation from the Herlihy\&Wing queue~\cite{journals/toplas/HerlihyW90} to $AbsQ$.
 
 \subsection{Abstract Queue Implementation}
+
+Before describing our abstract implementation $AbsQ$, we present an example of a queue implementation where only the dequeue methods have fixed linearization points, namely the Herlihy \& Wing queue~\cite{journals/toplas/HerlihyW90} listed in Figure~\ref{fig:HerlihyWing} (we assume that every statement is atomic).
+
+\begin{wrapfigure}{l}{5.7cm}
+%\vspace{-8mm}
+\begin{lstlisting}
+void enq(int x){
+  i = back++;
+  items[i] = x;
+}
+
+int deq() {
+  while (1) {
+    range = back - 1;
+    for (int i = 0; i <= range; i++){
+      x = swap(items[i],null);
+      if ( x != null ) return x;
+    }
+  }
+}
+  \end{lstlisting}
+\vspace{-5mm}
+\caption{Herlihy \& Wing queue.}
+\label{fig:HerlihyWing}
+\vspace{-6mm}
+\end{wrapfigure}
+The shared state consists of an array {\tt items} storing values in the queue and a counter {\tt back} storing the index of the first unused position in {\tt items}.
+Initially, all the positions in the array are {\tt null} and {\tt back} is 0.
+An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueue operations) and then, storing the input value {\tt x} at this position.
+The dequeue method traverses the array {\tt items} starting from the beginning and atomically swaps {\tt null} with the encountered value.
+If the value is not {\tt null}, then the dequeue returns that value.
+If it reaches the end of the array, then it restarts.
+
+The linearization points of the enqueue methods are not fixed, they depend on dequeue operations executing in the future.
+Consider the following trace with two concurrent enqueues (the local variables of operation $k$ are indexed by $k$):
+\begin{align*}
+inv(enq,x,1)\ \ \ inv(enq,y,2)\ \ \ {\tt i}_1 = \mbox{{\tt bck++}}\ \ \ {\tt i}_2 = \mbox{{\tt bck++}}\ \ \ {\tt items[i_2]} = y
+\end{align*}
+%\begin{align*}
+%inv(enq,x,1)\ inv(enq,y,2)\ ({\tt i}_1 = 0,{\tt bck} = 1)\ ({\tt i}_2 = 1,{\tt bck} = 2)\ ({\tt items[1]} = y)
+%\end{align*}
+Assuming that the linearization point corresponds to the assignment of {\tt i}, the history of this trace should be linearized to $inv(enq,x,1)\ ret(enq,1)\ inv(enq,y,2)\ ret(enq,2)$.
+However, a dequeue operation executing until completion after this trace will return value $y$ (only position $1$ is filled in the array {\tt items}) which is not consistent with this linearization.
+On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i_1]} = x$ and a completed dequeue that in this case returns $x$, leads to an incorrect linearization
+\begin{align*}
+inv(enq,y,2)\ ret(enq,2) inv(enq,x,1)\ ret(enq,1)\ inv(deq,3)\ ret(deq,1,3).
+\end{align*}
+
+The dequeue method has a fixed linearization point which corresponds to an execution of {\tt swap} returning a non-null value.
+This action alone contributes to the effect of that value being removed from the queue.
+This claim will be formally proved in Section~\ref{ssec:HerlihyWing}.
+
+Since the linearization points of the enqueues are not fixed, it is not possible to define a forward simulation from the Herlihy \& Wing queue to the standard abstract implementation $AbsQ_0$.
+In the following, we describe the abstract implementation $AbsQ$ we propose for which such a forward simulation does exist.
 
@@
 
-TODO DESCRIBE THE HERLIHY \& WING, EXPLAIN THAT ENQUEUES HAVE NO FIXED LINEARIZATION POINTS, GIVE THE LINEARIZATION POINTS OF DEQUEUES, SHOW HOW $AbsQ$ CAN SIMULATE SOME PROBLEMATIC HISTORIES
+\begin{wrapfigure}{l}{7cm}
+\includegraphics[width=7cm]{fig-queue1.pdf}
+\begin{center}
+(a)
+\end{center}
+
+\includegraphics[width=7cm]{fig-queue2.pdf}
+\begin{center}
+(b)
+\end{center}
+\caption{Simulating queue histories with $AbsQ$.}
+\label{fig:queueSim}
+\end{wrapfigure}
+SHOW HOW $AbsQ$ CAN SIMULATE SOME PROBLEMATIC HISTORIES
 
@@
       \xrightarrow{lin(deq,d,k)}
-      O\setminus \{k\},<\uparrow k,\ell,rv[k\mapsto d],cp[k\mapsto 2]
+      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto 2]
     }\hspace{5mm}
@@
 
+Both methods have a fixed linearization point when the update of the shared sequence happens.
+Let $AbsQ_0$ denote this implementation~\footnote{For $\<Methods>=\{enq,deq\}$, the alphabet of $AbsQ_0$ is $C\cup R\cup Lin$.} (formally defined in Appendix~\ref{app:absImplQueue}).
 The following result states that the library $AbsQ$ has exactly the same set of histories as the standard abstract library $AbsQ_0$ (see Appendix~\ref{app:absImplQueue} for a proof).
@@
 
-\subsection{A Correctness Proof For Herlihy\&Wing Queue}
+\subsection{A Correctness Proof For Herlihy\&Wing Queue}\label{ssec:HerlihyWing}

### 2017-01-16 draftCAV/refProofs.tex
more on queues

@@
 
-Refinement between two libraries $L_1$ and $L_2$ (and thus, linearizability) can be proved using simulation relations which roughly, require that 
-the library $L_2$ can mimic every step of $L_1$ using a (possibly empty) sequence of steps. Mainly, there are two kinds of simulation
-relations, forward or backward depending on whether the preservation of steps is proved starting from a similar state
-forward or backward. In general, forward simulations are easier to establish, but they are incomplete, i.e., refinement may hold while
-there exists no forward simulation. On the other hand, proofs based on backward simulations are more difficult, but modulo the
-addition of history variables (auxiliary variables recording events in the implementation) backward simulations are complete for 
-proving refinement~\footnote{Adding history variables has no impact on the incompleteness of forward simulations.}.
+Library refinement is the instance of a more general notion of refinement between LTSs
+which for some alphabet $\Gamma$ of \emph{observable actions} is defined as the inclusion of sets of 
+traces projected on $\Gamma$. Library refinement corresponds to the case $\Gamma=C\cup R$. 
+Typically, $\Gamma$-refinement between two LTSs $A$ and $B$ is proved using \emph{simulation relations} which roughly, require that 
+$B$ can mimic every step of $A$ using a (possibly empty) sequence of steps. Mainly, there are two kinds of simulation
+relations, forward or backward, depending on whether the preservation of steps is proved starting from a similar state
+forward or backward. It has been shown
+that $\Gamma$-refinement is equivalent to the existence of \emph{backward simulations}, modulo the addition of history variables
+that record events in the implementation, and to the existence of \emph{forward simulations} provided that the right-hand side
+LTS $B$ is $\Gamma$-deterministic~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}. 
+We focus on proofs based on forward simulations because they are easier to automatize.
 
-This section introduces notions used to enlarge the scope of proofs based on forward simulations, which is the main goal of this paper.
-We first introduce a stronger notion of refinement that compares the set of traces of two libraries projected
-on a larger alphabet of actions than the set of call/return actions. These actions are called \emph{observable actions},
-and they may represent for instance, linearization points which informally, are points in time where an operation 
-is conceptually effectuated. We then define a notion of forward simulation which is complete for proving
-this notion of refinement provided that the alphabet of the ``abstract'' library $L_2$ equals the set of 
-observable actions and $L_2$ is deterministic. We discuss an application of this result, namely,
-proving the correctness of libraries with ``fixed linearization points'', where the so-called ``linearization points''
-are fixed to particular implementation actions.
+In general, forward simulations are \emph{not} a complete proof method for library refinement because libraries are not 
+$C\cup R$-deterministic (the same sequence of call and 
+return actions can lead to different states depending on the interleaving of the internal actions).
+However, there are classes of atomic libraries, e.g., libraries with ``fixed linearization points'' (defined later in this section), 
+for which it is possible to identify a larger alphabet $\Gamma$ of observable actions (including call/return actions), 
+and implementations that are $\Gamma$-deterministic. When focusing on queues and stacks, 
+Section~\ref{sec:queues} and Section~\ref{sec:stacks} define other such classes of implementations that cover
+all the implementations that we are aware of.
+
+%Refinement between two libraries $L_1$ and $L_2$  can be proved using  In general, forward simulations are easier to establish, but they are incomplete, i.e., refinement may hold while
+%there exists no forward simulation. On the other hand, proofs based on backward simulations are more difficult, but modulo the
+%addition of history variables (auxiliary variables recording events in the implementation) backward simulations are complete for 
+%proving refinement~\footnote{Adding history variables has no impact on the incompleteness of forward simulations.}.
+
+%This section introduces notions used to enlarge the scope of proofs based on forward simulations, which is the main goal of this paper.
+%We first introduce a stronger notion of refinement that compares the set of traces of two libraries projected
+%on a larger alphabet of actions than the set of call/return actions. These actions are called \emph{observable actions},
+%and they may represent for instance, linearization points which informally, are points in time where an operation 
+%is conceptually effectuated. We then define a notion of forward simulation which is complete for proving
+%this notion of refinement provided that the alphabet of the ``abstract'' library $L_2$ equals the set of 
+%observable actions and $L_2$ is deterministic. We discuss an application of this result, namely,
+%proving the correctness of libraries with ``fixed linearization points'', where the so-called ``linearization points''
+%are fixed to particular implementation actions.
 
@@
 If there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$.
-Moreover, if $L_1$ $\Gamma$-refines $L_2$, $\Gamma=\Sigma_2$, and $L_2$ is deterministic, then there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$.
+Moreover, if $L_1$ $\Gamma$-refines $L_2$ and $L_2$ is $\Gamma$-deterministic, then there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$.
 \end{theorem}
@@
 reduces to proving that $L_1$ is a $(C\cup R\cup Lin)$-refinement of an abstract implementation $L_2$ of the same object where method
-bodies are guarded with a global-lock acquisition. Assuming that the abstract implementation is deterministic, which is typically the case,
+bodies are guarded with a global-lock acquisition. Assuming that the abstract implementation is $(C\cup R\cup Lin)$-deterministic, which is typically the case,
 by Theorem~\ref{th:forSim}, proving $(C\cup R\cup Lin)$-refinement is equivalent to finding a $(C\cup R\cup Lin)$-forward simulation 
-from $L_1$ to $L_2$~\footnote{Note that the alphabet of $L_2$ is exactly $C\cup R\cup Lin$.}.
+from $L_1$ to $L_2$.
 
-The next sections extend this result to implementations of queues and stacks where the linearization point of the methods 
+Section~\ref{sec:queues} and Section~\ref{sec:stacks} extend this result to implementations of queues and stacks where the linearization point of the methods 
 that \emph{add} values to the collection is \emph{not} fixed.

### 2017-01-16 draftCAV/stack.tex
more on queues

@@
 %!TEX root = draft.tex
-\section{Stacks With Fixed Pop Commit Points}
+\section{Stacks With Fixed Pop Commit Points}\label{sec:stacks}
 
@@
       \xrightarrow{com(pop,d,k)} \\
-      O\setminus \{k\},<\uparrow k,\ell,rv[k\mapsto d],cp[k\mapsto 2],be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}],ov
+      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto 2],be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}],ov
     }\hspace{5mm}

### 2017-01-16 draftCAV/queue.tex
half of the simulation relation

@@
 This covers all the queue implementations that we are aware of.
-We describe a forward simulation from the Herlihy\&Wing queue~\cite{journals/toplas/HerlihyW90} to $AbsQ$.
+We describe a forward simulation from the Herlihy\&Wing Queue~\cite{journals/toplas/HerlihyW90} to $AbsQ$.
 
@@
 
-Before describing our abstract implementation $AbsQ$, we present an example of a queue implementation where only the dequeue methods have fixed linearization points, namely the Herlihy \& Wing queue~\cite{journals/toplas/HerlihyW90} listed in Figure~\ref{fig:HerlihyWing} (we assume that every statement is atomic).
+Before defining the abstract implementation $AbsQ$, we describe a queue implementation listed in Figure~\ref{fig:HerlihyWing} and known as the Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short),
+where only the dequeue methods have fixed linearization points.
 
 \begin{wrapfigure}{l}{5.7cm}
-%\vspace{-8mm}
+\vspace{-8mm}
 \begin{lstlisting}
@@
 \vspace{-5mm}
-\caption{Herlihy \& Wing queue.}
+\caption{Herlihy \& Wing Queue. We assume that every statement is atomic.}
 \label{fig:HerlihyWing}
@@
 \end{wrapfigure}
-The shared state consists of an array {\tt items} storing values in the queue and a counter {\tt back} storing the index of the first unused position in {\tt items}.
+The shared state consists of an array {\tt items} storing the values in the queue and a counter {\tt back} storing the index of the first unused position in {\tt items}.
 Initially, all the positions in the array are {\tt null} and {\tt back} is 0.
-An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueue operations) and then, storing the input value {\tt x} at this position.
+An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueue operations) and then, stores the input value {\tt x} at this position.
 The dequeue method traverses the array {\tt items} starting from the beginning and atomically swaps {\tt null} with the encountered value.
@@
 The linearization points of the enqueue methods are not fixed, they depend on dequeue operations executing in the future.
-Consider the following trace with two concurrent enqueues (the local variables of operation $k$ are indexed by $k$):
+Consider the following trace with two concurrent enqueues (${\tt i}(k)$ represents the value of {\tt i} in operation $k$):
 \begin{align*}
-inv(enq,x,1)\ \ \ inv(enq,y,2)\ \ \ {\tt i}_1 = \mbox{{\tt bck++}}\ \ \ {\tt i}_2 = \mbox{{\tt bck++}}\ \ \ {\tt items[i_2]} = y
+inv(enq,x,1)\ \ \ inv(enq,y,2)\ \ \ {\tt i}(1) = \mbox{{\tt bck++}}\ \ \ {\tt i}(2) = \mbox{{\tt bck++}}\ \ \ {\tt items[i(2)]} = y
 \end{align*}
@@
 However, a dequeue operation executing until completion after this trace will return value $y$ (only position $1$ is filled in the array {\tt items}) which is not consistent with this linearization.
-On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i_1]} = x$ and a completed dequeue that in this case returns $x$, leads to an incorrect linearization
+On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i_1]} = x$ and a completed dequeue that in this case returns $x$, leads to the following incorrect linearization:
 \begin{align*}
@@
 
-Since the linearization points of the enqueues are not fixed, it is not possible to define a forward simulation from the Herlihy \& Wing queue to the standard abstract implementation $AbsQ_0$.
-In the following, we describe the abstract implementation $AbsQ$ we propose for which such a forward simulation does exist.
+Since the linearization points of the enqueues are not fixed, it is not possible to define a forward simulation from $\mathit{HWQ}$ to the standard abstract implementation $AbsQ_0$.
+In the following, we describe the abstract implementation $AbsQ$ for which such a forward simulation does exist.
 
@@
 \begin{wrapfigure}{l}{7cm}
+\vspace{-8mm}
 \includegraphics[width=7cm]{fig-queue1.pdf}
-\begin{center}
-(a)
-\end{center}
 
+\vspace{2mm}
 \includegraphics[width=7cm]{fig-queue2.pdf}
-\begin{center}
-(b)
-\end{center}
+\vspace{-5mm}
 \caption{Simulating queue histories with $AbsQ$.}
 \label{fig:queueSim}
+\vspace{-10mm}
 \end{wrapfigure}
-SHOW HOW $AbsQ$ CAN SIMULATE SOME PROBLEMATIC HISTORIES
+Figure 2 pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
+The state of $AbsQ$ after each action is pictured as a graph below the action.
+The nodes of this graph represent enqueue operations and the edges happens-before constraints.
+Each node is labeled by a value (the input of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed.
+For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with input $y$.
+Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.
 
-The states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed (this flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
+Formally, the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed (this flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
 The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.
@@
 \subsection{A Correctness Proof For Herlihy\&Wing Queue}\label{ssec:HerlihyWing}
+
+We describe a forward simulation $\mathit{fs}$ from $\mathit{HWQ}$ to $AbsQ$.
+Essentially, an $AbsQ$ state associated by $\mathit{fs}$ to a $\mathit{HWQ}$ state consists of all the  enqueue operations for which the input is still present in the array {\tt items} and all the pending enqueue operations that have at most reserved an array position, ordered by a relation $<$ satisfying the following:
+\begin{itemize}
+	\item $<$ is consistent with the order in which positions of {\tt items} have been reserved, i.e., for every two enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $k' \not< k$,
+	\item enqueues that haven't reserved an array position are maximal, i.e., for every two enqueues $k$ and $k'$ such that $k'$ has not executed the first statement, we have that $k'\not< k$,
+	\item an enqueue which has reserved an array position $i$ and executed only the first statement can't be ordered before another enqueue that has reserved a position $j > i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal, i.e., for every two enqueues $k$ and $k'$, and a dequeue $k_d$, such that ${\tt i}(k) < {\tt i}(k_d) \leq {\tt i}(k')$, ${\tt x}={\tt null}$, and ${\tt i}(k') < {\tt range}(k_d)$, we have that $k\not< k'$.
+\end{itemize}
+An enqueue is labeled by $(d,{\tt PEND})$ where $d$ is the input value if it's pending and by  $(d,{\tt COMP})$, otherwise.
+Also, for every dequeue operation $k$ such that ${\tt x}(k)=d\neq {\tt null}$, we have that $rv(k)=d$.
+
+We show that $\mathit{fs}$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in\mathit{fs}$.
+
+We first consider the steps of an enqueue operation.
+It is easy to see that transitions labeled by call and return actions can be mimicked by similar transitions of $AbsQ$.
+The internal steps are simulated by an empty sequence of $AbsQ$ transitions because the constraints defining $<$ are an invariant for these steps.
+For instance, let $s'$ be the state obtained from $s$ when a pending enqueue reserves an array position.
+This enqueue operation must be maximal in both $t$ and any state $t'$ related to $s'$.
+Notice that there is no dequeue that can observe this position without reaching the end of the current traversal.

### 2017-01-17 draftCAV/queue.tex
half of the simulation relation
simulation for HW

@@
 We describe a forward simulation from the Herlihy\&Wing Queue~\cite{journals/toplas/HerlihyW90} to $AbsQ$.
+
+TODO HOW CAN WE CONVINCE THAT OTHER IMPLEMENTATIONS HAVE THE SAME PROPERTY ?
 
@@
 \begin{itemize}
-	\item $<$ is consistent with the order in which positions of {\tt items} have been reserved, i.e., for every two enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $k' \not< k$,
-	\item enqueues that haven't reserved an array position are maximal, i.e., for every two enqueues $k$ and $k'$ such that $k'$ has not executed the first statement, we have that $k'\not< k$,
-	\item an enqueue which has reserved an array position $i$ and executed only the first statement can't be ordered before another enqueue that has reserved a position $j > i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal, i.e., for every two enqueues $k$ and $k'$, and a dequeue $k_d$, such that ${\tt i}(k) < {\tt i}(k_d) \leq {\tt i}(k')$, ${\tt x}={\tt null}$, and ${\tt i}(k') < {\tt range}(k_d)$, we have that $k\not< k'$.
+	\item[(a)] pending enqueues are maximal, i.e., for every two enqueues $k$ and $k'$ such that $k'$ is pending, we have that $k'\not< k$,
+	\item[(b)] $<$ is consistent with the order in which positions of {\tt items} have been reserved, i.e., for every two enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $k' \not< k$,
+	\item[(c)] an enqueue which has reserved an array position $i$ %and executed only the first statement 
+	can't be ordered before another enqueue that has reserved a position $j > i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal (we also require that the dequeue has just finished the execution of {\tt swap}), i.e., for every two enqueues $k$ and $k'$, and a dequeue $k_d$, such that
+
+	\noindent
+	{\small
+	\begin{align}
+	\hspace{-8mm}
+	{\tt x}(k_d)={\tt null} \land {\tt i}(k') \leq {\tt range}(k_d) \land {\tt i}(k) \leq {\tt i}(k_d) \leq {\tt i}(k')
+	 \land ({\tt i}(k) = {\tt i}(k_d) => k_d\atCP {\tt if}\text{-}{\tt inc}) \label{eq:inst}
+	\end{align}}
+	we have that $k \not< k'$.
+The predicate $k_d\atCP {\tt if}\text{-}{\tt inc}$ holds when the dequeue $k_d$ is at a control point after a {\tt swap} returning {\tt null} and before the increment of {\tt i}.
 \end{itemize}
+
 An enqueue is labeled by $(d,{\tt PEND})$ where $d$ is the input value if it's pending and by  $(d,{\tt COMP})$, otherwise.
@@
 We show that $\mathit{fs}$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in\mathit{fs}$. 
+We omit discussing the trivial case of transitions labeled by call and return actions (for the return a dequeue operation $k$, we use the equality between the local variable ${\tt x}(k)$ in $s$ and the component $rv(k)$ in $t$).
 
-We first consider the steps of an enqueue operation.
-It is easy to see that transitions labeled by call and return actions can be mimicked by similar transitions of $AbsQ$.
-The internal steps are simulated by an empty sequence of $AbsQ$ transitions because the constraints defining $<$ are an invariant for these steps.
-For instance, let $s'$ be the state obtained from $s$ when a pending enqueue reserves an array position.
-This enqueue operation must be maximal in both $t$ and any state $t'$ related to $s'$.
-Notice that there is no dequeue that can observe this position without reaching the end of the current traversal.
+We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an empty sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in\mathit{fs}$, then $(s',t)\in\mathit{fs}$ for each $AbsQ$ state $t$.
+We focus on the following essential property, called \emph{monotonicity}: the set of possible orders $<$ associated by $\mathit{fs}$ to $s'$ doesn't exclude any order $<$ associated to $s$.
+%Essentially, this boils down to showing that the constraints over $<$ in the definition of $\mathit{fs}$ are an invariant for these steps.
+
+Concerning enqueues, let $s'$ be the state obtained from $s$ when a pending enqueue $k$ reserves an array position.
+This enqueue operation must be maximal in both $t$ and any state $t'$ related to $s'$ (because it is pending).
+Moreover, there is no dequeue that can ``observe'' this position before restarting the array traversal.
+Therefore, item (c) in the definition of $<$ doesn't constrain the order between $k$ and some other enqueue neither in $s$ nor in $s'$.
+Since this transition doesn't affect the constraints on the order between enqueue operations different from $k$ (their local variables remain unchanged), we have that monotonicity holds.
+This property is trivially satisfied by the second step of enqueue which doesn't affect {\tt i}.
+
+To prove monotonicity in the case of dequeue internal steps different from its linearization point, it is important to track the non-trivial instantiations of item (c) in the definition of $<$ over the two states $s$ and $s'$., i.e., the triples $(k,k',k_d)$ for which (\ref{eq:inst}) holds.
+Instantiations that are enabled only in $s'$ may in principle lead to a violation of monotonicity (since they restrict the orders $<$ associated to $s'$).
+For the two steps that begin an array traversal, i.e., reading the index of the last used position and setting the iterator {\tt i} to $0$, there exist no  such new instantiations in $s'$ because the value of {\tt i} is either not set or $0$.
+% (it is trivial to notice that applying these steps doesn't disable such instantiations that were possible in $s$).
+%The same holds for the step incrementing the iterator {\tt i}. 
+%
+%The execution of {\tt swap} returning {\tt null} may introduce one new non-trivial instantiation $(k,k',k_d)$ of item (c).
+%We write ${\tt i}_s(k)$ to refer to the value of the variable {\tt i} of operation $k$ in state $s$.
+Assume that indeed, there exist two enqueue operations $k$ and $k'$ such that ${\tt i}_{s'}(k) < {\tt i}_{s'}(k_d) \leq {\tt i}_{s'}(k')$, ${\tt x}_{s'}(k_d)={\tt null}$, ${\tt i}_{s'}(k') \leq {\tt range}_{s'}(k_d)$ TODO SWAP.
+Since {\tt swap} returnes {\tt null}, the position ${\tt i}_{s'}(k_d)$
+%
+%
+% and ${\tt i}_{s}(k) = {\tt i}_{s'}(k_d)$. The latter constraint guarantees that this instantiation is not enabled in state $s$. The increment of {\tt i} being enabled, implies that 
+%
+The same is true for the increment of the iterator {\tt i} in a dequeue $k_d$ since the predicate $k_d\atCP {\tt if}\text{-}{\tt inc}$ holds in state $s$.
+The execution of {\tt swap} returning {\tt null} in a dequeue $k_d$ enables new instantiations $(k,k',k_d)$ in $s'$, thus adding potentially new constraints $k\not< k'$.
+We show however that these instantiations are vacuous because $k$ must be a pending enqueue in $s$ and thus maximal in any order relation $<$ associated by $\mathit{fs}$ to $s$.
+Let $k$ and $k'$ be two enqueue operations such that together with the dequeue $k_d$ they satisfy the property (\ref{eq:inst}) in $s'$ but not in $s$. 
+We write ${\tt i}_s(k)$ to refer to the value of the variable {\tt i} of operation $k$ in state $s$. 
+This implies that ${\tt i}_{s'}(k) = {\tt i}_{s'}(k_d) \leq {\tt i}_{s'}(k')$ and ${\tt items}[{\tt i}_{s'}(k_d)]={\tt null}$. The latter implies that the enqueue $k$ didn't executed
+the second statement and it is pending (because the position it reserved is still {\tt null}).
+Checking whether the value returned by {\tt swap} is {\tt null} doesn't modify any of the variables in property  (\ref{eq:inst}) and also, it doesn't change the valuation of the predicate $\atCP {\tt if}\text{-}{\tt inc}$.

### 2017-01-17 draftCAV/queue.tex
simulation for HW

@@
 the second statement and it is pending (because the position it reserved is still {\tt null}).
-Checking whether the value returned by {\tt swap} is {\tt null} doesn't modify any of the variables in property  (\ref{eq:inst}) and also, it doesn't change the valuation of the predicate $\atCP {\tt if}\text{-}{\tt inc}$.
+Finally, checking whether the value returned by {\tt swap} is {\tt null} doesn't modify any of the variables in property  (\ref{eq:inst}) and also, it doesn't change the valuation of the predicate $\atCP {\tt if}\text{-}{\tt inc}$.

### 2017-01-17 draftCAV/queue.tex
simulation for HW
small modif

@@
 	\item[(c)] an enqueue which has reserved an array position $i$ %and executed only the first statement 
-	can't be ordered before another enqueue that has reserved a position $j > i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal (we also require that the dequeue has just finished the execution of {\tt swap}), i.e., for every two enqueues $k$ and $k'$, and a dequeue $k_d$, such that
+	can't be ordered before another enqueue that has reserved a position $j > i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal, i.e., for every two enqueues $k$ and $k'$, and a dequeue $k_d$, such that
 
@@
 Finally, checking whether the value returned by {\tt swap} is {\tt null} doesn't modify any of the variables in property  (\ref{eq:inst}) and also, it doesn't change the valuation of the predicate $\atCP {\tt if}\text{-}{\tt inc}$.
+
+TODO LINEARIZATION POINT

### 2017-01-19 commit.tex
Comments on the draft document

--- 2017-01-14 draftCAV/app-absImplStack.tex
@@
-%!TEX root = draft.tex
-\section{Proof of Theorem~\ref{th:absImplStack}}\label{app:absImplStack}
-
+\section{Relaxation for the Data Structures Without Fixed Remove Linearization Points}
 We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$.
@@
 \item $(q, inv(pop,k),q') \in \delta_I$ iff for every open node $n \in nd_q$, $k$ is added to $n.con$ in $q'$ and for every maximally closed node $m \in nd_q$, $k$ is added to $m.mc$ in state $q'$. 
-\item $(q, com(pop,k,d), q') \in \delta_I$ iff there exists a node $n$ in state $q$ such that $n.d = d$ and either $k \in n.con$ or $k \in n.mc$, this node $n$ and all the nodes adjacent to it are removed in the state $q'$, $k$ is removed from all $con$ and $mc$ fields of all nodes in state $q'$ and for all other pop operations $k'$ in $n.mc$ or in $n.con$ and for all states $n' \in nd_q$ such that $n' \rightarrow n \in ed_q$ and for all $n'' \in nd_q$, $n' \rightarrow n'' \in ed_q$ implies $k' \in n''.con$ (implies $k \notin n''.mc$ but it is stronger than this condition: if there are three closed states $p,q,r$ s.t.
+\item $(q, com(pop,k,d), q') \in \delta_I$ iff there exists a node $n$ in state $q$ such that $n.d = d$ and either $k \in n.con$ or $k \in n.mc$, this node $n$ and all the edges adjacent to it are removed in the state $q'$, $k$ is removed from all $con$ and $mc$ fields of all nodes in state $q'$ and for all other pop operations $k'$ in $n.mc$ or in $n.con$ and for all states $n' \in nd_q$ such that $n' \rightarrow n \in ed_q$ and for all $n'' \in nd_q$, $n' \rightarrow n'' \in ed_q$ implies $k' \in n''.con$ (implies $k \notin n''.mc$ but it is stronger than this condition: if there are three closed states $p,q,r$ s.t.
 $p \rightarrow q$, $q \rightarrow r$, $p \rightarrow r$, $k'$ is only in $r.mc$ and we delete $r$, former condition only allows $k' \in q.mc$ whereas the latter one allows $k' \in p.mc$ in addition) we have $k' \in n'.mc$ in the state $q'$.

### 2017-01-19 draftCAV/queue.tex
small modif on the queue picture
Comments on the draft document

@@
 The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.
+\textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too.
+We should be able to return EMPTY if there are only pending nodes.
+Consider the following history of $AbsQ_0$: $inv(enq, d_1,k_1), inv(enq, d_2,k_2), inv(deq, k_3), lin(deq, \texttt{EMPTY}, k_3)$.
+This history should be reflected in $AbsQ$ by enabling lin \texttt{EMPTY} of dequeue when there are pending nodes.
+We also need to update the rules in figure.}
 
@@
 \vspace{-8mm}
-\includegraphics[width=7cm]{fig-queue1.pdf}
-
-\vspace{2mm}
-\includegraphics[width=7cm]{fig-queue2.pdf}
+\includegraphics[width=7cm]{fig-queue12.pdf}
+%
+%\vspace{2mm}
+%\includegraphics[width=7cm]{fig-queue2.pdf}
 \vspace{-5mm}
@@
   \caption{The transition relation of $AbsQ$.
-We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.}
+We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.\textcolor{red}{Call-Enq must have $d!= \texttt{EMPTY}$ as a premise.
+Also lin deq returning empty must be changed as before.}}
   \label{fig:transitions:AbsQ}
@@
 We omit discussing the trivial case of transitions labeled by call and return actions (for the return a dequeue operation $k$, we use the equality between the local variable ${\tt x}(k)$ in $s$ and the component $rv(k)$ in $t$).
+\textcolor{red}{ I think it is good to mention again that call/return actions in HWQ correspond to the same call/return actions in AbsQ (without any other internal action).
+I also think that invoke enqueue operation is non-trivial.
+Preservation of the strict partial order and all of the above items a, b and c needs to be rechecked.}
 

### 2017-01-19 draftCAV/refProofs.tex
Comments on the draft document

@@
 
-The following lemma shows the soundness and the completeness of $\Gamma$-forward simulations.
+The following lemma \textcolor{red}{(following is a theorem)} shows the soundness and the completeness of $\Gamma$-forward simulations.
 \begin{theorem}\label{th:forSim}
@@
 \emph{extended history}. A trace or extended history is called \emph{$Lin$-complete} when every completed operation has a linearization 
-point, i.e., TODO. A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup Lin\subseteq \Sigma$ 
+point, i.e., TODO \textcolor{red}{(So, our machine $AbsQ$ is not $Lin$-complete because add operations in its histories do not have linearization points?)} .
+A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup Lin\subseteq \Sigma$
 and every trace $@t\in Tr(L)$ is $Lin$-complete.
+\textcolor{red}{I did not understand the point of lin-completeness and with fixed linearization points.
+Did not we assume that libraries produce well-formed traces which produce lin-complete traces by definition?}
 
@@
 reduces to proving that $L_1$ is a $(C\cup R\cup Lin)$-refinement of an abstract implementation $L_2$ of the same object where method
-bodies are guarded with a global-lock acquisition. Assuming that the abstract implementation is $(C\cup R\cup Lin)$-deterministic, which is typically the case,
+bodies are guarded with a global-lock acquisition \textcolor{red}{Can we say that a stack implementation is correct although lock-protected-body version $L_2$ is not a correct stack implementation?
+Should not we pick $L_2$ the standart $AbsS_0$ or something equivalent to it?} .
+Assuming that the abstract implementation is $(C\cup R\cup Lin)$-deterministic, which is typically the case \textcolor{red}{typical or not typical?
+$AbsQ_0$ is the typical abstract queue which is not deterministic.},
 by Theorem~\ref{th:forSim}, proving $(C\cup R\cup Lin)$-refinement is equivalent to finding a $(C\cup R\cup Lin)$-forward simulation 

### 2017-01-19 queue.tex
Comments on the draft document

@@
 \item $(q, inv(deq,k), q') \in \delta_I$ iff $f_q(k) = N \wedge f_{q'}(k) = D_0$.
-\item $(q, lin(deq,d,k), q') \in \delta_I$ iff $f_q(k) = D_0 \wedge f_{q'}(k) = D_1 \wedge (d = \texttt{EMPTY} \wedge nd_q = ed_q = nd_{q'} = ed_{q'} \emptyset \vee d \neq \texttt{EMPTY} \wedge RemoveNode(so_q, so_{q'},d))$ where $RemoveNode(so_q, so_{q'},d)$ is true iff there is a minimal node $n \in so_q$ of which data value (second field) is $d$ and $so_{q'}$ is obtained from $so_q$ by removing $n$ and all the edges adjacent to it.
+\item $(q, lin(deq,d,k), q') \in \delta_I$ iff $f_q(k) = D_0 \wedge f_{q'}(k) = D_1 \wedge (d = \texttt{EMPTY} \wedge ed_q = ed_{q'} = \emptyset \wedge Pending(nd_q) \wedge Pending(nd_{q'} \vee d \neq \texttt{EMPTY} \wedge RemoveNode(so_q, so_{q'},d))$ where $RemoveNode(so_q, so_{q'},d)$ is true iff there is a minimal node $n \in so_q$ of which data value (second field) is $d$ and $so_{q'}$ is obtained from $so_q$ by removing $n$ and all the edges adjacent to it and $Pending(nd)$ iff all the nodes in $nd$ are pending.
 Note that linearization of dequeue could remove a \texttt{PENDING} node.
@@
 \begin{enumerate}
-\item For all $k \in O_e$, if $cp(k) \in \{1,2\}$, then $cp'(k) = 1$; if $cp(k) = 3$ then $cp'(k)=2$.
+\item For all $k \in O_e$, if $cp(k) \in \{1,2, 3\}$, then $cp'(k) = 1$; if $cp(k) = 4$ then $cp'(k)=2$.
 For all $k \in O_d$, if $cp(k) \in \{1,2,5\}$, then $cp'(k) = 1$; if $cp(k) = 3$, then $cp'(k)=2$ and if $cp(k)=4$, then $cp'(k)=3$.

### 2017-01-19 draftCAV/queue.tex
small modifs

--- 2017-01-17 draftCAV/queue.tex
@@
 }
-
 int deq() {
@@
 However, a dequeue operation executing until completion after this trace will return value $y$ (only position $1$ is filled in the array {\tt items}) which is not consistent with this linearization.
-On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i_1]} = x$ and a completed dequeue that in this case returns $x$, leads to the following incorrect linearization:
+On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i(1)]} = x$ and a completed dequeue that in this case returns $x$, leads to the following incorrect linearization:
 \begin{align*}
@@
 
-Formally, the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed (this flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
+Formally, the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed (this flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
 The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.
@@
 	\item[(c)] an enqueue which has reserved an array position $i$ %and executed only the first statement 
-	can't be ordered before another enqueue that has reserved a position $j > i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal, i.e., for every two enqueues $k$ and $k'$, and a dequeue $k_d$, such that
+	can't be ordered before another enqueue that has reserved a position $j \geq i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal, i.e., for every two enqueues $k$ and $k'$, and a dequeue $k_d$, such that
 
@@
 	\end{align}}
+
+	\noindent
 	we have that $k \not< k'$.
@@
 
-We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an empty sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in\mathit{fs}$, then $(s',t)\in\mathit{fs}$ for each $AbsQ$ state $t$.
+We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an \emph{empty} sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in\mathit{fs}$, then $(s',t)\in\mathit{fs}$ for each $AbsQ$ state $t$.
 We focus on the following essential property, called \emph{monotonicity}: the set of possible orders $<$ associated by $\mathit{fs}$ to $s'$ doesn't exclude any order $<$ associated to $s$.

### 2017-01-19 draftCAV/stack.tex

@@
   \caption{The transition relation of $AbsQ$.
-We use the following notions: $f[\forall k.guard(k)=> k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and $g(k)=expr(k)$ otherwise, $maxCo(O)$ is the set of maximally completed operations in $O$, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.}
+We use the following notions: $f[\forall k.guard(k)=> k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and $g(k)=expr(k)$ otherwise, $maxCo(O)$ is the set of maximally completed operations in $O$, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise.
+If I understand correctly, there is a problem change in $be$ in lin-pop1 rule.
+What I understand is we move $k''$ to every predecesoor of $k'$.
+But we should not move it to predecessor of $k'$ if this predecessor has a successor $m$ such that $m \in be(k'')$.
+Is this case reflected in your definition.
+Please read sub-bullet 4 of bullet 4 in the definition of $L_I$ in my document.
+Lin-pop2 seems ok for me.
+Last note: Bibliography seems not correct to me.} }
   \label{fig:transitions:AbsS}

### 2017-01-19 draftCAV/queue.tex
small modifs

--- 2017-01-19 draftCAV/queue.tex
@@
 The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
-When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores no enqueue operation.
+When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueue operation.
 The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.
+%\textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too.
+We should be able to return EMPTY if there are only pending nodes.
+Consider the following history of $AbsQ_0$: $inv(enq, d_1,k_1), inv(enq, d_2,k_2), inv(deq, k_3), lin(deq, \texttt{EMPTY}, k_3)$.
+This history should be reflected in $AbsQ$ by enabling lin \texttt{EMPTY} of dequeue when there are pending nodes.
+We also need to update the rules in figure.}
 
@@
 When it is enabled, it removes the enqueue adding $d$ from the state.
-The linearization point rule {\sc lin-deq2} corresponds to the case of dequeue operations linearized when the queue is empty.
+The linearization point rule {\sc lin-deq2} corresponds to the case of dequeue operations linearized with an {\tt EMPTY} return value.
 
@@
       k\not\in dom(cp) \\ 
+      d\neq {\tt EMPTY}
     }{
@@
        cp(k) = 1 \\
-       O=\emptyset
+       \forall o\in O. \ell_2(o)={\tt EMPTY}
     }{
@@
   \caption{The transition relation of $AbsQ$.
-We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.}
+We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.
+  %\textcolor{red}{Call-Enq must have $d!= \texttt{EMPTY}$ as a premise. Also lin deq returning empty must be changed as before.}
+  }
   \label{fig:transitions:AbsQ}

### 2017-01-19 draftCAV/queue.tex
started stacks

@@
 The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
-When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueue operation.
+When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueue operations (the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
 The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point. 
@@
 \end{wrapfigure}
-Figure 2 pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
+Figure~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
 The state of $AbsQ$ after each action is pictured as a graph below the action.
@@
 We show that $\mathit{fs}$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in\mathit{fs}$. 
-We omit discussing the trivial case of transitions labeled by call and return actions (for the return a dequeue operation $k$, we use the equality between the local variable ${\tt x}(k)$ in $s$ and the component $rv(k)$ in $t$).
-\textcolor{red}{ I think it is good to mention again that call/return actions in HWQ correspond to the same call/return actions in AbsQ (without any other internal action).
+We omit discussing the trivial case of transitions labeled by call and return actions which are simulated by similar transitions of $AbsQ$ (for the return a dequeue operation $k$, we use the equality between the local variable ${\tt x}(k)$ in $s$ and the component $rv(k)$ in $t$).
+%\textcolor{red}{ I think it is good to mention again that call/return actions in HWQ correspond to the same call/return actions in AbsQ (without any other internal action).
 I also think that invoke enqueue operation is non-trivial.

### 2017-01-19 draftCAV/stack.tex
small modifs
started stacks

@@
 Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order and continuously tracks the push operations which are overlapping with it.
-The commit point is enabled only if the return value corresponds to one of the push operations in the initial snapshot, or to a push happening earlier when values from the initial snapshot are removed by other pop operations, or to one of the push operations that overlaps with the pop.
+The commit point is enabled only if the return value corresponds to one of the push operations in the initial snapshot, or to a push happening earlier when the values from the initial snapshot are removed by other pop operations, or to one of the push operations that overlaps with the pop.
+TODO EMPTY
+
+\begin{wrapfigure}{l}{7cm}
+\vspace{-8mm}
+\includegraphics[width=7cm]{fig-stack.pdf}
+%
+%\vspace{2mm}
+%\includegraphics[width=7cm]{fig-queue2.pdf}
+\vspace{-5mm}
+\caption{Simulating stack histories with $AbsS$.}
+\label{fig:stackSim}
+\vspace{-6mm}
+\end{wrapfigure}
+Figure~\ref{fig:stackSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
+The state of $AbsQ$ after each action is pictured as a graph below the action.
+The nodes of this graph represent enqueue operations and the edges happens-before constraints.
+Each node is labeled by a value (the input of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed.
+For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with input $y$.
+Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.
 
@@
        d\neq{\tt EMPTY} \\
-       k'\in be(k)\cup ov(k)\land \ell_1(k')=d
+       k'\in be(k)\cup ov(k)\land \ell_1(k')=d \\
+       \forall k_1.\ k'\not\in be(k_1) \Rightarrow be'(k_1)=be(k_1) \\     \\
+       \forall k_1.\ k'\in be(k_1) \Rightarrow be'(k_1)=(be(k_1)\setminus\{k'\})\cup \{k_2: k_2\in pred_{<}(k') \land \forall k_3. k_2<k_3 => k_3\not\in be(k_1)\} \\
     }{
@@
       \xrightarrow{com(pop,d,k)} \\
-      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto 2],be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}],ov
+      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto 2],be',ov %be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}]
     }\hspace{5mm}
@@
   \caption{The transition relation of $AbsQ$.
-We use the following notions: $f[\forall k.guard(k)=> k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and $g(k)=expr(k)$ otherwise, $maxCo(O)$ is the set of maximally completed operations in $O$, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise.
+We use the following notions: $f[\forall k.guard(k)=> k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and $g(k)=expr(k)$ otherwise, $maxCo(O)$ is the set of maximally completed operations in $O$, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.%\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise.
 If I understand correctly, there is a problem change in $be$ in lin-pop1 rule.
@@
 Lin-pop2 seems ok for me.
-Last note: Bibliography seems not correct to me.} }
+Last note: Bibliography seems not correct to me.}
+  }
   \label{fig:transitions:AbsS}

### 2017-01-20 draft.tex
HW queue linearizability proof added
more on queues
Simplified TS-Stack Algorithm

@@
 
+\usepackage{amsfonts}
+\usepackage{xspace}
+\usepackage{latexsym}
+\usepackage{wasysym} % causing problems with llncs's bold vectors
+\usepackage{stmaryrd}
+\usepackage{alltt}
+\usepackage{mathpartir}
+\usepackage[ligature,reserved,inference]{semantic}
+\usepackage[lined]{algorithm2e}
+\def\ttat{\mtt{@}} % the at package clobbers this
+\usepackage{at}
+\usepackage{alltt}
+
+%\usepackage[numbers]{natbib}
+
+%\input macros
+
 \newtheorem{dfn}{Definition}
@@
 \input{queue}
+
 \section{Existence of Forward Simulations for Stack Implementations that have Fixed Pop Linearization Points}
 \input{commit}
+First, describe the TS-Stack algorithm:
+
+\begin{lstlisting}
+Type Node{
+ elt:Vals;
+ ts: int;
+ next: Node;
+ taken: bool;
+};
+
+pools: [Tid]Node; //Keeps top of pools, 
+//Initially they point to sentinel dummy nodes s s.t.
+//s.next = s and s.ts = -1 and s.taken=false
+TSGen: int;    //Timestamp generator initially 0
+
+void push(e:Vals, tid:Tid){
+  Node n = new Node(e, MAX_INT, NULL, false);
+  n.next = pools[tid];
+  pools[tid] = n;
+  i: int = TSGen++;
+  n.ts = i;
+}
+
+Vals pop(){
+ success: bool = false;
+ maxTS: int = -1;
+ youngest: Node = NULL;
+ while(!success){
+  maxTS: int = -1;
+  youngest = NULL;
+  for(current in pools){
+    n: Node = current;
+    while(n.taken && n.next != n)
+      n = n.next;
+    if(maxTS < n.ts)
+      youngest = n;
+  }
+  success = CAS(youngest.taken,false,true);
+ }
+ return youngest;
+}
+
+\end{lstlisting}
 \input{set}

### 2017-01-20 draftCAV/stack.tex
started stacks

@@
 
+\begin{wrapfigure}{l}{6.1cm}
+\vspace{-14mm}
+\begin{lstlisting}
+struct Node{
+  int data;
+  int ts;
+  Node* next;
+  boolean taken;
+};
+Node* pools[maxThreads];
+int TS = 0;
+
+void push(int x) {
+  Node* n = new Node(x,MAX_INT,
+                        null,false);
+  n->next = pools[myTID];
+  pools[myTID] = n;
+  int i = TS++;
+  n->ts = i;
+}
+int pop() {
+  boolean success = false;
+  int maxTS = -1;
+  Node* youngest = null;
+  while ( !success ) {
+    maxTS = -1; youngest = null;
+    for(int i = 0; i < maxThreads; i++){
+      Node* n = pools[i];
+      while (n->taken && n->next != null)
+        n = n->next;
+      if(maxTS < n->ts) {
+        maxTS = n->ts; youngest = n;
+      }
+    }
+    success=CAS(youngest->taken,false,true);
+  }
+  return youngest;
+}
+\end{lstlisting}
+\vspace{-6mm}
+\caption{Time-Stamped Stack. We assume that every statement is atomic.}
+\label{fig:HerlihyWing}
+\vspace{-1mm}
+\end{wrapfigure}
 We consider a set of actions $Com(pop)=\set{com(pop,d,k):d\in\<Vals>, k\in\<Ops>}$, called \emph{commit points}, representing actions of a pop operation where its return value is determined.
@@
 Such an action represents a commit point since the pop operation will return the value stored in this element.
+
+// Initially each pool[i] points to a 
+// sentinel dummy node with ts = -1, 
+// next = null, taken=false
 

### 2017-01-20 draftCAV/stack.tex
started stacks

--- 2017-01-19 draftCAV/stack.tex
@@
 
+\begin{wrapfigure}{l}{6.1cm}
+\vspace{-14mm}
+\begin{lstlisting}
+struct Node{
+  int data;
+  int ts;
+  Node* next;
+  boolean taken;
+};
+Node* pools[maxThreads];
+int TS = 0;
+
+void push(int x) {
+  Node* n = new Node(x,MAX_INT,
+                        null,false);
+  n->next = pools[myTID];
+  pools[myTID] = n;
+  int i = TS++;
+  n->ts = i;
+}
+int pop() {
+ boolean success = false;
+ int maxTS = -1;
+ Node* youngest = null;
+ while ( !success ) {
+   maxTS = -1; youngest = null;
+   for(int i = 0; i < maxThreads; i++){
+     Node* n = pools[i];
+     while (n->taken && n->next != null)
+       n = n->next;
+     if(maxTS < n->ts) {
+       maxTS = n->ts; youngest = n;
+     }
+   }
+   success=CAS(youngest->taken,false,true);
+ }
+ return youngest->data;
+}
+\end{lstlisting}
+\vspace{-6mm}
+\caption{Time-Stamped Stack. We assume that every statement is atomic.}
+\label{fig:TimeStamped}
+\vspace{-1mm}
+\end{wrapfigure}
+We explain the meaning of the commit points on a simplified version of the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} given in Figure~\ref{fig:TimeStamped}.
+This stack implementation maintains a vector of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
+Initially, each list contains a sentinel dummy node pointing to {\tt null} with timestamp $-1$ and the flag {\tt taken} set to {\tt false}.
+
+Pushing a value to the stack proceeds in several steps: adding a node with maximal timestamp, asking for a new timestamp (given by the value of the shared integer {\tt TS}), and updating the timestamp of the added node.
+Popping a value from the stack consists in traversing all the lists, finding the first element which doesn't represent a removed value (i.e., {\tt taken} is {\tt false}) in each list, and selecting the element with the maximal timestamp.
+A compare-and-swap (CAS) is used to set the {\tt taken} flag of this element to {\tt true}.
+The procedure restarts if the CAS fails.
+
 We consider a set of actions $Com(pop)=\set{com(pop,d,k):d\in\<Vals>, k\in\<Ops>}$, called \emph{commit points}, representing actions of a pop operation where its return value is determined.

### 2017-01-21 draftCAV/stack.tex
started explaining commit points

@@
 
-Pushing a value to the stack proceeds in several steps: adding a node with maximal timestamp, asking for a new timestamp (given by the value of the shared integer {\tt TS}), and updating the timestamp of the added node.
+Pushing a value to the stack proceeds in several steps: adding a node with maximal timestamp in the list associated to the thread executing the push (given by the special variable {\tt myTID}), asking for a new timestamp (given by the value of the shared integer {\tt TS}), and updating the timestamp of the added node.
 Popping a value from the stack consists in traversing all the lists, finding the first element which doesn't represent a removed value (i.e., {\tt taken} is {\tt false}) in each list, and selecting the element with the maximal timestamp.
@@
 The procedure restarts if the CAS fails.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=11cm]{fig-timeStamp.pdf}
+\caption{An execution of the Time-Stamped Stack.}
+\label{fig:commit}
+\end{figure}
+
+The push operations don't have a fixed linearization point because adding a node to a list and updating its timestamp are not executed in a single atomic step.
+The nodes can be added in an order which is not consistent with the order between the timestamps assigned later in the execution.
+Also, the value added by a push that just added an element to a list can be popped before the value added by a completed push (since it has a maximal timestamp).
+The same holds for pop operations.
+The only reasonable choice for a linearization point is a successful CAS (that results in updating the field {\tt taken}).
+
+TODO GIVE AN EXAMPLE TO SHOW THE ISSUE WITH POP LINEARIZATION POINTS, AND TO SHOW THAT COMMIT POINTS ARE RIGHT LIMITS FOR THE INTERVAL OF A LIN POINT, I.E., A POP WITH A LATER COMMIT POINT CAN BE LINEARIZED BEFORE.
+NEED ONLY 2 POPS IN AN EXECUTION
 

### 2017-01-21 draftCAV/prelim.tex
started stacks
started explaining commit points

@@
 which method calls and returns overlap. 
+For simplicity, Definition~\ref{def:libraries} assumes that every thread performs a single operation.
+The extension to multiple operations per thread is straightforward, e.g.
+the closure rules must assume that the actions $a$ and $c$ belong to different threads
 %A library which accesses the client's
@@
 
-\begin{dfn}
+%\textcolor{red}{We have two Definition 1. The problem occurs because we used different commands for definitions.}
+\begin{definition}
 Let $L_1$ and $L_2$ be two libraries. 
@@
 %for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
-\end{dfn}
+\end{definition}
 

### 2017-01-21 draftCAV/stack.tex
started explaining commit points

@@
 
-\begin{wrapfigure}{l}{6.1cm}
-\vspace{-14mm}
+\begin{wrapfigure}{l}{5.4cm}
+\vspace{-11mm}
 \begin{lstlisting}
@@
    maxTS = -1; youngest = null;
-   for(int i = 0; i < maxThreads; i++){
+   for(int i=0; i<maxThreads; i++){
      Node* n = pools[i];
-     while (n->taken && n->next != null)
+     while (n->taken && n->next != n)
        n = n->next;
@@
    }
-   success=CAS(youngest->taken,false,true);
+   if (youngest != null)
+     success=CAS(youngest->taken,
+                       false,true);
  }
@@
 This stack implementation maintains a vector of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
-Initially, each list contains a sentinel dummy node pointing to {\tt null} with timestamp $-1$ and the flag {\tt taken} set to {\tt false}.
+Initially, each list contains a sentinel dummy node pointing to itself with timestamp $-1$ and the flag {\tt taken} set to {\tt false}.
 
@@
 \includegraphics[width=11cm]{fig-timeStamp.pdf}
-\caption{An execution of the Time-Stamped Stack.}
+\caption{An execution of the Time-Stamped Stack.
+An operation is pictured by a line delimited by two circles denoting the call and respectively, the return action.
+For pop operations, we introduce another circle to represent a CAS returning {\tt true} which is the commit point of that operation.
+The state of the library after an execution prefix delimited at the right by a dotted line is pictured in the bottom part (the picture immediately to the left of the dotted line).
+Also, ${\tt i}_1$ denotes the value of {\tt i} in the pop operation with identifier 1.}
 \label{fig:commit}

### 2017-01-21 draftCAV/intro.tex
finished stacks except TS proof and completed queue stuff

%!TEX root = draft.tex
\section{Introduction}

\newpage~\newpage

### 2017-01-21 draftCAV/queue.tex
started explaining commit points
finished stacks except TS proof and completed queue stuff

@@
 \vspace{-5mm}
-\caption{Simulating queue histories with $AbsQ$.}
+\caption{Simulating queue histories with $AbsQ$. The order between actions is from left to right.}
 \label{fig:queueSim}
@@
 
-Formally, the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed (this flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
-The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.
+Formally, the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed (this flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue ($R_1,R_2,R_3$) operation.
+The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}.
 The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points.
-The latter are denoted by $lin(deq,d,k)$ where $d\in \<Vals>$ and $k\in\<Ops>$.
+The latter are denoted by $lin(deq,d,k)$ where $d\in \<Vals>$, $k\in\<Ops>$.
 $Lin(deq)$ is the set of all such actions.
@@
       %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
-      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
+      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1]
     }\hspace{5mm}
@@
       \xrightarrow{inv(deq,k)}
-      O,<,\ell,rv,cp[k\mapsto 1]
+      O,<,\ell,rv,cp[k\mapsto R_1]
     }\hspace{5mm}
     \inferrule[ret-deq]{
-       cp(k) = 2 \\
+       cp(k) = R_2 \\
        rv(k)=d  
@@
       \xrightarrow{ret(deq,d,k)}
-      O,<,\ell,rv,cp[k\mapsto 3]
+      O,<,\ell,rv,cp[k\mapsto R_3]
     }\hspace{5mm}
@@
     \inferrule[ret-enq1]{
-      cp(k) = 1 \\
+      cp(k) = A_1 \\
       k \in O \\
@@
       \xrightarrow{ret(enq,k)}
-      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto 2]
+      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2]
     }\hspace{5mm}
     \inferrule[ret-enq2]{
-      cp(k) = 1 \\
+      cp(k) = A_1 \\
       k \not\in O 
@@
       \xrightarrow{ret(enq,k)}
-      O,<,\ell,rv,cp[k\mapsto 2]
+      O,<,\ell,rv,cp[k\mapsto A_2]
     }\hspace{5mm}
@@
     \inferrule[lin-deq1]{
-       cp(k) = 1 \\
+       cp(k) = R_1 \\
        d\neq{\tt EMPTY} \\
-       k'\in min(O) \land \ell_1(k')=d
+       k'\in min(O) \\ 
+       \ell_1(k')=d
     }{
@@
       \xrightarrow{lin(deq,d,k)}
-      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto 2]
+      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2]
     }\hspace{5mm}
-
     \inferrule[lin-deq2]{
-       cp(k) = 1 \\
+       cp(k) = R_1 \\
        \forall o\in O. \ell_2(o)={\tt EMPTY}
@@
       \xrightarrow{lin(deq,{\tt EMPTY},k)}
-      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto 2]
+      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2]
     }\hspace{5mm}    
-
       \end{mathpar}
@@
 
-TODO LINEARIZATION POINT
+Finally, we show that the linearization point of a dequeue $k$ of $\mathit{HWQ}$, i.e., an execution of {\tt swap} returning a non-null value $d$, from state $s$ and leading to a state $s'$ is simulated by a transition labeled by $lin(deq,d,k)$ of $AbsQ$ from state $t$.
+By the definition of $\mathit{HWQ}$, there exists a unique enqueue $k_e$ which filled the position updated by $k$, i.e., ${\tt i}_s(k_e)=i_s(k)$ and ${\tt x}_{s'}(k)={\tt x}_s(k_e)$.
+We show that $k_e$ is minimal in the order relation $<$ of $t$ which implies that $lin(deq,d,k)$ is enabled in $t$.
+Thus, instantiating item (c) in the definition of $<$ with $k'=k_e$ and $k_d=k$ we get that every enqueue that reserved a position smaller than the one of $k_e$ can't be ordered before $k_e$ in the order $<$.
+Also, applying item (b) with $k=k_e$ we get the same for every enqueue that reserved a bigger position.
+An enqueue that didn't reserved a position is by definition maximal in the order $<$ and therefore, not a predecessor of $k_e$.
+The state $t'$ obtained from $t$ through a $lin(deq,d,k)$ transition is related to $s'$ because essentially, (1) the value added by $k_e$ is not anymore present in {\tt items} which implies that $k_e$ doesn't occur in any $AbsQ$ state related to $s'$, and (2) the value of ${\tt x}(k)$ is set to $d\neq {\tt null}$ which implies that $rv(k)$ is set to $d$ in every $AbsQ$ state related to $s'$.

### 2017-01-21 draftCAV/stack.tex
finished stacks except TS proof and completed queue stuff

@@
 
-While the abstract queue implementation in Section~\ref{sec:queues} can be adapted to stacks~\footnote{The only modification is that the linearization point of a pop $lin(pop,d,k)$ with $d\neq{\tt EMPTY}$ is enabled only if $k$ is added by a push which is maximal in the happens-before order stored in the state.}, it can not simulate (through forward simulations) existing stack implementations like the Time-stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} where the linearization points of the pop operations are not fixed.
+While the abstract queue implementation in Section~\ref{sec:queues} can be adapted to stacks~\footnote{The only modification is that the linearization point $lin(pop,d,k)$ with $d\neq{\tt EMPTY}$ is enabled only if $k$ is added by a push which is maximal in the happens-before order stored in the state.}, it can not simulate (through forward simulations) existing stack implementations like the Time-stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} where the linearization points of the pop operations are not fixed.
 In this section, we define an abstract implementation which simulates such implementations provided that the point in time where the return value of a pop operation is determined corresponds to a fixed action.
@@
 \label{fig:TimeStamped}
-\vspace{-1mm}
+\vspace{-8mm}
 \end{wrapfigure}
@@
 \includegraphics[width=11cm]{fig-timeStamp.pdf}
+\vspace{-4.5mm}
 \caption{An execution of the Time-Stamped Stack.
 An operation is pictured by a line delimited by two circles denoting the call and respectively, the return action.
-For pop operations, we introduce another circle to represent a CAS returning {\tt true} which is the commit point of that operation.
+Pop operations with identifier $k$ and removing value $d$ are denoted by $pop(d,k)$ and their representation includes another circle to represent a CAS returning {\tt true} which is the commit point of that operation.
 The state of the library after an execution prefix delimited at the right by a dotted line is pictured in the bottom part (the picture immediately to the left of the dotted line).
-Also, ${\tt i}_1$ denotes the value of {\tt i} in the pop operation with identifier 1.}
+A pair $(d,t)$ for some value $d$ and integer $t$ represents a list node with ${\tt data}=d$ and ${\tt ts}=t$, and ${\tt i}_1$ denotes the value of {\tt i} in the pop operation with identifier 1.
+We omit the nodes where {\tt taken} is set to {\tt true}.}
 \label{fig:commit}
+\vspace{-7mm}
 \end{figure}
@@
 The only reasonable choice for a linearization point is a successful CAS (that results in updating the field {\tt taken}).
+Fig~\ref{fig:commit} pictures an execution showing that this action doesn't correspond to a linearization point, i.e., an execution for which the pop operations in every correct linearization are not ordered according to the order between successful CASs.
+In every correct linearization of that execution, the pop operation removing $x$ is ordered before the one removing $z$ although they perform a successful CAS in the opposite order.
 
-TODO GIVE AN EXAMPLE TO SHOW THE ISSUE WITH POP LINEARIZATION POINTS, AND TO SHOW THAT COMMIT POINTS ARE RIGHT LIMITS FOR THE INTERVAL OF A LIN POINT, I.E., A POP WITH A LATER COMMIT POINT CAN BE LINEARIZED BEFORE.
+An interesting property of the successful CASs in pop operations is that they fix the return value, i.e., the return value is {\tt youngest->data} where {\tt youngest} is the node updated by the CAS.
+We call such actions \emph{commit points}.
+More generally, commit points are actions that access shared variables, from which every control-flow path leads to the return control point and contains no more accesses to the shared memory (i.e., the return value is computed using only local variables).
+
+TODO WHAT CAN WE SAY ABOUT COMMIT POINTS IN OTHER IMPLEMENTATIONS
+
+%Usually, a stack implementation stores the pushed values into a data structure, e.g., a singly-linked list, and a pop operation contains an action (typically, a compare-and-swap) that removes an element from this data structure (or sets a flag associated to this element to a ``deleted'' state).
+Such an action represents a commit point since the pop operation will return the value stored in this element.
+
+%TODO GIVE AN EXAMPLE TO SHOW THE ISSUE WITH POP LINEARIZATION POINTS, AND TO SHOW THAT COMMIT POINTS ARE RIGHT LIMITS FOR THE INTERVAL OF A LIN POINT, I.E., A POP WITH A LATER COMMIT POINT CAN BE LINEARIZED BEFORE.
 NEED ONLY 2 POPS IN AN EXECUTION
 
-We consider a set of actions $Com(pop)=\set{com(pop,d,k):d\in\<Vals>, k\in\<Ops>}$, called \emph{commit points}, representing actions of a pop operation where its return value is determined.
-Usually, a stack implementation stores the pushed values into a data structure, e.g., a singly-linked list, and a pop operation contains an action (typically, a compare-and-swap) that removes an element from this data structure (or sets a flag associated to this element to a ``deleted'' state).
-Such an action represents a commit point since the pop operation will return the value stored in this element.
-
-// Initially each pool[i] points to a 
-// sentinel dummy node with ts = -1, 
-// next = null, taken=false
-
-TODO DESCRIBE THE TIME-STAMPED STACK, EXPLAIN THAT NO METHOD HAS A FIXED LIN POINT, GIVE THE COMMIT POINTS, SHOW THAT COMMIT POINTS ARE NOT LIN POINTS
-
-COMMIT POINTS ARE LINEARIZATION POINTS WHEN THE LIBRARY IS INTERPRETED AS A MULTISET
+When the commit points of pop operations are fixed to particular implementation actions (e.g., a successful CAS) we assume that the library is an LTS over an alphabet that contains actions $com(pop,d,k)$ with $d\in\<Vals>$ and $k\in\<Ops>$.
+This action represents the commit point of the pop with identifier $k$ and returning value $d$.
+Let $Com(pop)$ denote the set of such actions.
+%We consider a set of actions $Com(pop)=\set{com(pop,d,k):d\in\<Vals>, k\in\<Ops>}$, called \emph{commit points}, representing actions of a pop operation where its return value is determined. 
+%TODO DESCRIBE THE TIME-STAMPED STACK, EXPLAIN THAT NO METHOD HAS A FIXED LIN POINT, GIVE THE COMMIT POINTS, SHOW THAT COMMIT POINTS ARE NOT LIN POINTS
+%
+%COMMIT POINTS ARE LINEARIZATION POINTS WHEN THE LIBRARY IS INTERPRETED AS A MULTISET
 
@@
 However, since the commit points are not necessarily linearization points, the pop operations are treated differently.
-Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order and continuously tracks the push operations which are overlapping with it.
-The commit point is enabled only if the return value corresponds to one of the push operations in the initial snapshot, or to a push happening earlier when the values from the initial snapshot are removed by other pop operations, or to one of the push operations that overlaps with the pop.
-TODO EMPTY
+Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order, and continuously tracks the push operations which are overlapping with it.
+The commit point $com(pop,d,k)$ with $d\neq {\tt EMPTY}$ is enabled only if $d$ was added by one of the push operations in the initial snapshot, or by a push happening earlier when all the values from the initial snapshot have been removed by other pops, or by one of the push operations that overlaps with pop $k$.
+The commit point $com(pop,{\tt EMPTY},k)$ is enabled only if all the values added by push operations ending before $k$ started have been removed (checking this condition relies again on the initial snapshot).
+The effect of the commit points is explained below through examples.
 
-\begin{wrapfigure}{l}{7cm}
+\begin{wrapfigure}{l}{6.8cm}
 \vspace{-8mm}
@@
 \end{wrapfigure}
-Figure~\ref{fig:stackSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
-The state of $AbsQ$ after each action is pictured as a graph below the action.
-The nodes of this graph represent enqueue operations and the edges happens-before constraints.
-Each node is labeled by a value (the input of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed.
-For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with input $y$.
-Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.
+Figure~\ref{fig:stackSim} pictures two executions of $AbsS$ for two extended histories (that include pop commit points).
+For readability, we give the state of $AbsS$ only after several execution prefixes delimited at the right by a dotted line.
+We focus on pop operations because the effect of push invocations and returns is similar to enqueue invocations and returns in $AbsQ$.
+Let us first consider the history in the top part of the figure.
+The first $AbsS$ state we give for this history is reached after the call of the pop operation with identifier $3$.
+This shows the effect of a pop invocation: the maximal completed pushes according to the current happens-before (in this case the push with identifier $1$) are marked as $be(3)$ (from ``before'' operation 3), and the pending pushes are marked as $ov(3)$ (from ``overlapping'' with operation 3).
+As a side remark, any other push operation that starts after pop $3$ would be also marked as $ov(3)$.
+The commit point $com(pop,x,3)$ (pictured with a red circle) is enabled because $x$ was added by a push marked as $be(3)$.
+In this case, the effect of the commit point is that the push $1$ is removed from the happens-before (a more complicated case can be seen in the second execution).
+For the second pop, the commit point $com(pop,y,4)$ is enabled because $y$ was added by a push marked as $ov(4)$.
+The second execution shows an example where the marking $be(k)$ for some pop $k$ can be updated at commit points.
+When the two pop operations $5$ and $6$ start, the pushes $3$ and $4$ are marked as $be(5)$ and $be(6)$.
+Then, $com(pop,t,5)$ is enabled because $t$ was added by $push(t,4)$ which is marked as $be(5)$.
+Besides removing $push(t,4)$ from the state, the commit point must produce a state where a pop committing later, e.g., pop $6$, can remove $y$ which was added by a predecessor of $push(t,4)$ in the happens-before ($y$ could become the top of the stack when $t$ is removed).
+This history remains valid because $push(y,2)$ can be linearized after $push(x,1)$ and $push(z,3)$.
+Therefore, push 2, a predecessor of the push which is removed, is marked as $be(6)$.
+The other predecessor of the removed push, i.e., push $1$, is not marked as $be(6)$ because it happens before another push, i.e., push 3, which is already marked as $be(6)$ (the value added by push 3 should be removed as well before the value added by push 1 could become the top of the stack).
 
-TODO AN EXAMPLE FOR EACH CASE (HISTORIES OF THE TIME-STAMPED STACK)
+The states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $<$ is a strict partial order over the set $O$ of operation identifiers, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a pending/completed flag, $rv:\<Ops> ~> \<Vals>$ records the return value of a pending pop fixed at its commit point, $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A_2$) or pop ($R_1,R_2,R_3$) operation, $be:\<Ops> ~> 2^O$ records push operations that were maximal before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
+The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.
 
-The states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked push operations), $<\subseteq O\times O$ is a strict partial order over $O$, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a flag that records whether the operation is pending or completed, $be:\<Ops> ~> 2^O$ records push operations that were maximal before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
-The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.
+The transition rules which don't correspond to commit point actions are similar to those for $AbsQ$.
+The rule {\sc com-pop1} for $com(pop,d,k)$ is enabled only if there exists a push $k'$ which added value $d$ and which belongs to $be(k)$ or $ov(k)$.
+When enabled, the push $k'$ is removed from the set $O$ (and the order $<$) and for every other pop $k_1$ such that $k'$ belongs to $be(k_1)$, $k'$ is replaced in $be(k_1)$ by its predecessors for which no successor belongs already to $be(k_1)$.
+Also, $rv(k)$ is set to $d$.
+The rule {\sc com-pop1} for $com(pop,{\tt EMPTY},k)$ is enabled only if $be(k)$ is empty (i.e., all the values added by pushes ending before $k$, if any, have been removed).
+Then, $rv(k)$ is set to ${\tt EMPTY}$.
 
@@
       k\not\in dom(cp) \\ 
-      d\neq {\tt EMPTY}
+      d\neq {\tt EMPTY} \\
+      \forall k'.\ ov'(k') = ov(k')\cup \{k\}
     }{
       O,<,\ell,rv,cp,be,ov 
-      \xrightarrow{inv(push,d,k)} \\
-      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1],be,ov[\forall k'.\ k'\mapsto ov(k')\cup \{k\}]
+      \xrightarrow{inv(push,d,k)} 
+      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1],be,ov'
     }\hspace{5mm}
@@
       O,<,\ell,rv,cp,be,ov
-      \xrightarrow{inv(pop,k)} \\
-      O,<,\ell,rv,cp[k\mapsto 1],be[k\mapsto maxCo(O)],ov[k\mapsto {\tt PEND}(O)]
+      \xrightarrow{inv(pop,k)} 
+      O,<,\ell,rv,cp[k\mapsto R_1],be[k\mapsto maxCo(O)],ov[k\mapsto {\tt PEND}(O)]
     }\hspace{5mm}
@@
     \inferrule[ret-pop]{
-       cp(k) = 2 \\
+       cp(k) = R_2 \\
        rv(k)=d  
@@
       \xrightarrow{ret(pop,d,k)}
-      O,<,\ell,rv,cp[k\mapsto 3],be,ov
+      O,<,\ell,rv,cp[k\mapsto R_3],be,ov
     }\hspace{5mm}
@@
     \inferrule[ret-push1]{
-      cp(k) = 1 \\
+      cp(k) = A_1 \\
       k \in O \\
@@
       \xrightarrow{ret(push,k)}
-      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto 2],be,ov
+      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2],be,ov
     }\hspace{5mm}
-
     \inferrule[ret-push2]{
-      cp(k) = 1 \\
+      cp(k) = A_1 \\
       k \not\in O 
@@
       \xrightarrow{ret(push,k)}
-      O,<,\ell,rv,cp[k\mapsto 2],be,ov
+      O,<,\ell,rv,cp[k\mapsto A_2],be,ov
     }\hspace{5mm}
 
-    \inferrule[lin-pop1]{
-       cp(k) = 1 \\
+    \inferrule[com-pop1]{
+       cp(k) = R_1 \\
        d\neq{\tt EMPTY} \\
-       k'\in be(k)\cup ov(k)\land \ell_1(k')=d \\
+       k'\in be(k)\cup ov(k) \\
+        \ell_1(k')=d \\
        \forall k_1.\ k'\not\in be(k_1) \Rightarrow be'(k_1)=be(k_1) \\     \\
@@
       \xrightarrow{com(pop,d,k)} \\
-      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto 2],be',ov %be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}]
+      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2],be',ov %be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}]
     }\hspace{5mm}
 
-    \inferrule[lin-pop2]{
-       cp(k) = 1 \\
+    \inferrule[com-pop2]{
+       cp(k) = R_1 \\
        be(k)=\emptyset
@@
       \xrightarrow{com(pop,{\tt EMPTY},k)}
-      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto 2],be,ov
+      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2],be,ov
     }\hspace{5mm}
@@
   \caption{The transition relation of $AbsQ$.
-We use the following notions: $f[\forall k.guard(k)=> k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and $g(k)=expr(k)$ otherwise, $maxCo(O)$ is the set of maximally completed operations in $O$, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.%\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise.
+We use the following notions: $maxCo(O)$ is the set of greatest operations in $O$ (w.r.t.
+$<$) which are completed, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.%\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise.
 If I understand correctly, there is a problem change in $be$ in lin-pop1 rule.
@@
 Lin-pop2 seems ok for me.
-Last note: Bibliography seems not correct to me.}
+Last note: Bibliography seems not correct to me.}  $f[\forall k.\ k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=expr(k)$,   $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and
   }
@@
 
-Let $AbsS_0$ be the standard abstract implementation of a concurrent stack (where elements are stored in a sequence, a push operation adds atomically an element at the beginning of the sequence, and a pop operation removes atomically an element from the beginning of the sequence)~\footnote{For $\<Methods>=\{push,pop\}$, the alphabet of $AbsS_0$ is $C\cup R\cup Lin$.}.
+Let $AbsS_0$ be the standard abstract implementation of a concurrent stack (where elements are stored in a sequence, a push operation adds atomically an element at the beginning of the sequence, and a pop operation removes atomically an element from the beginning of the sequence).
+For $\<Methods>=\{push,pop\}$, the alphabet of $AbsS_0$ is $C\cup R\cup Lin$.
 The following result states that the library $AbsS$ has exactly the same set of histories as $AbsS_0$ (see Appendix~\ref{app:absImplStack} for a proof).

### 2017-01-22 draft.tex
TS Stack fs relation and flow diagram

--- 2017-01-10 draft.tex
@@
 \input{queue}
+
 \section{Existence of Forward Simulations for Stack Implementations that have Fixed Pop Linearization Points}
 \input{commit}
-\input{set}
+First, we describe the TS-Stack algorithm:
+\begin{lstlisting}
+struct Node{
+  int data;
+  int ts;
+  Node* next;
+  boolean taken;
+};
+Node* pools[maxThreads];
+int TS = 0;
+
+void push(int x) {
+  Node* n = new Node(x,MAX_INT,
+                        null,false);
+  n->next = pools[myTID];
+  pools[myTID] = n;
+  int i = TS++;
+  n->ts = i;
+}
+int pop() {
+ boolean success = false;
+ int maxTS = -1;
+ Node* youngest = null;
+ while ( !success ) {
+   maxTS = -1; youngest = null;
+   for(int i=0; i<maxThreads; i++){
+     Node* n = pools[i];
+     while (n->taken && n->next != n)
+       n = n->next;
+     if(maxTS < n->ts) {
+       maxTS = n->ts; youngest = n;
+     }
+   }
+   if (youngest != null)
+     success=CAS(youngest->taken,
+                       false,true);
+ }
+ return youngest->data;
+}
+\end{lstlisting}
+Then, we begin to obtain the LTS ($L_C$) of TS Stack from the algorithm by defining control points and the actions among these control points as seen in Figure X \textcolor{red}{Flow diagram is referenced here}.
+To simplify the proof, we take the initializations of some local variables together as atomic.
+
+States of the TS-Stack contains the global variables and local variables as fields.
+Global variables are just elements of their domains and local variables are maps from operation identifiers to their domains.
+We say $i_q(k)$ for referencing the value of local variable $i$ of operation $k$ in state $q$.
+There is only one special local variable called $myTID$.
+Its value is unique to each pending operation in a state i.e., concurrent operations cannot have the same $myTID$ value.
+TS-Stack states also contains sets $O_a, O_r \in \mathbb{O}$ which are operation identifier sets of push and pops respectively, and the control point function $cp$ which is a map from operation identifiers to the control points set that are presented in the flow diagram Figure X  \textcolor{red}{Flow diagram is referenced here}.
+Transition relation of the TS-STack is presented in Figure~\ref{fig:transitions:TSS}.
+Next, we show that the linearizability of TS Stack.
+
+\begin{figure} [t]
+{\scriptsize
+  \centering
+  \begin{mathpar}
+    \inferrule[call-push]{
+      k\not\in dom(cp) \\ 
+      d\neq {\tt EMPTY} \\
+      \forall k'.\ ov'(k') = ov(k')\cup \{k\}
+    }{
+      O,<,\ell,rv,cp,be,ov 
+      \xrightarrow{inv(push,d,k)} 
+      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1],be,ov'
+    }\hspace{5mm}
+
+      \end{mathpar}
+  }
+ \vspace{-5mm}
+  \caption{The transition relation of TS-Stack
+  }
+  \label{fig:transitions:TSS}
+\vspace{-6mm}
+\end{figure}
+
+\begin{lem}
+$L_C$ is a Com(pop)- refinement of $AbsS$.
+\end{lem}
+\begin{proof}
+Since $AbsS$ is deterministic with respect to $C \cup R \cup Com(pop)$, there exists a Com(pop) forward simulation from $L_C$ to $AbsS$ iff $L_C$ is a Com(pop)-refinement of $AbsS$.
+Hence, we define the relation $fs$ and show that it is a Com(pop)-forward simulation relation.
+
+Let us make some clarifications before defining the relation.
+In order not to confuse nodes in TS Stack and nodes in $AbsS$, we call nodes of $AbsS$ as vertices from now on.
+We also define ordering relation (called traverse order) among the operations in a state of $TS$.
+It basically reflects the traverse order of pop operations.
+For two push operations $m,n \in O_a$ is state $s$ we say that $m <^{tr}_s n$ iff either $myTid(m) < myTid(n)$ or $myTid(m) = myTid(n)$ and $n_s(n)$ is reachable from $n_s(m)$ using next pointers.
+$\geq^{tr}$ is obtained from $<^{tr}$ in the usual way.
+
+The relation $fs \subseteq Q_C \rightarrow Q_{AbsS}$ contains $(s,t)$ iff the following are satisfied:
+\begin{itemize}
+\item[\emph{Nodes}] $k \in O_t$ iff $k$ is a push operation in $s$ ($k \in O_a$) such that either it has not inserted its node to the pool yet ( $cp_s(k) = A_i$ and $i<3$) or its node is not taken by a pop ($cp_s(k) = A_i$, $i\geq 3$ and $n_s(k)->taken = false$).
+\item[\emph{Pend/Comp}] A vertex $k \in O_t$ is pending ($\ell_t(k) = (d, \texttt{PEND})$) iff $k$ satisfies the previous condition, $x_s(k) = d$ and it is not completed in $s$ ($cp_s(k) = A_i$ and $i<6$).
+Similarly, this vertex is completed ($\ell_t(k) = (d, \texttt{COMP})$) iff $k$ satisfies the previous condition, $x_s(k) = d$ and it is completed in $s$ ($cp_s(k) = A_6$).
+\item[\emph{TSOrder}] If a node has a smaller timestamp than the other node in $s$, the operations that inserted them cannot be ordered reversely in $t$.
+More formally, let $k, k' \in O_t$ s.t.
+$n_s(k)-> ts < n_s(k')->ts$.
+Then, $k \nless_t k'$.
+\item[\emph{TidOrder}] Order among the nodes inserted by the same threads in $s$ must be preserved among the operations that inserted them in $t$.
+Let $k, k' \in O_t$ s.t.
+$myTid_s(k) = myTid_s(k')$ and $n_s(k)->ts < n_s(k')->ts$.
+Then, $k <_t k'$.
+\item[Frontiers] Every maximally closed or pending vertex can be removed by a pending pop.
+More formally, let $k \in O_t$ such that $\ell_t(k) = (\_,\texttt{PEND})$.
+Then, for all pops $p$, $k \in ov_t(p)$.
+In the other case, let $k \in O_t$ such that $\ell_t(k) = (\_,\texttt{COMP})$ and for all other $k' \in O_t$ such that $k<_t k'$, we know $\ell_t(k') = (\_,\texttt{PEND})$.
+Then, for all pop operations $p$, $k \in be_t(p)$ or $k \in ov_t(p)$.
+\item[\emph{MaximalOV}] If a push $k \in O_t$ is a candidate to be removed by a pop $p$, then every other push $k'$ invoked after $k$ is a candidate to be removed by $p$ since $k$ is concurrent with $p$.
+More formally, let $k, k' \in O_t$ such that $k <_t k'$ and there exists a pop $p$ such that $k \in be_t(p)$ or $k \in ov_t(p)$.
+Then, $k' \in ov_t(p)$.
+\item[\emph{MinimalBE}] If a push $k \in O_t$ has finished before the pop $p$ is invoked and yet $k$ is a candidate to be removed by $p$, other pushes completed before $k$ can not be candidates to be removed by $p$ at that state.
+More formally, let $k, k' \in O_t$ such that $k <_t k'$ and there exists a pop $p$ such that $k' \in be_t(p)$.
+Then, neither $k \in be_t(p)$ nor $k \in ov_t(b)$.
+\item[\emph{ReverseFrontiers}] If a push $k \in O_t$ is concurrent with the pop $p$ and there exists another push $k' \in O_t$ that is immediate predecessor of $k$, then $k'$ is either concurrent or maximally closed with respect to $p$.
+More formally, let $k, k' \in O_t$ such that $k' \in pred_{<_t}(k)$ and $k \in ov_t(p)$ for some pop $p$.
+Then, either $k' \in ov_t(p)$ or $k' \in be_t(p)$.
+\item[\emph{TraverseBefore}] If a pop operation $p$ is currently visiting node $n$ and there is a node $m$ coming before $n$ in the traverse order with a greater timestamp, then the operation that inserts $m$ must be concurrent with $p$.
+More formally, let $k \in O_t$ such that $n_s(k) <^{tr}_s n_s(p)$ and $n_s(k)->ts > n_s(p)->ts$.
+Then, $k \in ov_t(p)$.
+\item[\emph{TraverseAfter}] If a pop operation $p$ is currently visiting node $n$ and its youngest element $m$ is still not taken in state $s$, then either $m$ is a candidate to be removed by $p$ in $t$ or there exists a later node $m'$ than $n$ such that $m'$ is a candidate in $t$ it has a bigger timestamp than n.
+More formally, let $k, k' \in O_t$ such that $youngest_s(p) = n_s(k)$ and $n_s(k') = n_s(p)$.
+Then, either $k \in ov_t(p) \vee k \in be_t(p)$ or there exists $k'' \in O_t$ s.t.
+$n_s(k'')->ts > n_s(k) ->ts$ and $k'' \in ov_t(p) \vee k'' \in be_t(p)$ and either $k' <^{tr}_s k''$ or $n_s(p) = n_s(k'') \wedge cp_s(p) = R_j \wedge j>3$.
+\end{itemize}
+\end{proof}
+
+%\input{set}
 \end{document}

### 2017-01-22 draft.tex
Refined TSS fs relation definition

@@
 Similarly, this vertex is completed ($\ell_t(k) = (d, \texttt{COMP})$) iff $k$ satisfies the previous condition, $x_s(k) = d$ and it is completed in $s$ ($cp_s(k) = A_6$).
+Pending vertices are maximal with respect to $<_t$ i.e., if $k \in O_t$ is a pending vertex, then for all $k' \in O_t$ $k \nless_t k'$.
 \item[\emph{TSOrder}] If a node has a smaller timestamp than the other node in $s$, the operations that inserted them cannot be ordered reversely in $t$.
 More formally, let $k, k' \in O_t$ s.t.
-$n_s(k)-> ts < n_s(k')->ts$.
-Then, $k \nless_t k'$.
+$n_s(k)-> ts \leq n_s(k')->ts$.
+Then, $k' \nless_t k$.
 \item[\emph{TidOrder}] Order among the nodes inserted by the same threads in $s$ must be preserved among the operations that inserted them in $t$.
@@
 Then, $k \in ov_t(p)$.
-\item[\emph{TraverseAfter}] If a pop operation $p$ is currently visiting node $n$ and its youngest element $m$ is still not taken in state $s$, then either $m$ is a candidate to be removed by $p$ in $t$ or there exists a later node $m'$ than $n$ such that $m'$ is a candidate in $t$ it has a bigger timestamp than n.
-More formally, let $k, k' \in O_t$ such that $youngest_s(p) = n_s(k)$ and $n_s(k') = n_s(p)$.
+\item[\emph{TraverseAfter}] If a pop operation $p$ is currently visiting node $n$ that is not null and its youngest element $m$ is not null and still not taken in state $s$, then either $m$ is a candidate to be removed by $p$ in $t$ or there exists a later node $m'$ than $n$ such that $m'$ is a candidate in $t$ it has a bigger timestamp than n.
+More formally, assume that there exists $k, k' \in O_t$ such that $youngest_s(p) = n_s(k)$ and $n_s(k') = n_s(p)$.
 Then, either $k \in ov_t(p) \vee k \in be_t(p)$ or there exists $k'' \in O_t$ s.t.

### 2017-01-22 draftCAV/queue.tex
stack simulation

@@
 
-TODO NEEDS DATA INDEPENDENCE FOR THE LINEARIZATION POINT TRANSITIONS TO BE DETERMINISTIC
+%TODO NEEDS DATA INDEPENDENCE FOR THE LINEARIZATION POINT TRANSITIONS TO BE DETERMINISTIC
 
@@
 
-We describe a forward simulation $\mathit{fs}$ from $\mathit{HWQ}$ to $AbsQ$.
-Essentially, an $AbsQ$ state associated by $\mathit{fs}$ to a $\mathit{HWQ}$ state consists of all the  enqueue operations for which the input is still present in the array {\tt items} and all the pending enqueue operations that have at most reserved an array position, ordered by a relation $<$ satisfying the following:
+We describe a forward simulation $\mathit{fs}_1$ from $\mathit{HWQ}$ to $AbsQ$.
+Essentially, an $AbsQ$ state associated by $\mathit{fs}_1$ to a $\mathit{HWQ}$ state consists of all the  enqueue operations for which the input is still present in the array {\tt items} and all the pending enqueue operations that have at most reserved an array position, ordered by a relation $<$ satisfying the following:
 \begin{itemize}
@@
 
-We show that $\mathit{fs}$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in\mathit{fs}$. 
+We show that $\mathit{fs}_1$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in\mathit{fs}_1$. 
 We omit discussing the trivial case of transitions labeled by call and return actions which are simulated by similar transitions of $AbsQ$ (for the return a dequeue operation $k$, we use the equality between the local variable ${\tt x}(k)$ in $s$ and the component $rv(k)$ in $t$).
@@
 
-We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an \emph{empty} sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in\mathit{fs}$, then $(s',t)\in\mathit{fs}$ for each $AbsQ$ state $t$.
-We focus on the following essential property, called \emph{monotonicity}: the set of possible orders $<$ associated by $\mathit{fs}$ to $s'$ doesn't exclude any order $<$ associated to $s$.
+We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an \emph{empty} sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in\mathit{fs}_1$, then $(s',t)\in\mathit{fs}_1$ for each $AbsQ$ state $t$.
+We focus on the following essential property, called \emph{monotonicity}: the set of possible orders $<$ associated by $\mathit{fs}_1$ to $s'$ doesn't exclude any order $<$ associated to $s$.
 %Essentially, this boils down to showing that the constraints over $<$ in the definition of $\mathit{fs}$ are an invariant for these steps.
@@
 The execution of {\tt swap} returning {\tt null} in a dequeue $k_d$ enables new instantiations $(k,k',k_d)$ in $s'$, thus adding potentially new constraints $k\not< k'$.
-We show however that these instantiations are vacuous because $k$ must be a pending enqueue in $s$ and thus maximal in any order relation $<$ associated by $\mathit{fs}$ to $s$.
+We show however that these instantiations are vacuous because $k$ must be a pending enqueue in $s$ and thus maximal in any order relation $<$ associated by $\mathit{fs}_1$ to $s$.
 Let $k$ and $k'$ be two enqueue operations such that together with the dequeue $k_d$ they satisfy the property (\ref{eq:inst}) in $s'$ but not in $s$. 

### 2017-01-22 draftCAV/stack.tex
stack simulation

@@
 \end{wrapfigure}
-We explain the meaning of the commit points on a simplified version of the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} given in Figure~\ref{fig:TimeStamped}.
+We explain the meaning of the commit points on a simplified version of the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) given in Figure~\ref{fig:TimeStamped}.
 This stack implementation maintains a vector of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
@@
 More generally, commit points are actions that access shared variables, from which every control-flow path leads to the return control point and contains no more accesses to the shared memory (i.e., the return value is computed using only local variables).
+
+The complete version of $\mathit{TSS}$~\cite{DBLP:conf/popl/DoddsHK15} contains also a mechanism for elimination (a pop can remove a value without traversing all the lists if it has been added by a concurrent push) and emptiness checking.
+In both cases, the commit points can be identified in the code and correspond to certain boolean conditions evaluated to {\tt true}.
 
@@
 
-TODO NEEDS DATA INDEPENDENCE FOR THE COMMIT POINT TRANSITIONS TO BE DETERMINISTIC
+%TODO NEEDS DATA INDEPENDENCE FOR THE COMMIT POINT TRANSITIONS TO BE DETERMINISTIC
 
@@
 \subsection{A Correctness Proof For Time-Stamped Stack}
+
+We describe a forward simulation $\mathit{fs}_2$ from $\mathit{TSS}$ to $AbsS$.
+Except for the constraints on the components $be$ and $ov$ of a $AbsS$ state, it is similar to the simulation $\mathit{fs}_1$ from $\mathit{HWQ}$ to $AbsQ$.
+Thus, the $AbsS$ states $t=\tup{O,<,\ell,rv,cp,be,ov}$ associated by $\mathit{fs}_2$ to a $\mathit{TSS}$ state $s$ satisfy the following.
+The set $O$ consists of all the identifiers of pushes in $s$ which didn't added yet a node to {\tt pools} or for which the input is still present in {\tt pools} (i.e., the node created by the push has {\tt taken} set to {\tt false}).
+A push $k$ is labeled by $(d,{\tt PEND})$ where $d$ is the input value if it's pending and by $(d,{\tt COMP})$, otherwise.
+
+To describe the order relation $<$ we consider the following notations: ${\tt ts}_s(k)$, resp., ${\tt TID}_s(k)$, denotes the timestamp of the node created by the push $k$ in state $s$ (the {\tt ts} field of this node), resp., the id of the thread executing $k$.
+By an abuse of terminology, we call ${\tt ts}_s(k)$ the timestamp of $k$ in state $s$.
+Also, $k \leadsto_s k'$ when intuitively, a traversal of {\tt pools}  would encounter the node created by $k$ before the one created by $k'$.
+More precisely, $k \leadsto_s k'$ when ${\tt TID}_s(k) < {\tt TID}_s(k')$, or ${\tt TID}_s(k) = {\tt TID}_s(k')$ and the node created by $k'$ is reachable from the one created by $k$ in the list pointed to by ${\tt pools}[{\tt TID}_s(k)]$.
+The order relation $<$ satisfies the following: (1) pending pushes are maximal, (2) $<$ is consistent with the order between node timestamps, i.e., ${\tt ts}_s(k) \leq {\tt ts}_s(k')$ implies $k'\not< k$, and (3) $<$ includes the order between pushes executed in the same thread, i.e., ${\tt TID}_s(k) = {\tt TID}_s(k')$ and ${\tt ts}_s(k) < {\tt ts}_s(k')$ implies $k < k'$.
+
+The components $be$ and $ov$ satisfy the following constraints (their domain is the set of identifiers of pending pops):
+\begin{itemize}
+	\item every pop overlaps with all pending pushes, i.e., $k_1\in O$ is pending implies $k_1\in ov(k)$ for each $k, k_1$
+	\item completed pushes which are maximal in $<$ are either overlapping with a pop $k$ or they were maximal in $<$ when $k$ started, i.e., every such push belongs either to $be(k)$ or $ov(k)$ for each $k$
+	\item for every push that overlaps with a pop $k$ or was maximal in $<$ when $k$ started, its successors are overlapping with $k$, i.e., $k_1\in be(k)\cup ov(k)$ and $k_1 < k_2$ implies $k_2 \in ov(k)$ for each $k, k_1, k_2$
+	\item $be(k)$ and $ov(k)$ don't contain predecessors of pushes from $be(k)$, i.e., $k_1 < k_2$ and $k_2 \in be(k)$ implies $k_1\not\in be(k)\cup ov(k)$ for each $k, k_1, k_2$
+	\item immediate predecessors of pushes overlapping with a given pop $k$ are either overlapping with $k$ or they were maximal in $<$ when $k$ started, i.e., $k_2\in pred_{<}(k_1)$ and $k_1\in ov(k)$ implies $k_2\in be(k)\cup ov(k)$ for each $k,k_1,k_2$
+	\item a pop $k$ that reached a node with timestamp $\tau$ (its variable {\tt n} points to this node) overlaps with any push that has a timestamp bigger than $\tau$ and which created a node that occurs in {\tt pools} before the node reached by $k$, i.e., ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k_1)$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
+	\item if the variable {\tt youngest} of a pop $k$ points to a node which is not taken, then this node was created by a push in $be(k)\cup ov(k)$ or the node currently reached by $k$ is followed in {\tt pools} by another node which was created by a push in $be(k)\cup ov(k)$, i.e., ${\tt youngest}_s(k)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(k)={\tt n}_s(k_2)$ implies $k_1\in be(k)\cup ov(k)$ or that there exists $k_3\in O$ such that ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in be(k)\cup ov(k)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and $k$ is traversing the last list in the array {\tt pools}, for each $k, k_1,k_2$.
+\end{itemize}
+Finally, for every pop operation $k$ such that ${\tt success}(k)={\tt true}$, we have that $rv(k)={\tt youngest}(k)\text{\tt->data}$.
+
+The proof that $\mathit{fs}_2$ is indeed a forward simulation from $\mathit{TSS}$ to $AbsS$ follows the same lines as the one given for the Herlihy\&Wing Queue. It can be found in Appendix~\ref{}.

### 2017-01-23 draftCAV/prelim.tex
shrinking preliminaries

--- 2017-01-16 draftCAV/prelim.tex
@@
 parameter/return values. 
-
+%
 %\begin{example}
@@
 %\end{example}
-
-\noindent
+%
 We fix an arbitrary set $\<Ops>$ of operation identifiers, and for given sets
@@
 \emph{operation identifier} $k \in \<Ops>$. Operation identifiers are used to
-pair call and return actions. We denote the operation identifier of a
-call/return action $a$ by $\<op>(a)$. Call and return actions $c \in C$ and $r
-\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
-We may omit the second field from a call/return action $a$ for methods that have no inputs (e.g., the pop method of a stack) or return values (e.g., the push method of a stack).
-A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
-\emph{well formed} when:
-\begin{itemize}
-
-  \item Each return is preceded by a matching call: \\
-  $\tau_j \in R$ implies $\tau_i \match \tau_j$ for some $i < j$.
-
-  \item Each operation identifier is used in at most one call/return: \\
-  $\<op>(\tau_i) = \<op>(\tau_j)$ and $i < j$ implies $\tau_i \match \tau_j$.
-
-\end{itemize}
-We say that the well-formed word $\tau \in @S^*$ is \emph{sequential} when
-\begin{itemize}
-
-  \item Operations do not overlap: \\
-  $\tau_i, \tau_k \in C$ and $i < k$ implies $\tau_i \match \tau_j$ for some $i < j < k$.
-
-\end{itemize}
-Well-formed words represent traces of a library. We assume every set of well-formed
+pair call and return actions. 
+%We denote the operation identifier of a
+%call/return action $a$ by $\<op>(a)$. Call and return actions $c \in C$ and $r
+%\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
+We may omit the second field from a call/return action $a$ for methods that have no inputs or return values.
+%A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
+%\emph{well formed} when:
+%\begin{itemize}
+%
+%  \item Each return is preceded by a matching call: \\
+%  $\tau_j \in R$ implies $\tau_i \match \tau_j$ for some $i < j$.
+%
+%  \item Each operation identifier is used in at most one call/return: \\
+%  $\<op>(\tau_i) = \<op>(\tau_j)$ and $i < j$ implies $\tau_i \match \tau_j$.
+%
+%\end{itemize}
+%We say that the well-formed word $\tau \in @S^*$ is \emph{sequential} when
+%\begin{itemize}
+%
+%  \item Operations do not overlap: \\
+%  $\tau_i, \tau_k \in C$ and $i < k$ implies $\tau_i \match \tau_j$ for some $i < j < k$.
+%
+%\end{itemize}
+%Well-formed words represent traces of a library. 
+We assume every set of 
 words is closed under isomorphic renaming of operation identifiers. For
 notational convenience, we take $\<Ops>=\<Nats>$ for the rest of the paper.
-When the value of a certain field in a call/return action is not important we use 
-the placeholder $\_$, e.g., $inv(m,\_,k)$ instead of $inv(m,d,k)$ when the input  
-$d$ can take any value.
+%When the value of a certain field in a call/return action is not important we use 
+%the placeholder $\_$, e.g., $inv(m,\_,k)$ instead of $inv(m,d,k)$ when the input  
+%$d$ can take any value.
 
-An operation $k$ is called \emph{completed} in a well-formed trace $\tau$ when
-$ret(m,d,k)$ occurs in $\tau$, for some $m$ and $d$. Otherwise, it is called \emph{pending}.
 %An operation $o$ of an execution $e$ is \emph{completed}
@@
 
-Libraries dictate the execution of methods between their call and return
-points. Accordingly, a library cannot prevent a method from being called,
-though it can decide not to return. Furthermore, any library action performed
-in the interval between call and return points can also be performed should the
-call have been made earlier, and/or the return made later. 
-A library thus allows any sequence of
-invocations to its methods made by \emph{some} program.
+A \emph{library} is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$. 
+We assume that the traces of a library satisfy standard well-formedness properties, 
+e.g., return actions correspond to previous call actions, which for lack of 
+space are delegated to Appendix~\ref{app:prelim}. An operation $k$ is called \emph{completed} in a trace $\tau$ when
+$ret(m,d,k)$ occurs in $\tau$, for some $m$ and $d$. Otherwise, it is called \emph{pending}.
 
-\begin{definition}\label{def:libraries}
-A \emph{library} $L$ is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$
-and each trace $\tau \in Tr(L)$ is well formed, and
-  \begin{itemize}
-
-    \item Call actions $c \in C$ cannot be disabled: \\
-    $\tau \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot \tau' \in Tr(L)$
-    if $\tau \cdot c \cdot \tau'$ is well formed.
-
-    \item Call actions $c \in C$ cannot disable other actions: \\
-    $\tau \cdot a \cdot c \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot a \cdot \tau' \in Tr(L)$.
-
-    \item Return actions $r \in R$ cannot enable other actions: \\
-    $\tau \cdot r \cdot a \cdot \tau' \in Tr(L)$ implies $\tau \cdot a \cdot r \cdot \tau' \in Tr(L)$.
-
-  \end{itemize}
-
-\end{definition}
-
-The projection of a library trace over $C\cup R$ is called a \emph{history}. The set of histories of a library $L$ is denoted by $H(L)$.
-
-Note that even a library that implements \emph{atomic methods}, e.g.,~by
-guarding method bodies with a global-lock acquisition, admits executions in
-which method calls and returns overlap. 
-%A library which accesses the client's
-%thread identifiers can be modeled by taking thread identifiers as method
-%parameters.
+%Libraries dictate the execution of methods between their call and return
+%points. Accordingly, a library cannot prevent a method from being called,
+%though it can decide not to return. Furthermore, any library action performed
+%in the interval between call and return points can also be performed should the
+%call have been made earlier, and/or the return made later. 
+%A library thus allows any sequence of
+%invocations to its methods made by \emph{some} program.
+%
+%\begin{definition}\label{def:libraries}
+%A \emph{library} $L$ is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$
+%and each trace $\tau \in Tr(L)$ is well formed, and
+%  \begin{itemize}
+%
+%    \item Call actions $c \in C$ cannot be disabled: \\
+%    $\tau \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot \tau' \in Tr(L)$
+%    if $\tau \cdot c \cdot \tau'$ is well formed.
+%  
+%    \item Call actions $c \in C$ cannot disable other actions: \\
+%    $\tau \cdot a \cdot c \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot a \cdot \tau' \in Tr(L)$.
+%  
+%    \item Return actions $r \in R$ cannot enable other actions: \\
+%    $\tau \cdot r \cdot a \cdot \tau' \in Tr(L)$ implies $\tau \cdot a \cdot r \cdot \tau' \in Tr(L)$.
+%  
+%  \end{itemize}
+%
+%\end{definition}
+%
+%%Note that even a library that implements \emph{atomic methods}, e.g.,~by
+%%guarding method bodies with a global-lock acquisition, admits executions in
+%%which method calls and returns overlap. 
+%For simplicity, Definition~\ref{def:libraries} assumes that every thread performs a single operation.
+The extension to multiple operations per thread is straightforward, e.g.
+the closure rules must assume that the actions $a$ and $c$ belong to different threads
+%%A library which accesses the client's
+%%thread identifiers can be modeled by taking thread identifiers as method
+%%parameters.
 
@@
 Ignore this if I am wrong.}
+The projection of a library trace over $C\cup R$ is called a \emph{history}. The set of histories of a library $L$ is denoted by $H(L)$.
 Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
@@
 obtained from $h_1$ by appending return actions, and deleting call actions,
-such that:
+s.t.:
 \begin{quote}
@@
 \end{quote}
-A call action $inv(\_,\_,k)$ may happen before a return action $ret(\_,\_,k')$ in $h_1$ (the operations $k$ and $k'$ overlapping in time) while it can be permuted after that return action in $h_2$, thus introducing a new inter-operation ordering (operation $k'$ ending before $k$).
+%A call action $inv(\_,\_,k)$ may happen before a return action $ret(\_,\_,k')$ in $h_1$ (the operations $k$ and $k'$ overlapping in time) while it can be permuted after that return action in $h_2$, thus introducing a new inter-operation ordering (operation $k'$ ending before $k$).
 
-$\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
-history set $H$ under weakening.
-A direct consequence of Definition~\ref{def:libraries} is that $H(L)=\overline{H(L)}$ for
-every library $L$. A \emph{kernel} of L is any set $K$ such that $\overline{K} = H(L)$.
-A library $L$ is called atomic if it has a sequential kernel. 
+%$\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
+%history set $H$ under weakening.
+%We assume that $H(L)=\overline{H(L)}$ for
+%every library $L$. %A \emph{kernel} of L is any set $K$ such that $\overline{K} = H(L)$.
+A library $L$ is called \emph{atomic} when there exists a set $S$ of sequential histories such that 
+$H(L)$ contains every weakening of a history in $S$.
 Atomic libraries are often considered as specifications for concurrent objects. 
-In practice, libraries can be made atomic by guarding their methods bodies with global lock acquisitions.
+Libraries can be made atomic by guarding their methods bodies with global lock acquisitions.
 
-A library $L$ is called a \emph{queue implementation} when its alphabet contains call and return actions defined over $\<Methods>=\{enq,deq\}$ ($enq$ is the name of the methods that enqueue a value, and $deq$ is the name of the methods that dequeue a value) and $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$ where {\tt EMPTY} is the value returned by $deq$ when the queue is empty.
+A library $L$ is called a \emph{queue implementation} when $\<Methods>=\{enq,deq\}$ ($enq$ is the method that enqueues a value and $deq$ is the method removing a value) and $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$ where {\tt EMPTY} is the value returned by $deq$ when the queue is empty.
 Similarly, a library $L$ is called a \emph{stack implementation} when $\<Methods>=\{push,pop\}$ and  $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$.
+For queue and stack implementations, we assume that the same value is never added twice, i.e., for every trace $@t$ of such a library and every two call actions $inv(m,d_1,k_1)$ and $inv(m,d_2,k_2)$ where $m\in \{enq,push\}$ we have that $d_1\neq d_2$.
+As shown in several works~\cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, this assumption is without loss of generality for libraries that are data independent, i.e., their behaviors are not influenced by the values added to the collection, which is always the case in practice.
+On a technical note, this assumption is used to define ($\Gamma$-)deterministic abstract implementations of stacks and queues in Section~\ref{sec:queues} and Section~\ref{sec:stacks}.
 
@@
 
-\begin{dfn}
-Let $L_1$ and $L_2$ be two libraries. 
+%\textcolor{red}{We have two Definition 1. The problem occurs because we used different commands for definitions.}
+\begin{definition}
+%Let $L_1$ and $L_2$ be two libraries. 
 %agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets.
@@
 Note that $A\Sigma \subseteq \Sigma_{L_1}$ and $A\Sigma \subseteq \Sigma_{L_2}$.
-The library $L_1$ \emph{refines} $L_2$ if{f} $H(L_1) \subseteq H(L_2)$.
+A library $L_1$ \emph{refines} another library $L_2$ if{f} $H(L_1) \subseteq H(L_2)$.
 %We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
 %for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
-\end{dfn}
+\end{definition}
 

### 2017-01-23 draftCAV/refProofs.tex
small modifs
started stacks
shrinking preliminaries

--- 2017-01-16 draftCAV/refProofs.tex
@@
 We focus on proofs based on forward simulations because they are easier to automatize.
+
+%TODO SAY THAT THE EXTENSION TO $\Gamma$-determinism is new (the previous results were about determinism).
 
@@
 
-The following lemma shows the soundness and the completeness of $\Gamma$-forward simulations.
+The following theorem shows the soundness and the completeness of $\Gamma$-forward simulations.
+It is an instantiation of previous results~~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}.
 \begin{theorem}\label{th:forSim}
-Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$.
+Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$.
 Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
-If there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$.
-Moreover, if $L_1$ $\Gamma$-refines $L_2$ and $L_2$ is $\Gamma$-deterministic, then there exists a $\Gamma$-forward simulation from $L_1$ to $L_2$.
+If there is a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$.
+Moreover, if $L_1$ $\Gamma$-refines $L_2$ and $L_2$ is $\Gamma$-deterministic, then there is a $\Gamma$-forward simulation from $L_1$ to $L_2$.
 \end{theorem}
-\begin{proof}
-TODO
-\end{proof}
 
@@
 \emph{extended history}. A trace or extended history is called \emph{$Lin$-complete} when every completed operation has a linearization 
-point, i.e., TODO. A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup Lin\subseteq \Sigma$ 
+point, i.e., TODO. 
+A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup Lin\subseteq \Sigma$ 
 and every trace $@t\in Tr(L)$ is $Lin$-complete.

### 2017-01-23 draftCAV/prelim.tex
dealing with simulations

@@
 Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
-admit histories with weaker inter-operation ordering, in which calls may happen earlier, and/or returns later. This weakening relation
-between histories can be formalized as follows: a history $h_1$ is \emph{weaker} than a history $h_2$, written $h_1 \sqsubseteq h_2$, 
-if{f} there exists a well-formed history $h_1'$
+admit histories with weaker inter-operation ordering, in which calls may happen earlier, and/or returns later. A 
+history $h_1$ is \emph{weaker} than a history $h_2$, written $h_1 \sqsubseteq h_2$, 
+if{f} there exists a history $h_1'$
 obtained from $h_1$ by appending return actions, and deleting call actions,
 s.t.:
-\begin{quote}
-
   $h_2$ is a permutation of $h_1'$ that preserves the order between
@@
   call action in $h_1'$, then the same holds in $h_2$.
-
-\end{quote}
 %A call action $inv(\_,\_,k)$ may happen before a return action $ret(\_,\_,k')$ in $h_1$ (the operations $k$ and $k'$ overlapping in time) while it can be permuted after that return action in $h_2$, thus introducing a new inter-operation ordering (operation $k'$ ending before $k$).
@@
 interact with the library only through call and return actions, and thus refinement can be defined
-as history set inclusion. As shown in~\citet{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15},
-refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} when $L_2$ is an atomic library. 
+as history set inclusion. Refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} 
+when $L_2$ is an atomic library~\cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}.
+. 
 %We fix a set of methods $\<Methods>$ and values $\<Vals>$.

### 2017-01-23 draftCAV/refProofs.tex
dealing with simulations

@@
 \begin{itemize}
-\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$ 
-\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0<i<|@s|$.
-%$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $.
-The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
-\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0\leq i<|@s| -1$.
-% where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
-\item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $\delta_2(u,\gamma, u')$. 
-\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma_1\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0\leq i<|@s|$.
+\item $\mathit{fs}[s_0^1] = \{s_0^2 \}$.
+\item If $(s,\gamma,s') \in \delta_1$, for some $\gamma\in \Gamma$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_i=\gamma$, for some $i$, and $@s_j\in \Sigma_2\setminus\Gamma$, for each $j\neq i$.
+\item If $(s,e,s') \in \delta_1$, for some $e \in \Sigma_1\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $\sigma\in (\Sigma_2\setminus\Gamma)^*$.
 %where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
@@
 %If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
-A $\Gamma$-forward simulation shows that every step of $L_1$ corresponds to a (possibly empty) sequence of steps of $L_2$.
-In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action.
-The rest of the transitions in $L_1$ are matched to a possibly empty sequence of transitions of $L_2$ with arbitrary labels.
+A $\Gamma$-forward simulation requires that every step of $L_1$ corresponds to a sequence of steps of $L_2$.
+To imply $\Gamma$-refinement, every step of $L_1$ labeled by an observable action $\gamma\in \Gamma$ should be simulated by a sequence of steps of $L_2$ where exactly one transition is labeled by $\gamma$ and all the other transitions are labeled by non-observable actions.
 
-The following theorem shows the soundness and the completeness of $\Gamma$-forward simulations.
+The following theorem shows the soundness and the completeness of $\Gamma$-forward simulations (when $L_2$ is $\Gamma$-deterministic).
 It is an instantiation of previous results~~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}.

### 2017-01-23 draft.tex
TS Stack linearizability proof, push cases finished

@@
 \end{lstlisting}
-Then, we begin to obtain the LTS ($L_C$) of TS Stack from the algorithm by defining control points and the actions among these control points as seen in Figure X \textcolor{red}{Flow diagram is referenced here}.
+Then, we begin to obtain the LTS ($TSS$) of TS Stack from the algorithm by defining control points and the actions among these control points as seen in Figure X \textcolor{red}{Flow diagram is referenced here}.
 To simplify the proof, we take the initializations of some local variables together as atomic.
@@
 TS-Stack states also contains sets $O_a, O_r \in \mathbb{O}$ which are operation identifier sets of push and pops respectively, and the control point function $cp$ which is a map from operation identifiers to the control points set that are presented in the flow diagram Figure X  \textcolor{red}{Flow diagram is referenced here}.
-Transition relation of the TS-STack is presented in Figure~\ref{fig:transitions:TSS}.
+Transition relation of the TS-STack is presented in Figure~\ref{fig:transitions:TSSPush} (push rules) and Figure~\ref{fig:transitions:TSSPop} (pop rules).
 Next, we show that the linearizability of TS Stack.
@@
       k\not\in dom(cp) \\ 
-      d\neq {\tt EMPTY} \\
-      \forall k'.\ ov'(k') = ov(k')\cup \{k\}
+      d\neq {\tt null}
     }{
-      O,<,\ell,rv,cp,be,ov 
+      ..., O_a, x, cp,...
       \xrightarrow{inv(push,d,k)} 
-      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1],be,ov'
+     ..., O_a\cup\{k\}, x[k \mapsto d], cp[k \mapsto A_1], ...
+    }\hspace{5mm}
+
+    \inferrule[push1]{
+      cp(k) = A_1 \\ 
+      *n' = (x(k),\texttt{MAX\_INT}, \texttt{null}, \texttt{false})
+    }{
+      ..., n, cp,...
+      \xrightarrow{push1(k)} 
+     ..., n[k \mapsto n'], cp[k \mapsto A_2], ...
+    }\hspace{5mm}
+
+  \inferrule[push2]{
+      cp(k) = A_2 
+    }{
+      ..., pools, cp,...
+      \xrightarrow{push2(k)} 
+     ..., pools[myTid(k) \mapsto n(k)], cp[k \mapsto A_3], ...
+    }\hspace{5mm}
+
+      \inferrule[push3]{
+      cp(k) = A_3 
+    }{
+      ..., i, TS, cp,...
+      \xrightarrow{push3(k)} 
+     ..., i[k \mapsto TS], TS+1, cp[k \mapsto A_4], ...
+    }\hspace{5mm}
+
+   \inferrule[push4]{
+      cp(k) = A_4 \\
+      n'(k) = n(k)[ts \mapsto i(k)] \\
+      \forall k'. cp(k') = A_6 \implies i(k') < i(k)
+    }{
+      ..., n, cp,...
+      \xrightarrow{push4(k)} 
+     ..., n[k \mapsto n'(k)], cp[k \mapsto A_5], ...
+    }\hspace{5mm}
+
+   \inferrule[ret-push]{
+      cp(k) = A_5 
+    }{
+      ...,cp,...
+      \xrightarrow{ret(push,k)} 
+     ...,cp[k \mapsto A_6], ...
     }\hspace{5mm}
@@
  \vspace{-5mm}
-  \caption{The transition relation of TS-Stack
+  \caption{The push derivation rules of $TSS$.
+We only mention the state components that are modified.
+Unmentioned state components have the names in the algorithm in the prestate.
+$*n = (a,b,c,d)$ is shorthand for $n->data = a$, $n->ts = b$, ...
+$n' = n[ts \mapsto expr]$ is short for $n'->ts = expr$ and all the other fields of $n$ and $n'$ are the same.
   }
-  \label{fig:transitions:TSS}
+  \label{fig:transitions:TSSPush}
+\vspace{-6mm}
+\end{figure}
+
+\begin{figure} [t]
+{\scriptsize
+  \centering
+  \begin{mathpar}
+    \inferrule[call-push]{
+      k\not\in dom(cp) \\ 
+      d\neq {\tt null}
+    }{
+      ..., O_a, x, cp,...
+      \xrightarrow{inv(push,d,k)} 
+     ..., O_a\cup\{k\}, x[k \mapsto d], cp[k \mapsto A_1], ...
+    }\hspace{5mm}
+
+      \end{mathpar}
+  }
+ \vspace{-5mm}
+  \caption{The pop derivation rules of $TSS$.
+We only mention the state components that are modified.
+Unmentioned state components have the names in the algorithm in the prestate.
+$*n = (a,b,c,d)$ is shorthand for $n->data = a$, $n->ts = b$, ...
+$n' = n[ts \mapsto expr]$ is short for $n'->ts = expr$ and all the other fields of $n$ and $n'$ are the same.
+  }
+  \label{fig:transitions:TSSPop}
 \vspace{-6mm}
@@
 \begin{lem}
-$L_C$ is a Com(pop)- refinement of $AbsS$.
+$TSS$ is a Com(pop)- refinement of $AbsS$. 
 \end{lem}
 \begin{proof}
-Since $AbsS$ is deterministic with respect to $C \cup R \cup Com(pop)$, there exists a Com(pop) forward simulation from $L_C$ to $AbsS$ iff $L_C$ is a Com(pop)-refinement of $AbsS$.
+Since $AbsS$ is deterministic with respect to $C \cup R \cup Com(pop)$, there exists a Com(pop) forward simulation from $TSS$ to $AbsS$ iff $TSS$ is a Com(pop)-refinement of $AbsS$.
 Hence, we define the relation $fs$ and show that it is a Com(pop)-forward simulation relation.
@@
 \item[\emph{TraverseBefore}] If a pop operation $p$ is currently visiting node $n$ and there is a node $m$ coming before $n$ in the traverse order with a greater timestamp, then the operation that inserts $m$ must be concurrent with $p$.
-More formally, let $k \in O_t$ such that $n_s(k) <^{tr}_s n_s(p)$ and $n_s(k)->ts > n_s(p)->ts$.
+More formally, let $k \in O_t$ such that $n_s(k) <^{tr}_s n_s(p)$ and $n_s(k)->ts \geq n_s(p)->ts$.
 Then, $k \in ov_t(p)$.
-\item[\emph{TraverseAfter}] If a pop operation $p$ is currently visiting node $n$ that is not null and its youngest element $m$ is not null and still not taken in state $s$, then either $m$ is a candidate to be removed by $p$ in $t$ or there exists a later node $m'$ than $n$ such that $m'$ is a candidate in $t$ it has a bigger timestamp than n.
-More formally, assume that there exists $k, k' \in O_t$ such that $youngest_s(p) = n_s(k)$ and $n_s(k') = n_s(p)$.
+\item[\emph{TraverseAfter}] If a pop operation $p$ is currently visiting node $n$ that is not null and its youngest element $m$ is not null and still not taken in state $s$, then either $m$ is a candidate to be removed by $p$ in $t$ or there exists a later node $m'$ than $n$ such that $m'$ is a candidate in $t$ and it has a bigger timestamp than n.
+More formally, assume that there exists $k, k' \in O_t$ such that $youngest_s(p)->taken \neq false$, $youngest_s(p) = n_s(k)$ and $n_s(k') = n_s(p)$.
 Then, either $k \in ov_t(p) \vee k \in be_t(p)$ or there exists $k'' \in O_t$ s.t.
 $n_s(k'')->ts > n_s(k) ->ts$ and $k'' \in ov_t(p) \vee k'' \in be_t(p)$ and either $k' <^{tr}_s k''$ or $n_s(p) = n_s(k'') \wedge cp_s(p) = R_j \wedge j>3$.
+\end{itemize}
+Next, we will show that $fs$ is really a Com(pop)-forward simulation relation.
+Except the trivial base case, we case-split on the transition rules.
+We first assume $(s, \alpha s') \in \delta_C$ and $t \in fs[s]$.
+Then, we find  corresponding transition $\alpha' \in \Sigma_{AbsS}$ obeying the Com(pop)-forward simulation relation conditions and obtain $t'$ such that $(t, \alpha' t') \in \delta_{AbsS}$  and $t' \in fs[s']$.
+
+We observe that if $\alpha \in C \cup R \cup Com(pop)$, then the corresponding rule in $AbsS$ is $\alpha' = \alpha$. Otherwise, $\alpha' = \epsilon$.
+
+Let the following derivation rule of $TSS$ be the one for describing $\alpha$:
+\begin{mathpar}
+    \inferrule{
+      \psi
+    }{
+      s
+      \xrightarrow{\alpha} 
+      s'
+    }
+\end{mathpar}
+and the following one be the derivation rule of $AbsS$ describing $\alpha'$ if $\alpha' \neq \epsilon$ (equivalently $\alpha' = \alpha$):
+\begin{mathpar}
+    \inferrule{
+      \psi'
+    }{
+      q
+      \xrightarrow{\alpha'} 
+      q'
+    }
+\end{mathpar}
+For the cases $\alpha' = \alpha$, we first need to show $\alpha'$ is enabled in state $t$ i.e., $t$ satisfies $\psi'$.
+If this can not be directly obtained from $s$ satisfies $\psi$ and using one or two obvious conditions on $fs$ (since $t \in fs[s]$), we show the derivation in the proof.
+Then, $t'$ is obtained in a unique way since $AbsS$ is deterministic on its alphabet $ \Sigma_{AbsS} = C \cup R \cup Com(pop)$.
+The, only other thing to show is $t' \in fs[t']$.
+We show this by proving that $t'$ does not violate any of the conditions of the $fs$ described above.
+We only explain why the new instantiations due to the difference between $s'$ and $s$ or the difference between $t'$ and $t$ do not violate the conditions.
+We skip the instances that we assumed while relating $s$ to $t$.
+
+For the cases in which $\alpha' = \epsilon$, we have $t'=t$ and the only thing to show is $t \in fs[s']$.
+Again, we only explain why the new instantions due to the difference between $s'$ and $s$ do not violate the conditions.
+\begin{itemize}
+\item[\textsc{call-push}] The same derivation rule of $TSS$ is applied to $t$ obtain $t'$.
+The premise of the rule is satisfied by $t$ trivially in the sense explained before.
+The new vertex $k$ is added to the $O_t$ such that $k$ is maximal, pending and every completed vertex is ordered before $k$ in $t'$.
+Moreover, $k$ is overlapping with every pending pop.
+To see that $t' \in fs[s']$ we observe the following: \emph{Nodes} condition is preserved because $k \in O_{t'}$.
+Since the newly added vertex $k$ is maximal and pending in $t'$, \emph{Pend/Comp} condition is preserved.
+\emph{Frontiers} and \emph{MaximalOV} conditions are not violated since $k$ is added to $ov(p)$ set for every pending pop operation $p$.
+\item[\textsc{push1}] We have $t' = t$ and show $t \in fs[s']$.\emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex.
+\emph{TSOrder} is still preserved.
+Timestamp of $n_{s'}(k)$ is maximal and every other nodes of push operations with maximal timestamp in $s'$ are pending vertices in $t$.
+Hence there can be no ordering between those pushes and $k$ in $t$ that can violate \emph{TSOrder}.
+Moreover, $k$ is maximal in $t$ which means that it cannot be ordered before another push $k'$ of which node has a lower timestamp.
+\emph{TidOrder} is also satisfied.
+Since $k$ is ordered after every completed push in $t$ and every other push by the same thread is completed, ordering required by the \emph{TidOrder} is present.
+\item[\textsc{push2}] We have $t' = t$ and show $t \in fs[s']$.
+\emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex.
+One can also see that the \emph{TraverseBefore} condition is preserved.
+Let the pop $p$ visiting node $m$ and $n_{s'}(k) <^{tr}_{s'} m$.
+Since $k$ and $p$ are both pending in $s$ and $t \in fs[s]$, $k \in ov_t(p)$ (by the \emph{Frontiers} condition).
+Hence, \emph{TraverseBefore} is preserved.
+\item[\textsc{push3}] We have $t' = t$ and show $t \in fs[s']$.
+We consider two cases: $n_s(k)->taken$ is \texttt{true} or it is \texttt{false}.
+For the former case, $k \notin O_t$.
+The only new instantiation we check is $k \notin O_t$ does not violate \emph{Nodes} condition while relating $s'$ to $t$.
+For the latter case, we have $k \in O_t$.
+\emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex after changing $s$ to $s'$.
+\item[\textsc{push4}] We have $t' = t$ and show $t \in fs[s']$.
+We consider two cases: $n_s(k)->taken$ is \texttt{true} or it is \texttt{false}.
+For the former case, \emph{Nodes} condition is still satisfied since $k$ remains to be not a vertex.
+For the latter case \emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex.
+\emph{TSOrder} condition is still not violated since if $k'<_t k$, then $k'$ is a completed vertex in $s$ and $s'$.
+By the premise of the rule (which can be shown to hold for every operation at control point $A_4$) $i_s(k') < i_s(k)$ and consequently $n_{s'}(k')->ts < n_{s'}(k)->ts$.
+Since every other push by the thread of $k$ is completed, \emph{TidOrder} still continues to hold for the same reasons.
+\emph{TraverseAfter} condition is also preserved.
+Let $k'$ be the push and $p$ be the pop such that $n_s(k') = youngest_s(p)$, $n_s(k') \leq^{tr}_s n_s(k)$, $n_s(k')->ts < n_s(k)->ts$ and $k \in ov_t(p)$ or $k \in be_t(p)$.
+Assume $n_{s'}(k')->ts \geq N_{s'}(k)->ts$ after the action.
+Then, $k'$ must be a pending push both in $s$ and $s'$ by the premise of the derivation rule and $k' \in ov_t(p)$ must be true by \emph{Frontiers} condition and $t \in fs[s]$.
+Hence, the \emph{TraverseAfter} condition is preserved.
+\item[\textsc{ret-push}] We consider two cases, $n_s(k)->taken$ is \texttt{false} or \texttt{true}.
+For the former case, we obtain $t'$ by applying \textsc{ret-push1} rule of $AbsS$.
+\emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ becomes a completed vertex in $t'$.
+\emph{Frontiers} condition still holds since although $k$ become a maximally closed vertex in $t'$, we have $k \in ov_{t'}(p)$ for all pending nodes $p$ (due to \emph{Frontiers} condition, $t \in fs[s]$ and $k$ was a pending operation in state $t$, $k \in ov_t(p)$).
+
+For the latter case, we obtain $t'$ by applying \textsc{ret-push2} rule of $AbsS$. \emph{Nodes} condition is still satisfied since $k \notin O_{t'}$. 
 \end{itemize}

### 2017-01-23 draftCAV/app-backSim.tex

@@
 %!TEX root = draft.tex
-\section{Backward simulations}\label{app:backSim}
+\section{Normal Forward/Backward Simulations}\label{app:backSim}
+
+We define a class of forward/backward simulations, called \emph{normal simulations}, that are used in the proofs in Appendix~\ref{app:absImplQueue} and Appendix~\ref{app:absImplStack}.
+
+\begin{dfn}
+Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
+A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{normal $\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
+\begin{itemize}
+\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$ 
+\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0<i<|@s|$.
+%$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $.
+The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
+\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma_2\setminus\Gamma$, for each $0\leq i<|@s| -1$.
+% where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
+\item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $\delta_2(u,\gamma, u')$. 
+\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma_1\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $\sigma\in (\Sigma_2\setminus\Gamma)^*$.
+%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
+\end{itemize}
+\end{dfn}
+%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
+A $\Gamma$-forward simulation requires that every step of $L_1$ corresponds to a (possibly empty) sequence of steps of $L_2$.
+In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action.
+The rest of the transitions in $L_1$ are matched to a possibly empty sequence of transitions of $L_2$ with arbitrary labels.
 
@@
 Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
-A relation $bs \subseteq Q_1 \times Q_2$ is called a \emph{backward simulation} from $L_1$ to $L_2$ iff the following holds:
+A relation $bs \subseteq Q_1 \times Q_2$ is called a \emph{normal $\Gamma$-backward simulation} from $L_1$ to $L_2$ iff the following holds:
 \begin{itemize}
@@
 \item[(ii-c)] If $(s,\gamma, s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $\delta_2(u,\gamma,u')$
-\item[(ii-d)] If $(s,e,s') \in \delta_1$ for some $e \in \Sigma\setminus \Gamma$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$ and $@s_i\in \Sigma\setminus\Gamma$, for each $0\leq i<|@s|$.
+\item[(ii-d)] If $(s,e,s') \in \delta_1$ for some $e \in \Sigma\setminus \Gamma$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $u \xrightarrow{@s} u'$ and $\sigma\in (\Sigma_2\setminus\Gamma)^*$.
 %where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
@@
 
-\begin{lem}
-Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
+%\begin{lem}
+%Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
 $L_1$ $\Gamma$-refines $L_2$ if{f} there exists a $\Gamma$-backward simulation from $L_1$ to $L_2$.
-\end{lem}
-\begin{proof}
-Looks trivial and follows Lynch paper. Can be completed later.
-\end{proof}
+%\end{lem}
+%\begin{proof}
+%Looks trivial and follows Lynch paper. Can be completed later.
+%\end{proof}

### 2017-01-23 draftCAV/stack.tex

@@
 We call such actions \emph{commit points}.
-More generally, commit points are actions that access shared variables, from which every control-flow path leads to the return control point and contains no more accesses to the shared memory (i.e., the return value is computed using only local variables).
+More generally, commit points are actions that access shared variables, from which every control-flow path leads to the return control point and contains no more accesses to the shared memory (i.e., after a commit point, the return value is computed using only local variables).
 

### 2017-01-23 draftCAV/queue.tex
removing auxiliary files

@@
       if ( x != null ) return x;
-    }
-  }
-}
+}}}
   \end{lstlisting}

### 2017-01-23 draftCAV/prelim.tex
a pass until stacks

@@
 \section{Preliminaries}
-
+\vspace{-2mm}
 We formalize several abstraction relations between libraries using a simple
@@
 
+\vspace{-3.5mm}
 \subsection{Libraries}
-
+\vspace{-1.5mm}
 Programs interact with libraries by calling named library \emph{methods}, which
@@
 $\<Methods>$ and $\<Vals>$ of methods and values, we fix the sets
+\vspace{-2mm}
 \begin{align*}
   & C = \set{ inv(m,d,k) : m \in \<Methods>, d \in \<Vals>, k \in \<Ops> }
-  \text{, and } \\
-  & R = \set{ ret(m,d,k) : m \in \<Methods>, d \in \<Vals>, k \in \<Ops> }  
+  \text{ and } 
+  R = \set{ ret(m,d,k) : m \in \<Methods>, d \in \<Vals>, k \in \<Ops> }  
 \end{align*}
+
+\vspace{-2mm}
+\noindent
 of \emph{call actions} and \emph{return actions}; each call action $inv(m,d,k)$
@@
 %Well-formed words represent traces of a library. 
-We assume every set of 
-words is closed under isomorphic renaming of operation identifiers. For
+%We assume every set of 
+%words is closed under isomorphic renaming of operation identifiers. 
+For
 notational convenience, we take $\<Ops>=\<Nats>$ for the rest of the paper.
@@
 
+\vspace{-3.5mm}
 \subsection{Refinement and Linearizability}
-
+\vspace{-1.5mm}
 %\textcolor{red}{A general organizational idea for  this subsection: Linearizability is defined in terms of weakening in which observational actions are just call and return events.
@@
 Conformance of a library $L_1$ to a specification given as an ``abstract'' library $L_2$ 
-is formally captured by \emph{(observational) refinement}. Informally, given two libraries
-$L_1$ and $L_2$ implementing the methods of some concurrent object, we
+is formally captured by \emph{(observational) refinement}. Informally, 
+%given two libraries
+%$L_1$ and $L_2$ implementing the methods of some concurrent object, 
+we
 say $L_1$ refines $L_2$ if{f} every computation of every program
@@
 when $L_2$ is an atomic library~\cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}.
-. 
+
 %We fix a set of methods $\<Methods>$ and values $\<Vals>$.
-
+\vspace{-1.5mm}
 %\textcolor{red}{We have two Definition 1. The problem occurs because we used different commands for definitions.}
@@
 %for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
+\vspace{-1.5mm}
 \end{definition}
@@
 
+\vspace{-1.5mm}
 \begin{theorem}[\cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}]
@@
   if $L_2$ is atomic.
-
+\vspace{-1.5mm}
 \end{theorem}

### 2017-01-23 draftCAV/queue.tex
a pass until stacks

@@
 %!TEX root = draft.tex
+\vspace{-3.5mm}
 \section{Queues With Fixed Dequeue Linearization Points}\label{sec:queues}
-
-The typical abstract implementation of a concurrent queue, denoted by $AbsQ_0$, maintains a sequence of values, the enqueue method adds a value atomically to the beginning of the sequence, and the dequeue method removes a value from the end of the sequence (if any, otherwise it returns ``empty'').
-In this section, we describe another abstract implementation, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence.
-We show that there exists a forward simulation from any correct queue implementation where the \emph{dequeue} methods have fixed linearization points (the enqueue methods are unconstrained) to $AbsQ$.
-This covers all the queue implementations that we are aware of.
-We describe a forward simulation from the Herlihy\&Wing Queue~\cite{journals/toplas/HerlihyW90} to $AbsQ$.
+\vspace{-1.5mm}
+The typical abstract implementation of a concurrent queue, denoted as $AbsQ_0$, maintains a sequence of values, the enqueue method adds a value atomically to the beginning of the sequence, and the dequeue method removes a value from the end of the sequence (if any, otherwise it returns {\tt EMPTY}).
+Both methods have a fixed linearization point when the update of the shared sequence happens.
+%~\footnote{For $\<Methods>=\{enq,deq\}$, the alphabet of $AbsQ_0$ is $C\cup R\cup Lin$.}.
+For some queue implementations, e.g., the Herlihy\&Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), there exists no forward simulation to $AbsQ_0$ although they are a refinement of $AbsQ_0$.
+The main reason is that the enqueue methods in these implementations don't have a \emph{fixed} linearization point.
+In this section, we propose a new abstract implementation for queues, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence.
+We show that there exists a forward simulation from any correct queue implementation where only the \emph{dequeue} methods have fixed linearization points (the enqueue methods are unconstrained) to $AbsQ$.
+This covers all the queue implementations that we are aware of, in particular $\mathit{HWQ}$.
+ We also describe a forward simulation from $\mathit{HWQ}$ to $AbsQ$.
 
 TODO HOW CAN WE CONVINCE THAT OTHER IMPLEMENTATIONS HAVE THE SAME PROPERTY ?
-
-\subsection{Abstract Queue Implementation}
-
-Before defining the abstract implementation $AbsQ$, we describe a queue implementation listed in Figure~\ref{fig:HerlihyWing} and known as the Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short),
-where only the dequeue methods have fixed linearization points.
-
-\begin{wrapfigure}{l}{5.7cm}
+\vspace{-3.5mm}
+\subsection{Enqueue Methods With Non-Fixed Linearization Points}
+\vspace{-1.5mm}
+% describe a queue implementation listed in Figure~\ref{fig:HerlihyWing} and known as the Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), where only the dequeue methods have fixed linearization points.
+\begin{wrapfigure}{l}{5.3cm}
 \vspace{-8mm}
@@
 \label{fig:HerlihyWing}
-\vspace{-6mm}
+\vspace{-7mm}
 \end{wrapfigure}
+We describe $\mathit{HWQ}$ as an instance of a queue implementation where the linearization points of the enqueue methods are not fixed.
 The shared state consists of an array {\tt items} storing the values in the queue and a counter {\tt back} storing the index of the first unused position in {\tt items}.
 Initially, all the positions in the array are {\tt null} and {\tt back} is 0.
-An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueue operations) and then, stores the input value {\tt x} at this position.
+An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueues) and then, stores the input value {\tt x} at this position.
 The dequeue method traverses the array {\tt items} starting from the beginning and atomically swaps {\tt null} with the encountered value.
@@
 The linearization points of the enqueue methods are not fixed, they depend on dequeue operations executing in the future.
-Consider the following trace with two concurrent enqueues (${\tt i}(k)$ represents the value of {\tt i} in operation $k$):
-\begin{align*}
-inv(enq,x,1)\ \ \ inv(enq,y,2)\ \ \ {\tt i}(1) = \mbox{{\tt bck++}}\ \ \ {\tt i}(2) = \mbox{{\tt bck++}}\ \ \ {\tt items[i(2)]} = y
-\end{align*}
+Consider the following trace with two concurrent enqueues (${\tt i}(k)$ represents the value of {\tt i} in operation $k$): $inv(enq,x,1)$, $inv(enq,y,2)$, ${\tt i}(1) = \mbox{{\tt bck++}}$, ${\tt i}(2) = \mbox{{\tt bck++}}$, ${\tt items[i(}2{\tt )]} = y$.
+%\vspace{-1.5mm}
+%\begin{align*}
+%inv(enq,x,1)\ \ \ inv(enq,y,2)\ \ \ {\tt i}(1) = \mbox{{\tt bck++}}\ \ \ {\tt i}(2) = \mbox{{\tt bck++}}\ \ \ {\tt items[i(}2{\tt )]} = y
+%\vspace{-1.5mm}
+%\end{align*}
 %\begin{align*}
@@
 However, a dequeue operation executing until completion after this trace will return value $y$ (only position $1$ is filled in the array {\tt items}) which is not consistent with this linearization.
-On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i(1)]} = x$ and a completed dequeue that in this case returns $x$, leads to the following incorrect linearization:
-\begin{align*}
-inv(enq,y,2)\ ret(enq,2) inv(enq,x,1)\ ret(enq,1)\ inv(deq,3)\ ret(deq,1,3).
-\end{align*}
+On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i(}1{\tt )]} = x$ and a completed dequeue that in this case returns $x$, leads to the following incorrect linearization: $inv(enq,y,2)$, $ret(enq,2)$, $inv(enq,x,1)$, $ret(enq,1)$, $inv(deq,3)$, $ret(deq,x,3)$.
+%\vspace{-1.5mm}
+%\begin{align*}
+%inv(enq,y,2)\ ret(enq,2) inv(enq,x,1)\ ret(enq,1)\ inv(deq,3)\ ret(deq,x,3).
+%\vspace{-1.5mm}
+%\end{align*}
 
@@
 
-Since the linearization points of the enqueues are not fixed, it is not possible to define a forward simulation from $\mathit{HWQ}$ to the standard abstract implementation $AbsQ_0$.
+Since the linearization points of the enqueues are not fixed, there exists no forward simulation from $\mathit{HWQ}$ to $AbsQ_0$. 
 In the following, we describe the abstract implementation $AbsQ$ for which such a forward simulation does exist.
 
-Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation.
-The linearization point of a dequeue operation with return value $d\neq{\tt EMPTY}$ is enabled only if the happens-before order stored in the current state contains a minimal enqueue operation that adds the value $d$.
-The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
-When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueue operations (the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
-The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point. 
+\vspace{-3.5mm}
+\subsection{Abstract Queue Implementation}
+\vspace{-1.5mm}
 %\textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too.
@@
 We also need to update the rules in figure.}
-
-\begin{wrapfigure}{l}{7cm}
-\vspace{-8mm}
+\begin{wrapfigure}{l}{6.7cm}
+\vspace{-7mm}
 \includegraphics[width=7cm]{fig-queue12.pdf}
@@
 %\includegraphics[width=7cm]{fig-queue2.pdf}
-\vspace{-5mm}
+\vspace{-8mm}
 \caption{Simulating queue histories with $AbsQ$. The order between actions is from left to right.}
 \label{fig:queueSim}
-\vspace{-10mm}
+\vspace{-6mm}
 \end{wrapfigure}
+Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation.
+The linearization point of a dequeue operation with return value $d\neq{\tt EMPTY}$ is enabled only if the happens-before stored in the current state contains a minimal enqueue that adds the value $d$.
+The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
+When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueues (the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
+The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.
+
 Figure~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
@@
 
-Formally, the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed (this flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue ($R_1,R_2,R_3$) operation.
-The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}.
-The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points.
-The latter are denoted by $lin(deq,d,k)$ where $d\in \<Vals>$, $k\in\<Ops>$.
-$Lin(deq)$ is the set of all such actions.
+Formally, the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).
+All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}.
+The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points, denoted by $lin(deq,d,k)$.
+$Lin(deq)$ is the set of all actions $lin(deq,d,k)$.
 
-Concerning enqueue operations, the rule {\sc call-enq} orders the invoked operation after all the completed enqueue operations stored in the current state, and the rules {\sc ret-enq1}/{\sc ret-enq2} flip the corresponding flag from {\tt PEND} to {\tt COMP} provided that the operation is still present in the current state.
-For dequeue operations, {\sc call-deq} has no effect other than incrementing the control point and {\sc ret-deq} checks whether the return value is the same as the one fixed at the linearization point.
-The linearization point rule {\sc lin-deq1} corresponds to the case of a non-empty queue, showing that $lin(deq,d,k)$ is enabled only if the value $d$ has been added by an enqueue which is minimal in the current happens-before order.
-When it is enabled, it removes the enqueue adding $d$ from the state.
+Concerning enqueue operations, the rule {\sc call-enq} orders the invoked operation after all the completed enqueues in the current state, and the rules {\sc ret-enq1}/{\sc ret-enq2} flip the corresponding flag from {\tt PEND} to {\tt COMP} provided that the operation is still present in the current state.
+For dequeue operations, {\sc call-deq} only increments the control point and {\sc ret-deq} checks whether the return value is the same as the one fixed at the linearization point.
+The linearization point rule {\sc lin-deq1} corresponds to the case of a non-empty queue, showing that $lin(deq,d,k)$ is enabled only if $d$ has been added by an enqueue which is minimal in the current happens-before.
+When enabled, it removes the enqueue adding $d$ from the state.
 The linearization point rule {\sc lin-deq2} corresponds to the case of dequeue operations linearized with an {\tt EMPTY} return value.
@@
        cp(k) = R_1 \\
-       \forall o\in O. \ell_2(o)={\tt EMPTY}
+       \forall o\in O.\ \ell_2(o)={\tt PEND}
     }{
@@
   }
- \vspace{-5mm}
+ \vspace{-6mm}
   \caption{The transition relation of $AbsQ$.
@@
 
-Both methods have a fixed linearization point when the update of the shared sequence happens.
-Let $AbsQ_0$ denote this implementation~\footnote{For $\<Methods>=\{enq,deq\}$, the alphabet of $AbsQ_0$ is $C\cup R\cup Lin$.} (formally defined in Appendix~\ref{app:absImplQueue}).
+% Let $AbsQ_0$ denote this implementation (formally defined in Appendix~\ref{app:absImplQueue}).
 The following result states that the library $AbsQ$ has exactly the same set of histories as the standard abstract library $AbsQ_0$ (see Appendix~\ref{app:absImplQueue} for a proof).
 
+\vspace{-1.5mm}
 \begin{theorem}\label{th:absImplQueue}
 $AbsQ$ is a refinement of $AbsQ_0$ and vice-versa.
+\vspace{-2mm}
 \end{theorem}
 
-A trace of a queue implementation is called \emph{$Lin(deq)$-complete} when every completed dequeue has a linearization point, i.e., TODO.
-A queue implementation $L$ over alphabet $\Sigma$ is called \emph{with fixed dequeue linearization points} if{f} $C\cup R\cup Lin(deq)\subseteq \Sigma$
-and every trace $@t\in Tr(L)$ is $Lin(deq)$-complete.
+A trace of a queue implementation is called \emph{$Lin(deq)$-complete} when every completed dequeue has a linearization point, i.e., each return action $ret(deq,d,k)$ is preceded by an action $lin(deq,d,k)$.
+A queue implementation $L$ over alphabet $\Sigma$, such that $C\cup R\cup Lin(deq)\subseteq \Sigma$, is called \emph{with fixed dequeue linearization points} when every trace $@t\in Tr(L)$ is $Lin(deq)$-complete.
 
@@
 
+\vspace{-1.5mm}
 \begin{corollary}
 A queue implementation $L$ with fixed dequeue linearization points is a $C\cup R\cup Lin(deq)$-refinement of $AbsQ_0$ if{f} there exists a $C\cup R\cup Lin(deq)$-forward simulation from $L$ to $AbsQ$.
+\vspace{-1.5mm}
 \end{corollary}
 
+\vspace{-4.5mm}
 \subsection{A Correctness Proof For Herlihy\&Wing Queue}\label{ssec:HerlihyWing}
-
+\vspace{-1mm}
 We describe a forward simulation $\mathit{fs}_1$ from $\mathit{HWQ}$ to $AbsQ$.
-Essentially, an $AbsQ$ state associated by $\mathit{fs}_1$ to a $\mathit{HWQ}$ state consists of all the  enqueue operations for which the input is still present in the array {\tt items} and all the pending enqueue operations that have at most reserved an array position, ordered by a relation $<$ satisfying the following:
+A $\mathit{HWQ}$ state is related by $\mathit{fs}_1$ to an $AbsQ$ state that consists of all the  enqueue operations for which the input is still present in the array {\tt items} and all the pending enqueue operations that have at most reserved an array position, ordered by a relation $<$ satisfying the following:
+\vspace{-2mm}
 \begin{itemize}
@@
 	\item[(b)] $<$ is consistent with the order in which positions of {\tt items} have been reserved, i.e., for every two enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $k' \not< k$,
-	\item[(c)] an enqueue which has reserved an array position $i$ %and executed only the first statement 
+	\item[(c)] an enqueue which has reserved a position $i$ %and executed only the first statement 
 	can't be ordered before another enqueue that has reserved a position $j \geq i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal, i.e., for every two enqueues $k$ and $k'$, and a dequeue $k_d$, such that
 
+	\vspace{-2mm}
 	\noindent
@@
 
+	\vspace{-6mm}
 	\noindent
@@
 The predicate $k_d\atCP {\tt if}\text{-}{\tt inc}$ holds when the dequeue $k_d$ is at a control point after a {\tt swap} returning {\tt null} and before the increment of {\tt i}.
+\vspace{-2mm}
 \end{itemize}
 
+\noindent
 An enqueue is labeled by $(d,{\tt PEND})$ where $d$ is the input value if it's pending and by  $(d,{\tt COMP})$, otherwise.
@@
 
+%\setlength{\parskip}{0pt}
+
 We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an \emph{empty} sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in\mathit{fs}_1$, then $(s',t)\in\mathit{fs}_1$ for each $AbsQ$ state $t$.
-We focus on the following essential property, called \emph{monotonicity}: the set of possible orders $<$ associated by $\mathit{fs}_1$ to $s'$ doesn't exclude any order $<$ associated to $s$.
+Essentially, this consists in proving the following property, called \emph{monotonicity}: the set of possible orders $<$ associated by $\mathit{fs}_1$ to $s'$ doesn't exclude any order $<$ associated to $s$.
 %Essentially, this boils down to showing that the constraints over $<$ in the definition of $\mathit{fs}$ are an invariant for these steps.
@@
 Concerning enqueues, let $s'$ be the state obtained from $s$ when a pending enqueue $k$ reserves an array position.
-This enqueue operation must be maximal in both $t$ and any state $t'$ related to $s'$ (because it is pending).
+This enqueue must be maximal in both $t$ and any state $t'$ related to $s'$ (since it's pending).
 Moreover, there is no dequeue that can ``observe'' this position before restarting the array traversal.
 Therefore, item (c) in the definition of $<$ doesn't constrain the order between $k$ and some other enqueue neither in $s$ nor in $s'$.
-Since this transition doesn't affect the constraints on the order between enqueue operations different from $k$ (their local variables remain unchanged), we have that monotonicity holds.
+Since this transition doesn't affect the constraints on the order between enqueues different from $k$ (their local variables remain unchanged), monotonicity holds.
 This property is trivially satisfied by the second step of enqueue which doesn't affect {\tt i}.
@@
 Instantiations that are enabled only in $s'$ may in principle lead to a violation of monotonicity (since they restrict the orders $<$ associated to $s'$).
-For the two steps that begin an array traversal, i.e., reading the index of the last used position and setting the iterator {\tt i} to $0$, there exist no  such new instantiations in $s'$ because the value of {\tt i} is either not set or $0$.
+For the two steps that begin an array traversal, i.e., reading the index of the last used position and setting {\tt i} to $0$, there exist no  such new instantiations in $s'$ because the value of {\tt i} is either not set or $0$.
 % (it is trivial to notice that applying these steps doesn't disable such instantiations that were possible in $s$).
@@
 An enqueue that didn't reserved a position is by definition maximal in the order $<$ and therefore, not a predecessor of $k_e$.
-The state $t'$ obtained from $t$ through a $lin(deq,d,k)$ transition is related to $s'$ because essentially, (1) the value added by $k_e$ is not anymore present in {\tt items} which implies that $k_e$ doesn't occur in any $AbsQ$ state related to $s'$, and (2) the value of ${\tt x}(k)$ is set to $d\neq {\tt null}$ which implies that $rv(k)$ is set to $d$ in every $AbsQ$ state related to $s'$.
+Then, the state $t'$ obtained from $t$ through a $lin(deq,d,k)$ transition is related to $s'$ because essentially, (1) the value added by $k_e$ is not anymore present in {\tt items} which implies that $k_e$ doesn't occur in any $AbsQ$ state related to $s'$, and (2) the value of ${\tt x}(k)$ is set to $d\neq {\tt null}$ which implies that $rv(k)$ is set to $d$ in every $AbsQ$ state related to $s'$.
+%
+%\vfill
+%

### 2017-01-23 draftCAV/refProofs.tex
a pass until stacks

@@
 %!TEX root = draft.tex
-
+\vspace{-3.5mm}
 \section{Refinement Proofs}
-
+\vspace{-1.5mm}
 Library refinement is the instance of a more general notion of refinement between LTSs
@@
 In general, forward simulations are \emph{not} a complete proof method for library refinement because libraries are not 
-$C\cup R$-deterministic (the same sequence of call and 
-return actions can lead to different states depending on the interleaving of the internal actions).
+$C\cup R$-deterministic (the same sequence of call/return actions can lead to different states depending on the interleaving of the internal actions).
 However, there are classes of atomic libraries, e.g., libraries with ``fixed linearization points'' (defined later in this section), 
 for which it is possible to identify a larger alphabet $\Gamma$ of observable actions (including call/return actions), 
-and implementations that are $\Gamma$-deterministic. When focusing on queues and stacks, 
+and implementations that are $\Gamma$-deterministic. For queues and stacks, 
 Section~\ref{sec:queues} and Section~\ref{sec:stacks} define other such classes of implementations that cover
 all the implementations that we are aware of.
+
+Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over $\Sigma_1$ and $\Sigma_2$, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$.
+Also, let $\Gamma$ be a set of actions s.t.
+$C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
 
@@
 %are fixed to particular implementation actions.
-
-\begin{dfn}\label{def:gref}
-Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$.
+\vspace{-1.5mm}
+\begin{definition}\label{def:gref}
+%Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$.
 Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
@@
 %for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
-\end{dfn}
+\vspace{-1.5mm}
+\end{definition}
 
@@
 
-We define a notion of \emph{forward} simulation that can be used to prove $\Gamma$-refinement. 
-A dual notion of \emph{backward} simulation is defined in Appendix~\ref{app:backSim}. 
+We define a notion of \emph{forward} simulation that can be used to prove $\Gamma$-refinement (a dual notion of \emph{backward} simulation is defined in 
+Appendix~\ref{app:backSim}). 
 %We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement. 
-For a relation $R\subseteq A\times B$, $R[X]$ denotes the set of elements of $B$ related to elements of $X$ by $R$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.
+For a relation $R\subseteq A\times B$, $R[X]$ is the set of elements related by $R$ to elements of $X$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.
 
 %We can extend the relations between libraries by introducing simulation relations. We will later show that simulation relations imply refinement. 
-
-\begin{dfn}
-Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
-A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
+\vspace{-1.5mm}
+\begin{definition}
+%Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
+A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ if{f} $\mathit{fs}[s_0^1] = \{s_0^2 \}$ and:
+\vspace{-1.5mm}
 \begin{itemize}
-\item $\mathit{fs}[s_0^1] = \{s_0^2 \}$.
 \item If $(s,\gamma,s') \in \delta_1$, for some $\gamma\in \Gamma$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_i=\gamma$, for some $i$, and $@s_j\in \Sigma_2\setminus\Gamma$, for each $j\neq i$.
@@
 \end{itemize}
-\end{dfn}
+\vspace{-3.5mm}
+\end{definition}
 %If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
@@
 It is an instantiation of previous results~~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}.
+
+\vspace{-1.5mm}
 \begin{theorem}\label{th:forSim}
-Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$.
+%Let $L_1$ and $L_2$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$.
 Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
-If there is a $\Gamma$-forward simulation from $L_1$ to $L_2$, then $L_1$ $\Gamma$-refines $L_2$.
+$L_1$ $\Gamma$-refines $L_2$ when there is a $\Gamma$-forward simulation from $L_1$ to $L_2$.
 Moreover, if $L_1$ $\Gamma$-refines $L_2$ and $L_2$ is $\Gamma$-deterministic, then there is a $\Gamma$-forward simulation from $L_1$ to $L_2$.
+\vspace{-1.5mm}
 \end{theorem}
@@
 $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$. The action $lin(m,d,k)$ represents the linearization point of the operation $k$ 
-which is an invocation of method $m$ returning value $d$.
-Let $Lin$ denote the set of such actions. We extend the notion of
-well-formedness to sequences $@t$ over alphabet $\Sigma$,
-such that $C\cup R\cup Lin\subseteq \Sigma$, by adding the following constraints:
-\begin{itemize}
-	\item every linearization point corresponds to an operation that has been called, i.e., TODO
-	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
-	\item the return value of an operation is fixed by the linearization point, i.e., TODO
-\end{itemize}
-%The fact that a history $h_1$ is linearizable w.r.t. another sequential history $h_2$ can be stated as: there exists a way of choosing a linearization
-%point 
-%Formally, a \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by 
-%inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that:
+%which is an invocation of method $m$ 
+returning value $d$.
+Let $Lin$ denote the set of such actions. 
+%We extend the notion of
+%well-formedness to sequences $@t$ over alphabet $\Sigma$,
+%such that $C\cup R\cup Lin\subseteq \Sigma$, by adding the following constraints:
 %\begin{itemize}
@@
 %\end{itemize}
-We assume that libraries produce only well-formed traces. The projection of a library trace over $C\cup R\cup Lin$ is called an 
+%We assume that libraries produce only well-formed traces. 
+The projection of a library trace over $C\cup R\cup Lin$ is called an 
 \emph{extended history}. A trace or extended history is called \emph{$Lin$-complete} when every completed operation has a linearization 
-point, i.e., TODO. 
+point, i.e., each return action $ret(m,d,k)$ is preceded by an action $lin(m,d,k)$. 
 A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup Lin\subseteq \Sigma$ 
@@
 reduces to proving that $L_1$ is a $(C\cup R\cup Lin)$-refinement of an abstract implementation $L_2$ of the same object where method
-bodies are guarded with a global-lock acquisition. Assuming that the abstract implementation is $(C\cup R\cup Lin)$-deterministic, which is typically the case,
+bodies are guarded with a global-lock acquisition. Since the abstract implementation is usually $(C\cup R\cup Lin)$-deterministic,
 by Theorem~\ref{th:forSim}, proving $(C\cup R\cup Lin)$-refinement is equivalent to finding a $(C\cup R\cup Lin)$-forward simulation 
@@
 
-Section~\ref{sec:queues} and Section~\ref{sec:stacks} extend this result to implementations of queues and stacks where the linearization point of the methods 
-that \emph{add} values to the collection is \emph{not} fixed.
+Section~\ref{sec:queues} and Section~\ref{sec:stacks} extend this result to queue and stack implementations where the linearization point of the methods 
+\emph{adding} values to the collection is \emph{not} fixed.

### 2017-01-23 draftCAV/stack.tex
a pass until stacks

@@
 %!TEX root = draft.tex
+\vspace{-3.5mm}
 \section{Stacks With Fixed Pop Commit Points}\label{sec:stacks}
-
+\vspace{-1.5mm}
 While the abstract queue implementation in Section~\ref{sec:queues} can be adapted to stacks~\footnote{The only modification is that the linearization point $lin(pop,d,k)$ with $d\neq{\tt EMPTY}$ is enabled only if $k$ is added by a push which is maximal in the happens-before order stored in the state.}, it can not simulate (through forward simulations) existing stack implementations like the Time-stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} where the linearization points of the pop operations are not fixed.
@@
 
+\vspace{-3.5mm}
 \subsection{Abstract stack implementation}
-
+\vspace{-1.5mm}
 \begin{wrapfigure}{l}{5.4cm}

### 2017-01-23 draftCAV/app-prelim.tex
a pass until stacks

--- 2017-01-23 draftCAV/prelim.tex
@@
 %!TEX root = draft.tex
-\section{Preliminaries}
-
-We formalize several abstraction relations between libraries using a simple
-yet universal model of computation, namely labeled transition systems (LTS).
-This model captures shared-memory programs with an arbitrary number of threads,
-abstracting away the details of any particular programming system irrelevant to
-our development.
-
-A \emph{labeled transition system} (LTS) $A=(Q,\Sigma, s_0, \delta)$ over the 
-possibly-infinite alphabet $\Sigma$ is a possibly-infinite set $Q$ of states with
-initial state $s_0 \in Q$, and a transition relation $\delta \subseteq Q \times @S \times
-Q$. The $i$th symbol of a sequence $@t \in @S^*$ is denoted $@t_i$, and the empty
-sequence is denoted by $\epsilon$.
-An \emph{execution} of $A$ is an alternating sequence of states and transition labels (called also actions)
-$\rho = s_0, e_0,s_1\ldots e_{k-1},s_k$ for some $k>0$ such that $\delta(s_i, e_i, s_{i+1})$
-for each $i$ such that $0\leq i<k$. We write $s_i\xrightarrow{e_i\ldots e_{j-1}}_A s_j$ as shorthand for 
-the subsequence $s_i,e_i,...,s_{j-1},e_{j-1},s_j$ of $\rho$, for any $0\leq i\leq j <k$
-(in particular $s_i\xrightarrow{\epsilon}s_i$).
-The projection $@t| \Gamma$ of a sequence $@t$ is the maximum subsequence of $@t$ over
- alphabet $\Gamma$. This notation is extended to sets of sequences as usual.
-A \emph{trace} of $A$ is the projection $\rho | \Sigma$ of an execution $\rho$ of $A$. 
-The set of executions, resp., traces, of an LTS $A$ is denoted by $E(A)$, resp., $Tr(A)$.
-An LTS is \emph{deterministic} if for any state $s$ and any sequence $@t\in \Sigma^*$, there is at most
-one state $s'$ such that $s\xrightarrow{@t}s'$. More generally, for an alphabet $\Gamma\subseteq \Sigma$,
-an LTS is \emph{$\Gamma$-deterministic} if for any state s and any sequence $@t\in \Gamma^*$, there
-is at most one state $s'$ such that $s\xrightarrow{@t'}s'$ and $@t$ is a subsequence of $@t'$.
-
-\subsection{Libraries}
+\section{Libraries}\label{app:prelim}
 
@@
 parameter/return values.
-%
+
 %\begin{example}
@@
 %\end{example}
-%
+
+\noindent
 We fix an arbitrary set $\<Ops>$ of operation identifiers, and for given sets
@@
 pair call and return actions. 
-%We denote the operation identifier of a
-%call/return action $a$ by $\<op>(a)$. Call and return actions $c \in C$ and $r
-%\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
-We may omit the second field from a call/return action $a$ for methods that have no inputs or return values.
-%A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
-%\emph{well formed} when:
-%\begin{itemize}
-%
-%  \item Each return is preceded by a matching call: \\
-%  $\tau_j \in R$ implies $\tau_i \match \tau_j$ for some $i < j$.
-%
-%  \item Each operation identifier is used in at most one call/return: \\
-%  $\<op>(\tau_i) = \<op>(\tau_j)$ and $i < j$ implies $\tau_i \match \tau_j$.
-%
-%\end{itemize}
-%We say that the well-formed word $\tau \in @S^*$ is \emph{sequential} when
-%\begin{itemize}
-%
-%  \item Operations do not overlap: \\
-%  $\tau_i, \tau_k \in C$ and $i < k$ implies $\tau_i \match \tau_j$ for some $i < j < k$.
-%
-%\end{itemize}
-%Well-formed words represent traces of a library. 
 We assume every set of 
+words is closed under isomorphic renaming of operation identifiers. 
+We denote the operation identifier of a
+call/return action $a$ by $\<op>(a)$. Call and return actions $c \in C$ and $r
+\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
+We may omit the second field from a call/return action $a$ for methods that have no inputs (e.g., the pop method of a stack) or return values (e.g., the push method of a stack).
+A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
+\emph{well formed} when:
+\begin{itemize}
+
+  \item Each return is preceded by a matching call: \\
+  $\tau_j \in R$ implies $\tau_i \match \tau_j$ for some $i < j$.
+
+  \item Each operation identifier is used in at most one call/return: \\
+  $\<op>(\tau_i) = \<op>(\tau_j)$ and $i < j$ implies $\tau_i \match \tau_j$.
+
+\end{itemize}
+We say that the well-formed word $\tau \in @S^*$ is \emph{sequential} when
+\begin{itemize}
+
+  \item Operations do not overlap: \\
+  $\tau_i, \tau_k \in C$ and $i < k$ implies $\tau_i \match \tau_j$ for some $i < j < k$.
+
+\end{itemize}
+Well-formed words represent traces of a library. We assume every set of well-formed
 words is closed under isomorphic renaming of operation identifiers. For
 notational convenience, we take $\<Ops>=\<Nats>$ for the rest of the paper.
-%When the value of a certain field in a call/return action is not important we use 
-%the placeholder $\_$, e.g., $inv(m,\_,k)$ instead of $inv(m,d,k)$ when the input  
-%$d$ can take any value.
+When the value of a certain field in a call/return action is not important we use 
+the placeholder $\_$, e.g., $inv(m,\_,k)$ instead of $inv(m,d,k)$ when the input  
+$d$ can take any value.
 
+An operation $k$ is called \emph{completed} in a well-formed trace $\tau$ when
+$ret(m,d,k)$ occurs in $\tau$, for some $m$ and $d$. Otherwise, it is called \emph{pending}.
 %An operation $o$ of an execution $e$ is \emph{completed}
@@
 
-A \emph{library} is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$. 
-We assume that the traces of a library satisfy standard well-formedness properties, 
-e.g., return actions correspond to previous call actions, which for lack of 
-space are delegated to Appendix~\ref{app:prelim}. An operation $k$ is called \emph{completed} in a trace $\tau$ when
-$ret(m,d,k)$ occurs in $\tau$, for some $m$ and $d$. Otherwise, it is called \emph{pending}.
+Libraries dictate the execution of methods between their call and return
+points. Accordingly, a library cannot prevent a method from being called,
+though it can decide not to return. Furthermore, any library action performed
+in the interval between call and return points can also be performed should the
+call have been made earlier, and/or the return made later. 
+A library thus allows any sequence of
+invocations to its methods made by \emph{some} program.
 
-%Libraries dictate the execution of methods between their call and return
-%points. Accordingly, a library cannot prevent a method from being called,
-%though it can decide not to return. Furthermore, any library action performed
-%in the interval between call and return points can also be performed should the
-%call have been made earlier, and/or the return made later. 
-%A library thus allows any sequence of
-%invocations to its methods made by \emph{some} program.
-%
-%\begin{definition}\label{def:libraries}
-%A \emph{library} $L$ is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$
-%and each trace $\tau \in Tr(L)$ is well formed, and
-%  \begin{itemize}
-%
-%    \item Call actions $c \in C$ cannot be disabled: \\
-%    $\tau \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot \tau' \in Tr(L)$
-%    if $\tau \cdot c \cdot \tau'$ is well formed.
-%  
-%    \item Call actions $c \in C$ cannot disable other actions: \\
-%    $\tau \cdot a \cdot c \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot a \cdot \tau' \in Tr(L)$.
-%  
-%    \item Return actions $r \in R$ cannot enable other actions: \\
-%    $\tau \cdot r \cdot a \cdot \tau' \in Tr(L)$ implies $\tau \cdot a \cdot r \cdot \tau' \in Tr(L)$.
-%  
-%  \end{itemize}
-%
-%\end{definition}
-%
-%%Note that even a library that implements \emph{atomic methods}, e.g.,~by
-%%guarding method bodies with a global-lock acquisition, admits executions in
-%%which method calls and returns overlap. 
-%For simplicity, Definition~\ref{def:libraries} assumes that every thread performs a single operation.
+\begin{definition}\label{def:libraries}
+A \emph{library} $L$ is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$
+and each trace $\tau \in Tr(L)$ is well formed, and
+  \begin{itemize}
+
+    \item Call actions $c \in C$ cannot be disabled: \\
+    $\tau \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot \tau' \in Tr(L)$
+    if $\tau \cdot c \cdot \tau'$ is well formed.
+
+    \item Call actions $c \in C$ cannot disable other actions: \\
+    $\tau \cdot a \cdot c \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot a \cdot \tau' \in Tr(L)$.
+
+    \item Return actions $r \in R$ cannot enable other actions: \\
+    $\tau \cdot r \cdot a \cdot \tau' \in Tr(L)$ implies $\tau \cdot a \cdot r \cdot \tau' \in Tr(L)$.
+
+  \end{itemize}
+
+\end{definition}
+
+Note that even a library that implements \emph{atomic methods}, e.g.,~by
+guarding method bodies with a global-lock acquisition, admits executions in
+which method calls and returns overlap. 
+For simplicity, Definition~\ref{def:libraries} assumes that every thread performs a single operation.
 The extension to multiple operations per thread is straightforward, e.g.
 the closure rules must assume that the actions $a$ and $c$ belong to different threads
-%%A library which accesses the client's
-%%thread identifiers can be modeled by taking thread identifiers as method
-%%parameters.
+%A library which accesses the client's
+%thread identifiers can be modeled by taking thread identifiers as method
+%parameters.
 
@@
 Ignore this if I am wrong.}
-The projection of a library trace over $C\cup R$ is called a \emph{history}. The set of histories of a library $L$ is denoted by $H(L)$.
-Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
-admit histories with weaker inter-operation ordering, in which calls may happen earlier, and/or returns later. A 
-history $h_1$ is \emph{weaker} than a history $h_2$, written $h_1 \sqsubseteq h_2$, 
-if{f} there exists a history $h_1'$
-obtained from $h_1$ by appending return actions, and deleting call actions,
-s.t.:
-  $h_2$ is a permutation of $h_1'$ that preserves the order between
-  return and call actions, i.e.,~if a given return action occurs before a given
-  call action in $h_1'$, then the same holds in $h_2$.
-%A call action $inv(\_,\_,k)$ may happen before a return action $ret(\_,\_,k')$ in $h_1$ (the operations $k$ and $k'$ overlapping in time) while it can be permuted after that return action in $h_2$, thus introducing a new inter-operation ordering (operation $k'$ ending before $k$).
-
-%$\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
-%history set $H$ under weakening.
-%We assume that $H(L)=\overline{H(L)}$ for
-%every library $L$. %A \emph{kernel} of L is any set $K$ such that $\overline{K} = H(L)$.
-A library $L$ is called \emph{atomic} when there exists a set $S$ of sequential histories such that 
-$H(L)$ contains every weakening of a history in $S$.
-Atomic libraries are often considered as specifications for concurrent objects. 
-Libraries can be made atomic by guarding their methods bodies with global lock acquisitions.
-
-A library $L$ is called a \emph{queue implementation} when $\<Methods>=\{enq,deq\}$ ($enq$ is the method that enqueues a value and $deq$ is the method removing a value) and $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$ where {\tt EMPTY} is the value returned by $deq$ when the queue is empty.
-Similarly, a library $L$ is called a \emph{stack implementation} when $\<Methods>=\{push,pop\}$ and  $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$.
-For queue and stack implementations, we assume that the same value is never added twice, i.e., for every trace $@t$ of such a library and every two call actions $inv(m,d_1,k_1)$ and $inv(m,d_2,k_2)$ where $m\in \{enq,push\}$ we have that $d_1\neq d_2$.
-As shown in several works~\cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, this assumption is without loss of generality for libraries that are data independent, i.e., their behaviors are not influenced by the values added to the collection, which is always the case in practice.
-On a technical note, this assumption is used to define ($\Gamma$-)deterministic abstract implementations of stacks and queues in Section~\ref{sec:queues} and Section~\ref{sec:stacks}.
 
@@
 Then, traces $e$ and $e'$ are equivalent if $e'$ is obtained from $e$ by replacing every action $inv(m,d,k)$ with $inv(m,d,f(k))$ and every action $ret(m,d,k)$ with $ret(m,d,f(k))$.
-
-\subsection{Refinement and Linearizability}
-
-%\textcolor{red}{A general organizational idea for  this subsection: Linearizability is defined in terms of weakening in which observational actions are just call and return events.
-Then, we put forward the idea of enriching $\Gamma$ by adding new actions.
-In the current organization, linearizability is defined after refinement before which this $\Gamma$ extension argument is explained.
-Hence, people may confuse that notion of linearizability is based on $\Gamma$ set we picked.
-I propose to first introduce linearizability.
-Then explain refinement and give Theorem 1.}
-Conformance of a library $L_1$ to a specification given as an ``abstract'' library $L_2$ 
-is formally captured by \emph{(observational) refinement}. Informally, given two libraries
-$L_1$ and $L_2$ implementing the methods of some concurrent object, we
-say $L_1$ refines $L_2$ if{f} every computation of every program
-using $L_1$ would also be possible were $L_2$ used instead. We assume that a program can 
-interact with the library only through call and return actions, and thus refinement can be defined
-as history set inclusion. Refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} 
-when $L_2$ is an atomic library~\cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}.
-. 
-%We fix a set of methods $\<Methods>$ and values $\<Vals>$.
-
-%\textcolor{red}{We have two Definition 1. The problem occurs because we used different commands for definitions.}
-\begin{definition}
-%Let $L_1$ and $L_2$ be two libraries. 
-%agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets.
-We define the set $A\Sigma = \{inv(m,d,i)| m \in \mathcal{M} \wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\} \cup \{ret(m,d,i)| m \in \mathcal{M}\wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\}$ as abstract transition labels.
-Note that $A\Sigma \subseteq \Sigma_{L_1}$ and $A\Sigma \subseteq \Sigma_{L_2}$.
-A library $L_1$ \emph{refines} another library $L_2$ if{f} $H(L_1) \subseteq H(L_2)$.
-%We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
-%for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
-\end{definition}
-
-%Linearizability is also a relation between two libraries and it is stricter than refinement.
-It requires $e'$ in Definition 2 to be a sequential one.
-A trace $e$ is sequential iff following two conditions hold for its projection to abstract transition labels $e|A\Sigma = e_1, ...e_n$: $(i)$ $e_1 = inv(m,d,k)$ for some $m \in \mathcal{M}, d \in \mathcal{D} and k \in \mathbb{N}$ and $(ii)$ for all $i \in [1,n)$, either $e_i = inv(m,d,k)$ and $e_{i+1} = ret(m,d',k)$ or $e_i = ret(m,d,k)$ and $e_{i+1} = inv(m',d',k')$ for some $m,m' \in \mathcal{M}$, $d,d' \in \mathcal{D}$ and $k,k' \in \mathbb{N}$.
-
-Linearizability~\cite{journals/toplas/HerlihyW90} requires that every history of a concurrent library $L_1$ can be 
-``linearized'' to a sequential history admitted by a library $L_2$ used as a specification. 
-Formally, a sequential history $h_2$ with only complete operations is called a \emph{linearization} of a history $h_1$ when $h_1 \sqsubseteq h_2$.
-A history $h_1$ is \emph{linearizable} w.r.t.~a library $L_2$ if{f} there exists a linearization $h_2$ of $h_1$ such that 
-$h_2 \in H(L_2)$. A library $L_1$
-is \emph{linearizable} w.r.t. $L_2$, written $L_1 \sqsubseteq L_2$, if{f}
-each history $h_1 \in H(L_1)$ is linearizable w.r.t. $L_2$.
-
-\begin{theorem}[\cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}]
-  $L_1 \sqsubseteq L_2$ if{f} $L_1$ refines $L_2$,
-  if $L_2$ is atomic.
-
-\end{theorem}
-
-In the rest of the paper, we discuss methods for proving refinement (and thus, linearizability) focusing mainly on queue and stack implementations.

### 2017-01-23 draftCAV/prelim-long.tex
a pass until stacks

--- 2017-01-23 draftCAV/prelim.tex
@@
 \emph{operation identifier} $k \in \<Ops>$. Operation identifiers are used to
-pair call and return actions. 
-%We denote the operation identifier of a
-%call/return action $a$ by $\<op>(a)$. Call and return actions $c \in C$ and $r
-%\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
-We may omit the second field from a call/return action $a$ for methods that have no inputs or return values.
-%A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
-%\emph{well formed} when:
-%\begin{itemize}
-%
-%  \item Each return is preceded by a matching call: \\
-%  $\tau_j \in R$ implies $\tau_i \match \tau_j$ for some $i < j$.
-%
-%  \item Each operation identifier is used in at most one call/return: \\
-%  $\<op>(\tau_i) = \<op>(\tau_j)$ and $i < j$ implies $\tau_i \match \tau_j$.
-%
-%\end{itemize}
-%We say that the well-formed word $\tau \in @S^*$ is \emph{sequential} when
-%\begin{itemize}
-%
-%  \item Operations do not overlap: \\
-%  $\tau_i, \tau_k \in C$ and $i < k$ implies $\tau_i \match \tau_j$ for some $i < j < k$.
-%
-%\end{itemize}
-%Well-formed words represent traces of a library. 
-We assume every set of 
+pair call and return actions. We denote the operation identifier of a
+call/return action $a$ by $\<op>(a)$. Call and return actions $c \in C$ and $r
+\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
+We may omit the second field from a call/return action $a$ for methods that have no inputs (e.g., the pop method of a stack) or return values (e.g., the push method of a stack).
+A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
+\emph{well formed} when:
+\begin{itemize}
+
+  \item Each return is preceded by a matching call: \\
+  $\tau_j \in R$ implies $\tau_i \match \tau_j$ for some $i < j$.
+
+  \item Each operation identifier is used in at most one call/return: \\
+  $\<op>(\tau_i) = \<op>(\tau_j)$ and $i < j$ implies $\tau_i \match \tau_j$.
+
+\end{itemize}
+We say that the well-formed word $\tau \in @S^*$ is \emph{sequential} when
+\begin{itemize}
+
+  \item Operations do not overlap: \\
+  $\tau_i, \tau_k \in C$ and $i < k$ implies $\tau_i \match \tau_j$ for some $i < j < k$.
+
+\end{itemize}
+Well-formed words represent traces of a library. We assume every set of well-formed
 words is closed under isomorphic renaming of operation identifiers. For
 notational convenience, we take $\<Ops>=\<Nats>$ for the rest of the paper.
-%When the value of a certain field in a call/return action is not important we use 
-%the placeholder $\_$, e.g., $inv(m,\_,k)$ instead of $inv(m,d,k)$ when the input  
-%$d$ can take any value.
+When the value of a certain field in a call/return action is not important we use 
+the placeholder $\_$, e.g., $inv(m,\_,k)$ instead of $inv(m,d,k)$ when the input  
+$d$ can take any value.
 
+An operation $k$ is called \emph{completed} in a well-formed trace $\tau$ when
+$ret(m,d,k)$ occurs in $\tau$, for some $m$ and $d$. Otherwise, it is called \emph{pending}.
 %An operation $o$ of an execution $e$ is \emph{completed}
@@
 
-A \emph{library} is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$. 
-We assume that the traces of a library satisfy standard well-formedness properties, 
-e.g., return actions correspond to previous call actions, which for lack of 
-space are delegated to Appendix~\ref{app:prelim}. An operation $k$ is called \emph{completed} in a trace $\tau$ when
-$ret(m,d,k)$ occurs in $\tau$, for some $m$ and $d$. Otherwise, it is called \emph{pending}.
+Libraries dictate the execution of methods between their call and return
+points. Accordingly, a library cannot prevent a method from being called,
+though it can decide not to return. Furthermore, any library action performed
+in the interval between call and return points can also be performed should the
+call have been made earlier, and/or the return made later. 
+A library thus allows any sequence of
+invocations to its methods made by \emph{some} program.
 
-%Libraries dictate the execution of methods between their call and return
-%points. Accordingly, a library cannot prevent a method from being called,
-%though it can decide not to return. Furthermore, any library action performed
-%in the interval between call and return points can also be performed should the
-%call have been made earlier, and/or the return made later. 
-%A library thus allows any sequence of
-%invocations to its methods made by \emph{some} program.
-%
-%\begin{definition}\label{def:libraries}
-%A \emph{library} $L$ is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$
-%and each trace $\tau \in Tr(L)$ is well formed, and
-%  \begin{itemize}
-%
-%    \item Call actions $c \in C$ cannot be disabled: \\
-%    $\tau \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot \tau' \in Tr(L)$
-%    if $\tau \cdot c \cdot \tau'$ is well formed.
-%  
-%    \item Call actions $c \in C$ cannot disable other actions: \\
-%    $\tau \cdot a \cdot c \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot a \cdot \tau' \in Tr(L)$.
-%  
-%    \item Return actions $r \in R$ cannot enable other actions: \\
-%    $\tau \cdot r \cdot a \cdot \tau' \in Tr(L)$ implies $\tau \cdot a \cdot r \cdot \tau' \in Tr(L)$.
-%  
-%  \end{itemize}
-%
-%\end{definition}
-%
-%%Note that even a library that implements \emph{atomic methods}, e.g.,~by
-%%guarding method bodies with a global-lock acquisition, admits executions in
-%%which method calls and returns overlap. 
-%For simplicity, Definition~\ref{def:libraries} assumes that every thread performs a single operation.
+\begin{definition}\label{def:libraries}
+A \emph{library} $L$ is an LTS over alphabet $\Sigma$ such that $C \u R\subseteq \Sigma$
+and each trace $\tau \in Tr(L)$ is well formed, and
+  \begin{itemize}
+
+    \item Call actions $c \in C$ cannot be disabled: \\
+    $\tau \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot \tau' \in Tr(L)$
+    if $\tau \cdot c \cdot \tau'$ is well formed.
+
+    \item Call actions $c \in C$ cannot disable other actions: \\
+    $\tau \cdot a \cdot c \cdot \tau' \in Tr(L)$ implies $\tau \cdot c \cdot a \cdot \tau' \in Tr(L)$.
+
+    \item Return actions $r \in R$ cannot enable other actions: \\
+    $\tau \cdot r \cdot a \cdot \tau' \in Tr(L)$ implies $\tau \cdot a \cdot r \cdot \tau' \in Tr(L)$.
+
+  \end{itemize}
+
+\end{definition}
+
+%Note that even a library that implements \emph{atomic methods}, e.g.,~by
+%guarding method bodies with a global-lock acquisition, admits executions in
+%which method calls and returns overlap. 
+For simplicity, Definition~\ref{def:libraries} assumes that every thread performs a single operation.
 The extension to multiple operations per thread is straightforward, e.g.
 the closure rules must assume that the actions $a$ and $c$ belong to different threads
-%%A library which accesses the client's
-%%thread identifiers can be modeled by taking thread identifiers as method
-%%parameters.
+%A library which accesses the client's
+%thread identifiers can be modeled by taking thread identifiers as method
+%parameters.
 
@@
 
-%$\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
-%history set $H$ under weakening.
-%We assume that $H(L)=\overline{H(L)}$ for
-%every library $L$. %A \emph{kernel} of L is any set $K$ such that $\overline{K} = H(L)$.
-A library $L$ is called \emph{atomic} when there exists a set $S$ of sequential histories such that 
-$H(L)$ contains every weakening of a history in $S$.
+$\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
+history set $H$ under weakening.
+A direct consequence of Definition~\ref{def:libraries} is that $H(L)=\overline{H(L)}$ for
+every library $L$. %A \emph{kernel} of L is any set $K$ such that $\overline{K} = H(L)$.
+A library $L$ is called atomic if $H(L)=\overline{S}$ for some set of sequential histories $S$.
 Atomic libraries are often considered as specifications for concurrent objects. 
@@
 
-A library $L$ is called a \emph{queue implementation} when $\<Methods>=\{enq,deq\}$ ($enq$ is the method that enqueues a value and $deq$ is the method removing a value) and $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$ where {\tt EMPTY} is the value returned by $deq$ when the queue is empty.
+A library $L$ is called a \emph{queue implementation} when its alphabet contains call and return actions defined over $\<Methods>=\{enq,deq\}$ ($enq$ is the name of the methods that enqueue a value, and $deq$ is the name of the methods that dequeue a value) and $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$ where {\tt EMPTY} is the value returned by $deq$ when the queue is empty.
 Similarly, a library $L$ is called a \emph{stack implementation} when $\<Methods>=\{push,pop\}$ and  $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$.

### 2017-01-23 draftCAV/intro.tex
intro sketch

@@
 
+Efficient implementations of concurrent objects such as semaphores, locks, and atomic collections including stacks and queues are vital to modern computer systems.
+Programming them is however error prone.
+To minimize synchronization overhead between concurrent object-method invocations, implementors avoid blocking operations like lock acquisition, allowing methods to execute concurrently.
+However, concurrency risks unintended inter-operation interference, and risks conformance to atomic reference implementations.
+Conformance is formally captured by \emph{observational refinement}, which assures that all behaviors of programs
+using these efficient implementations would also be possible were the atomic
+reference implementations used instead.
+An equivalent property is \emph{linearizability}~\cite{journals/toplas/HerlihyW90} which requires that each concurrent execution
+has a linearization which is a valid sequential execution according to a
+specification, given by an abstract data type or reference implementation.
+
+Existing proof methods for refinement/linearizability combine reductions to assertion checking with compositional reasoning techniques for proving the assertions.
+
+The reductions to assertion checking are based on simulation relations: the assertions describe a simulation relation to standard atomic reference implementations
+
+No surprise, simulations are the standard approach for proving refinement
+
+This works smoothly for particular classes of implementations, e.g., with fixed linearization points.
+The explanation is that in this case, forward simulations are enough and the reference implementation is deterministic when the linearization points are observable
+
+Why they are enough ? Forward simulations are not complete for proving refinement: the reference implementation is not deterministic when projected over call and returns.
+
+However, in this case, the set of observable actions includes the linearization points. And the reference implementations becomes deterministic for this larger set of observable actions.
+
+Determinism also simplifies the assertions
+
+When linearization points are not observable, reference implementations are not deterministic and one may need backward simulations.
+There is a paper arguing for this fact ~\cite{DBLP:conf/cav/SchellhornWD12}.
+
+However, using backward simulations is hard, they require prophecy variables. They are not good for automation.
+
+Question: We want to avoid backward simulations, so what can we do about implementations without fixed linearization points, e.g, Herlihy\&Wing, Time-Stamped Stack ?
+Can they be proved using forward simulations ?
+
+We answer positively for queues and stacks.
+The idea is to define new abstract implementations over a set of observable actions bigger than the set of call/returns, but smaller than all linearization points.
+
+Queues: only dequeues have fixed linearization points.
+
+Stacks: pops have fixed commit points
+
 \newpage~\newpage

### 2017-01-23 draftCAV/queue.tex
intro sketch

@@
 
-To prove monotonicity in the case of dequeue internal steps different from its linearization point, it is important to track the non-trivial instantiations of item (c) in the definition of $<$ over the two states $s$ and $s'$., i.e., the triples $(k,k',k_d)$ for which (\ref{eq:inst}) holds.
+To prove monotonicity in the case of dequeue internal steps different from its linearization point, it is important to track the non-trivial instantiations of item (c) in the definition of $<$ over the two states $s$ and $s'$, i.e., the triples $(k,k',k_d)$ for which (\ref{eq:inst}) holds.
 Instantiations that are enabled only in $s'$ may in principle lead to a violation of monotonicity (since they restrict the orders $<$ associated to $s'$).
@@
 %
-The same is true for the increment of the iterator {\tt i} in a dequeue $k_d$ since the predicate $k_d\atCP {\tt if}\text{-}{\tt inc}$ holds in state $s$.
+The same is true for the increment of {\tt i} in a dequeue $k_d$ since the predicate $k_d\atCP {\tt if}\text{-}{\tt inc}$ holds in state $s$.
 The execution of {\tt swap} returning {\tt null} in a dequeue $k_d$ enables new instantiations $(k,k',k_d)$ in $s'$, thus adding potentially new constraints $k\not< k'$.
-We show however that these instantiations are vacuous because $k$ must be a pending enqueue in $s$ and thus maximal in any order relation $<$ associated by $\mathit{fs}_1$ to $s$.
-Let $k$ and $k'$ be two enqueue operations such that together with the dequeue $k_d$ they satisfy the property (\ref{eq:inst}) in $s'$ but not in $s$. 
-We write ${\tt i}_s(k)$ to refer to the value of the variable {\tt i} of operation $k$ in state $s$. 
-This implies that ${\tt i}_{s'}(k) = {\tt i}_{s'}(k_d) \leq {\tt i}_{s'}(k')$ and ${\tt items}[{\tt i}_{s'}(k_d)]={\tt null}$. The latter implies that the enqueue $k$ didn't executed
-the second statement and it is pending (because the position it reserved is still {\tt null}).
-Finally, checking whether the value returned by {\tt swap} is {\tt null} doesn't modify any of the variables in property  (\ref{eq:inst}) and also, it doesn't change the valuation of the predicate $\atCP {\tt if}\text{-}{\tt inc}$.
+We show that these instantiations are however vacuous because $k$ must be pending in $s$ and thus maximal in every order $<$ associated by $\mathit{fs}_1$ to $s$.
+Let $k$ and $k'$ be two enqueues such that together with the dequeue $k_d$ they satisfy the property (\ref{eq:inst}) in $s'$ but not in $s$. 
+We write ${\tt i}_s(k)$ for the value of the variable {\tt i} of operation $k$ in state $s$. 
+We have that ${\tt i}_{s'}(k) = {\tt i}_{s'}(k_d) \leq {\tt i}_{s'}(k')$ and ${\tt items}[{\tt i}_{s'}(k_d)]={\tt null}$. The latter implies that the enqueue $k$ didn't executed
+the second statement (since the position it reserved is still {\tt null}) and it is pending in $s$.
+The step that checks that the value returned by {\tt swap} is {\tt null} doesn't modify the variables in property  (\ref{eq:inst}) and also, it doesn't change the valuation of the predicate $\atCP {\tt if}\text{-}{\tt inc}$.
 
 Finally, we show that the linearization point of a dequeue $k$ of $\mathit{HWQ}$, i.e., an execution of {\tt swap} returning a non-null value $d$, from state $s$ and leading to a state $s'$ is simulated by a transition labeled by $lin(deq,d,k)$ of $AbsQ$ from state $t$.
-By the definition of $\mathit{HWQ}$, there exists a unique enqueue $k_e$ which filled the position updated by $k$, i.e., ${\tt i}_s(k_e)=i_s(k)$ and ${\tt x}_{s'}(k)={\tt x}_s(k_e)$.
-We show that $k_e$ is minimal in the order relation $<$ of $t$ which implies that $lin(deq,d,k)$ is enabled in $t$.
+By the definition of $\mathit{HWQ}$, there is a unique enqueue $k_e$ which filled the position updated by $k$, i.e., ${\tt i}_s(k_e)=i_s(k)$ and ${\tt x}_{s'}(k)={\tt x}_s(k_e)$.
+We show that $k_e$ is minimal in the order $<$ of $t$ which implies that $lin(deq,d,k)$ is enabled in $t$.
 Thus, instantiating item (c) in the definition of $<$ with $k'=k_e$ and $k_d=k$ we get that every enqueue that reserved a position smaller than the one of $k_e$ can't be ordered before $k_e$ in the order $<$.
 Also, applying item (b) with $k=k_e$ we get the same for every enqueue that reserved a bigger position.
-An enqueue that didn't reserved a position is by definition maximal in the order $<$ and therefore, not a predecessor of $k_e$.
-Then, the state $t'$ obtained from $t$ through a $lin(deq,d,k)$ transition is related to $s'$ because essentially, (1) the value added by $k_e$ is not anymore present in {\tt items} which implies that $k_e$ doesn't occur in any $AbsQ$ state related to $s'$, and (2) the value of ${\tt x}(k)$ is set to $d\neq {\tt null}$ which implies that $rv(k)$ is set to $d$ in every $AbsQ$ state related to $s'$.
+An enqueue that didn't reserved a position is by definition maximal in $<$ and therefore, not a predecessor of $k_e$.
+Then, the state $t'$ obtained from $t$ through a $lin(deq,d,k)$ transition is related to $s'$ because (1) the value added by $k_e$ is not anymore present in {\tt items} which implies that $k_e$ doesn't occur in any $AbsQ$ state related to $s'$, and (2) the value of ${\tt x}(k)$ is set to $d\neq {\tt null}$ which implies that $rv(k)$ is set to $d$ in every $AbsQ$ state related to $s'$.
 %

### 2017-01-23 draftCAV/stack.tex
intro sketch

@@
 \vspace{-1.5mm}
-While the abstract queue implementation in Section~\ref{sec:queues} can be adapted to stacks~\footnote{The only modification is that the linearization point $lin(pop,d,k)$ with $d\neq{\tt EMPTY}$ is enabled only if $k$ is added by a push which is maximal in the happens-before order stored in the state.}, it can not simulate (through forward simulations) existing stack implementations like the Time-stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} where the linearization points of the pop operations are not fixed.
-In this section, we define an abstract implementation which simulates such implementations provided that the point in time where the return value of a pop operation is determined corresponds to a fixed action.
+While the abstract queue in Section~\ref{sec:queues} can be adapted to stacks (the linearization point $lin(pop,d,k)$ with $d\neq{\tt EMPTY}$ is enabled when $k$ is added by a push which is maximal in the happens-before order stored in the state), it can't simulate (through forward simulations) existing stack implementations like the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) where the linearization points of the pop operations are not fixed.
+Exploiting particular properties of the stack semantics, we refine the ideas used in $AbsQ$ and define
+a new abstract implementation for stacks, denoted as $AbsQ$, which is able to simulate such implementations.
+Forward simulations to $AbsS$ are complete for proving the correctness of stack implementations provided that the point in time where the return value of a pop operation is determined, called \emph{commit point}, corresponds to a fixed action.
+% provided that the point in time where the return value of a pop operation is determined corresponds to a fixed action.
 To demonstrate the use of this abstract implementation we provide a correctness proof for a simplified version of Time-Stamped Stack that preserves its most complex behaviors.
@@
 \vspace{-3.5mm}
-\subsection{Abstract stack implementation}
-\vspace{-1.5mm}
-\begin{wrapfigure}{l}{5.4cm}
-\vspace{-11mm}
+\subsection{Pop Methods With Fixed Commit Points}
+\vspace{-1mm}
+\begin{wrapfigure}{l}{5.2cm}
+\vspace{-10.5mm}
 \begin{lstlisting}
@@
 \vspace{-6mm}
-\caption{Time-Stamped Stack. We assume that every statement is atomic.}
+\caption{Time-Stamped Stack.} %We assume that every statement is atomic.}
 \label{fig:TimeStamped}
-\vspace{-8mm}
+\vspace{-7mm}
 \end{wrapfigure}
 We explain the meaning of the commit points on a simplified version of the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) given in Figure~\ref{fig:TimeStamped}.
-This stack implementation maintains a vector of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
+This  implementation maintains an array of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
 Initially, each list contains a sentinel dummy node pointing to itself with timestamp $-1$ and the flag {\tt taken} set to {\tt false}.
 
-Pushing a value to the stack proceeds in several steps: adding a node with maximal timestamp in the list associated to the thread executing the push (given by the special variable {\tt myTID}), asking for a new timestamp (given by the value of the shared integer {\tt TS}), and updating the timestamp of the added node.
+Pushing a value to the stack proceeds in several steps: adding a node with maximal timestamp in the list associated to the thread executing the push (given by the special variable {\tt myTID}), asking for a new timestamp (given by the shared variable {\tt TS}), and updating the timestamp of the added node.
 Popping a value from the stack consists in traversing all the lists, finding the first element which doesn't represent a removed value (i.e., {\tt taken} is {\tt false}) in each list, and selecting the element with the maximal timestamp.
@@
 \centering
-\includegraphics[width=11cm]{fig-timeStamp.pdf}
-\vspace{-4.5mm}
-\caption{An execution of the Time-Stamped Stack.
+\includegraphics[width=11.5cm]{fig-tss.pdf}
+\vspace{-4mm}
+\caption{An execution of $\mathit{TSS}$.
 An operation is pictured by a line delimited by two circles denoting the call and respectively, the return action.
-Pop operations with identifier $k$ and removing value $d$ are denoted by $pop(d,k)$ and their representation includes another circle to represent a CAS returning {\tt true} which is the commit point of that operation.
-The state of the library after an execution prefix delimited at the right by a dotted line is pictured in the bottom part (the picture immediately to the left of the dotted line).
-A pair $(d,t)$ for some value $d$ and integer $t$ represents a list node with ${\tt data}=d$ and ${\tt ts}=t$, and ${\tt i}_1$ denotes the value of {\tt i} in the pop operation with identifier 1.
-We omit the nodes where {\tt taken} is set to {\tt true}.}
+Pop operations with identifier $k$ and removing value $d$ are labeled $pop(d,k)$.
+Their representation includes another circle that stands for a successful CAS which is their commit point.
+The library state after an execution prefix delimited at the right by a dotted line is pictured in the bottom part (the picture immediately to the left of the dotted line).
+A pair $(d,t)$ represents a list node with ${\tt data}=d$ and ${\tt ts}=t$, and ${\tt i}(1)$ denotes the value of {\tt i} in the pop with identifier 1.
+We omit the nodes where the field {\tt taken} is {\tt true}.}
 \label{fig:commit}
-\vspace{-7mm}
+\vspace{-6mm}
 \end{figure}
@@
 Also, the value added by a push that just added an element to a list can be popped before the value added by a completed push (since it has a maximal timestamp).
-The same holds for pop operations.
-The only reasonable choice for a linearization point is a successful CAS (that results in updating the field {\tt taken}).
-Fig~\ref{fig:commit} pictures an execution showing that this action doesn't correspond to a linearization point, i.e., an execution for which the pop operations in every correct linearization are not ordered according to the order between successful CASs.
+The same holds for pop operations: The only reasonable choice for a linearization point is a successful CAS (that results in updating the field {\tt taken}).
+Fig.~\ref{fig:commit} pictures an execution showing that this action doesn't correspond to a linearization point, i.e., an execution for which the pop operations in every correct linearization are not ordered according to the order between successful CASs.
 In every correct linearization of that execution, the pop operation removing $x$ is ordered before the one removing $z$ although they perform a successful CAS in the opposite order.
@@
 
-When the commit points of pop operations are fixed to particular implementation actions (e.g., a successful CAS) we assume that the library is an LTS over an alphabet that contains actions $com(pop,d,k)$ with $d\in\<Vals>$ and $k\in\<Ops>$.
-This action represents the commit point of the pop with identifier $k$ and returning value $d$.
-Let $Com(pop)$ denote the set of such actions.
+When the commit points of pop operations are fixed to particular implementation actions (e.g., a successful CAS) we assume that the library is an LTS over an alphabet that contains actions $com(pop,d,k)$ with $d\in\<Vals>$ and $k\in\<Ops>$ (denoting the commit point of the pop with identifier $k$ and returning $d$).
+Let $Com(pop)$ be the set of such actions.
 %We consider a set of actions $Com(pop)=\set{com(pop,d,k):d\in\<Vals>, k\in\<Ops>}$, called \emph{commit points}, representing actions of a pop operation where its return value is determined. 
@@
 
-We define an abstract stack implementation $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the push operations whose value has not been yet removed.
+\vspace{-3mm}
+\subsection{Abstract stack implementation}
+\vspace{-1mm}
+We define an abstract stack $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the push operations whose value has not been yet removed.
 However, since the commit points are not necessarily linearization points, the pop operations are treated differently.
 Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order, and continuously tracks the push operations which are overlapping with it.
-The commit point $com(pop,d,k)$ with $d\neq {\tt EMPTY}$ is enabled only if $d$ was added by one of the push operations in the initial snapshot, or by a push happening earlier when all the values from the initial snapshot have been removed by other pops, or by one of the push operations that overlaps with pop $k$.
-The commit point $com(pop,{\tt EMPTY},k)$ is enabled only if all the values added by push operations ending before $k$ started have been removed (checking this condition relies again on the initial snapshot).
+The commit point $com(pop,d,k)$ with $d\neq {\tt EMPTY}$ is enabled only if $d$ was added by one of the push operations in the initial snapshot, or by a push happening earlier when all the values from the initial snapshot have been removed, or by one of the push operations that overlaps with pop $k$.
+The commit point $com(pop,{\tt EMPTY},k)$ is enabled only if all the values added by push operations ending before $k$ started have been removed.
 The effect of the commit points is explained below through examples.
@@
 \begin{wrapfigure}{l}{6.8cm}
-\vspace{-8mm}
+\vspace{-7mm}
 \includegraphics[width=7cm]{fig-stack.pdf}
@@
 %\includegraphics[width=7cm]{fig-queue2.pdf}
-\vspace{-5mm}
+\vspace{-7mm}
 \caption{Simulating stack histories with $AbsS$.}
 \label{fig:stackSim}
-\vspace{-6mm}
+\vspace{-5.5mm}
 \end{wrapfigure}
@@
 For readability, we give the state of $AbsS$ only after several execution prefixes delimited at the right by a dotted line.
-We focus on pop operations because the effect of push invocations and returns is similar to enqueue invocations and returns in $AbsQ$.
+We focus on pop operations because the effect of push calls and returns is similar to enqueue calls and returns in $AbsQ$.
 Let us first consider the history in the top part of the figure.
-The first $AbsS$ state we give for this history is reached after the call of the pop operation with identifier $3$.
+The first state we give for this history is reached after the call of the pop operation with identifier $3$.
 This shows the effect of a pop invocation: the maximal completed pushes according to the current happens-before (in this case the push with identifier $1$) are marked as $be(3)$ (from ``before'' operation 3), and the pending pushes are marked as $ov(3)$ (from ``overlapping'' with operation 3).
@@
 For the second pop, the commit point $com(pop,y,4)$ is enabled because $y$ was added by a push marked as $ov(4)$.
-The second execution shows an example where the marking $be(k)$ for some pop $k$ can be updated at commit points.
+The second execution shows an example where the marking $be(k)$ for some pop $k$ is updated at commit points.
 When the two pop operations $5$ and $6$ start, the pushes $3$ and $4$ are marked as $be(5)$ and $be(6)$.
@@
 Therefore, push 2, a predecessor of the push which is removed, is marked as $be(6)$.
-The other predecessor of the removed push, i.e., push $1$, is not marked as $be(6)$ because it happens before another push, i.e., push 3, which is already marked as $be(6)$ (the value added by push 3 should be removed as well before the value added by push 1 could become the top of the stack).
+The other predecessor of the removed push, i.e., push $1$, is not marked as $be(6)$ because it happens before another push, i.e., push 3, which is already marked as $be(6)$ (the value added by push 3 should be removed before the value added by push 1 could become the top of the stack).
 
-The states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $<$ is a strict partial order over the set $O$ of operation identifiers, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a pending/completed flag, $rv:\<Ops> ~> \<Vals>$ records the return value of a pending pop fixed at its commit point, $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A_2$) or pop ($R_1,R_2,R_3$) operation, $be:\<Ops> ~> 2^O$ records push operations that were maximal before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
-The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.
+Formally, the states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $<$ is a strict partial order over the set $O$ of operation identifiers, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a pending/completed flag, $rv:\<Ops> ~> \<Vals>$ records the return value of a pending pop fixed at its commit point, $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A_2$) or pop operation ($R_1,R_2,R_3$), $be:\<Ops> ~> 2^O$ records push operations that were maximal before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
+%The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.
+All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsS}.
 
@@
 \begin{figure} [t]
+\vspace{-2mm}
 {\scriptsize
@@
     \inferrule[call-pop]{
-      k\not\in dom(cp) \\
+      k\not\in dom(cp) %\\ k\not\in dom(cp) \\ k\not\in dom(cp)
     }{
@@
 We use the following notions: $maxCo(O)$ is the set of greatest operations in $O$ (w.r.t.
-$<$) which are completed, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.%\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise.
+$<$) which are completed, i.e., $maxCo(O)=\set{k\in O: \ell_2(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_2(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.%\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise.
 If I understand correctly, there is a problem change in $be$ in lin-pop1 rule.
@@
   \label{fig:transitions:AbsS}
-\vspace{-6mm}
+\vspace{-5mm}
 \end{figure}
 
-Let $AbsS_0$ be the standard abstract implementation of a concurrent stack (where elements are stored in a sequence, a push operation adds atomically an element at the beginning of the sequence, and a pop operation removes atomically an element from the beginning of the sequence).
+Let $AbsS_0$ be the standard abstract implementation of a stack (where elements are stored in a sequence; push, resp., pop operations add, resp., remove, an element from the beginning of the sequence in one atomic step).
 For $\<Methods>=\{push,pop\}$, the alphabet of $AbsS_0$ is $C\cup R\cup Lin$.
@@
 
+\vspace{-1mm}
 \begin{theorem}\label{th:absImplStack}
 $AbsS$ is a refinement of $AbsS_0$ and vice-versa.
+\vspace{-1mm}
 \end{theorem}
 
-A trace of a queue implementation is called \emph{$Com(pop)$-complete} when every completed pop has a commit point, i.e., TODO.
-A stack implementation $L$ over alphabet $\Sigma$ is called \emph{with fixed pop commit points} if{f} $C\cup R\cup Com(pop)\subseteq \Sigma$
-and every trace $@t\in Tr(L)$ is $Com(pop)$-complete.
+A trace of a stack implementation is called \emph{$Com(pop)$-complete} when every completed pop has a commit point, i.e., each return $ret(pop,d,k)$ is preceded by an action $com(pop,d,k)$.
+A stack implementation $L$ over $\Sigma$, such that $C\cup R\cup Com(pop)\subseteq \Sigma$, is called \emph{with fixed pop commit points} when every trace $@t\in Tr(L)$ is $Com(pop)$-complete.
 
@@
 
+\vspace{-1.5mm}
 \begin{corollary}
 A stack implementation $L$ with fixed pop commit points is a $C\cup R\cup Com(pop)$-refinement of $AbsS$ if{f} there exists a $C\cup R\cup Com(pop)$-forward simulation from $L$ to $AbsS$.
+\vspace{-1.5mm}
 \end{corollary}
 
+\vspace{-3.5mm}
 \subsection{A Correctness Proof For Time-Stamped Stack}
-
+\vspace{-1mm}
 We describe a forward simulation $\mathit{fs}_2$ from $\mathit{TSS}$ to $AbsS$.

### 2017-01-24 draftCAV/queue.tex
spaces

@@
 \vspace{-1.5mm}
-The typical abstract implementation of a concurrent queue, denoted as $AbsQ_0$, maintains a sequence of values, the enqueue method adds a value atomically to the beginning of the sequence, and the dequeue method removes a value from the end of the sequence (if any, otherwise it returns {\tt EMPTY}).
-Both methods have a fixed linearization point when the update of the shared sequence happens.
+The typical abstract implementation of a concurrent queue, denoted as $AbsQ_0$, maintains a sequence of values, the enqueue adds a value atomically to the beginning of the sequence, and the dequeue removes a value from the end of the sequence (if any, otherwise it returns {\tt EMPTY}).
+Both methods have a fixed linearization point when the update of the sequence happens.
 %~\footnote{For $\<Methods>=\{enq,deq\}$, the alphabet of $AbsQ_0$ is $C\cup R\cup Lin$.}.
 For some queue implementations, e.g., the Herlihy\&Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), there exists no forward simulation to $AbsQ_0$ although they are a refinement of $AbsQ_0$.
-The main reason is that the enqueue methods in these implementations don't have a \emph{fixed} linearization point.
+The main reason is that the enqueue methods don't have a \emph{fixed} linearization point.
 In this section, we propose a new abstract implementation for queues, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence.
 We show that there exists a forward simulation from any correct queue implementation where only the \emph{dequeue} methods have fixed linearization points (the enqueue methods are unconstrained) to $AbsQ$.
-This covers all the queue implementations that we are aware of, in particular $\mathit{HWQ}$.
+This covers all the queue implementations that we are aware of, in particular $\mathit{HWQ}$, Baskets Queue~\cite{DBLP:conf/opodis/HoffmanSS07}, LCRQ~\cite{DBLP:conf/ppopp/MorrisonA13}, or Time-Stamped Queue~\cite{DBLP:conf/popl/DoddsHK15} (where the enqueues don't have fixed linearization points).
 We also describe a forward simulation from $\mathit{HWQ}$ to $AbsQ$.
 
-TODO HOW CAN WE CONVINCE THAT OTHER IMPLEMENTATIONS HAVE THE SAME PROPERTY ?
+%TODO HOW CAN WE CONVINCE THAT OTHER IMPLEMENTATIONS HAVE THE SAME PROPERTY ?
 \vspace{-3.5mm}
 \subsection{Enqueue Methods With Non-Fixed Linearization Points}
-\vspace{-1.5mm}
+\vspace{-1mm}
 % describe a queue implementation listed in Figure~\ref{fig:HerlihyWing} and known as the Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), where only the dequeue methods have fixed linearization points.
 \begin{wrapfigure}{l}{5.3cm}
-\vspace{-8mm}
+\vspace{-10mm}
 \begin{lstlisting}
@@
 \end{wrapfigure}
-We describe $\mathit{HWQ}$ as an instance of a queue implementation where the linearization points of the enqueue methods are not fixed.
+We describe $\mathit{HWQ}$ which is a queue implementation where the linearization points of the enqueue methods are not fixed.
 The shared state consists of an array {\tt items} storing the values in the queue and a counter {\tt back} storing the index of the first unused position in {\tt items}.
@@
 This action alone contributes to the effect of that value being removed from the queue.
-This claim will be formally proved in Section~\ref{ssec:HerlihyWing}.
+Every concurrent history can be linearized to a sequential history where dequeues occur in the order of their linearization points in the concurrent history.
+This claim is formally proved in Section~\ref{ssec:HerlihyWing}.
 

### 2017-01-24 draftCAV/stack.tex
spaces

@@
 
-TODO WHAT CAN WE SAY ABOUT COMMIT POINTS IN OTHER IMPLEMENTATIONS
+%TODO WHAT CAN WE SAY ABOUT COMMIT POINTS IN OTHER IMPLEMENTATIONS
 
@@
 We define an abstract stack $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the push operations whose value has not been yet removed.
-However, since the commit points are not necessarily linearization points, the pop operations are treated differently.
+Pop operations are treated differently since the commit points are not necessarily linearization points.
 Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order, and continuously tracks the push operations which are overlapping with it.
@@
 The effect of the commit points is explained below through examples.
+\vspace{-.4mm}
 
 \begin{wrapfigure}{l}{6.8cm}
-\vspace{-7mm}
+\vspace{-6mm}
 \includegraphics[width=7cm]{fig-stack.pdf}
@@
 %\includegraphics[width=7cm]{fig-queue2.pdf}
-\vspace{-7mm}
+\vspace{-8mm}
 \caption{Simulating stack histories with $AbsS$.}
 \label{fig:stackSim}
-\vspace{-5.5mm}
+\vspace{-6.5mm}
 \end{wrapfigure}
@@
 For readability, we give the state of $AbsS$ only after several execution prefixes delimited at the right by a dotted line.
-We focus on pop operations because the effect of push calls and returns is similar to enqueue calls and returns in $AbsQ$.
-Let us first consider the history in the top part of the figure.
-The first state we give for this history is reached after the call of the pop operation with identifier $3$.
-This shows the effect of a pop invocation: the maximal completed pushes according to the current happens-before (in this case the push with identifier $1$) are marked as $be(3)$ (from ``before'' operation 3), and the pending pushes are marked as $ov(3)$ (from ``overlapping'' with operation 3).
+We focus on pop operations -- the effect of push calls and returns is similar to enqueue calls and returns in $AbsQ$.
+Let us first consider the history on the top part.
+The first state we give is reached after the call of pop with identifier $3$.
+This shows the effect of a pop invocation: the maximal completed pushes according to the current happens-before (here, the push with identifier $1$) are marked as $be(3)$ (from ``before'' operation 3), and the pending pushes are marked as $ov(3)$ (from ``overlapping'' with operation 3).
 As a side remark, any other push operation that starts after pop $3$ would be also marked as $ov(3)$.
 The commit point $com(pop,x,3)$ (pictured with a red circle) is enabled because $x$ was added by a push marked as $be(3)$.
-In this case, the effect of the commit point is that the push $1$ is removed from the happens-before (a more complicated case can be seen in the second execution).
+The effect of the commit point is that push $1$ is removed from the state (the execution on the bottom shows a more complicated case).
 For the second pop, the commit point $com(pop,y,4)$ is enabled because $y$ was added by a push marked as $ov(4)$.
-The second execution shows an example where the marking $be(k)$ for some pop $k$ is updated at commit points.
-When the two pop operations $5$ and $6$ start, the pushes $3$ and $4$ are marked as $be(5)$ and $be(6)$.
-Then, $com(pop,t,5)$ is enabled because $t$ was added by $push(t,4)$ which is marked as $be(5)$.
-Besides removing $push(t,4)$ from the state, the commit point must produce a state where a pop committing later, e.g., pop $6$, can remove $y$ which was added by a predecessor of $push(t,4)$ in the happens-before ($y$ could become the top of the stack when $t$ is removed).
-This history remains valid because $push(y,2)$ can be linearized after $push(x,1)$ and $push(z,3)$.
-Therefore, push 2, a predecessor of the push which is removed, is marked as $be(6)$.
-The other predecessor of the removed push, i.e., push $1$, is not marked as $be(6)$ because it happens before another push, i.e., push 3, which is already marked as $be(6)$ (the value added by push 3 should be removed before the value added by push 1 could become the top of the stack).
+The execution on the bottom shows an example where the marking $be(k)$ for some pop $k$ is updated at commit points.
+The pushes $3$ and $4$ are marked as $be(5)$ and $be(6)$
+when the pops $5$ and $6$ start.
+Then, $com(pop,t,5)$ is enabled since $t$ was added by $push(t,4)$ which is marked as $be(5)$.
+Besides removing $push(t,4)$, the commit point produces a state where a pop committing later, e.g., pop $6$, can remove $y$ which was added by a predecessor of $push(t,4)$ in the happens-before ($y$ could become the top of the stack when $t$ is removed).
+This history is valid because $push(y,2)$ can be linearized after $push(x,1)$ and $push(z,3)$.
+Thus, push 2, a predecessor of the push which is removed, is marked as $be(6)$.
+Push $1$ which is also a predecessor of the removed push is not marked as $be(6)$ because it happens before another push, i.e., push 3, which is already marked as $be(6)$ (the value added by push 3 should be removed before the value added by push 1 could become the top of the stack).
 

### 2017-01-24 draftCAV/related.tex
fixed an issue in AbsS definition

%!TEX root = draft.tex
\vspace{-3.5mm}
\section{Related Work}
\vspace{-1.5mm}
Many techniques for linearizability verification, e.g.,~\cite{conf/ppopp/VafeiadisHHS06,conf/cav/AmitRRSY07,conf/vmcai/Vafeiadis09,conf/tacas/AbdullaHHJR13}, are based on forward simulation arguments, and typically only work for libraries where the linearization point of every invocation of a method $m$ is fixed to a particular statement in the code of $m$. The works in~\cite{conf/cav/Vafeiadis10,Derrick2011,conf/cav/DragoiGH13,DBLP:conf/cav/ZhuPJ15} deal with \emph{external} linearization points where the action of an operation $k$ can be the linearization point of a concurrently executing operation $k'$. We say that the linearization point of $k'$ is external. This situation arises in read-only methods like the {\tt contains} method of an optimistic set~\cite{conf/podc/OHearnRVYY10}, libraries based on the elimination back-off scheme, e.g.,~\cite{conf/spaa/HendlerSY04}, or flat combining~\cite{DBLP:conf/spaa/HendlerIST10,DBLP:conf/podc/GorelikH13}. 
In these implementations, an operation can do an update on the shared state that becomes the linearization point of a concurrent read-only method (e.g., a {\tt contains} returning {\tt true} may be linearized when an {\tt add} method adds a new value to the shared state) or an operation may update the data structure on behalf of other concurrently executing operations (whose updates are published in the shared state). In all these cases, every linearization point can still be associated syntactically to a statement in the code of a method and doesn't depend on operations executed in the future (unlike $\mathit{HWQ}$ and $\mathit{TSS}$). However, identifying the set of operations for which such a statement is a linearization point can only be done by looking at the whole program state (the local states of all the active operations). This poses a problem in the context of compositional reasoning (where auxiliary variables are required), but still admits a forward simulation argument. For manual proofs, such implementations with external linearization points can still be defined as LTSs that produce $Lin$-complete traces and thus still fall in the class of implementations for which forward simulations are enough for proving refinement. All these works are not able to deal with implementations such as $\mathit{HWQ}$ or $\mathit{TSS}$.

There also exist linearizability proof techniques based on backward simulations or alternatively, prophecy variables, e.g.,~\cite{phd/Vafeiadis08,DBLP:conf/cav/SchellhornWD12,DBLP:conf/pldi/LiangF13}. These works can deal with implementations where the linearization points are not fixed, but the proofs are conceptually more complex and less amenable to automation.

The works in~\cite{conf/concur/HenzingerSV13,DBLP:conf/icalp/BouajjaniEEH15} propose reductions of linearizability to assertion checking where the idea is to define finite-state automata that recognize violations of concurrent queues and stacks. These automata are simple enough in the case of queues and there is a proof of $\mathit{HWQ}$ based on this reduction~\cite{conf/concur/HenzingerSV13}. However, in the case of stacks, the automata become much more complicated and we are not aware of a proof for an implementation such as $\mathit{TSS}$ which is based on this reduction.

### 2017-01-24 draftCAV/stack.tex
fixed an issue in AbsS definition

@@
 
-The complete version of $\mathit{TSS}$~\cite{DBLP:conf/popl/DoddsHK15} contains also a mechanism for elimination (a pop can remove a value without traversing all the lists if it has been added by a concurrent push) and emptiness checking.
+%The complete version of $\mathit{TSS}$~\cite{DBLP:conf/popl/DoddsHK15} contains also a mechanism for elimination (a pop can remove a value without traversing all the lists if it has been added by a concurrent push) and emptiness checking.
 In both cases, the commit points can be identified in the code and correspond to certain boolean conditions evaluated to {\tt true}.
@@
 The rule {\sc com-pop1} for $com(pop,d,k)$ is enabled only if there exists a push $k'$ which added value $d$ and which belongs to $be(k)$ or $ov(k)$.
-When enabled, the push $k'$ is removed from the set $O$ (and the order $<$) and for every other pop $k_1$ such that $k'$ belongs to $be(k_1)$, $k'$ is replaced in $be(k_1)$ by its predecessors for which no successor belongs already to $be(k_1)$.
+When enabled, the push $k'$ is removed from the set $O$ (and the order $<$) and for every other pop $k_1$ such that $k'$ belongs to $be(k_1)$, $k'$ is replaced in $be(k_1)$ by its predecessors which are followed exclusively by pushes overlapping with $k_1$ (these predecessors become maximal closed pushes once $k'$ is removed).
 Also, $rv(k)$ is set to $d$.
@@
 
+    \vspace{-1mm}
     \inferrule[call-pop]{
@@
 
+    \vspace{-1mm}
     \inferrule[ret-pop]{
@@
 
+    \vspace{-1mm}
     \inferrule[ret-push1]{
@@
        \forall k_1.\ k'\not\in be(k_1) \Rightarrow be'(k_1)=be(k_1) \\     \\
-       \forall k_1.\ k'\in be(k_1) \Rightarrow be'(k_1)=(be(k_1)\setminus\{k'\})\cup \{k_2: k_2\in pred_{<}(k') \land \forall k_3. k_2<k_3 => k_3\not\in be(k_1)\} \\
+       \forall k_1.\ k'\in be(k_1) \Rightarrow be'(k_1)=(be(k_1)\setminus\{k'\})\cup \{k_2: k_2\in pred_{<}(k') \land \forall k_3. (k_2\in pred_{<}(k_3) \land k_3\neq k') => k_3\in ov(k_1)\} \\
     }{
@@
 
+    %\vspace{-1mm}
     \inferrule[com-pop2]{
@@
 
-\vspace{-1mm}
+\vspace{-2mm}
 \begin{theorem}\label{th:absImplStack}
 $AbsS$ is a refinement of $AbsS_0$ and vice-versa.
-\vspace{-1mm}
+\vspace{-2mm}
 \end{theorem}
@@
 
-TODO EXPLAIN THAT THIS IS DIFFERENT W.R.T. QUEUES ($Abs_0$ doesn't have commit points).
+%TODO EXPLAIN THAT THIS IS DIFFERENT W.R.T. QUEUES ($Abs_0$ doesn't have commit points).
 
@@
 \begin{corollary}
-A stack implementation $L$ with fixed pop commit points is a $C\cup R\cup Com(pop)$-refinement of $AbsS$ if{f} there exists a $C\cup R\cup Com(pop)$-forward simulation from $L$ to $AbsS$.
+A stack $L$ with fixed pop commit points is a $C\cup R\cup Com(pop)$-refinement of $AbsS$ if{f} there is a $C\cup R\cup Com(pop)$-forward simulation from $L$ to $AbsS$.
 \vspace{-1.5mm}
@@
 
-\vspace{-3.5mm}
+Linearization points can also be seen as commit points and thus the following holds.
+
+\vspace{-1.5mm}
+\begin{corollary}
+A stack implementation $L$ with fixed pop linearization points where transition labels $lin(pop,d,k)$ are substituted with $com(pop,d,k)$ is a $C\cup R\cup Com(pop)$-refinement of $AbsS_0$ if{f} there is a $C\cup R\cup Com(pop)$-forward simulation from $L$ to $AbsS$.
+\vspace{-1.5mm}
+\end{corollary}
+
+\vspace{-6mm}
 \subsection{A Correctness Proof For Time-Stamped Stack}
@@
 The components $be$ and $ov$ satisfy the following constraints (their domain is the set of identifiers of pending pops):
+\vspace{-2mm}
 \begin{itemize}
@@
 	\item a pop $k$ that reached a node with timestamp $\tau$ (its variable {\tt n} points to this node) overlaps with any push that has a timestamp bigger than $\tau$ and which created a node that occurs in {\tt pools} before the node reached by $k$, i.e., ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k_1)$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
-	\item if the variable {\tt youngest} of a pop $k$ points to a node which is not taken, then this node was created by a push in $be(k)\cup ov(k)$ or the node currently reached by $k$ is followed in {\tt pools} by another node which was created by a push in $be(k)\cup ov(k)$, i.e., ${\tt youngest}_s(k)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(k)={\tt n}_s(k_2)$ implies $k_1\in be(k)\cup ov(k)$ or that there exists $k_3\in O$ such that ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in be(k)\cup ov(k)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and $k$ is traversing the last list in the array {\tt pools}, for each $k, k_1,k_2$.
+	\item if the variable {\tt youngest} of a pop $k$ points to a node which is not taken, then this node was created by a push in $be(k)\cup ov(k)$ or the node currently reached by $k$ is followed in {\tt pools} by another node which was created by a push in $be(k)\cup ov(k)$, i.e., ${\tt youngest}_s(k)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(k)={\tt n}_s(k_2)$ implies $k_1\in be(k)\cup ov(k)$ or that there exists $k_3\in O$ such that ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in be(k)\cup ov(k)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and TODO $k$ is traversing the last list in the array {\tt pools}, for each $k, k_1,k_2$.
+\vspace{-2mm}
 \end{itemize}

### 2017-01-24 draft.tex
At last, TSS proof finished

@@
   \begin{mathpar}
-    \inferrule[call-push]{
-      k\not\in dom(cp) \\ 
-      d\neq {\tt null}
+    \inferrule[call-pop]{
+      k\not\in dom(cp) 
     }{
-      ..., O_a, x, cp,...
-      \xrightarrow{inv(push,d,k)} 
-     ..., O_a\cup\{k\}, x[k \mapsto d], cp[k \mapsto A_1], ...
+      ..., O_r, cp,...
+      \xrightarrow{inv(pop,k)} 
+     ..., O_r\cup\{k\}, cp[k \mapsto R_1], ...
+    }\hspace{5mm}
+
+    \inferrule[pop1]{
+      cp(k) = R_1 \\
+      maxThreads > 0
+    }{
+      ..., suc, ygst, mTS, i, cp
+      \xrightarrow{pop1(k)} 
+     ..., suc[k \mapsto \texttt{false}], ygst[k \mapsto \texttt{null}], mTS[k \mapsto -1], i[k \mapsto 0],  cp[k \mapsto R_2]
+    }\hspace{5mm}
+
+    \inferrule[pop2]{
+      cp(k) = R_2 \\
+      0 \leq i(k) < \texttt{maxThreads}
+    }{
+      ..., n, cp, ...
+      \xrightarrow{pop2(k)} 
+     ..., n[k \mapsto pools(i(k))], cp[k \mapsto R_3],...
+    }\hspace{5mm}
+
+    \inferrule[pop3]{
+      cp(k) = R_3 \\
+      n(k) \neq \texttt{null}\\
+      n(k)->taken = \texttt{true}\\
+      n(k)->next \neq n(k)
+    }{
+      ..., n, ...
+      \xrightarrow{pop3(k)} 
+     ..., n[k \mapsto n(k)->next],...
+    }\hspace{5mm}
+
+    \inferrule[pop4]{
+      cp(k) = R_3 \\
+      n(k) \neq \texttt{null}\\
+      n(k)->taken = \texttt{false}\\
+      n(k)->ts > maxTS(k)
+    }{
+      ..., maxTS, cp ...
+      \xrightarrow{pop4(k)} 
+     ..., maxTS[k \mapsto n(k)->ts], cp[k \mapsto R_4]...
+    }\hspace{5mm}
+
+    \inferrule[pop5]{
+      cp(k) = R_4 
+    }{
+      ..., youngest, cp ...
+      \xrightarrow{pop5(k)} 
+     ..., youngest[k \mapsto n(k)], cp[k \mapsto R_5]...
+    }\hspace{5mm}
+
+    \inferrule[pop6]{
+      cp(k) = R_3 \\
+      n(k) \neq \texttt{null}\\
+      n(k)->taken = \texttt{false}\\
+      n(k)->ts \leq maxTS(k)
+    }{
+      ..., cp, ...
+      \xrightarrow{pop6(k)} 
+     ...,cp[k \mapsto R_5],...
+    }\hspace{5mm}
+
+    \inferrule[pop7]{
+      cp(k) = R_5 \\
+      i(k) < \texttt{maxThreads}-1
+    }{
+      ...,i, cp, ...
+      \xrightarrow{pop7(k)} 
+     ...,i[k \mapsto i(k)+1], cp[k \mapsto R_2],...
+    }\hspace{5mm}
+
+    \inferrule[pop8]{
+      cp(k) = R_5 \\
+      youngest(k) = \texttt{null} \vee (youngest(k) \neq \texttt{null} \wedge youngest->taken)
+    }{
+      ...,success, cp, ...
+      \xrightarrow{pop7(k)} 
+     ...,success[k \mapsto \texttt{false}], cp[k \mapsto R_6],...
+    }\hspace{5mm}
+
+    \inferrule[com-pop]{
+      cp(k) = R_5 \\
+      youngest(k) \neq \texttt{null} \\
+      youngest(k) = m \\
+      d = m->data\\
+      m->taken = false \\
+      m' = m[taken \mapsto true]
+    }{
+      ...,success, youngest, cp, ...
+      \xrightarrow{com(pop,d,k)} 
+     ...,success[k\mapsto true], youngest[k \mapsto m'], cp[k \mapsto R_6],...
+    }\hspace{5mm}
+
+    \inferrule[pop9]{
+      cp(k) = R_6 \\
+      success(k) = \texttt{false}
+    }{
+      ..., cp, ...
+      \xrightarrow{pop9(k)} 
+     ..., cp[k \mapsto R_1],...
+    }\hspace{5mm}
+    \inferrule[ret-pop]{
+      cp(k) = R_6 \\
+      suc(k) = \texttt{false} \\ 
+      d = yst(k)->data
+    }{
+      ..., cp, ...
+      \xrightarrow{ret(pop,d,k)} 
+     ..., cp[k \mapsto R_7],...
     }\hspace{5mm}
@@
 We only mention the state components that are modified.
-Unmentioned state components have the names in the algorithm in the prestate.
-$*n = (a,b,c,d)$ is shorthand for $n->data = a$, $n->ts = b$, ...
-$n' = n[ts \mapsto expr]$ is short for $n'->ts = expr$ and all the other fields of $n$ and $n'$ are the same.
+Unmentioned state components have the names in the algorithm in the pre-state.
+$n' = n[taken \mapsto expr]$ is short for $n'->taken = expr$ and all the other fields of $n$ and $n'$ are the same.
   }
@@
 Then, neither $k \in be_t(p)$ nor $k \in ov_t(b)$.
-\item[\emph{ReverseFrontiers}] If a push $k \in O_t$ is concurrent with the pop $p$ and there exists another push $k' \in O_t$ that is immediate predecessor of $k$, then $k'$ is either concurrent or maximally closed with respect to $p$.
+\item[\emph{ReverseFrontiers}] If all immediate followers $k' \in O_t$ of a push $k \in O_t$ are concurrent with pop $p$, then $k$ is either concurrent or maximally closed with respect to $p$.
+More formally, let $k \in O_t$ and for all $k' \in O_t$ such that $k \in pred_{<_t}(k')$, $k' \in ov_t(p)$, where $p$ is a pop operation.
+Then, $k \in ov_t(p) \cup be_t(p)$.
+%If a push $k \in O_t$ is concurrent with the pop $p$ and there exists another push $k' \in O_t$ that is immediate predecessor of $k$, then $k'$ is either concurrent or maximally closed with respect to $p$.
 More formally, let $k, k' \in O_t$ such that $k' \in pred_{<_t}(k)$ and $k \in ov_t(p)$ for some pop $p$.
 Then, either $k' \in ov_t(p)$ or $k' \in be_t(p)$.
-\item[\emph{TraverseBefore}] If a pop operation $p$ is currently visiting node $n$ and there is a node $m$ coming before $n$ in the traverse order with a greater timestamp, then the operation that inserts $m$ must be concurrent with $p$.
-More formally, let $k \in O_t$ such that $n_s(k) <^{tr}_s n_s(p)$ and $n_s(k)->ts \geq n_s(p)->ts$.
+\item[\emph{FixReturn}] If a pop $p$ is after its commit point action in $s$, then the $rv$ value of this operation in $t$ is fixed to $youngest_s(p)->data$.
+More formally, Let $p$ be the pop operation such that $cp_s(p) = R_6$ and $success_s(p) = \texttt{true}$.
+Then, $rv_t(p) = youngest_s(p)->data$.
+\item[\emph{TraverseBefore}] If a pop operation $p$ is currently visiting node $n$, it has non-null node $y$ as the $youngest$ and there is a non-null not taken node $m$ coming before $n$ in the traverse order with a greater timestamp than $y$, then the operation that inserts $m$ must be concurrent with $p$.
+More formally, assume $youngest_s(p) = y$ and $ y \neq \texttt{null}$.
+Let $k \in O_t$ such that $n_s(k) \neq \texttt{null}$, $n_s(k)->taken = \texttt{false}$, $n_s(k) <^{tr}_s n_s(p)$ and $n_s(k)->ts \geq y->ts$.
+Then, $k \in ov_t(p)$.
+\item[\emph{TraverseBeforeNull}] If a pop operation $p$ is currently visiting node $n$, and its $youngest$ field is \texttt{null}, then every other node $m$ coming before $n$ in the traverse order must be concurrent with $p$.
+More formally , let $youngest_s(p) = \texttt{null}$ and assume there exists  an operation $k \in O_t$ such that $n_s(k) \neq \texttt{null}$, $n_s(k)->taken = \texttt{false}$ and  $n_s(k) <^{tr}_s n_s(p)$.
 Then, $k \in ov_t(p)$.
@@
 Then, either $k \in ov_t(p) \vee k \in be_t(p)$ or there exists $k'' \in O_t$ s.t.
-$n_s(k'')->ts > n_s(k) ->ts$ and $k'' \in ov_t(p) \vee k'' \in be_t(p)$ and either $k' <^{tr}_s k''$ or $n_s(p) = n_s(k'') \wedge cp_s(p) = R_j \wedge j>3$.
+$n_s(k'')->ts > n_s(k) ->ts$ and $k'' \in ov_t(p) \vee k'' \in be_t(p)$ and either $k' <^{tr}_s k''$ or $n_s(p) = n_s(k'') \wedge cp_s(p) = R_j \wedge j<5$.
 \end{itemize}
@@
 \begin{itemize}
-\item[\textsc{call-push}] The same derivation rule of $TSS$ is applied to $t$ obtain $t'$.
+\item[\textsc{init}] $fs[{q_0}_{TSS}] =\{{q_0}_{AbsS}\}$
+\item[\textsc{call-push}] The same derivation rule of $TSS$ is applied to $t$  to obtain $t'$.
 The premise of the rule is satisfied by $t$ trivially in the sense explained before.
@@
 The only new instantiation we check is $k \notin O_t$ does not violate \emph{Nodes} condition while relating $s'$ to $t$.
-For the latter case, we have $k \in O_t$.
-\emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex after changing $s$ to $s'$.
+
+For the latter case, we have $k \in O_t$. \emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex after changing $s$ to $s'$.
 \item[\textsc{push4}] We have $t' = t$ and show $t \in fs[s']$.
-We consider two cases: $n_s(k)->taken$ is \texttt{true} or it is \texttt{false}.
-For the former case, \emph{Nodes} condition is still satisfied since $k$ remains to be not a vertex.
+
+We consider two cases: $n_s(k)->taken$ is \texttt{true} or it is \texttt{false}. For the former case, \emph{Nodes} condition is still satisfied since $k$ remains to be not a vertex.
+
 For the latter case \emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex.
@@
 For the latter case, we obtain $t'$ by applying \textsc{ret-push2} rule of $AbsS$. \emph{Nodes} condition is still satisfied since $k \notin O_{t'}$. 
+\item[\textsc{call-pop}] The same derivation rule of $TSS$ is applied to $t$  to obtain $t'$.
+\emph{Frontiers} condition holds for $p = k$ relating $s'$ to $t'$ since $k' \in ov_{t'}(k)$ for every pending vertex $k'$ and $k'' \in be_{t'}(p)$ for all completed vertex $k''$.
+$t'$ due to action $inv(pop,k)$ applied on $t$.
+\emph{MaximalOV} condition holds for $p = k$ since pending vertices are maximal in $t'$ and for any maximally closed vertex $k'$ in $t'$, if $k'$ is ordered before other vertex $k''$, then $k''$ is a pending operation by definition of being maximally closed and $k'' \in ov_t(k)$ due to the changes by \textsc{inv-pop} action on $t$.
+\emph{MinimalBE} condition holds while relating $s'$ to $t'$ for the pop $ p = k$ because only maximally closed vertices are in $be(k)$ and if a push $k'$ is ordered before a maximally closed push $k''$ in $t$, neither $k'' \in be_{t'}(k)$ (since $k''$ is not maximally closed) nor $k'' \in ov_{t'}$ (since $k''$ cannot be pending).
+\emph{ReverseFrontiers} condition holds while relating $s'$ to $t'$ for the pop $p=k$ because, if $k'' \in ov_{t'}(k)$ for all immediate successors of $k'$ in $t$, then $k'' $ are pending vertices (due to \emph{call-pop} action of $AbsS$), $k'$ is a maximally closed vertex and $k' \in be_{t'}(k)$ (due to \emph{call-pop} action of $AbsS$).
+\item[\textsc{pop1}]We have $t' = t$ and $t \in fs[s']$.
+\item[\textsc{pop2}]We have $t' = t$ and show $t \in fs[s']$.
+\emph{TraverseBefore} condition while relating $s'$ to $t$ still holds for $p=k$.
+Assume $youngest_{s'}(k) = y$ is a non-null node.
+Then, for all nodes $m$ in $s'$ such that $n_s(k) \leq^{tr}_{s'} m <^{tr}_{s'} n_{s'}(k)$ we have $m->ts < y->ts$ in $s'$ because  $n_s(k)->ts > m->ts$ (since $n_s(k)$ is added to the pool after $m$ by the same thread) and $y->ts \geq n_s(k)->ts$ in $s'$ (since either $youngest_{s'}(k) = n_s(k)$ or $youngest_{s'}(k)->ts > n_s(k)->ts$).
+\emph{TraverseAfter} does not have any new instatiations since the guard mentions the nodes after $n_s(k)$ while relating $s$ to $t$ whereas it mentions nodes after or including $n_{s'}(k)$ which contains the all nodes in the former case.
+\item[\textsc{pop3}]We have $t' = t$ and $t \in fs[s']$.
+\item[\textsc{pop4}]We have $t' = t$ and $t \in fs[s']$.
+\item[\textsc{pop5}]We have $t' = t$ and show $t \in fs[s']$. 
+\emph{TraverseBefore} condition while relating $s'$ to $t$ still holds for $p=k$ since $youngest_s(k)->ts < youngest_{s'}(k)->ts$ and \emph{TraverseBefore} holds while relating $s$ to $t$.
+
+\emph{TraverseAfter} condition also continues to hold for $p=k$. There are two possible cases: $youngest_s(k) = \texttt{null}$ or not.
+
+First, consider the former case.
+Since \emph{TraverseBeforeNull} is satisfied while relating $s$ to $t$, for every operation $k', k'' \in O_t$ such that  $k'' <^{tr}_s k'$ and $n_s(k') = youngest_{s'}(k)$ we have $k'' \in ov_t(k)$.
+Consider all such $k''$ such that $n_s(k'')->ts > n_s(k')->ts$.
+If there exists such a $k''$ such that $k' \in pred_{<_t}(k'')$, then $k' \in ov_t(k) \cup be_t(k)$ since \emph{ReverseFrontiers} condition holds relating $s$ to $t$.
+Otherwise, either $k'$ is maximal in $t$ or all the vertices $v$ ordered after $k'$ in $t$ we have $v >^{tr}_s k'$.
+Then, either $k'$ or one of these $v$ vertices must be maximal in $t$ and must be in $be_t(k) \cup ov_t(k)$ since \emph{Frontiers} condition holds (one of them is maximal in $t$) while relating $s$ to $t$.
+
+Second, assume there exists push operations $j, k'$ such that $n_s(j) = youngest_s(k) \neq \texttt{null}$ and $n_s(k') = n_s(k) = youngest_{s'}(k)$ .
+Since \emph{TraverseBefore} is satisfied while relating $s$ to $t$, if there exists a push $k'' <^{tr}_s k'$ such that $n_s(k'')$ is not taken and $n_s(k'')->ts \geq n_s(j)->ts$, then $k'' \in ov_t(k)$.
+Then, for all $k'' <^{tr}_s k'$ such that $n_s(k'')$ is not taken and $n_s(k'')->ts \geq n_s(k')->ts$, then $k'' \in ov_t(k)$ since $n_s(k')->ts \geq n_s(j)->ts$.
+If there exists such a $k''$ such that $k' \in pred_{<_t}(k'')$, then $k' \in ov_t(k) \cup be_t(k)$ since \emph{ReverseFrontiers} condition holds relating $s$ to $t$.
+Otherwise, either $k'$ is maximal in $t$ or all the vertices $v$ ordered after $k'$ in $t$ we have $v >^{tr}_s k'$.
+Then, either $k'$ or one of these $v$ vertices must be maximal in $t$ and must be in $be_t(k) \cup ov_t(k)$ since \emph{Frontiers} condition holds (one of them is maximal in $t$) while relating $s$ to $t$.
+
+\item[\textsc{pop6}] We have $t' = t$ and show $t \in fs[s']$.
+\emph{TraverseAfter} continues to hold while relating $s'$ to $t$ for $p=k$.
+Let $k', k'' \in O_t$ such that $youngest_s(k) = n_s(k')$, $n_s(k) = n_s(k'')$ and $k' \notin  ov_t(k) \cup be_t(k)$.
+Note that $k' <^{tr}_s k''$.
+Then, $n_s(k'')->ts < n_s(k')->ts$ since $n_s(k'')->ts < maxTS(k)$ and $maxTS(k) = n_s(k')->TS$ ($n_s(k')->ts$ cannot be \texttt{MAX\_INT} since $k'$ would be pending and $k' \in ov_t(k)$ otherwise).
+Hence, there exists another push $j$ such that $j >^{tr}_s$ and $j \in ov_t(k) \cup be_t(k)$.
+
+\item[\textsc{pop7}] We have $t' = t$ and $t \in fs[s']$.
+\item[\textsc{pop8}] We have $t' = t$ and $t \in fs[s']$.
+\item[\textsc{com-pop}] $t'$ is obtained by applying \textsc{com-pop1} rule of $AbsS$.
+We first show that precondition of \textsc{com-pop1} rule of $AbsS$ si satisfied by $t$.
+If $com(pop,d,k)$ removes a  node $n$ such that there exists a push $k'$ such that $n_s(k') =n$ in $s$, then $k' \in O_t$ since it is non-null and not taken.
+Moreover, $k' \in ov_t(k) \cup be_t(k)$ since \emph{TraverseAfter} is preserved while relating $s$ to $t$ and all the nodes that come after $n_s(k)$ in terms of traverse order in $s$ have lower timestamp values than $n_s(k)->ts$ and $n_s(k)->ts \leq youngest_s(k)->ts$.
+
+Next, we show that $t' \in fs[s']$. We case split on the conditions of $fs$ considering new instantiations.
+
+\emph{Nodes} condition is still preserved after $k$ removes the node pushed by operation $k'$ in $s$ since $k' \notin O_{t'}$ anymore by due to $com(pop,d,k)$ action.
+
+\emph{Frontiers} condition is still preserved if $k$ removes the vertex $k'$ and makes another $k''$ maximally closed in $t$.
+Since all the other nodes $j$ ordered after $k''$ (except possibly $k'$) in $t$ are pending, $j \in ov_t(p)$ (due to \emph{Frontiers} condition while relating $s$ to $t$) for some pending pop $p \neq k$.
+Then, $k'' \in be_{t'}(p)$ by $com(pop,d,k)$ action.
+
+For the \emph{MinimalBE} condition, we do not have a new instance.
+If $k' \in be_{t'}(p)$ becomes true although $k' \notin be_t(p)$, we cannot have $k'' \in O_{t'}$ such that $k' \in pred_{<_{t'}}(k'')$ and $k'' \in be_{t'}(p)$ since $com(pop,d,k)$ does not add $k''$ to $ov(p)$ if its successor is not pending with respect to $p$.
+
+\emph{ReverseFrontiers} condition is still preserved.
+If $k$ removes the vertex $k'$ and there exists an immediate predecessor $k''$ of $k'$ such that all of immediate successors of $k''$ are in $ov_{t'}(p)$, then $k'' \in ov_{t'}(p)$ due to the action $com(pop,d,k)$.
+
+\emph{TraverseAfter} condition is still preserved after $k$ removes the node of push $k'$.
+Let $p \neq k$ be another pop operation such that $n_s(j) = youngest_s(p)$ for some push $j$ and $n_s(k')$ be the only node such that $n_s(k')->ts > youngest_s(p)->ts$ and $n_s(k')$ comes after $n_s(p)$ in the traverse order of $s$ and $k' \in ov_t(p) \cup be_t(p)$.
+Hence, there is no $k''$ such that $n_s(k'')$ comes after $n_s(p)$ in the traverse order and $j <_t k''$ except $k'$ (i).
+In other direction, if for all $k'' \in O_t$ such that $n_s(k'')$ comes before $n_s(p)$ in the traverse order and $n_s(k'')->ts > youngest_s(p)->ts$ , then $k'' \in ov_t(p)$ since \emph{TraverseBefore} condition holds while relating $s$ to $t$.
+Then, for all $k'' \in O_t$ such that $n_s(k'')$ comes before $n_s(p)$ in the traverse order of $s$ and $k'' >_t j$ implies $k'' \in ov_t(p)$ since $n_s(k'')->ts > n_s(j)->ts$ if $k'' >_t j$ (ii).
+Then, for all $k'' \in O_t$ such that if $k'' >_t j$, then $k'' \in ov_t(p)$ except $k'$ due to (i) and (ii).
+If $j \nless_t k'$, then $j \in ov_t(p) \cup be_t(p)$ since \emph{ReverseFrontiers} hold while relating $s$ to $t$ and $j \in ov_{t'} \cup be_{t'}$ after applying the action $com(pop,d,k)$.
+Otherwise, if $j <_t k'$, then $k \in be_{t'}$ after applying $com(pop,d,k)$.
+
+\emph{FixReturn} condition continues to hold.
+If $com(pop,d,k)$ removes the node pushed by $k'$ in $s$, then $com(pop,d,k)$ removes the vertex $k'$ (assuming data independece) and $youngest_s(k')->data = \ell_t(k')_1$.
+Then, $youngest_{s'}(p)->data = rv_{t'}(p)$ after applying commit actions at both sides.
+
+\item[\textsc{pop9}]We have $t' = t$ and $t \in fs[s']$.
+\item[\textsc{ret-pop}] $t'$ is obtained by applying \textsc{ret-pop} rule of $AbsS$ and $t' \in fs[s']$.
 \end{itemize}
 \end{proof}
-
+\textcolor{red}{TODO: Check ReverseFrontiers for the cases before compop.}
+\textcolor{red}{TODO: Check TraverseBefore for the cases before pop2.}
+\textcolor{red}{TODO: CheckTraverseBeforeNull for the cases befoere pop 5.}
+\textcolor{red}{TODO: Apply Dr. Enea's comments.}
+\textcolor{red}{TODO: Try case split on fs conditions instantiations check.}
+\textcolor{red}{TODO: Add Flow Diagrams to the paper.} 
+\textcolor{red}{TODO: }
 %\input{set}

### 2017-01-24 draftCAV/app-absImplQueue.tex
queues

@@
 
-In this section, we will show that for any concurrent queue implementation library $L_C$ for which we know the linearization points of the dequeue operation and that is linearizable with respect to the reference implementation library $L_A$, there exists a forward simulation $fs$ relating $L_C$ to $L_I$ where $L_I$ is an intermediate library equivalent to $L_A$ i.e.
+\begin{figure} [t]
+{\scriptsize
+  \centering
+  \begin{mathpar}
+    \inferrule[call-enq]{
+      k\not\in dom(cp^0) \\ 
+      d\neq {\tt EMPTY}
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{inv(enq,d,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
+      \sigma,in^0[k\mapsto d], rv^0,cp^0[k\mapsto A_1]
+    }\hspace{5mm}
+    \inferrule[lin-enq]{
+      cp^0(k)=A_1
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{lin(enq,d,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      d\cdot\sigma,in^0,rv^0,cp^0[k\mapsto A]
+    }\hspace{5mm}
+
+        \inferrule[ret-enq]{
+      cp^0(k)=A
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{ret(enq,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma,in^0,rv^0,cp^0[k\mapsto A_2]
+    }\hspace{5mm}
+
+    \inferrule[call-deq]{
+      k\not\in dom(cp^0) \\ 
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{inv(deq,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma,in^0,rv^0,cp^0[k\mapsto R_1]
+    }\hspace{5mm}
+        \inferrule[lin-deq1]{
+      cp^0(k)=R_1 \\
+      \sigma = \sigma'\cdot d 
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{lin(deq,d,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma',in^0,rv^0[k\mapsto d],cp^0[k\mapsto R_2]
+    }\hspace{5mm}
+
+        \inferrule[lin-deq2]{
+      cp^0(k)=R_1 \\
+      \sigma = \epsilon
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{lin(deq,{\tt EMPTY},k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma,in^0,rv^0[k\mapsto {\tt EMPTY}],cp^0[k\mapsto R_2]
+    }\hspace{5mm}
+    \inferrule[ret-deq]{
+      cp^0(k)=R_2 \\
+      rv^0(k) = d
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{ret(deq,d,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma,in^0,rv^0,cp^0[k\mapsto R_3]
+    }\hspace{5mm}
+          \end{mathpar}
+  }
+ \vspace{-4mm}
+  \caption{The transition relation of $AbsQ_0$. 
+  %\textcolor{red}{Call-Enq must have $d!= \texttt{EMPTY}$ as a premise. Also lin deq returning empty must be changed as before.}
+  }
+  \label{fig:transitions:AbsQ_0}
+\vspace{-2mm}
+\end{figure}
+
+We show that $AbsQ$ and $AbsQ_0$ refine each other. We start by giving a formal definition of the standard reference implementation $AbsQ_0$.
+Thus, the states of $AbsQ_0$ are tuples $\tup{\sigma,in^0,rv^0,cp^0}$ where $\sigma\in\<Vals>^*$ is a sequence of values, $in^0:\<Ops> ~> \<Vals>$ records the input value of an enqueue, $rv^0:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp^0:\<Ops> ~> \{A_1,A,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A,A_2$) or dequeue operation ($R_1,R_2,R_3$).
+All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ_0}.
+The alphabet of $AbsQ$ contains call/return actions and enqueue/dequeue linearization points.
+
+To prove that $AbsQ$ is a refinement of $AbsQ_0$ we define a normal $C\cup R\cup Lin(deq)$-backward simulation (i.e, a backward simulation as in Definition~\ref{def:back_app}) from $AbsQ$ to $AbsQ_0$.
+The reverse is shown using a normal $C\cup R\cup Lin(deq)$-forward simulation (i.e, a forward simulation as in Definition~\ref{def:for_app}).
+
+%In this section, we will show that for any concurrent queue implementation library $L_C$ for which we know the linearization points of the dequeue operation and that is linearizable with respect to the reference implementation library $L_A$, there exists a forward simulation $fs$ relating $L_C$ to $L_I$ where $L_I$ is an intermediate library equivalent to $L_A$ i.e.
 $L_I$ refines $L_A$ and vice versa.
-
-For all the queue libraries, we fix $\mathcal{M} = \{ enq, deq \}$ and $\mathcal{D} = \mathbb{N} \cup \{\texttt{EMPTY} \}$.
+%
+%For all the queue libraries, we fix $\mathcal{M} = \{ enq, deq \}$ and $\mathcal{D} = \mathbb{N} \cup \{\texttt{EMPTY} \}$.
 Since we know the linearization point of dequeues, we extend the definitions of the previous sections adding this information.
@@
 We also change definition of forward and backward simulation relations by replacing $A\Sigma$ with $AQ\Sigma$ in the original definitions and adding a new condition $(ii-d)$ to each of them:
-\begin{itemize}
-\item[Forward Simulation: (ii-d)] If $(s, lin(deq,d,k), s') \in \delta_{L_1}$ and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $(u, lin(deq,d,k), u') \in \delta_{L_2}$
-\item[Backward Simulation: (ii-d)] If $(s, lin(deq,d,k), s') \in \delta_{L_1}$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $(u, lin(deq,d,k), u') \in \delta_{L_2}$
-\end{itemize}
-Lemma 1 of the previous section still holds if we replace refinement with q-refinement and use new definitions of backward and forward simulations.
-
-Next, we define the abstract library $L_A$ as follows:
-\begin{itemize}
-\item A queue state consists of a finite sequence of natural numbers representing the queue content and a program counter for each operation.
+%\begin{itemize}
+%\item[Forward Simulation: (ii-d)] If $(s, lin(deq,d,k), s') \in \delta_{L_1}$ and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $(u, lin(deq,d,k), u') \in \delta_{L_2}$
+%\item[Backward Simulation: (ii-d)] If $(s, lin(deq,d,k), s') \in \delta_{L_1}$ and $u' \in bs[s']$, then there exists $u \in bs[s]$ such that $(u, lin(deq,d,k), u') \in \delta_{L_2}$
+%\end{itemize}
+%Lemma 1 of the previous section still holds if we replace refinement with q-refinement and use new definitions of backward and forward simulations.
+%
+%Next, we define the abstract library $L_A$ as follows:
+%\begin{itemize}
+%\item A queue state consists of a finite sequence of natural numbers representing the queue content and a program counter for each operation.
 More formally: $Q_A \subseteq \mathbb{N}^*  \times (\mathbb{N} \rightarrow Lbl_A)$ where $Lbl_A = \{N, E_0, E_1, E_2, D_0, D_1, D_2 \}$ is the set of transition labels of the operations.
@@
 For a state $q$, we denote the contents of the queue (first field) with $s_q$ and the function that maps operations to labels (the second field) with $f_q$.
-\item Transition labels consists of invocations, returns and linearizations of enqueue and dequeue operations: $\Sigma_A = AQ\Sigma \cup \{lin(enq,d,k)| d \in \mathcal{D}, k \in \mathbb{N}\}$.
+%\item Transition labels consists of invocations, returns and linearizations of enqueue and dequeue operations: $\Sigma_A = AQ\Sigma \cup \{lin(enq,d,k)| d \in \mathcal{D}, k \in \mathbb{N}\}$.
 Invocation of $deq$ operation does not have any input and return of $enq$ operation does not have any output value.
 We omit second fields from these labels.
-\item Initial state is the empty queue: ${q_0}_A = (\langle \rangle, f_{{q_0}_A})$ where $f_{{q_0}_A}(i) = N$ for all $i \in \mathbb{N}$.
-\item Each operation consists of invocation, linearization and return steps.
+%\item Initial state is the empty queue: ${q_0}_A = (\langle \rangle, f_{{q_0}_A})$ where $f_{{q_0}_A}(i) = N$ for all $i \in \mathbb{N}$.
+%\item Each operation consists of invocation, linearization and return steps.
 These are reflected in the transition relation.
 For the below definitions, unchanged parts of the state are omitted from the formulae:
-\begin{itemize}
-\item $(q, inv(enq,d,k), q') \in \delta_A$ iff $ d \neq \texttt{EMPTY} \wedge f_q(k) = N \wedge f_{q'}(k) = E_0$,
-\item $(q,lin(enq,d,k),q') \in \delta_A$ iff $f_q(k) =E_0 \wedge s_{q'} = s_q \circ \langle d \rangle \wedge f_{q'}(k) = E_1$ where $\circ$ is the operation that concatenates two finite sequences,
-\item $(q,ret(enq,k),q') \in \delta_A$ iff $f_q(k) = E_1 \wedge f_{q'}(k) = E_2$,
-\item $(q, inv(deq,k), q') \in \delta_A$ iff $f_q(k) = N \wedge f_{q'}(k) = D_0$
-\item $(q,lin(deq,d,k),q') \in \delta_A$ iff $f_q(k) =D_0 \wedge (d \neq \texttt{EMPTY} \wedge s_q = \langle d \rangle \circ s_{q'} \vee d=\texttt{EMPTY} \wedge s_q = s_{q'} = \langle \rangle )\wedge f_{q'}(k) = D_1$,
-\item $(q,ret(deq,d,k),q') \in \delta_A$ iff $f_q(k) = D_1 \wedge f_{q'}(k) = D_2$.
-\end{itemize}
-
-\end{itemize} 
-We restrict traces generated by this LTS by neglecting the invalid traces.
+%\begin{itemize}
+%\item $(q, inv(enq,d,k), q') \in \delta_A$ iff $ d \neq \texttt{EMPTY} \wedge f_q(k) = N \wedge f_{q'}(k) = E_0$,
+%\item $(q,lin(enq,d,k),q') \in \delta_A$ iff $f_q(k) =E_0 \wedge s_{q'} = s_q \circ \langle d \rangle \wedge f_{q'}(k) = E_1$ where $\circ$ is the operation that concatenates two finite sequences,
+%\item $(q,ret(enq,k),q') \in \delta_A$ iff $f_q(k) = E_1 \wedge f_{q'}(k) = E_2$,
+%\item $(q, inv(deq,k), q') \in \delta_A$ iff $f_q(k) = N \wedge f_{q'}(k) = D_0$
+%\item $(q,lin(deq,d,k),q') \in \delta_A$ iff $f_q(k) =D_0 \wedge (d \neq \texttt{EMPTY} \wedge s_q = \langle d \rangle \circ s_{q'} \vee d=\texttt{EMPTY} \wedge s_q = s_{q'} = \langle \rangle )\wedge f_{q'}(k) = D_1$,
+%\item $(q,ret(deq,d,k),q') \in \delta_A$ iff $f_q(k) = D_1 \wedge f_{q'}(k) = D_2$.
+%\end{itemize}
+%
+%\end{itemize} 
+%We restrict traces generated by this LTS by neglecting the invalid traces.
 If we project a trace to some operation identifier $k$ and obtain a sequence $\langle ..., inv(enq,d,k), lin(enq,d',k),...\rangle$ or $\langle ..., lin(deq,d,k), ret(deq,d',k),...\rangle$ where $d \neq d'$ we say that this trace is invalid.
-
-We want to introduce $L_I$ as the next step.
+%
+%We want to introduce $L_I$ as the next step.
 States of $L_I$ consists of strict orders of enqueue operations.
@@
 Then, the LTS of $L_I$ is defined as follows:
-\begin{itemize}
-\item A state consists of a partial order and a function keeping the program counter.
+%\begin{itemize}
+%\item A state consists of a partial order and a function keeping the program counter.
 More formally $Q_I \subseteq ND \times ED \times (\mathbb{N} \rightarrow Lbl_I)$ where $Lbl_I = \{N, E_0, E_1, D_0, D_1, D_2\} $ is the set of transition labels.
 $nd_q$ $ed_q$ and $f_q$ denotes the nodes of the strict order in state $q$, edges of the strict order in state $q$ and the function that maps operation to labels in state $q$, respectively.
-\item The transition label set consists of invocation and return action of both methods and linearization action of only the dequeue method: $\Sigma_I = AQ\Sigma$.
+%\item The transition label set consists of invocation and return action of both methods and linearization action of only the dequeue method: $\Sigma_I = AQ\Sigma$.
 Invocation of $deq$ operation does not have any input and return of $enq$ operation does not have any output value.
 We omit second fields from these labels.
-\item Initial state consists of an empty strict order and a function mapping every operation to $N$: ${q_0}_{I} = (so_{{q_0}_I}, f_{{q_0}_I}$ where $so_{{q_0}_I} =(\emptyset, \emptyset)$ and $f_{{q_0}_I}(i) = N$ for all $i \in \mathbb{N}$.
-\item While defining $\delta_I$, we again omit mentioning about the parts that has not changed:
+%\item Initial state consists of an empty strict order and a function mapping every operation to $N$: ${q_0}_{I} = (so_{{q_0}_I}, f_{{q_0}_I}$ where $so_{{q_0}_I} =(\emptyset, \emptyset)$ and $f_{{q_0}_I}(i) = N$ for all $i \in \mathbb{N}$.
+%\item While defining $\delta_I$, we again omit mentioning about the parts that has not changed:
+%\begin{itemize}
+%\item $(q, inv(enq,d,k), q') \in \delta_I$ iff $f_q(k) = N \wedge f_{q'}(k) = E_0 \wedge d \neq \texttt{EMPTY} \wedge (k,\_,\_) \notin nd_q \wedge nd_{q'} = nd_q \cup \{ (k,d,\texttt{PENDING}\} \wedge AddNode(ed_{q'}, ed_q,k)$ where $AddNode(ed_{q'}, ed_q,k)$ is true iff $ed_{q'}$ is obtained from $ed_q$ by adding edges from every closed node of $ed_q$ to the node with identifier $k$.
+%\item $(q, ret(enq,k), q') \in \delta_I$ iff $f_q(k) =E_0 \wedge f_{q'}(k) = E_1 \wedge ( (k,\_,\texttt{PENDING}) \notin nd_q \wedge so_q = so_{q'} \vee UpdateNode(so_{q'}, so_q, k) )$ where $UpdateNode(so_{q'}, so_q, k)$ is true iff $(k,d,\texttt{PENDING}) \in nd_q$ for some $d \in \mathcal{D}$, it is replaced with node $(k,d,\texttt{CLOSED}$ in the state $q'$ and all the edges adjacent to $(k,d,\texttt{PENDING})$ in state $q$ are replaced with edges adjacent to $(k,d,\texttt{CLOSED})$ in the state $q'$.
+%\item $(q, inv(deq,k), q') \in \delta_I$ iff $f_q(k) = N \wedge f_{q'}(k) = D_0$.
+%\item $(q, lin(deq,d,k), q') \in \delta_I$ iff $f_q(k) = D_0 \wedge f_{q'}(k) = D_1 \wedge (d = \texttt{EMPTY} \wedge nd_q = ed_q = nd_{q'} = ed_{q'} \emptyset \vee d \neq \texttt{EMPTY} \wedge RemoveNode(so_q, so_{q'},d))$ where $RemoveNode(so_q, so_{q'},d)$ is true iff there is a minimal node $n \in so_q$ of which data value (second field) is $d$ and $so_{q'}$ is obtained from $so_q$ by removing $n$ and all the edges adjacent to it.
+Note that linearization of dequeue could remove a \texttt{PENDING} node.
+%\item $(q, ret(deq,d,k) q') \in \delta_I$ iff $f_q(k) = D_1 \wedge f_{q'}(k) = D_2$.
+%\end{itemize}
+%\end{itemize}
+%Again, we omit the inconsistent traces from $L_I$ as in $L_A$. Note that the library $L_I$ is deterministic with respect to alphabe $AQ\Sigma$.
+%
+%Next, we show equivalence of $L_A$ and $L_I$ in terms of q-refinement.
+\begin{lemma} 
+$AbsQ$ is a refinement of $AbsQ_0$.
+\end{lemma}
+\begin{proof}
+We define a normal $C\cup R\cup Lin(deq)$-backward simulation $bs$ from $AbsQ$ to $AbsQ_0$ as follows.
+Given an $AbsQ$ state $s=\tup{O,<,\ell,rv,cp}$ and an $AbsQ_0$ state $t=\tup{\sigma,in^0,rv^0,cp^0}$ we have that $(s,t)\in bs$ iff the following hold:
 \begin{itemize}
-\item $(q, inv(enq,d,k), q') \in \delta_I$ iff $f_q(k) = N \wedge f_{q'}(k) = E_0 \wedge d \neq \texttt{EMPTY} \wedge (k,\_,\_) \notin nd_q \wedge nd_{q'} = nd_q \cup \{ (k,d,\texttt{PENDING}\} \wedge AddNode(ed_{q'}, ed_q,k)$ where $AddNode(ed_{q'}, ed_q,k)$ is true iff $ed_{q'}$ is obtained from $ed_q$ by adding edges from every closed node of $ed_q$ to the node with identifier $k$.
-\item $(q, ret(enq,k), q') \in \delta_I$ iff $f_q(k) =E_0 \wedge f_{q'}(k) = E_1 \wedge ( (k,\_,\texttt{PENDING}) \notin nd_q \wedge so_q = so_{q'} \vee UpdateNode(so_{q'}, so_q, k) )$ where $UpdateNode(so_{q'}, so_q, k)$ is true iff $(k,d,\texttt{PENDING}) \in nd_q$ for some $d \in \mathcal{D}$, it is replaced with node $(k,d,\texttt{CLOSED}$ in the state $q'$ and all the edges adjacent to $(k,d,\texttt{PENDING})$ in state $q$ are replaced with edges adjacent to $(k,d,\texttt{CLOSED})$ in the state $q'$.
-\item $(q, inv(deq,k), q') \in \delta_I$ iff $f_q(k) = N \wedge f_{q'}(k) = D_0$.
-\item $(q, lin(deq,d,k), q') \in \delta_I$ iff $f_q(k) = D_0 \wedge f_{q'}(k) = D_1 \wedge (d = \texttt{EMPTY} \wedge nd_q = ed_q = nd_{q'} = ed_{q'} \emptyset \vee d \neq \texttt{EMPTY} \wedge RemoveNode(so_q, so_{q'},d))$ where $RemoveNode(so_q, so_{q'},d)$ is true iff there is a minimal node $n \in so_q$ of which data value (second field) is $d$ and $so_{q'}$ is obtained from $so_q$ by removing $n$ and all the edges adjacent to it.
-Note that linearization of dequeue could remove a \texttt{PENDING} node.
-\item $(q, ret(deq,d,k) q') \in \delta_I$ iff $f_q(k) = D_1 \wedge f_{q'}(k) = D_2$.
+	\item the sequence $\sigma$ is a linearization of a partial order $(D,\prec)$ where $D$ contains values labeling elements of $O$ and all the values corresponding to completed enqueues, i.e., $\ell_1({\tt COMP}(O))\subseteq D\subseteq \ell_1(O)$ ordered according to the happens-before order between the enqueues that added them, i.e., $d_1\prec d_2$ if{f} there exists $k_1,k_2$ such that $\ell_1(k_1)=d_1$, $\ell_1(k_2)=d_2$, and $k_1 < k_2$.
+	\item the return values fixed at dequeue linearization points are the same, i.e., for every $k$, $rv(k)=rv^0(k)$,
+	\item every dequeue is at the same control point in both $s$ and $t$, i.e., for every $k$ and $i\in \{1,2,3\}$, $cp(k)=R_i$ iff $cp^0(k)=R_i$,
+	\item every pending enqueue has the same input value in both $s$ and $t$, i.e., for every $k$, $\ell_1(k)=in^0(k)$,
+	\item a pending enqueue from $O$ has been linearized whenever its value is contained in $\sigma$, i.e., for every $k$, $cp^0(k)=A$ iff $\ell_1(k)\in D$ and $\ell_2(k)={\tt PEND}$, 
+	\item a pending enqueue from $O$ hasn't been linearized whenever its value is not in $\sigma$, i.e., for every $k$, $cp^0(k)=A_1$ iff $\ell_1(k)\not\in D$ and $\ell_2(k)={\tt PEND}$, 
+	\item a pending enqueue which is not in $O$ has been linearized, i.e., for every $k$, $cp^0(k)=A$ iff $k\not\in O$ and $cp(k)=A_1$, 
+	\item an enqueue is completed in $s$ whenever it is completed in $t$, i.e., for every $k$, $cp(k)=A_2$ iff $cp^0(k)=A_2$,
 \end{itemize}
-\end{itemize}
-Again, we omit the inconsistent traces from $L_I$ as in $L_A$. Note that the library $L_I$ is deterministic with respect to alphabe $AQ\Sigma$.
 
-Next, we show equivalence of $L_A$ and $L_I$ in terms of q-refinement.
-\begin{lem} 
-$L_I$ is a q-refinement of $L_A$.
-\end{lem}
-\begin{proof}
+In the following, we show that indeed $bs$ is a normal $C\cup R\cup Lin(deq)$-backward simulation from $AbsQ$ to $AbsQ_0$.
+
+TODO FILL IN THE STEPS
+
 We will provide a backward simulation relation btw $L_I$ and $L_A$.
@@
 
-\begin{lem}
-$L_A$ is a q-refinement of $L_I$.
-\end{lem}
+%\begin{lem}
+%$L_A$ is a q-refinement of $L_I$.
+%\end{lem}
+%\begin{proof}
+\begin{lemma} 
+$AbsQ_0$ is a refinement of $AbsQ$.
+\end{lemma}
 \begin{proof}
+We define a normal $C\cup R\cup Lin(deq)$-forward simulation $fs$ from $AbsQ_0$ to $AbsQ$ as follows. 
+Given $AbsQ_0$ state $t=\tup{\sigma,in^0,rv^0,cp^0}$ and an $AbsQ$ state $s=\tup{O,<,\ell,rv,cp}$ we have that $(t,s)\in fs$ iff the following hold:
+\begin{itemize}
+	\item the sequence $\sigma$ is a linearization of a partial order $(D,\prec)$ where $D$ contains values labeling elements of $O$ and all the values corresponding to completed enqueues, i.e., $\ell_1({\tt COMP}(O))\subseteq D\subseteq \ell_1(O)$ ordered according to the happens-before order between the enqueues that added them, i.e., $d_1\prec d_2$ if{f} there exists $k_1,k_2$ such that $\ell_1(k_1)=d_1$, $\ell_1(k_2)=d_2$, and $k_1 < k_2$.
+	\item every dequeue is at the same control point in both $s$ and $t$, i.e., for every $k$ and $i\in \{1,2,3\}$, $cp(k)=R_i$ iff $cp^0(k)=R_i$,
+	\item every enqueue is pending in $s$ whenever it is pending in $t$, i.e., for every $k$, $cp(k)=A_1$ iff $cp^0(k)\in \{A_1,A\}$,
+	\item every enqueue is completed in $s$ whenever it is completed in $t$, i.e., for every $k$, $cp(k)=A_2$ iff $cp^0(k) = A_2$,
+	\item every pending enqueue which is not linearized or whose value is present in $\sigma$ is a member of $O$, i.e., for every $k$, 
+	\begin{align*}
+	&k\in O\land \ell(k)=(d,{\tt PEND})\mbox{ iff } \\
+	&\hspace{2cm}(cp^0(k)=A_1\land in^0(k)=d)\vee (\exists i.\ \sigma_i =d\land cp^0(k)=A\land in^0(k)=d)
+	\end{align*}
+	\item every completed enqueue whose value is present in $\sigma$ is a member of $O$, i.e., for every $k$, 
+	\begin{align*}
+	&k\in O\land \ell(k)=(d,{\tt COMP})\mbox{ iff } \exists i.\ \sigma_i =d\land cp^0(k)=A_2\land in^0(k)=d
+	\end{align*}
+	\item pending enqueues are maximal in $<$, i.e., for every $k$ and $k'$, $k \not< k'$ if $\ell_2(k)={\tt PEND}$,
+\end{itemize}
+In the following, we show that indeed $fs$ is a normal $C\cup R\cup Lin(deq)$-forward simulation from $AbsQ_0$ to $AbsQ$.
+
+TODO FILL IN THE STEPS
+
 Since $L_I$ is deterministic with respect to the alphabet $AQ\Sigma$, we should be able to find a forward simulation between $L_A$ and $L_I$ if our claim is correct.

### 2017-01-24 draftCAV/app-absImplStack.tex
dealing with simulations
queues

@@
 
-We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$.
+We show that $AbsS$ and $AbsS_0$ refine each other.
+The standard reference implementation $AbsS_0$ is defined exactly as the one for queues, $AbsQ_0$, except that pop linearization points extract values from the beginning of the sequence stored in the state.
+
+\begin{figure} [t]
+{\scriptsize
+  \centering
+  \begin{mathpar}
+    \inferrule[call-push]{
+      k\not\in dom(cp^0) \\ 
+      d\neq {\tt EMPTY}
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{inv(push,d,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
+      \sigma,in^0[k\mapsto d], rv^0,cp^0[k\mapsto A_1]
+    }\hspace{5mm}
+    \inferrule[lin-push]{
+      cp^0(k)=A_1
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{lin(push,d,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      d\cdot\sigma,in^0,rv^0,cp^0[k\mapsto A]
+    }\hspace{5mm}
+
+        \inferrule[ret-push]{
+      cp^0(k)=A
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{ret(push,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma,in^0,rv^0,cp^0[k\mapsto A_2]
+    }\hspace{5mm}
+
+    \inferrule[call-pop]{
+      k\not\in dom(cp^0) \\ 
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{inv(pop,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma,in^0,rv^0,cp^0[k\mapsto R_1]
+    }\hspace{5mm}
+        \inferrule[lin-pop1]{
+      cp^0(k)=R_1 \\
+      \sigma = d\cdot \sigma' 
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{lin(pop,d,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma',in^0,rv^0[k\mapsto d],cp^0[k\mapsto R_2]
+    }\hspace{5mm}
+
+        \inferrule[lin-pop2]{
+      cp^0(k)=R_1 \\
+      \sigma = \epsilon
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{lin(pop,{\tt EMPTY},k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma,in^0,rv^0[k\mapsto {\tt EMPTY}],cp^0[k\mapsto R_2]
+    }\hspace{5mm}
+    \inferrule[ret-pop]{
+      cp^0(k)=R_2 \\
+      rv^0(k) = d
+    }{
+      \sigma,in^0,rv^0,cp^0
+      \xrightarrow{ret(pop,d,k)}
+      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv^0,cp[k\mapsto 1]
+      \sigma,in^0,rv^0,cp^0[k\mapsto R_3]
+    }\hspace{5mm}
+          \end{mathpar}
+  }
+ \vspace{-4mm}
+  \caption{The transition relation of $AbsS_0$. 
+  %\textcolor{red}{Call-Enq must have $d!= \texttt{EMPTY}$ as a premise. Also lin deq returning empty must be changed as before.}
+  }
+  \label{fig:transitions:AbsS_0}
+\vspace{-2mm}
+\end{figure}
+
+Thus, the states of $AbsS_0$ are tuples $\tup{\sigma,in^0,rv^0,cp^0}$ where $\sigma\in\<Vals>^*$ is a sequence of values, $in^0:\<Ops> ~> \<Vals>$ records the input value of a push, $rv^0:\<Ops> ~> \<Vals>$ records the return value of a pop fixed at its linearization point ($~>$ denotes a partial function), and $cp^0:\<Ops> ~> \{A_1,A,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A,A_2$) or pop operation ($R_1,R_2,R_3$).
+All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsS_0}.
+The alphabet of $AbsS$ contains call/return actions and push/pop linearization points.
+
+To prove that $AbsS$ is a refinement of $AbsS_0$ we define a normal $C\cup R$-backward simulation (i.e, a backward simulation as in Definition~\ref{def:back_app}) from $AbsS$ to $AbsS_0$.
+The reverse is shown using a normal $C\cup R$-forward simulation (i.e, a forward simulation as in Definition~\ref{def:for_app}).
+
+%We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$.
 However, we observe that, these implementations contain some points in their pop methods that logically removes the element from the pool.
@@
 Fixed linearization point of a pop preserves the following properties:
-\begin{itemize}
-\item If a $ret$ comes before a linearization point in the concrete execution, this order is preserved in the linearization of this execution.
-\item If a $lin$ pop comes before another $lin$ pop in the concrete execution, this order is preserved in the linearization of this execution.
-\item If a $lin$ pop comes before an $inv$ in the concrete execution, this order is preserved in the linearization of this execution. 
-\end{itemize} 
-A commit point is weaker than a fixed linearization point in the sense that it does not need to satisfy the first and the second conditions.
-
-To our intuition, if a pop (remove) method has multiple finite linearization points, commit point is the latest element in this set.
+%\begin{itemize}
+%\item If a $ret$ comes before a linearization point in the concrete execution, this order is preserved in the linearization of this execution.
+%\item If a $lin$ pop comes before another $lin$ pop in the concrete execution, this order is preserved in the linearization of this execution.
+%\item If a $lin$ pop comes before an $inv$ in the concrete execution, this order is preserved in the linearization of this execution. 
+%\end{itemize} 
+%A commit point is weaker than a fixed linearization point in the sense that it does not need to satisfy the first and the second conditions. 
+%
+%To our intuition, if a pop (remove) method has multiple finite linearization points, commit point is the latest element in this set.
 We have never observed an implementation in which a pop is linearized after it logically removes the element from the data structure.
-
-From now on, we will restrict ourselves to stacks, since our example implementation that we will show linearizability of is the time-stamped (TS) stack.
+%
+%From now on, we will restrict ourselves to stacks, since our example implementation that we will show linearizability of is the time-stamped (TS) stack.
 However, the notions and the machines we will introduce can be extended to queues easily.
-
-We fix $\mathcal{M} = \{push, pop\}$ and $\mathcal{D} = \mathbb{N} \cup \{\texttt{EMPTY}\}$.
+%
+%We fix $\mathcal{M} = \{push, pop\}$ and $\mathcal{D} = \mathbb{N} \cup \{\texttt{EMPTY}\}$.
 We extend the alphabet $A\Sigma$ for stacks with commit points as $ACS\Sigma = A\Sigma \cup \{com(pop,d,k)|d \in \mathcal{D}, k \in \mathcal{M}\}$.
@@
 Lemma 1 of the first section still holds with new simulation relation definitions and cs-linearizability and cs-linearizability implies the original linearizability definition.
-
-Our road map for this section is as follows: We will first introduce an intermediate stack machine $L_I$ that will be deterministic with respect to the alphabet $ACS\Sigma$.
+%
+%Our road map for this section is as follows: We will first introduce an intermediate stack machine $L_I$ that will be deterministic with respect to the alphabet $ACS\Sigma$.
 We will show that $L_I$ is equivalent to the standard abstract stack $L_A$ defined in the previous section with respect to the language $A\Sigma$.
@@
 As an example, we will pick $L_C$ as time-stamped stack and establish a forward simulation relation between it and our $L_I$ machine.
-
-Let us continue with defining $L_I$ first:
-\begin{itemize}
-\item A state of $L_I$ again consists of a partial strict order and a program counter: $Q_I \subseteq ND \times ED \times (\mathbb{N} \rightarrow Lbl_I)$ where $Lbl_I = \{N, A_0, A_1, R_0, R_1, R_2\}$ is the set of transition labels for the operations.
+%
+%Let us continue with defining $L_I$ first:
+%\begin{itemize}
+%\item A state of $L_I$ again consists of a partial strict order and a program counter: $Q_I \subseteq ND \times ED \times (\mathbb{N} \rightarrow Lbl_I)$ where $Lbl_I = \{N, A_0, A_1, R_0, R_1, R_2\}$ is the set of transition labels for the operations.
 Different from the fixed linearization point case, this time nodes are not triples but 5-tuples of the form $(k,d,st,mc,con)$.
@@
 the either this node was open when the pop started or this node is created when the pop was pending.
-\item The transition labels consist of invocation and return actions of both methods and commit action for only pop method.
+%\item The transition labels consist of invocation and return actions of both methods and commit action for only pop method.
 Hence $\Sigma_I = ACS\Sigma$.
@@
 They are of the form: $com(pop,d,k)$ where $d \in \mathcal{D}$ and $k \in \mathbb{N}$.
-\item Initial state consists of an empty strict partial order and a function mapping ever operation to $N$: ${q_0}_I = (\emptyset, \emptyset, f_{{q_0}_I})$ where  $f_{{q_0}_I}(i) =N$ for all $i \in \mathbb{N}$.
-\item We define $\delta_I$ less formally, by giving verbal explanations to the transitions, omitting the obvious updates on the $f$ part and not mentioning about the parts of the nodes that does not change:
-\begin{itemize}
-\item $(q, inv(push,d,k),q') \in \delta_I$ iff a new node $n=(k,d,\texttt{PENDING}, \emptyset, con_n)$ is added to $nd_{q'}$, where $con_n = \{ i \in \mathbb{N}| f_q(i) = R_0\}$; $n' \rightarrow n$ will be added to $ed_{q'}$ if $n'$ is a closed node at the state $q$.
-\item $(q, ret(push,k),q') \in \delta_I$ iff either there is a \texttt{PENDING} node $n$ in state $q$ and this node becomes \texttt{CLOSED} in state $q'$ or there is no node with identifier $k$ in $q$ and nothing else than $f$ field changes in $q'$.
-\item $(q, inv(pop,k),q') \in \delta_I$ iff for every open node $n \in nd_q$, $k$ is added to $n.con$ in $q'$ and for every maximally closed node $m \in nd_q$, $k$ is added to $m.mc$ in state $q'$. 
-\item $(q, com(pop,k,d), q') \in \delta_I$ iff there exists a node $n$ in state $q$ such that $n.d = d$ and either $k \in n.con$ or $k \in n.mc$, this node $n$ and all the nodes adjacent to it are removed in the state $q'$, $k$ is removed from all $con$ and $mc$ fields of all nodes in state $q'$ and for all other pop operations $k'$ in $n.mc$ or in $n.con$ and for all states $n' \in nd_q$ such that $n' \rightarrow n \in ed_q$ and for all $n'' \in nd_q$, $n' \rightarrow n'' \in ed_q$ implies $k' \in n''.con$ (implies $k \notin n''.mc$ but it is stronger than this condition: if there are three closed states $p,q,r$ s.t.
+%\item Initial state consists of an empty strict partial order and a function mapping ever operation to $N$: ${q_0}_I = (\emptyset, \emptyset, f_{{q_0}_I})$ where  $f_{{q_0}_I}(i) =N$ for all $i \in \mathbb{N}$.
+%\item We define $\delta_I$ less formally, by giving verbal explanations to the transitions, omitting the obvious updates on the $f$ part and not mentioning about the parts of the nodes that does not change:
+%\begin{itemize}
+%\item $(q, inv(push,d,k),q') \in \delta_I$ iff a new node $n=(k,d,\texttt{PENDING}, \emptyset, con_n)$ is added to $nd_{q'}$, where $con_n = \{ i \in \mathbb{N}| f_q(i) = R_0\}$; $n' \rightarrow n$ will be added to $ed_{q'}$ if $n'$ is a closed node at the state $q$.
+%\item $(q, ret(push,k),q') \in \delta_I$ iff either there is a \texttt{PENDING} node $n$ in state $q$ and this node becomes \texttt{CLOSED} in state $q'$ or there is no node with identifier $k$ in $q$ and nothing else than $f$ field changes in $q'$.
+%\item $(q, inv(pop,k),q') \in \delta_I$ iff for every open node $n \in nd_q$, $k$ is added to $n.con$ in $q'$ and for every maximally closed node $m \in nd_q$, $k$ is added to $m.mc$ in state $q'$. 
+%\item $(q, com(pop,k,d), q') \in \delta_I$ iff there exists a node $n$ in state $q$ such that $n.d = d$ and either $k \in n.con$ or $k \in n.mc$, this node $n$ and all the nodes adjacent to it are removed in the state $q'$, $k$ is removed from all $con$ and $mc$ fields of all nodes in state $q'$ and for all other pop operations $k'$ in $n.mc$ or in $n.con$ and for all states $n' \in nd_q$ such that $n' \rightarrow n \in ed_q$ and for all $n'' \in nd_q$, $n' \rightarrow n'' \in ed_q$ implies $k' \in n''.con$ (implies $k \notin n''.mc$ but it is stronger than this condition: if there are three closed states $p,q,r$ s.t.
 $p \rightarrow q$, $q \rightarrow r$, $p \rightarrow r$, $k'$ is only in $r.mc$ and we delete $r$, former condition only allows $k' \in q.mc$ whereas the latter one allows $k' \in p.mc$ in addition) we have $k' \in n'.mc$ in the state $q'$.
 Note that we need to assume data independence to make this action deterministic.
-\item $(q, ret(pop,k,d), q') \in \delta_I$ iff $q=q'$ ignoring the $f$ fields.
+%\item $(q, ret(pop,k,d), q') \in \delta_I$ iff $q=q'$ ignoring the $f$ fields.
+%\end{itemize}
+%\end{itemize}
+%$L_A$ is the same machine defined in the previous section. The common alphabet between $L_A$ and $L_I$ is $A\Sigma$. We will show that they are equivalent in terms of this alphabet.
+
+\begin{lemma} 
+$AbsS$ is a refinement of $AbsS_0$.
+\end{lemma}
+\begin{proof}
+We define a normal $C\cup R$-backward simulation $bs$ from $AbsS$ to $AbsS_0$ as follows.
+Given an $AbsS$ state $s=\tup{O,<,\ell,rv,cp}$ and an $AbsS_0$ state $t=\tup{\sigma,in^0,rv^0,cp^0}$ we have that $(s,t)\in bs$ iff the following hold:
+\begin{itemize}
+	\item if a pop has committed or respectively, it has returned in $s$, then it had been linearized or respectively, it has returned in $t$, i.e., for every $k$, if $cp(k)\in\{R_2,R_3\}$ then $cp^0(k)=cp(k)$,
+	\item a push is completed in $s$ whenever the same is true in $t$, i.e., for every $k$, $cp(k)=A_2$ iff $cp^0(k)=A_2$,
+	\item a push is pending in $s$ iff either it is a non-linearized pending push in $t$ or its linearization point has been executed, i.e., for every $k$, $cp(k)=A_1$ iff $cp^0(k)=A_1$ and $in^0(k)$ doesn't occur in $\sigma$, or $cp(k)=A$,
+	\item if a pop didn't commit in $s$ then it is pending in $t$ and it may have been linearized, i.e., for every $k$, if $cp(k)=R_1$ then $cp^0(k)\in \{R_1,R_2\}$,
+	\item let ${\tt UNCOM\_POP}=\{k: cp(k)=R_1\}$ be the set of pending pops in $s$ that didn't commit. Then, there exists a function
+
+	\item the sequence $\sigma$ is a linearization of a partial order $(D,\prec)$ where $D$ contains values labeling elements of $O$ and all the values corresponding to completed enqueues, i.e., $\ell_1({\tt COMP}(O))\subseteq D\subseteq \ell_1(O)$ ordered according to the happens-before order between the enqueues that added them, i.e., $d_1\prec d_2$ if{f} there exists $k_1,k_2$ such that $\ell_1(k_1)=d_1$, $\ell_1(k_2)=d_2$, and $k_1 < k_2$.
+	\item the return values fixed at dequeue linearization points are the same, i.e., for every $k$, $rv(k)=rv^0(k)$,
+	\item every dequeue is at the same control point in both $s$ and $t$, i.e., for every $k$ and $i\in \{1,2,3\}$, $cp(k)=R_i$ iff $cp^0(k)=R_i$,
+	\item every pending enqueue has the same input value in both $s$ and $t$, i.e., for every $k$, $\ell_1(k)=in^0(k)$,
+	\item a pending enqueue from $O$ has been linearized whenever its value is contained in $\sigma$, i.e., for every $k$, $cp^0(k)=A$ iff $\ell_1(k)\in D$ and $\ell_2(k)={\tt PEND}$, 
+	\item a pending enqueue from $O$ hasn't been linearized whenever its value is not in $\sigma$, i.e., for every $k$, $cp^0(k)=A_1$ iff $\ell_1(k)\not\in D$ and $\ell_2(k)={\tt PEND}$, 
+	\item a pending enqueue which is not in $O$ has been linearized, i.e., for every $k$, $cp^0(k)=A$ iff $k\not\in O$ and $cp(k)=A_1$, 
+	\item an enqueue is completed in $s$ whenever it is completed in $t$, i.e., for every $k$, $cp(k)=A_2$ iff $cp^0(k)=A_2$,
 \end{itemize}
-\end{itemize}
-$L_A$ is the same machine defined in the previous section. The common alphabet between $L_A$ and $L_I$ is $A\Sigma$. We will show that they are equivalent in terms of this alphabet.
 
-\begin{lem}
-$L_I$ is a refinement of $L_A$.
-\end{lem}
-\begin{proof}
+In the following, we show that indeed $bs$ is a normal $C\cup R\cup Lin(deq)$-backward simulation from $AbsQ$ to $AbsQ_0$.
+
 We will provide a backward simulation relation $bs$ between states of $L_I$ and states of $L_A$.
@@
 \end{proof}
-
-Now, we show that TS-Stack is linearizable by showing the concrete TS-Stack implementation $L_C$ is a cs-refinement of $L_I$.
-As $L_C$, we pick the simplest version that omits the \texttt{EMPTY} returns of the pop  methods, does not allow unlinking and elimination.

### 2017-01-24 draftCAV/app-backSim.tex
queues

@@
 
-\begin{dfn}
+\begin{definition}\label{def:for_app}
 Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
@@
 \end{itemize}
-\end{dfn}
+\end{definition}
 %If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
-A $\Gamma$-forward simulation requires that every step of $L_1$ corresponds to a (possibly empty) sequence of steps of $L_2$.
-In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action.
+With normal $\Gamma$-forward simulations, a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action.
 The rest of the transitions in $L_1$ are matched to a possibly empty sequence of transitions of $L_2$ with arbitrary labels.
@@
 A dual notion of forward simulation is the backward simulation:
-\begin{dfn}
+\begin{definition}\label{def:back_app}
 Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$.
@@
 \end{itemize}
-\end{dfn}
+\end{definition}
 

### 2017-01-24 draftCAV/app-absImplStack.tex
stacks

@@
 	\item if a pop didn't commit in $s$ then it is pending in $t$ and it may have been linearized, i.e., for every $k$, if $cp(k)=R_1$ then $cp^0(k)\in \{R_1,R_2\}$,
-	\item let ${\tt UNCOM\_POP}=\{k: cp(k)=R_1\}$ be the set of pending pops in $s$ that didn't commit. Then, there exists a function
-
-	\item the sequence $\sigma$ is a linearization of a partial order $(D,\prec)$ where $D$ contains values labeling elements of $O$ and all the values corresponding to completed enqueues, i.e., $\ell_1({\tt COMP}(O))\subseteq D\subseteq \ell_1(O)$ ordered according to the happens-before order between the enqueues that added them, i.e., $d_1\prec d_2$ if{f} there exists $k_1,k_2$ such that $\ell_1(k_1)=d_1$, $\ell_1(k_2)=d_2$, and $k_1 < k_2$.
-	\item the return values fixed at dequeue linearization points are the same, i.e., for every $k$, $rv(k)=rv^0(k)$,
-	\item every dequeue is at the same control point in both $s$ and $t$, i.e., for every $k$ and $i\in \{1,2,3\}$, $cp(k)=R_i$ iff $cp^0(k)=R_i$,
-	\item every pending enqueue has the same input value in both $s$ and $t$, i.e., for every $k$, $\ell_1(k)=in^0(k)$,
-	\item a pending enqueue from $O$ has been linearized whenever its value is contained in $\sigma$, i.e., for every $k$, $cp^0(k)=A$ iff $\ell_1(k)\in D$ and $\ell_2(k)={\tt PEND}$, 
-	\item a pending enqueue from $O$ hasn't been linearized whenever its value is not in $\sigma$, i.e., for every $k$, $cp^0(k)=A_1$ iff $\ell_1(k)\not\in D$ and $\ell_2(k)={\tt PEND}$, 
-	\item a pending enqueue which is not in $O$ has been linearized, i.e., for every $k$, $cp^0(k)=A$ iff $k\not\in O$ and $cp(k)=A_1$, 
-	\item an enqueue is completed in $s$ whenever it is completed in $t$, i.e., for every $k$, $cp(k)=A_2$ iff $cp^0(k)=A_2$,
+	\item there exists a partial function $g: \{k: cp(k)=R_1\} ~> O$ which associates uncommitted pops to pushes in $O$ such that:
+	\begin{itemize}
+		\item for every $k$, $k\in dom(g)$ iff $g(k)\in be(k)\cup ov(k)$
+		\item  the sequence $\sigma$ is the mirror of a linearization of a partial order $(D,\prec)$ where $D$ contains values labeling elements of $O$ except for those in the range of $g$, and all the values corresponding to completed pushes which are not in the range of $g$, i.e., $\ell_1({\tt COMP}(O)\setminus range(g))\subseteq D\subseteq \ell_1(O\setminus range(g))$ ordered according to the happens-before order between the pushes that added them, i.e., $d_1\prec d_2$ if{f} there exists $k_1,k_2$ such that $\ell_1(k_1)=d_1$, $\ell_1(k_2)=d_2$, and $k_1 < k_2$
+		\item every pop in the domain of $g$ has been linearized, i.e., for every $k$, $k\in dom(g)$ implies $cp^0(k)=R_2$,
+		\item every pop which is not in the domain of $g$ hasn't been linearized, i.e., for every $k$, $k\not\in dom(g)$ implies $cp^0(k)=R_1$,
+		\item every push in the range of $g$ has been linearized, i.e., for every $k$, $k\in range(g)$ implies $cp^0(k)=A$,
+		\item a pending enqueue from $O$ has been linearized when its value is contained in $\sigma$, i.e., for every $k$, if $\ell_1(k)\in D$ and $\ell_2(k)={\tt PEND}$, then $cp^0(k)=A$.
+	\end{itemize}
+	\item the return values fixed at pop commit points are the same, i.e., for every $k$, if $rv(k)$ is defined, then $rv(k)=rv^0(k)$,
+	\item every pending push has the same input value in both $s$ and $t$, i.e., for every $k$, $\ell_1(k)=in^0(k)$,
 \end{itemize}
 
-In the following, we show that indeed $bs$ is a normal $C\cup R\cup Lin(deq)$-backward simulation from $AbsQ$ to $AbsQ_0$.
+In the following, we show that indeed $bs$ is a normal $C\cup R$-backward simulation from $AbsS$ to $AbsS_0$.
+
+TODO FILL IN THE STEPS
 
@@
 
-\begin{lem}
-$L_A$ is a refinement of $L_I$.
-\end{lem}
+\begin{lemma} 
+$AbsS_0$ is a refinement of $AbsS$.
+\end{lemma}
 \begin{proof}
-We will construct a forward simulation relation $fs$ between $L_A$ and $L_I$.
+We define a normal $C\cup R$-forward simulation $fs$ from $AbsS_0$ to $AbsS$ as follows.
+Given an $AbsS_0$ state $t=\tup{\sigma,in^0,rv^0,cp^0}$ and an $AbsS$ state $s=\tup{O,<,\ell,rv,cp}$ we have that $(t,s)\in fs$ iff the following hold:
+\begin{enumerate}
+	\item every pop is at the same control point in both $t$ and $s$, i.e., for every $k$ and $i\in \{1,2,3\}$, $cp^0(k)=R_i$ iff $cp(k)=R_i$,
+	\item a push has been invoked in $t$ whenever it has been invoked in $s$, i.e., for every $k$, $cp^0(k)=A_1$ iff $cp(k)=A_1$,
+	\item a push which is linearized in $t$ has been invoked in $s$, i.e., for every $k$, if $cp^0(k)=A$ then $cp(k)=A_0$,
+	\item a push is completed in $t$ iff the same holds in $s$, i.e., for every $k$, $cp^0(k)=A_2$ iff $cp(k)=A_2$,
+	\item the pair $(O,\ell)$ in $s$ satisfies the following:
+	\begin{itemize}
+		\item for every $k$, if $in^0(k)=d$, $cp^0(k)\in \{A_1,A\}$, and $d$ occurs in $\sigma$, then $k\in O$ and $\ell(k)=(d,{\tt PEND})$,
+		\item for every $k$, if $in^0(k)=d$, $cp^0(k) = A_2$, and $d$ occurs in $\sigma$, then $k\in O$ and $\ell(k)=(d,{\tt COMP})$,
+	\end{itemize}
+	\item every pending push in $O$ is overlapping with every non-linearized pop, i.e., for every $k$, if $cp^0(k)=R_1$ then $\{k': k'\in O\land \ell_2(k')={\tt PEND}\}\subseteq ov(k)$.
+	\item every completed push is either overlapping or was the greatest completed push before a non-linearized pop started, i.e., for every $k$, if $cp^0(k)=R_1$, then ${\tt COMP}(O)\subseteq ov(k)\cup be(k)$,
+	\item for every push that overlaps with a pop $k$ or was maximal in $<$ when $k$ started, its successors are overlapping with $k$, i.e., $k_1\in be(k)\cup ov(k)$ and $k_1 < k_2$ implies $k_2 \in ov(k)$ for each $k, k_1, k_2$
+	\item predecessors of pushes in $be(k)$ for a given pop $k$ are neither overlapping with $k$ nor in $be(k)$, i.e., $k_1 < k_2$ and $k_2\in be(k)$ implies $k_1\not\in ov(k)\cup be(k)$ for each $k,k_1,k_2$
+	\item pending pushes are maximal in $<$, for every $k$ and $k'$, $k \not< k'$ if $\ell_2(k)={\tt PEND}$,
+	\item\label{item:stack_fs_lin} the sequence $\sigma$ is the mirror of a linearization of a partial order $(D,\prec)$ where $D$ contains values labeling elements of $O$ and all the values corresponding to completed pushes, i.e., $\ell_1({\tt COMP}(O))\subseteq D\subseteq \ell_1(O)$ ordered according to the happens-before order between the pushes that added them, i.e., $d_1\prec d_2$ if{f} there exists $k_1,k_2$ such that $\ell_1(k_1)=d_1$, $\ell_1(k_2)=d_2$, and $k_1 < k_2$.
+	\item the return values fixed at pop linearization/commit points are the same, i.e., for every $k$, $rv(k)=rv^0(k)$,
+	\item every pending push has the same input value in both $s$ and $t$, i.e., for every $k$, $\ell_1(k)=in^0(k)$,
+\end{enumerate}
+
+In the following, we show that indeed $fs$ is a normal $C\cup R$-backward simulation from $AbsS_0$ to $AbsS$:
+\begin{itemize}
+	\item Let $t\xrightarrow{inv(push,d,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$.
+We have that $(t',s')\in fs$ where $s\xrightarrow{inv(push,d,k)} s'$ (recall that $AbsS$ is deterministic).
+Since the push $k$ is non-linearized in $t'$, the component $\sigma$ of both $t$ and $t'$ are the same and $\ell_2(k)={\tt PEND}$ in $s'$.
+Then, the component $\sigma$ in $Abs_0$ states related by $fs$ to $s'$ is allowed to exclude values added by pushes in $s'$ which are labeled as pending.
+The effect of $inv(push,d,k)$ in $AbsS$ implies that $k$ overlaps with all pending pops.
+	\item Let $t\xrightarrow{inv(pop,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$.
+We have that $(t',s')\in fs$ where $s\xrightarrow{inv(pop,k)} s'$.
+The only difference between $s$ and $s'$ is that the components $be(k)$ and $ov(k)$ in $s'$ contain the greatest completed pushes in $s$ and the pending pushes in $s$, respectively (these components were undefined in $s$).
+The relation $fs$ doesn't exclude this particular choice for $be(k)$ and $ov(k)$ when applied to $t'$ and $s'$.
+	\item Let $t\xrightarrow{lin(push,d,k)} t'$ be a transition in $Abs_0$ and $(t,s)\in fs$.
+We have that $(t',s)\in fs$, i.e., the $AbsS_0$ transition is simulated by an empty sequence of $AbsS$ transitions, because essentially the component $\sigma$ of $t'$ still corresponds to a linearization of the pushes in $s$ according to item \ref{item:stack_fs_lin} in the definition of $fs$.
+The sequence $\sigma$ in $t'$ contains the value added by the push $k$ at the end, but this is allowed by $fs$ since $k$ is labeled as pending in $s$.
+	\item Let $t\xrightarrow{lin(pop,d,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$.
+We have that $(t',s')\in fs$ where $s\xrightarrow{com(pop,d,k)} s'$.
+The transition labeled by $com(pop,d,k)$ is enabled in $AbsS_0$ because $d$ was the first value in the sequence $\sigma$ of $t$.
+Indeed, this implies that $d$ was added by a push $k'$ which is maximal in the happens-before stored in $s$.
+This clearly implies that $k'\in be(k)\cup ov(k)$.
+In addition, the sequence $\sigma$ in $t'$ does correspond to a linearization of the pushes in $s'$ (which don't contain $k$ anymore) because $\sigma$ in $t$ had this property with respect to $s$ and $\sigma$ in $t'$ is obtained by deleting the first value in the sequence $\sigma$ of $t$.
+	\item Let $t\xrightarrow{ret(push,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$.
+We have that $(t',s')\in fs$ where $s\xrightarrow{ret(push,k)} s'$.
+There are two cases depending on whether the value added by $k$ is still present in the sequence $\sigma$ of $t$.
+If it is not, then the push $k$ doesn't occur in the happens-before from $s$, and the only effect of these two transitions is changing the control point of $k$.
+Therefore, $(t',s')\in fs$ clearly holds.
+When this value is still present, the effect of $ret(push,k)$ in $AbsS$ is changing the flag of push $k$ from ${\tt PEND}$ to ${\tt COMP}$.
+Since the order between pushes doesn't change, we have that $(t',s')\in fs$.
+	\item Let $t\xrightarrow{ret(pop,d,k)} t'$ be a transition in $AbsS_0$ and $(t,s)\in fs$.
+We have that $(t',s')\in fs$ where $s\xrightarrow{ret(pop,d,k)} s'$.
+This case is obvious, the only change between $s$ and $s'$ being the control point of $k$.
+\end{itemize}
+
+%We will construct a forward simulation relation $fs$ between $L_A$ and $L_I$.
 Our relation $fs \subseteq Q_A \times Q_I$ relates state $q = (s_q, f_q) \in Q_A$ to a state $q' =(so_{q'}, f_{q'}) \in Q_I$ iff (i) for all operation identifiers $k \in \mathbb{N}$, if $f_q(k) \in \{N, A_0, R_0, R_1, R_2\}$ then $f_{q'}(k) = f_q(k)$; if $f_q(k) = A_1$, then $f_{q'}(k) = A_0$; if $f_q(k) = A_2$, then $f_{q'}(k)= A_1$; (ii) We form nodes of $so_{q'}$ ($ND_{q'}$ as follows:  If $k$ is a push operation adding data value $d$ and either $f_q(k)= R_0$ or $f_q(k)=R_1$ and data added by this push exists in $s_q$, then there is a \texttt{PENDING} node in $ND_{q'}$ with identifier $k$ and data value $d$.
@@
 Note that we do not need $g$ function for keeping track of linearized pops unlike the previous proof.
-
-Next, we will show that $fs$ is a forward simulation relation.
-\begin{itemize}
-\item[$\langle i \rangle$] $fs[{q_0}_A] = \{{q_0}_I\}
-$ 
-\item[$\langle ii-a-push \rangle$] Let $(q, inv(push,d,k), q') \in \delta_A$ and $t \in fs[q]$.
+%
+%Next, we will show that $fs$ is a forward simulation relation.
+%\begin{itemize}
+%\item[$\langle i \rangle$] $fs[{q_0}_A] = \{{q_0}_I\}
+%$ 
+%\item[$\langle ii-a-push \rangle$] Let $(q, inv(push,d,k), q') \in \delta_A$ and $t \in fs[q]$.
 Pick $t'$ such that $(t, inv(push,d,k), t') \in \delta_I$.
@@
 By checking the other conditions, one can observe that $t' \in fs[q']$.
-\item[$\langle ii-a-pop \rangle$] Let $(q, inv(pop,k), q') \in \delta_A$ and $t \in fs[q]$.
+%\item[$\langle ii-a-pop \rangle$] Let $(q, inv(pop,k), q') \in \delta_A$ and $t \in fs[q]$.
 Pick $t'$ such that $(t, inv(pop,k), t') \in \delta_I$.
@@
 Since this new addition obeys our forward simulation definition, $t' \in fs[q']$.
-\item[$\langle ii-c-push \rangle$] Let $(q, lin(push,d,k) q') \in \delta_A$ and $t \in fs[q]$.
+%\item[$\langle ii-c-push \rangle$] Let $(q, lin(push,d,k) q') \in \delta_A$ and $t \in fs[q]$.
 Pick $t'=t$.
@@
 So, $t' \in fs[q']$ holds.
-\item[$\langle ii-c-pop \rangle$] Let $(q, lin(pop,d,k), q') \in \delta_A$ and $t \in fs[q]$.
+%\item[$\langle ii-c-pop \rangle$] Let $(q, lin(pop,d,k), q') \in \delta_A$ and $t \in fs[q]$.
 Pick $t'$ such that $(t, com(pop,d,k) t') \in \delta_I$.
@@
 Hence, $t' \in fs[q']$ holds.
-\item[$\langle ii-b-push \rangle$] Let $(q, ret(push,k), q') \in \delta_A$ and $t \in fs[q]$.
+%\item[$\langle ii-b-push \rangle$] Let $(q, ret(push,k), q') \in \delta_A$ and $t \in fs[q]$.
 We will consider two cases: either the element inserted by the push with identifier $k$ is removed by a concurrent pop or the element is still in $s_q$.
@@
 Since the edges are the same, $s_{q'}$ is a valid linearization of $s_{t'}$ and $t' \in fs[q']$ holds.
-\item[$langle ii-b-pop \rangle$] Let $(q, ret(pop,d,k), q') \in \delta_A$ and $t \in fs[q]$.
+%\item[$langle ii-b-pop \rangle$] Let $(q, ret(pop,d,k), q') \in \delta_A$ and $t \in fs[q]$.
 We pick $t'$ such that $(t, ret(pop,d,k) t') \in \delta_I$.
@@
 By checking the other conditions, we see that $t' \in fs[q']$.
-\end{itemize}
+%\end{itemize}
 \end{proof}

### 2017-01-24 draftCAV/stack.tex
stacks

@@
 Pop operations are treated differently since the commit points are not necessarily linearization points.
-Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order, and continuously tracks the push operations which are overlapping with it.
+Intuitively, each pop operation starts by taking a snapshot of the greatest completed push operations in the happens-before order, and continuously tracks the push operations which are overlapping with it.
 The commit point $com(pop,d,k)$ with $d\neq {\tt EMPTY}$ is enabled only if $d$ was added by one of the push operations in the initial snapshot, or by a push happening earlier when all the values from the initial snapshot have been removed, or by one of the push operations that overlaps with pop $k$.
@@
 The first state we give is reached after the call of pop with identifier $3$.
-This shows the effect of a pop invocation: the maximal completed pushes according to the current happens-before (here, the push with identifier $1$) are marked as $be(3)$ (from ``before'' operation 3), and the pending pushes are marked as $ov(3)$ (from ``overlapping'' with operation 3).
+This shows the effect of a pop invocation: the greatest completed pushes according to the current happens-before (here, the push with identifier $1$) are marked as $be(3)$ (from ``before'' operation 3), and the pending pushes are marked as $ov(3)$ (from ``overlapping'' with operation 3).
 As a side remark, any other push operation that starts after pop $3$ would be also marked as $ov(3)$.
@@
 
-Formally, the states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $<$ is a strict partial order over the set $O$ of operation identifiers, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a pending/completed flag, $rv:\<Ops> ~> \<Vals>$ records the return value of a pending pop fixed at its commit point, $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A_2$) or pop operation ($R_1,R_2,R_3$), $be:\<Ops> ~> 2^O$ records push operations that were maximal before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
+Formally, the states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $<$ is a strict partial order over the set $O$ of operation identifiers, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a pending/completed flag, $rv:\<Ops> ~> \<Vals>$ records the return value of a pending pop fixed at its commit point, $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A_2$) or pop operation ($R_1,R_2,R_3$), $be:\<Ops> ~> 2^O$ records the greatest completed push operations before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
 %The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.
@@
 	\item every pop overlaps with all pending pushes, i.e., $k_1\in O$ is pending implies $k_1\in ov(k)$ for each $k, k_1$
-	\item completed pushes which are maximal in $<$ are either overlapping with a pop $k$ or they were maximal in $<$ when $k$ started, i.e., every such push belongs either to $be(k)$ or $ov(k)$ for each $k$
+	\item the greatest completed pushes in $<$ are either overlapping with a pop $k$ or they were the greatest completed pushes when $k$ started, i.e., every such push belongs either to $be(k)$ or $ov(k)$ for each $k$
 	\item for every push that overlaps with a pop $k$ or was maximal in $<$ when $k$ started, its successors are overlapping with $k$, i.e., $k_1\in be(k)\cup ov(k)$ and $k_1 < k_2$ implies $k_2 \in ov(k)$ for each $k, k_1, k_2$

### 2017-01-25 draftCAV/stack.tex
small things

@@
 \vspace{-1mm}
-We define an abstract stack $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the push operations whose value has not been yet removed.
-Pop operations are treated differently since the commit points are not necessarily linearization points.
-Intuitively, each pop operation starts by taking a snapshot of the greatest completed push operations in the happens-before order, and continuously tracks the push operations which are overlapping with it.
+We define an abstract stack $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the pushes whose value has not been yet removed.
+Pops are treated differently since the commit points are not necessarily linearization points, intuitively, a pop can be linearized before its commit.
+Each pop operation starts by taking a snapshot of the greatest completed push operations in the happens-before order, and continuously tracks the push operations which are overlapping with it.
 The commit point $com(pop,d,k)$ with $d\neq {\tt EMPTY}$ is enabled only if $d$ was added by one of the push operations in the initial snapshot, or by a push happening earlier when all the values from the initial snapshot have been removed, or by one of the push operations that overlaps with pop $k$.

### 2017-01-25 draft.tex
small things

--- 2017-01-22 draft.tex
@@
 \input{commit}
-First, we describe the TS-Stack algorithm:
-\begin{lstlisting}
-struct Node{
-  int data;
-  int ts;
-  Node* next;
-  boolean taken;
-};
-Node* pools[maxThreads];
-int TS = 0;
-
-void push(int x) {
-  Node* n = new Node(x,MAX_INT,
-                        null,false);
-  n->next = pools[myTID];
-  pools[myTID] = n;
-  int i = TS++;
-  n->ts = i;
-}
-int pop() {
- boolean success = false;
- int maxTS = -1;
- Node* youngest = null;
- while ( !success ) {
-   maxTS = -1; youngest = null;
-   for(int i=0; i<maxThreads; i++){
-     Node* n = pools[i];
-     while (n->taken && n->next != n)
-       n = n->next;
-     if(maxTS < n->ts) {
-       maxTS = n->ts; youngest = n;
-     }
-   }
-   if (youngest != null)
-     success=CAS(youngest->taken,
-                       false,true);
- }
- return youngest->data;
-}
-\end{lstlisting}
-Then, we begin to obtain the LTS ($L_C$) of TS Stack from the algorithm by defining control points and the actions among these control points as seen in Figure X \textcolor{red}{Flow diagram is referenced here}.
-To simplify the proof, we take the initializations of some local variables together as atomic.
-
-States of the TS-Stack contains the global variables and local variables as fields.
-Global variables are just elements of their domains and local variables are maps from operation identifiers to their domains.
-We say $i_q(k)$ for referencing the value of local variable $i$ of operation $k$ in state $q$.
-There is only one special local variable called $myTID$.
-Its value is unique to each pending operation in a state i.e., concurrent operations cannot have the same $myTID$ value.
-TS-Stack states also contains sets $O_a, O_r \in \mathbb{O}$ which are operation identifier sets of push and pops respectively, and the control point function $cp$ which is a map from operation identifiers to the control points set that are presented in the flow diagram Figure X  \textcolor{red}{Flow diagram is referenced here}.
-Transition relation of the TS-STack is presented in Figure~\ref{fig:transitions:TSS}.
-Next, we show that the linearizability of TS Stack.
-
-\begin{figure} [t]
-{\scriptsize
-  \centering
-  \begin{mathpar}
-    \inferrule[call-push]{
-      k\not\in dom(cp) \\ 
-      d\neq {\tt EMPTY} \\
-      \forall k'.\ ov'(k') = ov(k')\cup \{k\}
-    }{
-      O,<,\ell,rv,cp,be,ov 
-      \xrightarrow{inv(push,d,k)} 
-      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1],be,ov'
-    }\hspace{5mm}
-
-      \end{mathpar}
-  }
- \vspace{-5mm}
-  \caption{The transition relation of TS-Stack
-  }
-  \label{fig:transitions:TSS}
-\vspace{-6mm}
-\end{figure}
-
-\begin{lem}
-$L_C$ is a Com(pop)- refinement of $AbsS$.
-\end{lem}
-\begin{proof}
-Since $AbsS$ is deterministic with respect to $C \cup R \cup Com(pop)$, there exists a Com(pop) forward simulation from $L_C$ to $AbsS$ iff $L_C$ is a Com(pop)-refinement of $AbsS$.
-Hence, we define the relation $fs$ and show that it is a Com(pop)-forward simulation relation.
-
-Let us make some clarifications before defining the relation.
-In order not to confuse nodes in TS Stack and nodes in $AbsS$, we call nodes of $AbsS$ as vertices from now on.
-We also define ordering relation (called traverse order) among the operations in a state of $TS$.
-It basically reflects the traverse order of pop operations.
-For two push operations $m,n \in O_a$ is state $s$ we say that $m <^{tr}_s n$ iff either $myTid(m) < myTid(n)$ or $myTid(m) = myTid(n)$ and $n_s(n)$ is reachable from $n_s(m)$ using next pointers.
-$\geq^{tr}$ is obtained from $<^{tr}$ in the usual way.
-
-The relation $fs \subseteq Q_C \rightarrow Q_{AbsS}$ contains $(s,t)$ iff the following are satisfied:
-\begin{itemize}
-\item[\emph{Nodes}] $k \in O_t$ iff $k$ is a push operation in $s$ ($k \in O_a$) such that either it has not inserted its node to the pool yet ( $cp_s(k) = A_i$ and $i<3$) or its node is not taken by a pop ($cp_s(k) = A_i$, $i\geq 3$ and $n_s(k)->taken = false$).
-\item[\emph{Pend/Comp}] A vertex $k \in O_t$ is pending ($\ell_t(k) = (d, \texttt{PEND})$) iff $k$ satisfies the previous condition, $x_s(k) = d$ and it is not completed in $s$ ($cp_s(k) = A_i$ and $i<6$).
-Similarly, this vertex is completed ($\ell_t(k) = (d, \texttt{COMP})$) iff $k$ satisfies the previous condition, $x_s(k) = d$ and it is completed in $s$ ($cp_s(k) = A_6$).
-Pending vertices are maximal with respect to $<_t$ i.e., if $k \in O_t$ is a pending vertex, then for all $k' \in O_t$ $k \nless_t k'$.
-\item[\emph{TSOrder}] If a node has a smaller timestamp than the other node in $s$, the operations that inserted them cannot be ordered reversely in $t$.
-More formally, let $k, k' \in O_t$ s.t.
-$n_s(k)-> ts \leq n_s(k')->ts$.
-Then, $k' \nless_t k$.
-\item[\emph{TidOrder}] Order among the nodes inserted by the same threads in $s$ must be preserved among the operations that inserted them in $t$.
-Let $k, k' \in O_t$ s.t.
-$myTid_s(k) = myTid_s(k')$ and $n_s(k)->ts < n_s(k')->ts$.
-Then, $k <_t k'$.
-\item[Frontiers] Every maximally closed or pending vertex can be removed by a pending pop.
-More formally, let $k \in O_t$ such that $\ell_t(k) = (\_,\texttt{PEND})$.
-Then, for all pops $p$, $k \in ov_t(p)$.
-In the other case, let $k \in O_t$ such that $\ell_t(k) = (\_,\texttt{COMP})$ and for all other $k' \in O_t$ such that $k<_t k'$, we know $\ell_t(k') = (\_,\texttt{PEND})$.
-Then, for all pop operations $p$, $k \in be_t(p)$ or $k \in ov_t(p)$.
-\item[\emph{MaximalOV}] If a push $k \in O_t$ is a candidate to be removed by a pop $p$, then every other push $k'$ invoked after $k$ is a candidate to be removed by $p$ since $k$ is concurrent with $p$.
-More formally, let $k, k' \in O_t$ such that $k <_t k'$ and there exists a pop $p$ such that $k \in be_t(p)$ or $k \in ov_t(p)$.
-Then, $k' \in ov_t(p)$.
-\item[\emph{MinimalBE}] If a push $k \in O_t$ has finished before the pop $p$ is invoked and yet $k$ is a candidate to be removed by $p$, other pushes completed before $k$ can not be candidates to be removed by $p$ at that state.
-More formally, let $k, k' \in O_t$ such that $k <_t k'$ and there exists a pop $p$ such that $k' \in be_t(p)$.
-Then, neither $k \in be_t(p)$ nor $k \in ov_t(b)$.
-\item[\emph{ReverseFrontiers}] If a push $k \in O_t$ is concurrent with the pop $p$ and there exists another push $k' \in O_t$ that is immediate predecessor of $k$, then $k'$ is either concurrent or maximally closed with respect to $p$.
-More formally, let $k, k' \in O_t$ such that $k' \in pred_{<_t}(k)$ and $k \in ov_t(p)$ for some pop $p$.
-Then, either $k' \in ov_t(p)$ or $k' \in be_t(p)$.
-\item[\emph{TraverseBefore}] If a pop operation $p$ is currently visiting node $n$ and there is a node $m$ coming before $n$ in the traverse order with a greater timestamp, then the operation that inserts $m$ must be concurrent with $p$.
-More formally, let $k \in O_t$ such that $n_s(k) <^{tr}_s n_s(p)$ and $n_s(k)->ts > n_s(p)->ts$.
-Then, $k \in ov_t(p)$.
-\item[\emph{TraverseAfter}] If a pop operation $p$ is currently visiting node $n$ that is not null and its youngest element $m$ is not null and still not taken in state $s$, then either $m$ is a candidate to be removed by $p$ in $t$ or there exists a later node $m'$ than $n$ such that $m'$ is a candidate in $t$ it has a bigger timestamp than n.
-More formally, assume that there exists $k, k' \in O_t$ such that $youngest_s(p) = n_s(k)$ and $n_s(k') = n_s(p)$.
-Then, either $k \in ov_t(p) \vee k \in be_t(p)$ or there exists $k'' \in O_t$ s.t.
-$n_s(k'')->ts > n_s(k) ->ts$ and $k'' \in ov_t(p) \vee k'' \in be_t(p)$ and either $k' <^{tr}_s k''$ or $n_s(p) = n_s(k'') \wedge cp_s(p) = R_j \wedge j>3$.
-\end{itemize}
-\end{proof}
 

### 2017-01-25 draftCAV/app-tss.tex
adding appendix tss

%!TEX root = draft.tex
\section{Proving the correctness of $TSS$}\label{app:tss}

%First, we describe the TS-Stack algorithm:
%\begin{lstlisting}
%struct Node{
%  int data;
%  int ts;
%  Node* next;
%  boolean taken;
%};
%Node* pools[maxThreads];
%int TS = 0;   
%
%void push(int x) {
%  Node* n = new Node(x,MAX_INT,
%                        null,false);
%  n->next = pools[myTID];
%  pools[myTID] = n;
%  int i = TS++;
%  n->ts = i;
%}
%int pop() {
% boolean success = false;
% int maxTS = -1;
% Node* youngest = null;
% while ( !success ) {
%   maxTS = -1; youngest = null;
%   for(int i=0; i<maxThreads; i++){
%     Node* n = pools[i];
%     while (n->taken && n->next != n)
%       n = n->next;
%     if(maxTS < n->ts) {
%       maxTS = n->ts; youngest = n;
%     }
%   }
%   if (youngest != null)
%     success=CAS(youngest->taken,
%                       false,true);
% }
% return youngest->data;
%}
%\end{lstlisting}
The LTS corresponding to the description of $TSS$ given in Fig.~\ref{fig:TimeStamped} is defined as usual. The control points and transition labels we use in the following proof are pictured in Fig.~\ref{} \textcolor{red}{Flow diagram is referenced here}. To simplify the proof, we take the initializations of some local variables together as atomic.

States of the TS-Stack contains the global variables and local variables as fields. Global variables are just elements of their domains and local variables are maps from operation identifiers to their domains. We say $i_q(k)$ for referencing the value of local variable $i$ of operation $k$ in state $q$. There is only one special local variable called $myTID$. Its value is unique to each pending operation in a state i.e., concurrent operations cannot have the same $myTID$ value. TS-Stack states also contains sets $O_a, O_r \in \mathbb{O}$ which are operation identifier sets of push and pops respectively, and the control point function $cp$ which is a map from operation identifiers to the control points set that are presented in the flow diagram Figure X  \textcolor{red}{Flow diagram is referenced here}. Transition relation of the TS-STack is presented in Figure~\ref{fig:transitions:TSSPush} (push rules) and Figure~\ref{fig:transitions:TSSPop} (pop rules).
Next, we show that the linearizability of TS Stack.

\begin{figure} [t]
{\scriptsize
  \centering
  \begin{mathpar}
    \inferrule[call-push]{
      k\not\in dom(cp) \\ 
      d\neq {\tt null}
    }{
      ..., O_a, x, cp,...
      \xrightarrow{inv(push,d,k)} 
     ..., O_a\cup\{k\}, x[k \mapsto d], cp[k \mapsto A_1], ...
    }\hspace{5mm}

    \inferrule[push1]{
      cp(k) = A_1 \\ 
      *n' = (x(k),\texttt{MAX\_INT}, \texttt{null}, \texttt{false})
    }{
      ..., n, cp,...
      \xrightarrow{push1(k)} 
     ..., n[k \mapsto n'], cp[k \mapsto A_2], ...
    }\hspace{5mm}

  \inferrule[push2]{
      cp(k) = A_2 
    }{
      ..., pools, cp,...
      \xrightarrow{push2(k)} 
     ..., pools[myTid(k) \mapsto n(k)], cp[k \mapsto A_3], ...
    }\hspace{5mm}

      \inferrule[push3]{
      cp(k) = A_3 
    }{
      ..., i, TS, cp,...
      \xrightarrow{push3(k)} 
     ..., i[k \mapsto TS], TS+1, cp[k \mapsto A_4], ...
    }\hspace{5mm}

   \inferrule[push4]{
      cp(k) = A_4 \\
      n'(k) = n(k)[ts \mapsto i(k)] \\
      \forall k'. cp(k') = A_6 \implies i(k') < i(k)
    }{
      ..., n, cp,...
      \xrightarrow{push4(k)} 
     ..., n[k \mapsto n'(k)], cp[k \mapsto A_5], ...
    }\hspace{5mm}

   \inferrule[ret-push]{
      cp(k) = A_5 
    }{
      ...,cp,...
      \xrightarrow{ret(push,k)} 
     ...,cp[k \mapsto A_6], ...
    }\hspace{5mm}

      \end{mathpar}
  }
 \vspace{-5mm}
  \caption{The push derivation rules of $TSS$. We only mention the state components that are modified. Unmentioned state components have the names in the algorithm in the prestate. $*n = (a,b,c,d)$ is shorthand for $n->data = a$, $n->ts = b$, ... $n' = n[ts \mapsto expr]$ is short for $n'->ts = expr$ and all the other fields of $n$ and $n'$ are the same.
  }
  \label{fig:transitions:TSSPush}
\vspace{-2mm}
\end{figure}

\begin{figure} [t]
{\scriptsize
  \centering
  \begin{mathpar}
    \inferrule[call-pop]{
      k\not\in dom(cp) 
    }{
      ..., O_r, cp,...
      \xrightarrow{inv(pop,k)} 
     ..., O_r\cup\{k\}, cp[k \mapsto R_1], ...
    }\hspace{5mm}

    \inferrule[pop1]{
      cp(k) = R_1 \\
      maxThreads > 0
    }{
      ..., suc, ygst, mTS, i, cp
      \xrightarrow{pop1(k)} 
     ..., suc[k \mapsto \texttt{false}], ygst[k \mapsto \texttt{null}], mTS[k \mapsto -1], i[k \mapsto 0],  cp[k \mapsto R_2]
    }\hspace{5mm}

    \inferrule[pop2]{
      cp(k) = R_2 \\
      0 \leq i(k) < \texttt{maxThreads}
    }{
      ..., n, cp, ...
      \xrightarrow{pop2(k)} 
     ..., n[k \mapsto pools(i(k))], cp[k \mapsto R_3],...
    }\hspace{5mm}

    \inferrule[pop3]{
      cp(k) = R_3 \\
      n(k) \neq \texttt{null}\\
      n(k)->taken = \texttt{true}\\
      n(k)->next \neq n(k)
    }{
      ..., n, ...
      \xrightarrow{pop3(k)} 
     ..., n[k \mapsto n(k)->next],...
    }\hspace{5mm}

    \inferrule[pop4]{
      cp(k) = R_3 \\
      n(k) \neq \texttt{null}\\
      n(k)->taken = \texttt{false}\\
      n(k)->ts > maxTS(k)
    }{
      ..., maxTS, cp ...
      \xrightarrow{pop4(k)} 
     ..., maxTS[k \mapsto n(k)->ts], cp[k \mapsto R_4]...
    }\hspace{5mm}

    \inferrule[pop5]{
      cp(k) = R_4 
    }{
      ..., youngest, cp ...
      \xrightarrow{pop5(k)} 
     ..., youngest[k \mapsto n(k)], cp[k \mapsto R_5]...
    }\hspace{5mm}

    \inferrule[pop6]{
      cp(k) = R_3 \\
      n(k) \neq \texttt{null}\\
      n(k)->taken = \texttt{false}\\
      n(k)->ts \leq maxTS(k)
    }{
      ..., cp, ...
      \xrightarrow{pop6(k)} 
     ...,cp[k \mapsto R_5],...
    }\hspace{5mm}

    \inferrule[pop7]{
      cp(k) = R_5 \\
      i(k) < \texttt{maxThreads}-1
    }{
      ...,i, cp, ...
      \xrightarrow{pop7(k)} 
     ...,i[k \mapsto i(k)+1], cp[k \mapsto R_2],...
    }\hspace{5mm}

    \inferrule[pop8]{
      cp(k) = R_5 \\
      youngest(k) = \texttt{null} \vee (youngest(k) \neq \texttt{null} \wedge youngest->taken)
    }{
      ...,success, cp, ...
      \xrightarrow{pop7(k)} 
     ...,success[k \mapsto \texttt{false}], cp[k \mapsto R_6],...
    }\hspace{5mm}

    \inferrule[com-pop]{
      cp(k) = R_5 \\
      youngest(k) \neq \texttt{null} \\
      youngest(k) = m \\
      d = m->data\\
      m->taken = false \\
      m' = m[taken \mapsto true]
    }{
      ...,success, youngest, cp, ...
      \xrightarrow{com(pop,d,k)} 
     ...,success[k\mapsto true], youngest[k \mapsto m'], cp[k \mapsto R_6],...
    }\hspace{5mm}

    \inferrule[pop9]{
      cp(k) = R_6 \\
      success(k) = \texttt{false}
    }{
      ..., cp, ...
      \xrightarrow{pop9(k)} 
     ..., cp[k \mapsto R_1],...
    }\hspace{5mm}
    \inferrule[ret-pop]{
      cp(k) = R_6 \\
      suc(k) = \texttt{false} \\ 
      d = yst(k)->data
    }{
      ..., cp, ...
      \xrightarrow{ret(pop,d,k)} 
     ..., cp[k \mapsto R_7],...
    }\hspace{5mm}

    \end{mathpar}
  }
 \vspace{-5mm}
  \caption{The pop derivation rules of $TSS$. We only mention the state components that are modified. Unmentioned state components have the names in the algorithm in the pre-state. $n' = n[taken \mapsto expr]$ is short for $n'->taken = expr$ and all the other fields of $n$ and $n'$ are the same.
  }
  \label{fig:transitions:TSSPop}
\vspace{-6mm}
\end{figure}

\begin{lem}
$TSS$ is a $C\cup R\cup Com(pop)$-refinement of $AbsS$. 
\end{lem}
\begin{proof}
We show that the relation $\mathit{fs}_2$ defined in Section~\ref{sec:corr_tss} is a $C\cup R\cup Com(pop)$-forward simulation  from $TSS$ to $AbsS$. For readability, we recall the definition of $\mathit{fs}_2$.

Let us make some clarifications before defining the relation. In order not to confuse nodes in TS Stack and nodes in $AbsS$, we call nodes of $AbsS$ as vertices from now on. We also define ordering relation (called traverse order) among the operations in a state of $TS$. It basically reflects the traverse order of pop operations. For two push operations $m,n \in O_a$ is state $s$ we say that $m <^{tr}_s n$ iff either $myTid(m) < myTid(n)$ or $myTid(m) = myTid(n)$ and $n_s(n)$ is reachable from $n_s(m)$ using next pointers. $\geq^{tr}$ is obtained from $<^{tr}$ in the usual way.

The relation $\mathit{fs}_2 \subseteq Q_C \rightarrow Q_{AbsS}$ contains $(s,t)$ iff the following are satisfied:
\begin{itemize}
\item[\emph{Nodes}] $k \in O_t$ iff $k$ is a push operation in $s$ ($k \in O_a$) such that either it has not inserted its node to the pool yet ( $cp_s(k) = A_i$ and $i<3$) or its node is not taken by a pop ($cp_s(k) = A_i$, $i\geq 3$ and $n_s(k)->taken = false$). 
\item[\emph{Pend/Comp}] A vertex $k \in O_t$ is pending ($\ell_t(k) = (d, \texttt{PEND})$) iff $k$ satisfies the previous condition, $x_s(k) = d$ and it is not completed in $s$ ($cp_s(k) = A_i$ and $i<6$). Similarly, this vertex is completed ($\ell_t(k) = (d, \texttt{COMP})$) iff $k$ satisfies the previous condition, $x_s(k) = d$ and it is completed in $s$ ($cp_s(k) = A_6$). Pending vertices are maximal with respect to $<_t$ i.e., if $k \in O_t$ is a pending vertex, then for all $k' \in O_t$ $k \nless_t k'$.
\item[\emph{TSOrder}] If a node has a smaller timestamp than the other node in $s$, the operations that inserted them cannot be ordered reversely in $t$. More formally, let $k, k' \in O_t$ s.t. $n_s(k)-> ts \leq n_s(k')->ts$. Then, $k' \nless_t k$.
\item[\emph{TidOrder}] Order among the nodes inserted by the same threads in $s$ must be preserved among the operations that inserted them in $t$. Let $k, k' \in O_t$ s.t. $myTid_s(k) = myTid_s(k')$ and $n_s(k)->ts < n_s(k')->ts$. Then, $k <_t k'$.
\item[Frontiers] Every maximally closed or pending vertex can be removed by a pending pop. More formally, let $k \in O_t$ such that $\ell_t(k) = (\_,\texttt{PEND})$. Then, for all pops $p$, $k \in ov_t(p)$. In the other case, let $k \in O_t$ such that $\ell_t(k) = (\_,\texttt{COMP})$ and for all other $k' \in O_t$ such that $k<_t k'$, we know $\ell_t(k') = (\_,\texttt{PEND})$. Then, for all pop operations $p$, $k \in be_t(p)$ or $k \in ov_t(p)$. 
\item[\emph{MaximalOV}] If a push $k \in O_t$ is a candidate to be removed by a pop $p$, then every other push $k'$ invoked after $k$ is a candidate to be removed by $p$ since $k$ is concurrent with $p$. More formally, let $k, k' \in O_t$ such that $k <_t k'$ and there exists a pop $p$ such that $k \in be_t(p)$ or $k \in ov_t(p)$. Then, $k' \in ov_t(p)$.
\item[\emph{MinimalBE}] If a push $k \in O_t$ has finished before the pop $p$ is invoked and yet $k$ is a candidate to be removed by $p$, other pushes completed before $k$ can not be candidates to be removed by $p$ at that state. More formally, let $k, k' \in O_t$ such that $k <_t k'$ and there exists a pop $p$ such that $k' \in be_t(p)$. Then, neither $k \in be_t(p)$ nor $k \in ov_t(b)$.
\item[\emph{ReverseFrontiers}] If all immediate followers $k' \in O_t$ of a push $k \in O_t$ are concurrent with pop $p$, then $k$ is either concurrent or maximally closed with respect to $p$. More formally, let $k \in O_t$ and for all $k' \in O_t$ such that $k \in pred_{<_t}(k')$, $k' \in ov_t(p)$, where $p$ is a pop operation. Then, $k \in ov_t(p) \cup be_t(p)$. 
%If a push $k \in O_t$ is concurrent with the pop $p$ and there exists another push $k' \in O_t$ that is immediate predecessor of $k$, then $k'$ is either concurrent or maximally closed with respect to $p$. More formally, let $k, k' \in O_t$ such that $k' \in pred_{<_t}(k)$ and $k \in ov_t(p)$ for some pop $p$. Then, either $k' \in ov_t(p)$ or $k' \in be_t(p)$. 
\item[\emph{FixReturn}] If a pop $p$ is after its commit point action in $s$, then the $rv$ value of this operation in $t$ is fixed to $youngest_s(p)->data$. More formally, Let $p$ be the pop operation such that $cp_s(p) = R_6$ and $success_s(p) = \texttt{true}$. Then, $rv_t(p) = youngest_s(p)->data$. 
\item[\emph{TraverseBefore}] If a pop operation $p$ is currently visiting node $n$, it has non-null node $y$ as the $youngest$ and there is a non-null not taken node $m$ coming before $n$ in the traverse order with a greater timestamp than $y$, then the operation that inserts $m$ must be concurrent with $p$. More formally, assume $youngest_s(p) = y$ and $ y \neq \texttt{null}$. Let $k \in O_t$ such that $n_s(k) \neq \texttt{null}$, $n_s(k)->taken = \texttt{false}$, $n_s(k) <^{tr}_s n_s(p)$ and $n_s(k)->ts \geq y->ts$. Then, $k \in ov_t(p)$.
\item[\emph{TraverseBeforeNull}] If a pop operation $p$ is currently visiting node $n$, and its $youngest$ field is \texttt{null}, then every other node $m$ coming before $n$ in the traverse order must be concurrent with $p$. More formally , let $youngest_s(p) = \texttt{null}$ and assume there exists  an operation $k \in O_t$ such that $n_s(k) \neq \texttt{null}$, $n_s(k)->taken = \texttt{false}$ and  $n_s(k) <^{tr}_s n_s(p)$. Then, $k \in ov_t(p)$. 
\item[\emph{TraverseAfter}] If a pop operation $p$ is currently visiting node $n$ that is not null and its youngest element $m$ is not null and still not taken in state $s$, then either $m$ is a candidate to be removed by $p$ in $t$ or there exists a later node $m'$ than $n$ such that $m'$ is a candidate in $t$ and it has a bigger timestamp than n. More formally, assume that there exists $k, k' \in O_t$ such that $youngest_s(p)->taken \neq false$, $youngest_s(p) = n_s(k)$ and $n_s(k') = n_s(p)$. Then, either $k \in ov_t(p) \vee k \in be_t(p)$ or there exists $k'' \in O_t$ s.t. $n_s(k'')->ts > n_s(k) ->ts$ and $k'' \in ov_t(p) \vee k'' \in be_t(p)$ and either $k' <^{tr}_s k''$ or $n_s(p) = n_s(k'') \wedge cp_s(p) = R_j \wedge j<5$. 
\end{itemize}
Next, we will show that $\mathit{fs}_2$ is really a Com(pop)-forward simulation relation. Except the trivial base case, we case-split on the transition rules. We first assume $(s, \alpha s') \in \delta_C$ and $t \in \mathit{fs}_2[s]$. Then, we find  corresponding transition $\alpha' \in \Sigma_{AbsS}$ obeying the Com(pop)-forward simulation relation conditions and obtain $t'$ such that $(t, \alpha' t') \in \delta_{AbsS}$  and $t' \in \mathit{fs}_2[s']$.

We observe that if $\alpha \in C \cup R \cup Com(pop)$, then the corresponding rule in $AbsS$ is $\alpha' = \alpha$. Otherwise, $\alpha' = \epsilon$.

Let the following derivation rule of $TSS$ be the one for describing $\alpha$:
\begin{mathpar}
    \inferrule{
      \psi
    }{
      s
      \xrightarrow{\alpha} 
      s'
    }
\end{mathpar}
and the following one be the derivation rule of $AbsS$ describing $\alpha'$ if $\alpha' \neq \epsilon$ (equivalently $\alpha' = \alpha$):
\begin{mathpar}
    \inferrule{
      \psi'
    }{
      q
      \xrightarrow{\alpha'} 
      q'
    }
\end{mathpar}
For the cases $\alpha' = \alpha$, we first need to show $\alpha'$ is enabled in state $t$ i.e., $t$ satisfies $\psi'$. If this can not be directly obtained from $s$ satisfies $\psi$ and using one or two obvious conditions on $\mathit{fs}_2$ (since $t \in \mathit{fs}_2[s]$), we show the derivation in the proof. Then, $t'$ is obtained in a unique way since $AbsS$ is deterministic on its alphabet $ \Sigma_{AbsS} = C \cup R \cup Com(pop)$. The, only other thing to show is $t' \in \mathit{fs}_2[t']$. We show this by proving that $t'$ does not violate any of the conditions of the $\mathit{fs}_2$ described above. We only explain why the new instantiations due to the difference between $s'$ and $s$ or the difference between $t'$ and $t$ do not violate the conditions. We skip the instances that we assumed while relating $s$ to $t$.

For the cases in which $\alpha' = \epsilon$, we have $t'=t$ and the only thing to show is $t \in \mathit{fs}_2[s']$. Again, we only explain why the new instantions due to the difference between $s'$ and $s$ do not violate the conditions.
\begin{itemize}
\item[\textsc{init}] $\mathit{fs}_2[{q_0}_{TSS}] =\{{q_0}_{AbsS}\}$
\item[\textsc{call-push}] The same derivation rule of $TSS$ is applied to $t$  to obtain $t'$. The premise of the rule is satisfied by $t$ trivially in the sense explained before. The new vertex $k$ is added to the $O_t$ such that $k$ is maximal, pending and every completed vertex is ordered before $k$ in $t'$. Moreover, $k$ is overlapping with every pending pop. To see that $t' \in \mathit{fs}_2[s']$ we observe the following: \emph{Nodes} condition is preserved because $k \in O_{t'}$. Since the newly added vertex $k$ is maximal and pending in $t'$, \emph{Pend/Comp} condition is preserved. \emph{Frontiers} and \emph{MaximalOV} conditions are not violated since $k$ is added to $ov(p)$ set for every pending pop operation $p$. 
\item[\textsc{push1}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$.\emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex. \emph{TSOrder} is still preserved. Timestamp of $n_{s'}(k)$ is maximal and every other nodes of push operations with maximal timestamp in $s'$ are pending vertices in $t$. Hence there can be no ordering between those pushes and $k$ in $t$ that can violate \emph{TSOrder}. Moreover, $k$ is maximal in $t$ which means that it cannot be ordered before another push $k'$ of which node has a lower timestamp. \emph{TidOrder} is also satisfied. Since $k$ is ordered after every completed push in $t$ and every other push by the same thread is completed, ordering required by the \emph{TidOrder} is present.
\item[\textsc{push2}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$. \emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex. One can also see that the \emph{TraverseBefore} condition is preserved. Let the pop $p$ visiting node $m$ and $n_{s'}(k) <^{tr}_{s'} m$. Since $k$ and $p$ are both pending in $s$ and $t \in \mathit{fs}_2[s]$, $k \in ov_t(p)$ (by the \emph{Frontiers} condition). Hence, \emph{TraverseBefore} is preserved. 
\item[\textsc{push3}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$. We consider two cases: $n_s(k)->taken$ is \texttt{true} or it is \texttt{false}. For the former case, $k \notin O_t$. The only new instantiation we check is $k \notin O_t$ does not violate \emph{Nodes} condition while relating $s'$ to $t$.

For the latter case, we have $k \in O_t$. \emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex after changing $s$ to $s'$.
\item[\textsc{push4}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$.

We consider two cases: $n_s(k)->taken$ is \texttt{true} or it is \texttt{false}. For the former case, \emph{Nodes} condition is still satisfied since $k$ remains to be not a vertex.

For the latter case \emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex. \emph{TSOrder} condition is still not violated since if $k'<_t k$, then $k'$ is a completed vertex in $s$ and $s'$. By the premise of the rule (which can be shown to hold for every operation at control point $A_4$) $i_s(k') < i_s(k)$ and consequently $n_{s'}(k')->ts < n_{s'}(k)->ts$. Since every other push by the thread of $k$ is completed, \emph{TidOrder} still continues to hold for the same reasons. \emph{TraverseAfter} condition is also preserved. Let $k'$ be the push and $p$ be the pop such that $n_s(k') = youngest_s(p)$, $n_s(k') \leq^{tr}_s n_s(k)$, $n_s(k')->ts < n_s(k)->ts$ and $k \in ov_t(p)$ or $k \in be_t(p)$. Assume $n_{s'}(k')->ts \geq N_{s'}(k)->ts$ after the action. Then, $k'$ must be a pending push both in $s$ and $s'$ by the premise of the derivation rule and $k' \in ov_t(p)$ must be true by \emph{Frontiers} condition and $t \in \mathit{fs}_2[s]$. Hence, the \emph{TraverseAfter} condition is preserved.
\item[\textsc{ret-push}] We consider two cases, $n_s(k)->taken$ is \texttt{false} or \texttt{true}. For the former case, we obtain $t'$ by applying \textsc{ret-push1} rule of $AbsS$. \emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ becomes a completed vertex in $t'$. \emph{Frontiers} condition still holds since although $k$ become a maximally closed vertex in $t'$, we have $k \in ov_{t'}(p)$ for all pending nodes $p$ (due to \emph{Frontiers} condition, $t \in \mathit{fs}_2[s]$ and $k$ was a pending operation in state $t$, $k \in ov_t(p)$).

For the latter case, we obtain $t'$ by applying \textsc{ret-push2} rule of $AbsS$. \emph{Nodes} condition is still satisfied since $k \notin O_{t'}$. 
\item[\textsc{call-pop}] The same derivation rule of $TSS$ is applied to $t$  to obtain $t'$. \emph{Frontiers} condition holds for $p = k$ relating $s'$ to $t'$ since $k' \in ov_{t'}(k)$ for every pending vertex $k'$ and $k'' \in be_{t'}(p)$ for all completed vertex $k''$. $t'$ due to action $inv(pop,k)$ applied on $t$. \emph{MaximalOV} condition holds for $p = k$ since pending vertices are maximal in $t'$ and for any maximally closed vertex $k'$ in $t'$, if $k'$ is ordered before other vertex $k''$, then $k''$ is a pending operation by definition of being maximally closed and $k'' \in ov_t(k)$ due to the changes by \textsc{inv-pop} action on $t$. \emph{MinimalBE} condition holds while relating $s'$ to $t'$ for the pop $ p = k$ because only maximally closed vertices are in $be(k)$ and if a push $k'$ is ordered before a maximally closed push $k''$ in $t$, neither $k'' \in be_{t'}(k)$ (since $k''$ is not maximally closed) nor $k'' \in ov_{t'}$ (since $k''$ cannot be pending). \emph{ReverseFrontiers} condition holds while relating $s'$ to $t'$ for the pop $p=k$ because, if $k'' \in ov_{t'}(k)$ for all immediate successors of $k'$ in $t$, then $k'' $ are pending vertices (due to \emph{call-pop} action of $AbsS$), $k'$ is a maximally closed vertex and $k' \in be_{t'}(k)$ (due to \emph{call-pop} action of $AbsS$).
\item[\textsc{pop1}]We have $t' = t$ and $t \in \mathit{fs}_2[s']$.
\item[\textsc{pop2}]We have $t' = t$ and show $t \in \mathit{fs}_2[s']$. \emph{TraverseBefore} condition while relating $s'$ to $t$ still holds for $p=k$. Assume $youngest_{s'}(k) = y$ is a non-null node. Then, for all nodes $m$ in $s'$ such that $n_s(k) \leq^{tr}_{s'} m <^{tr}_{s'} n_{s'}(k)$ we have $m->ts < y->ts$ in $s'$ because  $n_s(k)->ts > m->ts$ (since $n_s(k)$ is added to the pool after $m$ by the same thread) and $y->ts \geq n_s(k)->ts$ in $s'$ (since either $youngest_{s'}(k) = n_s(k)$ or $youngest_{s'}(k)->ts > n_s(k)->ts$). \emph{TraverseAfter} does not have any new instatiations since the guard mentions the nodes after $n_s(k)$ while relating $s$ to $t$ whereas it mentions nodes after or including $n_{s'}(k)$ which contains the all nodes in the former case.
\item[\textsc{pop3}]We have $t' = t$ and $t \in \mathit{fs}_2[s']$.
\item[\textsc{pop4}]We have $t' = t$ and $t \in \mathit{fs}_2[s']$.
\item[\textsc{pop5}]We have $t' = t$ and show $t \in \mathit{fs}_2[s']$. 
\emph{TraverseBefore} condition while relating $s'$ to $t$ still holds for $p=k$ since $youngest_s(k)->ts < youngest_{s'}(k)->ts$ and \emph{TraverseBefore} holds while relating $s$ to $t$.

\emph{TraverseAfter} condition also continues to hold for $p=k$. There are two possible cases: $youngest_s(k) = \texttt{null}$ or not.

First, consider the former case. Since \emph{TraverseBeforeNull} is satisfied while relating $s$ to $t$, for every operation $k', k'' \in O_t$ such that  $k'' <^{tr}_s k'$ and $n_s(k') = youngest_{s'}(k)$ we have $k'' \in ov_t(k)$. Consider all such $k''$ such that $n_s(k'')->ts > n_s(k')->ts$. If there exists such a $k''$ such that $k' \in pred_{<_t}(k'')$, then $k' \in ov_t(k) \cup be_t(k)$ since \emph{ReverseFrontiers} condition holds relating $s$ to $t$. Otherwise, either $k'$ is maximal in $t$ or all the vertices $v$ ordered after $k'$ in $t$ we have $v >^{tr}_s k'$. Then, either $k'$ or one of these $v$ vertices must be maximal in $t$ and must be in $be_t(k) \cup ov_t(k)$ since \emph{Frontiers} condition holds (one of them is maximal in $t$) while relating $s$ to $t$.

Second, assume there exists push operations $j, k'$ such that $n_s(j) = youngest_s(k) \neq \texttt{null}$ and $n_s(k') = n_s(k) = youngest_{s'}(k)$ . Since \emph{TraverseBefore} is satisfied while relating $s$ to $t$, if there exists a push $k'' <^{tr}_s k'$ such that $n_s(k'')$ is not taken and $n_s(k'')->ts \geq n_s(j)->ts$, then $k'' \in ov_t(k)$. Then, for all $k'' <^{tr}_s k'$ such that $n_s(k'')$ is not taken and $n_s(k'')->ts \geq n_s(k')->ts$, then $k'' \in ov_t(k)$ since $n_s(k')->ts \geq n_s(j)->ts$. If there exists such a $k''$ such that $k' \in pred_{<_t}(k'')$, then $k' \in ov_t(k) \cup be_t(k)$ since \emph{ReverseFrontiers} condition holds relating $s$ to $t$. Otherwise, either $k'$ is maximal in $t$ or all the vertices $v$ ordered after $k'$ in $t$ we have $v >^{tr}_s k'$. Then, either $k'$ or one of these $v$ vertices must be maximal in $t$ and must be in $be_t(k) \cup ov_t(k)$ since \emph{Frontiers} condition holds (one of them is maximal in $t$) while relating $s$ to $t$.

\item[\textsc{pop6}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$. \emph{TraverseAfter} continues to hold while relating $s'$ to $t$ for $p=k$. Let $k', k'' \in O_t$ such that $youngest_s(k) = n_s(k')$, $n_s(k) = n_s(k'')$ and $k' \notin  ov_t(k) \cup be_t(k)$. Note that $k' <^{tr}_s k''$. Then, $n_s(k'')->ts < n_s(k')->ts$ since $n_s(k'')->ts < maxTS(k)$ and $maxTS(k) = n_s(k')->TS$ ($n_s(k')->ts$ cannot be \texttt{MAX\_INT} since $k'$ would be pending and $k' \in ov_t(k)$ otherwise). Hence, there exists another push $j$ such that $j >^{tr}_s$ and $j \in ov_t(k) \cup be_t(k)$.

\item[\textsc{pop7}] We have $t' = t$ and $t \in \mathit{fs}_2[s']$.
\item[\textsc{pop8}] We have $t' = t$ and $t \in \mathit{fs}_2[s']$.
\item[\textsc{com-pop}] $t'$ is obtained by applying \textsc{com-pop1} rule of $AbsS$.
We first show that precondition of \textsc{com-pop1} rule of $AbsS$ si satisfied by $t$. If $com(pop,d,k)$ removes a  node $n$ such that there exists a push $k'$ such that $n_s(k') =n$ in $s$, then $k' \in O_t$ since it is non-null and not taken. Moreover, $k' \in ov_t(k) \cup be_t(k)$ since \emph{TraverseAfter} is preserved while relating $s$ to $t$ and all the nodes that come after $n_s(k)$ in terms of traverse order in $s$ have lower timestamp values than $n_s(k)->ts$ and $n_s(k)->ts \leq youngest_s(k)->ts$.

Next, we show that $t' \in \mathit{fs}_2[s']$. We case split on the conditions of $\mathit{fs}_2$ considering new instantiations.

\emph{Nodes} condition is still preserved after $k$ removes the node pushed by operation $k'$ in $s$ since $k' \notin O_{t'}$ anymore by due to $com(pop,d,k)$ action.

\emph{Frontiers} condition is still preserved if $k$ removes the vertex $k'$ and makes another $k''$ maximally closed in $t$. Since all the other nodes $j$ ordered after $k''$ (except possibly $k'$) in $t$ are pending, $j \in ov_t(p)$ (due to \emph{Frontiers} condition while relating $s$ to $t$) for some pending pop $p \neq k$. Then, $k'' \in be_{t'}(p)$ by $com(pop,d,k)$ action.

For the \emph{MinimalBE} condition, we do not have a new instance. If $k' \in be_{t'}(p)$ becomes true although $k' \notin be_t(p)$, we cannot have $k'' \in O_{t'}$ such that $k' \in pred_{<_{t'}}(k'')$ and $k'' \in be_{t'}(p)$ since $com(pop,d,k)$ does not add $k''$ to $ov(p)$ if its successor is not pending with respect to $p$.

\emph{ReverseFrontiers} condition is still preserved. If $k$ removes the vertex $k'$ and there exists an immediate predecessor $k''$ of $k'$ such that all of immediate successors of $k''$ are in $ov_{t'}(p)$, then $k'' \in ov_{t'}(p)$ due to the action $com(pop,d,k)$.

\emph{TraverseAfter} condition is still preserved after $k$ removes the node of push $k'$. Let $p \neq k$ be another pop operation such that $n_s(j) = youngest_s(p)$ for some push $j$ and $n_s(k')$ be the only node such that $n_s(k')->ts > youngest_s(p)->ts$ and $n_s(k')$ comes after $n_s(p)$ in the traverse order of $s$ and $k' \in ov_t(p) \cup be_t(p)$. Hence, there is no $k''$ such that $n_s(k'')$ comes after $n_s(p)$ in the traverse order and $j <_t k''$ except $k'$ (i). In other direction, if for all $k'' \in O_t$ such that $n_s(k'')$ comes before $n_s(p)$ in the traverse order and $n_s(k'')->ts > youngest_s(p)->ts$ , then $k'' \in ov_t(p)$ since \emph{TraverseBefore} condition holds while relating $s$ to $t$. Then, for all $k'' \in O_t$ such that $n_s(k'')$ comes before $n_s(p)$ in the traverse order of $s$ and $k'' >_t j$ implies $k'' \in ov_t(p)$ since $n_s(k'')->ts > n_s(j)->ts$ if $k'' >_t j$ (ii). Then, for all $k'' \in O_t$ such that if $k'' >_t j$, then $k'' \in ov_t(p)$ except $k'$ due to (i) and (ii). If $j \nless_t k'$, then $j \in ov_t(p) \cup be_t(p)$ since \emph{ReverseFrontiers} hold while relating $s$ to $t$ and $j \in ov_{t'} \cup be_{t'}$ after applying the action $com(pop,d,k)$. Otherwise, if $j <_t k'$, then $k \in be_{t'}$ after applying $com(pop,d,k)$.

\emph{FixReturn} condition continues to hold. If $com(pop,d,k)$ removes the node pushed by $k'$ in $s$, then $com(pop,d,k)$ removes the vertex $k'$ (assuming data independece) and $youngest_s(k')->data = \ell_t(k')_1$. Then, $youngest_{s'}(p)->data = rv_{t'}(p)$ after applying commit actions at both sides.

\item[\textsc{pop9}]We have $t' = t$ and $t \in \mathit{fs}_2[s']$.
\item[\textsc{ret-pop}] $t'$ is obtained by applying \textsc{ret-pop} rule of $AbsS$ and $t' \in \mathit{fs}_2[s']$.
\end{itemize}
\end{proof}
\textcolor{red}{TODO: Check ReverseFrontiers for the cases before compop.}
\textcolor{red}{TODO: Check TraverseBefore for the cases before pop2.}
\textcolor{red}{TODO: CheckTraverseBeforeNull for the cases befoere pop 5.}
\textcolor{red}{TODO: Apply Dr. Enea's comments.}
\textcolor{red}{TODO: Try case split on fs conditions instantiations check.}
\textcolor{red}{TODO: Add Flow Diagrams to the paper.} 
\textcolor{red}{TODO: }

### 2017-01-25 draftCAV/app-absImplStack.tex
finished stacks

@@
 	\item if a pop didn't commit in $s$ then it is pending in $t$ and it may have been linearized, i.e., for every $k$, if $cp(k)=R_1$ then $cp^0(k)\in \{R_1,R_2\}$,
-	\item there exists a partial function $g: \{k: cp(k)=R_1\} ~> O$ which associates uncommitted pops to pushes in $O$ such that:
+	\item there exists a partial injective function $g: \{k: cp(k)=R_1\} ~> O$ which associates uncommitted pops to pushes in $O$ such that:
 	\begin{itemize}
@@
 
-In the following, we show that indeed $bs$ is a normal $C\cup R$-backward simulation from $AbsS$ to $AbsS_0$.
+In the following, we show that indeed $bs$ is a normal $C\cup R$-backward simulation from $AbsS$ to $AbsS_0$:
+\begin{itemize}
+	\item Let $s\xrightarrow{inv(push,d,k)} s'$ be a transition in $AbsS$ and $(s',t')\in bs$.
+We consider two cases depending on whether the value $d$ occurs on a position $i$ in the sequence $\sigma$ of $t'$ or not.
+If it occurs, let $t$ be a $AbsS_0$ state where essentially, the component $\sigma$ is the prefix of the sequence $\sigma$ of $t'$ that contains the first $i-1$ positions (except for some set of pushes that will be defined hereafter, all operations are at the same control point).
+Let $@t$ be the following $AbsS_0$ trace:
+	\begin{align*}
+	@t=inv(push,d,k),\ lin(push,d,k),\ lin(push,d_{i+1},k_{i+1}),\ldots,lin(push,d_{n-1},k_{n-1})
+	\end{align*}
+	where $d_j$ is the value on position $j$ in the sequence $\sigma$ of $t'$ and $n$ is the length of this sequence (we assume that positions are indexed starting from $0$). Let $k_i=k$.
+	For every $k_j$ with $i\leq j\leq n-1$, we must have that $cp(k_j)=A$ in $t'$. We take $cp(k_j)=A_1$ in $t$ for every $j\geq i+1$ and $cp$ undefined for $k_i$.
+	We have that $t\xrightarrow{@t} t'$ is a valid sequence of transitions of $AbsS_0$ and $(s,t)\in bs$ (the latter can be proved by taking the same function $g$ used in establishing that $(s',t')\in bs$).
+	Now, assume that the value $d$ is not in the sequence $\sigma$ of $t'$.
+We consider an $AbsS_0$ state $t$ where the component $\sigma$ is the same as the one in $t'$.
+There are two sub-cases depending on whether there exists a pending pop $k'$ such that $g(k')=k$ when establishing that $(s',t')\in bs$.
+If it exists, the operations are at the same control point in both $t$ and $t'$ except for the push $k$ for which $cp(k)$ is undefined in $t$, and the the pop $k'$ for which we take $cp(k')=R_0$ in $t$.
+We have that
+	\begin{align*}
+	t\xrightarrow{inv(push,d,k),\ lin(push,d,k)\ lin(pop,d,k')} t'
+	\end{align*}
+	in $AbsS_0$. If there exists no such pop $k'$, it can be easily seen that there exists $t$ such that $t\xrightarrow{inv(push,d,k)} t'$ and $(s,t)\in bs$.
 
-TODO FILL IN THE STEPS
+	\item Let $s\xrightarrow{inv(pop,k)} s'$ be a transition in $AbsS$ and $(s',t')\in bs$.
+We consider two cases depending on whether in the function $g$ used to relate $s'$ to $t'$ we have that $k\in dom(g)$.
+In other words, either the newly invoked pop operation $k$ did not linearize yet ($k\not\in dom(g)$) or it linearizes and removes an element inserted by a linearized push ($k\in dom(g)$).
+The second case also splits into two sub-cases: The value removed by pop $k$ is inserted by a push $k'=g(k)$ that is still pending or the push has returned.
+We will look at all three cases separately.
+	The easiest one is the first case. There exists some $t$ where essentially the component $\sigma$ is the same as the one in $t'$, such that $t\xrightarrow{inv(pop,k)} t'$ and $(s,t)\in bs$.
 
-We will provide a backward simulation relation $bs$ between states of $L_I$ and states of $L_A$.
+	For the first sub-case of the second case, we take an $AbsS_0$ state $t$ where $\sigma_t = d\cdot\sigma_{t'}$ (we use $\sigma_t$ to denote the component $\sigma$ in $t$).
+It must happen that $cp^0_{t'}(k)=R_2$.
+The operations are at the same control point in both $t$ and $t'$, except for $k$ in which case $cp^0$ is undefined.
+We have that $t\xrightarrow{inv(pop,k),\ lin(pop,d,k')} t'$ and $(s,t)\in bs$.
+The latter holds because essentially, $k'$ is a maximal node in $s'$ (since it is pending).
+
+	For the second sub-case, we define an $AbsS_0$ state $t$ where the sequence $\sigma$ is the minimal prefix of $\sigma_{t'}$ that includes the value $d$ added by $k'$.
+Let $i$ be the index of this value in $\sigma_{t'}$ and $k_{j}$ with $i<j$ the identifiers of the pushes that added the values following $d$ in $\sigma_{t'}$.
+Let $@t$ be the following $AbsS_0$ trace:
+	\begin{align*}
+	@t=inv(pop,k),\ lin(pop,d,k),\ lin(push,d_{i+1},k_{i+1}),\ldots,lin(push,d_{n-1},k_{n-1})
+	\end{align*}
+	where $d_j$ is the value on position $j$ in the sequence $\sigma_{t'}$ and $n$ is the length of this sequence.
+We have that $t\xrightarrow{@t} t'$ is a valid sequence of transitions of $AbsS_0$ and $(s,t)\in bs$.
+The latter relies on the fact that $k'$ is a greatest completed push in $s$ and all pushes $k_j$ with $j>i$ are pending in $s$.
+
+	\item Let $s\xrightarrow{com(pop,d,k)} s'$ be a transition in $AbsS$ and $(s',t')\in bs$.
+When this transition results in removing a greatest completed push or a pending push in $s$, then there exists an $AbsS_0$ state $t$ such that $t\xrightarrow{@t} t'$ is a valid sequence of $AbsS_0$ transitions and $(s,t)\in bs$, for some $t$ and $@t$ defined as in the second case of $inv(pop,k)$.
+When it removes a completed push which is followed by other completed pushes (in the happens-before in $s$), then we pick $t=t'$.
+We have that $t\xrightarrow{\epsilon} t'$ and $(s,t)\in bs$ (for the latter we must choose a function $g$ such that $g(k)=k'$ where $k'$ is the push removed by the $AbsS$ transition.
+
+	\item Let $s\xrightarrow{ret(push,k)} s'$ be a transition in $AbsS$ and $(s',t')\in bs$.
+We consider two cases depending on whether the happens-before in $s$ contains push $k$.
+If it contains push $k$, there are two sub-cases: (1) if its input is present in $\sigma_{t'}$ then there exists an $AbsS_0$ state $t$ such that $t\xrightarrow{ret(push,k)} t'$ is a valid sequence of $AbsS_0$ transitions and $(s,t)\in bs$, and (2) otherwise, we take a state $t$ where essentially, $\sigma_t = d\cdot\sigma_{t'}$ for which we have that $t\xrightarrow{lin(pop,d,k),\ ret(push,k)} t'$ and $(s,t)\in bs$.
+If the happens-before in $s$ doesn't contain push $k$, then there exists an $AbsS_0$ state $t$ such that $t\xrightarrow{ret(push,k)} t'$.
+
+	\item The case of pop returns $ret(pop,k)$ is trivial. Such transitions are simulated by $ret(pop,k)$ transitions of $AbsS_0$.
+\end{itemize}
+
+%TODO FILL IN THE STEPS
+%
+%
+%We will provide a backward simulation relation $bs$ between states of $L_I$ and states of $L_A$.
 Our relation $bs \subseteq Q_I \times Q_A$ relates state $q=(so_q, f_q)$ to $q' =(s_{q'}, f_{q'}$ iff (i) for all operation identifiers $k \in \mathbb{N}$, if $f_q(k) \in \{N,R_1,R_2\}$, then $f_q(k) = f_{q'}(k)$; if $f_q(k) = A_0$, then $f_{q'}(k) = A_0$ and the data value $d$ associated with this add operation is not inserted to $s_{q'}$ or $f_{q'}(k) = A_1$;  if $f_q(k) = A_1$, then $f_{q'}(k) = A_2$; if $f_q(k) = R_0$, then $f_{q'}(k)=R_0$ or $f_{q'}(k)=R_1$; (ii) Let us call a pop operation pending if $f_{q}(k) = R_0$ and $PP_q$ be set of pending pops.
 There there exists a function $g: PP_q \rightarrow ND_q \cup \{\texttt{NONE}\}$ such that $k \in g(k).mc$ or $k \in g(k).con$ for all pop operation identifiers $k$ such that $g(k) \neq \texttt{NONE}$ and $g$ is one-to-one if we neglect \texttt{NONE}; $s_{q'}$ is obtained by extending $so_q$ to a total order in which pending nodes may not take place and nodes (open or closed) $n$ such that there exists a pop operation with identifier $k$ so that $g(k)=n$ surely do not take place, (iii) $g(k) = \texttt{NONE}$ implies $f_{q'}(k) = R_0$ and $g(k) \neq \texttt{NONE}$ implies $f_{q'}(k) =R_1$, (iv) if $g_{q'}(k) =n$ and $n$ is a pending node, then $f_{q'}(k') = A_1$ where $k'$ is the operation identifier part of $n$, (v) if there is a pending node with identifier $k$ and it takes part in the linearization, then $f_{q'}(k) = A_1$.
-
-Now, we will show that $bs$ is a backward simulation relation:
-\begin{itemize}
-\item[$\langle i \rangle$] $bs[{q_0}_I] =\{{q_0}_A\}$
-\item[$\langle ii-a-push \rangle$] Let $(q,inv(push,d,k),q') \in \delta_I$ and $t' \in bs[q']$.
+%
+%Now, we will show that $bs$ is a backward simulation relation:
+%\begin{itemize}
+%\item[$\langle i \rangle$] $bs[{q_0}_I] =\{{q_0}_A\}$
+%\item[$\langle ii-a-push \rangle$] Let $(q,inv(push,d,k),q') \in \delta_I$ and $t' \in bs[q']$.
 We consider two cases: Either the newly added node in $q'$ takes place in the linearization and there exists an index $i$ such that $s_{t'}(i) = d$ or this new node does not exist in the linearization.
@@
 Then, it is easy to see that $t \xrightarrow{inv(push,d,k)} t'$ holds and $t \in bs[q]$.
-\item[$\langle ii-a-pop \rangle$] Let $(q, inv(pop,k),q') \in \delta_A$ and $t' \in bs[q']$.
+%\item[$\langle ii-a-pop \rangle$] Let $(q, inv(pop,k),q') \in \delta_A$ and $t' \in bs[q']$.
 We will again consider two cases: When relating $t'$ to $q'$, either $g(k) = \texttt{NONE}$ or $g(k)$ is a node in $q'$.
@@
 It is true because $so_q = so_{q'}$, $k'$ is a maximally closed node in $so_q$ and all the nodes with identifier $k_j$ ($j>i$) are pending nodes.
-\item[$\langle ii-c \rangle$] Let $(q, com(pop,d,k), q') \in \delta_I$ and $t' \in bs[q']$.
+%\item[$\langle ii-c \rangle$] Let $(q, com(pop,d,k), q') \in \delta_I$ and $t' \in bs[q']$.
 We will consider two cases.
@@
 Then, $t \xrightarrow{\epsilon} t'$ is the path in $L_A$ and one can show that $t \in bs[q]$ by choosing $g(k)$ as the node that is removed.
-\item[$\langle ii-b-push \rangle$] Let $(q,ret(push,k),q') \in \delta_I$ and $t' \in bs[q']$.
+%\item[$\langle ii-b-push \rangle$] Let $(q,ret(push,k),q') \in \delta_I$ and $t' \in bs[q']$.
 We consider two cases.
@@
 Then, $t \xrightarrow{ret(push,k)} t'$ is a path in $L_A$ and $t \in bs[q]$.
-\item[$\langle ii-b-pop \rangle$] Let $(q,ret(pop,d,k),q') \in \delta_I$ and $t' \in bs[q']$.
+%\item[$\langle ii-b-pop \rangle$] Let $(q,ret(pop,d,k),q') \in \delta_I$ and $t' \in bs[q']$.
 We pick $s_t = s_{t'}$ and $f_t = f_{t'}$ except that $f_t(k) =R_1$.
 One can see that $t \xrightarrow{ret(pop,d,k)} t'$ is a valid action in $L_A$ and $t \in bs[q]$.
-\end{itemize}
+%\end{itemize}
 \end{proof}

### 2017-01-25 draftCAV/queue.tex
finished stacks

@@
 The linearization points of the enqueue methods are not fixed, they depend on dequeue operations executing in the future.
-Consider the following trace with two concurrent enqueues (${\tt i}(k)$ represents the value of {\tt i} in operation $k$): $inv(enq,x,1)$, $inv(enq,y,2)$, ${\tt i}(1) = \mbox{{\tt bck++}}$, ${\tt i}(2) = \mbox{{\tt bck++}}$, ${\tt items[i(}2{\tt )]} = y$.
+Consider the following trace with two concurrent enqueues (that is, ${\tt i}(k)$ represents the value of {\tt i} in operation $k$): $inv(enq,x,1)$, $inv(enq,y,2)$, ${\tt i}(1) = \mbox{{\tt bck++}}$, ${\tt i}(2) = \mbox{{\tt bck++}}$, ${\tt items[i(}2{\tt )]} = y$.
 %\vspace{-1.5mm}

### 2017-01-25 draftCAV/stack.tex
small things
finished stacks

@@
 \vspace{-6mm}
-\subsection{A Correctness Proof For Time-Stamped Stack}
+\subsection{A Correctness Proof For Time-Stamped Stack}\label{sec:corr_tss}
 \vspace{-1mm}
@@
 \begin{itemize}
-	\item every pop overlaps with all pending pushes, i.e., $k_1\in O$ is pending implies $k_1\in ov(k)$ for each $k, k_1$
-	\item the greatest completed pushes in $<$ are either overlapping with a pop $k$ or they were the greatest completed pushes when $k$ started, i.e., every such push belongs either to $be(k)$ or $ov(k)$ for each $k$
-	\item for every push that overlaps with a pop $k$ or was maximal in $<$ when $k$ started, its successors are overlapping with $k$, i.e., $k_1\in be(k)\cup ov(k)$ and $k_1 < k_2$ implies $k_2 \in ov(k)$ for each $k, k_1, k_2$
-	\item $be(k)$ and $ov(k)$ don't contain predecessors of pushes from $be(k)$, i.e., $k_1 < k_2$ and $k_2 \in be(k)$ implies $k_1\not\in be(k)\cup ov(k)$ for each $k, k_1, k_2$
-	\item immediate predecessors of pushes overlapping with a given pop $k$ are either overlapping with $k$ or they were maximal in $<$ when $k$ started, i.e., $k_2\in pred_{<}(k_1)$ and $k_1\in ov(k)$ implies $k_2\in be(k)\cup ov(k)$ for each $k,k_1,k_2$
-	\item a pop $k$ that reached a node with timestamp $\tau$ (its variable {\tt n} points to this node) overlaps with any push that has a timestamp bigger than $\tau$ and which created a node that occurs in {\tt pools} before the node reached by $k$, i.e., ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k_1)$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
-	\item if the variable {\tt youngest} of a pop $k$ points to a node which is not taken, then this node was created by a push in $be(k)\cup ov(k)$ or the node currently reached by $k$ is followed in {\tt pools} by another node which was created by a push in $be(k)\cup ov(k)$, i.e., ${\tt youngest}_s(k)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(k)={\tt n}_s(k_2)$ implies $k_1\in be(k)\cup ov(k)$ or that there exists $k_3\in O$ such that ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in be(k)\cup ov(k)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and TODO $k$ is traversing the last list in the array {\tt pools}, for each $k, k_1,k_2$.
+	\item a pop $k$ with ${\tt youngest}\neq {\tt null}$ that reached a node with timestamp $\tau$ (its variable {\tt n} points to this node) overlaps with every push that created a node with a timestamp bigger than $\tau$ and which occurs in {\tt pools} before the node reached by $k$, i.e., ${\tt youngest}_s(k)\neq{\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k_1)$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
+	\item a pop $k$ with ${\tt youngest}={\tt null}$ overlaps with every push that created a node which occurs in {\tt pools} before the node reached by $k$,
+i.e., ${\tt youngest}_s(k)={\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
+	\item if the variable {\tt youngest} of a pop $k$ points to a node which is not taken, then this node was created by a push in $be(k)\cup ov(k)$ or the node currently reached by $k$ is followed in {\tt pools} by another node which was created by a push in $be(k)\cup ov(k)$, i.e., ${\tt youngest}_s(k)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(k)={\tt n}_s(k_2)$ implies $k_1\in be(k)\cup ov(k)$ or that there exists $k_3\in O$ such that ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in be(k)\cup ov(k)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and TODO $k$ is traversing the last list in the array {\tt pools}, for each $k, k_1,k_2$
+%	
+%	\item every pop overlaps with all pending pushes, i.e., $k_1\in O$ is pending implies $k_1\in ov(k)$ for each $k, k_1$
+%	\item the greatest completed pushes in $<$ are either overlapping with a pop $k$ or they were the greatest completed pushes when $k$ started, i.e., every such push belongs either to $be(k)$ or $ov(k)$ for each $k$
+%	\item for every push that overlaps with a pop $k$ or was maximal in $<$ when $k$ started, its successors are overlapping with $k$, i.e., $k_1\in be(k)\cup ov(k)$ and $k_1 < k_2$ implies $k_2 \in ov(k)$ for each $k, k_1, k_2$
+%	\item $be(k)$ and $ov(k)$ don't contain predecessors of pushes from $be(k)$, i.e., $k_1 < k_2$ and $k_2 \in be(k)$ implies $k_1\not\in be(k)\cup ov(k)$ for each $k, k_1, k_2$
+%	\item if all immediate successors of a given push $k_1$ are overlapping with a pop $k$, then $k_1$ is either overlapping with $k$ or it was a greatest completed push when $k$ started, 
+%	if $k_2\in ov(k)$, for every $k_2$ s.t. $k_1 \in pred_{<}(k_2)$, then $k_1\in be(k)\cup ov(k)$ for each $k,k_1$
 \vspace{-2mm}
 \end{itemize}
+There are some more constraints on $be$ and $ov$ that can be seen as invariants of $AbsS$, i.e., $be(k)$ and $ov(k)$ don't contain predecessors of pushes from $be(k)$ (for each $k, k_1, k_2$, $k_1 < k_2$ and $k_2 \in be(k)$ implies $k_1\not\in be(k)\cup ov(k)$).
+They can be found in Appendix~\ref{app:tss}.
+
 Finally, for every pop operation $k$ such that ${\tt success}(k)={\tt true}$, we have that $rv(k)={\tt youngest}(k)\text{\tt->data}$.
 
-The proof that $\mathit{fs}_2$ is indeed a forward simulation from $\mathit{TSS}$ to $AbsS$ follows the same lines as the one given for the Herlihy\&Wing Queue. It can be found in Appendix~\ref{}.
+The proof that $\mathit{fs}_2$ is indeed a forward simulation from $\mathit{TSS}$ to $AbsS$ follows the same lines as the one given for the Herlihy\&Wing Queue.
+It can be found in Appendix~\ref{app:tss}.

### 2017-01-25 draftCAV/intro.tex
fixed an issue in AbsS definition
first version of the intro

--- 2017-01-21 draftCAV/intro.tex
@@
 
-\newpage~\newpage
+Efficient concurrent implementations of atomic collections including stacks and queues are vital to modern computer systems.
+Programming them is however error prone.
+To minimize synchronization overhead between concurrent method invocations, implementors avoid blocking operations like lock acquisition, allowing methods to execute concurrently.
+However, concurrency risks unintended inter-operation interference, and risks conformance to atomic reference implementations.
+Conformance is formally captured by \emph{observational refinement}, which assures that all behaviors of programs
+using these efficient implementations would also be possible were the atomic reference implementations used instead.
+
+Observational refinement can be formalized as a trace inclusion problem, and the latter can itself be reduced to an assertion (invariant) checking problem, but this requires in general the consideration of additional history and prophecy variables.
+Alternatively, verifying observational refinement requires in general establishing both a forward simulation {\em and} a backward simulation.
+While simulations are natural concepts, reasoning about backward simulations, corresponding to the use of prophecy variables, is however less intuitive, and from the practical point of view, backward reasoning is in general hard and complex for programs manipulating data structures.
+Therefore, a crucial issue is to understand the limits of forward reasoning in establishing observational refinement.
+More precisely, we consider that an important question is to determine for which concurrent abstract data structures, and for which types of their implementation, is it possible to carry out an observational refinement proof using only forward reasoning.
+
+In order to get rid of backward simulations (and equivalently of prophecy variables), it is necessary to have reference implementations of the considered abstract data structures that are {\em deterministic}.
+Interestingly, determinism allows also to simplify the forward simulation checking problem.
+Indeed, in this case, this problem can be reduced to an invariant checking problem.
+Basically, the simulation relation can be seen as an invariant of the system composed by the two compared programs.
+Therefore, existing methods and tools for invariant checking can be leveraged in the context of proving observational refinement.
+
+But, in order to determine precisely what is meant by determinism, an important point is to fix the alphabet of observable events along computations. 
+Typically, to reason about observational refinement between two library implementations, the only observable events are the calls and returns corresponding to the method invocations along computations.
+This means that only the external interface of the library is considered to compare behaviors, and nothing else from the implementations is exposed.
+Unfortunately, it can be shown that in this case, it is impossible to have deterministic atomic reference implementations for common data structures such as stacks and queues (see, e.g., \cite{DBLP:conf/cav/SchellhornWD12}).
+Then, an important question is what is the necessary amount of information that should be exposed by the implementations to overcome this problem ?
+
+One approach addressing this question is based on the notion of linearizability \cite{journals/toplas/HerlihyW90} and its correspondence with observational refinement (w.r.t.
+atomic reference implementations) \cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}.
+Linearizability of a computation (of some concurrent implementation) means that each of the method invocations can be seen as happening at some point, called {\em linearization point}, occurring somewhere between the call and return events of that invocation, such that the obtained sequence of linearization points along the computation defines a sequence of operations that is possible in the atomic reference implementation.
+Proving the existence of such sequences of linearization points, for all the computations of a concurrent library, is a complex problem \cite{journals/iandc/AlurMP00,conf/esop/BouajjaniEEH13,DBLP:conf/netys/Hamza15}.
+%
+However, proving linearizability becomes less complex when linearization points are fixed for each method, i.e., associated (once and for all) with the execution of a designated statement in its source code \cite{conf/esop/BouajjaniEEH13}.
+In this case, we can consider that libraries expose not only the call and returns events, but also events signaling linearization points.
+By extending this way the alphabet of observable events, it becomes straightforward to define {\em deterministic} atomic reference implementations.
+Therefore, proving linearizability can be carried out as an invariant checking problem when linearization pointst are fixed, e.g.,~\cite{conf/ppopp/VafeiadisHHS06,conf/cav/AmitRRSY07,conf/vmcai/Vafeiadis09,conf/tacas/AbdullaHHJR13}.
+%
+While this approach can be useful in some cases, it is unfortunately not applicable to many efficient implementations such as the LCRQ queue~\cite{DBLP:conf/ppopp/MorrisonA13} (based on the principle of the Herlihy\&Wing queue \cite{journals/toplas/HerlihyW90}), and the Time-Stamped Stack \cite{DBLP:conf/popl/DoddsHK15}.
+The proofs of linearizability of these implementations are highly nontrivial, very involved, and hard to read, understand and automatize.
+Therefore, the crucial question we address is what is precisely the kind of information that is necessary to expose in order to obtain deterministic atomic reference implementations for such data structures, allowing to derive simple and natural linearizability proofs for such complex implementations, which can be carried out as systematic invariant checking proofs that are amenable to automation?
+
+We observe that the main difficulty in reasoning about these implementations is that, linearization points of enqueue/push operations occurring along some given computation, depend in general on the linearization points of dequeue/pop operations that are occurring later in that computation, at arbitrary far positions.
+Therefore, since linearization points for enqueue/push operations cannot be determined in advance, the information that could be fixed and exposed can concern only the dequeue/pop operations.
+
+One first idea is to consider that linearization points are fixed for dequeue/pop methods and only for these methods.
+We show that under the assumption that implementations expose linearizations points for these methods, it is possible to define deterministic atomic reference implementations for both queues and stacks.
+We show that this is indeed useful by providing a simple proof of the Herlihy\&Wing queue (based on establishing a forward simulation) that can be carried out as an invariant checking proof.
+
+However, in the case of the Time-Stamped Stack, fixing linearization points of pop operations is actually too restrictive.
+Nevertheless, we show that our approach can be generalized to handle this case.
+The key idea is to reason about what we call {\em commit points}, and  that correspond roughly speaking to the last point a method accesses to the shared data structure during its execution.
+We prove that by exposing commit points (instead of linearization points) for pop methods, we can still provide deterministic reference implementations.
+We show that using this approach leads to a quite simple proof of the Time-Stamped Stack.
+
+%Efficient implementations of concurrent objects such as semaphores, locks, and atomic collections including stacks and queues are vital to modern computer systems.
+Programming them is however error prone.
+To minimize synchronization overhead between concurrent object-method invocations, implementors avoid blocking operations like lock acquisition, allowing methods to execute concurrently.
+However, concurrency risks unintended inter-operation interference, and risks conformance to atomic reference implementations.
+Conformance is formally captured by \emph{observational refinement}, which assures that all behaviors of programs
+%using these efficient implementations would also be possible were the atomic
+%reference implementations used instead.
+%An equivalent property is \emph{linearizability}~\cite{journals/toplas/HerlihyW90} which requires that each concurrent execution
+%has a linearization which is a valid sequential execution according to a
+%specification, given by an abstract data type or reference implementation.
+%
+%Existing proof methods for refinement/linearizability combine reductions to assertion checking with compositional reasoning techniques for proving the assertions.
+%
+%The reductions to assertion checking are based on simulation relations: the assertions describe a simulation relation to standard atomic reference implementations
+%
+%No surprise, simulations are the standard approach for proving refinement
+%
+%This works smoothly for particular classes of implementations, e.g., with fixed linearization points.
+The explanation is that in this case, forward simulations are enough and the reference implementation is deterministic when the linearization points are observable
+%
+%Why they are enough ? Forward simulations are not complete for proving refinement: the reference implementation is not deterministic when projected over call and returns. 
+%
+%However, in this case, the set of observable actions includes the linearization points. And the reference implementations becomes deterministic for this larger set of observable actions.
+%
+%Determinism also simplifies the assertions
+%
+%When linearization points are not observable, reference implementations are not deterministic and one may need backward simulations.
+There is a paper arguing for this fact ~\cite{DBLP:conf/cav/SchellhornWD12}.
+%
+%However, using backward simulations is hard, they require prophecy variables. They are not good for automation.
+%
+%Question: We want to avoid backward simulations, so what can we do about implementations without fixed linearization points, e.g, Herlihy\&Wing, Time-Stamped Stack ?
+Can they be proved using forward simulations ?
+%
+%We answer positively for queues and stacks.
+The idea is to define new abstract implementations over a set of observable actions bigger than the set of call/returns, but smaller than all linearization points.
+%
+%Queues: only dequeues have fixed linearization points.
+%
+%Stacks: pops have fixed commit points
+%
+%\newpage%~\newpage

### 2017-01-25 draftCAV/app-absImplQueue.tex
Revisited Queue proofs in the appendix

@@
 	\item every pending enqueue has the same input value in both $s$ and $t$, i.e., for every $k$, $\ell_1(k)=in^0(k)$,
-	\item a pending enqueue from $O$ has been linearized whenever its value is contained in $\sigma$, i.e., for every $k$, $cp^0(k)=A$ iff $\ell_1(k)\in D$ and $\ell_2(k)={\tt PEND}$, 
+	\item a pending enqueue from $O$ has been linearized whenever its value is contained in $\sigma$, i.e., for every $k$, $cp^0(k)=A$ if $\ell_1(k)\in D$ and $\ell_2(k)={\tt PEND}$, 
 	\item a pending enqueue from $O$ hasn't been linearized whenever its value is not in $\sigma$, i.e., for every $k$, $cp^0(k)=A_1$ iff $\ell_1(k)\not\in D$ and $\ell_2(k)={\tt PEND}$, 
-	\item a pending enqueue which is not in $O$ has been linearized, i.e., for every $k$, $cp^0(k)=A$ iff $k\not\in O$ and $cp(k)=A_1$, 
+	\item a pending enqueue which is not in $O$ has been linearized, i.e., for every $k$, $cp^0(k)=A$ if $k\not\in O$ and $cp(k)=A_1$, 
 	\item an enqueue is completed in $s$ whenever it is completed in $t$, i.e., for every $k$, $cp(k)=A_2$ iff $cp^0(k)=A_2$,
@@
 
+For the conditions described above, if we fix the set $D$ and $\sigma_t$, then the state $t$ related to $s$ becomes unique.
+We use this fact in the proof.
+In some places, we only give $D$, $\sigma_t$ and $s$ without explicitly defining $t$ or show that there exists $t$ with the given $\sigma_t$ that is related to $s$ by just finding a $D$ such that $\sigma_t$ is a linearization of $(D, \prec)$ where $\prec$ is induced from $<_s$.
+
+or $\sigma_t$ and not describing $t$ explicitly.
+
 In the following, we show that indeed $bs$ is a normal $C\cup R\cup Lin(deq)$-backward simulation from $AbsQ$ to $AbsQ_0$.
+% mainly focusing on how to pick $\sigma$ fields of related states in $AbsQ_0$ and how the corresponding actions of $AbsQ_0$ modifies $\sigma$.
 
-TODO FILL IN THE STEPS
-
-We will provide a backward simulation relation btw $L_I$ and $L_A$.
+%TODO FILL IN THE STEPS
+%We will provide a backward simulation relation btw $L_I$ and $L_A$.
 Given a state $q = (so_q, f_q) \in Q_I$,  $q' \in bs[q]$ iff (i) $s_{q'}$ is a linearization of $so_q$ projected to $d$ fields (second field).
@@
 \begin{itemize}
-\item[$\langle i \rangle$] $bs[{q_0}_I] = \{ {q_0}_A \}$.
-\item[$\langle ii-a-enq \rangle$] Let $(q, inv(enq,d,k), q') \in \delta_I$ and $t' \in bs[q']$.
+\item[$\langle i \rangle$] $bs[s^{AbsQ}_0] = \{ s^{AbsQ_0}_0 \}$.
+
+\item[\textsc{call-enq}] Let $s \xrightarrow{inv(enq,d,k)}_{AbsQ} s'$ and $t' \in bs[s']$. Either $k \in D_{t'}$ or not.
+
+First consider the former case.
+We know that $\ell_{s'}(k) = (d, \texttt{PEND})$ and $k$ is maximal in $s'$.
+Hence $\sigma_{t'} = \rho \circ \langle d \rangle \circ \pi$ where $\pi$ contains linearization of pending elements in $O_{s'}$.
+Then, pick $\sigma_t = \rho$.
+We can find such a $t \in bs[s]$ with $\sigma_t$.
+Let $(D, \prec)$ be the partial order that is used while constructing $\sigma_{t'}$ from $O_s$ and $<_s$.
+We can find $(D', \prec')$ for relating $s$ to $t$ such that $D'$ does not contain the values of pending elements that formed $\pi$ suffix of $s_{t'}$ and $d$ coming from linearization of $k \in O_{s'}$.
+
+One can also see that $t \xrightarrow{\alpha}_{AbsQ_0} t'$ where $\alpha = inv(enq,d,k), lin(enq,d,k), \\lin(enq,d_1,k_1), ..., lin(enq,d_j,k_j)$ such that $\pi = d_1,...,d_j$ and $k_1,...,k_j \in O_{s'}$ are the pending elements that are  linearized to form $\pi$.
+Note that $\alpha$ obeys the definition of normal backward simulation definition.
+
+For the second case, pick $t$ such that $\sigma_t = \sigma_{t'}$.
+We can find a $t$ with $\sigma_t$ related to $s$ by $bs$ using the same $(D,\prec)$ partial order that is used while relating $s'$ to $t'$.
+$\ell_1(\texttt{COMP}(O_s)) \subseteq D$ holds because $\texttt{COMP}(O_s) = \texttt{COMP}(O_{s'})$.
+%\item[$\langle ii-a-enq \rangle$] Let $(q, inv(enq,d,k), q') \in \delta_I$ and $t' \in bs[q']$.
 We know that $(k,d,\texttt{PENDING}$ is a maximal element in $so_{q'}$ and either $s_{t'}$ does not linearize it or $s_{t'} = \rho \circ \langle d \rangle \circ \pi$ where $\pi$ consists of only linearization of \texttt{PENDING} elements.
@@
 The $f_t$ could be obtained easily by observing the sequence $a$ and it can be checked that such $t \in bs[q]$ exists.
-\item[$\langle ii-a-deq \rangle$] Let $(q, inv(deq,k), q') \in \delta_I$ and $t' \in bs[q']$.
+
+\item[\textsc{call-deq}] Let $s \xrightarrow{inv(deq,d,k)}_{AbsQ} s'$ and $t' \in bs[s']$.
+Pick $t$ such that it is equal to $t'$ in every field except that $k \notin dom(cp^0_t)$.
+Then, $t \in bs[s]$ and $t \xrightarrow{inv(deq,d,k)_{AbsQ_0}} t'$.
+%\item[$\langle ii-a-deq \rangle$] Let $(q, inv(deq,k), q') \in \delta_I$ and $t' \in bs[q']$.
 We know that $f_{t'}(k) = D_0$.
 We know that there is a $t \in bs[q]$ such that $s_t = s_{t'}$ (since $so_q = so_{q'}$), $f_t(k) = N$ and $(t,inv(deq,k),t') \in \delta_A$.
-\item[$\langle ii-d \rangle$] Let $(q,lin(deq,d,k),q') \in \delta_I$ and $t' \in bs[q']$.
+
+\item[\textsc{lin-deq1}]  Let $s \xrightarrow{lin(deq,d,k)}_{AbsQ} s'$, $t' \in bs[s']$ and $d \neq \texttt{EMPTY}$.
+We pick $t$ such that $\sigma_t = \langle d \rangle \circ \sigma_{t'}$.
+We first show that $t \in bs[s]$.
+Let $(D, \prec)$ be the partial order that is linearized to obtain $\sigma_{t'}$ and $k' \in O_s$ be the element such that ${\ell_s}_1(k') = d$.
+We know that $k'$ is minimal in $<_s$ due to the premise of the rule \textsc{lin-deq1}.
+Hence, we can obtain $(D', \prec')$ such that $D' = D \cup \{{\ell_s}_1(k')\}$ and $\sigma_t$ is a linearization of it.
+
+In addition, $t \xrightarrow{lin(deq,d,k)}_{AbsQ_0} t'$.
+The action $lin(deq,d,k)$ is enabled in state $t$ since $d$ is the minimum element of $\sigma_t$.
+Note that the transition relating $t$ to $t'$ obeys the definition of normal forward simulation.
+
+\item[\textsc{lin-deq2}]  Let $s \xrightarrow{lin(deq,\texttt{EMPTY},k)}_{AbsQ} s'$ and $t' \in bs[s']$.
+We pick $(D, \prec)$ for relating $s$ to $t$ such that $D = \emptyset$.
+Such a $D$ is a valid choice since all the elements $O_s$ are pending.
+Then, $\sigma_t = \langle \rangle$ is the only linearization of $(D, \prec)$.
+Hence, $lin(deq,\texttt{EMPTY},k)$ action is enabled in $AbsQ_0$ and $t \xrightarrow{lin(deq,\texttt{EMPTY},k)}_{AbsQ_0} t'$ holds.
+%\item[$\langle ii-d \rangle$] Let $(q,lin(deq,d,k),q') \in \delta_I$ and $t' \in bs[q']$.
 First consider the case $d \neq \texttt{EMPTY}$.
@@
 Hence $(t,lin(deq,d,k),t') \in \delta_A$.
-\item[$\langle ii-b-enq \rangle$] Let $(q,ret(enq,k),q') \in \delta_I$ and $t' \in bs[q']$.
+
+\item[\textsc{ret-enq1}]  Let $s \xrightarrow{ret(enq,k)}_{AbsQ} s'$, $\ell_s(k) = (d,\texttt{PEND})$ and $t' \in bs[s']$.
+Assume $(D, \prec)$ be the partial order of which linearization is $\sigma_{t'}$.
+Pick $D'=D$.
+Then, $\ell_1(\texttt{COMP}(O_s)) \subseteq D \subseteq \ell_1(O_s)$ holds since $\texttt{COMP}(O_s) = \texttt{COMP}(O_{s'}) \setminus \{k\}$ and $k \in \texttt{PEND}(O_s)$.
+Construct $t \in bs[s]$ such that $\sigma_t = \sigma_{t'}$ is obtained by linearizing the partial order $(D',\prec')$.
+Then, $t \xrightarrow{ret(enq,k)}_{AbsQ_0} t'$ holds and it is a valid action with respect to normal backward-simulation relation definition.
+
+\item[\textsc{ret-enq2}]  Let $s \xrightarrow{ret(enq,k)}_{AbsQ} s'$, $k \notin O_s$ and $t' \in bs[s']$.
+Since $O_s = O_{s'}$ and $\texttt{COMP}(O_s) = \texttt{COMP}(O_{s'})$, we can pick $D' = D$ where $(D, \prec)$ is the strict partial order such that $\sigma_{t'}$ is its linearization.
+Construct $t \in bs[s]$ such that $\sigma_t = \sigma_{t'}$ is obtained by linearizing the partial order $(D',\prec')$.
+Then, $t \xrightarrow{ret(enq,k)}_{AbsQ_0} t'$ holds and it is a valid action with respect to normal backward-simulation relation definition.
+%\item[$\langle ii-b-enq \rangle$] Let $(q,ret(enq,k),q') \in \delta_I$ and $t' \in bs[q']$.
 There are two possible cases: There exists a node $(k,d,\texttt{PENDING}) \in nd_q$ or for all nodes $n = (k,\_,\_)$, $n \notin nd_q$.
@@
 Hence, $(t,ret(enq,k),t') \in \delta_A$.
-\item[$\langle ii-b-deq \rangle$] Let $(q,ret(deq,d,k), q') \in \delta_I$ and $t' \in bs[q']$.
+
+\item[\textsc{ret-deq}]  Let $s \xrightarrow{ret(deq,d,k)}_{AbsQ} s'$ and $t' \in bs[s']$.
+Assume $(D, prec)$ is the partial order of which linearization is $\sigma_{t'}$.
+Construct $t \in bs[s]$ such that $\sigma_t = \sigma_{t'}$ and $(D, \prec
+)$ is the partial order $\sigma_t$ is obtained from.
+$\texttt{COMP}(O_s) \subseteq D \subseteq O_s$ since $\texttt{COMP}(O_s) = \texttt{COMP}(O_{s'})$ and $O_s = O_{s'}$.
+Then, $t \xrightarrow{ret(deq,d,k)}_{AbsQ_0} t'$ holds.
+We have $rv_s(k) = rv^0_t(k)$ since $t \in bs[s]$.
+Hence the $ret(deq,d,k)$ is enabled in $t$.
+Moreover, $ret(deq,d,k)$ is a valid transition with respect to the normal backward simulation relation definition.
+%\item[$\langle ii-b-deq \rangle$] Let $(q,ret(deq,d,k), q') \in \delta_I$ and $t' \in bs[q']$.
 Then, we know that $f_{t'}(k) = D_2$ and there exists a $t \in bs[q]$ such that $s_t = s_{t'}$ (since $so_q = so_{q'})$ and $f_t(k) = D_1$.
@@
 	\item pending enqueues are maximal in $<$, i.e., for every $k$ and $k'$, $k \not< k'$ if $\ell_2(k)={\tt PEND}$,
+	\item the return values fixed at dequeue linearization points are the same, i.e., for every $k$, $rv(k)=rv^0(k)$.
 \end{itemize}
@@
 
-TODO FILL IN THE STEPS
-
-Since $L_I$ is deterministic with respect to the alphabet $AQ\Sigma$, we should be able to find a forward simulation between $L_A$ and $L_I$ if our claim is correct.
+%TODO FILL IN THE STEPS
+%Since $L_I$ is deterministic with respect to the alphabet $AQ\Sigma$, we should be able to find a forward simulation between $L_A$ and $L_I$ if our claim is correct.
 We propose a forward simulation by adding some auxiliary varibles to the state of $L_A$.
@@
 Elements of $InvEnq_q$ are pairs of the form $(d,k)$ where  $d \in \mathcal{D}\backslash \{ \texttt{EMPTY} \}$ is a value and $k \in \mathbb{N}$ is an operation identifier.
-Then our forward simulation $fs \subseteq Q_A \times Q_I$ relates the state $q = (s_q, f_q) \in Q_A$ to a state $q' =(so_{q'}, f_{q'}) \in Q_I$ iff (i) $f_{q'}(k) = f_q(k)$ for all $f_q(k) \in \{N, D_0, D_1, D_2\}$, (ii) $f_q(k) = E_0$ or $f_q(k) = E_1$ implies $f_{q'}(k) = E_0$, (iii) $f_q(k) = E_2$ implies $f_{q'}(k) = E_1$, (iv) If $(d,k) \in InvEnq_q$ or there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ and $f_q(k) = E_1$, then $(k,d,\texttt{PENDING})$ is a node in $so_{q'}$, (v) if there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ and $f_q(k) = E_2$ then $(k,d,\texttt{CLOSED})$ is a node in $so_{q'}$, (vi) open nodes in $so_{q'}$ are maximal and (vii) there exists a linearization (total order) $lo_{q'}$ of $so_{q'}$ that may omit some open nodes of $so_{q'}$ such that if we project nodes of
-$lo_{q'}$ to the first two fields, this linear order is equal to the sequence $s_q$.
+Then our forward simulation $fs \subseteq Q_A \times Q_I$ relates the state $q = (s_q, f_q) \in Q_A$ to a state $q' =(so_{q'}, f_{q'}) \in Q_I$ iff (i) $f_{q'}(k) = f_q(k)$ for all $f_q(k) \in \{N, D_0, D_1, D_2\}$, (ii) $f_q(k) = E_0$ or $f_q(k) = E_1$ implies $f_{q'}(k) = E_0$, (iii) $f_q(k) = E_2$ implies $f_{q'}(k) = E_1$, (iv) If $(d,k) \in InvEnq_q$ or there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ and $f_q(k) = E_1$, then $(k,d,\texttt{PENDING})$ is a node in $so_{q'}$, (v) if there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ and $f_q(k) = E_2$ then $(k,d,\texttt{CLOSED})$ is a node in $so_{q'}$, (vi) open nodes in $so_{q'}$ are maximal and (vii) there exists a linearization (total order) $lo_{q'}$ of $so_{q'}$ that may omit some open nodes of $so_{q'}$ such that if we project nodes of $lo_{q'}$ to the first two fields, this linear order is equal to the sequence $s_q$.
+%Next, we will show that $fs$ is a forward simulation relation:
 
-Next, we will show that $fs$ is a forward simulation relation:
 \begin{itemize}
-\item[$\langle i \rangle$] $fs[(q_0)_A] = \{(q_0)_I\}$
-\item[$\langle ii-a-enq \rangle$] Let $(q,inv(enq,d,k),q') \in \delta_A$ and $t \in fs[q]$.
+\item[$\langle i \rangle$] $fs[s_0^{AbsQ_0}] = \{s_0^{AbsQ}\}$
+\item[\textsc{call-enq}] Let $t \xrightarrow{inv(enq,d,k)}_{AbsQ_0} t'$ and $s \in fs[t]$.
+Then, $inv(enq,d,k)$ is an enabled action in $AbsQ$ since premise of \textsc{call-enq} holds in $t$ and $s \in fs[t]$.
+Obtain $s'$ such that $s \xrightarrow{inv(enq,d,k)}_{AbsQ} s'$.
+Note that $s'$ is unique since $AbsQ$ is deterministic with respect to $C \cup R \cup Lin(deq)$.
+
+Next, we show that $s' \in fs[t']$.
+Let $(D, \prec)$ be the partial order used while relating $t$ to $s$.
+Same partial order can be used while relating $\sigma_{s'}$ to $t'$ since $\texttt{COMP}(O_s) = \texttt{COMP}(O_{s'}$, $O_s \subseteq O_{s'}$ and $<_s \subseteq <_{s'}$.
+The only change we have in control point fields after the actions is that $cp^0_{s'}(k)= A_1$ and $cp_{t'}(k)=A_1$ which satisfies the conditions on $fs$.
+Moreover $k$ is a maximal pending node in $t'$ as required by the $fs$ conditions.
+Consequently, $s' \in fs[t']$.
+%\item[$\langle ii-a-enq \rangle$] Let $(q,inv(enq,d,k),q') \in \delta_A$ and $t \in fs[q]$.
 We can obtain $t'$ such that $(t,inv(enq,d,k),t') \in \delta_I$.
@@
 We omit the node $(k,d)$ while linearizing $so_{t'}$ and obtain $s_{q'}$.
-\item[$\langle ii-a-deq \rangle$] Let $(q, inv(deq,k), q') \in \delta_A$ and $t \in fs[q]$.
+
+\item[\textsc{call-deq}] Let $t \xrightarrow{inv(deq,k)}_{AbsQ_0} t'$ and $s \in fs[t]$.
+Then, $inv(deq,k)$ is an enabled action in $AbsQ$ since premise of \textsc{call-deq} holds in $t$ and $s \in fs[t]$.
+Obtain $s'$ such that $s \xrightarrow{inv(deq,k)}_{AbsQ} s'$.
+Note that $s'$ is unique since $AbsQ$ is deterministic with respect to $C \cup R \cup Lin(deq)$.
+
+Next, we show that $s' \in fs[t']$.
+Since $\sigma_s =\sigma_{s'}$, $O_s = O_{s'}$ and $\texttt{COMP}(O_s) = \texttt{COMP}(O_{s'}$, we can pick same $(D,\prec)$  partial order in $s'$ and show that $\sigma_{t'}$ is a linearization of it.
+The only change in control points after the transitions is that $cp^0_{t'}(k) = cp_{s'}(k) = R_1$ which does not violate any condition in $fs$.
+Consequently, $s' \in fs[t']$.
+%\item[$\langle ii-a-deq \rangle$] Let $(q, inv(deq,k), q') \in \delta_A$ and $t \in fs[q]$.
 We obtain $t'$ such that $(t, inv(deq,k),t') \in \delta_I$.
@@
 We can again see that $t' \in fs[q']$.
-\item[$\langle ii-c \rangle$] Let $(q,lin(enq,d,k),q') \in \delta_A$ and $t \in fs[q]$.
+
+\item[\textsc{lin-enq}] Let $t \xrightarrow{lin(enq,d,k)}_{AbsQ_0} t'$ and $s \in fs[t]$.
+Then, pick $s' = s$ such that $s \xrightarrow{\epsilon}_{AbsQ} s'$.
+Note that $\epsilon$ is a valid transition with respect to the normal forward simulation relation definition.
+We show that $s \in fs[t']$.
+If $(D, \prec)$ is the partial order in $s$ of which one linearization is $\sigma_t$, we pick $D' = D \cup \{k\} \subseteq O_s$.
+$(D', \prec')$ can be linearized to $\sigma_{t'}$ since $k$ is a maximal pending node and can be linearized at the end.
+Moreover, the only change in control point $cp^0_{t'}(k) = A$ which does not violate the $fs$ conditions.
+%\item[$\langle ii-c \rangle$] Let $(q,lin(enq,d,k),q') \in \delta_A$ and $t \in fs[q]$.
 Pick $t' = t$.
@@
 While obtaining linearization of $so_{t'}$, we do not neglect the node $(k,d, \texttt{PENDING})$ this time, although we neglect it in the linearization of $so_t$.
-\item[$\langle ii-d \rangle$] Let $(q, lin(deq,d,k), q') \in \delta_A$ and $t \in fs[q]$.
+
+\item[\textsc{lin-deq1}] Let $t \xrightarrow{lin(deq,d,k)}_{AbsQ_0} t'$, $d \neq \texttt{EMPTY}$ and $s \in fs[t]$.
+Then, $lin(deq,d,k)$ is an enabled action in $AbsQ$.
+There must exist $d \in D \subseteq \ell_1(O_s)$ such that ${\ell_s}_1(k') = d$ and $k'$ is minimal in $D$ (since $d$ is linearized as the minimum element in $\sigma_t$ according to premise of \textsc{lin-deq1} of $AbsQ$).
+Obtain $s'$ such that $s \xrightarrow{lin(deq,d,k)}_{AbsQ} s'$.
+Note that $s'$ is unique since $AbsQ$ is deterministic with respect to $C \cup R \cup Lin(deq)$.
+
+Next, we show that $s' \in fs[t']$.
+Let $(D, \prec)$ be the partial order used while relating $t$ to $s$ such that $\sigma_t$ is a linearization of this partial order.
+Since we have shown that $k'$ is minimal in that partial order, $\sigma_{t'}$ is a linearization of $(D', \prec')$ where $D' = D \setminus \{\ell_1(k')\}$.
+Note that $\ell_1(\texttt{COMP}(O_{s'})) \subseteq D' \subseteq \ell_1(O_{s'})$ holds.
+The only change in control points is that $cp^0_{t'}(k) = cp_{s'}(k) = R_2$ which does not violate the conditions for relating $t'$ to $s'$.
+Note that the fifth condition of $fs$ still holds for $k'$ while relating $t'$ to $s'$.
+After transitions $rv^0_{t'}(k) = rv_{s'}(k) = d$ and the last condition on $fs$ is preserved.
+
+\item[\textsc{lin-deq2}] Let $t \xrightarrow{lin(deq,\texttt{EMPTY},k)}_{AbsQ_0} t'$ and $s \in fs[t]$.
+Then, $lin(deq,d,k)$ is an enabled action in $AbsQ$.
+If $\texttt{COMP}(O_t) \neq \emptyset$, then $D$ use for linearization cannot be $\emptyset$ $\sigma_t = \langle\rangle$ cannot be a linearization of $(D, \prec)$.
+Obtain $s'$ such that $s \xrightarrow{lin(deq,\texttt{EMPTY},k)}_{AbsQ} s'$.
+Note that $s'$ is unique since $AbsQ$ is deterministic with respect to $C \cup R \cup Lin(deq)$.
+
+Next, we show that $s' \in fs[t']$.
+Let $(D, \prec)$ be the partial order used while relating $t$ to $s$ such that $\sigma_t = \langle \rangle $ is a linearization of this partial order.
+We can use the same $(D, \prec)$ for relating $t'$ to $s'$ because $\sigma$ field is the same for both $s$, $s'$; and $O$, $<$, $\ell$ fields are same for both $t$ and $t'$.
+The only change in control points is that $cp^0_{t'}(k) = cp_{s'}(k) = R_2$ which does not violate the conditions for relating $t'$ to $s'$.
+After transitions $rv^0_{t'}(k) = rv_{s'}(k) = d$ and the last condition on $fs$ is preserved.
+%\item[$\langle ii-d \rangle$] Let $(q, lin(deq,d,k), q') \in \delta_A$ and $t \in fs[q]$.
 Obtain $t'$ such that $(t, lin(deq,d,k), t') \in \delta_I$.
@@
 We can see that $t' \in fs[q']$ by checking the seven conditions.
-\item[$\langle ii-b-enq \rangle$] Let $(q, ret(enq,k), q') \in \delta_A$ and $t \in fs[q]$.
+
+\item[\textsc{ret-enq}] Let $t \xrightarrow{ret(enq,k)}_{AbsQ_0} t'$ and $s \in fs[t]$.
+Then, there are two cases assuming data independence: (i) $in^0_t(k)=d$ and $\exists i.
+\sigma_t(i)= d$ (ii) or not.
+
+First, consider the former case.
+Then, \textsc{ret-enq1} rule of $AbSQ$ is applicable.
+Its precondition holds since fifth condition of $fs$ holds while relating $t$ to $s$.
+Apply this rule ($ret(enq,k)$) to obtain $s'$.
+Note that $s'$ is unique since $AbsQ$ is deterministic with respect to $C \cup R \cup Lin(deq)$ and it is a valid action according to the normal forward-simulation relation definition.
+
+Next, we show that $s' \in fs[t']$.
+Since $\exists i.
+\sigma_t(i) = d$, we know that $d \in D$ where $(D, \prec)$ is the partial order satisfying first condition of $fs$ while relating $t$ to $s$, and $k \in O_s$ takes part in the linearization i.e., ${\ell_t}_1(k) \in D$.
+We can use the same partial order $(D,\prec)$ for relating $t'$ to $s'$ such that it satisfies the first condition of $fs$.
+The only change in control points is that $cp^0_{t'}(k) = cp_{s'}(k) = A_2$ which does not violate the conditions for relating $t'$ to $s'$.
+Note that the sixth condition of $fs$ also continue to hold for $k$ for the post-states.
+
+Second, consider the latter case:  $in^0_t(k)=d$, but $\forall i.
+\sigma_t(i)\neq d$.
+Since $(t,s) \in fs$, $k \notin O_s$ by the fifth and sixth conditions.
+Hence, the pre-condition of \textsc{ret-enq2} is satisfied by $t$.
+Apply this rule ($ret(enq,k)$) to obtain $s'$.
+Note that $s'$ is unique since $AbsQ$ is deterministic with respect to $C \cup R \cup Lin(deq)$ and it is a valid action according to the normal forward-simulation relation definition.
+
+Next, we show that $s' \in fs[t']$.
+For satisfying the first condition, one can use the same $(D, \prec)$ partial order that is used for relating pre-states since $\sigma$ fields of $t$, $t'$ and $O$, $<$, $\ell$ fields of $s$ and $s'$ are the same.
+The only change in control points is that $cp^0_{t'}(k) = cp_{s'}(k) = A_2$ which does not violate the conditions for relating $t'$ to $s'$.
+%\item[$\langle ii-b-enq \rangle$] Let $(q, ret(enq,k), q') \in \delta_A$ and $t \in fs[q]$.
 There are two cases: Either there exists an index $i \in \mathbb{N}$ such that $s_q(i) = (k,d)$ or there is no such $i$.
@@
 Again, $so_t = so_{t'}$ and $s_q = s_{q'}$ holds and $t' \in fs[q']$ can be observed.
-\item[$\langle ii-b-deq \rangle$] Let $(q, ret(deq,d,k), q') \in \delta_A$ and $t \in fs[q]$.
+
+\item[\textsc{ret-deq}] Let $t \xrightarrow{ret(deq,d,k)}_{AbsQ_0} t'$ and $s \in fs[t]$.
+Then, $ret(deq,d,k)$ is an enabled action in $AbsQ$ due to premise \textsc{ret-deq} of $AbsQ_0$ and the last condition on $fs$ (since $(t,s) \in fs$).
+Obtain $s'$ such that $s \xrightarrow{ret(deq,d,k)}_{AbsQ} s'$.
+Note that $s'$ is unique since $AbsQ$ is deterministic with respect to $C \cup R \cup Lin(deq)$.
+
+We see that $s' \in fs[t']$.
+Pre-states are equal to the post-states with the only exception in the control points such that $cp^0_{t'}(k) = cp_{s'}(k) = R_3$.
+All the conditions except the third one continues to hold in the post states since they hold in the pre-states.
+The third rule regarding the control points of dequeues also continue to the hold since changes in the control point of $k$ does not violate it.
+%\item[$\langle ii-b-deq \rangle$] Let $(q, ret(deq,d,k), q') \in \delta_A$ and $t \in fs[q]$.
 We pick $t'$ such that $(t, ret(deq,d,k), t') \in \delta_I$.

### 2017-01-25 draftCAV/draft.tex
structure stuff
more on queues
simulation for HW
title
a pass until stacks
small things
intro

@@
 \usepackage{alltt}
+\usepackage{listings}
+\usepackage[lined]{algorithm2e}
+\usepackage{wrapfig}
+%\newcommand{\xRightarrow}[2][]{\ext@arrow 0359\Rightarrowfill@{#1}{#2}}
 
-%\newcommand{\xRightarrow}[2][]{\ext@arrow 0359\Rightarrowfill@{#1}{#2}}
+\newcommand{\atCP}{\makeatletter @\makeatother}
+
+\usepackage{pslatex}
 
@@
 \input macros
+\input macros-programs
+
+\lstset{
+  language={Java},
+  basicstyle=\ttfamily\scriptsize
+}
 
@@
 
-\title{Linearizability Proofs Based On Forward Simulations}
+\title{Proving linearizability using forward simulations}
 \author{ }
@@
 
+%\raggedbottom
+
 \maketitle
 
+\vspace{-10mm}
+\begin{abstract}
+Efficient concurrent implementations of abstract data structures such as stacks and queues are especially susceptible to programming errors, and necessitate automatic verification.
+Linearizability is the correctness criterion in this context. It allows to establish observational refinement between a concurrent implementation and an atomic reference implementation.
+Proving linearizability, or equivalently observational refinement, requires identifying linearization points for each method invocation along all possible computations, leading to valid sequential executions, or alternatively, establishing forward and backward simulations.
+In both cases, carrying out proofs is hard and complex in general.
+In particular, backward reasoning is difficult in the context of programs with data structures, and strategies for identifying statically linearization points cannot be defined for all existing implementations.
+In this paper, we show that, contrary to common belief, many such complex implementations, including, e.g., the Herlihy\&Wing queue and the Time-Stamped Stack, can be proved correct using only forward simulation arguments.
+This leads to simple and natural correctness proofs for these implementations that are amenable to automation.
+
+%Efficient implementations of atomic objects such as concurrent stacks and queues are especially susceptible to programming errors, and necessitate automatic verification.
+Unfortunately their correctness criteria -- conformance to atomic reference implementations -- are hard to verify.
+Conformance can be formalized as a trace inclusion problem, also known as observational refinement, which in general, can be proved using compositions of forward and backward simulations.
+Contradicting common belief, we show that many complex implementations, e.g., the Herlihy and Wing queue and the Time-Stamped Stack, can be proved correct using only forward simulation arguments.
+We define classes of queue and stack implementations for which forward simulations are sound and complete for proving their correctness.
+\vspace{-2mm}
+\end{abstract}
+
+\input{intro}
 \input{prelim}
@@
 
+\input{related}
+
 \bibliographystyle{abbrvnat}
@@
 
+\input{app-prelim}
 \input{app-backSim}
@@
 \input{app-absImplStack}
+\input{app-tss}
 

### 2017-01-25 draftCAV/intro.tex
intro

@@
 \section{Introduction}
-
+\vspace{-1mm}
 Efficient concurrent implementations of atomic collections including stacks and queues are vital to modern computer systems.
@@
 However, concurrency risks unintended inter-operation interference, and risks conformance to atomic reference implementations.
-Conformance is formally captured by \emph{observational refinement}, which assures that all behaviors of programs
+Conformance is formally captured by \emph{(observational) refinement}, which assures that all behaviors of programs
 using these efficient implementations would also be possible were the atomic reference implementations used instead.
 
-Observational refinement can be formalized as a trace inclusion problem, and the latter can itself be reduced to an assertion (invariant) checking problem, but this requires in general the consideration of additional history and prophecy variables.
-Alternatively, verifying observational refinement requires in general establishing both a forward simulation {\em and} a backward simulation.
+Observational refinement can be formalized as a trace inclusion problem, and the latter can itself be reduced to an assertion (invariant) checking problem, but this requires in general the consideration of additional history and prophecy variables~\cite{DBLP:journals/tcs/AbadiL91}.
+Alternatively, verifying refinement requires in general establishing both a forward simulation {\em and} a backward simulation~\cite{DBLP:journals/iandc/LynchV95}.
 While simulations are natural concepts, reasoning about backward simulations, corresponding to the use of prophecy variables, is however less intuitive, and from the practical point of view, backward reasoning is in general hard and complex for programs manipulating data structures.
-Therefore, a crucial issue is to understand the limits of forward reasoning in establishing observational refinement.
-More precisely, we consider that an important question is to determine for which concurrent abstract data structures, and for which types of their implementation, is it possible to carry out an observational refinement proof using only forward reasoning.
+Therefore, a crucial issue is to understand the limits of forward reasoning in establishing refinement.
+More precisely, an important question is to determine for which concurrent abstract data structures, and for which classes of implementations, it is possible to carry out a refinement proof using only forward simulations.
 
-In order to get rid of backward simulations (and equivalently of prophecy variables), it is necessary to have reference implementations of the considered abstract data structures that are {\em deterministic}.
+In order to get rid of backward simulations (and equivalently of prophecy variables) while preserving completeness w.r.t.
+refinement, it is necessary to have reference implementations of the considered abstract data structures that are {\em deterministic}.
 Interestingly, determinism allows also to simplify the forward simulation checking problem.
 Indeed, in this case, this problem can be reduced to an invariant checking problem.
-Basically, the simulation relation can be seen as an invariant of the system composed by the two compared programs.
-Therefore, existing methods and tools for invariant checking can be leveraged in the context of proving observational refinement.
+Basically, the simulation relation can be seen as an invariant of the system composed of the two compared programs.
+Therefore, existing methods and tools for invariant checking can be leveraged in this context.
+% of proving observational refinement.
 
 But, in order to determine precisely what is meant by determinism, an important point is to fix the alphabet of observable events along computations. 
-Typically, to reason about observational refinement between two library implementations, the only observable events are the calls and returns corresponding to the method invocations along computations.
+Typically, to reason about refinement between two library implementations, the only observable events are the calls and returns corresponding to the method invocations along computations.
 This means that only the external interface of the library is considered to compare behaviors, and nothing else from the implementations is exposed.
@@
 
-One approach addressing this question is based on the notion of linearizability \cite{journals/toplas/HerlihyW90} and its correspondence with observational refinement (w.r.t.
+One approach addressing this question is based on the notion of linearizability \cite{journals/toplas/HerlihyW90} and its correspondence with refinement (w.r.t.
 atomic reference implementations) \cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}.
@@
 By extending this way the alphabet of observable events, it becomes straightforward to define {\em deterministic} atomic reference implementations.
-Therefore, proving linearizability can be carried out as an invariant checking problem when linearization pointst are fixed, e.g.,~\cite{conf/ppopp/VafeiadisHHS06,conf/cav/AmitRRSY07,conf/vmcai/Vafeiadis09,conf/tacas/AbdullaHHJR13}.
+Therefore, proving linearizability can be carried out using forward simulations when linearization points are fixed, e.g.,~\cite{conf/ppopp/VafeiadisHHS06,conf/cav/AmitRRSY07,conf/vmcai/Vafeiadis09,conf/tacas/AbdullaHHJR13}.
 %
@@
 The proofs of linearizability of these implementations are highly nontrivial, very involved, and hard to read, understand and automatize.
-Therefore, the crucial question we address is what is precisely the kind of information that is necessary to expose in order to obtain deterministic atomic reference implementations for such data structures, allowing to derive simple and natural linearizability proofs for such complex implementations, which can be carried out as systematic invariant checking proofs that are amenable to automation?
+Therefore, the crucial question we address is what is precisely the kind of information that is necessary to expose in order to obtain deterministic atomic reference implementations for such data structures, allowing to derive simple and natural linearizability proofs for such complex implementations, based on forward simulations, that are amenable to automation ?
 
-We observe that the main difficulty in reasoning about these implementations is that, linearization points of enqueue/push operations occurring along some given computation, depend in general on the linearization points of dequeue/pop operations that are occurring later in that computation, at arbitrary far positions.
+We observe that the main difficulty in reasoning about these implementations is that, linearization points of enqueue/push operations occurring along some given computation, depend in general on the linearization points of dequeue/pop operations that occur arbitrarily far in the future.
 Therefore, since linearization points for enqueue/push operations cannot be determined in advance, the information that could be fixed and exposed can concern only the dequeue/pop operations.
@@
 
-However, in the case of the Time-Stamped Stack, fixing linearization points of pop operations is actually too restrictive.
+However, in the case of Time-Stamped Stack, fixing linearization points of pop operations is actually too restrictive.
 Nevertheless, we show that our approach can be generalized to handle this case.
@@
 We prove that by exposing commit points (instead of linearization points) for pop methods, we can still provide deterministic reference implementations.
-We show that using this approach leads to a quite simple proof of the Time-Stamped Stack.
+We show that using this approach leads to a quite simple proof of the Time-Stamped Stack, based on forward simulations.
 

### 2017-01-25 draftCAV/app-tss.tex
Flow diagrams are added to the appendix

@@
 %\end{lstlisting}
+\begin{figure}[t]
+\centering
+\includegraphics[width=10cm]{tssPushFlow.pdf}
+\includegraphics[width=12cm]{tssPopFlow.pdf}
+\vspace{-8mm}
+\caption{The flow diagram for the pop and push methods of the Time-Stamped Stack algorithm. The blue points show the control points roughly and the arrows show the possible transitions.}
+\label{fig:tssFlow}
+\end{figure}
+
 The LTS corresponding to the description of $TSS$ given in Fig.~\ref{fig:TimeStamped} is defined as usual.
-The control points and transition labels we use in the following proof are pictured in Fig.~\ref{} \textcolor{red}{Flow diagram is referenced here}.
+The control points and transition labels we use in the following proof are pictured in Fig.~\ref{fig:tssFlow}.
 To simplify the proof, we take the initializations of some local variables together as atomic.
@@
 Its value is unique to each pending operation in a state i.e., concurrent operations cannot have the same $myTID$ value.
-TS-Stack states also contains sets $O_a, O_r \in \mathbb{O}$ which are operation identifier sets of push and pops respectively, and the control point function $cp$ which is a map from operation identifiers to the control points set that are presented in the flow diagram Figure X  \textcolor{red}{Flow diagram is referenced here}.
+TS-Stack states also contains sets $O_a, O_r \in \mathbb{O}$ which are operation identifier sets of push and pops respectively, and the control point function $cp$ which is a map from operation identifiers to the control points set that are presented in the flow diagram Figure ~\ref{fig:tssFlow}.
 Transition relation of the TS-STack is presented in Figure~\ref{fig:transitions:TSSPush} (push rules) and Figure~\ref{fig:transitions:TSSPop} (pop rules).
@@
 \end{proof}
-\textcolor{red}{TODO: Check ReverseFrontiers for the cases before compop.}
-\textcolor{red}{TODO: Check TraverseBefore for the cases before pop2.}
-\textcolor{red}{TODO: CheckTraverseBeforeNull for the cases befoere pop 5.}
-\textcolor{red}{TODO: Apply Dr. Enea's comments.}
-\textcolor{red}{TODO: Try case split on fs conditions instantiations check.}
-\textcolor{red}{TODO: Add Flow Diagrams to the paper.} 
-\textcolor{red}{TODO: }
+%\textcolor{red}{TODO: Check ReverseFrontiers for the cases before compop.}
+%\textcolor{red}{TODO: Check TraverseBefore for the cases before pop2.}
+%\textcolor{red}{TODO: CheckTraverseBeforeNull for the cases befoere pop 5.}
+%\textcolor{red}{TODO: Apply Dr. Enea's comments.}
+%\textcolor{red}{TODO: Try case split on fs conditions instantiations check.}
+%\textcolor{red}{TODO: Add Flow Diagrams to the paper.} 
+%\textcolor{red}{TODO: }

### 2017-01-25 draftCAV/draft.tex
fixed some spaces

@@
 
-\vspace{-10mm}
+\vspace{-9mm}
 \begin{abstract}
-Efficient concurrent implementations of abstract data structures such as stacks and queues are especially susceptible to programming errors, and necessitate automatic verification.
-Linearizability is the correctness criterion in this context. It allows to establish observational refinement between a concurrent implementation and an atomic reference implementation.
-Proving linearizability, or equivalently observational refinement, requires identifying linearization points for each method invocation along all possible computations, leading to valid sequential executions, or alternatively, establishing forward and backward simulations.
+%Efficient concurrent implementations of abstract data structures such as stacks and queues are especially susceptible to programming errors, and necessitate automatic verification.
+Linearizability is the standard correctness criterion concurrent data structures such as stacks and queues. %
+%in this context. 
+It allows to establish observational refinement between a concurrent implementation and an atomic reference implementation.
+%or equivalently observational refinement, 
+Proving linearizability requires identifying linearization points for each method invocation along all possible computations, leading to valid sequential executions, or alternatively, establishing forward \emph{and} backward simulations.
 In both cases, carrying out proofs is hard and complex in general.
 In particular, backward reasoning is difficult in the context of programs with data structures, and strategies for identifying statically linearization points cannot be defined for all existing implementations.
-In this paper, we show that, contrary to common belief, many such complex implementations, including, e.g., the Herlihy\&Wing queue and the Time-Stamped Stack, can be proved correct using only forward simulation arguments.
+In this paper, we show that, contrary to common belief, many such complex implementations, including, e.g., the Herlihy\&Wing Queue and the Time-Stamped Stack, can be proved correct using only forward simulation arguments.
 This leads to simple and natural correctness proofs for these implementations that are amenable to automation.
@@
 \end{abstract}
+
+%Linearizability is the standard correctness criterion concurrent data structures such as stacks and queues.
+It allows to establish observational refinement between a concurrent implementation and an atomic reference implementation.Proving linearizability requires identifying linearization points for each method invocation along all possible computations, leading to valid sequential executions, or alternatively, establishing forward and backward simulations.
+In both cases, carrying out proofs is hard and complex in general.
+In particular, backward reasoning is difficult in the context of programs with data structures, and strategies for identifying statically linearization points cannot be defined for all existing implementations.
+In this paper, we show that, contrary to common belief, many such complex implementations, including, e.g., the Herlihy&Wing Queue and the Time-Stamped Stack, can be proved correct using only forward simulation arguments.
+This leads to simple and natural correctness proofs for these implementations that are amenable to automation.
 

### 2017-01-25 draftCAV/intro.tex
fixed some spaces

@@
 \vspace{-1mm}
-Efficient concurrent implementations of atomic collections including stacks and queues are vital to modern computer systems.
-Programming them is however error prone.
+Programming efficient concurrent implementations of atomic collections, e.g., stacks and queues, is error prone.
 To minimize synchronization overhead between concurrent method invocations, implementors avoid blocking operations like lock acquisition, allowing methods to execute concurrently.
@@
 
-Observational refinement can be formalized as a trace inclusion problem, and the latter can itself be reduced to an assertion (invariant) checking problem, but this requires in general the consideration of additional history and prophecy variables~\cite{DBLP:journals/tcs/AbadiL91}.
-Alternatively, verifying refinement requires in general establishing both a forward simulation {\em and} a backward simulation~\cite{DBLP:journals/iandc/LynchV95}.
-While simulations are natural concepts, reasoning about backward simulations, corresponding to the use of prophecy variables, is however less intuitive, and from the practical point of view, backward reasoning is in general hard and complex for programs manipulating data structures.
+Observational refinement can be formalized as a trace inclusion problem, and the latter can itself be reduced to an invariant checking problem, but this requires in general introducing history and prophecy variables~\cite{DBLP:journals/tcs/AbadiL91}.
+Alternatively, verifying refinement requires in general establishing a forward simulation {\em and} a backward simulation~\cite{DBLP:journals/iandc/LynchV95}.
+While simulations are natural concepts, backward reasoning, corresponding to the use of prophecy variables, is in general hard and complex for programs manipulating data structures.
+%reasoning about backward simulations, corresponding to the use of prophecy variables, is however less intuitive, and from the practical point of view, backward reasoning is in general hard and complex for programs manipulating data structures.
 Therefore, a crucial issue is to understand the limits of forward reasoning in establishing refinement.
@@
 
-In order to get rid of backward simulations (and equivalently of prophecy variables) while preserving completeness w.r.t.
-refinement, it is necessary to have reference implementations of the considered abstract data structures that are {\em deterministic}.
+To get rid of backward simulations (or prophecy variables) while preserving completeness w.r.t. refinement, it is necessary to have reference implementations 
+%of the considered abstract data structures 
+that are {\em deterministic}.
 Interestingly, determinism allows also to simplify the forward simulation checking problem.
@@
 
-One approach addressing this question is based on the notion of linearizability \cite{journals/toplas/HerlihyW90} and its correspondence with refinement (w.r.t.
-atomic reference implementations) \cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}.
-Linearizability of a computation (of some concurrent implementation) means that each of the method invocations can be seen as happening at some point, called {\em linearization point}, occurring somewhere between the call and return events of that invocation, such that the obtained sequence of linearization points along the computation defines a sequence of operations that is possible in the atomic reference implementation.
+One approach addressing this question is based on linearizability \cite{journals/toplas/HerlihyW90} and its correspondence with refinement % (w.r.t. atomic reference implementations)
+\cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}.
+Linearizability of a computation (of some implementation) means that each of the method invocations can be seen as happening at some point, called {\em linearization point}, occurring somewhere between the call and return events of that invocation.
+The obtained sequence of linearization points along the computation should define a sequence of operations that is possible in the atomic reference implementation.
 Proving the existence of such sequences of linearization points, for all the computations of a concurrent library, is a complex problem \cite{journals/iandc/AlurMP00,conf/esop/BouajjaniEEH13,DBLP:conf/netys/Hamza15}.
 %
-However, proving linearizability becomes less complex when linearization points are fixed for each method, i.e., associated (once and for all) with the execution of a designated statement in its source code \cite{conf/esop/BouajjaniEEH13}.
-In this case, we can consider that libraries expose not only the call and returns events, but also events signaling linearization points.
+However, proving linearizability becomes less complex when linearization points are fixed for each method, i.e., associated with the execution of a designated statement in its source code \cite{conf/esop/BouajjaniEEH13}.
+In this case, we can consider that libraries expose in addition to calls and returns, events signaling linearization points.
 By extending this way the alphabet of observable events, it becomes straightforward to define {\em deterministic} atomic reference implementations.
@@
 %
-While this approach can be useful in some cases, it is unfortunately not applicable to many efficient implementations such as the LCRQ queue~\cite{DBLP:conf/ppopp/MorrisonA13} (based on the principle of the Herlihy\&Wing queue \cite{journals/toplas/HerlihyW90}), and the Time-Stamped Stack \cite{DBLP:conf/popl/DoddsHK15}.
+Unfortunately, this approach is not applicable to efficient implementations such as the LCRQ queue~\cite{DBLP:conf/ppopp/MorrisonA13} (based on the principle of the Herlihy\&Wing queue \cite{journals/toplas/HerlihyW90}), and the Time-Stamped Stack \cite{DBLP:conf/popl/DoddsHK15}.
 The proofs of linearizability of these implementations are highly nontrivial, very involved, and hard to read, understand and automatize.

### 2017-01-25 draftCAV/queue.tex
fixed some spaces

--- 2017-01-24 draftCAV/queue.tex
@@
 % describe a queue implementation listed in Figure~\ref{fig:HerlihyWing} and known as the Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), where only the dequeue methods have fixed linearization points.
+We describe $\mathit{HWQ}$ where the linearization points of the enqueue methods are not fixed.
+The shared state consists of an array {\tt items} storing the values in the queue and a counter {\tt back} storing the index of the first unused position in {\tt items}.
+Initially, all the positions in the array are {\tt null} and {\tt back} is 0.
+An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueues) and then, stores the input value {\tt x} at this position.
+The dequeue method traverses the array {\tt items} starting from the beginning and atomically swaps {\tt null} with the encountered value.
+If the value is not {\tt null}, then the dequeue returns that value.
+If it reaches the end of the array, then it restarts.
+
 \begin{wrapfigure}{l}{5.3cm}
-\vspace{-10mm}
+\vspace{-9mm}
 \begin{lstlisting}
@@
   \end{lstlisting}
-\vspace{-5mm}
+\vspace{-5.5mm}
 \caption{Herlihy \& Wing Queue. We assume that every statement is atomic.}
 \label{fig:HerlihyWing}
-\vspace{-7mm}
+\vspace{-6mm}
 \end{wrapfigure}
-We describe $\mathit{HWQ}$ which is a queue implementation where the linearization points of the enqueue methods are not fixed.
-The shared state consists of an array {\tt items} storing the values in the queue and a counter {\tt back} storing the index of the first unused position in {\tt items}.
-Initially, all the positions in the array are {\tt null} and {\tt back} is 0.
-An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueues) and then, stores the input value {\tt x} at this position.
-The dequeue method traverses the array {\tt items} starting from the beginning and atomically swaps {\tt null} with the encountered value.
-If the value is not {\tt null}, then the dequeue returns that value.
-If it reaches the end of the array, then it restarts.
-
-The linearization points of the enqueue methods are not fixed, they depend on dequeue operations executing in the future.
+The linearization points of the enqueues are not fixed, they depend on dequeues executing in the future.
 Consider the following trace with two concurrent enqueues (${\tt i}(k)$ represents the value of {\tt i} in operation $k$): $inv(enq,x,1)$, $inv(enq,y,2)$, ${\tt i}(1) = \mbox{{\tt bck++}}$, ${\tt i}(2) = \mbox{{\tt bck++}}$, ${\tt items[i(}2{\tt )]} = y$.
@@
 %\end{align*}
-Assuming that the linearization point corresponds to the assignment of {\tt i}, the history of this trace should be linearized to $inv(enq,x,1)\ ret(enq,1)\ inv(enq,y,2)\ ret(enq,2)$.
-However, a dequeue operation executing until completion after this trace will return value $y$ (only position $1$ is filled in the array {\tt items}) which is not consistent with this linearization.
-On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i(}1{\tt )]} = x$ and a completed dequeue that in this case returns $x$, leads to the following incorrect linearization: $inv(enq,y,2)$, $ret(enq,2)$, $inv(enq,x,1)$, $ret(enq,1)$, $inv(deq,3)$, $ret(deq,x,3)$.
+Assuming that the linearization point corresponds to the assignment of {\tt i}, the history of this trace should be linearized to $inv(enq,x,1)$, $ret(enq,1)$, $inv(enq,y,2)$, $ret(enq,2)$.
+However, a dequeue executing until completion after this trace will return $y$ (only position $1$ is filled in the array {\tt items}) which is not consistent with this linearization.
+On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i(}1{\tt )]} = x$ and a completed dequeue that in this case returns $x$, leads to the incorrect linearization: $inv(enq,y,2)$, $ret(enq,2)$, $inv(enq,x,1)$, $ret(enq,1)$, $inv(deq,3)$, $ret(deq,x,3)$.
 %\vspace{-1.5mm}
@@
 \vspace{-1.5mm}
-%\textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too.
-We should be able to return EMPTY if there are only pending nodes.
-Consider the following history of $AbsQ_0$: $inv(enq, d_1,k_1), inv(enq, d_2,k_2), inv(deq, k_3), lin(deq, \texttt{EMPTY}, k_3)$.
-This history should be reflected in $AbsQ$ by enabling lin \texttt{EMPTY} of dequeue when there are pending nodes.
-We also need to update the rules in figure.}
 \begin{wrapfigure}{l}{6.7cm}
-\vspace{-7mm}
-\includegraphics[width=7cm]{fig-queue12.pdf}
+\vspace{-6mm}
+\includegraphics[width=6.8cm]{fig-queue12.pdf}
 %
@@
 \end{wrapfigure}
+%\textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too.
+We should be able to return EMPTY if there are only pending nodes.
+Consider the following history of $AbsQ_0$: $inv(enq, d_1,k_1), inv(enq, d_2,k_2), inv(deq, k_3), lin(deq, \texttt{EMPTY}, k_3)$.
+This history should be reflected in $AbsQ$ by enabling lin \texttt{EMPTY} of dequeue when there are pending nodes.
+We also need to update the rules in figure.}
 Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation.
@@
 The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
-When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueues (the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
+When
+%\noindent
+the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueues (the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
 The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.

### 2017-01-25 draftCAV/app-tss.tex
fixed some spaces
TSS appendix revised

@@
 \centering
-\includegraphics[width=10cm]{tssPushFlow.pdf}
-\includegraphics[width=12cm]{tssPopFlow.pdf}
+\includegraphics[width=7cm]{tssPushFlow.pdf}
+\includegraphics[width=16cm]{tssPopFlow.pdf}
 \vspace{-8mm}
@@
 \end{itemize}
-Next, we will show that $\mathit{fs}_2$ is really a Com(pop)-forward simulation relation.
+Next, we will show that $\mathit{fs}_2$ is really a $C\cup R\cup Com(pop)$-forward simulation relation.
 Except the trivial base case, we case-split on the transition rules.
-We first assume $(s, \alpha s') \in \delta_C$ and $t \in \mathit{fs}_2[s]$.
-Then, we find  corresponding transition $\alpha' \in \Sigma_{AbsS}$ obeying the Com(pop)-forward simulation relation conditions and obtain $t'$ such that $(t, \alpha' t') \in \delta_{AbsS}$  and $t' \in \mathit{fs}_2[s']$.
+We first assume $s \xrightarrow{\alpha}_{TSS} s' $ and $t \in \mathit{fs}_2[s]$.
+Then, we find  corresponding transition $\alpha' \in \Sigma_{AbsS}$ obeying the $C\cup R\cup Com(pop)$-forward simulation relation conditions and obtain $t'$ such that $t \xrightarrow{\alpha'}_{AbsS} st $  and $t' \in \mathit{fs}_2[s']$.
 
@@
 
-Let the following derivation rule of $TSS$ be the one for describing $\alpha$:
-\begin{mathpar}
-    \inferrule{
-      \psi
-    }{
-      s
-      \xrightarrow{\alpha} 
-      s'
-    }
-\end{mathpar}
-and the following one be the derivation rule of $AbsS$ describing $\alpha'$ if $\alpha' \neq \epsilon$ (equivalently $\alpha' = \alpha$):
-\begin{mathpar}
-    \inferrule{
-      \psi'
-    }{
-      q
-      \xrightarrow{\alpha'} 
-      q'
-    }
-\end{mathpar}
+Let the following describe $\alpha$: $\psi \triangleright s\xrightarrow{\alpha}_{TSS} s'$ where $\psi$ is the precondition (guard) that needs to be satisfied for enabling $\alpha$ and $\psi' \triangleright t \xrightarrow{\alpha'}_{AbsS} t'$ describe the $\alpha'$ if $\alpha' \neq \epsilon$ (equivalently $\alpha' = \alpha$).
+
 For the cases $\alpha' = \alpha$, we first need to show $\alpha'$ is enabled in state $t$ i.e., $t$ satisfies $\psi'$.
-If this can not be directly obtained from $s$ satisfies $\psi$ and using one or two obvious conditions on $\mathit{fs}_2$ (since $t \in \mathit{fs}_2[s]$), we show the derivation in the proof.
+If this can not be directly obtained from the information that $s$ satisfies $\psi$ and using one or two obvious conditions on $\mathit{fs}_2$ (since $t \in \mathit{fs}_2[s]$), we show the derivation in the proof.
 Then, $t'$ is obtained in a unique way since $AbsS$ is deterministic on its alphabet $ \Sigma_{AbsS} = C \cup R \cup Com(pop)$.
@@
 We show this by proving that $t'$ does not violate any of the conditions of the $\mathit{fs}_2$ described above.
+Suppose conditions on $\mathit{fs}_2$ are of the form $\forall \overline{k}.
+guard_{s,t}(\overline{k}) \triangleright \phi_{s,t}(\overline{k})$ where the $overline{k}$ is a vector of operation identifiers and $\phi$ defined on states $s$ and $t$ must hold if the guard defined on $s$ and $t$ holds.
+We say that a vector $\overline{k_1}$ is a new instantiation of the condtion if $\overline{k_1}$ not satisfy the $guard_{s,t}$ while relating pre-states, but it satisfies $guard_{s',t'}$ while relating post-states and $\phi_{s',t'}$ needs to be checked non-trivially.
+
 We only explain why the new instantiations due to the difference between $s'$ and $s$ or the difference between $t'$ and $t$ do not violate the conditions.
@@
 For the cases in which $\alpha' = \epsilon$, we have $t'=t$ and the only thing to show is $t \in \mathit{fs}_2[s']$.
-Again, we only explain why the new instantions due to the difference between $s'$ and $s$ do not violate the conditions.
+Again, we only explain why the new instantiations due to the difference between $s'$ and $s$ do not violate the conditions.
+
+In the following, we show that $\mathit{fs}_2$ is a $C \cup R \cup Com(pop)$-forward simulation relation.
 \begin{itemize}
@@
 %\textcolor{red}{TODO: Try case split on fs conditions instantiations check.}
-%\textcolor{red}{TODO: Add Flow Diagrams to the paper.} 
-%\textcolor{red}{TODO: }

### 2017-01-25 draftCAV/app-tss.tex
fixed a typo

@@
 Suppose conditions on $\mathit{fs}_2$ are of the form $\forall \overline{k}.
-guard_{s,t}(\overline{k}) \triangleright \phi_{s,t}(\overline{k})$ where the $overline{k}$ is a vector of operation identifiers and $\phi$ defined on states $s$ and $t$ must hold if the guard defined on $s$ and $t$ holds.
-We say that a vector $\overline{k_1}$ is a new instantiation of the condtion if $\overline{k_1}$ not satisfy the $guard_{s,t}$ while relating pre-states, but it satisfies $guard_{s',t'}$ while relating post-states and $\phi_{s',t'}$ needs to be checked non-trivially.
+guard_{s,t}(\overline{k}) \triangleright \phi_{s,t}(\overline{k})$ where the $\overline{k}$ is a vector of operation identifiers and $\phi$ defined on states $s$ and $t$ must hold if the guard defined on $s$ and $t$ holds.
+We say that a vector $\overline{k_1}$ is a new instantiation of the condtion if $\overline{k_1}$ does not satisfy the $guard_{s,t}$ while relating pre-states, but it satisfies $guard_{s',t'}$ while relating post-states.
 

### 2017-05-03 draftCAV/queue.tex
changes on the abstract queue implementation

@@
 \vspace{-1.5mm}
+
+%\textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too.
+We should be able to return EMPTY if there are only pending nodes.
+Consider the following history of $AbsQ_0$: $inv(enq, d_1,k_1), inv(enq, d_2,k_2), inv(deq, k_3), lin(deq, \texttt{EMPTY}, k_3)$.
+This history should be reflected in $AbsQ$ by enabling lin \texttt{EMPTY} of dequeue when there are pending nodes.
+We also need to update the rules in figure.}
+Informally, $AbsQ$ records the set of enqueue operations, whose input value has not yet been removed by a matching dequeue operation.
+In addition, it records the happens-before order between those enqueue operations: this is a partial order ordering an enqueue $k_1$ before another enqueue $k_2$ iff $k_1$ returned before $k_2$ was invoked.
+%Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation. 
+The linearization point of a dequeue can either remove a minimal enqueue $k$ (w.r.t.
+the happens-before stored in the state) and fix the return value to the value $d$ added by $k$, or fix the return value to ${\tt EMPTY}$ provided that the current state stores only pending enqueues (intuitively, the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
+%with return value $d\neq{\tt EMPTY}$ is enabled only if the happens-before stored in the current state contains a minimal enqueue that adds the value $d$.
+The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
+%When 
+%%\noindent
+%the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueues (because, the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
+%The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.
+
 \begin{wrapfigure}{l}{6.7cm}
@@
 \end{wrapfigure}
-%\textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too.
-We should be able to return EMPTY if there are only pending nodes.
-Consider the following history of $AbsQ_0$: $inv(enq, d_1,k_1), inv(enq, d_2,k_2), inv(deq, k_3), lin(deq, \texttt{EMPTY}, k_3)$.
-This history should be reflected in $AbsQ$ by enabling lin \texttt{EMPTY} of dequeue when there are pending nodes.
-We also need to update the rules in figure.}
-Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation.
-The linearization point of a dequeue operation with return value $d\neq{\tt EMPTY}$ is enabled only if the happens-before stored in the current state contains a minimal enqueue that adds the value $d$.
-The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
-When
-%\noindent
-the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueues (the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
-The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.
-
 Figure~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
@@
 
-Formally, the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).
-All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}.
+For readability, we define $AbsQ$ as an abstract state machine, which is given in Fig.~\ref{fig:transitions:AbsQ}.
+The states of $AbsQ$ consist of several boolean predicates indicating whether the value added by an enqueue has not been removed yet (the predicate $\mathsf{present})$, whether an enqueue is pending (the predicate $\mathsf{pending}$), the happens-before order (the predicate $\mathsf{before}$), and a function giving the value added by an enqueue (the function $\mathsf{arg}$).
+%the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).
+%All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}.
 The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points, denoted by $lin(deq,d,k)$.
 $Lin(deq)$ is the set of all actions $lin(deq,d,k)$.
+%
+The enqueue operations consist of two macro rules: the rule {\tt inv(enq,x,k)} orders the invoked operation after all the completed enqueues present in the current state ({\tt k} is a fresh operation identifier associated to the current operation), and the rule {\tt ret(enq,k)} sets $\mathsf{pending}${\tt (k)} to {\tt false}, provided that the operation is still present in the current state.
+The dequeue operations consist of two rules {\tt inv(deq,k)} and {\tt ret(deq,y,k)} associated to the invocation and respectively, the return action, whose definition is obvious, and the rule {\tt lin(deq,y,k)} corresponding to its linearization point.
+When a non-empty set of pending enqueues are present, the rule {\tt lin(deq,y,k)} is non-deterministic.
+It can set the return value {\tt y} either to {\tt EMPTY}, or to a value {\tt d} $\neq$ {\tt EMPTY}, provided that {\tt d} has been added by an enqueue {\tt k} which is minimal in the current happens-before.
+In the latter case, $\mathsf{present}${\tt (k)} is set to {\tt false} to indicate that {\tt k}'s input has been removed.
+When there exists at least one completed enqueue whose value has not been removed, the rule {\tt lin(deq,y,k)} sets {\tt y} to the value added by a minimal enqueue (w.r.t.
+the current happens-before).
 
-Concerning enqueue operations, the rule {\sc call-enq} orders the invoked operation after all the completed enqueues in the current state, and the rules {\sc ret-enq1}/{\sc ret-enq2} flip the corresponding flag from {\tt PEND} to {\tt COMP} provided that the operation is still present in the current state.
+%{\sc call-enq} orders the invoked operation after all the completed enqueues in the current state, and the rules {\sc ret-enq1}/{\sc ret-enq2} flip the corresponding flag from {\tt PEND} to {\tt COMP} provided that the operation is still present in the current state.
 For dequeue operations, {\sc call-deq} only increments the control point and {\sc ret-deq} checks whether the return value is the same as the one fixed at the linearization point.
@@
 
-\begin{figure} [t]
-{\scriptsize
-  \centering
-  \begin{mathpar}
-    \inferrule[call-enq]{
-      k\not\in dom(cp) \\ 
-      d\neq {\tt EMPTY}
-    }{
-      O,<,\ell,rv,cp
-      \xrightarrow{inv(enq,d,k)}
-      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
-      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1]
-    }\hspace{5mm}
+%\begin{figure} [t]
+\begin{figure}[t]
+\hspace{-8mm}
+\begin{minipage}[t]{7.5cm}
+\scriptsize{
+{	%\centering
+	\begin{align*}
+%		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
+		& \text{\underline{Shared State}} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \mathsf{Operations} \rightarrow \mathsf{Bool} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \mathsf{Operations} \rightarrow \mathsf{Bool} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \mathsf{Operations} \times \mathsf{Operations} \rightarrow \mathsf{Bool} \\
+%		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
+%		& \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} &\mbox{ // the argument of an enqueue} \\
+	\end{align*}
+}}
+\end{minipage}
+\begin{minipage}[t]{7.5cm}
+\begin{lstlisting}
+void enq($\mathsf{Values}$ x) {
+  atomic rule inv(enq,x,k = new);
+  atomic rule ret(enq,k);
+}
 
-    \inferrule[call-deq]{
-      k\not\in dom(cp) \\ 
-    }{
-      O,<,\ell,rv,cp
-      \xrightarrow{inv(deq,k)}
-      O,<,\ell,rv,cp[k\mapsto R_1]
-    }\hspace{5mm}
-    \inferrule[ret-deq]{
-       cp(k) = R_2 \\
-       rv(k)=d  
-    }{
-      O,<,\ell,rv,cp
-      \xrightarrow{ret(deq,d,k)}
-      O,<,\ell,rv,cp[k\mapsto R_3]
-    }\hspace{5mm}
+$\mathsf{Values}$ deq() {
+  atomic rule inv(deq,k = new);
+  atomic rule lin(deq,y,k);
+  atomic rule ret(deq,y,k);
+}
+\end{lstlisting}
+\end{minipage}
 
-    \inferrule[ret-enq1]{
-      cp(k) = A_1 \\
-      k \in O \\
-      \ell(k) = (d,{\tt PEND}) 
-    }{
-      O,<,\ell,rv,cp
-      \xrightarrow{ret(enq,k)}
-      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2]
-    }\hspace{5mm}
-    \inferrule[ret-enq2]{
-      cp(k) = A_1 \\
-      k \not\in O 
-    }{
-      O,<,\ell,rv,cp
-      \xrightarrow{ret(enq,k)}
-      O,<,\ell,rv,cp[k\mapsto A_2]
-    }\hspace{5mm}
+\begin{minipage}[t]{7.5cm}
+\begin{lstlisting}
+rule inv(enq,x,k) = {
+  $\mathsf{present}$(k) = true;
+  $\mathsf{pending}$(k) = true;
+  $\mathsf{arg}$(k) = x;
+  forall k' with $\mathsf{present}$(k)$\land \neg \mathsf{pending}$(k') do
+    $\mathsf{before}$(k',k) = true;
+}
 
-    \inferrule[lin-deq1]{
-       cp(k) = R_1 \\
-       d\neq{\tt EMPTY} \\
-       k'\in min(O) \\ 
-       \ell_1(k')=d
-    }{
-      O,<,\ell,rv,cp
-      \xrightarrow{lin(deq,d,k)}
-      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2]
-    }\hspace{5mm}
-    \inferrule[lin-deq2]{
-       cp(k) = R_1 \\
-       \forall o\in O.\ \ell_2(o)={\tt PEND}
-    }{
-      O,<,\ell,rv,cp
-      \xrightarrow{lin(deq,{\tt EMPTY},k)}
-      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2]
-    }\hspace{5mm}    
-      \end{mathpar}
+rule lin(deq,y,k) = {
+  choose a $\in$ {true, false};
+  if ((a || $\forall k.\ \mathsf{present}(k)$=false) && $\forall k.\ \mathsf{present}(k)\Rightarrow \mathsf{pending}(k)$) 
+    y = EMPTY;
+  else 
+    choose k $\in$ $\mathsf{Operations}$ with $\mathsf{present}$(k)$\land\neg \exists k'.\ \mathsf{present}(k')\land \mathsf{before}(k',$k$)$;
+    $\mathsf{present}$(k) = false;
+    y = $\mathsf{arg}$(k);
 }
+\end{lstlisting}
+\end{minipage}
+\begin{minipage}[t]{5cm}
+\vspace{-1mm}
+\begin{lstlisting}
+rule ret(enq,k) = {
+  $\mathsf{pending}$(k) = false;
+  return;
+}
+rule inv(deq,k) = {
+  skip;
+}
+rule ret(deq,y,k) = {
+  return y;
+}
+\end{lstlisting}
+\end{minipage}
+%	\caption{The abstract queue implementation $AbsQ$.}
+%	\label{fig:signatures}
+%\end{figure}
+
+%{\scriptsize
+%  \centering
+%  \begin{mathpar}
+%    \inferrule[call-enq]{
+%      k\not\in dom(cp) \\ 
+%      d\neq {\tt EMPTY}
+%    }{
+%      O,<,\ell,rv,cp
+%      \xrightarrow{inv(enq,d,k)}
+%      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
+%      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1]
+%    }\hspace{5mm}
+%
+%    \inferrule[call-deq]{
+%      k\not\in dom(cp) \\ 
+%    }{
+%      O,<,\ell,rv,cp
+%      \xrightarrow{inv(deq,k)}
+%      O,<,\ell,rv,cp[k\mapsto R_1]
+%    }\hspace{5mm}
+%    \inferrule[ret-deq]{
+%       cp(k) = R_2 \\
+%       rv(k)=d  
+%    }{
+%      O,<,\ell,rv,cp
+%      \xrightarrow{ret(deq,d,k)}
+%      O,<,\ell,rv,cp[k\mapsto R_3]
+%    }\hspace{5mm}
+%
+%    \inferrule[ret-enq1]{
+%      cp(k) = A_1 \\
+%      k \in O \\
+%      \ell(k) = (d,{\tt PEND}) 
+%    }{
+%      O,<,\ell,rv,cp
+%      \xrightarrow{ret(enq,k)}
+%      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2]
+%    }\hspace{5mm}
+%    \inferrule[ret-enq2]{
+%      cp(k) = A_1 \\
+%      k \not\in O 
+%    }{
+%      O,<,\ell,rv,cp
+%      \xrightarrow{ret(enq,k)}
+%      O,<,\ell,rv,cp[k\mapsto A_2]
+%    }\hspace{5mm}
+%
+%    \inferrule[lin-deq1]{
+%       cp(k) = R_1 \\
+%       d\neq{\tt EMPTY} \\
+%       k'\in min(O) \\ 
+%       \ell_1(k')=d
+%    }{
+%      O,<,\ell,rv,cp
+%      \xrightarrow{lin(deq,d,k)}
+%      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2]
+%    }\hspace{5mm}
+%    \inferrule[lin-deq2]{
+%       cp(k) = R_1 \\
+%       \forall o\in O.\ \ell_2(o)={\tt PEND}
+%    }{
+%      O,<,\ell,rv,cp
+%      \xrightarrow{lin(deq,{\tt EMPTY},k)}
+%      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2]
+%    }\hspace{5mm}    
+%      \end{mathpar}
+%  }
  \vspace{-6mm}
-  \caption{The transition relation of $AbsQ$.
+  \caption{The abstract queue implementations $AbsQ$. Each macro rule is executed atomically in one step (indicated by the keyword {\ttfamily \bfseries atomic}).
+%  transition relation of $AbsQ$.
 We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.
@@
 % Let $AbsQ_0$ denote this implementation (formally defined in Appendix~\ref{app:absImplQueue}).
-The following result states that the library $AbsQ$ has exactly the same set of histories as the standard abstract library $AbsQ_0$ (see Appendix~\ref{app:absImplQueue} for a proof).
+The following result states that the library $AbsQ$ has exactly the same set of histories as the standard abstract library $AbsQ_0$. % (see Appendix~\ref{app:absImplQueue} for a proof).
 
@@
 \end{theorem}
+
+The abstract state machine in Fig.~\ref{fig:transitions:AbsQ} defines an LTS over an alphabet $\Sigma$ such that $C\cup R\cup Lin(deq)\subseteq \Sigma$.
+The transitions corresponding to the rule {\tt lin(deq,y,k)} are labeled by $lin(deq,y,k)$ actions.
+Also, we assume that the transition corresponding to the invocation of a method is done atomically with the first step of the same invocation (the macro rule {\tt inv(...)}), and similarly for transitions corresponding to returning from a method, they are done atomically with the last step (the macro rule {\tt ret(...)}).
+They are labeled as expected with call/return actions (borrowing the operation identifier which occurs as argument of the macro rules).
+It can be easily proved that this assumption doesn't modify the set of traces projected on $C\cup R\cup Lin(deq)$.
+
+%~\footnote{Without this assumption, while proving forward simulations, a call/return action in the concrete implementation is simulated by two steps of the abstract implementation, the call/return action together with the first/last macro rule.}.
 

### 2017-05-03 draftCAV/refProofs.tex
fixed some spaces
changes on the abstract queue implementation

@@
 traces projected on $\Gamma$. Library refinement corresponds to the case $\Gamma=C\cup R$. 
-Typically, $\Gamma$-refinement between two LTSs $A$ and $B$ is proved using \emph{simulation relations} which roughly, require that 
-$B$ can mimic every step of $A$ using a (possibly empty) sequence of steps. Mainly, there are two kinds of simulation
+Typically, $\Gamma$-refinement between two LTSs $A_1$ and $A_2$ is proved using \emph{simulation relations} which roughly, require that 
+$A_2$ can mimic every step of $A_1$ using a (possibly empty) sequence of steps. Mainly, there are two kinds of simulation
 relations, forward or backward, depending on whether the preservation of steps is proved starting from a similar state
@@
 that record events in the implementation, and to the existence of \emph{forward simulations} provided that the right-hand side
-LTS $B$ is $\Gamma$-deterministic~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}. 
+LTS $A_2$ is $\Gamma$-deterministic~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}. 
 We focus on proofs based on forward simulations because they are easier to automatize.
@@
 
-Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over $\Sigma_1$ and $\Sigma_2$, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$.
-Also, let $\Gamma$ be a set of actions s.t.
-$C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
+Let $A_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $A_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two LTSs over $\Sigma_1$ and $\Sigma_2$, respectively, such that $\Gamma\subseteq \Sigma_1\cap\Sigma_2$.
 
@@
 Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
-The library $L_1$ \emph{$\Gamma$-refines} $L_2$ if{f} $Tr(L_1) | \Gamma \subseteq Tr(L_2) | \Gamma$.
+The LTS $A_1$ \emph{$\Gamma$-refines} $A_2$ if{f} $Tr(A_1) | \Gamma \subseteq Tr(A_2) | \Gamma$.
 %We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
@@
 
-Notice that $\Gamma$-refinement implies refinement for any $\Gamma$ as in Definition~\ref{def:gref}.
+The notion of $\Gamma$-refinement instantiated to libraries (i.e., to LTSs defining libraries) implies the notion of refinement in Definition~\ref{def:lib_ref}, for every $\Gamma$ such that $C\cup R \subseteq \Gamma$.
+%, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions s.t. $C\cup R\subseteq $
+%
+%When $A_1$ and $A_2$ are libraries, $\Gamma$-refinement implies the notion of library refinement  for any $\Gamma$ as in Definition~\ref{def:gref}.
 
-We define a notion of \emph{forward} simulation that can be used to prove $\Gamma$-refinement (a dual notion of \emph{backward} simulation is defined in 
-Appendix~\ref{app:backSim}). 
-%We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement. 
+We define a notion of \emph{forward} simulation that can be used to prove $\Gamma$-refinement %We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement.
 For a relation $R\subseteq A\times B$, $R[X]$ is the set of elements related by $R$ to elements of $X$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.
@@
 %Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
-A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ if{f} $\mathit{fs}[s_0^1] = \{s_0^2 \}$ and:
+A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $A_1$ to $A_2$ if{f} $\mathit{fs}[s_0^1] = \{s_0^2 \}$ and:
 \vspace{-1.5mm}
@@
 %If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
-A $\Gamma$-forward simulation requires that every step of $L_1$ corresponds to a sequence of steps of $L_2$.
-To imply $\Gamma$-refinement, every step of $L_1$ labeled by an observable action $\gamma\in \Gamma$ should be simulated by a sequence of steps of $L_2$ where exactly one transition is labeled by $\gamma$ and all the other transitions are labeled by non-observable actions.
+A $\Gamma$-forward simulation requires that every step of $A_1$ corresponds to a sequence of steps of $A_2$.
+To imply $\Gamma$-refinement, every step of $A_1$ labeled by an observable action $\gamma\in \Gamma$ should be simulated by a sequence of steps of $A_2$ where exactly one transition is labeled by $\gamma$ and all the other transitions are labeled by non-observable actions.
+The dual notion of \emph{backward} simulation where steps are simulated backwards can be defined in a similar way.
 
-The following theorem shows the soundness and the completeness of $\Gamma$-forward simulations (when $L_2$ is $\Gamma$-deterministic).
+The following shows the soundness and the completeness of $\Gamma$-forward simulations (when $A_2$ is $\Gamma$-deterministic).
 It is an instantiation of previous results~~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}.
@@
 Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
-$L_1$ $\Gamma$-refines $L_2$ when there is a $\Gamma$-forward simulation from $L_1$ to $L_2$.
-Moreover, if $L_1$ $\Gamma$-refines $L_2$ and $L_2$ is $\Gamma$-deterministic, then there is a $\Gamma$-forward simulation from $L_1$ to $L_2$.
+$A_1$ $\Gamma$-refines $A_2$ when there is a $\Gamma$-forward simulation from $A_1$ to $A_2$.
+Moreover, if $A_1$ $\Gamma$-refines $A_2$ and $A_2$ is $\Gamma$-deterministic, then there is a $\Gamma$-forward simulation from $A_1$ to $A_2$.
 \vspace{-1.5mm}

### 2017-05-04 draftCAV/draft.tex
the abstract queue
changes on the abstract queue implementation
removing appendices
stack implementation

@@
 \lstset{
-  language={Java},
+  language={program},
   basicstyle=\ttfamily\scriptsize
@@
 \title{Proving linearizability using forward simulations}
-\author{ }
-\institute{ } 
+\author{Ahmed Bouajjani\inst{1}\and Michael Emmi\inst{2}\and Constantin Enea\inst{1} \and Suha Orhun Mutluergil\inst{3}}
+\institute{IRIF, Univ. Paris Diderot, \email{\{abou,cenea\}@irif.fr} \and Nokia Bell Labs, \email{michael.emmi@nokia.com}\and Koc University, \email{smutluergil@ku.edu.tr}}
 \begin{document}
@@
 
-\vspace{-9mm}
+%\vspace{-9mm}
 \begin{abstract}
 %Efficient concurrent implementations of abstract data structures such as stacks and queues are especially susceptible to programming errors, and necessitate automatic verification.
-Linearizability is the standard correctness criterion concurrent data structures such as stacks and queues. %
+Linearizability is the standard correctness criterion for concurrent data structures such as stacks and queues. %
 %in this context. 
@@
 We define classes of queue and stack implementations for which forward simulations are sound and complete for proving their correctness.
-\vspace{-2mm}
+%\vspace{-2mm}
 \end{abstract}
@@
 
+%\input{abstract_queues}
+
 \input{queue}
@@
 
-\newpage
-\appendix
-
-\input{app-prelim}
-\input{app-backSim}
-\input{app-absImplQueue}
-\input{app-absImplStack}
-\input{app-tss}
+%\newpage
+%\appendix
+%
+%\input{app-prelim}
+%\input{app-backSim}
+%\input{app-absImplQueue}
+%\input{app-absImplStack}
+%\input{app-tss}
 

### 2017-05-04 draftCAV/queue.tex
changes on the abstract queue implementation
stack implementation

@@
 
-Since the linearization points of the enqueues are not fixed, there exists no forward simulation from $\mathit{HWQ}$ to $AbsQ_0$. 
+Since the linearization points of the enqueues are determined by future dequeue invocations, there exists no forward simulation from $\mathit{HWQ}$ to $AbsQ_0$.  
 In the following, we describe the abstract implementation $AbsQ$ for which such a forward simulation does exist.
@@
 \vspace{-8mm}
-\caption{Simulating queue histories with $AbsQ$. The order between actions is from left to right.}
+\caption{Simulating queue histories with $AbsQ$.
+An operation is pictured by a line delimited by two circles denoting the call and respectively, the return action.
+The representation of a dequeue operation includes a red circle that stands for a {\tt swap} returning a non-null value, which is their linearization point.}
 \label{fig:queueSim}
@@
 %
-The enqueue operations consist of two macro rules: the rule {\tt inv(enq,x,k)} orders the invoked operation after all the completed enqueues present in the current state ({\tt k} is a fresh operation identifier associated to the current operation), and the rule {\tt ret(enq,k)} sets $\mathsf{pending}${\tt (k)} to {\tt false}, provided that the operation is still present in the current state.
+The enqueue operations consist of two macro rules: the rule {\tt inv(enq,x,k)} orders the invoked operation after all the completed enqueues present in the current state ({\tt k} is a fresh operation identifier associated to the current operation), and the rule {\tt ret(enq,k)} sets $\mathsf{pending}${\tt (k)} to {\tt false}.
+% provided that the operation is still present in the current state.
 The dequeue operations consist of two rules {\tt inv(deq,k)} and {\tt ret(deq,y,k)} associated to the invocation and respectively, the return action, whose definition is obvious, and the rule {\tt lin(deq,y,k)} corresponding to its linearization point.
 When a non-empty set of pending enqueues are present, the rule {\tt lin(deq,y,k)} is non-deterministic.
-It can set the return value {\tt y} either to {\tt EMPTY}, or to a value {\tt d} $\neq$ {\tt EMPTY}, provided that {\tt d} has been added by an enqueue {\tt k} which is minimal in the current happens-before.
-In the latter case, $\mathsf{present}${\tt (k)} is set to {\tt false} to indicate that {\tt k}'s input has been removed.
+It can set the return value {\tt y} either to {\tt EMPTY}, or to a value {\tt d} $\neq$ {\tt EMPTY}, provided that {\tt d} has been added by an enqueue {\tt k1} which is minimal in the current happens-before.
+In the latter case, $\mathsf{present}${\tt (k1)} is set to {\tt false} to indicate that {\tt k1}'s input has been removed.
 When there exists at least one completed enqueue whose value has not been removed, the rule {\tt lin(deq,y,k)} sets {\tt y} to the value added by a minimal enqueue (w.r.t.
@@
 \begin{minipage}[t]{7.5cm}
+\vspace{-3mm}
 \begin{lstlisting}
@@
   $\mathsf{arg}$(k) = x;
-  forall k' with $\mathsf{present}$(k)$\land \neg \mathsf{pending}$(k') do
-    $\mathsf{before}$(k',k) = true;
+  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1) do
+    $\mathsf{before}$(k1,k) = true;
 }
@@
   choose a $\in$ {true, false};
-  if ((a || $\forall k.\ \mathsf{present}(k)$=false) && $\forall k.\ \mathsf{present}(k)\Rightarrow \mathsf{pending}(k)$) 
+  if ( $($a$\ \vee\ \forall k'.\ \mathsf{present}(k')$=false$)\ \land\ \forall k'.\ \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$ ) 
     y = EMPTY;
   else 
-    choose k $\in$ $\mathsf{Operations}$ with $\mathsf{present}$(k)$\land\neg \exists k'.\ \mathsf{present}(k')\land \mathsf{before}(k',$k$)$;
-    $\mathsf{present}$(k) = false;
-    y = $\mathsf{arg}$(k);
+    choose k1 $\in$ $\mathsf{Operations}$ with $\mathsf{present}$(k1)$\land\neg \exists k'.\ \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$;
+    $\mathsf{present}$(k1) = false;
+    y = $\mathsf{arg}$(k1);
 }
@@
 \begin{minipage}[t]{5cm}
-\vspace{-1mm}
+\vspace{-3mm}
 \begin{lstlisting}
@@
 %  }
- \vspace{-6mm}
+ \vspace{-4mm}
   \caption{The abstract queue implementations $AbsQ$. Each macro rule is executed atomically in one step (indicated by the keyword {\ttfamily \bfseries atomic}).

### 2017-05-04 draftCAV/refProofs.tex
stack implementation

@@
 
-The notion of $\Gamma$-refinement instantiated to libraries (i.e., to LTSs defining libraries) implies the notion of refinement in Definition~\ref{def:lib_ref}, for every $\Gamma$ such that $C\cup R \subseteq \Gamma$.
+The notion of $\Gamma$-refinement instantiated to libraries (i.e., to LTSs defining libraries) implies the notion of refinement in Definition~\ref{def:lib_ref} for every $\Gamma$ such that $C\cup R \subseteq \Gamma$.
 %, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions s.t. $C\cup R\subseteq $
@@
 %Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
-A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $A_1$ to $A_2$ if{f} $\mathit{fs}[s_0^1] = \{s_0^2 \}$ and:
+A relation $F \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $A_1$ to $A_2$ if{f} $F(s_0^1,s_0^2)$ and:
 \vspace{-1.5mm}
 \begin{itemize}
-\item If $(s,\gamma,s') \in \delta_1$, for some $\gamma\in \Gamma$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_i=\gamma$, for some $i$, and $@s_j\in \Sigma_2\setminus\Gamma$, for each $j\neq i$.
-\item If $(s,e,s') \in \delta_1$, for some $e \in \Sigma_1\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $\sigma\in (\Sigma_2\setminus\Gamma)^*$.
+\item For all $s,s'\in Q_1$, $\gamma\in \Gamma$, and $u\in Q_2$, such that $(s,\gamma,s') \in \delta_1$ and $F(s,u)$, we have that there exists $u'\in Q_2$ such that $F(s',u')$ and $u \xrightarrow{@s} u'$ where $@s_i=\gamma$, for some $i$, and $@s_j\in \Sigma_2\setminus\Gamma$, for all $j\neq i$.
+\item For all $s,s'\in Q_1$, $e \in \Sigma_1\setminus \Gamma$, and $u\in Q_2$, such that $(s,e,s') \in \delta_1$ and $F(s,u)$, we have that there exists $u'\in Q_2$ such that $F(s',u')$ and $u \xrightarrow{\sigma} u'$ where $\sigma\in (\Sigma_2\setminus\Gamma)^*$.
 %where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
@@
 reduces to proving that $L_1$ is a $(C\cup R\cup Lin)$-refinement of an abstract implementation $L_2$ of the same object where method
-bodies are guarded with a global-lock acquisition. Since the abstract implementation is usually $(C\cup R\cup Lin)$-deterministic,
-by Theorem~\ref{th:forSim}, proving $(C\cup R\cup Lin)$-refinement is equivalent to finding a $(C\cup R\cup Lin)$-forward simulation 
+bodies are guarded with a global-lock acquisition. As a direct consequence of Theorem~\ref{th:forSim}, since the abstract implementation is 
+ $(C\cup R\cup Lin)$-deterministic, proving $(C\cup R\cup Lin)$-refinement is equivalent to finding a $(C\cup R\cup Lin)$-forward simulation 
 from $L_1$ to $L_2$.

### 2017-05-04 draftCAV/stack.tex
stack implementation

@@
 Exploiting particular properties of the stack semantics, we refine the ideas used in $AbsQ$ and define
-a new abstract implementation for stacks, denoted as $AbsQ$, which is able to simulate such implementations.
+a new abstract implementation for stacks, denoted as $AbsS$, which is able to simulate such implementations.
 Forward simulations to $AbsS$ are complete for proving the correctness of stack implementations provided that the point in time where the return value of a pop operation is determined, called \emph{commit point}, corresponds to a fixed action.
@@
 \vspace{-2mm}
-{\scriptsize
-  \centering
-  \begin{mathpar}
-    \inferrule[call-push]{
-      k\not\in dom(cp) \\ 
-      d\neq {\tt EMPTY} \\
-      \forall k'.\ ov'(k') = ov(k')\cup \{k\}
-    }{
-      O,<,\ell,rv,cp,be,ov 
-      \xrightarrow{inv(push,d,k)} 
-      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1],be,ov'
-    }\hspace{5mm}
+\begin{minipage}[t]{7.5cm}
+\scriptsize{
+{	%\centering
+	\begin{align*}
+%		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
+		& \text{\underline{Shared State}} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \mathsf{Operations} \rightarrow \mathsf{Bool} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \mathsf{Operations} \rightarrow \mathsf{Bool} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \mathsf{Operations} \times \mathsf{Operations} \rightarrow \mathsf{Bool} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{maxAtInvoc} : \mathsf{Operations} \rightarrow 2^\mathsf{Operations} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{overlap} : \mathsf{Operations} \rightarrow 2^\mathsf{Operations} \\
+%		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
+%		& \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} &\mbox{ // the argument of an enqueue} \\
+	\end{align*}
+}}
+\end{minipage}
+\begin{minipage}[t]{7.5cm}
+\begin{lstlisting}
+void push($\mathsf{Values}$ x) {
+  atomic rule inv(push,x,k = new);
+  atomic rule ret(push,k);
+}
 
-    \vspace{-1mm}
-    \inferrule[call-pop]{
-      k\not\in dom(cp) %\\ k\not\in dom(cp) \\ k\not\in dom(cp)
-    }{
-      O,<,\ell,rv,cp,be,ov
-      \xrightarrow{inv(pop,k)} 
-      O,<,\ell,rv,cp[k\mapsto R_1],be[k\mapsto maxCo(O)],ov[k\mapsto {\tt PEND}(O)]
-    }\hspace{5mm}
+$\mathsf{Values}$ pop() {
+  atomic rule inv(pop,k = new);
+  atomic rule com(pop,y,k);
+  atomic rule ret(pop,y,k);
+}
+\end{lstlisting}
+\end{minipage}
 
-    \vspace{-1mm}
-    \inferrule[ret-pop]{
-       cp(k) = R_2 \\
-       rv(k)=d  
-    }{
-      O,<,\ell,rv,cp,be,ov
-      \xrightarrow{ret(pop,d,k)}
-      O,<,\ell,rv,cp[k\mapsto R_3],be,ov
-    }\hspace{5mm}
+\vspace{-5mm}
+\begin{minipage}[t]{6.5cm}
+\begin{lstlisting}
+rule inv(push,x,k) = {
+  $\mathsf{present}$(k) = true;
+  $\mathsf{pending}$(k) = true;
+  $\mathsf{arg}$(k) = x;
+  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1) do
+    $\mathsf{before}$(k1,k) = true;
+  forall k1 do
+    $\mathsf{overlap}$(k1) = $\mathsf{overlap}$(k1) $\cup$ $\{$k$\}$;
+}
+rule ret(push,k) = {
+  $\mathsf{pending}$(k) = false;
+  return;
+}
+\end{lstlisting}
+\end{minipage}
+\begin{minipage}[t]{5cm}
+%\vspace{-3mm}
+\begin{lstlisting}
+rule inv(pop,k) = {
+  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1)$\land$
+                 $\neg \exists k'.\ \mathsf{present}(k')\land \neg \mathsf{pending}(k')\land$ 
+                 $\mathsf{before}$(k1,$k'$) do
+    $\mathsf{maxAtInvoc}$(k) = $\mathsf{maxAtInvoc}$(k) $\cup$ $\{$k1$\}$;
+  forall k1 with $\mathsf{present}$(k1)$\land \mathsf{pending}$(k1) do
+    $\mathsf{overlap}$(k) = $\mathsf{overlap}$(k) $\cup$ $\{$k1$\}$;
+}
+rule ret(pop,y,k) = {
+  return y;
+}
+\end{lstlisting}
+\end{minipage}
 
-    \vspace{-1mm}
-    \inferrule[ret-push1]{
-      cp(k) = A_1 \\
-      k \in O \\
-      \ell(k) = (d,{\tt PEND}) 
-    }{
-      O,<,\ell,rv,cp,be,ov
-      \xrightarrow{ret(push,k)}
-      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2],be,ov
-    }\hspace{5mm}
-    \inferrule[ret-push2]{
-      cp(k) = A_1 \\
-      k \not\in O 
-    }{
-      O,<,\ell,rv,cp,be,ov
-      \xrightarrow{ret(push,k)}
-      O,<,\ell,rv,cp[k\mapsto A_2],be,ov
-    }\hspace{5mm}
+\vspace{-4mm}
+\begin{lstlisting}
+rule com(pop,y,k) = {
+  choose a $\in$ {true, false};
+  if ((a || $\forall k'.\ \mathsf{present}(k')$=false) && $\mathsf{maxAtInvoc}$(k)=$\,\emptyset$) 
+    y = EMPTY;
+  else 
+    choose k1 $\in$ $\mathsf{Operations}$ with $\mathsf{present}$(k1)$\land$ k1$\in\mathsf{maxAtInvoc}$(k)$\cup\,\mathsf{overlap}$(k) do
+    $\mathsf{present}$(k1) = false;
+    y = $\mathsf{arg}$(k1);
+    forall k2 with k1$\in\mathsf{maxAtInvoc}$(k2) do
+      $\mathsf{maxAtInvoc}$(k2) = $\mathsf{maxAtInvoc}$(k2)$\setminus\{$k1$\}$;
+      forall k3 with $\mathsf{before_1}$(k3,k1)$\land \forall k'.\ (\mathsf{before_1}($k3$,k')\land k'\neq $k1$)\Rightarrow k'\in \mathsf{overlap}($k2$)$ do
+        $\mathsf{maxAtInvoc}$(k2) = $\mathsf{maxAtInvoc}$(k2)$\cup\{$k3$\}$;
+}
+\end{lstlisting}
+ %\neg\exists k'.\ \mathsf{present}(k')\,=\,$true$\,\land\, \mathsf{before}($k3,$k')\land \mathsf{before}(k',$k1$)$
 
-    \inferrule[com-pop1]{
-       cp(k) = R_1 \\
-       d\neq{\tt EMPTY} \\
-       k'\in be(k)\cup ov(k) \\
-        \ell_1(k')=d \\
-       \forall k_1.\ k'\not\in be(k_1) \Rightarrow be'(k_1)=be(k_1) \\     \\
-       \forall k_1.\ k'\in be(k_1) \Rightarrow be'(k_1)=(be(k_1)\setminus\{k'\})\cup \{k_2: k_2\in pred_{<}(k') \land \forall k_3. (k_2\in pred_{<}(k_3) \land k_3\neq k') => k_3\in ov(k_1)\} \\
-    }{
-      O,<,\ell,rv,cp,be,ov
-      \xrightarrow{com(pop,d,k)} \\
-      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2],be',ov %be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}]
-    }\hspace{5mm}
-
-    %\vspace{-1mm}
-    \inferrule[com-pop2]{
-       cp(k) = R_1 \\
-       be(k)=\emptyset
-    }{
-      O,<,\ell,rv,cp,be,ov
-      \xrightarrow{com(pop,{\tt EMPTY},k)}
-      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2],be,ov
-    }\hspace{5mm}
-
-      \end{mathpar}
-  }
- \vspace{-6mm}
-  \caption{The transition relation of $AbsQ$.
+%{\scriptsize
+%  \centering
+%  \begin{mathpar}
+%    \inferrule[call-push]{
+%      k\not\in dom(cp) \\ 
+%      d\neq {\tt EMPTY} \\
+%      \forall k'.\ ov'(k') = ov(k')\cup \{k\}
+%    }{
+%      O,<,\ell,rv,cp,be,ov 
+%      \xrightarrow{inv(push,d,k)} 
+%      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1],be,ov'
+%    }\hspace{5mm}
+%
+%    \vspace{-1mm}
+%    \inferrule[call-pop]{
+%      k\not\in dom(cp) %\\ k\not\in dom(cp) \\ k\not\in dom(cp)
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{inv(pop,k)} 
+%      O,<,\ell,rv,cp[k\mapsto R_1],be[k\mapsto maxCo(O)],ov[k\mapsto {\tt PEND}(O)]
+%    }\hspace{5mm}
+%    
+%    \vspace{-1mm}
+%    \inferrule[ret-pop]{
+%       cp(k) = R_2 \\
+%       rv(k)=d  
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{ret(pop,d,k)}
+%      O,<,\ell,rv,cp[k\mapsto R_3],be,ov
+%    }\hspace{5mm}
+%
+%    \vspace{-1mm}
+%    \inferrule[ret-push1]{
+%      cp(k) = A_1 \\
+%      k \in O \\
+%      \ell(k) = (d,{\tt PEND}) 
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{ret(push,k)}
+%      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2],be,ov
+%    }\hspace{5mm}
+%    \inferrule[ret-push2]{
+%      cp(k) = A_1 \\
+%      k \not\in O 
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{ret(push,k)}
+%      O,<,\ell,rv,cp[k\mapsto A_2],be,ov
+%    }\hspace{5mm}
+%
+%    \inferrule[com-pop1]{
+%       cp(k) = R_1 \\
+%       d\neq{\tt EMPTY} \\
+%       k'\in be(k)\cup ov(k) \\
+%        \ell_1(k')=d \\
+%       \forall k_1.\ k'\not\in be(k_1) \Rightarrow be'(k_1)=be(k_1) \\     \\
+%       \forall k_1.\ k'\in be(k_1) \Rightarrow be'(k_1)=(be(k_1)\setminus\{k'\})\cup \{k_2: k_2\in pred_{<}(k') \land \forall k_3. (k_2\in pred_{<}(k_3) \land k_3\neq k') => k_3\in ov(k_1)\} \\
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{com(pop,d,k)} \\
+%      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2],be',ov %be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}]
+%    }\hspace{5mm}
+%
+%    %\vspace{-1mm}
+%    \inferrule[com-pop2]{
+%       cp(k) = R_1 \\
+%       be(k)=\emptyset
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{com(pop,{\tt EMPTY},k)}
+%      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2],be,ov
+%    }\hspace{5mm}
+%
+%    
+%      \end{mathpar}
+%  }
+ \vspace{-4mm}
+  \caption{The abstract stack implementation $AbsS$.
+We use $\mathsf{before_1}$ to denote the immediate successor relation induced by $\mathsf{before}$, i.e., $\mathsf{before}_1(k,k')={\tt true}$ iff $\neg\exists k''.\ \mathsf{present}(k'')\land \mathsf{before}(k,k'')\land \mathsf{before}(k'',k')$.
+%  The transition relation of $AbsQ$.
 We use the following notions: $maxCo(O)$ is the set of greatest operations in $O$ (w.r.t.
@@
 Thus, the $AbsS$ states $t=\tup{O,<,\ell,rv,cp,be,ov}$ associated by $\mathit{fs}_2$ to a $\mathit{TSS}$ state $s$ satisfy the following.
-The set $O$ consists of all the identifiers of pushes in $s$ which didn't added yet a node to {\tt pools} or for which the input is still present in {\tt pools} (i.e., the node created by the push has {\tt taken} set to {\tt false}).
+The set $O$ consists of all the identifiers of pushes in $s$ which haven't added yet a node to {\tt pools} or for which the input is still present in {\tt pools} (i.e., the node created by the push has {\tt taken} set to {\tt false}).
 A push $k$ is labeled by $(d,{\tt PEND})$ where $d$ is the input value if it's pending and by $(d,{\tt COMP})$, otherwise.
@@
 i.e., ${\tt youngest}_s(k)={\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
-	\item if the variable {\tt youngest} of a pop $k$ points to a node which is not taken, then this node was created by a push in $be(k)\cup ov(k)$ or the node currently reached by $k$ is followed in {\tt pools} by another node which was created by a push in $be(k)\cup ov(k)$, i.e., ${\tt youngest}_s(k)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(k)={\tt n}_s(k_2)$ implies $k_1\in be(k)\cup ov(k)$ or that there exists $k_3\in O$ such that ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in be(k)\cup ov(k)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and TODO $k$ is traversing the last list in the array {\tt pools}, for each $k, k_1,k_2$
+	\item if the variable {\tt youngest} of a pop $k$ points to a node which is not taken, then this node was created by a push in $be(k)\cup ov(k)$ or the node currently reached by $k$ is followed in {\tt pools} by another node which was created by a push in $be(k)\cup ov(k)$, i.e., ${\tt youngest}_s(k)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(k)={\tt n}_s(k_2)$ implies $k_1\in be(k)\cup ov(k)$ or that there exists $k_3\in O$ such that ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in be(k)\cup ov(k)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and $k$ is traversing the last list in the array {\tt pools}, for each $k, k_1,k_2$
 %	

### 2017-05-04 draftCAV/prelim.tex
changes on the abstract queue implementation
stack implementation
some minor issues

@@
 This model captures shared-memory programs with an arbitrary number of threads,
-abstracting away the details of any particular programming system irrelevant to
+abstracting away the details of any particular programming language irrelevant to
 our development.
@@
 initial state $s_0 \in Q$, and a transition relation $\delta \subseteq Q \times @S \times
-Q$. The $i$th symbol of a sequence $@t \in @S^*$ is denoted $@t_i$, and the empty
-sequence is denoted by $\epsilon$.
-An \emph{execution} of $A$ is an alternating sequence of states and transition labels (called also actions)
+Q$. The $i$th symbol of a sequence $@t \in @S^*$ is denoted $@t_i$, and $\epsilon$ denotes the empty
+sequence.
+An \emph{execution} of $A$ is an alternating sequence of states and transition labels (also called actions)
 $\rho = s_0, e_0,s_1\ldots e_{k-1},s_k$ for some $k>0$ such that $\delta(s_i, e_i, s_{i+1})$
@@
 Programs interact with libraries by calling named library \emph{methods}, which
-receive \emph{parameter values} and yield \emph{return values} upon completion.
+receive \emph{arguments} and yield \emph{return values} upon completion.
 We fix arbitrary sets $\<Methods>$ and $\<Vals>$ of method names and
-parameter/return values. 
+argument/return values. 
 %
@@
 %\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
-We may omit the second field from a call/return action $a$ for methods that have no inputs or return values.
+We may omit the second field from a call/return action for methods that have no arguments or return values.
 %A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
@@
 We assume that the traces of a library satisfy standard well-formedness properties, 
-e.g., return actions correspond to previous call actions, which for lack of 
-space are delegated to Appendix~\ref{app:prelim}. An operation $k$ is called \emph{completed} in a trace $\tau$ when
+e.g., return actions correspond to previous call actions.
+Given a standard library description as a set of methods, the LTS represents the executions of its most general client (that calls an arbitrary set of methods with an arbitrary set of threads in an unbounded loop).
+The states of this LTS consist of the shared state of the library together with the local state of each thread.
+The transitions correspond to statements in the methods' bodies, or call and return actions.
+%, which for lack of 
+%space are delegated to Appendix~\ref{app:prelim}. 
+An operation $k$ is called \emph{completed} in a trace $\tau$ when
 $ret(m,d,k)$ occurs in $\tau$, for some $m$ and $d$. Otherwise, it is called \emph{pending}.
@@
 For queue and stack implementations, we assume that the same value is never added twice, i.e., for every trace $@t$ of such a library and every two call actions $inv(m,d_1,k_1)$ and $inv(m,d_2,k_2)$ where $m\in \{enq,push\}$ we have that $d_1\neq d_2$.
-As shown in several works~\cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, this assumption is without loss of generality for libraries that are data independent, i.e., their behaviors are not influenced by the values added to the collection, which is always the case in practice.
+As shown in several works~\cite{conf/tacas/AbdullaHHJR13,DBLP:conf/icalp/BouajjaniEEH15}, this assumption is without loss of generality for libraries that are data independent, i.e., their behaviors are not influenced by the values added to the collection.
+All the queue and stack implementations that we are aware of are data independent.
 On a technical note, this assumption is used to define ($\Gamma$-)deterministic abstract implementations of stacks and queues in Section~\ref{sec:queues} and Section~\ref{sec:stacks}.
@@
 %\textcolor{red}{We have two Definition 1. The problem occurs because we used different commands for definitions.}
-\begin{definition}
+\begin{definition}\label{def:lib_ref}
 %Let $L_1$ and $L_2$ be two libraries. 
@@
 \begin{theorem}[\cite{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15}]
-  $L_1 \sqsubseteq L_2$ if{f} $L_1$ refines $L_2$,
-  if $L_2$ is atomic.
+Let $L_1$ and $L_2$ be two libraries, such that $L_2$ is atomic. Then, $L_1 \sqsubseteq L_2$ if{f} $L_1$ refines $L_2$.
 \vspace{-1.5mm}

### 2017-05-04 draftCAV/queue.tex
some minor issues

@@
 \vspace{-1mm}
-We describe a forward simulation $\mathit{fs}_1$ from $\mathit{HWQ}$ to $AbsQ$.
-A $\mathit{HWQ}$ state is related by $\mathit{fs}_1$ to an $AbsQ$ state that consists of all the  enqueue operations for which the input is still present in the array {\tt items} and all the pending enqueue operations that have at most reserved an array position, ordered by a relation $<$ satisfying the following:
+We describe a forward simulation $F_1$ from $\mathit{HWQ}$ to $AbsQ$.
+The description of $\mathit{HWQ}$ in Fig.~\ref{fig:HerlihyWing} defines an LTS whose states contain the shared array ${\tt items}$ and the shared counter ${\tt back}$ together with a valuation for the local variables ${\tt i}$, ${\tt x}$, and ${\tt range}$.
+A transition is either a call or a return action, or a statement in one of the two methods ${\tt enq}$ or ${\tt deq}$.
+A $\mathit{HWQ}$ state is related by $F_1$ to an $AbsQ$ state where $\mathsf{present}$ is true for all the enqueues whose argument is  stored in the array ${\tt items}$, and all the pending enqueues that have not yet written to the array ${\tt items}$ (and only for these enqueues).
+The order relation $\mathsf{before}$ between these enqueues is defined as follows:
 \vspace{-2mm}
 \begin{itemize}
-	\item[(a)] pending enqueues are maximal, i.e., for every two enqueues $k$ and $k'$ such that $k'$ is pending, we have that $k'\not< k$,
-	\item[(b)] $<$ is consistent with the order in which positions of {\tt items} have been reserved, i.e., for every two enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $k' \not< k$,
+	\item[(a)] pending enqueues are maximal, i.e., for every two enqueues $k$ and $k'$ such that $k'$ is pending, we have that $\neg \mathsf{before}(k',k)$, % $k'\not< k$,
+	\item[(b)] $\mathsf{before}$ is consistent with the order in which positions of ${\tt items}$ have been reserved, i.e., for every two enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $\neg \mathsf{before}(k',k)$, %$k' \not< k$,
 	\item[(c)] an enqueue which has reserved a position $i$ %and executed only the first statement 
@@
 	\noindent
-	we have that $k \not< k'$.
-The predicate $k_d\atCP {\tt if}\text{-}{\tt inc}$ holds when the dequeue $k_d$ is at a control point after a {\tt swap} returning {\tt null} and before the increment of {\tt i}.
+	we have that $\neg \mathsf{before}(k,k')$.
+The predicate $k_d\atCP {\tt if}\text{-}{\tt inc}$ holds when the dequeue $k_d$ is at a control point after a ${\tt swap}$ returning ${\tt null}$ and before the increment of ${\tt i}$.
 \vspace{-2mm}
@@
 \noindent
-An enqueue is labeled by $(d,{\tt PEND})$ where $d$ is the input value if it's pending and by  $(d,{\tt COMP})$, otherwise.
-Also, for every dequeue operation $k$ such that ${\tt x}(k)=d\neq {\tt null}$, we have that $rv(k)=d$.
+We have that $\mathsf{pending}(k)$ is ${\tt true}$ whenever $k$ is a pending enqueue, and $\mathsf{arg}(k)=d$ whenever the argument of the enqueue $k$ is $d$. 
+%An enqueue is labeled by $(d,{\tt PEND})$ where $d$ is the input value if it's pending and by  $(d,{\tt COMP})$, otherwise. 
+Also, for every dequeue operation $k$ such that ${\tt x}(k)=d\neq {\tt null}$, we have that ${\tt y}(k)=d$ (recall that ${\tt y}$ is a local variable of the dequeue method in $AbsQ$).
 
-We show that $\mathit{fs}_1$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in\mathit{fs}_1$. 
+TODO I FIXED THE NOTATIONS ONLY UNTIL THIS POINT
+
+We show that $F_1$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in F_1$. 
 We omit discussing the trivial case of transitions labeled by call and return actions which are simulated by similar transitions of $AbsQ$ (for the return a dequeue operation $k$, we use the equality between the local variable ${\tt x}(k)$ in $s$ and the component $rv(k)$ in $t$).
@@
 
-We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an \emph{empty} sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in\mathit{fs}_1$, then $(s',t)\in\mathit{fs}_1$ for each $AbsQ$ state $t$.
-Essentially, this consists in proving the following property, called \emph{monotonicity}: the set of possible orders $<$ associated by $\mathit{fs}_1$ to $s'$ doesn't exclude any order $<$ associated to $s$.
+We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an \emph{empty} sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in F_1$, then $(s',t)\in F_1$ for each $AbsQ$ state $t$.
+Essentially, this consists in proving the following property, called \emph{monotonicity}: the set of possible orders $<$ associated by $F_1$ to $s'$ doesn't exclude any order $<$ associated to $s$.
 %Essentially, this boils down to showing that the constraints over $<$ in the definition of $\mathit{fs}$ are an invariant for these steps.
@@
 The execution of {\tt swap} returning {\tt null} in a dequeue $k_d$ enables new instantiations $(k,k',k_d)$ in $s'$, thus adding potentially new constraints $k\not< k'$.
-We show that these instantiations are however vacuous because $k$ must be pending in $s$ and thus maximal in every order $<$ associated by $\mathit{fs}_1$ to $s$.
+We show that these instantiations are however vacuous because $k$ must be pending in $s$ and thus maximal in every order $<$ associated by $F_1$ to $s$.
 Let $k$ and $k'$ be two enqueues such that together with the dequeue $k_d$ they satisfy the property (\ref{eq:inst}) in $s'$ but not in $s$. 

### 2017-05-04 draftCAV/refProofs.tex
some minor issues

@@
 \vspace{-1.5mm}
-Library refinement is the instance of a more general notion of refinement between LTSs
+Library refinement is an instance of a more general notion of refinement between LTSs,
 which for some alphabet $\Gamma$ of \emph{observable actions} is defined as the inclusion of sets of 
@@
 that record events in the implementation, and to the existence of \emph{forward simulations} provided that the right-hand side
-LTS $A_2$ is $\Gamma$-deterministic~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}. 
+LTS, $A_2$, is $\Gamma$-deterministic~\cite{DBLP:journals/tcs/AbadiL91,DBLP:journals/iandc/LynchV95}. 
 We focus on proofs based on forward simulations because they are easier to automatize.
@@
 
-Let $A_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $A_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two LTSs over $\Sigma_1$ and $\Sigma_2$, respectively, such that $\Gamma\subseteq \Sigma_1\cap\Sigma_2$.
+Let $A_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $A_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two LTSs over $\Sigma_1$ and $\Sigma_2$, respectively, and $\Gamma$ an alphabet, such that $\Gamma\subseteq \Sigma_1\cap\Sigma_2$.
 
@@
 We define a notion of \emph{forward} simulation that can be used to prove $\Gamma$-refinement %We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement.
-For a relation $R\subseteq A\times B$, $R[X]$ is the set of elements related by $R$ to elements of $X$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.
+%For a relation $R\subseteq A\times B$, $R[X]$ is the set of elements related by $R$ to elements of $X$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.
 

### 2017-05-04 draftCAV/refProofs.tex
small modifs

@@
 an operation in an execution is a point in time where the operation is conceptually effectuated; given the linearization points of 
-each operation, the linearization of a concurrent history is the sequential history which takes operations in order of their linearization points.
-For some libraries, the linearization points correspond to a fixed set of actions. For instance, in the case of atomic libraries  
+each operation, the linearization of a concurrent history is the sequential history which takes operations in the order of their linearization points.
+For some libraries, the linearization points of all the invocations of a method $m$ correspond to the execution of a fixed statement in $m$'s body. For instance, in the case of atomic libraries  
 where method bodies are guarded with a global-lock acquisition, the linearization point of every method invocation corresponds to the execution 

### 2017-05-04 draftCAV/stack.tex
small modifs

@@
 \vspace{-1.5mm}
-While the abstract queue in Section~\ref{sec:queues} can be adapted to stacks (the linearization point $lin(pop,d,k)$ with $d\neq{\tt EMPTY}$ is enabled when $k$ is added by a push which is maximal in the happens-before order stored in the state), it can't simulate (through forward simulations) existing stack implementations like the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) where the linearization points of the pop operations are not fixed.
+The abstract implementation in Section~\ref{sec:queues} can be adapted to stacks, the main modification being that the linearization point $lin(pop,d,k)$ with $d\neq{\tt EMPTY}$ is enabled when $k$ is added by a push which is maximal in the happens-before order stored in the state.
+However, there are stack implementations, e.g., Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short), which cannot be proved correct using forward simulations to this abstract implementation because the linearization points of the pop operations are not fixed.
+%this stack implementation cannot simulate (through forward simulations) existing stack implementations like the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) where the linearization points of the pop operations are not fixed.
 Exploiting particular properties of the stack semantics, we refine the ideas used in $AbsQ$ and define 
 a new abstract implementation for stacks, denoted as $AbsS$, which is able to simulate such implementations.
-Forward simulations to $AbsS$ are complete for proving the correctness of stack implementations provided that the point in time where the return value of a pop operation is determined, called \emph{commit point}, corresponds to a fixed action.
+Forward simulations to $AbsS$ are complete for proving the correctness of stack implementations provided that the point in time where the return value of a pop operation is determined, called \emph{commit point}, corresponds to a fixed statement in the pop method.
 % provided that the point in time where the return value of a pop operation is determined corresponds to a fixed action.

### 2017-05-04 draftCAV/queue.tex
small modifs

@@
 For some queue implementations, e.g., the Herlihy\&Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), there exists no forward simulation to $AbsQ_0$ although they are a refinement of $AbsQ_0$.
-The main reason is that the enqueue methods don't have a \emph{fixed} linearization point.
+The main reason is that the linearization points of the enqueue operations are not fixed, they depend on future dequeue invocations.
 In this section, we propose a new abstract implementation for queues, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence.
@@
 For readability, we define $AbsQ$ as an abstract state machine, which is given in Fig.~\ref{fig:transitions:AbsQ}.
-The states of $AbsQ$ consist of several boolean predicates indicating whether the value added by an enqueue has not been removed yet (the predicate $\mathsf{present})$, whether an enqueue is pending (the predicate $\mathsf{pending}$), the happens-before order (the predicate $\mathsf{before}$), and a function giving the value added by an enqueue (the function $\mathsf{arg}$).
+The shared state of $AbsQ$ consists of several boolean predicates indicating whether the value added by an enqueue has not been removed yet (the predicate $\mathsf{present})$, whether an enqueue is pending (the predicate $\mathsf{pending}$), the happens-before order (the predicate $\mathsf{before}$), and a function giving the value added by an enqueue (the function $\mathsf{arg}$).
 %the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).

### 2017-05-04 draftCAV/draft.tex
small modifs

@@
 
-\title{Proving linearizability using forward simulations}
+\title{Proving linearizability using forward simulations\thanks{An extended version of this paper including the missing proofs can be found at \cite{extended}.}}
 \author{Ahmed Bouajjani\inst{1}\and Michael Emmi\inst{2}\and Constantin Enea\inst{1} \and Suha Orhun Mutluergil\inst{3}}
@@
 
+\section{Acknowledgements}
+
+This work is supported in part by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation program (grant agreement No 678177).
+
 \bibliographystyle{abbrvnat}

### 2017-05-04 draftCAV/queue.tex
small modifs

@@
 The shared state of $AbsQ$ consists of several boolean predicates indicating whether the value added by an enqueue has not been removed yet (the predicate $\mathsf{present})$, whether an enqueue is pending (the predicate $\mathsf{pending}$), the happens-before order (the predicate $\mathsf{before}$), and a function giving the value added by an enqueue (the function $\mathsf{arg}$).
+In the initial state, the domain of every function (predicate) is empty.
 %the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).
@@
 \begin{figure}[t]
-\hspace{-8mm}
+\hspace{-10mm}
 \begin{minipage}[t]{7.5cm}
@@
 		& \text{\underline{Shared State}} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \mathsf{Operations} \rightarrow \mathsf{Bool} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \mathsf{Operations} \rightarrow \mathsf{Bool} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \mathsf{Operations} \times \mathsf{Operations} \rightarrow \mathsf{Bool} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
 %		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
@@
 \begin{lstlisting}
-void enq($\mathsf{Values}$ x) {
+void enq($\<Vals>$ x) {
   atomic rule inv(enq,x,k = new);
@@
 
-$\mathsf{Values}$ deq() {
+$\<Vals>$ deq() {
   atomic rule inv(deq,k = new);
@@
   else 
-    choose k1 $\in$ $\mathsf{Operations}$ with $\mathsf{present}$(k1)$\land\neg \exists k'.\ \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$;
+    choose k1 $\in$ $\<Ops>$ with $\mathsf{present}$(k1)$\land\neg \exists k'.\ \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$;
     $\mathsf{present}$(k1) = false;
@@
 We describe a forward simulation $F_1$ from $\mathit{HWQ}$ to $AbsQ$.
-The description of $\mathit{HWQ}$ in Fig.~\ref{fig:HerlihyWing} defines an LTS whose states contain the shared array ${\tt items}$ and the shared counter ${\tt back}$ together with a valuation for the local variables ${\tt i}$, ${\tt x}$, and ${\tt range}$.
+The description of $\mathit{HWQ}$ in Fig.~\ref{fig:HerlihyWing} defines an LTS whose states contain the shared array ${\tt items}$ and the shared counter ${\tt back}$ together with a valuation for the local variables ${\tt i}$, ${\tt x}$, and ${\tt range}$, and the control location of each operation.
 A transition is either a call or a return action, or a statement in one of the two methods ${\tt enq}$ or ${\tt deq}$.

### 2017-05-04 draftCAV/stack.tex
small modifs

@@
 \vspace{-1mm}
-We define an abstract stack $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the pushes whose value has not been yet removed.
-Pops are treated differently since the commit points are not necessarily linearization points, intuitively, a pop can be linearized before its commit.
-Each pop operation starts by taking a snapshot of the greatest completed push operations in the happens-before order, and continuously tracks the push operations which are overlapping with it.
-The commit point $com(pop,d,k)$ with $d\neq {\tt EMPTY}$ is enabled only if $d$ was added by one of the push operations in the initial snapshot, or by a push happening earlier when all the values from the initial snapshot have been removed, or by one of the push operations that overlaps with pop $k$.
-The commit point $com(pop,{\tt EMPTY},k)$ is enabled only if all the values added by push operations ending before $k$ started have been removed.
+We define an abstract stack $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the pushes whose value has not yet been  removed by a matching pop.
+Pop operations are treated differently since the commit points are not necessarily linearization points.
+Intuitively, a pop can be linearized before its commit point.
+Each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before, more precisely, which don't happen before another completed push operation.
+Also, the library maintains the set of push operations overlapping with each pop operation.
+The commit point $com(pop,d,k)$ with $d\neq {\tt EMPTY}$ is enabled if either $d$ was added by one of the push operations in the initial snapshot, or by a push happening earlier when arguments of pushes from the initial snapshot have been removed, or by one of the push operations that overlaps with pop $k$.
+The commit point $com(pop,{\tt EMPTY},k)$ is enabled if all the values added by push operations happening before $k$ have been removed.
 The effect of the commit points is explained below through examples.
@@
 
-Formally, the states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $<$ is a strict partial order over the set $O$ of operation identifiers, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a pending/completed flag, $rv:\<Ops> ~> \<Vals>$ records the return value of a pending pop fixed at its commit point, $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A_2$) or pop operation ($R_1,R_2,R_3$), $be:\<Ops> ~> 2^O$ records the greatest completed push operations before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
-%The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.
-All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsS}.
-
-The transition rules which don't correspond to commit point actions are similar to those for $AbsQ$.
-The rule {\sc com-pop1} for $com(pop,d,k)$ is enabled only if there exists a push $k'$ which added value $d$ and which belongs to $be(k)$ or $ov(k)$.
-When enabled, the push $k'$ is removed from the set $O$ (and the order $<$) and for every other pop $k_1$ such that $k'$ belongs to $be(k_1)$, $k'$ is replaced in $be(k_1)$ by its predecessors which are followed exclusively by pushes overlapping with $k_1$ (these predecessors become maximal closed pushes once $k'$ is removed).
-Also, $rv(k)$ is set to $d$.
-The rule {\sc com-pop1} for $com(pop,{\tt EMPTY},k)$ is enabled only if $be(k)$ is empty (i.e., all the values added by pushes ending before $k$, if any, have been removed).
-Then, $rv(k)$ is set to ${\tt EMPTY}$.
-
 \begin{figure} [t]
-\vspace{-2mm}
+%\vspace{-2mm}
+\hspace{-15mm}
 \begin{minipage}[t]{7.5cm}
@@
 		& \text{\underline{Shared State}} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \mathsf{Operations} \rightarrow \mathsf{Bool} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \mathsf{Operations} \rightarrow \mathsf{Bool} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \mathsf{Operations} \times \mathsf{Operations} \rightarrow \mathsf{Bool} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{maxAtInvoc} : \mathsf{Operations} \rightarrow 2^\mathsf{Operations} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{overlap} : \mathsf{Operations} \rightarrow 2^\mathsf{Operations} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \mathsf{Values} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{maxAtInvoc} : \<Ops> \rightarrow 2^{\<Ops>} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{overlap} : \<Ops> \rightarrow 2^{\<Ops>} \\
 %		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
@@
 \end{minipage}
-\begin{minipage}[t]{7.5cm}
+\begin{minipage}[t]{7.3cm}
 \begin{lstlisting}
-void push($\mathsf{Values}$ x) {
+void push($\<Vals>$ x) {
   atomic rule inv(push,x,k = new);
@@
 
-$\mathsf{Values}$ pop() {
+$\<Vals>$ pop() {
   atomic rule inv(pop,k = new);
@@
 \vspace{-5mm}
-\begin{minipage}[t]{6.5cm}
+\begin{minipage}[t]{6.2cm}
 \begin{lstlisting}
@@
   forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1)$\land$
-                 $\neg \exists k'.\ \mathsf{present}(k')\land \neg \mathsf{pending}(k')\land$ 
+                 $\neg \exists k'.\ \mathsf{present}(k')\land \neg \mathsf{pending}(k')\,\land$ 
                  $\mathsf{before}$(k1,$k'$) do
@@
   else 
-    choose k1 $\in$ $\mathsf{Operations}$ with $\mathsf{present}$(k1)$\land$ k1$\in\mathsf{maxAtInvoc}$(k)$\cup\,\mathsf{overlap}$(k) do
+    choose k1 $\in\<Ops>$ with $\mathsf{present}$(k1)$\land$ k1$\in\mathsf{maxAtInvoc}$(k)$\cup\,\mathsf{overlap}$(k) do
     $\mathsf{present}$(k1) = false;
@@
 
-Let $AbsS_0$ be the standard abstract implementation of a stack (where elements are stored in a sequence; push, resp., pop operations add, resp., remove, an element from the beginning of the sequence in one atomic step).
+The description of $AbsS$ as an abstract state machine is given in Fig.~\ref{fig:transitions:AbsS}.
+Compared to the queue implementation $AbsQ$, the shared state contains two more functions $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$.
+For each pop operation $k$, $\mathsf{maxAtInvoc}(k)$ records the set of completed push operations which were maximal in the happens-before (defined by $\mathsf{before}$) when pop $k$ was invoked, or happening earlier provided that the values of all the pushes happening later than one of these maximal ones have been removed.
+Also, for each pop operation $k$, $\mathsf{overlap}(k)$ contains the push operations overlapping with pop $k$.
+%Formally, the states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $<$ is a strict partial order over the set $O$ of operation identifiers, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a pending/completed flag, $rv:\<Ops> ~> \<Vals>$ records the return value of a pending pop fixed at its commit point, $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A_2$) or pop operation ($R_1,R_2,R_3$), $be:\<Ops> ~> 2^O$ records the greatest completed push operations before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
+%%The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.
+%All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsS}.
+In the initial state, the domain of every function (predicate) is empty.
+The push methods are similar to the enqueue methods in $AbsQ$.
+The only difference is that every newly invoked push operation {\tt k} is added to the set $\mathsf{overlap}(${\tt k1}$)$ for all pop operations {\tt k1} (since {\tt k} overlaps with all the currently pending pops).
+The rule {\tt inv(pop,k)}, marking the invocation of a pop, sets $\mathsf{maxAtInvoc}(${\tt k}$)$ and $\mathsf{overlap}(${\tt k}$)$ as explained above.
+The rule {\tt com(pop,y,k)} results in setting the return value {\tt y} to {\tt EMPTY} if the set $\mathsf{maxAtInvoc}(${\tt k}$)$ is empty (otherwise, there would be push operations happening before pop {\tt k} which makes the answer {\tt EMPTY} incorrect).
+It results in setting the return value to a value $d\neq\,${\tt EMPTY} when $d$ was added by a push {\tt k1} which belongs to $\mathsf{maxAtInvoc}(${\tt k}$)$ or $\mathsf{overlap}(${\tt k}$)$.
+In this latter case, the rule {\tt com(pop,y,k)} may also update $\mathsf{maxAtInvoc}(${\tt k2}$)$ for other pending pops {\tt k2}.
+More precisely, whenever $\mathsf{maxAtInvoc}(${\tt k2}$)$ contains the push {\tt k1}, the latter is replaced by the immediate predecessors of {\tt k1} (according to $\mathsf{before}$) that are followed exclusively by pushes overlapping with {\tt k2}.
+%The transition rules which don't correspond to commit point actions are similar to those for $AbsQ$.
+The rule {\sc com-pop1} for $com(pop,d,k)$ is enabled only if there exists a push $k'$ which added value $d$ and which belongs to $be(k)$ or $ov(k)$.
+When enabled, the push $k'$ is removed from the set $O$ (and the order $<$) and for every other pop $k_1$ such that $k'$ belongs to $be(k_1)$, $k'$ is replaced in $be(k_1)$ by its predecessors which are followed exclusively by pushes overlapping with $k_1$ (these predecessors become maximal closed pushes once $k'$ is removed).
+Also, $rv(k)$ is set to $d$.
+The rule {\sc com-pop1} for $com(pop,{\tt EMPTY},k)$ is enabled only if $be(k)$ is empty (i.e., all the values added by pushes ending before $k$, if any, have been removed).
+Then, $rv(k)$ is set to ${\tt EMPTY}$.
+
+The abstract state machine in Fig.~\ref{fig:transitions:AbsS} defines an LTS over the alphabet $\Sigma=C\cup R\cup Com(pop)$.
+As in the case of $AbsQ$, we assume that the transition corresponding to the invocation, resp., the return, of a method, is done atomically with the fist, resp., the last, macro rule in its body.
+These transitions are labeled as expected with call and return actions (borrowing the operation identifier which occurs as argument of the macro rule), and the transitions corresponding to the rule {\tt com(pop,y,k)} are labeled by $com(pop,d,k)$.
+
+Let $AbsS_0$ be the standard abstract implementation of a stack where elements are stored in a sequence, push and pop operations adding and removing an element from the beginning of the sequence in one atomic step, respectively.
 For $\<Methods>=\{push,pop\}$, the alphabet of $AbsS_0$ is $C\cup R\cup Lin$.
-The following result states that the library $AbsS$ has exactly the same set of histories as $AbsS_0$ (see Appendix~\ref{app:absImplStack} for a proof).
+The following result states that the library $AbsS$ has exactly the same set of histories as $AbsS_0$.
 

### 2017-05-04 draftCAV/stack.tex
small modifs

@@
 \end{wrapfigure}
-We explain the meaning of the commit points on a simplified version of the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) given in Figure~\ref{fig:TimeStamped}.
+We explain the meaning of the commit points on a simplified version of the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) given in Fig.~\ref{fig:TimeStamped}.
 This  implementation maintains an array of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
@@
 
-\begin{wrapfigure}{l}{6.8cm}
-\vspace{-6mm}
-\includegraphics[width=7cm]{fig-stack.pdf}
+%\begin{wrapfigure}{l}{7.8cm}
+%\vspace{-6mm}
+\begin{figure}[t]
+\centering
+\includegraphics[width=10cm]{fig-stack.pdf}
 %
@@
 %\includegraphics[width=7cm]{fig-queue2.pdf}
-\vspace{-8mm}
+\vspace{-2mm}
 \caption{Simulating stack histories with $AbsS$.}
@@
 \vspace{-6.5mm}
-\end{wrapfigure}
-Figure~\ref{fig:stackSim} pictures two executions of $AbsS$ for two extended histories (that include pop commit points).
+\end{figure}
+
+Fig.~\ref{fig:stackSim} pictures two executions of $AbsS$ for two extended histories (that include pop commit points).
 For readability, we give the state of $AbsS$ only after several execution prefixes delimited at the right by a dotted line.

### 2017-05-04 draftCAV/queue.tex
small modifs

@@
 
-\begin{wrapfigure}{l}{6.7cm}
+\begin{wrapfigure}{l}{7.2cm}
 \vspace{-6mm}
-\includegraphics[width=6.8cm]{fig-queue12.pdf}
+\includegraphics[width=7.3cm]{fig-queue12.pdf}
 %
@@
 \end{wrapfigure}
-Figure~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
+Fig.~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
 The state of $AbsQ$ after each action is pictured as a graph below the action.
@@
 It is obtained from Theorem~\ref{th:absImplQueue} and Theorem~\ref{th:forSim} using the fact that the alphabet of $AbsQ$ is exactly $C\cup R\cup Lin(deq)$ and $AbsQ$ is deterministic.
+The determinism of $AbsQ$ relies on the assumption that every value is added at most once.
+Without this assumption, $AbsQ$ may reach a state with two enqueues adding the same value being both minimal in the happens-before.
+A transition corresponding to the linearization point of a dequeue from this state can remove any of these two enqueues leading to two different states.
+Therefore, $AbsQ$ becomes non-deterministic.
+Note that this is independent of the fact that $AbsQ$ manipulates  operation identifiers.
 
@@
 We write ${\tt i}_s(k)$ for the value of the variable {\tt i} of operation $k$ in state $s$. 
-We have that ${\tt i}_{s'}(k) = {\tt i}_{s'}(k_d) \leq {\tt i}_{s'}(k')$ and ${\tt items}[{\tt i}_{s'}(k_d)]={\tt null}$. The latter implies that the enqueue $k$ didn't executed
+We have that ${\tt i}_{s'}(k) = {\tt i}_{s'}(k_d) \leq {\tt i}_{s'}(k')$ and ${\tt items}[{\tt i}_{s'}(k_d)]={\tt null}$. The latter implies that the enqueue $k$ didn't execute
 the second statement (since the position it reserved is still {\tt null}) and it is pending in $s$.
@@
 Also, applying item (b) with $k=k_e$ we get the same for every enqueue that reserved a bigger position.
-An enqueue that didn't reserved a position is by definition maximal in $<$ and therefore, not a predecessor of $k_e$.
+An enqueue that didn't reserve a position is by definition maximal in $<$ and therefore, not a predecessor of $k_e$.
 Then, the state $t'$ obtained from $t$ through a $lin(deq,d,k)$ transition is related to $s'$ because (1) the value added by $k_e$ is not anymore present in {\tt items} which implies that $k_e$ doesn't occur in any $AbsQ$ state related to $s'$, and (2) the value of ${\tt x}(k)$ is set to $d\neq {\tt null}$ which implies that $rv(k)$ is set to $d$ in every $AbsQ$ state related to $s'$.

### 2017-05-05 draftCAV/queue.tex
small modifs

@@
 Initially, all the positions in the array are {\tt null} and {\tt back} is 0.
-An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueues) and then, stores the input value {\tt x} at this position.
+An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueues) and then, stores the argument {\tt x} at this position.
 The dequeue method traverses the array {\tt items} starting from the beginning and atomically swaps {\tt null} with the encountered value.
@@
 We also need to update the rules in figure.}
-Informally, $AbsQ$ records the set of enqueue operations, whose input value has not yet been removed by a matching dequeue operation.
+Informally, $AbsQ$ records the set of enqueue operations, whose argument has not yet been removed by a matching dequeue operation.
 In addition, it records the happens-before order between those enqueue operations: this is a partial order ordering an enqueue $k_1$ before another enqueue $k_2$ iff $k_1$ returned before $k_2$ was invoked.
@@
 The nodes of this graph represent enqueue operations and the edges happens-before constraints.
-Each node is labeled by a value (the input of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed.
-For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with input $y$.
+Each node is labeled by a value (the argument of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed.
+For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with argument $y$.
 Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.
@@
 It can set the return value {\tt y} either to {\tt EMPTY}, or to a value {\tt d} $\neq$ {\tt EMPTY}, provided that {\tt d} has been added by an enqueue {\tt k1} which is minimal in the current happens-before.
-In the latter case, $\mathsf{present}${\tt (k1)} is set to {\tt false} to indicate that {\tt k1}'s input has been removed.
+In the latter case, $\mathsf{present}${\tt (k1)} is set to {\tt false} to indicate that {\tt k1}'s argument has been removed.
 When there exists at least one completed enqueue whose value has not been removed, the rule {\tt lin(deq,y,k)} sets {\tt y} to the value added by a minimal enqueue (w.r.t.
@@
 
-The abstract state machine in Fig.~\ref{fig:transitions:AbsQ} defines an LTS over an alphabet $\Sigma$ such that $C\cup R\cup Lin(deq)\subseteq \Sigma$.
-The transitions corresponding to the rule {\tt lin(deq,y,k)} are labeled by $lin(deq,y,k)$ actions.
+The abstract state machine in Fig.~\ref{fig:transitions:AbsQ} defines an LTS over the alphabet $C\cup R\cup Lin(deq)$.
+The transitions corresponding to the rule {\tt lin(deq,y,k)} are labeled by  actions $lin(deq,y,k)$.
 Also, we assume that the transition corresponding to the invocation of a method is done atomically with the first step of the same invocation (the macro rule {\tt inv(...)}), and similarly for transitions corresponding to returning from a method, they are done atomically with the last step (the macro rule {\tt ret(...)}).
 They are labeled as expected with call/return actions (borrowing the operation identifier which occurs as argument of the macro rules).
-It can be easily proved that this assumption doesn't modify the set of traces projected on $C\cup R\cup Lin(deq)$.
+It can be easily proved that this assumption doesn't modify the set of traces projected on $C\cup R\cup Lin(deq)$ (compared to an LTS where these steps are not atomic).
 

### 2017-05-05 draftCAV/stack.tex
small modifs

@@
 The first state we give is reached after the call of pop with identifier $3$.
-This shows the effect of a pop invocation: the greatest completed pushes according to the current happens-before (here, the push with identifier $1$) are marked as $be(3)$ (from ``before'' operation 3), and the pending pushes are marked as $ov(3)$ (from ``overlapping'' with operation 3).
-As a side remark, any other push operation that starts after pop $3$ would be also marked as $ov(3)$.
-The commit point $com(pop,x,3)$ (pictured with a red circle) is enabled because $x$ was added by a push marked as $be(3)$.
+This shows the effect of a pop invocation: the greatest completed pushes according to the current happens-before (here, the push with identifier $1$) are marked as $\mathsf{maxAtInvoc}(3)$, and the pending pushes are marked as $\mathsf{overlap}(3)$.
+As a side remark, any other push operation that starts after pop $3$ would be also marked as $\mathsf{overlap}(3)$.
+The commit point $com(pop,x,3)$ (pictured with a red circle) is enabled because $x$ was added by a push marked as $\mathsf{maxAtInvoc}(3)$.
 The effect of the commit point is that push $1$ is removed from the state (the execution on the bottom shows a more complicated case).
-For the second pop, the commit point $com(pop,y,4)$ is enabled because $y$ was added by a push marked as $ov(4)$.
-The execution on the bottom shows an example where the marking $be(k)$ for some pop $k$ is updated at commit points.
-The pushes $3$ and $4$ are marked as $be(5)$ and $be(6)$
+For the second pop, the commit point $com(pop,y,4)$ is enabled because $y$ was added by a push marked as $\mathsf{overlap}(4)$.
+The execution on the bottom shows an example where the marking $\mathsf{maxAtInvoc}(k)$ for some pop $k$ is updated at commit points.
+The pushes $3$ and $4$ are marked as $\mathsf{maxAtInvoc}(5)$ and $\mathsf{maxAtInvoc}(6)$
 when the pops $5$ and $6$ start.
-Then, $com(pop,t,5)$ is enabled since $t$ was added by $push(t,4)$ which is marked as $be(5)$.
+Then, $com(pop,t,5)$ is enabled since $t$ was added by $push(t,4)$ which is marked as $\mathsf{maxAtInvoc}(5)$.
 Besides removing $push(t,4)$, the commit point produces a state where a pop committing later, e.g., pop $6$, can remove $y$ which was added by a predecessor of $push(t,4)$ in the happens-before ($y$ could become the top of the stack when $t$ is removed).
 This history is valid because $push(y,2)$ can be linearized after $push(x,1)$ and $push(z,3)$.
-Thus, push 2, a predecessor of the push which is removed, is marked as $be(6)$.
-Push $1$ which is also a predecessor of the removed push is not marked as $be(6)$ because it happens before another push, i.e., push 3, which is already marked as $be(6)$ (the value added by push 3 should be removed before the value added by push 1 could become the top of the stack).
+Thus, push 2, a predecessor of the push which is removed, is marked as $\mathsf{maxAtInvoc}(6)$.
+Push $1$ which is also a predecessor of the removed push is not marked as $\mathsf{maxAtInvoc}(6)$ because it happens before another push, i.e., push 3, which is already marked as $\mathsf{maxAtInvoc}(6)$ (the value added by push 3 should be removed before the value added by push 1 could become the top of the stack).
 
@@
 		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \mathsf{Values} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
 		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
@@
   choose a $\in$ {true, false};
-  if ((a || $\forall k'.\ \mathsf{present}(k')$=false) && $\mathsf{maxAtInvoc}$(k)=$\,\emptyset$) 
+  if ((a $\vee$ $\forall k'.\ \mathsf{present}(k')$=false) $\land$ $\mathsf{maxAtInvoc}$(k)=$\,\emptyset$) 
     y = EMPTY;
   else 
-    choose k1 $\in\<Ops>$ with $\mathsf{present}$(k1)$\land$ k1$\in\mathsf{maxAtInvoc}$(k)$\cup\,\mathsf{overlap}$(k) do
+    choose k1 $\in\<Ops>$ with $\mathsf{present}$(k1)$\land$ k1$\in\mathsf{maxAtInvoc}$(k)$\cup\,\mathsf{overlap}$(k);
     $\mathsf{present}$(k1) = false;
@@
 Compared to the queue implementation $AbsQ$, the shared state contains two more functions $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$.
-For each pop operation $k$, $\mathsf{maxAtInvoc}(k)$ records the set of completed push operations which were maximal in the happens-before (defined by $\mathsf{before}$) when pop $k$ was invoked, or happening earlier provided that the values of all the pushes happening later than one of these maximal ones have been removed.
+For each pop operation $k$, $\mathsf{maxAtInvoc}(k)$ records the set of completed push operations which were maximal in the happens-before (defined by $\mathsf{before}$) when pop $k$ was invoked, or happening earlier provided that the values of all the pushes happening later than one of these maximal ones and before pop $k$ have been removed.
 Also, for each pop operation $k$, $\mathsf{overlap}(k)$ contains the push operations overlapping with pop $k$.
@@
 It results in setting the return value to a value $d\neq\,${\tt EMPTY} when $d$ was added by a push {\tt k1} which belongs to $\mathsf{maxAtInvoc}(${\tt k}$)$ or $\mathsf{overlap}(${\tt k}$)$.
-In this latter case, the rule {\tt com(pop,y,k)} may also update $\mathsf{maxAtInvoc}(${\tt k2}$)$ for other pending pops {\tt k2}.
+In this case, the rule {\tt com(pop,y,k)} may also update $\mathsf{maxAtInvoc}(${\tt k2}$)$ for other pending pops {\tt k2}.
 More precisely, whenever $\mathsf{maxAtInvoc}(${\tt k2}$)$ contains the push {\tt k1}, the latter is replaced by the immediate predecessors of {\tt k1} (according to $\mathsf{before}$) that are followed exclusively by pushes overlapping with {\tt k2}.
@@
 
-The abstract state machine in Fig.~\ref{fig:transitions:AbsS} defines an LTS over the alphabet $\Sigma=C\cup R\cup Com(pop)$.
+The abstract state machine in Fig.~\ref{fig:transitions:AbsS} defines an LTS over the alphabet $C\cup R\cup Com(pop)$.
 As in the case of $AbsQ$, we assume that the transition corresponding to the invocation, resp., the return, of a method, is done atomically with the fist, resp., the last, macro rule in its body.

### 2017-05-05 draftCAV/queue.tex
Changed HWQ Proof (4.3.)
Added intuition to F_1.

@@
 A transition is either a call or a return action, or a statement in one of the two methods ${\tt enq}$ or ${\tt deq}$.
-A $\mathit{HWQ}$ state is related by $F_1$ to an $AbsQ$ state where $\mathsf{present}$ is true for all the enqueues whose argument is  stored in the array ${\tt items}$, and all the pending enqueues that have not yet written to the array ${\tt items}$ (and only for these enqueues).
+
+An enqueue operation in a $\mathit{HWQ}$ state is pending and present if and only if its argument is stored in the array ${\tt items}$ or it has not yet written to the array  ${\tt items}$.
+Those operations should be $\mathsf{pending}$ and $\mathsf{present}$ in the related $AbsQ$ states.
+In addition to this, the forward simulation $F_1$ imposes following restrictions:
+%A $\mathit{HWQ}$ state is related by $F_1$ to an $AbsQ$ state where $\mathsf{present}$ is true for all the enqueues whose argument is  stored in the array ${\tt items}$, and all the pending enqueues that have not yet written to the array ${\tt items}$ (and only for these enqueues).
 The order relation $\mathsf{before}$ between these enqueues is defined as follows:
@@
 \begin{itemize}
-	\item[(a)] pending enqueues are maximal, i.e., for every two enqueues $k$ and $k'$ such that $k'$ is pending, we have that $\neg \mathsf{before}(k',k)$, % $k'\not< k$,
-	\item[(b)] $\mathsf{before}$ is consistent with the order in which positions of ${\tt items}$ have been reserved, i.e., for every two enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $\neg \mathsf{before}(k',k)$, %$k' \not< k$,
+	\item[(a)] pending and present enqueues are maximal, i.e., for every two $\mathsf{present}$ enqueues $k$ and $k'$ such that $k'$ is $\mathsf{pending}$, we have that $\neg \mathsf{before}(k',k)$, % $k'\not< k$,
+	\item[(b)] $\mathsf{before}$ is consistent with the order in which positions of ${\tt items}$ have been reserved, i.e., for every two present enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $\neg \mathsf{before}(k',k)$, %$k' \not< k$,
 	\item[(c)] an enqueue which has reserved a position $i$ %and executed only the first statement 
-	can't be ordered before another enqueue that has reserved a position $j \geq i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal, i.e., for every two enqueues $k$ and $k'$, and a dequeue $k_d$, such that
+	can't be ordered before another enqueue that has reserved a position $j \geq i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal, i.e., for every two present enqueues $k$ and $k'$, and a dequeue $k_d$, such that
 
@@
 	\hspace{-8mm}
-	{\tt x}(k_d)={\tt null} \land {\tt i}(k') \leq {\tt range}(k_d) \land {\tt i}(k) \leq {\tt i}(k_d) \leq {\tt i}(k')
-	 \land ({\tt i}(k) = {\tt i}(k_d) => k_d\atCP {\tt if}\text{-}{\tt inc}) \label{eq:inst}
+	{\tt canRemove}(k_d,k) \land (i(k') < i(k_d) \lor (i(k')=i(k_d) \land {\tt afterSwapNull}(k_d)))
+\label{eq:inst}
 	\end{align}}
@@
 	we have that $\neg \mathsf{before}(k,k')$.
-The predicate $k_d\atCP {\tt if}\text{-}{\tt inc}$ holds when the dequeue $k_d$ is at a control point after a ${\tt swap}$ returning ${\tt null}$ and before the increment of ${\tt i}$.
+The predicate ${\tt canRemove}(k_d,k)$ holds when $k_d$ has currently visited a {\tt null} item in {\tt items} and the present enqueue index $i(k)$ is in the range of $(k_d)$ i.e., $\mathsf{present}(k) \land (x(k_d) = {\tt null} \land i(k) < {\tt range}(k_d) \land i(k_d) < i(k)) \lor (i(k_d) = i(k) \land {\tt beforeSwap}(k_d) \land {\tt items}[i(k)] != {\tt null})$.
+The predicates ${\tt afterSwapNull}(k_d)$ (resp., ${\tt beforeSwap}(k_d)$) holds when the dequeue $k_d$ is at the control point after a ${\tt swap}$ returning ${\tt null}$ (resp., before a ${\tt swap}$).
 \vspace{-2mm}
@@
 
-TODO I FIXED THE NOTATIONS ONLY UNTIL THIS POINT
+The restrictions of $F_1$ aim to ensure two important points:  (i) identifier of a pending and present enqueue method in the $\mathit{HWQ}$ state should be maximal according to $\mathsf{before}$ predicate in the related $AbsQ$ state and (ii) identifiera of a present enqueue method in $\mathit{HWQ}$ state of which data value is about to be removed by a dequeue operation should be minimal in the related $AbsQ$ state.
 
-We show that $F_1$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in F_1$. 
+The first point is ensured by the restriction (a).
+Present and pending enqueue identifiers in the $\mathit{HWQ}$ state are $\mathsf{pending}$ in the related $AbsQ$ state and $\mathsf{pending}$ enqueues are maximal in a valid $AbsQ$ state.
+
+Restrictions (b) and (c) ensure the point (ii).
+Consider a present enqueue $k$ that inserted its argument to ${\tt items}$ and  there exists a pending dequeue $k_d$ such that ${\tt canRemove}(k_d, k)$ and $k_d$ is just before its swap action at the reserved position of $k$ i.e., $i(k_d) = i(k)$.
+A pending enqueue operation cannot be ordered $\mathsf{before}$ $k$ since pending enqueues are maximal by (a).
+Regarding the completed and present enqueues $k'$, we consider two cases: $i(k') > i(k)$ and $i(k') < i(k)$.
+For the former case, the restriction (b) ensures $\neg \mathsf{before}(k',k)$ and for the latter case the restriction (c) ensures $\neg \mathsf{before}(k',k)$.
+Consequently, $k$ is a minimal element in $\mathsf{before}$ relation just before $k_d$ removes its data value.
+
+Next, we show that $F_1$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in F_1$. 
 We omit discussing the trivial case of transitions labeled by call and return actions which are simulated by similar transitions of $AbsQ$ (for the return a dequeue operation $k$, we use the equality between the local variable ${\tt x}(k)$ in $s$ and the component $rv(k)$ in $t$).
@@
 %\setlength{\parskip}{0pt}
-
 We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an \emph{empty} sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in F_1$, then $(s',t)\in F_1$ for each $AbsQ$ state $t$.
-Essentially, this consists in proving the following property, called \emph{monotonicity}: the set of possible orders $<$ associated by $F_1$ to $s'$ doesn't exclude any order $<$ associated to $s$.
+Essentially, this consists in proving the following property, called \emph{monotonicity}: the set of possible $\mathsf{before}$ relations associated by $F_1$ to $s'$ doesn't exclude any order $\mathsf{before}$ associated to $s$.
 %Essentially, this boils down to showing that the constraints over $<$ in the definition of $\mathit{fs}$ are an invariant for these steps.
 
-Concerning enqueues, let $s'$ be the state obtained from $s$ when a pending enqueue $k$ reserves an array position.
+Concerning enqueue rules, let $s'$ be the state obtained from $s$ when a pending enqueue $k$ reserves an array position.
 This enqueue must be maximal in both $t$ and any state $t'$ related to $s'$ (since it's pending).
 Moreover, there is no dequeue that can ``observe'' this position before restarting the array traversal.
-Therefore, item (c) in the definition of $<$ doesn't constrain the order between $k$ and some other enqueue neither in $s$ nor in $s'$.
+Therefore, item (c) in the definition of $F_1$ doesn't constrain the order between $k$ and some other enqueue neither in $s$ nor in $s'$.
 Since this transition doesn't affect the constraints on the order between enqueues different from $k$ (their local variables remain unchanged), monotonicity holds.
@@
 
-To prove monotonicity in the case of dequeue internal steps different from its linearization point, it is important to track the non-trivial instantiations of item (c) in the definition of $<$ over the two states $s$ and $s'$, i.e., the triples $(k,k',k_d)$ for which (\ref{eq:inst}) holds.
-Instantiations that are enabled only in $s'$ may in principle lead to a violation of monotonicity (since they restrict the orders $<$ associated to $s'$).
+To prove monotonicity in the case of dequeue internal steps different from its linearization point, it is important to track the non-trivial instantiations of item (c) in the definition of $\mathsf{before}$ over the two states $s$ and $s'$, i.e., the triples $(k,k',k_d)$ for which (\ref{eq:inst}) holds.
+Instantiations that are enabled only in $s'$ may in principle lead to a violation of monotonicity (since they restrict the orders $\mathsf{before}$ associated to $s'$).
 For the two steps that begin an array traversal, i.e., reading the index of the last used position and setting {\tt i} to $0$, there exist no  such new instantiations in $s'$ because the value of {\tt i} is either not set or $0$.
@@
 %
-The same is true for the increment of {\tt i} in a dequeue $k_d$ since the predicate $k_d\atCP {\tt if}\text{-}{\tt inc}$ holds in state $s$.
-The execution of {\tt swap} returning {\tt null} in a dequeue $k_d$ enables new instantiations $(k,k',k_d)$ in $s'$, thus adding potentially new constraints $k\not< k'$.
-We show that these instantiations are however vacuous because $k$ must be pending in $s$ and thus maximal in every order $<$ associated by $F_1$ to $s$.
+The same is true for the increment of {\tt i} in a dequeue $k_d$ since the predicate ${\tt afterSwapNull}(k_d)$ holds in state $s$.
+The execution of {\tt swap} returning {\tt null} in a dequeue $k_d$ enables new instantiations $(k,k',k_d)$ in $s'$, thus adding potentially new constraints $\neg \mathsf{before}(k,k')$.
+We show that these instantiations are however vacuous because $k$ must be pending in $s$ and thus maximal in every order $\mathsf{before}$ associated by $F_1$ to $s$.
 Let $k$ and $k'$ be two enqueues such that together with the dequeue $k_d$ they satisfy the property (\ref{eq:inst}) in $s'$ but not in $s$. 
@@
 We have that ${\tt i}_{s'}(k) = {\tt i}_{s'}(k_d) \leq {\tt i}_{s'}(k')$ and ${\tt items}[{\tt i}_{s'}(k_d)]={\tt null}$. The latter implies that the enqueue $k$ didn't execute
-the second statement (since the position it reserved is still {\tt null}) and it is pending in $s$.
-The step that checks that the value returned by {\tt swap} is {\tt null} doesn't modify the variables in property  (\ref{eq:inst}) and also, it doesn't change the valuation of the predicate $\atCP {\tt if}\text{-}{\tt inc}$.
+the second statement (since the position it reserved is still {\tt null}) and it is pending in $s'$.
+The step that swaps the null item does not modify anything except the control point of $k_d$ that makes ${\tt afterSwapNull}(k_d)$ true in $s'$ .
+Hence, $i_s(k) = i_s(k_d)\leq i_s(k')$ and ${\tt items}[i_s(k_d)] = {\tt null}$ is also true.
+Therefore, $k$ is pending in $s$ and maximal.
+Hence, $\mathsf{before}(k,k')$ is not true in both $s$ and $s'$.
 
@@
 By the definition of $\mathit{HWQ}$, there is a unique enqueue $k_e$ which filled the position updated by $k$, i.e., ${\tt i}_s(k_e)=i_s(k)$ and ${\tt x}_{s'}(k)={\tt x}_s(k_e)$.
-We show that $k_e$ is minimal in the order $<$ of $t$ which implies that $lin(deq,d,k)$ is enabled in $t$.
-Thus, instantiating item (c) in the definition of $<$ with $k'=k_e$ and $k_d=k$ we get that every enqueue that reserved a position smaller than the one of $k_e$ can't be ordered before $k_e$ in the order $<$.
+
+We show that $k_e$ is minimal in the order $\mathsf{before}$ of $t$ which implies that $k_e$ could be chosen by $lin(deq,d,k)$ step applied on $t$.
+As explained previously, instantiating item (c) in the definition of $\mathsf{before}$ with $k'=k_e$ and $k_d=k$; and instantiating item (b) with $k=k_e$, we ensures the minimality of $k_e$.
+Moreover, the state $t'$ obtained from $t$ through a $lin(deq,d,k)$ transition is related to $s'$ because the value added by $k_e$ is not anymore present in {\tt items} and $\mathsf{present}(k_e)$ doesn't hold in $t'$.
+%Thus, instantiating item (c) in the definition of $\mathsf{before}$ with $k'=k_e$ and $k_d=k$ we get that every enqueue that reserved a position smaller than the one of $k_e$ can't be ordered before $k_e$ in the order $\mathsf{before}$.
 Also, applying item (b) with $k=k_e$ we get the same for every enqueue that reserved a bigger position.

### 2017-05-05 draftCAV/queue.tex
Clarity and concision in Section 4.0.

@@
 \vspace{-1.5mm}
-The typical abstract implementation of a concurrent queue, denoted as $AbsQ_0$, maintains a sequence of values, the enqueue adds a value atomically to the beginning of the sequence, and the dequeue removes a value from the end of the sequence (if any, otherwise it returns {\tt EMPTY}).
-Both methods have a fixed linearization point when the update of the sequence happens.
-%~\footnote{For $\<Methods>=\{enq,deq\}$, the alphabet of $AbsQ_0$ is $C\cup R\cup Lin$.}.
-For some queue implementations, e.g., the Herlihy\&Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), there exists no forward simulation to $AbsQ_0$ although they are a refinement of $AbsQ_0$.
-The main reason is that the linearization points of the enqueue operations are not fixed, they depend on future dequeue invocations.
-In this section, we propose a new abstract implementation for queues, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence.
-We show that there exists a forward simulation from any correct queue implementation where only the \emph{dequeue} methods have fixed linearization points (the enqueue methods are unconstrained) to $AbsQ$.
-This covers all the queue implementations that we are aware of, in particular $\mathit{HWQ}$, Baskets Queue~\cite{DBLP:conf/opodis/HoffmanSS07}, LCRQ~\cite{DBLP:conf/ppopp/MorrisonA13}, or Time-Stamped Queue~\cite{DBLP:conf/popl/DoddsHK15} (where the enqueues don't have fixed linearization points).
-We also describe a forward simulation from $\mathit{HWQ}$ to $AbsQ$.
+
+The classical abstract queue implementation, denoted $AbsQ_0$, maintains a
+sequence of enqueued values; dequeues return the oldest non-dequeued value, at
+the time of their linearization points, or {\tt EMPTY}. Some implementations,
+like the queue of \citet{journals/toplas/HerlihyW90}, denoted {\sc hwq}, are not
+simulated by $AbsQ_0$, even though they refine $AbsQ_0$, since the order in
+which their enqueues are linearized to form $AbsQ_0$’s sequence is not
+determined until later, when their values are dequeued.
+
+In this section we develop an abstract queue implementation, denoted $AbsQ$,
+which maintains a partial order of enqueues, rather than a linear sequence.
+Since $AbsQ$ does not force refining implementations to eagerly pick among
+linearizations of their enqueues, it simulates many more queue implementations.
+In fact, $AbsQ$ simulates all queue implementations of which we are aware that
+are not simulated by $AbsQ_0$, including {\sc hwq}, The Baskets
+Queue~\cite{DBLP:conf/opodis/HoffmanSS07}, The Linked Concurrent Ring Queue
+({\sc lcrq})~\cite{DBLP:conf/ppopp/MorrisonA13}, and The Time-Stamped
+Queue~\cite{DBLP:conf/popl/DoddsHK15}.
 

### 2017-05-05 draftCAV/queue.tex
Using "forward simulates."

@@
 like the queue of \citet{journals/toplas/HerlihyW90}, denoted {\sc hwq}, are not
-simulated by $AbsQ_0$, even though they refine $AbsQ_0$, since the order in
-which their enqueues are linearized to form $AbsQ_0$’s sequence is not
+forward-simulated by $AbsQ_0$, even though they refine $AbsQ_0$, since the order
+in which their enqueues are linearized to form $AbsQ_0$’s sequence is not
 determined until later, when their values are dequeued.
@@
 Since $AbsQ$ does not force refining implementations to eagerly pick among
-linearizations of their enqueues, it simulates many more queue implementations.
-In fact, $AbsQ$ simulates all queue implementations of which we are aware that
-are not simulated by $AbsQ_0$, including {\sc hwq}, The Baskets
-Queue~\cite{DBLP:conf/opodis/HoffmanSS07}, The Linked Concurrent Ring Queue
-({\sc lcrq})~\cite{DBLP:conf/ppopp/MorrisonA13}, and The Time-Stamped
-Queue~\cite{DBLP:conf/popl/DoddsHK15}.
+linearizations of their enqueues, it forward-simulates many more queue
+implementations. In fact, $AbsQ$ forward-simulates all queue implementations of
+which we are aware that are not forward-simulated by $AbsQ_0$, including {\sc
+hwq}, The Baskets Queue~\cite{DBLP:conf/opodis/HoffmanSS07}, The Linked
+Concurrent Ring Queue ({\sc lcrq})~\cite{DBLP:conf/ppopp/MorrisonA13}, and The
+Time-Stamped Queue~\cite{DBLP:conf/popl/DoddsHK15}.
 

### 2017-05-05 draftCAV/queue.tex
Syntax of abstract state machines.

@@
 %		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
-		& \text{\underline{Shared State}} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{\underline{Abstract State}} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \mathbb{B} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \mathbb{B} \\
 		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \mathbb{B} \\
 %		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
@@
 \begin{lstlisting}
-void enq($\<Vals>$ x) {
-  atomic rule inv(enq,x,k = new);
-  atomic rule ret(enq,k);
-}
+void enq($\<Vals>$ x):
+  atomic rule inv(enq,x,k = new)
+  atomic rule ret(enq,k)
 
-$\<Vals>$ deq() {
-  atomic rule inv(deq,k = new);
-  atomic rule lin(deq,y,k);
-  atomic rule ret(deq,y,k);
-}
+$\<Vals>$ deq():
+  atomic rule inv(deq,k = new)
+  atomic rule lin(deq,y,k)
+  atomic rule ret(deq,y,k)
+
 \end{lstlisting}
@@
 \begin{lstlisting}
-rule inv(enq,x,k) = {
-  $\mathsf{present}$(k) = true;
-  $\mathsf{pending}$(k) = true;
-  $\mathsf{arg}$(k) = x;
-  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1) do
-    $\mathsf{before}$(k1,k) = true;
-}
+rule inv(enq,x,k):
+  $\mathsf{present}$(k) := true
+  $\mathsf{pending}$(k) := true
+  $\mathsf{arg}$(k) := x
+  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1):
+    $\mathsf{before}$(k1,k) := true
 
-rule lin(deq,y,k) = {
-  choose a $\in$ {true, false};
-  if ( $($a$\ \vee\ \forall k'.\ \mathsf{present}(k')$=false$)\ \land\ \forall k'.\ \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$ )
-    y = EMPTY;
-  else
-    choose k1 $\in$ $\<Ops>$ with $\mathsf{present}$(k1)$\land\neg \exists k'.\ \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$;
-    $\mathsf{present}$(k1) = false;
-    y = $\mathsf{arg}$(k1);
-}
+rule lin(deq,y,k):
+  choose a $\in$ $\mathbb{B}$
+  if $($a$\vee\forall k'. \mathsf{present}(k')$=false$) \land \forall k'. \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$:
+    y := EMPTY
+  else:
+    choose k1 $\in$ $\<Ops>$ with $\mathsf{present}$(k1)$\land\neg \exists k'. \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$
+    $\mathsf{present}$(k1) := false
+    y := $\mathsf{arg}$(k1)
+
 \end{lstlisting}
@@
 \begin{lstlisting}
-rule ret(enq,k) = {
-  $\mathsf{pending}$(k) = false;
-  return;
-}
-rule inv(deq,k) = {
-  skip;
-}
-rule ret(deq,y,k) = {
-  return y;
-}
+rule ret(enq,k):
+  $\mathsf{pending}$(k) := false
+
+rule inv(deq,k):
+  skip
+
+rule ret(deq,y,k):
+  return y
+
 \end{lstlisting}

### 2017-05-05 draftCAV/queue.tex
Deterministic rules.

@@
 \begin{lstlisting}
-rule inv(enq,x,k):
+rule inv(enq,v):
+  let k = new $\mathbb{O}$
   $\mathsf{present}$(k) := true
   $\mathsf{pending}$(k) := true
-  $\mathsf{arg}$(k) := x
-  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1):
-    $\mathsf{before}$(k1,k) := true
+  $\mathsf{arg}$(k) := v
+  forall k' with $\mathsf{present}$(k')$\land \neg \mathsf{pending}$(k'):
+    $\mathsf{before}$(k',k) := true
 
-rule lin(deq,y,k):
-  choose a $\in$ $\mathbb{B}$
-  if $($a$\vee\forall k'. \mathsf{present}(k')$=false$) \land \forall k'. \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$:
-    y := EMPTY
-  else:
-    choose k1 $\in$ $\<Ops>$ with $\mathsf{present}$(k1)$\land\neg \exists k'. \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$
-    $\mathsf{present}$(k1) := false
-    y := $\mathsf{arg}$(k1)
+rule lin(enq,v):
+  pass
+
+rule ret(enq,v):
+  let k with $\mathsf{arg}$(k) = v
+  $\mathsf{pending}$(k) := false
 
@@
 \begin{lstlisting}
-rule ret(enq,k):
-  $\mathsf{pending}$(k) := false
+rule inv(deq,v):
+  pass
 
-rule inv(deq,k):
-  skip
+rule lin(deq,v):
+  if v = EMPTY:
+    forall k with $\mathsf{present}$(k):
+      assert $\mathsf{pending}$(k)
+  else:
+    assert $\mathsf{present}$(k)
+    let k with $\mathsf{arg}$(k) = v
+    forall k' with $\mathsf{present}$(k'):
+      assert $\lnot\mathsf{before}$(k', k)
+    $\mathsf{present}$(k) := false
 
-rule ret(deq,y,k):
-  return y
+rule ret(deq,v):
+  pass
 

### 2017-05-05 draftCAV/queue.tex
More rule syntax.

@@
 \begin{lstlisting}
-void enq($\<Vals>$ x):
-  atomic rule inv(enq,x,k = new)
-  atomic rule ret(enq,k)
+void enq($\<Vals>$ v):
+  atomic rule inv(enq,v)
+  atomic rule lin(enq,v)
+  atomic rule ret(enq,v)
 
 $\<Vals>$ deq():
-  atomic rule inv(deq,k = new)
-  atomic rule lin(deq,y,k)
-  atomic rule ret(deq,y,k)
+  let v with ...
+  atomic rule inv(deq,v)
+  atomic rule lin(deq,v)
+  atomic rule ret(deq,v)
+  return v
 

### 2017-05-05 draftCAV/queue.tex
Syntax of abstract state machines.

@@
 \begin{minipage}[t]{7.5cm}
-\scriptsize{
-{	%\centering
-	\begin{align*}
-%		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
-		& \text{\underline{Abstract State}} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \mathbb{B} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \mathbb{B} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \mathbb{B} \\
-%		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
-%		& \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} &\mbox{ // the argument of an enqueue} \\
-	\end{align*}
-}}
-\end{minipage}
-\begin{minipage}[t]{7.5cm}
 \begin{lstlisting}
@@
 
+\end{lstlisting}
+\end{minipage}
+\begin{minipage}[t]{7.5cm}
+\begin{lstlisting}
 $\<Vals>$ deq():
@@
 \begin{lstlisting}
+function $\mathsf{present}$: $\mathbb{O} \to \mathbb{B}$
+function $\mathsf{pending}$: $\mathbb{O} \to \mathbb{B}$
+function $\mathsf{arg}$: $\mathbb{O} \to \mathbb{V}$
+function $\mathsf{before}$: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$
+
 rule inv(enq,v):

### 2017-05-05 draftCAV/queue.tex
Syntax of abstract state machines.

@@
 \begin{lstlisting}
-function $\mathsf{present}$: $\mathbb{O} \to \mathbb{B}$
-function $\mathsf{pending}$: $\mathbb{O} \to \mathbb{B}$
-function $\mathsf{arg}$: $\mathbb{O} \to \mathbb{V}$
-function $\mathsf{before}$: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$
+function present: $\mathbb{O} \to \mathbb{B}$
+function pending: $\mathbb{O} \to \mathbb{B}$
+function arg: $\mathbb{O} \to \mathbb{V}$
+function before: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$
 
@@
   let k = new $\mathbb{O}$
-  $\mathsf{present}$(k) := true
-  $\mathsf{pending}$(k) := true
-  $\mathsf{arg}$(k) := v
-  forall k' with $\mathsf{present}$(k')$\land \neg \mathsf{pending}$(k'):
-    $\mathsf{before}$(k',k) := true
+  present(k) := true
+  pending(k) := true
+  arg(k) := v
+  forall k' with present(k')$\land \lnot$pending(k'):
+    before(k',k) := true
 
@@
 rule ret(enq,v):
-  let k with $\mathsf{arg}$(k) = v
-  $\mathsf{pending}$(k) := false
+  let k with arg(k) = v
+  pending(k) := false
 
@@
   if v = EMPTY:
-    forall k with $\mathsf{present}$(k):
-      assert $\mathsf{pending}$(k)
+    forall k with present(k):
+      assert pending(k)
   else:
-    assert $\mathsf{present}$(k)
-    let k with $\mathsf{arg}$(k) = v
-    forall k' with $\mathsf{present}$(k'):
-      assert $\lnot\mathsf{before}$(k', k)
-    $\mathsf{present}$(k) := false
+    assert present(k)
+    let k with arg(k) = v
+    forall k' with present(k'):
+      assert $\lnot$before(k', k)
+    present(k) := false
 

### 2017-05-05 draftCAV/queue.tex
first part of H&Q proof

--- 2017-05-05 draftCAV/queue.tex
@@
 \begin{lstlisting}
-void enq($\<Vals>$ v):
-  atomic rule inv(enq,v)
-  atomic rule lin(enq,v)
-  atomic rule ret(enq,v)
+void enq($\<Vals>$ x):
+  atomic rule inv(enq,x,k = new)
+  atomic rule ret(enq,k)
 
 $\<Vals>$ deq():
-  let v with ...
-  atomic rule inv(deq,v)
-  atomic rule lin(deq,v)
-  atomic rule ret(deq,v)
-  return v
+  atomic rule inv(deq,k = new)
+  atomic rule lin(deq,y,k)
+  atomic rule ret(deq,y,k)
 
@@
 \begin{lstlisting}
-rule inv(enq,v):
-  let k = new $\mathbb{O}$
+rule inv(enq,x,k):
   $\mathsf{present}$(k) := true
   $\mathsf{pending}$(k) := true
-  $\mathsf{arg}$(k) := v
-  forall k' with $\mathsf{present}$(k')$\land \neg \mathsf{pending}$(k'):
-    $\mathsf{before}$(k',k) := true
+  $\mathsf{arg}$(k) := x
+  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1):
+    $\mathsf{before}$(k1,k) := true
 
-rule lin(enq,v):
-  pass
-
-rule ret(enq,v):
-  let k with $\mathsf{arg}$(k) = v
-  $\mathsf{pending}$(k) := false
+rule lin(deq,y,k):
+  choose a $\in$ $\mathbb{B}$
+  if $($a$\vee\forall k'. \mathsf{present}(k')$=false$) \land \forall k'. \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$:
+    y := EMPTY
+  else:
+    choose k1 $\in$ $\<Ops>$ with $\mathsf{present}$(k1)$\land\neg \exists k'. \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$
+    $\mathsf{present}$(k1) := false
+    y := $\mathsf{arg}$(k1)
 
@@
 \begin{lstlisting}
-rule inv(deq,v):
-  pass
+rule ret(enq,k):
+  $\mathsf{pending}$(k) := false
 
-rule lin(deq,v):
-  if v = EMPTY:
-    forall k with $\mathsf{present}$(k):
-      assert $\mathsf{pending}$(k)
-  else:
-    assert $\mathsf{present}$(k)
-    let k with $\mathsf{arg}$(k) = v
-    forall k' with $\mathsf{present}$(k'):
-      assert $\lnot\mathsf{before}$(k', k)
-    $\mathsf{present}$(k) := false
+rule inv(deq,k):
+  skip
 
-rule ret(deq,v):
-  pass
+rule ret(deq,y,k):
+  return y
 
@@
 
-An enqueue operation in a $\mathit{HWQ}$ state is pending and present if and only if its argument is stored in the array ${\tt items}$ or it has not yet written to the array  ${\tt items}$.
+%An enqueue operation in a $\mathit{HWQ}$ state is pending and present if and only if its argument is stored in the array ${\tt items}$ or it has not yet written to the array  ${\tt items}$.
 Those operations should be $\mathsf{pending}$ and $\mathsf{present}$ in the related $AbsQ$ states.
 In addition to this, the forward simulation $F_1$ imposes following restrictions:
-%A $\mathit{HWQ}$ state is related by $F_1$ to an $AbsQ$ state where $\mathsf{present}$ is true for all the enqueues whose argument is  stored in the array ${\tt items}$, and all the pending enqueues that have not yet written to the array ${\tt items}$ (and only for these enqueues).
-The order relation $\mathsf{before}$ between these enqueues is defined as follows:
+An $\mathit{HWQ}$ state $s$ is related by $F_1$ to $AbsQ$ states $t$ where the predicate 
+$\mathsf{present}$ is true for all the enqueues in $s$ whose argument is stored in the array ${\tt items}$, and all the pending enqueues that have not yet written to the array ${\tt items}$ (and only for these enqueues).
+We refer to such enqueues in $s$ as $\mathsf{present}$ enqueues.
+Also, $\mathsf{pending}(k)$ is ${\tt true}$ in $t$ whenever $k$ is a pending enqueue in $s$, $\mathsf{arg}(d)=k$ in $t$ whenever the argument of the enqueue $k$ in $s$ is $d$, and for every dequeue operation $k$ such that ${\tt x}(k)=d\neq {\tt null}$, we have that ${\tt y}(k)=d$ (recall that ${\tt y}$ is a local variable of the dequeue method in $AbsQ$).
+The order relation $\mathsf{before}$ in $t$ satisfies the following constraints:
 \vspace{-2mm}
 \begin{itemize}
-	\item[(a)] pending and present enqueues are maximal, i.e., for every two $\mathsf{present}$ enqueues $k$ and $k'$ such that $k'$ is $\mathsf{pending}$, we have that $\neg \mathsf{before}(k',k)$, % $k'\not< k$,
+	\item[(a)] pending enqueues are maximal, i.e., for every two $\mathsf{present}$ enqueues $k$ and $k'$ such that $k'$ is $\mathsf{pending}$, we have that $\neg \mathsf{before}(k',k)$, % $k'\not< k$,
 	\item[(b)] $\mathsf{before}$ is consistent with the order in which positions of ${\tt items}$ have been reserved, i.e., for every two present enqueues $k$ and $k'$ such that ${\tt i}(k) < {\tt i}(k')$, we have that $\neg \mathsf{before}(k',k)$, %$k' \not< k$,
-	\item[(c)] an enqueue which has reserved a position $i$ %and executed only the first statement
-	can't be ordered before another enqueue that has reserved a position $j \geq i$ when the position $i$ has been ``observed'' by a non-linearized dequeue that may ``observe'' $j$ in the current array traversal, i.e., for every two present enqueues $k$ and $k'$, and a dequeue $k_d$, such that
+	\item[(c)] if the position $i$ reserved by an enqueue $k$ has been ``observed'' by a non-linearized dequeue that in the current array traversal may ``observe'' a later position $j$ reserved by another enqueue $k'$, then $k$ can't be ordered before $k'$,
+%	an enqueue which has reserved a position $i$ %and executed only the first statement
+%	can't be ordered before another enqueue that has reserved a position $j \geq i$ , 
+	i.e., for every two present enqueues $k$ and $k'$, and a dequeue $k_d$, such that
 
@@
 	\hspace{-8mm}
-	{\tt canRemove}(k_d,k) \land (i(k') < i(k_d) \lor (i(k')=i(k_d) \land {\tt afterSwapNull}(k_d)))
+	{\tt canRemove}(k_d,k') \land ({\tt i}(k) < {\tt i}(k_d) \lor ({\tt i}(k)={\tt i}(k_d) \land {\tt afterSwapNull}(k_d)))
 \label{eq:inst}
@@
 	we have that $\neg \mathsf{before}(k,k')$.
-The predicate ${\tt canRemove}(k_d,k)$ holds when $k_d$ has currently visited a {\tt null} item in {\tt items} and the present enqueue index $i(k)$ is in the range of $(k_d)$ i.e., $\mathsf{present}(k) \land (x(k_d) = {\tt null} \land i(k) < {\tt range}(k_d) \land i(k_d) < i(k)) \lor (i(k_d) = i(k) \land {\tt beforeSwap}(k_d) \land {\tt items}[i(k)] != {\tt null})$.
-The predicates ${\tt afterSwapNull}(k_d)$ (resp., ${\tt beforeSwap}(k_d)$) holds when the dequeue $k_d$ is at the control point after a ${\tt swap}$ returning ${\tt null}$ (resp., before a ${\tt swap}$).
+The predicate ${\tt canRemove}(k_d,k')$ holds when $k_d$ visited a {\tt null} item in {\tt items} and the position $i(k')$ reserved by $k'$ is in the range of $(k_d)$ i.e., $({\tt x}(k_d) = {\tt null} \land {\tt i}(k_d) < {\tt i}(k') \leq {\tt range}(k_d)) \lor ({\tt i}(k_d) = {\tt i}(k') \land {\tt beforeSwap}(k_d) \land {\tt items}[{\tt i}(k')] != {\tt null})$.
+The predicate ${\tt afterSwapNull}(k_d)$ (resp., ${\tt beforeSwap}(k_d)$) holds when the dequeue $k_d$ is at the control point after a ${\tt swap}$ returning ${\tt null}$ (resp., before a ${\tt swap}$).
 \vspace{-2mm}
 \end{itemize}
-
-\noindent
-We have that $\mathsf{pending}(k)$ is ${\tt true}$ whenever $k$ is a pending enqueue, and $\mathsf{arg}(k)=d$ whenever the argument of the enqueue $k$ is $d$.
+%\noindent
 %An enqueue is labeled by $(d,{\tt PEND})$ where $d$ is the input value if it's pending and by  $(d,{\tt COMP})$, otherwise.
-Also, for every dequeue operation $k$ such that ${\tt x}(k)=d\neq {\tt null}$, we have that ${\tt y}(k)=d$ (recall that ${\tt y}$ is a local variable of the dequeue method in $AbsQ$).
-
-The restrictions of $F_1$ aim to ensure two important points:  (i) identifier of a pending and present enqueue method in the $\mathit{HWQ}$ state should be maximal according to $\mathsf{before}$ predicate in the related $AbsQ$ state and (ii) identifiera of a present enqueue method in $\mathit{HWQ}$ state of which data value is about to be removed by a dequeue operation should be minimal in the related $AbsQ$ state.
-
-The first point is ensured by the restriction (a).
-Present and pending enqueue identifiers in the $\mathit{HWQ}$ state are $\mathsf{pending}$ in the related $AbsQ$ state and $\mathsf{pending}$ enqueues are maximal in a valid $AbsQ$ state.
-
-Restrictions (b) and (c) ensure the point (ii).
-Consider a present enqueue $k$ that inserted its argument to ${\tt items}$ and  there exists a pending dequeue $k_d$ such that ${\tt canRemove}(k_d, k)$ and $k_d$ is just before its swap action at the reserved position of $k$ i.e., $i(k_d) = i(k)$.
-A pending enqueue operation cannot be ordered $\mathsf{before}$ $k$ since pending enqueues are maximal by (a).
-Regarding the completed and present enqueues $k'$, we consider two cases: $i(k') > i(k)$ and $i(k') < i(k)$.
-For the former case, the restriction (b) ensures $\neg \mathsf{before}(k',k)$ and for the latter case the restriction (c) ensures $\neg \mathsf{before}(k',k)$.
-Consequently, $k$ is a minimal element in $\mathsf{before}$ relation just before $k_d$ removes its data value.
+%The constraint (a) ensures  on $\mathsf{before}$ ensure that: (i) every present and pending enqueue is maximal  and (ii) a present enqueue whose argument is about to be removed by a dequeue operation is minimal.
+%
+%The first point is ensured by the constraint (a). 
+%%Present and pending enqueue identifiers in the $\mathit{HWQ}$ state are $\mathsf{pending}$ in the related $AbsQ$ state and $\mathsf{pending}$ enqueues are maximal in a valid $AbsQ$ state.
+%
+The constraints on $\mathsf{before}$ ensure that a present enqueue whose argument is about to be removed by a dequeue operation is minimal.
+Thus, let $k'$ be a present enqueue that inserted its argument to ${\tt items}$, and $k_d$ a pending dequeue such that ${\tt canRemove}(k_d, k')$ holds and $k_d$ is just before its swap action at the reserved position of $k'$ i.e., ${\tt i}(k_d) = {\tt i}(k')$.
+Another pending enqueue $k$ cannot be ordered before $k'$ since pending enqueues are maximal by (a).
+Regarding the completed and present enqueues $k$, we consider two cases: $i(k) > i(k')$ and $i(k) < i(k')$.
+For the former case, the constraint (b) ensures $\neg \mathsf{before}(k,k')$ and for the latter case the constraint (c) ensures $\neg \mathsf{before}(k,k')$.
+Consequently, $k'$ is a minimal element w.r.t.
+$\mathsf{before}$ just before $k_d$ removes its argument.
 

### 2017-05-05 draftCAV/queue.tex

@@
 \begin{minipage}[t]{7.5cm}
-\scriptsize{
-{	%\centering
-	\begin{align*}
-%		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
-		& \text{\underline{Abstract State}} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \mathbb{B} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \mathbb{B} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \mathbb{B} \\
-%		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
-%		& \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} &\mbox{ // the argument of an enqueue} \\
-	\end{align*}
-}}
+\begin{lstlisting}
+void enq($\<Vals>$ v):
+  atomic rule inv(enq,v)
+  atomic rule lin(enq,v)
+  atomic rule ret(enq,v)
+
+\end{lstlisting}
 \end{minipage}
@@
 \begin{lstlisting}
-void enq($\<Vals>$ x):
-  atomic rule inv(enq,x,k = new)
-  atomic rule ret(enq,k)
-
 $\<Vals>$ deq():
-  atomic rule inv(deq,k = new)
-  atomic rule lin(deq,y,k)
-  atomic rule ret(deq,y,k)
+  let v with ...
+  atomic rule inv(deq,v)
+  atomic rule lin(deq,v)
+  atomic rule ret(deq,v)
+  return v
 
@@
 \begin{lstlisting}
-rule inv(enq,x,k):
-  $\mathsf{present}$(k) := true
-  $\mathsf{pending}$(k) := true
-  $\mathsf{arg}$(k) := x
-  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1):
-    $\mathsf{before}$(k1,k) := true
+function present: $\mathbb{O} \to \mathbb{B}$
+function pending: $\mathbb{O} \to \mathbb{B}$
+function arg: $\mathbb{O} \to \mathbb{V}$
+function before: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$
 
-rule lin(deq,y,k):
-  choose a $\in$ $\mathbb{B}$
-  if $($a$\vee\forall k'. \mathsf{present}(k')$=false$) \land \forall k'. \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$:
-    y := EMPTY
-  else:
-    choose k1 $\in$ $\<Ops>$ with $\mathsf{present}$(k1)$\land\neg \exists k'. \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$
-    $\mathsf{present}$(k1) := false
-    y := $\mathsf{arg}$(k1)
+rule inv(enq,v):
+  let k = new $\mathbb{O}$
+  present(k) := true
+  pending(k) := true
+  arg(k) := v
+  forall k' with present(k')$\land \lnot$pending(k'):
+    before(k',k) := true
+
+rule lin(enq,v):
+  pass
+
+rule ret(enq,v):
+  let k with arg(k) = v
+  pending(k) := false
 
@@
 \begin{lstlisting}
-rule ret(enq,k):
-  $\mathsf{pending}$(k) := false
+rule inv(deq,v):
+  pass
 
-rule inv(deq,k):
-  skip
+rule lin(deq,v):
+  if v = EMPTY:
+    forall k with present(k):
+      assert pending(k)
+  else:
+    assert present(k)
+    let k with arg(k) = v
+    forall k' with present(k'):
+      assert $\lnot$before(k', k)
+    present(k) := false
 
-rule ret(deq,y,k):
-  return y
+rule ret(deq,v):
+  pass
 

### 2017-05-05 draftCAV/queue.tex
A better abstract state machine.

@@
 
-%\begin{figure} [t]
 \begin{figure}[t]
-\hspace{-10mm}
-\begin{minipage}[t]{7.5cm}
-\begin{lstlisting}
-void enq($\<Vals>$ v):
-  atomic rule inv(enq,v)
-  atomic rule lin(enq,v)
-  atomic rule ret(enq,v)
-
-\end{lstlisting}
-\end{minipage}
-\begin{minipage}[t]{7.5cm}
-\begin{lstlisting}
-$\<Vals>$ deq():
-  let v with ...
-  atomic rule inv(deq,v)
-  atomic rule lin(deq,v)
-  atomic rule ret(deq,v)
-  return v
-
-\end{lstlisting}
-\end{minipage}
-
-\begin{minipage}[t]{7.5cm}
-\vspace{-3mm}
+  \begin{minipage}[t]{0.4\linewidth}
     \begin{lstlisting}
@@
 function pending: $\mathbb{O} \to \mathbb{B}$
-function arg: $\mathbb{O} \to \mathbb{V}$
+function arg: $\mathbb{V} \to \mathbb{O}$
 function before: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$
@@
 rule inv(enq,v):
-  let k = new $\mathbb{O}$
+  let k := new $\mathbb{O}$
   present(k) := true
@@
   arg(k) := v
-  forall k' with present(k')$\land \lnot$pending(k'):
+  forall k' with present(k'):
+    if $\lnot$pending(k'):
       before(k',k) := true
@@
 rule ret(enq,v):
-  let k with arg(k) = v
+  let k := arg(k)
   pending(k) := false
 
-\end{lstlisting}
-\end{minipage}
-\begin{minipage}[t]{5cm}
-\vspace{-3mm}
-\begin{lstlisting}
 rule inv(deq,v):
@@
     assert present(k)
-    let k with arg(k) = v
+    let k := arg(v)
     forall k' with present(k'):
@@
   pass
-
     \end{lstlisting}
+    \caption{The $AbsQ$ implementation.}
+    \label{fig:transitions:AbsQ}
   \end{minipage}
-%	\caption{The abstract queue implementation $AbsQ$.}
-%	\label{fig:signatures}
-%\end{figure}
-
-%{\scriptsize
-%  \centering
-%  \begin{mathpar}
-%    \inferrule[call-enq]{
-%      k\not\in dom(cp) \\
-%      d\neq {\tt EMPTY}
-%    }{
-%      O,<,\ell,rv,cp
-%      \xrightarrow{inv(enq,d,k)}
-%      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
-%      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1]
-%    }\hspace{5mm}
-%
-%    \inferrule[call-deq]{
-%      k\not\in dom(cp) \\
-%    }{
-%      O,<,\ell,rv,cp
-%      \xrightarrow{inv(deq,k)}
-%      O,<,\ell,rv,cp[k\mapsto R_1]
-%    }\hspace{5mm}
-%    \inferrule[ret-deq]{
-%       cp(k) = R_2 \\
-%       rv(k)=d
-%    }{
-%      O,<,\ell,rv,cp
-%      \xrightarrow{ret(deq,d,k)}
-%      O,<,\ell,rv,cp[k\mapsto R_3]
-%    }\hspace{5mm}
-%
-%    \inferrule[ret-enq1]{
-%      cp(k) = A_1 \\
-%      k \in O \\
-%      \ell(k) = (d,{\tt PEND})
-%    }{
-%      O,<,\ell,rv,cp
-%      \xrightarrow{ret(enq,k)}
-%      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2]
-%    }\hspace{5mm}
-%    \inferrule[ret-enq2]{
-%      cp(k) = A_1 \\
-%      k \not\in O
-%    }{
-%      O,<,\ell,rv,cp
-%      \xrightarrow{ret(enq,k)}
-%      O,<,\ell,rv,cp[k\mapsto A_2]
-%    }\hspace{5mm}
-%
-%    \inferrule[lin-deq1]{
-%       cp(k) = R_1 \\
-%       d\neq{\tt EMPTY} \\
-%       k'\in min(O) \\
-%       \ell_1(k')=d
-%    }{
-%      O,<,\ell,rv,cp
-%      \xrightarrow{lin(deq,d,k)}
-%      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2]
-%    }\hspace{5mm}
-%    \inferrule[lin-deq2]{
-%       cp(k) = R_1 \\
-%       \forall o\in O.\ \ell_2(o)={\tt PEND}
-%    }{
-%      O,<,\ell,rv,cp
-%      \xrightarrow{lin(deq,{\tt EMPTY},k)}
-%      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2]
-%    }\hspace{5mm}
-%      \end{mathpar}
-%  }
- \vspace{-4mm}
-  \caption{The abstract queue implementations $AbsQ$. Each macro rule is executed atomically in one step (indicated by the keyword {\ttfamily \bfseries atomic}).
-%  transition relation of $AbsQ$.
-We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.
-  %\textcolor{red}{Call-Enq must have $d!= \texttt{EMPTY}$ as a premise. Also lin deq returning empty must be changed as before.}
-  }
-  \label{fig:transitions:AbsQ}
-\vspace{-6mm}
 \end{figure}

### 2017-05-05 draftCAV/queue.tex
Typos in abstract state machine.

@@
   pending(k) := true
-  arg(k) := v
+  arg(v) := k
   forall k' with present(k'):
@@
 rule ret(enq,v):
-  let k := arg(k)
+  let k := arg(v)
   pending(k) := false
@@
   else:
+    let k := arg(v)
     assert present(k)
-    let k := arg(v)
     forall k' with present(k'):

### 2017-05-05 draftCAV/queue.tex
first part of H&Q proof

--- 2017-05-05 draftCAV/queue.tex
@@
 Next, we show that $F_1$ is indeed a $C\cup R\cup Lin(deq)$-forward simulation. Let $s$ and $t$ be states of $\mathit{HWQ}$ and $AbsQ$, respectively, such that $(s,t)\in F_1$.
-We omit discussing the trivial case of transitions labeled by call and return actions which are simulated by similar transitions of $AbsQ$ (for the return a dequeue operation $k$, we use the equality between the local variable ${\tt x}(k)$ in $s$ and the component $rv(k)$ in $t$).
+We omit discussing the trivial case of transitions labeled by call and return actions which are simulated by similar transitions of $AbsQ$.
+% (for the return a dequeue operation $k$, we use the equality between the local variable ${\tt x}(k)$ in $s$ and the component $rv(k)$ in $t$).
 %\textcolor{red}{ I think it is good to mention again that call/return actions in HWQ correspond to the same call/return actions in AbsQ (without any other internal action).
@@
 %\setlength{\parskip}{0pt}
-We show that each internal step of an enqueue or dequeue, except the execution of {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an \emph{empty} sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in F_1$, then $(s',t)\in F_1$ for each $AbsQ$ state $t$.
+We show that each internal step of an enqueue or dequeue, except a {\tt swap} returning a non-null value in dequeue (which represents its linearization point), is simulated by an \emph{empty} sequence of $AbsQ$ transitions, i.e., for every state $s'$ obtained through one of these steps, if $(s,t)\in F_1$, then $(s',t)\in F_1$ for each $AbsQ$ state $t$.
 Essentially, this consists in proving the following property, called \emph{monotonicity}: the set of possible $\mathsf{before}$ relations associated by $F_1$ to $s'$ doesn't exclude any order $\mathsf{before}$ associated to $s$.
@@
 We show that $k_e$ is minimal in the order $\mathsf{before}$ of $t$ which implies that $k_e$ could be chosen by $lin(deq,d,k)$ step applied on $t$.
-As explained previously, instantiating item (c) in the definition of $\mathsf{before}$ with $k'=k_e$ and $k_d=k$; and instantiating item (b) with $k=k_e$, we ensures the minimality of $k_e$.
+As explained previously, instantiating item (c) in the definition of $\mathsf{before}$ with $k'=k_e$ and $k_d=k$, and instantiating item (b) with $k=k_e$, we ensure the minimality of $k_e$.
 Moreover, the state $t'$ obtained from $t$ through a $lin(deq,d,k)$ transition is related to $s'$ because the value added by $k_e$ is not anymore present in {\tt items} and $\mathsf{present}(k_e)$ doesn't hold in $t'$.

### 2017-05-05 draftCAV/queue.tex
Consolidated figures.

@@
 
-\begin{wrapfigure}{l}{7.2cm}
-\vspace{-6mm}
-\includegraphics[width=7.3cm]{fig-queue12.pdf}
-%
-%\vspace{2mm}
-%\includegraphics[width=7cm]{fig-queue2.pdf}
-\vspace{-8mm}
-\caption{Simulating queue histories with $AbsQ$.
-An operation is pictured by a line delimited by two circles denoting the call and respectively, the return action.
-The representation of a dequeue operation includes a red circle that stands for a {\tt swap} returning a non-null value, which is their linearization point.}
-\label{fig:queueSim}
-\vspace{-6mm}
-\end{wrapfigure}
 Fig.~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
@@
 
-For readability, we define $AbsQ$ as an abstract state machine, which is given in Fig.~\ref{fig:transitions:AbsQ}.
-The shared state of $AbsQ$ consists of several boolean predicates indicating whether the value added by an enqueue has not been removed yet (the predicate $\mathsf{present})$, whether an enqueue is pending (the predicate $\mathsf{pending}$), the happens-before order (the predicate $\mathsf{before}$), and a function giving the value added by an enqueue (the function $\mathsf{arg}$).
-In the initial state, the domain of every function (predicate) is empty.
-%the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).
-%All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}.
-The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points, denoted by $lin(deq,d,k)$.
-$Lin(deq)$ is the set of all actions $lin(deq,d,k)$.
-%
-The enqueue operations consist of two macro rules: the rule {\tt inv(enq,x,k)} orders the invoked operation after all the completed enqueues present in the current state ({\tt k} is a fresh operation identifier associated to the current operation), and the rule {\tt ret(enq,k)} sets $\mathsf{pending}${\tt (k)} to {\tt false}.
-% provided that the operation is still present in the current state.
-The dequeue operations consist of two rules {\tt inv(deq,k)} and {\tt ret(deq,y,k)} associated to the invocation and respectively, the return action, whose definition is obvious, and the rule {\tt lin(deq,y,k)} corresponding to its linearization point.
-When a non-empty set of pending enqueues are present, the rule {\tt lin(deq,y,k)} is non-deterministic.
-It can set the return value {\tt y} either to {\tt EMPTY}, or to a value {\tt d} $\neq$ {\tt EMPTY}, provided that {\tt d} has been added by an enqueue {\tt k1} which is minimal in the current happens-before.
-In the latter case, $\mathsf{present}${\tt (k1)} is set to {\tt false} to indicate that {\tt k1}'s argument has been removed.
-When there exists at least one completed enqueue whose value has not been removed, the rule {\tt lin(deq,y,k)} sets {\tt y} to the value added by a minimal enqueue (w.r.t.
-the current happens-before).
+\begin{figure}[t]
+  \begin{minipage}[b]{0.55\linewidth}
+    \begin{lstlisting}
+void enq(int x) {
+  i = back++; items[i] = x;
+}
+int deq() {
+  while (1) {
+    range = back - 1;
+    for (int i = 0; i <= range; i++) {
+      x = swap(items[i],null);
+      if (x != null) return x;
+} } }
+    \end{lstlisting}
+    \caption{The Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90}.}
+    \label{fig:HerlihyWing}
+    \vspace{1em}
 
-%{\sc call-enq} orders the invoked operation after all the completed enqueues in the current state, and the rules {\sc ret-enq1}/{\sc ret-enq2} flip the corresponding flag from {\tt PEND} to {\tt COMP} provided that the operation is still present in the current state.
-For dequeue operations, {\sc call-deq} only increments the control point and {\sc ret-deq} checks whether the return value is the same as the one fixed at the linearization point.
-The linearization point rule {\sc lin-deq1} corresponds to the case of a non-empty queue, showing that $lin(deq,d,k)$ is enabled only if $d$ has been added by an enqueue which is minimal in the current happens-before.
-When enabled, it removes the enqueue adding $d$ from the state.
-The linearization point rule {\sc lin-deq2} corresponds to the case of dequeue operations linearized with an {\tt EMPTY} return value.
-
-\begin{figure}[t]
-  \begin{minipage}[t]{0.4\linewidth}
+    \includegraphics[width=7cm]{fig-queue12}
+    \caption{Forward simulation with $AbsQ$. Lines depict operations,
+    and circles depict call, return, and linearization point actions.}
+    \label{fig:queueSim}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.4\linewidth}
     \begin{lstlisting}
@@
 \end{figure}
+
+For readability, we define $AbsQ$ as an abstract state machine, which is given in Fig.~\ref{fig:transitions:AbsQ}.
+The shared state of $AbsQ$ consists of several boolean predicates indicating whether the value added by an enqueue has not been removed yet (the predicate $\mathsf{present})$, whether an enqueue is pending (the predicate $\mathsf{pending}$), the happens-before order (the predicate $\mathsf{before}$), and a function giving the value added by an enqueue (the function $\mathsf{arg}$).
+In the initial state, the domain of every function (predicate) is empty.
+%the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).
+%All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}.
+The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points, denoted by $lin(deq,d,k)$.
+$Lin(deq)$ is the set of all actions $lin(deq,d,k)$.
+%
+The enqueue operations consist of two macro rules: the rule {\tt inv(enq,x,k)} orders the invoked operation after all the completed enqueues present in the current state ({\tt k} is a fresh operation identifier associated to the current operation), and the rule {\tt ret(enq,k)} sets $\mathsf{pending}${\tt (k)} to {\tt false}.
+% provided that the operation is still present in the current state.
+The dequeue operations consist of two rules {\tt inv(deq,k)} and {\tt ret(deq,y,k)} associated to the invocation and respectively, the return action, whose definition is obvious, and the rule {\tt lin(deq,y,k)} corresponding to its linearization point.
+When a non-empty set of pending enqueues are present, the rule {\tt lin(deq,y,k)} is non-deterministic.
+It can set the return value {\tt y} either to {\tt EMPTY}, or to a value {\tt d} $\neq$ {\tt EMPTY}, provided that {\tt d} has been added by an enqueue {\tt k1} which is minimal in the current happens-before.
+In the latter case, $\mathsf{present}${\tt (k1)} is set to {\tt false} to indicate that {\tt k1}'s argument has been removed.
+When there exists at least one completed enqueue whose value has not been removed, the rule {\tt lin(deq,y,k)} sets {\tt y} to the value added by a minimal enqueue (w.r.t.
+the current happens-before).
+
+%{\sc call-enq} orders the invoked operation after all the completed enqueues in the current state, and the rules {\sc ret-enq1}/{\sc ret-enq2} flip the corresponding flag from {\tt PEND} to {\tt COMP} provided that the operation is still present in the current state.
+For dequeue operations, {\sc call-deq} only increments the control point and {\sc ret-deq} checks whether the return value is the same as the one fixed at the linearization point.
+The linearization point rule {\sc lin-deq1} corresponds to the case of a non-empty queue, showing that $lin(deq,d,k)$ is enabled only if $d$ has been added by an enqueue which is minimal in the current happens-before.
+When enabled, it removes the enqueue adding $d$ from the state.
+The linearization point rule {\sc lin-deq2} corresponds to the case of dequeue operations linearized with an {\tt EMPTY} return value.
 

### 2017-05-05 draftCAV/queue.tex
Figure placement.

@@
 If it reaches the end of the array, then it restarts.
-
-\begin{wrapfigure}{l}{5.3cm}
-\vspace{-9mm}
-\begin{lstlisting}
-void enq(int x){
-  i = back++;
-  items[i] = x;
-}
-int deq() {
-  while (1) {
-    range = back - 1;
-    for (int i = 0; i <= range; i++){
-      x = swap(items[i],null);
-      if ( x != null ) return x;
-}}}
-  \end{lstlisting}
-\vspace{-5.5mm}
-\caption{Herlihy \& Wing Queue. We assume that every statement is atomic.}
-\label{fig:HerlihyWing}
-\vspace{-6mm}
-\end{wrapfigure}
-The linearization points of the enqueues are not fixed, they depend on dequeues executing in the future.
-Consider the following trace with two concurrent enqueues (${\tt i}(k)$ represents the value of {\tt i} in operation $k$): $inv(enq,x,1)$, $inv(enq,y,2)$, ${\tt i}(1) = \mbox{{\tt bck++}}$, ${\tt i}(2) = \mbox{{\tt bck++}}$, ${\tt items[i(}2{\tt )]} = y$.
-%\vspace{-1.5mm}
-%\begin{align*}
-%inv(enq,x,1)\ \ \ inv(enq,y,2)\ \ \ {\tt i}(1) = \mbox{{\tt bck++}}\ \ \ {\tt i}(2) = \mbox{{\tt bck++}}\ \ \ {\tt items[i(}2{\tt )]} = y
-%\vspace{-1.5mm}
-%\end{align*}
-%\begin{align*}
-%inv(enq,x,1)\ inv(enq,y,2)\ ({\tt i}_1 = 0,{\tt bck} = 1)\ ({\tt i}_2 = 1,{\tt bck} = 2)\ ({\tt items[1]} = y)
-%\end{align*}
-Assuming that the linearization point corresponds to the assignment of {\tt i}, the history of this trace should be linearized to $inv(enq,x,1)$, $ret(enq,1)$, $inv(enq,y,2)$, $ret(enq,2)$.
-However, a dequeue executing until completion after this trace will return $y$ (only position $1$ is filled in the array {\tt items}) which is not consistent with this linearization.
-On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i(}1{\tt )]} = x$ and a completed dequeue that in this case returns $x$, leads to the incorrect linearization: $inv(enq,y,2)$, $ret(enq,2)$, $inv(enq,x,1)$, $ret(enq,1)$, $inv(deq,3)$, $ret(deq,x,3)$.
-%\vspace{-1.5mm}
-%\begin{align*}
-%inv(enq,y,2)\ ret(enq,2) inv(enq,x,1)\ ret(enq,1)\ inv(deq,3)\ ret(deq,x,3).
-%\vspace{-1.5mm}
-%\end{align*}
-
-The dequeue method has a fixed linearization point which corresponds to an execution of {\tt swap} returning a non-null value.
-This action alone contributes to the effect of that value being removed from the queue.
-Every concurrent history can be linearized to a sequential history where dequeues occur in the order of their linearization points in the concurrent history.
-This claim is formally proved in Section~\ref{ssec:HerlihyWing}.
-
-Since the linearization points of the enqueues are determined by future dequeue invocations, there exists no forward simulation from $\mathit{HWQ}$ to $AbsQ_0$.
-In the following, we describe the abstract implementation $AbsQ$ for which such a forward simulation does exist.
-
-\vspace{-3.5mm}
-\subsection{Abstract Queue Implementation}
-\vspace{-1.5mm}
-
-%\textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too.
-We should be able to return EMPTY if there are only pending nodes.
-Consider the following history of $AbsQ_0$: $inv(enq, d_1,k_1), inv(enq, d_2,k_2), inv(deq, k_3), lin(deq, \texttt{EMPTY}, k_3)$.
-This history should be reflected in $AbsQ$ by enabling lin \texttt{EMPTY} of dequeue when there are pending nodes.
-We also need to update the rules in figure.}
-Informally, $AbsQ$ records the set of enqueue operations, whose argument has not yet been removed by a matching dequeue operation.
-In addition, it records the happens-before order between those enqueue operations: this is a partial order ordering an enqueue $k_1$ before another enqueue $k_2$ iff $k_1$ returned before $k_2$ was invoked.
-%Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation.
-The linearization point of a dequeue can either remove a minimal enqueue $k$ (w.r.t.
-the happens-before stored in the state) and fix the return value to the value $d$ added by $k$, or fix the return value to ${\tt EMPTY}$ provided that the current state stores only pending enqueues (intuitively, the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
-%with return value $d\neq{\tt EMPTY}$ is enabled only if the happens-before stored in the current state contains a minimal enqueue that adds the value $d$.
-The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
-%When
-%%\noindent
-%the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueues (because, the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
-%The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.
-
-Fig.~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
-The state of $AbsQ$ after each action is pictured as a graph below the action.
-The nodes of this graph represent enqueue operations and the edges happens-before constraints.
-Each node is labeled by a value (the argument of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed.
-For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with argument $y$.
-Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.
 
@@
 \end{figure}
+
+The linearization points of the enqueues are not fixed, they depend on dequeues executing in the future.
+Consider the following trace with two concurrent enqueues (${\tt i}(k)$ represents the value of {\tt i} in operation $k$): $inv(enq,x,1)$, $inv(enq,y,2)$, ${\tt i}(1) = \mbox{{\tt bck++}}$, ${\tt i}(2) = \mbox{{\tt bck++}}$, ${\tt items[i(}2{\tt )]} = y$.
+%\vspace{-1.5mm}
+%\begin{align*}
+%inv(enq,x,1)\ \ \ inv(enq,y,2)\ \ \ {\tt i}(1) = \mbox{{\tt bck++}}\ \ \ {\tt i}(2) = \mbox{{\tt bck++}}\ \ \ {\tt items[i(}2{\tt )]} = y
+%\vspace{-1.5mm}
+%\end{align*}
+%\begin{align*}
+%inv(enq,x,1)\ inv(enq,y,2)\ ({\tt i}_1 = 0,{\tt bck} = 1)\ ({\tt i}_2 = 1,{\tt bck} = 2)\ ({\tt items[1]} = y)
+%\end{align*}
+Assuming that the linearization point corresponds to the assignment of {\tt i}, the history of this trace should be linearized to $inv(enq,x,1)$, $ret(enq,1)$, $inv(enq,y,2)$, $ret(enq,2)$.
+However, a dequeue executing until completion after this trace will return $y$ (only position $1$ is filled in the array {\tt items}) which is not consistent with this linearization.
+On the other hand, assuming that enqueues should be linearized at the assignment of {\tt items[i]} and extending the trace with ${\tt items[i(}1{\tt )]} = x$ and a completed dequeue that in this case returns $x$, leads to the incorrect linearization: $inv(enq,y,2)$, $ret(enq,2)$, $inv(enq,x,1)$, $ret(enq,1)$, $inv(deq,3)$, $ret(deq,x,3)$.
+%\vspace{-1.5mm}
+%\begin{align*}
+%inv(enq,y,2)\ ret(enq,2) inv(enq,x,1)\ ret(enq,1)\ inv(deq,3)\ ret(deq,x,3).
+%\vspace{-1.5mm}
+%\end{align*}
+
+The dequeue method has a fixed linearization point which corresponds to an execution of {\tt swap} returning a non-null value.
+This action alone contributes to the effect of that value being removed from the queue.
+Every concurrent history can be linearized to a sequential history where dequeues occur in the order of their linearization points in the concurrent history.
+This claim is formally proved in Section~\ref{ssec:HerlihyWing}.
+
+Since the linearization points of the enqueues are determined by future dequeue invocations, there exists no forward simulation from $\mathit{HWQ}$ to $AbsQ_0$.
+In the following, we describe the abstract implementation $AbsQ$ for which such a forward simulation does exist.
+
+\vspace{-3.5mm}
+\subsection{Abstract Queue Implementation}
+\vspace{-1.5mm}
+
+%\textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too.
+We should be able to return EMPTY if there are only pending nodes.
+Consider the following history of $AbsQ_0$: $inv(enq, d_1,k_1), inv(enq, d_2,k_2), inv(deq, k_3), lin(deq, \texttt{EMPTY}, k_3)$.
+This history should be reflected in $AbsQ$ by enabling lin \texttt{EMPTY} of dequeue when there are pending nodes.
+We also need to update the rules in figure.}
+Informally, $AbsQ$ records the set of enqueue operations, whose argument has not yet been removed by a matching dequeue operation.
+In addition, it records the happens-before order between those enqueue operations: this is a partial order ordering an enqueue $k_1$ before another enqueue $k_2$ iff $k_1$ returned before $k_2$ was invoked.
+%Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation.
+The linearization point of a dequeue can either remove a minimal enqueue $k$ (w.r.t.
+the happens-before stored in the state) and fix the return value to the value $d$ added by $k$, or fix the return value to ${\tt EMPTY}$ provided that the current state stores only pending enqueues (intuitively, the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
+%with return value $d\neq{\tt EMPTY}$ is enabled only if the happens-before stored in the current state contains a minimal enqueue that adds the value $d$.
+The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state.
+%When
+%%\noindent
+%the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores only pending enqueues (because, the dequeue overlaps with all the enqueue operations stored in the current state and it can be linearized before all of them).
+%The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.
+
+Fig.~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points).
+The state of $AbsQ$ after each action is pictured as a graph below the action.
+The nodes of this graph represent enqueue operations and the edges happens-before constraints.
+Each node is labeled by a value (the argument of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed.
+For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with argument $y$.
+Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.
 

### 2017-05-05 draftCAV/queue.tex
Updated queue simulation figures.
Figure placement.

@@
 the time of their linearization points, or {\tt EMPTY}. Some implementations,
-like the queue of \citet{journals/toplas/HerlihyW90}, denoted {\sc hwq}, are not
-forward-simulated by $AbsQ_0$, even though they refine $AbsQ_0$, since the order
-in which their enqueues are linearized to form $AbsQ_0$’s sequence is not
-determined until later, when their values are dequeued.
-
-In this section we develop an abstract queue implementation, denoted $AbsQ$,
-which maintains a partial order of enqueues, rather than a linear sequence.
-Since $AbsQ$ does not force refining implementations to eagerly pick among
-linearizations of their enqueues, it forward-simulates many more queue
-implementations. In fact, $AbsQ$ forward-simulates all queue implementations of
-which we are aware that are not forward-simulated by $AbsQ_0$, including {\sc
-hwq}, The Baskets Queue~\cite{DBLP:conf/opodis/HoffmanSS07}, The Linked
-Concurrent Ring Queue ({\sc lcrq})~\cite{DBLP:conf/ppopp/MorrisonA13}, and The
-Time-Stamped Queue~\cite{DBLP:conf/popl/DoddsHK15}.
-
-%TODO HOW CAN WE CONVINCE THAT OTHER IMPLEMENTATIONS HAVE THE SAME PROPERTY ?
-\vspace{-3.5mm}
-\subsection{Enqueue Methods With Non-Fixed Linearization Points}
-\vspace{-1mm}
-% describe a queue implementation listed in Figure~\ref{fig:HerlihyWing} and known as the Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), where only the dequeue methods have fixed linearization points.
-We describe $\mathit{HWQ}$ where the linearization points of the enqueue methods are not fixed.
-The shared state consists of an array {\tt items} storing the values in the queue and a counter {\tt back} storing the index of the first unused position in {\tt items}.
-Initially, all the positions in the array are {\tt null} and {\tt back} is 0.
-An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueues) and then, stores the argument {\tt x} at this position.
-The dequeue method traverses the array {\tt items} starting from the beginning and atomically swaps {\tt null} with the encountered value.
-If the value is not {\tt null}, then the dequeue returns that value.
-If it reaches the end of the array, then it restarts.
+like the queue of \citet{journals/toplas/HerlihyW90}, denoted {\sc hwq} and
+listed in Figure~\ref{fig:HerlihyWing}, are not forward-simulated by $AbsQ_0$,
+even though they refine $AbsQ_0$, since the order in which their enqueues are
+linearized to form $AbsQ_0$’s sequence is not determined until later, when their
+values are dequeued.
 
@@
 
-    \includegraphics[width=7cm]{fig-queue12}
+    \includegraphics[width=7cm]{fig-queue1}
+    \includegraphics[width=7cm]{fig-queue2}
     \caption{Forward simulation with $AbsQ$. Lines depict operations,
@@
 \end{figure}
+
+In this section we develop an abstract queue implementation, denoted $AbsQ$,
+which maintains a partial order of enqueues, rather than a linear sequence.
+Since $AbsQ$ does not force refining implementations to eagerly pick among
+linearizations of their enqueues, it forward-simulates many more queue
+implementations. In fact, $AbsQ$ forward-simulates all queue implementations of
+which we are aware that are not forward-simulated by $AbsQ_0$, including {\sc
+hwq}, The Baskets Queue~\cite{DBLP:conf/opodis/HoffmanSS07}, The Linked
+Concurrent Ring Queue ({\sc lcrq})~\cite{DBLP:conf/ppopp/MorrisonA13}, and The
+Time-Stamped Queue~\cite{DBLP:conf/popl/DoddsHK15}.
+
+%TODO HOW CAN WE CONVINCE THAT OTHER IMPLEMENTATIONS HAVE THE SAME PROPERTY ?
+\vspace{-3.5mm}
+\subsection{Enqueue Methods With Non-Fixed Linearization Points}
+\vspace{-1mm}
+% describe a queue implementation listed in Figure~\ref{fig:HerlihyWing} and known as the Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), where only the dequeue methods have fixed linearization points.
+We describe $\mathit{HWQ}$ where the linearization points of the enqueue methods are not fixed.
+The shared state consists of an array {\tt items} storing the values in the queue and a counter {\tt back} storing the index of the first unused position in {\tt items}.
+Initially, all the positions in the array are {\tt null} and {\tt back} is 0.
+An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueues) and then, stores the argument {\tt x} at this position.
+The dequeue method traverses the array {\tt items} starting from the beginning and atomically swaps {\tt null} with the encountered value.
+If the value is not {\tt null}, then the dequeue returns that value.
+If it reaches the end of the array, then it restarts.
 

### 2017-05-05 draftCAV/queue.tex
Separating stacked figures.
Figure separation.
Locations in the abstract state machine rules.

@@
     \label{fig:HerlihyWing}
-    \vspace{1em}
+
+    \smallskip
+    \hrule
+    \smallskip
 
@@
   \hfill
-  \begin{minipage}[b]{0.4\linewidth}
+  \begin{minipage}[b]{0.38\linewidth}
     \begin{lstlisting}
-function present: $\mathbb{O} \to \mathbb{B}$
-function pending: $\mathbb{O} \to \mathbb{B}$
-function arg: $\mathbb{V} \to \mathbb{O}$
+function loc: $\mathbb{O} \to \set{\tt inv, lin, ret, \bot}$
+function arg, ret: $\mathbb{O} \to \mathbb{V}$
+function present, pending: $\mathbb{O} \to \mathbb{B}$
 function before: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$
 
-rule inv(enq,v):
-  let k := new $\mathbb{O}$
+rule inv(enq,v,k):
+  arg(k) := v
   present(k) := true
   pending(k) := true
-  arg(v) := k
-  forall k' with present(k'):
-    if $\lnot$pending(k'):
-      before(k',k) := true
+  forall k1 with present(k1):
+    if $\lnot$pending(k1):
+      before(k1,k) := true
 
-rule lin(enq,v):
+rule ret(enq,k):
+  pending(k) := false
+
+rule inv(deq,k):
   pass
 
-rule ret(enq,v):
-  let k := arg(v)
-  pending(k) := false
+rule lin(deq,v,k):
+  ret(k) := v
+  if v = EMPTY:
+    forall k' with present(k'):
+      assert pending(k')
+  else:
+    let k1 = arg$^{-1}$(v)
+    assert present(k1)
+    forall k2 with present(k2):
+      assert $\lnot$before(k2,k1)
+    present(k1) := false
 
-rule inv(deq,v):
-  pass
-
-rule lin(deq,v):
-  if v = EMPTY:
-    forall k with present(k):
-      assert pending(k)
-  else:
-    let k := arg(v)
-    assert present(k)
-    forall k' with present(k'):
-      assert $\lnot$before(k', k)
-    present(k) := false
-
-rule ret(deq,v):
-  pass
+rule ret(deq,v,k):
+  assert ret(k) = v
     \end{lstlisting}
-    \caption{The $AbsQ$ implementation.}
+    \caption{The $AbsQ$ implementation; each rule \lstinline|$\alpha$(\_,k)|
+    implicitly begins with \lstinline|assert loc(k)=$\alpha$| and ends with
+    the appropriate \lstinline|loc(k):=$\beta$|.}
     \label{fig:transitions:AbsQ}

### 2017-05-05 draftCAV/queue.tex
first part of H&Q proof

@@
 We refer to such enqueues in $s$ as $\mathsf{present}$ enqueues.
-Also, $\mathsf{pending}(k)$ is ${\tt true}$ in $t$ whenever $k$ is a pending enqueue in $s$, $\mathsf{arg}(d)=k$ in $t$ whenever the argument of the enqueue $k$ in $s$ is $d$, and for every dequeue operation $k$ such that ${\tt x}(k)=d\neq {\tt null}$, we have that ${\tt y}(k)=d$ (recall that ${\tt y}$ is a local variable of the dequeue method in $AbsQ$).
+Also, $\mathsf{pending}(k)$ is ${\tt true}$ in $t$ whenever $k$ is a pending enqueue in $s$, $\mathsf{arg}(k)=d$ in $t$ whenever the argument of the enqueue $k$ in $s$ is $d$, and for every dequeue operation $k$ such that ${\tt x}(k)=d\neq {\tt null}$, we have that ${\tt y}(k)=d$ (recall that ${\tt y}$ is a local variable of the dequeue method in $AbsQ$).
 The order relation $\mathsf{before}$ in $t$ satisfies the following constraints:

### 2017-05-05 draftCAV/stack.tex
Updated AbsS figure.

@@
 
-\begin{figure} [t]
-%\vspace{-2mm}
-\hspace{-15mm}
-\begin{minipage}[t]{7.5cm}
-\scriptsize{
-{	%\centering
-	\begin{align*}
-%		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
-		& \text{\underline{Shared State}} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{maxAtInvoc} : \<Ops> \rightarrow 2^{\<Ops>} \\
-		& \text{{\ttfamily \bfseries function}}\ \mathsf{overlap} : \<Ops> \rightarrow 2^{\<Ops>} \\
-%		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
-%		& \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} &\mbox{ // the argument of an enqueue} \\
-	\end{align*}
-}}
-\end{minipage}
-\begin{minipage}[t]{7.3cm}
+\begin{figure}[t]
+  \begin{minipage}[t]{0.39\linewidth}
     \begin{lstlisting}
-void push($\<Vals>$ x) {
-  atomic rule inv(push,x,k = new);
-  atomic rule ret(push,k);
-}
+function loc: $\mathbb{O} \to \set{\tt inv, com, ret, \bot}$
+function arg, ret: $\mathbb{O} \to \mathbb{V}$
+function present, pending: $\mathbb{O} \to \mathbb{B}$
+function before: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$
+function maxAtInvoc: $\mathbb{O} \to \pow{\mathbb{O}}$
+function overlap: $\mathbb{O} \to \pow{\mathbb{O}}$
 
-$\<Vals>$ pop() {
-  atomic rule inv(pop,k = new);
-  atomic rule com(pop,y,k);
-  atomic rule ret(pop,y,k);
-}
+rule inv(push,x,k):
+  present(k) := true
+  pending(k) := true
+  arg(k) := x
+  forall k1 with present(k1):
+    if $\lnot$pending(k1):
+      before(k1,k) := true
+  forall k1 with pending(k1):
+    overlap(k1) := overlap(k1)$\cup$$\{$k$\}$
+
+rule ret(push,k):
+  pending(k) := false
+
+rule ret(pop,y,k):
+  assert ret(k) = y
     \end{lstlisting}
   \end{minipage}
+  \begin{minipage}[t]{0.59\linewidth}
+    \begin{lstlisting}
+rule inv(pop,k):
+  forall k1 with present(k1):
+    if pending(k1):
+      overlap(k) := overlap(k) $\cup$ $\{$k1$\}$
+    else:
+      if $\forall$k2.present(k2)$\land$before(k1,k2)$\Rightarrow$pending(k2):
+        maxAtInvoc(k) = maxAtInvoc(k) $\cup$ $\{$k1$\}$
 
-\vspace{-5mm}
-\begin{minipage}[t]{6.2cm}
-\begin{lstlisting}
-rule inv(push,x,k) = {
-  $\mathsf{present}$(k) = true;
-  $\mathsf{pending}$(k) = true;
-  $\mathsf{arg}$(k) = x;
-  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1) do
-    $\mathsf{before}$(k1,k) = true;
-  forall k1 do
-    $\mathsf{overlap}$(k1) = $\mathsf{overlap}$(k1) $\cup$ $\{$k$\}$;
-}
-rule ret(push,k) = {
-  $\mathsf{pending}$(k) = false;
-  return;
-}
+rule com(pop,y,k):
+  ret(k) := y
+  if y = EMPTY:
+    assert maxAtInvoc(k) = $\emptyset$
+  else:
+    let k1 := arg$^{-1}$(y)
+    assert present(k1)
+    assert k1 $\in$ maxAtInvoc(k) $\cup$ overlap(k)
+    present(k1) := false
+    forall k2 with k1 $\in$ maxAtInvoc(k2):
+      maxAtInvoc(k2) := maxAtInvoc(k2) $\setminus$ $\{$k1$\}$
+      forall k3 with before$_1$(k3,k1):
+        if $\forall$k4$\neq$k1.before$_1$(k3,k4) $\Rightarrow$ k4$\in$overlap(k2):
+          maxAtInvoc(k2) := maxAtInvoc(k2) $\cup$ $\{$k3$\}$
     \end{lstlisting}
   \end{minipage}
-\begin{minipage}[t]{5cm}
-%\vspace{-3mm}
-\begin{lstlisting}
-rule inv(pop,k) = {
-  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1)$\land$
-                 $\neg \exists k'.\ \mathsf{present}(k')\land \neg \mathsf{pending}(k')\,\land$ 
-                 $\mathsf{before}$(k1,$k'$) do
-    $\mathsf{maxAtInvoc}$(k) = $\mathsf{maxAtInvoc}$(k) $\cup$ $\{$k1$\}$;
-  forall k1 with $\mathsf{present}$(k1)$\land \mathsf{pending}$(k1) do
-    $\mathsf{overlap}$(k) = $\mathsf{overlap}$(k) $\cup$ $\{$k1$\}$;
-}
-rule ret(pop,y,k) = {
-  return y;
-}
-\end{lstlisting}
-\end{minipage}
-
-\vspace{-4mm}
-\begin{lstlisting}
-rule com(pop,y,k) = {
-  choose a $\in$ {true, false};
-  if ((a $\vee$ $\forall k'.\ \mathsf{present}(k')$=false) $\land$ $\mathsf{maxAtInvoc}$(k)=$\,\emptyset$) 
-    y = EMPTY;
-  else 
-    choose k1 $\in\<Ops>$ with $\mathsf{present}$(k1)$\land$ k1$\in\mathsf{maxAtInvoc}$(k)$\cup\,\mathsf{overlap}$(k);
-    $\mathsf{present}$(k1) = false;
-    y = $\mathsf{arg}$(k1);
-    forall k2 with k1$\in\mathsf{maxAtInvoc}$(k2) do
-      $\mathsf{maxAtInvoc}$(k2) = $\mathsf{maxAtInvoc}$(k2)$\setminus\{$k1$\}$;
-      forall k3 with $\mathsf{before_1}$(k3,k1)$\land \forall k'.\ (\mathsf{before_1}($k3$,k')\land k'\neq $k1$)\Rightarrow k'\in \mathsf{overlap}($k2$)$ do
-        $\mathsf{maxAtInvoc}$(k2) = $\mathsf{maxAtInvoc}$(k2)$\cup\{$k3$\}$;
-}
-\end{lstlisting}
- %\neg\exists k'.\ \mathsf{present}(k')\,=\,$true$\,\land\, \mathsf{before}($k3,$k')\land \mathsf{before}(k',$k1$)$
-
-%{\scriptsize
-%  \centering
-%  \begin{mathpar}
-%    \inferrule[call-push]{
-%      k\not\in dom(cp) \\ 
-%      d\neq {\tt EMPTY} \\
-%      \forall k'.\ ov'(k') = ov(k')\cup \{k\}
-%    }{
-%      O,<,\ell,rv,cp,be,ov 
-%      \xrightarrow{inv(push,d,k)} 
-%      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1],be,ov'
-%    }\hspace{5mm}
-%
-%    \vspace{-1mm}
-%    \inferrule[call-pop]{
-%      k\not\in dom(cp) %\\ k\not\in dom(cp) \\ k\not\in dom(cp)
-%    }{
-%      O,<,\ell,rv,cp,be,ov
-%      \xrightarrow{inv(pop,k)} 
-%      O,<,\ell,rv,cp[k\mapsto R_1],be[k\mapsto maxCo(O)],ov[k\mapsto {\tt PEND}(O)]
-%    }\hspace{5mm}
-%    
-%    \vspace{-1mm}
-%    \inferrule[ret-pop]{
-%       cp(k) = R_2 \\
-%       rv(k)=d  
-%    }{
-%      O,<,\ell,rv,cp,be,ov
-%      \xrightarrow{ret(pop,d,k)}
-%      O,<,\ell,rv,cp[k\mapsto R_3],be,ov
-%    }\hspace{5mm}
-%
-%    \vspace{-1mm}
-%    \inferrule[ret-push1]{
-%      cp(k) = A_1 \\
-%      k \in O \\
-%      \ell(k) = (d,{\tt PEND}) 
-%    }{
-%      O,<,\ell,rv,cp,be,ov
-%      \xrightarrow{ret(push,k)}
-%      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2],be,ov
-%    }\hspace{5mm}
-%    \inferrule[ret-push2]{
-%      cp(k) = A_1 \\
-%      k \not\in O 
-%    }{
-%      O,<,\ell,rv,cp,be,ov
-%      \xrightarrow{ret(push,k)}
-%      O,<,\ell,rv,cp[k\mapsto A_2],be,ov
-%    }\hspace{5mm}
-%
-%    \inferrule[com-pop1]{
-%       cp(k) = R_1 \\
-%       d\neq{\tt EMPTY} \\
-%       k'\in be(k)\cup ov(k) \\
-%        \ell_1(k')=d \\
-%       \forall k_1.\ k'\not\in be(k_1) \Rightarrow be'(k_1)=be(k_1) \\     \\
-%       \forall k_1.\ k'\in be(k_1) \Rightarrow be'(k_1)=(be(k_1)\setminus\{k'\})\cup \{k_2: k_2\in pred_{<}(k') \land \forall k_3. (k_2\in pred_{<}(k_3) \land k_3\neq k') => k_3\in ov(k_1)\} \\
-%    }{
-%      O,<,\ell,rv,cp,be,ov
-%      \xrightarrow{com(pop,d,k)} \\
-%      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2],be',ov %be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}]
-%    }\hspace{5mm}
-%
-%    %\vspace{-1mm}
-%    \inferrule[com-pop2]{
-%       cp(k) = R_1 \\
-%       be(k)=\emptyset
-%    }{
-%      O,<,\ell,rv,cp,be,ov
-%      \xrightarrow{com(pop,{\tt EMPTY},k)}
-%      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2],be,ov
-%    }\hspace{5mm}
-%
-%    
-%      \end{mathpar}
-%  }
- \vspace{-4mm}
-  \caption{The abstract stack implementation $AbsS$.
-We use $\mathsf{before_1}$ to denote the immediate successor relation induced by $\mathsf{before}$, i.e., $\mathsf{before}_1(k,k')={\tt true}$ iff $\neg\exists k''.\ \mathsf{present}(k'')\land \mathsf{before}(k,k'')\land \mathsf{before}(k'',k')$.
-%  The transition relation of $AbsQ$.
-We use the following notions: $maxCo(O)$ is the set of greatest operations in $O$ (w.r.t.
-$<$) which are completed, i.e., $maxCo(O)=\set{k\in O: \ell_2(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_2(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.%\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise.
-If I understand correctly, there is a problem change in $be$ in lin-pop1 rule.
-What I understand is we move $k''$ to every predecesoor of $k'$.
-But we should not move it to predecessor of $k'$ if this predecessor has a successor $m$ such that $m \in be(k'')$.
-Is this case reflected in your definition.
-Please read sub-bullet 4 of bullet 4 in the definition of $L_I$ in my document.
-Lin-pop2 seems ok for me.
-Last note: Bibliography seems not correct to me.}  $f[\forall k.\ k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=expr(k)$,   $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and
-  }
+  \caption{The $AbsS$ implementation; each rule \lstinline|$\alpha$(\_,k)|
+  implicitly begins with \lstinline|assert loc(k)=$\alpha$| and ends with the
+  appropriate \lstinline|loc(k):=$\beta$|; and \lstinline|before$_1$| denotes
+  the transitive reduction of \lstinline|before|.}
   \label{fig:transitions:AbsS}
-\vspace{-5mm}
 \end{figure}

### 2017-05-05 draftCAV/stack.tex
TSS figure.

@@
 \vspace{-1mm}
-\begin{wrapfigure}{l}{5.2cm}
-\vspace{-10.5mm}
+
+We explain the meaning of the commit points on a simplified version of the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) given in Fig.~\ref{fig:TimeStamped}.
+This  implementation maintains an array of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
+Initially, each list contains a sentinel dummy node pointing to itself with timestamp $-1$ and the flag {\tt taken} set to {\tt false}.
+
+\begin{figure}[t]
+  \begin{minipage}[t]{0.49\linewidth}
     \begin{lstlisting}
@@
   Node* next;
-  boolean taken;
+  bool taken;
 };
+
+bool CAS(bool data, bool a, bool b);
+
 Node* pools[maxThreads];
@@
 void push(int x) {
-  Node* n = new Node(x,MAX_INT,
-                        null,false);
+  Node* n =
+    new Node(x,MAX_INT, null,false);
   n->next = pools[myTID];
@@
 }
+    \end{lstlisting}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[t]{0.49\linewidth}
+    \begin{lstlisting}
 int pop() {
- boolean success = false;
+  bool success = false;
   int maxTS = -1;
@@
     maxTS = -1; youngest = null;
-   for(int i=0; i<maxThreads; i++){
+    for(int i=0; i<maxThreads; i++) {
       Node* n = pools[i];
@@
     if (youngest != null)
-     success=CAS(youngest->taken,
-                       false,true);
+      success =
+        CAS(youngest->taken, false, true);
   }
@@
     \end{lstlisting}
-\vspace{-6mm}
-\caption{Time-Stamped Stack.} %We assume that every statement is atomic.}
+  \end{minipage}
+
+  \caption{The Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15}.}
   \label{fig:TimeStamped}
-\vspace{-7mm}
-\end{wrapfigure}
-We explain the meaning of the commit points on a simplified version of the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) given in Fig.~\ref{fig:TimeStamped}.
-This  implementation maintains an array of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
-Initially, each list contains a sentinel dummy node pointing to itself with timestamp $-1$ and the flag {\tt taken} set to {\tt false}.
+\end{figure}
 

### 2017-05-05 draftCAV/stack.tex
Boolean is a proper name.

@@
 We explain the meaning of the commit points on a simplified version of the Time-Stamped Stack~\cite{DBLP:conf/popl/DoddsHK15} ($\mathit{TSS}$, for short) given in Fig.~\ref{fig:TimeStamped}.
-This  implementation maintains an array of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
+This  implementation maintains an array of singly-linked lists, one for each thread, where list nodes contain a data value (field {\tt data}), a timestamp (field {\tt ts}), the next pointer (field {\tt next}), and a Boolean flag indicating whether the node represents a value removed from the stack (field {\tt taken}).
 Initially, each list contains a sentinel dummy node pointing to itself with timestamp $-1$ and the flag {\tt taken} set to {\tt false}.

### 2017-05-08 draftCAV/stack.tex
Updated TSS subsection of Stacks Section
Adapted section to new notation and added intuition on F_2 relation.

--- 2017-05-05 draftCAV/stack.tex
@@
 
-\begin{figure}[t]
-  \begin{minipage}[t]{0.39\linewidth}
+\begin{figure} [t]
+%\vspace{-2mm}
+\hspace{-15mm}
+\begin{minipage}[t]{7.5cm}
+\scriptsize{
+{	%\centering
+	\begin{align*}
+%		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
+		& \text{\underline{Shared State}} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{maxAtInvoc} : \<Ops> \rightarrow 2^{\<Ops>} \\
+		& \text{{\ttfamily \bfseries function}}\ \mathsf{overlap} : \<Ops> \rightarrow 2^{\<Ops>} \\
+%		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
+%		& \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} &\mbox{ // the argument of an enqueue} \\
+	\end{align*}
+}}
+\end{minipage}
+\begin{minipage}[t]{7.3cm}
 \begin{lstlisting}
-function loc: $\mathbb{O} \to \set{\tt inv, com, ret, \bot}$
-function arg, ret: $\mathbb{O} \to \mathbb{V}$
-function present, pending: $\mathbb{O} \to \mathbb{B}$
-function before: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$
-function maxAtInvoc: $\mathbb{O} \to \pow{\mathbb{O}}$
-function overlap: $\mathbb{O} \to \pow{\mathbb{O}}$
+void push($\<Vals>$ x) {
+  atomic rule inv(push,x,k = new);
+  atomic rule ret(push,k);
+}
 
-rule inv(push,x,k):
-  present(k) := true
-  pending(k) := true
-  arg(k) := x
-  forall k1 with present(k1):
-    if $\lnot$pending(k1):
-      before(k1,k) := true
-  forall k1 with pending(k1):
-    overlap(k1) := overlap(k1)$\cup$$\{$k$\}$
-
-rule ret(push,k):
-  pending(k) := false
-
-rule ret(pop,y,k):
-  assert ret(k) = y
+$\<Vals>$ pop() {
+  atomic rule inv(pop,k = new);
+  atomic rule com(pop,y,k);
+  atomic rule ret(pop,y,k);
+}
 \end{lstlisting}
 \end{minipage}
-  \begin{minipage}[t]{0.59\linewidth}
+
+\vspace{-5mm}
+\begin{minipage}[t]{6.2cm}
 \begin{lstlisting}
-rule inv(pop,k):
-  forall k1 with present(k1):
-    if pending(k1):
-      overlap(k) := overlap(k) $\cup$ $\{$k1$\}$
-    else:
-      if $\forall$k2.present(k2)$\land$before(k1,k2)$\Rightarrow$pending(k2):
-        maxAtInvoc(k) = maxAtInvoc(k) $\cup$ $\{$k1$\}$
-
-rule com(pop,y,k):
-  ret(k) := y
-  if y = EMPTY:
-    assert maxAtInvoc(k) = $\emptyset$
-  else:
-    let k1 := arg$^{-1}$(y)
-    assert present(k1)
-    assert k1 $\in$ maxAtInvoc(k) $\cup$ overlap(k)
-    present(k1) := false
-    forall k2 with k1 $\in$ maxAtInvoc(k2):
-      maxAtInvoc(k2) := maxAtInvoc(k2) $\setminus$ $\{$k1$\}$
-      forall k3 with before$_1$(k3,k1):
-        if $\forall$k4$\neq$k1.before$_1$(k3,k4) $\Rightarrow$ k4$\in$overlap(k2):
-          maxAtInvoc(k2) := maxAtInvoc(k2) $\cup$ $\{$k3$\}$
+rule inv(push,x,k) = {
+  $\mathsf{present}$(k) = true;
+  $\mathsf{pending}$(k) = true;
+  $\mathsf{arg}$(k) = x;
+  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1) do
+    $\mathsf{before}$(k1,k) = true;
+  forall k1 do
+    $\mathsf{overlap}$(k1) = $\mathsf{overlap}$(k1) $\cup$ $\{$k$\}$;
+}
+rule ret(push,k) = {
+  $\mathsf{pending}$(k) = false;
+  return;
+}
 \end{lstlisting}
 \end{minipage}
-  \caption{The $AbsS$ implementation; each rule \lstinline|$\alpha$(\_,k)|
-  implicitly begins with \lstinline|assert loc(k)=$\alpha$| and ends with the
-  appropriate \lstinline|loc(k):=$\beta$|; and \lstinline|before$_1$| denotes
-  the transitive reduction of \lstinline|before|.}
+\begin{minipage}[t]{5cm}
+%\vspace{-3mm}
+\begin{lstlisting}
+rule inv(pop,k) = {
+  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1)$\land$
+                 $\neg \exists k'.\ \mathsf{present}(k')\land \neg \mathsf{pending}(k')\,\land$ 
+                 $\mathsf{before}$(k1,$k'$) do
+    $\mathsf{maxAtInvoc}$(k) = $\mathsf{maxAtInvoc}$(k) $\cup$ $\{$k1$\}$;
+  forall k1 with $\mathsf{present}$(k1)$\land \mathsf{pending}$(k1) do
+    $\mathsf{overlap}$(k) = $\mathsf{overlap}$(k) $\cup$ $\{$k1$\}$;
+}
+rule ret(pop,y,k) = {
+  return y;
+}
+\end{lstlisting}
+\end{minipage}
+
+\vspace{-4mm}
+\begin{lstlisting}
+rule com(pop,y,k) = {
+  choose a $\in$ {true, false};
+  if ((a $\vee$ $\forall k'.\ \mathsf{present}(k')$=false) $\land$ $\mathsf{maxAtInvoc}$(k)=$\,\emptyset$) 
+    y = EMPTY;
+  else 
+    choose k1 $\in\<Ops>$ with $\mathsf{present}$(k1)$\land$ k1$\in\mathsf{maxAtInvoc}$(k)$\cup\,\mathsf{overlap}$(k);
+    $\mathsf{present}$(k1) = false;
+    y = $\mathsf{arg}$(k1);
+    forall k2 with k1$\in\mathsf{maxAtInvoc}$(k2) do
+      $\mathsf{maxAtInvoc}$(k2) = $\mathsf{maxAtInvoc}$(k2)$\setminus\{$k1$\}$;
+      forall k3 with $\mathsf{before_1}$(k3,k1)$\land \forall k'.\ (\mathsf{before_1}($k3$,k')\land k'\neq $k1$)\Rightarrow k'\in \mathsf{overlap}($k2$)$ do
+        $\mathsf{maxAtInvoc}$(k2) = $\mathsf{maxAtInvoc}$(k2)$\cup\{$k3$\}$;
+}
+\end{lstlisting}
+ %\neg\exists k'.\ \mathsf{present}(k')\,=\,$true$\,\land\, \mathsf{before}($k3,$k')\land \mathsf{before}(k',$k1$)$
+
+%{\scriptsize
+%  \centering
+%  \begin{mathpar}
+%    \inferrule[call-push]{
+%      k\not\in dom(cp) \\ 
+%      d\neq {\tt EMPTY} \\
+%      \forall k'.\ ov'(k') = ov(k')\cup \{k\}
+%    }{
+%      O,<,\ell,rv,cp,be,ov 
+%      \xrightarrow{inv(push,d,k)} 
+%      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1],be,ov'
+%    }\hspace{5mm}
+%
+%    \vspace{-1mm}
+%    \inferrule[call-pop]{
+%      k\not\in dom(cp) %\\ k\not\in dom(cp) \\ k\not\in dom(cp)
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{inv(pop,k)} 
+%      O,<,\ell,rv,cp[k\mapsto R_1],be[k\mapsto maxCo(O)],ov[k\mapsto {\tt PEND}(O)]
+%    }\hspace{5mm}
+%    
+%    \vspace{-1mm}
+%    \inferrule[ret-pop]{
+%       cp(k) = R_2 \\
+%       rv(k)=d  
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{ret(pop,d,k)}
+%      O,<,\ell,rv,cp[k\mapsto R_3],be,ov
+%    }\hspace{5mm}
+%
+%    \vspace{-1mm}
+%    \inferrule[ret-push1]{
+%      cp(k) = A_1 \\
+%      k \in O \\
+%      \ell(k) = (d,{\tt PEND}) 
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{ret(push,k)}
+%      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2],be,ov
+%    }\hspace{5mm}
+%    \inferrule[ret-push2]{
+%      cp(k) = A_1 \\
+%      k \not\in O 
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{ret(push,k)}
+%      O,<,\ell,rv,cp[k\mapsto A_2],be,ov
+%    }\hspace{5mm}
+%
+%    \inferrule[com-pop1]{
+%       cp(k) = R_1 \\
+%       d\neq{\tt EMPTY} \\
+%       k'\in be(k)\cup ov(k) \\
+%        \ell_1(k')=d \\
+%       \forall k_1.\ k'\not\in be(k_1) \Rightarrow be'(k_1)=be(k_1) \\     \\
+%       \forall k_1.\ k'\in be(k_1) \Rightarrow be'(k_1)=(be(k_1)\setminus\{k'\})\cup \{k_2: k_2\in pred_{<}(k') \land \forall k_3. (k_2\in pred_{<}(k_3) \land k_3\neq k') => k_3\in ov(k_1)\} \\
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{com(pop,d,k)} \\
+%      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2],be',ov %be[\forall k''.\ k'\in be(k'') \Rightarrow k'' \mapsto (be(k'')\cup pred_{<}(k'))\setminus\{k'\}]
+%    }\hspace{5mm}
+%
+%    %\vspace{-1mm}
+%    \inferrule[com-pop2]{
+%       cp(k) = R_1 \\
+%       be(k)=\emptyset
+%    }{
+%      O,<,\ell,rv,cp,be,ov
+%      \xrightarrow{com(pop,{\tt EMPTY},k)}
+%      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2],be,ov
+%    }\hspace{5mm}
+%
+%    
+%      \end{mathpar}
+%  }
+ \vspace{-4mm}
+  \caption{The abstract stack implementation $AbsS$.
+We use $\mathsf{before_1}$ to denote the immediate successor relation induced by $\mathsf{before}$, i.e., $\mathsf{before}_1(k,k')={\tt true}$ iff $\neg\exists k''.\ \mathsf{present}(k'')\land \mathsf{before}(k,k'')\land \mathsf{before}(k'',k')$.
+%  The transition relation of $AbsQ$.
+We use the following notions: $maxCo(O)$ is the set of greatest operations in $O$ (w.r.t.
+$<$) which are completed, i.e., $maxCo(O)=\set{k\in O: \ell_2(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_2(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.%\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise.
+If I understand correctly, there is a problem change in $be$ in lin-pop1 rule.
+What I understand is we move $k''$ to every predecesoor of $k'$.
+But we should not move it to predecessor of $k'$ if this predecessor has a successor $m$ such that $m \in be(k'')$.
+Is this case reflected in your definition.
+Please read sub-bullet 4 of bullet 4 in the definition of $L_I$ in my document.
+Lin-pop2 seems ok for me.
+Last note: Bibliography seems not correct to me.}  $f[\forall k.\ k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=expr(k)$,   $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and
+  }
   \label{fig:transitions:AbsS}
+\vspace{-5mm}
 \end{figure}
@@
 \vspace{-1mm}
-We describe a forward simulation $\mathit{fs}_2$ from $\mathit{TSS}$ to $AbsS$.
-Except for the constraints on the components $be$ and $ov$ of a $AbsS$ state, it is similar to the simulation $\mathit{fs}_1$ from $\mathit{HWQ}$ to $AbsQ$.
-Thus, the $AbsS$ states $t=\tup{O,<,\ell,rv,cp,be,ov}$ associated by $\mathit{fs}_2$ to a $\mathit{TSS}$ state $s$ satisfy the following.
+We describe a forward simulation $F_2$ from $\mathit{TSS}$ to $AbsS$.
+From the description of $\mathit{TSS}$ in Fig.~\ref{fig:TimeStamped}, we can obtain the LTS with global variables {\tt pools}, {\tt TS}; local variables {\tt n}, {\tt youngest} (just for pops) etc.
+and transitions corresponding to statements in $\mathit{TSS}$.
+Except for the $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ fields, states of $AbsS$ LTS is the same as $AbsQ$ with transitions corresponding to rules of pops and pushes depicted in Fig.~\ref{fig:transitions:AbsS}.
+
+Similar to enqueue operations in a $\mathit{HWQ}$ state, if a $\mathit{TSS}$ state $s$ is related to $AbsS$ state $t$ by $F_2$, then for every push operation $k$ in $s$ such that $k$ has not yet added a node to {\tt pools} or its node is still present in {\tt pools} (i.e., the node created by the push has {\tt taken} set to {\tt false}), $\mathsf{present}(k)$ must be true in $t$.
+Similarly, $\mathsf{pending}(k)$ must be true in $t$ if $k$ is invoked but has not yet executed its last statement in $s$.
+%Except for the constraints on the components $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ of a $AbsS$ state, it is similar to the simulation $F_1$ from $\mathit{HWQ}$ to $AbsQ$.
+Thus, the $AbsS$ states $t=\tup{O,<,\ell,rv,cp,\mathsf{maxAtInvoc},\mathsf{overlap}}$ associated by $F_2$ to a $\mathit{TSS}$ state $s$ satisfy the following.
 The set $O$ consists of all the identifiers of pushes in $s$ which haven't added yet a node to {\tt pools} or for which the input is still present in {\tt pools} (i.e., the node created by the push has {\tt taken} set to {\tt false}).
@@
 
-To describe the order relation $<$ we consider the following notations: ${\tt ts}_s(k)$, resp., ${\tt TID}_s(k)$, denotes the timestamp of the node created by the push $k$ in state $s$ (the {\tt ts} field of this node), resp., the id of the thread executing $k$.
+To describe the restrictions of $F_2$ on order relation $\mathsf{before}$ and the sets $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$, we consider the following notations: ${\tt ts}_s(k)$, resp., ${\tt TID}_s(k)$, denotes the timestamp of the node created by the push $k$ in state $s$ (the {\tt ts} field of this node), resp., the id of the thread executing $k$.
 By an abuse of terminology, we call ${\tt ts}_s(k)$ the timestamp of $k$ in state $s$.
@@
 More precisely, $k \leadsto_s k'$ when ${\tt TID}_s(k) < {\tt TID}_s(k')$, or ${\tt TID}_s(k) = {\tt TID}_s(k')$ and the node created by $k'$ is reachable from the one created by $k$ in the list pointed to by ${\tt pools}[{\tt TID}_s(k)]$.
-The order relation $<$ satisfies the following: (1) pending pushes are maximal, (2) $<$ is consistent with the order between node timestamps, i.e., ${\tt ts}_s(k) \leq {\tt ts}_s(k')$ implies $k'\not< k$, and (3) $<$ includes the order between pushes executed in the same thread, i.e., ${\tt TID}_s(k) = {\tt TID}_s(k')$ and ${\tt ts}_s(k) < {\tt ts}_s(k')$ implies $k < k'$.
 
-The components $be$ and $ov$ satisfy the following constraints (their domain is the set of identifiers of pending pops):
+The order relation $\mathsf{before}$ satisfies the following: (1) pending pushes are maximal as in $F_1$, (2) $\mathsf{before}$ is consistent with the order between node timestamps, i.e., ${\tt ts}_s(k) \leq {\tt ts}_s(k')$ implies $\neg \mathsf{before}(k',k)$, and (3) $\mathsf{before}$ includes the order between pushes executed in the same thread, i.e., ${\tt TID}_s(k) = {\tt TID}_s(k')$ and ${\tt ts}_s(k) < {\tt ts}_s(k')$ implies $\mathsf{before}(k < k')$.
+
+The components $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ satisfy the following constraints (their domain is the set of identifiers of pending pops):
 \vspace{-2mm}
 \begin{itemize}
-	\item a pop $k$ with ${\tt youngest}\neq {\tt null}$ that reached a node with timestamp $\tau$ (its variable {\tt n} points to this node) overlaps with every push that created a node with a timestamp bigger than $\tau$ and which occurs in {\tt pools} before the node reached by $k$, i.e., ${\tt youngest}_s(k)\neq{\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k_1)$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
-	\item a pop $k$ with ${\tt youngest}={\tt null}$ overlaps with every push that created a node which occurs in {\tt pools} before the node reached by $k$,
-i.e., ${\tt youngest}_s(k)={\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
-	\item if the variable {\tt youngest} of a pop $k$ points to a node which is not taken, then this node was created by a push in $be(k)\cup ov(k)$ or the node currently reached by $k$ is followed in {\tt pools} by another node which was created by a push in $be(k)\cup ov(k)$, i.e., ${\tt youngest}_s(k)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(k)={\tt n}_s(k_2)$ implies $k_1\in be(k)\cup ov(k)$ or that there exists $k_3\in O$ such that ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in be(k)\cup ov(k)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and $k$ is traversing the last list in the array {\tt pools}, for each $k, k_1,k_2$
+    \item[\emph{Frontiers}] Every maximally-closed (maximal or all successors are pending) push $k$ can be removed by a pending pop $p$ in $t$.
+Formally, the push $k$ is maximally-closed if $\mathsf{present}(k) \land (\mathsf{pending}(k) \lor (\forall k'.
+\mathsf{present}(k') \land \mathsf{before}(k,k') \rightarrow \mathsf{pending}(k'))$.
+Then, for all pending pop operations $p$, $k \in \mathsf{overlap}(p) \cup \mathsf{maxAtInvoc}(p)$.
+	\item[\emph{TraverseBefore}] a pop $p$ with ${\tt youngest}(p)\neq {\tt null}$ that reached the node ${\tt n}$ overlaps with every present push that created a node with a timestamp greater than ${\tt youngest}(p){\tt ->ts}$ and which occurs in  {\tt pools} before the node $n$, i.e., ${\tt youngest}_s(p)= {\tt n}_s(k) \neq{\tt null}$, ${\tt n}_s(p)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, $\mathsf{present}(k_2)$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k)$ implies $k_2\in \mathsf{overlap}(p)$, for each $p, k_1, k_2$
+	%with timestamp $\tau$ (its variable {\tt n} points to this node) overlaps with every push that created a node with a timestamp bigger than $\tau$ and which occurs in {\tt pools} before the node reached by $k$, i.e., ${\tt youngest}_s(k)\neq{\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k_1)$ implies $k_2\in \mathsf{overlap}(k)$, for each $k, k_1, k_2$
+	\item[\emph{TraverseBeforeNull}] a pop $p$ with ${\tt youngest}(p)={\tt null}$ overlaps with every push that created a node which occurs in {\tt pools} before the node reached by $p$,
+i.e., ${\tt youngest}_s(p)={\tt null}$, ${\tt n}_s(p)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and $\mathsf{present}(k_2)$ implies $k_2\in \mathsf{overlap}(p)$, for each $p, k_1, k_2$
+	\item[\emph{TraverseAfter}] if the variable {\tt youngest} of a pop $p$ points to a node which is not taken, then this node was created by a push in $\mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$ or the node currently reached by $p$ is followed in {\tt pools} by another node which was created by a push in $\mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$, i.e., ${\tt youngest}_s(p)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(p)={\tt n}_s(k_2)$ implies $k_1\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$ or that there exists $k_3\in O$ such that $\mathsf{present}(k_3)$, ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and $p$ is on the control point that is before the assignment statement that changes the ${\tt youngest}$ field, for each $p, k_1,k_2$.
 %	
@@
 \end{itemize}
-There are some more constraints on $be$ and $ov$ that can be seen as invariants of $AbsS$, i.e., $be(k)$ and $ov(k)$ don't contain predecessors of pushes from $be(k)$ (for each $k, k_1, k_2$, $k_1 < k_2$ and $k_2 \in be(k)$ implies $k_1\not\in be(k)\cup ov(k)$).
-They can be found in Appendix~\ref{app:tss}.
+There are some more constraints on $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ that can be seen as invariants of $AbsS$, i.e., $\mathsf{maxAtInvoc}(p)$ and $\mathsf{overlap}(p)$ do not contain predecessors of pushes from $\mathsf{maxAtInvoc}(p)$ (for each $p, k_1, k_2$, $\mathsf{before}(k_1, k_2)$ and $k_2 \in \mathsf{maxAtInvoc}(p)$ implies $k_1\not\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$).
+They can be found in the extended version of the paper (\cite{extended}).
 
-Finally, for every pop operation $k$ such that ${\tt success}(k)={\tt true}$, we have that $rv(k)={\tt youngest}(k)\text{\tt->data}$.
+The rationale behind the construction of $F_2$ has some similarities with $F_1$ that relates $\mathit{HWQ}$ and $AbsQ$.
+As in $F_1$, operation identifiers of pending push operations should be maximal in the related $\mathit{TSS}$ states.
+This is ensured by labeling them as $\mathsf{pending}$ in the related $AbsS$ states as in $F_1$.
 
-The proof that $\mathit{fs}_2$ is indeed a forward simulation from $\mathit{TSS}$ to $AbsS$ follows the same lines as the one given for the Herlihy\&Wing Queue.
-It can be found in Appendix~\ref{app:tss}.
+The dual condition of ensuring maximal-closedness of remove candidate push operations is not valid for $F_2$, since linearization points of pop operations are not fixed and they can linearize in any point between their invocation and commit points.
+Hence, any push operation that is maximally-closed between these points are removable by the pop and they might not be maximally-closed in the current state.
+Unfortunately, $\mathit{TSS}$ does not keep history of pop operations in the state and we cannot infer all the push operations that are concurrent or maximally closed with respect to a given pop operation from the state information.
+
+Nevertheless, we can obtain partial information from the state that is enough for showing once-maximality of a push that is removed by commit action of a pop.
+First, the push operations that are maximally-closed in the current state are remove candidates by all pending pops since a pending pop operation may linearize at this point.
+This restriction is described by the \emph{Frontiers} condition.
+
+To obtain further information on $\mathsf{overlap}$ and $\mathsf{maxAtInvoc}$ sets of a pending pop $p$ in state $s$, we look at $youngest_s(p)$ and $n_s(p)$ which show the current youngest element $p$ has seen and the current element $p$ is visiting, respectively.
+\emph{TraverseBefore} and \emph{TraverseBeforeNull} conditions make sure that if threads add new nodes to their pools after $p$ passes them during its travels, these pushes must be concurrent with $p$.
+The \emph{TraverseAfter} condition makes sure that either the current youngest node $youngest_s(p)$ is in $\mathsf{overlap}_s(p) \cup \mathsf{maxAtInvoc}_s(p)$ or there is a future node that $p$ will visit and this future node is in $\mathsf{overlap}_s(p) \cup \mathsf{maxAtInvoc}_s(p)$.
+
+Combining \emph{TraverseBefore} and \emph{TraverseAfter} restrictions we make sure that when a new node $n_s(p)$ is visited, then either $youngest_s(p)$ was maximally-closed before new nodes that precede $n_s(p)$ in the traverse order were added, or there is a successor node of $n_s(p)$ in the traverse order that is maximally-closed.
+Hence, when $p$ finishes its traversal of pools with a non-null and present $youngest$ node, this node is guaranteed to be once-maximal between invocation and commit points of $p$.
+
+The proof that $F_2$ is indeed a forward simulation from $\mathit{TSS}$ to $AbsS$ follows the same lines as the one given for the Herlihy\&Wing Queue. It can be found in \cite{extended}.

### 2017-05-08 draftCAV/stack.tex

--- 2017-05-05 draftCAV/stack.tex
@@
 
-To describe the order relation $<$ we consider the following notations: ${\tt ts}_s(k)$, resp., ${\tt TID}_s(k)$, denotes the timestamp of the node created by the push $k$ in state $s$ (the {\tt ts} field of this node), resp., the id of the thread executing $k$.
+Similar to enqueue operations in a $\mathit{HWQ}$ state, if a $\mathit{TSS}$ state $s$ is related to $AbsS$ state $t$ by $F_2$, then for every push operation $k$ in $s$ such that $k$ has not yet added a node to {\tt pools} or its node is still present in {\tt pools} (i.e., the node created by the push has {\tt taken} set to {\tt false}), $\mathsf{present}(k)$ must be true in $t$.
+Similarly, $\mathsf{pending}(k)$ must be true in $t$ if $k$ is invoked but has not yet executed its last statement in $s$.
+%Except for the constraints on the components $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ of a $AbsS$ state, it is similar to the simulation $F_1$ from $\mathit{HWQ}$ to $AbsQ$.
+Thus, the $AbsS$ states $t=\tup{O,<,\ell,rv,cp,\mathsf{maxAtInvoc},\mathsf{overlap}}$ associated by $F_2$ to a $\mathit{TSS}$ state $s$ satisfy the following.
+The set $O$ consists of all the identifiers of pushes in $s$ which haven't added yet a node to {\tt pools} or for which the input is still present in {\tt pools} (i.e., the node created by the push has {\tt taken} set to {\tt false}).
+A push $k$ is labeled by $(d,{\tt PEND})$ where $d$ is the input value if it's pending and by $(d,{\tt COMP})$, otherwise.
+
+To describe the restrictions of $F_2$ on order relation $\mathsf{before}$ and the sets $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$, we consider the following notations: ${\tt ts}_s(k)$, resp., ${\tt TID}_s(k)$, denotes the timestamp of the node created by the push $k$ in state $s$ (the {\tt ts} field of this node), resp., the id of the thread executing $k$.
 By an abuse of terminology, we call ${\tt ts}_s(k)$ the timestamp of $k$ in state $s$.
@@
 More precisely, $k \leadsto_s k'$ when ${\tt TID}_s(k) < {\tt TID}_s(k')$, or ${\tt TID}_s(k) = {\tt TID}_s(k')$ and the node created by $k'$ is reachable from the one created by $k$ in the list pointed to by ${\tt pools}[{\tt TID}_s(k)]$.
-The order relation $<$ satisfies the following: (1) pending pushes are maximal, (2) $<$ is consistent with the order between node timestamps, i.e., ${\tt ts}_s(k) \leq {\tt ts}_s(k')$ implies $k'\not< k$, and (3) $<$ includes the order between pushes executed in the same thread, i.e., ${\tt TID}_s(k) = {\tt TID}_s(k')$ and ${\tt ts}_s(k) < {\tt ts}_s(k')$ implies $k < k'$.
 
-The components $be$ and $ov$ satisfy the following constraints (their domain is the set of identifiers of pending pops):
+The order relation $\mathsf{before}$ satisfies the following: (1) pending pushes are maximal as in $F_1$, (2) $\mathsf{before}$ is consistent with the order between node timestamps, i.e., ${\tt ts}_s(k) \leq {\tt ts}_s(k')$ implies $\neg \mathsf{before}(k',k)$, and (3) $\mathsf{before}$ includes the order between pushes executed in the same thread, i.e., ${\tt TID}_s(k) = {\tt TID}_s(k')$ and ${\tt ts}_s(k) < {\tt ts}_s(k')$ implies $\mathsf{before}(k < k')$.
+
+The components $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ satisfy the following constraints (their domain is the set of identifiers of pending pops):
 \vspace{-2mm}
 \begin{itemize}
-	\item a pop $k$ with ${\tt youngest}\neq {\tt null}$ that reached a node with timestamp $\tau$ (its variable {\tt n} points to this node) overlaps with every push that created a node with a timestamp bigger than $\tau$ and which occurs in {\tt pools} before the node reached by $k$, i.e., ${\tt youngest}_s(k)\neq{\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k_1)$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
-	\item a pop $k$ with ${\tt youngest}={\tt null}$ overlaps with every push that created a node which occurs in {\tt pools} before the node reached by $k$,
-i.e., ${\tt youngest}_s(k)={\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$ implies $k_2\in ov(k)$, for each $k, k_1, k_2$
-	\item if the variable {\tt youngest} of a pop $k$ points to a node which is not taken, then this node was created by a push in $be(k)\cup ov(k)$ or the node currently reached by $k$ is followed in {\tt pools} by another node which was created by a push in $be(k)\cup ov(k)$, i.e., ${\tt youngest}_s(k)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(k)={\tt n}_s(k_2)$ implies $k_1\in be(k)\cup ov(k)$ or that there exists $k_3\in O$ such that ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in be(k)\cup ov(k)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and $k$ is traversing the last list in the array {\tt pools}, for each $k, k_1,k_2$
+\item[\emph{Frontiers}] Every maximally-closed (maximal or all successors are pending) push $k$ can be removed by a pending pop $p$ in $t$.
+Formally, the push $k$ is maximally-closed if $\mathsf{present}(k) \land (\mathsf{pending}(k) \lor (\forall k'.
+\mathsf{present}(k') \land \mathsf{before}(k,k') \rightarrow \mathsf{pending}(k'))$.
+Then, for all pending pop operations $p$, $k \in \mathsf{overlap}(p) \cup \mathsf{maxAtInvoc}(p)$.
+	\item[\emph{TraverseBefore}] a pop $p$ with ${\tt youngest}(p)\neq {\tt null}$ that reached the node ${\tt n}$ overlaps with every present push that created a node with a timestamp greater than ${\tt youngest}(p){\tt ->ts}$ and which occurs in  {\tt pools} before the node $n$, i.e., ${\tt youngest}_s(p)= {\tt n}_s(k) \neq{\tt null}$, ${\tt n}_s(p)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, $\mathsf{present}(k_2)$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k)$ implies $k_2\in \mathsf{overlap}(p)$, for each $p, k_1, k_2$
+	%with timestamp $\tau$ (its variable {\tt n} points to this node) overlaps with every push that created a node with a timestamp bigger than $\tau$ and which occurs in {\tt pools} before the node reached by $k$, i.e., ${\tt youngest}_s(k)\neq{\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k_1)$ implies $k_2\in \mathsf{overlap}(k)$, for each $k, k_1, k_2$
+	\item[\emph{TraverseBeforeNull}] a pop $p$ with ${\tt youngest}(p)={\tt null}$ overlaps with every push that created a node which occurs in {\tt pools} before the node reached by $p$,
+i.e., ${\tt youngest}_s(p)={\tt null}$, ${\tt n}_s(p)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and $\mathsf{present}(k_2)$ implies $k_2\in \mathsf{overlap}(p)$, for each $p, k_1, k_2$
+	\item[\emph{TraverseAfter}] if the variable {\tt youngest} of a pop $p$ points to a node which is not taken, then this node was created by a push in $\mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$ or the node currently reached by $p$ is followed in {\tt pools} by another node which was created by a push in $\mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$, i.e., ${\tt youngest}_s(p)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(p)={\tt n}_s(k_2)$ implies $k_1\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$ or that there exists $k_3\in O$ such that $\mathsf{present}(k_3)$, ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and $p$ is on the control point that is before the assignment statement that changes the ${\tt youngest}$ field, for each $p, k_1,k_2$.
 %
+
 %	\item every pop overlaps with all pending pushes, i.e., $k_1\in O$ is pending implies $k_1\in ov(k)$ for each $k, k_1$

### 2017-05-08 draftCAV/stack.tex
started TSS simulation

@@
 \vspace{-1mm}
-We describe a forward simulation $\mathit{fs}_2$ from $\mathit{TSS}$ to $AbsS$.
-Except for the constraints on the components $be$ and $ov$ of a $AbsS$ state, it is similar to the simulation $\mathit{fs}_1$ from $\mathit{HWQ}$ to $AbsQ$.
+We describe a forward simulation $F_2$ from $\mathit{TSS}$ to $AbsS$, which is similar to the one from $\mathit{HWQ}$ to $AbsQ$ for the components of an $AbsS$ state which exist also in $AbsQ$ (i.e., different from $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$).
+% Except for the constraints on the components $be$ and $ov$ of a $AbsS$ state, it is similar to the simulation $\mathit{fs}_1$ from $\mathit{HWQ}$ to $AbsQ$.
 Thus, the $AbsS$ states $t=\tup{O,<,\ell,rv,cp,be,ov}$ associated by $\mathit{fs}_2$ to a $\mathit{TSS}$ state $s$ satisfy the following.
@@
 
-Similar to enqueue operations in a $\mathit{HWQ}$ state, if a $\mathit{TSS}$ state $s$ is related to $AbsS$ state $t$ by $F_2$, then for every push operation $k$ in $s$ such that $k$ has not yet added a node to {\tt pools} or its node is still present in {\tt pools} (i.e., the node created by the push has {\tt taken} set to {\tt false}), $\mathsf{present}(k)$ must be true in $t$.
-Similarly, $\mathsf{pending}(k)$ must be true in $t$ if $k$ is invoked but has not yet executed its last statement in $s$.
+Thus, a $\mathit{TSS}$ state $s$ is related by $F_2$ to $AbsS$ states $t$ where $\mathsf{present}(k)$ is true for every push operation $k$ in $s$ such that $k$ has not yet added a node to {\tt pools} or its node is still present in {\tt pools} (i.e., the node created by the push has {\tt taken} set to {\tt false}).
+Also, $\mathsf{pending}(k)$ is true in $t$ iff $k$ is pending in $s$.
 %Except for the constraints on the components $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ of a $AbsS$ state, it is similar to the simulation $F_1$ from $\mathit{HWQ}$ to $AbsQ$.
@@
 
-To describe the restrictions of $F_2$ on order relation $\mathsf{before}$ and the sets $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$, we consider the following notations: ${\tt ts}_s(k)$, resp., ${\tt TID}_s(k)$, denotes the timestamp of the node created by the push $k$ in state $s$ (the {\tt ts} field of this node), resp., the id of the thread executing $k$.
+To describe the constraints on the order relation $\mathsf{before}$ and the sets $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ in $t$, we consider the following notations: ${\tt ts}_s(k)$, resp., ${\tt TID}_s(k)$, denotes the timestamp of the node created by the push $k$ in state $s$ (the {\tt ts} field of this node), resp., the id of the thread executing $k$.
 By an abuse of terminology, we call ${\tt ts}_s(k)$ the timestamp of $k$ in state $s$.
@@
 
-The order relation $\mathsf{before}$ satisfies the following: (1) pending pushes are maximal as in $F_1$, (2) $\mathsf{before}$ is consistent with the order between node timestamps, i.e., ${\tt ts}_s(k) \leq {\tt ts}_s(k')$ implies $\neg \mathsf{before}(k',k)$, and (3) $\mathsf{before}$ includes the order between pushes executed in the same thread, i.e., ${\tt TID}_s(k) = {\tt TID}_s(k')$ and ${\tt ts}_s(k) < {\tt ts}_s(k')$ implies $\mathsf{before}(k < k')$.
+The order relation $\mathsf{before}$ satisfies the following: (1) pending pushes are maximal, (2) $\mathsf{before}$ is consistent with the order between node timestamps, i.e., ${\tt ts}_s(k) \leq {\tt ts}_s(k')$ implies $\neg \mathsf{before}(k',k)$, and (3) $\mathsf{before}$ includes the order between pushes executed in the same thread, i.e., ${\tt TID}_s(k) = {\tt TID}_s(k')$ and ${\tt ts}_s(k) < {\tt ts}_s(k')$ implies $\mathsf{before}(k, k')$.
 
 The components $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ satisfy the following constraints (their domain is the set of identifiers of pending pops):
-\vspace{-2mm}
-\begin{itemize}
-\item[\emph{Frontiers}] Every maximally-closed (maximal or all successors are pending) push $k$ can be removed by a pending pop $p$ in $t$.
-Formally, the push $k$ is maximally-closed if $\mathsf{present}(k) \land (\mathsf{pending}(k) \lor (\forall k'.
+\vspace{-1mm}
+\begin{description}
+\item[\emph{Frontiers}:] By the definition of $TSS$, a pending pop $p$ in $s$ could, in the future, remove the value added by a push $k$ which is maximal (w.r.t.
+$\mathsf{before}$) or a push $k$ which is completed but followed only by pending pushes (in the order relation $\mathsf{before}$).
+Therefore, for all pop operations $p$ which are pending in $s$, we have that $k \in \mathsf{overlap}(p) \cup \mathsf{maxAtInvoc}(p)$, for every push $k$ such that $\mathsf{present}(k) \land (\mathsf{pending}(k) \lor (\forall k'.
+\mathsf{present}(k') \land \mathsf{before}(k,k') \rightarrow \mathsf{pending}(k'))$.
+%Every maximal push $k$ or completed push  (maximal or all successors are pending) push $k$ can be removed by . Formally, the 
+%push $k$ such that $\mathsf{present}(k) \land (\mathsf{pending}(k) \lor (\forall k'.
 \mathsf{present}(k') \land \mathsf{before}(k,k') \rightarrow \mathsf{pending}(k'))$.
 Then, for all pending pop operations $p$, $k \in \mathsf{overlap}(p) \cup \mathsf{maxAtInvoc}(p)$.
-	\item[\emph{TraverseBefore}] a pop $p$ with ${\tt youngest}(p)\neq {\tt null}$ that reached the node ${\tt n}$ overlaps with every present push that created a node with a timestamp greater than ${\tt youngest}(p){\tt ->ts}$ and which occurs in  {\tt pools} before the node $n$, i.e., ${\tt youngest}_s(p)= {\tt n}_s(k) \neq{\tt null}$, ${\tt n}_s(p)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, $\mathsf{present}(k_2)$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k)$ implies $k_2\in \mathsf{overlap}(p)$, for each $p, k_1, k_2$
+	\item[\emph{TraverseBefore}:] a pop $p$ with ${\tt youngest}(p)\neq {\tt null}$ that reached the node ${\tt n}$ overlaps with every present push that created a node with a timestamp greater than ${\tt youngest}(p){\tt ->ts}$ and which occurs in  {\tt pools} before the node $n$.
+Formally, if ${\tt youngest}_s(p)= {\tt n}_s(k) \neq{\tt null}$, ${\tt n}_s(p)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, $\mathsf{present}(k_2)$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k)$, then $k_2\in \mathsf{overlap}(p)$, for each $p, k_1, k_2$.
 	%with timestamp $\tau$ (its variable {\tt n} points to this node) overlaps with every push that created a node with a timestamp bigger than $\tau$ and which occurs in {\tt pools} before the node reached by $k$, i.e., ${\tt youngest}_s(k)\neq{\tt null}$, ${\tt n}_s(k)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, ${\tt n}_s(k_2)\text{\tt ->taken}={\tt false}$, and ${\tt ts}_s(k_2) \geq {\tt ts}_s(k_1)$ implies $k_2\in \mathsf{overlap}(k)$, for each $k, k_1, k_2$
-	\item[\emph{TraverseBeforeNull}] a pop $p$ with ${\tt youngest}(p)={\tt null}$ overlaps with every push that created a node which occurs in {\tt pools} before the node reached by $p$,
-i.e., ${\tt youngest}_s(p)={\tt null}$, ${\tt n}_s(p)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and $\mathsf{present}(k_2)$ implies $k_2\in \mathsf{overlap}(p)$, for each $p, k_1, k_2$
-	\item[\emph{TraverseAfter}] if the variable {\tt youngest} of a pop $p$ points to a node which is not taken, then this node was created by a push in $\mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$ or the node currently reached by $p$ is followed in {\tt pools} by another node which was created by a push in $\mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$, i.e., ${\tt youngest}_s(p)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(p)={\tt n}_s(k_2)$ implies $k_1\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$ or that there exists $k_3\in O$ such that $\mathsf{present}(k_3)$, ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, $k_3\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and $p$ is on the control point that is before the assignment statement that changes the ${\tt youngest}$ field, for each $p, k_1,k_2$.
+	\item[\emph{TraverseBeforeNull}:] a pop $p$ with ${\tt youngest}(p)={\tt null}$ overlaps with every push that created a node which occurs in {\tt pools} before the node reached by $p$,
+i.e., ${\tt youngest}_s(p)={\tt null}$, ${\tt n}_s(p)={\tt n}_s(k_1)$, $k_2\leadsto_s k_1$, and $\mathsf{present}(k_2)$ implies $k_2\in \mathsf{overlap}(p)$, for each $p, k_1, k_2$.
+	\item[\emph{TraverseAfter}:] if the variable {\tt youngest} of a pop $p$ points to a node which is not taken, then this node was created by a push in $\mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$ or the node currently reached by $p$ is followed in {\tt pools} by another node which was created by a push in $\mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$.
+Formally,  for each $p, k_1,k_2$, if ${\tt youngest}_s(p)={\tt n}_s(k_1)$, ${\tt n}_s(k_1)\text{{\tt ->taken}}={\tt false}$, and ${\tt n}_s(p)={\tt n}_s(k_2)$, then one of the following holds:
+	\begin{itemize}
+		\item $k_1\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$, or 
+		\item there exists a push $k_3$ in $s$ such that $\mathsf{present}(k_3)$, $k_3\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$, ${\tt ts}_s(k_3) > {\tt ts}_s(k_1)$, and either $k_2\leadsto_s k_3$ or ${\tt n}_s(k_2)={\tt n}_s(k_3)$ and $p$ is at a control point before the assignment statement that changes the variable ${\tt youngest}$.
+	\end{itemize}
 %
@@
 \vspace{-2mm}
-\end{itemize}
-There are some more constraints on $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ that can be seen as invariants of $AbsS$, i.e., $\mathsf{maxAtInvoc}(p)$ and $\mathsf{overlap}(p)$ do not contain predecessors of pushes from $\mathsf{maxAtInvoc}(p)$ (for each $p, k_1, k_2$, $\mathsf{before}(k_1, k_2)$ and $k_2 \in \mathsf{maxAtInvoc}(p)$ implies $k_1\not\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$).
+\end{description}
+There are some more constraints on $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ that can be seen as invariants of $AbsS$, e.g., $\mathsf{maxAtInvoc}(p)$ and $\mathsf{overlap}(p)$ do not contain predecessors of pushes from $\mathsf{maxAtInvoc}(p)$ (for each $p, k_1, k_2$, $\mathsf{before}(k_1, k_2)$ and $k_2 \in \mathsf{maxAtInvoc}(p)$ implies $k_1\not\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$).
 They can be found in the extended version of the paper (\cite{extended}).
 
-The rationale behind the construction of $F_2$ has some similarities with $F_1$ that relates $\mathit{HWQ}$ and $AbsQ$.
+%The rationale behind the construction of $F_2$ has some similarities with $F_1$ that relates $\mathit{HWQ}$ and $AbsQ$.
 As in $F_1$, operation identifiers of pending push operations should be maximal in the related $\mathit{TSS}$ states.
@@
 
-The dual condition of ensuring maximal-closedness of remove candidate push operations is not valid for $F_2$, since linearization points of pop operations are not fixed and they can linearize in any point between their invocation and commit points.
-Hence, any push operation that is maximally-closed between these points are removable by the pop and they might not be maximally-closed in the current state.
+Note that $F_2$ cannot satisfy the reverse of {\bf \emph{Frontiers}}, i.e., every push in $\mathsf{overlap}(p) \cup \mathsf{maxAtInvoc}(p)$, for some $p$, is maximal or followed only by pending pushes (w.r.t., $\mathsf{before}$).
+This is because
+%The dual condition of ensuring maximal-closedness of remove candidate push operations is not valid for $F_2$, since 
+the linearization points of pop operations are not fixed and they can occur anywhere in between their invocation and commit points.
+Hence, any push operation which was maximal or followed only by pending pushes in the happens-before
+%that is maximally-closed 
+in between the invocation and the commit 
+can be removed by a pop. And such a push may not satisfy the same properties in the state $s$. 
+% and they might not be maximally-closed in the current state.
 Unfortunately, $\mathit{TSS}$ does not keep history of pop operations in the state and we cannot infer all the push operations that are concurrent or maximally closed with respect to a given pop operation from the state information.
-
-Nevertheless, we can obtain partial information from the state that is enough for showing once-maximality of a push that is removed by commit action of a pop.
-First, the push operations that are maximally-closed in the current state are remove candidates by all pending pops since a pending pop operation may linearize at this point.
+%Nevertheless, we can obtain partial information from the state that is enough for showing once-maximality of a push that is removed by commit action of a pop.
+First, the push operations tht are maximally-closed in the current state are remove candidates by all pending pops since a pending pop operation may linearize at this point.
 This restriction is described by the \emph{Frontiers} condition.

### 2017-05-08 draftCAV/stack.tex
TSS simulation

@@
 in between the invocation and the commit 
-can be removed by a pop. And such a push may not satisfy the same properties in the state $s$. 
+can be removed by a pop. And such a push may no longer satisfy the same properties in the state $s$.
+
+Based on the values stored in $youngest_s(p)$ and $n_s(p)$, for some pop $p$, %which give the current youngest element $p$ has seen and the current element $p$ is visiting, 
+the other three constraints identify other push operations that overlap with $p$, or they were followed only by pending pushes when $p$ was invoked. 
 % and they might not be maximally-closed in the current state.
@@
 This restriction is described by the \emph{Frontiers} condition.
+%To obtain further information on $\mathsf{overlap}$ and $\mathsf{maxAtInvoc}$ sets of a pending pop $p$ in state $s$, we look at  which show the current youngest element $p$ has seen and the current element $p$ is visiting, respectively.
+{\bf \emph{TraverseBefore}} and {\bf \emph{TraverseBeforeNull}} state that pushes which add new nodes to the pools seen by $p$ in the past,  are overlapping with $p$.
+{\bf \emph{TraverseAfter}} states that either the push adding the current youngest node $youngest_s(p)$ is in $\mathsf{overlap}_s(p) \cup \mathsf{maxAtInvoc}_s(p)$, or there is a node that $p$ will visit in the future which is in $\mathsf{overlap}_s(p) \cup \mathsf{maxAtInvoc}_s(p)$.
 
- To obtain further information on $\mathsf{overlap}$ and $\mathsf{maxAtInvoc}$ sets of a pending pop $p$ in state $s$, we look at $youngest_s(p)$ and $n_s(p)$ which show the current youngest element $p$ has seen and the current element $p$ is visiting, respectively.
-\emph{TraverseBefore} and \emph{TraverseBeforeNull} conditions make sure that if threads add new nodes to their pools after $p$ passes them during its travels, these pushes must be concurrent with $p$.
-The \emph{TraverseAfter} condition makes sure that either the current youngest node $youngest_s(p)$ is in $\mathsf{overlap}_s(p) \cup \mathsf{maxAtInvoc}_s(p)$ or there is a future node that $p$ will visit and this future node is in $\mathsf{overlap}_s(p) \cup \mathsf{maxAtInvoc}_s(p)$.
-
-Combining \emph{TraverseBefore} and \emph{TraverseAfter} restrictions we make sure that when a new node $n_s(p)$ is visited, then either $youngest_s(p)$ was maximally-closed before new nodes that precede $n_s(p)$ in the traverse order were added, or there is a successor node of $n_s(p)$ in the traverse order that is maximally-closed.
-Hence, when $p$ finishes its traversal of pools with a non-null and present $youngest$ node, this node is guaranteed to be once-maximal between invocation and commit points of $p$.
+%Combining \emph{TraverseBefore} and \emph{TraverseAfter} restrictions we make sure that when a new node $n_s(p)$ is visited, then either $youngest_s(p)$ was maximal or followed only by pending pushes before new nodes that precede $n_s(p)$ in the traverse order were added, or there is a successor node of $n_s(p)$ in the traverse order that is maximal or followed only by pending pushes.
+Hence, when a pop $p$ finishes its traversal of the pools with a non-null and present $youngest$ node, this node is guaranteed to be once-maximal between invocation and commit points of $p$.
 

### 2017-05-08 draftCAV/queue.tex
Updated AbsQ text.

@@
 
-For readability, we define $AbsQ$ as an abstract state machine, which is given in Fig.~\ref{fig:transitions:AbsQ}.
-The shared state of $AbsQ$ consists of several boolean predicates indicating whether the value added by an enqueue has not been removed yet (the predicate $\mathsf{present})$, whether an enqueue is pending (the predicate $\mathsf{pending}$), the happens-before order (the predicate $\mathsf{before}$), and a function giving the value added by an enqueue (the function $\mathsf{arg}$).
-In the initial state, the domain of every function (predicate) is empty.
-%the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).
-%All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}.
-The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points, denoted by $lin(deq,d,k)$.
-$Lin(deq)$ is the set of all actions $lin(deq,d,k)$.
-%
-The enqueue operations consist of two macro rules: the rule {\tt inv(enq,x,k)} orders the invoked operation after all the completed enqueues present in the current state ({\tt k} is a fresh operation identifier associated to the current operation), and the rule {\tt ret(enq,k)} sets $\mathsf{pending}${\tt (k)} to {\tt false}.
-% provided that the operation is still present in the current state.
-The dequeue operations consist of two rules {\tt inv(deq,k)} and {\tt ret(deq,y,k)} associated to the invocation and respectively, the return action, whose definition is obvious, and the rule {\tt lin(deq,y,k)} corresponding to its linearization point.
-When a non-empty set of pending enqueues are present, the rule {\tt lin(deq,y,k)} is non-deterministic.
-It can set the return value {\tt y} either to {\tt EMPTY}, or to a value {\tt d} $\neq$ {\tt EMPTY}, provided that {\tt d} has been added by an enqueue {\tt k1} which is minimal in the current happens-before.
-In the latter case, $\mathsf{present}${\tt (k1)} is set to {\tt false} to indicate that {\tt k1}'s argument has been removed.
-When there exists at least one completed enqueue whose value has not been removed, the rule {\tt lin(deq,y,k)} sets {\tt y} to the value added by a minimal enqueue (w.r.t.
-the current happens-before).
+We define $AbsQ$ with the abstract state machine given in
+Figure~\ref{fig:transitions:AbsQ} which defines an LTS over the alphabet $C\cup
+R\cup Lin(deq)$. The state of $AbsQ$ consists of several updatable functions:
+{\tt loc} indicates the abstract control point of a given operation; {\tt arg}
+and {\tt ret} indicate the argument or return value of a given operation,
+respectively; {\tt present} indicates whether a given enqueue operation has yet
+to be removed, and {\tt pending} indicates whether it has yet to complete; {\tt
+before} indicates the happens-before order between operations. Initially, {\tt
+loc(k) = inv} for all {\tt k}, and {\tt present(k1) = pending(k1) =
+before(k1,k2) = false} for all {\tt k1, k2}. Each rule determines the next state
+of $AbsQ$ for the corresponding action. For instance, the {\tt lin(deq,v,k)}
+rule updates the state of $AbsQ$ for the linearization action of a dequeue
+operation with identifier {\tt k} returning value {\tt v}: when {\tt v = EMPTY}
+then $AbsQ$ insists via an assertion that any still-present enqueue must still
+be pending; otherwise, when {\tt v $\neq$ k} then $AbsQ$ insists that a
+corresponding enqueue is present, and that it is minimal in the happens-before
+order, before marking that enqueue as not present. Updates to the {\tt loc}
+function, implicit in Figure~\ref{fig:transitions:AbsQ}, ensure that the
+invocation, linearization-point, and return actions of each operation occur in
+the correct order.
 
-%{\sc call-enq} orders the invoked operation after all the completed enqueues in the current state, and the rules {\sc ret-enq1}/{\sc ret-enq2} flip the corresponding flag from {\tt PEND} to {\tt COMP} provided that the operation is still present in the current state.
-For dequeue operations, {\sc call-deq} only increments the control point and {\sc ret-deq} checks whether the return value is the same as the one fixed at the linearization point.
-The linearization point rule {\sc lin-deq1} corresponds to the case of a non-empty queue, showing that $lin(deq,d,k)$ is enabled only if $d$ has been added by an enqueue which is minimal in the current happens-before.
-When enabled, it removes the enqueue adding $d$ from the state.
-The linearization point rule {\sc lin-deq2} corresponds to the case of dequeue operations linearized with an {\tt EMPTY} return value.
+The following result states that the library $AbsQ$ has exactly the same set of
+histories as the standard abstract library $AbsQ_0$.
 
-% Let $AbsQ_0$ denote this implementation (formally defined in Appendix~\ref{app:absImplQueue}).
-The following result states that the library $AbsQ$ has exactly the same set of histories as the standard abstract library $AbsQ_0$. % (see Appendix~\ref{app:absImplQueue} for a proof).
+\begin{theorem}
+  \label{th:absImplQueue}
 
-\vspace{-1.5mm}
-\begin{theorem}\label{th:absImplQueue}
   $AbsQ$ is a refinement of $AbsQ_0$ and vice-versa.
-\vspace{-2mm}
+
 \end{theorem}
-
-The abstract state machine in Fig.~\ref{fig:transitions:AbsQ} defines an LTS over the alphabet $C\cup R\cup Lin(deq)$.
-The transitions corresponding to the rule {\tt lin(deq,y,k)} are labeled by  actions $lin(deq,y,k)$.
-Also, we assume that the transition corresponding to the invocation of a method is done atomically with the first step of the same invocation (the macro rule {\tt inv(...)}), and similarly for transitions corresponding to returning from a method, they are done atomically with the last step (the macro rule {\tt ret(...)}).
-They are labeled as expected with call/return actions (borrowing the operation identifier which occurs as argument of the macro rules).
-It can be easily proved that this assumption doesn't modify the set of traces projected on $C\cup R\cup Lin(deq)$ (compared to an LTS where these steps are not atomic).
-
-%~\footnote{Without this assumption, while proving forward simulations, a call/return action in the concrete implementation is simulated by two steps of the abstract implementation, the call/return action together with the first/last macro rule.}.
 

### 2017-05-09 draftCAV/queue.tex
final version

@@
 %!TEX root = draft.tex
-\vspace{-3.5mm}
+\vspace{-1mm}
 \section{Queues With Fixed Dequeue Linearization Points}\label{sec:queues}
-\vspace{-1.5mm}
+%\vspace{-1.5mm}
 
@@
 %TODO HOW CAN WE CONVINCE THAT OTHER IMPLEMENTATIONS HAVE THE SAME PROPERTY ?
-\vspace{-3.5mm}
+\vspace{-2.5mm}
 \subsection{Enqueue Methods With Non-Fixed Linearization Points}
-\vspace{-1mm}
+%\vspace{-1mm}
 % describe a queue implementation listed in Figure~\ref{fig:HerlihyWing} and known as the Herlihy \& Wing Queue~\cite{journals/toplas/HerlihyW90} ($\mathit{HWQ}$ for short), where only the dequeue methods have fixed linearization points.
@@
 Initially, all the positions in the array are {\tt null} and {\tt back} is 0.
-An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position can't be used by other enqueues) and then, stores the argument {\tt x} at this position.
+An enqueue method starts by reserving a position in {\tt items} ({\tt i} stores the index of this position and {\tt back} is incremented so the same position cannot be used by other enqueues) and then, stores the argument {\tt x} at this position.
 The dequeue method traverses the array {\tt items} starting from the beginning and atomically swaps {\tt null} with the encountered value.
@@
 
-\vspace{-3.5mm}
+\vspace{-2.5mm}
 \subsection{Abstract Queue Implementation}
-\vspace{-1.5mm}
+%\vspace{-1.5mm}
 
@@
 
-\vspace{-4.5mm}
+\vspace{-2.5mm}
 \subsection{A Correctness Proof For Herlihy\&Wing Queue}\label{ssec:HerlihyWing}
-\vspace{-1mm}
+%\vspace{-1mm}
 We describe a forward simulation $F_1$ from $\mathit{HWQ}$ to $AbsQ$.

### 2017-07-04 draftCAV/prelim.tex

@@
 %!TEX root = draft.tex
+\vspace{-1mm}
 \section{Preliminaries}
-\vspace{-2mm}
+
 We formalize several abstraction relations between libraries using a simple
@@
 
-\vspace{-3.5mm}
+\vspace{-2.5mm}
 \subsection{Libraries}
-\vspace{-1.5mm}
+%\vspace{-1.5mm}
 Programs interact with libraries by calling named library \emph{methods}, which
@@
 
-\vspace{-3.5mm}
+\vspace{-2.5mm}
 \subsection{Refinement and Linearizability}
-\vspace{-1.5mm}
+%\vspace{-1.5mm}
 %\textcolor{red}{A general organizational idea for  this subsection: Linearizability is defined in terms of weakening in which observational actions are just call and return events.

### 2017-07-04 draftCAV/refProofs.tex

@@
 %!TEX root = draft.tex
-\vspace{-3.5mm}
+\vspace{-1mm}
 \section{Refinement Proofs}
-\vspace{-1.5mm}
+%\vspace{-1.5mm}
 Library refinement is an instance of a more general notion of refinement between LTSs,
@@
 \end{itemize}
-\vspace{-3.5mm}
+\vspace{-2.5mm}
 \end{definition}
 %If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
-A $\Gamma$-forward simulation requires that every step of $A_1$ corresponds to a sequence of steps of $A_2$.
+A $\Gamma$-forward simulation states that every step of $A_1$ is simulated by a sequence of steps of $A_2$.
 To imply $\Gamma$-refinement, every step of $A_1$ labeled by an observable action $\gamma\in \Gamma$ should be simulated by a sequence of steps of $A_2$ where exactly one transition is labeled by $\gamma$ and all the other transitions are labeled by non-observable actions.
-The dual notion of \emph{backward} simulation where steps are simulated backwards can be defined in a similar way.
+The dual notion of \emph{backward} simulation where steps are simulated backwards can be defined similarly.
 
@@
 
-\vspace{-1.5mm}
+\vspace{-1mm}
 \begin{theorem}\label{th:forSim}
@@
 Also, let $\Gamma$ be a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$.
-$A_1$ $\Gamma$-refines $A_2$ when there is a $\Gamma$-forward simulation from $A_1$ to $A_2$.
-Moreover, if $A_1$ $\Gamma$-refines $A_2$ and $A_2$ is $\Gamma$-deterministic, then there is a $\Gamma$-forward simulation from $A_1$ to $A_2$.
-\vspace{-1.5mm}
+If there is a $\Gamma$-forward simulation from $A_1$ to $A_2$, then $A_1$ $\Gamma$-refines $A_2$.
+Also, if $A_1$ $\Gamma$-refines $A_2$ and $A_2$ is $\Gamma$-deterministic, then there is a $\Gamma$-forward simulation from $A_1$ to $A_2$.
+\vspace{-1mm}
 \end{theorem}
@@
 each operation, the linearization of a concurrent history is the sequential history which takes operations in the order of their linearization points.
-For some libraries, the linearization points of all the invocations of a method $m$ correspond to the execution of a fixed statement in $m$'s body. For instance, in the case of atomic libraries  
-where method bodies are guarded with a global-lock acquisition, the linearization point of every method invocation corresponds to the execution 
+For some libraries, the linearization points of all the invocations of a method $m$ correspond to the execution of a fixed statement in $m$'s body.
+For instance, when method bodies are guarded with a global-lock acquisition, the linearization point of every method invocation corresponds to the execution
 of the body. When the linearization points are fixed, we assume that the library is an LTS over an alphabet that includes actions 
-$lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$. The action $lin(m,d,k)$ represents the linearization point of the operation $k$ 
+$lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$, representing the linearization point of the operation $k$ 
 %which is an invocation of method $m$ 

### 2017-07-04 draftCAV/related.tex
small things

@@
 %!TEX root = draft.tex
-\vspace{-3.5mm}
+\vspace{-1mm}
 \section{Related Work}
-\vspace{-1.5mm}
+%\vspace{-1.5mm}
 Many techniques for linearizability verification, e.g.,~\cite{conf/ppopp/VafeiadisHHS06,conf/cav/AmitRRSY07,conf/vmcai/Vafeiadis09,conf/tacas/AbdullaHHJR13}, are based on forward simulation arguments, and typically only work for libraries where the linearization point of every invocation of a method $m$ is fixed to a particular statement in the code of $m$.
@@
 For manual proofs, such implementations with external linearization points can still be defined as LTSs that produce $Lin$-complete traces and thus still fall in the class of implementations for which forward simulations are enough for proving refinement.
-All these works are not able to deal with implementations such as $\mathit{HWQ}$ or $\mathit{TSS}$.
+These proof methods are not complete and they are not able to deal with implementations like $\mathit{HWQ}$ or $\mathit{TSS}$.
 

### 2017-07-04 draftCAV/stack.tex

@@
 %!TEX root = draft.tex
-\vspace{-3.5mm}
+\vspace{-1mm}
 \section{Stacks With Fixed Pop Commit Points}\label{sec:stacks}
-\vspace{-1.5mm}
+%\vspace{-1.5mm}
 The abstract implementation in Section~\ref{sec:queues} can be adapted to stacks, the main modification being that the linearization point $lin(pop,d,k)$ with $d\neq{\tt EMPTY}$ is enabled when $k$ is added by a push which is maximal in the happens-before order stored in the state.
@@
 
-\vspace{-3.5mm}
+\vspace{-2.5mm}
 \subsection{Pop Methods With Fixed Commit Points}
-\vspace{-1mm}
+%\vspace{-1mm}
 
@@
 \includegraphics[width=11.5cm]{fig-tss.pdf}
-\vspace{-4mm}
+\vspace{-3mm}
 \caption{An execution of $\mathit{TSS}$.
@@
 \label{fig:commit}
-\vspace{-6mm}
+\vspace{-4mm}
 \end{figure}
@@
 
-\vspace{-3mm}
+\vspace{-2.5mm}
 \subsection{Abstract stack implementation}
-\vspace{-1mm}
+%\vspace{-1mm}
 We define an abstract stack $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the pushes whose value has not yet been  removed by a matching pop.
@@
 \label{fig:stackSim}
-\vspace{-6.5mm}
+\vspace{-4mm}
 \end{figure}
@@
       before(k1,k) := true
-  forall k1 with pending(k1):
+  forall k1:
     overlap(k1) := overlap(k1)$\cup$$\{$k$\}$
@@
 The description of $AbsS$ as an abstract state machine is given in Fig.~\ref{fig:transitions:AbsS}.
-Compared to the queue implementation $AbsQ$, the shared state contains two more functions $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$.
+Compared to $AbsQ$, the state contains two more updatable functions $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$.
 For each pop operation $k$, $\mathsf{maxAtInvoc}(k)$ records the set of completed push operations which were maximal in the happens-before (defined by $\mathsf{before}$) when pop $k$ was invoked, or happening earlier provided that the values of all the pushes happening later than one of these maximal ones and before pop $k$ have been removed.
-Also, for each pop operation $k$, $\mathsf{overlap}(k)$ contains the push operations overlapping with pop $k$.
+Also, $\mathsf{overlap}(k)$ contains the push operations overlapping with a pop $k$.
 %Formally, the states of $AbsS$ are tuples $\tup{O,<,\ell,rv,cp,be,ov}$ where $<$ is a strict partial order over the set $O$ of operation identifiers, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a pending/completed flag, $rv:\<Ops> ~> \<Vals>$ records the return value of a pending pop fixed at its commit point, $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every push ($A_1, A_2$) or pop operation ($R_1,R_2,R_3$), $be:\<Ops> ~> 2^O$ records the greatest completed push operations before a pop started or happening earlier provided that the values of all the push happening later have been removed, and $ov: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
@@
 %All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsS}.
-In the initial state, the domain of every function (predicate) is empty.
-The push methods are similar to the enqueue methods in $AbsQ$.
-The only difference is that every newly invoked push operation {\tt k} is added to the set $\mathsf{overlap}(${\tt k1}$)$ for all pop operations {\tt k1} (since {\tt k} overlaps with all the currently pending pops).
-The rule {\tt inv(pop,k)}, marking the invocation of a pop, sets $\mathsf{maxAtInvoc}(${\tt k}$)$ and $\mathsf{overlap}(${\tt k}$)$ as explained above.
-The rule {\tt com(pop,y,k)} results in setting the return value {\tt y} to {\tt EMPTY} if the set $\mathsf{maxAtInvoc}(${\tt k}$)$ is empty (otherwise, there would be push operations happening before pop {\tt k} which makes the answer {\tt EMPTY} incorrect).
-It results in setting the return value to a value $d\neq\,${\tt EMPTY} when $d$ was added by a push {\tt k1} which belongs to $\mathsf{maxAtInvoc}(${\tt k}$)$ or $\mathsf{overlap}(${\tt k}$)$.
-In this case, the rule {\tt com(pop,y,k)} may also update $\mathsf{maxAtInvoc}(${\tt k2}$)$ for other pending pops {\tt k2}.
+Initially, {\tt loc(k) = inv} for all {\tt k}, {\tt present(k1) = pending(k1) = before(k1,k2) = false}, and {\tt maxAtInvoc(k1) = overlap(k1) = }$\emptyset$, for all {\tt k1, k2}.
+%In the initial state, the domain of every function (predicate) is empty. 
+The rules for actions of push methods are similar to those for enqueues in $AbsQ$, except that every newly invoked push operation {\tt k} is added to the set $\mathsf{overlap}(${\tt k1}$)$ for all pop operations {\tt k1} (since {\tt k} overlaps with all the currently pending pops).
+The rule {\tt inv(pop,k)}, marking the invocation of a pop, sets $\mathsf{maxAtInvoc}(${\tt k}$)$ and $\mathsf{overlap}(${\tt k}$)$ as explained above. The rule {\tt com(pop,EMPTY,k)} is enabled when 
+%results in setting the return value {\tt y} to {\tt EMPTY} if 
+the set $\mathsf{maxAtInvoc}(${\tt k}$)$ is empty (otherwise, there would be push operations happening before pop {\tt k} which makes the return value {\tt EMPTY} incorrect).
+Also, {\tt com(pop,y,k)} with
+{\tt y}$\,\neq\,${\tt EMPTY} is enabled when {\tt y} was added by a push {\tt k1} which belongs to $\mathsf{maxAtInvoc}({\tt k})\cup\mathsf{overlap}({\tt k})$.
+This rule may also update $\mathsf{maxAtInvoc}(${\tt k2}$)$ for other pending pops {\tt k2}.
 More precisely, whenever $\mathsf{maxAtInvoc}(${\tt k2}$)$ contains the push {\tt k1}, the latter is replaced by the immediate predecessors of {\tt k1} (according to $\mathsf{before}$) that are followed exclusively by pushes overlapping with {\tt k2}.
@@
 The abstract state machine in Fig.~\ref{fig:transitions:AbsS} defines an LTS over the alphabet $C\cup R\cup Com(pop)$. 
-As in the case of $AbsQ$, we assume that the transition corresponding to the invocation, resp., the return, of a method, is done atomically with the fist, resp., the last, macro rule in its body.
+%As in the case of $AbsQ$, we assume that the transition corresponding to the invocation, resp., the return, of a method, is done atomically with the fist, resp., the last, macro rule in its body.
 These transitions are labeled as expected with call and return actions (borrowing the operation identifier which occurs as argument of the macro rule), and the transitions corresponding to the rule {\tt com(pop,y,k)} are labeled by $com(pop,d,k)$.
-
 Let $AbsS_0$ be the standard abstract implementation of a stack where elements are stored in a sequence, push and pop operations adding and removing an element from the beginning of the sequence in one atomic step, respectively.
@@
 
-\vspace{-2mm}
+\vspace{-1mm}
 \begin{theorem}\label{th:absImplStack}
 $AbsS$ is a refinement of $AbsS_0$ and vice-versa.
-\vspace{-2mm}
+\vspace{-1mm}
 \end{theorem}
@@
 
-\vspace{-1.5mm}
+\vspace{-1mm}
 \begin{corollary}
 A stack $L$ with fixed pop commit points is a $C\cup R\cup Com(pop)$-refinement of $AbsS$ if{f} there is a $C\cup R\cup Com(pop)$-forward simulation from $L$ to $AbsS$.
-\vspace{-1.5mm}
+\vspace{-1mm}
 \end{corollary}
@@
 
-\vspace{-1.5mm}
+\vspace{-1mm}
 \begin{corollary}
 A stack implementation $L$ with fixed pop linearization points where transition labels $lin(pop,d,k)$ are substituted with $com(pop,d,k)$ is a $C\cup R\cup Com(pop)$-refinement of $AbsS_0$ if{f} there is a $C\cup R\cup Com(pop)$-forward simulation from $L$ to $AbsS$.
-\vspace{-1.5mm}
+\vspace{-1mm}
 \end{corollary}
 
-\vspace{-6mm}
+\vspace{-2.5mm}
 \subsection{A Correctness Proof For Time-Stamped Stack}\label{sec:corr_tss}
-\vspace{-1mm}
+%\vspace{-1mm}
 We describe a forward simulation $F_2$ from $\mathit{TSS}$ to $AbsS$, which is similar to the one from $\mathit{HWQ}$ to $AbsQ$ for the components of an $AbsS$ state which exist also in $AbsQ$ (i.e., different from $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$).
@@
 \end{description}
-There are some more constraints on $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ that can be seen as invariants of $AbsS$, e.g., $\mathsf{maxAtInvoc}(p)$ and $\mathsf{overlap}(p)$ do not contain predecessors of pushes from $\mathsf{maxAtInvoc}(p)$ (for each $p, k_1, k_2$, $\mathsf{before}(k_1, k_2)$ and $k_2 \in \mathsf{maxAtInvoc}(p)$ implies $k_1\not\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$).
-They can be found in the extended version of the paper (\cite{extended}).
+%There are some more constraints on 
+The functions $\mathsf{maxAtInvoc}$ and $\mathsf{overlap}$ satisfy more constraints which can be seen as invariants of $AbsS$, e.g., $\mathsf{maxAtInvoc}(p)$ and $\mathsf{overlap}(p)$ do not contain predecessors of pushes from $\mathsf{maxAtInvoc}(p)$ (for each $p, k_1, k_2$, $\mathsf{before}(k_1, k_2)$ and $k_2 \in \mathsf{maxAtInvoc}(p)$ implies $k_1\not\in \mathsf{maxAtInvoc}(p)\cup \mathsf{overlap}(p)$).
+They can be found in~\cite{extended}.
 
